### 2016-11-07 pkg/glmertree/paper/treatment_interactions.Rnw
Several changes in paper
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1013 255f4249-2263-4417-b1ea-8498779e27a7

\documentclass[nobf,doc]{apa}
\usepackage{amssymb}
\usepackage{amsmath,amsthm,tabularx,curves,texdraw,psfrag}
\usepackage{apacite}
\usepackage[english]{babel}
\usepackage{lscape,dcolumn,hhline}
\usepackage{graphicx,epic,eepic,rotating}
\usepackage[normalem]{ulem}
\usepackage{fancyhdr}
\usepackage{eurosym,bbding}
\usepackage{verbatim}
\usepackage{bigfoot}
\usepackage{ctable}
\usepackage{rotating}
\usepackage{lineno}
\usepackage{placeins}
\linespread{1.2}
\usepackage{tabto}

%% for \usepackage{Sweave}
\SweaveOpts{echo=FALSE, results=hide}
\setkeys{Gin}{width=0.8\textwidth}

%% for comments
\newcommand{\readme}[1]{\emph{\marginpar{README} (#1)}}

<<preliminaries>>=
library("foreign")
library("glmertree")
library("lattice")
@

\title{Detecting Treatment-Subgroup Interactions in Clustered Data with Generalized Linear Mixed-Effects Model Trees}
\shorttitle{Detecting Treatment-Subgroup Interactions in Clustered Data with GLMM Trees}
\leftheader{Treatment-Subgroup Interactions with GLMM Trees}
\rightheader{Treatment-Subgroup Interactions with GLMM Trees}
\author{M. Fokkema$^1$, N. Smits$^2$, A. Zeileis$^3$, T. Hothorn$^4$, H. Kelderman$^5$}
\affiliation{$^1$Universiteit Leiden, $^2$Universiteit van Amsterdam, $^3$Universit\"{a}t Innsbruck, $^4$Universit\"{a}t Z\"{u}rich, $^5$Universiteit Leiden and Vrije Universiteit, Amsterdam}

\acknowledgements{The authors would like to thank Prof.~Pim Cuijpers, Prof.~Jeanne Miranda, Dr.~Boadie Dunlop, Prof.~Rob DeRubeis, Prof.~Zindel Segal, Dr.~Sona Dimidjian, Prof.~Steve Hollon and Erica Weitz for granting access to the dataset for the application. The work for this paper was partially done while MF, AZ and TH were visiting the Institute for Mathematical Sciences, National University of Singapore in 2014. The visit was supported by the Institute.}

\begin{document}

\maketitle
\pagewiselinenumbers

\section{Abstract}
Identification of subgroups of patients for which treatment A is more effective than treatment B, and vice versa, is of key importance to the development of personalized medicine. Tree-based algorithms are helpful tools for the  detection of such interactions, but none of the available tree-based algorithms allow for taking into account a clustered or nested structure. Therefore, we propose the generalized linear mixed-effects model trees (GLMM tree) algorithm, which allows for the detection of treatment-subgroup interactions, while accounting for the clustered structure of a dataset. The algorithm uses model-based recursive partitioning (MOB) to detect treatment-subgroup interactions, and a GLMM for the estimation of random-effects parameters. In a simulation study, we evaluate the performance of GLMM trees and compare it with that of trees without random effects, and GLMMs with pre-specified interactions. GLMM tree was found to accurately recover treatment-subgroup interactions in 90\% of the simulated datasets, whereas trees without random effects only did so in 61\% of datasets. GLMM tree also outperformed trees without random effects in terms of predictive accuracy and Type-I error rates. Furthermore, compared to GLMMs with pre-specified interaction effects, GLMM tree showed better predictive accuracy. We illustrate the application of GLMM trees on an individual patient-level data meta-analysis on treatments for depression. We conclude that GLMM tree is a promising exploratory tool for the detection of treatment-subgroup interactions in clustered datasets.\\
\\
\textit{Keywords}: model-based recursive partitioning, treatment-subgroup interactions, random effects, generalized linear mixed-effects model, classification and regression trees

\section{Introduction}

In research on the efficacy of treatments for somatic and psychological disorders, the one-size-fits-all paradigm is slowly losing ground, and stratified (or personalized) medicine is becoming increasingly important. Stratified medicine presents the challenge of finding which patients respond best to which treatments. This can be referred to as the detection of treatment-subgroup interactions \cite<e.g., >{DoovyDuss14}. Often, treatment-subgroup interactions are studied using linear models, such as factorial analysis of variance techniques, in which potential moderators have to be specified a-priori, have to be checked one at a time, and continuous moderator variables have to be discretized. This may hamper identification of which treatment works best for whom, especially when there are no a-priori hypotheses about treatment-subgroup interactions. As noted by \citeA{KreayFran06}, there is a need for methods that generate instead of test such hypotheses.

Tree-based methods are such hypothesis-generating methods, as they can automatically detect subgroups which differ in the expected outcomes for one or more treatments. Due to their flexibility, tree-based methods are particularly useful for exploratory purposes, as they can handle many potential predictor variables at once and can automatically detect (higher order) interactions between predictor variables \cite{StroyMall09}. As such, tree-based methods are preeminently suited to the detection of treatment-subgroup interactions. Several tree-based algorithms for the detection of treatment-subgroup interactions have already been developed (\citeNP{DussyDoov15, DussyMeul04, SuyTsai09, FostyTayl11, LipkyDmit11, ZeilyHoth08, SeibyZeil15}; see \citeNP{DoovyDuss14} for an overview). Also, \citeA{ZhanyTsia12Bio,ZhanyTsia12Stat} have developed a flexible classification-based approach, allowing users to select from a range of statistical methods, including trees.

In many instances, researchers may want to detect treatment-subgroup interactions in a generalized linear mixed-effects (GLMM) type model. For example, in individual-level patient data meta-analysis, where datasets of multiple clinical trials on the same treatments are pooled \cite<e.g., >{KoopyHeij07}. In such analyses, the nested or clustered structure of the dataset should be taken into account by including study-specific random effects in the model, prompting the need for a mixed-effects model \cite<e.g., >{CoopyPatt09,HiggyWhit01}. In linear models, ignoring the clustered structure may lead, for example, to biased inference due to underestimated standard errors in linear models \cite<e.g., >{BrykyRaud92,NooryOpde05}. For tree-based methods, ignoring the clustered structure has been found to result in the detection of spurious subgroups and inaccurate predictor variable selection \cite<e.g., >{SelaySimo12, Mart15}. However, none of the purely tree-based methods for treatment-subgroup interaction detection allow for taking into account the clustered structure of a dataset. Therefore, in the current paper, we present a tree-based algorithm which can be used for the detection of (treatment-subgroup) interactions and non-linearities in GLMM type models: generalized linear mixed-effects model tree, or GLMM tree.

The GLMM tree algorithm builds on model-based recursive partitioning (MOB, \citeNP{ZeilyHoth08}), which offers a flexible framework for subgroup detection. For example, GLM-based MOB has been applied to detect treatment-subgroup interactions for the treatment of depression \cite{DrieySmit16} and amyotrophic lateral sclerosis \cite{SeibyZeil15}. In contrast to other purely tree-based methods \cite<e.g., >{ZeilyHoth08, SuyTsai09, DussyMech14}), GLMM tree allows for taking into account the clustered structure of datasets. In contrast to previously suggested regression trees with random effects, GLMM tree allows for treatment effect estimation, with continuous as well as non-continuous response variables \cite<e.g., >{HajjyBell11,SelaySimo12}.

In what follows, we first introduce an artificial motivating dataset, which will be used to illustrate and explain treatment-subgroup interaction detection with GLM-based MOB and GLMM tree. In the Simulation-based Evaluation, we evaluate the performance of GLMM tree in simulated datasets, and compare it with that of GLM-based MOB and GLMMs with pre-specified interaction effects. The comparison with GLM-based MOB allows for assessing the extent to which including random-effects estimation improves the performance of MOB trees. The comparison with GLMMs with pre-specified interactions will allow for assessing the extent to which using a tree-based method will improve detection of treatment-subgroup interactions. In the Application, we apply the algorithm to an existing dataset of a patient-level meta-analysis on the effects of psycho- and pharmacotherapy for depression. Finally, in the Discussion we summarize the results and point out some directions for future research.

\subsection{Artificial motivating dataset}

<<artificial-data>>=
set.seed(123)
treatment <- rbinom(n=150,size=1,prob=1/2)
duration <- round(matrix(rnorm(150,mean=5,sd=3),ncol=1))
duration[duration<0] <- 1
anxiety <- round(rnorm(150,mean=10,sd=3))
age <- round(rnorm(150,mean=30,sd=10))
age[age<18] <- 18
age[age>75] <- 75
cutscore1 <- 5+.44*3
cutscore2 <- 10
node1t1 <- ifelse(duration<=cutscore1 & anxiety<=cutscore2 & treatment==0, 1, 0)
node1t2 <- ifelse(duration<=cutscore1 & anxiety<=cutscore2 & treatment==1, 1, 0)
node2t1 <- ifelse(duration<=cutscore1 & anxiety>cutscore2 & treatment==0, 1, 0)
node2t2 <- ifelse(duration<=cutscore1 & anxiety>cutscore2 & treatment==1, 1, 0)
node3t1 <- ifelse(duration>cutscore1 & treatment==0, 1, 0)
node3t2 <- ifelse(duration>cutscore1 & treatment==1, 1, 0)
beta1 <- -2
beta2 <- 0
beta3 <- 2.5
error <- rnorm(150,0,2)
cluster <- error+rnorm(150, 0, 6)
int_values <- rnorm(10, 0, 1)
tmp <- rep(int_values,each=15)
tmp <- tmp[order(tmp)]
rand_int <- rep(NA, times=150)
rand_int[order(cluster)] <- tmp
cluster[order(cluster)] <- rep(1:10, each=15)
cor(error,rand_int)
true_error <- error - rand_int
true_eff <- round(9 + beta1*node1t1 - beta1*node1t2 + beta2*node2t1 - beta2*node2t2 + beta3*node3t1 - beta3*node3t2 + 
     + .4*treatment)
outcome <- round(9 + beta1*node1t1 - beta1*node1t2 + beta2*node2t1 - beta2*node2t2 + beta3*node3t1 - beta3*node3t2 + 
                      + .4*treatment + true_error + rand_int)
outcome[outcome<0] <- 1
example <- data.frame(treatment, outcome, duration, anxiety, cluster, age)
example$treatment <- factor(example$treatment, labels = c("Treatment 1", "Treatment 2"))
@

We created a dataset representing observations on 150 participants in a randomized clinical trial. Every participant was randomly assigned to Treatment~1 or Treatment~2, and has a value for the response variable, with which the effect of treatment is assessed: the pos-ttreatment total score on a depression inventory. For all participants, three covariate values are available: duration of depressive symptoms prior to treatment in months (duration, range 0--15); age in years (age, range 18--75); anxiety inventory total score (anxiety, range 3--18).

The simulated dataset has 3 subgroups with differential treatment effectiveness. The first subgroup consists of observations with duration $\leq 6$ and anxiety $\leq 10$. In this subgroup, Treatment~1 is more effective than Treatment~2: the mean of the response variable is 7 for Treatment~1, and 11 for Treatment~2. The second subgroup consists of observations with duration $\leq 6$ and anxiety $> 10$. In this subgroup, both therapies are equally effective: the mean value of the response variable is 9 for Treatment~1, and 9 for Treatment~2. The third subgroup consists of observations with duration $>6$. In this subgroup, Treatment~2 is more effective than Treatment~1: the mean value of the response variable is 12 for Treatment~1, and 7 for Treatment~2.

<<example-trees>>=
exampleglm <- lmtree(outcome ~ treatment | age, data = example)
exampleglmtree <- lmtree(outcome ~ treatment | age + duration + anxiety, data = example)
exampleglmmtree <- lmertree(outcome ~ treatment | cluster | age + duration + anxiety, data = example)
@

\begin{figure}[t!]
\centering
<<exampleglmmtree, fig=TRUE, width=9, height=7>>=
plot(exampleglmmtree)
@
\caption{GLMM tree of the motivating example dataset. Three covariates (anxiety questionnaire score, duration of depressive symptoms at baseline in months and age) were used as potential splitting variables, and the clustering structure was taken into account by estimating random intercepts.}
\label{fig:example_glimmertree}
\end{figure}

Participants were part of one of ten clusters, each with a different value for the random intercept. Data were generated such that covariates and cluster-specific intercepts were uncorrelated. 43\% of variance in post-treatment depression scores was due to treatment-subgroup interactions, while 8\% of variance was due to cluster-specific variation.

See Figure~\ref{fig:example_glimmertree} for the GLMM tree fitted to the simulated data set which recovers the true underlying structure.

\subsection{Model-based recursive partitioning}

The rationale behind MOB is that a global parametric model may not describe the data well, and when additional covariates are available it may be possible to partition the dataset with respect to these covariates, and find a better model in each cell of the partition \cite{ZeilyHoth08}. This is reminiscent of the classification and regression tree (CART) algorithm of \citeA{BreiyFrie84}, which splits the dataset into subsets, for which the distributions of the outcome variable are most different. However, CART trees detect differences in the mean (or median) value across terminal nodes, whereas MOB trees detect differences in parameters of more complex models across terminal nodes.

For example, let us take a global GLM to estimate the overall treatment effect in the motivating dataset. The expectation $\mu_i$ of outcome $y_i$ given the treatment regressors $x_i$ is modeled through a linear predictor and suitable link function:
%
\begin{eqnarray}
	\label{eq:expected_value}
	E[y_i | x_i] & = & \mu_i, \\
	\label{eq:fixedeffects}
	g(\mu_{i}) & = & x_{i}^{\top}\beta,
\end{eqnarray}
%
where $x_{i}^{\top}\beta$ is the linear predictor for observation $i$ and $g$ is the link function. $\beta$ is a vector of fixed-effects regression coefficients, the first element representing the intercept, corresponding to the mean value of the linear predictor in the first treatment group, and the second element representing the slope, which is the mean difference in the linear predictor between the first and second treatment groups. Thus, for simplicity we assume $x_{i}$ and $\beta$ to have length 2 in the current paper. Also, for the continuous response variable in the motivating data set, we employ a Gaussian distribution with identity link and denote the error by $\epsilon_i = y_i - \mu_i$ with variance $\sigma_{\epsilon}^2$. However, the model can easily accommodate additional treatment conditions and covariates, and binary or count outcome variables.

\begin{figure}[t!]
\centering
\setkeys{Gin}{width=0.5\textwidth}
<<exampleglm, fig=TRUE, height=5, width=5>>=
plot(exampleglm, tp_args = list(mainlab = "Full sample (N = 150)"))
@
\caption{Example of a global GLM for treatment outcomes, based on the artificial motivating dataset ($N=150$). The dot for Treatment~1 represents the first, and the slope of the regression line represents the second element of $\beta$.}
\label{fig:fixedeffects}
\end{figure}

The result of fitting a global GLM to the motivating dataset is depicted in Figure~\ref{fig:fixedeffects}; the boxplots show the distribution of the posttreatment depression scores in both treatment groups. The global model does not describe the data well: there is substantial residual variance and the slope of the regression line is nearly zero. This does not necessarily mean that posttreatment depression score and treatment type are unrelated, as the effect of treatment may be moderated by variables not yet included in the model.

The MOB algorithm can be used to detect such moderation, by testing for parameter stability over a set of auxiliary covariates, also known as \textit{partitioning variables}. When the partitioning is based on a GLM, instabilities are differences in $\hat{\beta}$ across partitions of the dataset, which are defined by one or more auxiliary covariates not included in the linear predictor. To find these partitions, the MOB algorithm cycles iteratively through the following steps \cite{ZeilyHoth08}: (1) fit the parametric model to the dataset, (2) test for parameter instability over a set of partitioning variables, (3) if there is some overall parameter instability, split the dataset with respect to the variable associated with the highest instability, (4) repeat the procedure in each of the resulting subgroups.

More specifically, in step (2), to test for parameter instability, the so-called \textit{scores} are computed, using the score function. When the model is correctly specified, the expected value of the score for each observation equals zero. Therefore, under the null hypothesis of parameter stability, the scores do not systematically deviate from the expected value of zero, when observations are ordered by the values of a potential partitioning variable $U_k$ \cite<c.f., >{MerkyZeil13}. To statistically test whether the scores systematically deviate from zero with respect to variable $U_k$, the class of generalized M-fluctuation tests is used \cite{Zeil05,ZeilyHorn07}.

If the null hypothesis of parameter stability in step (2) can be rejected, that is, if at least one of the partitioning variables $U_{k}$ yields a $p$-value for the M-fluctuation test below the pre-specified significance level $\alpha$, the dataset is partitioned into two subsets in step (3). This partition is created using $U_{{k^*}}$, the variable with the minimal $p$-value in step (2). The split point for $U_{{k^*}}$ is selected, by taking the value that minimizes the sum of the values of the objective function in both partitions \cite{ZeilyHoth08}. In step (4), steps (1) through (3) are repeated in each partition, until the null hypothesis of parameter stability can no longer be rejected (or the subsets become too small).

\begin{figure}[t!]
\centering
\setkeys{Gin}{width=\textwidth}
<<exampleglmtree, fig=TRUE, width=12, height=9>>=
plot(exampleglmtree)
@
\caption{Example of a tree representation of model-based recursive partition, based on the artificial motivating dataset. Three additional covariates (anxiety questionnaire score, duration of depressive symptoms at baseline in months and age) were used as potential splitting variables.}
\label{fig:example_mobtree}
\end{figure}

Due to the binary recursive nature of MOB, the resulting partition can be represented as a binary tree. If the partitioning is based on a GLM, the result is a GLM tree, with a local fixed-effects regression model in every $j$-th ($j = 1,\dots,J$) terminal node or subgroup:
%
\begin{equation}
	\label{eq:fixedeffects_MOB}
	g(\mu_{ij}) = x_{i}^{\top}\beta_{j}
\end{equation}
%
Note that, if the recursive subgroup structure (i.e., the partition) were known, the tree could be estimated as a single GLM where all $x$-variables interact with the subgroup indicator. Somewhat more formally, the model could then be written: $g(\mu_{i}) = x_{i}^{* \top}\beta^{*}$, where $x_{i}^{*}$ are the values of the $2J$ interactions between the subgroups from the tree, and the elements of $x_{i}$. The corresponding $\beta^{*}$ would have length $2J$ as well, containing subgroup-specific fixed-effects coefficients.

Figure~\ref{fig:example_mobtree} shows the GLM tree grown on the motivating dataset. By using the three auxiliary covariates (anxiety, duration and age), MOB partitioned the observations into four subgroups, each with a different estimate for $\beta_j$. Age was correctly not detected as a partitioning variable, and the left- and rightmost subgroups are in accordance with the true treatment-subgroup interactions described above. However, the two subgroups in the middle do not represent true subgroups, which is due to the clustered structure of the dataset not being taken into account.

\subsection{Generalized linear mixed-effects model trees}

For datasets containing observations from multiple clusters (e.g., trials or research centers), application of a GLMM would be more appropriate. The model in Equation~\ref{eq:fixedeffects} is then extended to include cluster-specific, or random effects:
%
\begin{equation}
	\label{eq:mixedeffects}
	g(\mu_{i}) = x_{i}^{\top}\beta + z_{i}^{\top}b
\end{equation}
%
For a simple random intercept $z_{i}$ is a unit vector of length $M$, of which the $m$-th element takes a value of 1, and all other elements take a value of 0; $m$ ($m=1,\dots,M$) denotes the cluster which observation $i$ is part of. Further, $b$ is a random vector of length $M$, each $m$-th element representing the random intercept for cluster $m$. For simplicity we only employ a cluster-specific intercept here, but further random effects can be easily included in $z_i$. Furthermore, within the GLMM it is assumed that $b$ is normally distributed, with mean zero and variance $\sigma^{2}_{b}$. The parameters of the GLMM can be estimated with, for example, maximum likelihood (ML) and restricted ML (REML).

Note that, if the random-effects coefficients were known, the model could be estimated by a simple GLM as in Equation \ref{eq:fixedeffects} where $z_{i}^{\top}b$ would only be added as an offset (i.e., a variable with a fixed coefficient of 1) to the linear predictor.

Although the random-effects part of the GLMM in Equation \ref{eq:mixedeffects} accounts for the nested structure of the dataset, the global fixed-effects part may not describe the data well. Therefore, we propose the GLMM tree model, in which the fixed-effects part may be partitioned as in Equation \ref{eq:fixedeffects_MOB} while still adjusting for random effects: 
%
\begin{equation}
	\label{eq:glimmertree}
	g(\mu_{i}) = x_{i}^{\top}\beta_{j} + z_{i}^{\top}b
\end{equation}
%
To estimate the parameters of this model, we take an approach similar to that of the mixed-effects regression tree (MERT) approach of \citeA{HajjyBell11} and \citeA{SelaySimo12}. In the MERT approach, the fixed-effects part of a GLMM is replaced by a CART tree with constant fits in the nodes, and the random-effects part is estimated as usual. To estimate a MERT, an iterative approach is taken, alternating between (1) assuming random effects known, allowing for estimation of the CART tree, and (2) assuming the CART tree known, allowing for estimation of the random effects.

For estimating GLMM trees, we take this approach two steps further: (1)~Instead of a CART tree with constant fits to estimate the fixed-effects part of the GLMM, we use a GLM tree. This allows not only for detection of differences in intercepts across terminal nodes, but also for detection of differences in slopes such as treatment effects. (2)~By using generalized linear (mixed) models, the response may also be a binary or count variable instead of a continuous variable. The GLMM tree algorithm takes the following steps to estimate the model in Equation~\ref{eq:glimmertree}:
\begin{description}
  \item[Step 0:] Initialize by setting $r$ and all values $\hat{b}_{(r)}$ to 0.

  \item[Step 1:] Set $r = r+1$. Estimate GLM tree ($x_{i}^{\top}\hat{\beta}_{j(r)}$), with $z_{i}^{\top}\hat{b}_{(r-1)}$ as an offset.

  \item[Step 2:] Estimate random effects in the mixed-effects model $x_{i}^{\top}\hat{\beta}_{j(r)} + z_{i}^{\top}\hat{b}_{(r)}$ with subgroups $j(r)$ from the GLM tree.

  \item[Step 3:] Repeat Steps~1 and~2 until convergence.
\end{description}

The algorithm initializes by setting $b$ to $0$, since the random effects are initially unknown. In every iteration, the GLM tree and coefficients $\beta_{j(r)}$ and $b_{(r)}$ are re-estimated. The GLM tree is estimated, given the estimated random effects from the last iteration, and the random effects are estimated, given the estimated GLM tree from the current iteration. Iterations are continued until convergence, which is monitored by computing the log-likelihood criterion of the mixed-effects model in Equation~\ref{eq:glimmertree}. Typically, this converges if the tree does not change from one iteration to the next.

In Figure~\ref{fig:example_glimmertree}, the result of applying the GLMM tree algorithm to the motivating dataset is presented. By taking into account the clustering of observations, the spurious split involving the anxiety variable no longer appears in the tree and the true treatment subgroups have been recovered.

\section{Simulation-based evaluation}

To assess the performance of GLMM tree, we carried out three simulation studies: In Studies I and II, we compared the performance of GLMM tree with that of GLM tree. In Study I, we assessed the accuracy of both tree-based algorithms in datasets with treatment-subgroup interactions. In Study II, we assessed the Type-I error of both tree-based algorithms in datasets without treatment-subgroup interactions. In Study III, we compared the performance of GLMM tree in detecting treatment interactions with that of a more classical approach of interaction detection: a GLMM with pre-specified interactions.

\subsection{General simulation design}

In all simulation studies, we varied the following data-generating parameters:

\begin{enumerate} 
\item Three levels for sample size: $N=200$, $N=500$, $N=1000$.
\item Two levels for the number of potential partitioning covariates $U_1$ through $U_K$: $K=5$ and $K=15$.
\item Two levels for the intercorrelation between the potential partitioning covariates $U_1$ through $U_K$: $\rho_{U_k,U_{k'}}=0.0$, $\rho_{U_k,U_{k'}}=0.3$.
\item Three levels for the number of clusters: $M=5$, $M=10$, $M=25$.
\item Three levels for the population standard deviation of the normal distribution from which the cluster specific intercepts were drawn: $\sigma_{b}=0$, $\sigma_{b}=5$, $\sigma_{b}=10$.
\item Two levels for the intercorrelation between $b$ and one of the $U_k$ variables: $b$ and all $U_k$ covariates uncorrelated, $b$ correlated with one of the $U_k$ covariates.  
\end{enumerate}

In addition, each simulation study required a different design for main and interaction effects, which will be described in their respective sections below.

Following the approach of \citeA{DussyMech14}, all partitioning covariates $U_1$ through $U_{K}$ were drawn from a multivariate normal distribution with means ${\mu_U}_1 = 10$, ${\mu_U}_2 = 30$, ${\mu_U}_4 = -40$, and ${\mu_U}_5 = 70$. Means for the other covariates (i.e., ${\mu_U}_3$, and ${\mu_U}_6$ through ${\mu_U}_{15}$) were drawn from a discrete uniform distribution on the interval $[-70,70]$. All covariates $U_1$ through $U_{15}$ had the same standard deviation: ${\sigma_U}_k = 10$. Correlations between $U$ variables vary according to the third facet of the data-generating design described above.

To generate the cluster-specific intercepts $b_m$, we partitioned the sample into $M$ equally-sized clusters, conditional on one of the variables $U_1$ through $U_5$, producing the correlations in the sixth facet of the simulation design. For each cluster, a single value $b_m$ was drawn from a normal distribution with mean 0 and the value of $\sigma_{b}$ given by the fifth facet of the simulation design. When $b$ was correlated with one of the potential partitioning variables, this variable was randomly selected.

For every observation, we generated a binomial variable (with probability .5) as an indicator for treatment type. Random errors $\epsilon$ were drawn from a normal distribution with $\mu_{\epsilon} = 0$ and $\sigma_{\epsilon} = 5$. The value of the outcome variable $y_i$ was calculated as the sum of the random intercept, (node-specific) fixed effects and the random error term.

\subsection{Software}

R \cite{R} was used for the generation and analysis of all datasets. The \verb|partykit| package (version 1.0-2; \citeNP{HothyZeil15}) was employed for estimating GLM trees, using the \verb|lmtree| function for normal linear regressions and \verb|glmtree| for other response distributions. For estimating GLMMs the \verb|lmer| function (or \verb|glmer| function, respectively) from the \verb|lme4| package (version 1.1-7; \citeNP{BateyMeac12}) was employed, using restricted maximum likelihood (REML) estimation.

For estimation of GLMM trees the former two packages were combined in a new package \verb|glmertree| (version 0.1-0; \citeNP{FokkyZeil15}; available from R-Forge). This package provides functions \verb|lmertree| and \verb|glmertree| that iterate between estimation of the \verb|lmtree|/\verb|glmtree| model and the \verb|lmer|/\verb|glmer| model.

The significance level $\alpha$ for the parameter instability tests in the trees was set to .05, with a Bonferroni correction applied for multiple testing, in all simulations. The minimum number of observations per node in the tree was set to 20 and the maximum tree depth was set to four, thus limiting the number of potential subgroups to eight.

In Study III, the \verb|lmerTest| package (version 2.0-32; \citeNP{KuznyBroc16}) was used to assess statistical significance of fixed-effects predictors in GLMMs. The \verb|lmerTest| package calculates effective degrees of freedom and $p$-values based on Satterthwaite approximations.

\subsection{Study I: Method}

\begin{figure}[t!]
\centering
\setkeys{Gin}{width=0.8\textwidth}
<<dgp-tree, fig=TRUE, height=5, width=7>>=
fig4 <- party(
partynode(1L,
split = partysplit(2L, breaks = 30),
kids = list(
partynode(2L,
split = partysplit(1L, breaks = 17),
kids = list(
partynode(3L, info = c(
expression(''),
expression(beta[j0] == '17.5'),
expression(''),
expression(beta[j1] == -'5.0'),
expression(''),
expression(d[j] == -'1.0')
)),
partynode(4L, info = c(
expression(''),
expression(beta[j0] == '30.0'),
expression(''),
expression(beta[j1] == '0.0'),
expression(''),
expression(d[j] == '0.0')
)))),
partynode(5L,
split = partysplit(3L, breaks = 63),
kids = list(
partynode(6L, info = c(
expression(''),
expression(beta[j0] == '30.0'),
expression(''),
expression(beta[j1] == '0.0'),
expression(''),
expression(d[j] == '0.0')
)),
partynode(7L, info = c(
expression(''),
expression(beta[j0] == '42.5'),
expression(''),
expression(beta[j1] == '5.0'),
expression(''),
expression(d[j] == '1.0')
)))))),
data.frame(U1 = numeric(0), U2 = numeric(0), U5 = numeric(0))
)
plot(fig4, tp_args = list(FUN = identity, width = 9), tnex = 1.5)
@
\caption{Data-generating model for treatment-subgroup interactions. Parameter $d$ denotes the standardized mean difference between the outcomes of Treatment~1 and~2 (i.e., $\beta_{j1} / \sigma_{\epsilon}$).}
\label{fig:modelC}
\end{figure}

\subsubsection{Treatment-subgroup interaction design}

For generating datasets with treatment-subgroup (or piecewise) interactions, we used a design from \citeA{DussyMech14} which is depicted in Figure~\ref{fig:modelC}. Figure~\ref{fig:modelC} shows four subgroups, characterized by values of the (true) partitioning variables $U_2$, and $U_1$ or $U_5$. Two of the subgroups have mean differences in treatment outcome, indicated by a non-zero value of $\beta_{j1}$, and two subgroups do not have mean differences in treatment outcome, indicated by a $\beta_{j1}$ value of 0.

In this simulation study, some of the potential partitioning covariates were true partitioning covariates ($U_1$, $U_2$ and $U_5$), whereas others were not (e.g., $U_3$ and $U_4$). Therefore, an extra level was added to the sixth facet of the \textit{General simulation design} described above. Also, to assess the effect of the size of the treatment-effect differences in nodes 3 and 7, an additional facet was added:

\begin{enumerate}
\item[6.] Three levels for the intercorrelation between $b$ and one of the $U_k$ variables: $b$ and all $U_k$ covariates uncorrelated, $b$ correlated with one of the true partitioning covariates ($U_1$, $U_2$ or $U_5$), $b$ correlated with one of the noise variables ($U_3$ or $U_4$.  
\item[7.] Two levels for $\beta_{j1}$, the unstandardized mean difference in treatment outcomes. The absolute value of the treatment-effect difference was varied to be $|\beta_{j1}| = 2.5$ (corresponding to a medium effect size, Cohen's $d = 0.5$; \citeNP{Cohe92}) and $|\beta_{j1}| = 5.0$ (corresponding to a large effect size; Cohen's $d = 1.0$). 
\end{enumerate}

Again, 50 datasets were generated for each cell of the design. As a result, $50 \times 3 \times 2 \times 2 \times 3 \times 3 \times 3 \times 2 = 32,400$ datasets were generated. In every dataset, the outcome variable was calculated as $y_{i} = x_{i}^{\top} \beta_{j} + z_{i}^{\top} b_{m} + \epsilon_{i}$.

\subsubsection{Assessment of performance}

Performance of the algorithms was assessed by means of the total number of nodes in estimated GLM and GLMM trees, the accuracy of the resulting trees and predictive accuracy. An accurately recovered tree was defined as a tree with (1) a total of seven nodes (2) the first split involving variable $U_2$ and a value of $30 \pm 5$, (3) the next split on the left involving variable $U_1$ and a value of $17 \pm 5$, and (4) the next split on the right involving variable $U_5$ and a value of $63 \pm 5$. The allowance of $\pm 5$ equals plus or minus half the population standard deviation of the partitioning variable ($\sigma_{U_k}$).

Predictive accuracy of each method was assessed by calculating the correlation between true and predicted treatment-effect differences in every dataset. To prevent overly optimistic estimates of predictive accuracy \cite{HastyTibs09}, predictive accuracy was assessed using test datasets. Test datasets were generated from the same population as training datasets, but test observations were not drawn from the same clusters as the training observations, but from `new' clusters.

To identify the most important predictors of tree size, tree accuracy and predictive accuracy, we used ANOVAs and GLMs with algorithm type and the parameters of the data-generating process as independent variables. First-order interactions between algorithm type and each of the data-generating parameters were also included. The most important predictors were further assessed with graphical displays.

\subsection{Study I: Results}

<<simulation-data>>=
load("treespecs_long.dat")
treespecs.long$N <- factor(as.numeric(as.character(treespecs.long$N)), ordered=T)
treespecs.long$sigmab <- factor(as.numeric(as.character(treespecs.long$sigmab)), ordered=T)
treespecs.long$treatdiff <- factor(as.numeric(as.character(treespecs.long$treatdiff)), ordered=T)
treespecs.long$truetree.values <- as.numeric(treespecs.long$truetree.values)-1
levels(treespecs.long$corUb)[levels(treespecs.long$corUb)=="bi and splitting U correlated"] <- "b correlated with splitting U"
levels(treespecs.long$corUb)[levels(treespecs.long$corUb)=="bi and non-splitting U correlated"] <- "b correlated with non-splitting U"
levels(treespecs.long$corUb)[levels(treespecs.long$corUb)=="uncorrelated"] <- "b and U uncorrelated"
treespecs.long <- treespecs.long[treespecs.long$correlation.ind!="GLMM",]
treespecs.long$correlation.ind <- factor(treespecs.long$correlation.ind)
treespecs.long$treesize.ind <- factor(treespecs.long$treesize.ind)
@

\subsubsection{Tree size}

\begin{figure}[t!]
\centering
<<treesize-xyplot, fig=TRUE, width=9, height=9>>=
aggdata.size <- aggregate(formula=treesize.values ~ treesize.ind + N + sigmab + corUb, FUN=mean, data=treespecs.long)	
print(xyplot(treesize.values ~ sigmab | N + corUb, data = aggdata.size, groups=treesize.ind, 
type="b", ylab="tree size", xlab=expression(sigma[b]), par.settings=standard.theme("pdf",color=F), 
abline=c(7,0), auto.key=list(space="top", columns=2, title=" ", cex.title=1,lines=T, points=T)))
@
\caption{Average tree size of GLM and GLMM trees for datasets with treatment-subgroup interactions. Rows represent levels of dependence between random effects ($b$) and one of the partitioning variables $U_k$; columns represent levels of sample size.}
\label{fig:xyplot_treesize_interact}
\end{figure}

GLMM tree yielded trees with an average size of 7.15 nodes ($\mathrm{SD}=0.61$), whereas GLM tree yielded an trees with an average size of 8.15 nodes ($\mathrm{SD}=2.05$), indicating that GLM trees involved more spurious splits than GLMM trees on average. The estimated probability that a dataset was erroneously not partitioned (the Type-II error) was 0 for both algorithms.

An ANOVA indicated that the most important predictors of tree size were algorithm type, sample size, variance of the random intercept, and the correlation between partitioning variables. The graphical display in Figure~\ref{fig:xyplot_treesize_interact} indicates that GLMM tree grows trees of about the true tree size in all conditions. In the absence of random effects (i.e., $\sigma_{b}=0$), GLM and GLMM trees grow trees of about the same size. However, clear differences in tree size were observed when $\sigma_b > 0$. When sample size is small and the random intercept is not correlated with a partitioning covariate, GLM tree lacks power and grows trees that are too small, on average. When sample size is larger and random intercepts are not correlated with a partitioning covariate, GLM and GLMM trees are about equally sized. However, when random intercepts are correlated with a partitioning covariate, GLM starts to create spurious splits, especially with larger sample sizes and larger $\sigma_b$ values.

\subsubsection{Accuracy of recovered trees}

\begin{figure}[t!]
\centering
<<accuracy-xyplot, fig=TRUE, width=9, height=9>>=
aggdata.acc <- aggregate(formula=truetree.values ~ correlation.ind + N + sigmab + corUb, FUN=mean, data=treespecs.long)
print(xyplot(truetree.values ~ sigmab | N + corUb, data = aggdata.acc, groups=correlation.ind, type="b", 
ylab="tree accuracy", xlab=expression(sigma[b]), par.settings=standard.theme("pdf",color=FALSE), 
auto.key=list(space="top", columns=2, title=" ", cex.title=1,lines=TRUE, points=TRUE))
)
@
\caption{Tree accuracy of GLM and GLMM tree. Tree accuracy is defined as the proportion of datasets in which the true tree was accurately recovered. Rows represent levels of dependence between random effects ($b$) and one of the partitioning variables $U_k$, columns represent levels of sample size.}
\label{fig:xyplot_treeaccuracy}
\end{figure}

To assess the accuracy of trees recovered by both algorithms, we inspected variables and values that were selected for partitioning. For the first split, GLMM tree selected the true partitioning variable ($U_2$) in all datasets, and GLM tree in all but one datasets. The splitting value for $U_2$ selected for the first split was 29.94 for both GLM and GLMM tree, which is very close to the true splitting value of 30 (Figure~\ref{fig:modelC}).

Further splits were more accurately recovered by GLMM tree, which accurately recovered the partition in 90.40\% of datasets, and GLM tree in only 61.44\% of datasets. A GLM with tree accuracy as the outcome variable indicates that the most important predictors of tree accuracy were algorithm type, sample size, variance of the random intercept, and level of dependence between the partitioning variables and the random intercept. The graphical display in Figure~\ref{fig:xyplot_treeaccuracy} indicates that in the absence of random effects, the trees recovered by GLM and GLMM tree were about equally accurate. In the presence of random effects, GLM trees were much less accurate than GLMM trees when random effects were correlated with a partitioning covariate. When random intercepts were not correlated with one of the $U_k$ variables, GLMM tree outperformed GLM tree when sample size was small (i.e., $N = 200$).

\subsubsection{Predictive accuracy}

\begin{figure}[t!]
\centering
<<correlation-xyplot, fig=TRUE, width=9, height=7>>=
aggdata.cor <- aggregate(formula=correlation.values ~ correlation.ind + N + sigmab + treatdiff, FUN=mean, data=treespecs.long)
print(xyplot(correlation.values ~ sigmab | N + treatdiff, data = aggdata.cor, groups=correlation.ind, 
type="b", ylab="correlation", xlab=expression(sigma[b]), par.settings=standard.theme("pdf",color=F), 
auto.key=list(space="top", columns=2, title=" ", cex.title=1,lines=T, points=T),
scales=list(y=list(limits = c(-.1,1.1))))
)
@
\caption{Average predictive accuracy of GLM and GLMM trees. Predictive accuracy of trees is defined as the correlation between the true and predicted outcome differences between Treatment~1 and~2. Rows represent the levels of the absolute value of the unstandardized treatment-effect difference in subgroups with treatment-effect differences, columns represent levels of sample size.}
\label{fig:xyplot_correlations}
\end{figure}

The predicted treatment-effect differences of GLMM tree were closer to the true differences than those of GLM tree, with a mean correlation of .93 ($\mathrm{SD}=0.12$) for GLMM tree and .88 ($\mathrm{SD}=0.19$) for GLM tree. An ANOVA indicated that the most important predictors of predictive accuracy were algorithm type, sample size, variance of the random intercept, and the effect size of the treatment difference. The graphical display in Figure~\ref{fig:xyplot_correlations} shows clear main effects of sample size and treatment-effect differences for both algorithms. In the absence of random effects (i.e., $\sigma_b = 0$), predictions of GLM and GLMM tree were about equally accurate. In the presence of random effects, GLM tree predictions were clearly less accurate than those of GLMM tree when treatment-effect differences were smaller (i.e., Cohen's $d = .5$). This effect was also observed, but much less pronounced, when treatment-effect differences were larger (i.e., Cohen's $d = 1.0$). In other words, GLMM tree outperforms GLM tree in the presence of random effects, especially when sample size and treatment-effect differences are smaller.

\subsection{Study II: Method}

\subsubsection{Design}
In the second simulation study we assessed the Type-I error of GLM and GLMM tree. Type-I error was defined as the proportion of datasets without treatment-subgroup interactions which were erroneously partitioned by the algorithm. In the simulated datasets in this study there was only a main effect of treatment in the population. Put differently, there was only a single global value of $\beta_j = \beta$ in every dataset.

To assess the effect of the treatment-effect difference $\beta$, an additional facet was added to the \textit{General simulation design} described above: 
\begin{enumerate}
\item[7.] Two levels for $\beta$, the unstandardized mean difference in treatment outcomes. The absolute value of the treatment-effect difference was varied to be $|\beta_{j1}| = 2.5$ (corresponding to a medium effect size, Cohen's $d = 0.5$) and $|\beta_{j1}| = 5.0$ (corresponding to a large effect size; Cohen's $d = 1.0$). 
\end{enumerate}

For each cell of the \textit{General simulation design} described above, 50 datasets were generated. As a result, $50 \times 3 \times 2 \times 2 \times 3 \times 3 \times 2 \times 2 = 21,600$ datasets without treatment-subgroup interactions were generated. In every dataset, the outcome variable was calculated as $y_{i} = x_{i}^{\top} \beta + z_{i}^{\top} b_{m} + \epsilon_{i}$.

\subsubsection{Assessment of performance}

The total number of nodes in estimated GLM and GLMM tree were calculated. A tree of size $>1$ was classified as a Type-I error. To identify the most important predictors of Type-I error, we used a GLMs with algorithm type and the parameters of the data-generating process as independent variables. First-order interactions between algorithm type and each of the data-generating parameters were also included. The most important predictors were further assessed with a graphical display.

\subsection{Study II: Results}

In datasets without treatment-subgroup interactions, average tree size was 1.09 ($\mathrm{SD}=0.44$) for GLMM tree, and 2.02 ($\mathrm{SD}=1.68$) for GLM tree. The average Type-I error rate was only .04 for GLMM tree, and .33 for GLM tree. Main predictors of type-I error were found to be sample size, variance of the random intercept, and dependence between the random intercept and one of the partitioning variables.

The graphical display in Figure~\ref{fig:xyplot_treesize_nointeract} indicates that GLMM tree has a Type-I error rate somewhat below the pre-specified $\alpha$ level under all circumstances. The same goes for GLM tree, whenever the random intercept has 0 variance, or the random intercept is not correlated to one of the partitioning covariates. However, when the random intercept is correlated to one of the potential partitioning covariates, the type-I error rapidly increases for GLM tree. With large sample sizes and large variance of the random intercept, GLM tree will almost certainly create spurious splits.

\begin{figure}[t!]
<<treesize-nointeract-xyplot, fig=TRUE, width=9, height=7>>=
load("noint_treesizes_long.dat")
treesizes.long$sigmabm <- factor(as.numeric(as.character(treesizes.long$sigmabm)), ordered = TRUE)
treesizes.long$N <- factor(as.numeric(as.character(treesizes.long$N)), ordered = TRUE)
treesizes.long$treesize.valuesD <- as.numeric(treesizes.long$treesize.values>1)
treesize.glm <- glm(treesize.valuesD ~ treesize.ind + N + rho + np + treatdiff + numbclus + 
sigmabm + corUbm + treesize.ind*(N + rho + np + treatdiff + numbclus + sigmabm + corUbm), 
data=treesizes.long, family="binomial")
summary(treesize.glm)
#N, corUb & sigmab have strongest main and/or interaction effects
aggdata.size <- aggregate(formula=treesize.valuesD ~ treesize.ind + N + sigmabm + corUbm, 
FUN=mean, data=treesizes.long)
levels(aggdata.size$corUbm)[levels(aggdata.size$corUbm)=="b and random U correlated"] <- 
"b and U correlated"
levels(aggdata.size$corUbm)[levels(aggdata.size$corUbm)=="uncorrelated"] <- 
"b and U uncorrelated"
xyplot(treesize.valuesD ~ sigmabm | N + corUbm, data = aggdata.size, groups=treesize.ind, 
type="b", ylab="Type-I error", xlab=expression(sigma[b]), par.settings=standard.theme("pdf",color=F), 
auto.key=list(space="top", columns=2, title=" ", cex.title=1,lines=T, points=T))
@
\caption{Type-I error of GLM and GLMM trees. Rows represent dependence between random effects ($b$) and one of the partitioning variables $U_k$; columns represent sample size.}
\label{fig:xyplot_treesize_nointeract}
\end{figure}

\subsection{Study III: Method}

\subsubsection{Interaction design}

The treatment-subgroup interactions in Study I (Figure~\ref{fig:modelC}) can be referred to as piecewise interactions, as their effect is a stepwise function of the moderator (partitioning) variables $U_1$, $U_2$ and $U_5$. Trees are pre-eminently suited for recovering such piecewise interactions, but may have difficulty when the true interactions are continuous functions of moderator variables (for example, $U_1 \cdot U_2$). At the same time, linear regression models with pre-specified interaction terms may perform well in recovering continuous interactions, but may have difficulty in recovering piecewise interactions. Therefore, in the third simulation study, we added a seventh facet to the \textit{General simulation design} described above:

\begin{enumerate}
	\item[7.] Three levels for interaction type: continuous, piecewise and a combination of both. 
\end{enumerate}

For datasets with purely piecewise interactions, the same partition as in Study II (depicted in Figure \ref{fig:modelC}) was used. In other words, the outcome variable in this design was calculated as $y_i = x_{i}^{\top}\beta_{j} + z_{i}^{\top}b + \epsilon_i$, with the value of $\beta_j$ depending on the values of $U_2$, $U_1$ and $U_5$.

For the datasets with both piecewise and continuous interactions, the partition as depicted in Figure \ref{fig:modelC} was used. However, the fixed-effects part $x_{i}^{\top}\beta_{j}$ in each of the terminal nodes comprised continuous main and (treatment) interaction effects of the partitioning variables. These node-specific effects are described in Table \ref{tab:continuous_terms}. $\beta_j$ values were chosen to yield the same treatment-subgroup means as in Figure \ref{fig:modelC}. Continuous interactions were created using centered $U_k$ variables, calculated by subtracting the variable mean.

For datasets with purely continuous interactions, the outcome variable was calculated as $y_i = x_{i}^{\top}\beta + z_{i}^{\top}b + \epsilon_i$. Here, $\beta$ has a global value and no subscript, and comprised purely continuous main and (treatment) interaction effects of the moderator variables $U$, as described in Table \ref{tab:continuous_terms}.

\begin{table}
\caption{Fixed-effects terms in simulations with continuous and combined continuous and piecewise interaction designs.}
\begin{tabular}{lccccc}
\thickline
term							&	$\beta_3$&	$\beta_4$&	$\beta_6$&	$\beta_7$&	$\beta$		\\
\hline
intercept						&	27		&	27		&	27		&	27		&	27			\\
$U_2$							&	.1		&	.1		&	.1		&	.1		&	.1			\\
$U_2 \cdot U_1$ 				&	-.357	&	0		&	0		&	0		&	-.357		\\
$U_2 \cdot U_5$	 				&	0		&	0		&	0		&	.357	&	.357		\\
$U_2 \cdot U_1 \cdot treatment$	&	-.151	&	0		&	0		&	0		&	-.151		\\
$U_2 \cdot U_5 \cdot treatment$	&	0		&	0		&	0		&	.151	&	.151		\\
\hline
\multicolumn{6}{l}{\small{\textit{Note. }Subscripted $\beta$ values refer to the combined}}\\	\multicolumn{6}{l}{\small piecewise and continuous interaction design, with the}\\
\multicolumn{6}{l}{\small values of the subscripts denoting the terminal nodes in}\\
\multicolumn{6}{l}{\small Figure \ref{fig:modelC}. $\beta$ without a subscript refers to the global}\\
\multicolumn{6}{l}{\small coefficients in the purely continuous interaction design.}\\
\end{tabular}	
\label{tab:continuous_terms}
\end{table}

Furthermore, in this simulation study, the number of cells in the design was reduced by limiting the fourth facet of the data-generating design to a single level ($M=25$ clusters), as Study I and II indicated no effects of the number of clusters. The fifth facet of the data-generating design was limited to two levels ($\sigma_b = 2.5$ and $\sigma_b = 7.5$). As a result, $50 \times 3 \times 2 \times 2 \times 1 \times 2 \times 2 \times 3$ = 7,200 training datasets were generated for this study.

\subsubsection{GLMMs with pre-specified interactions}
GLMMs were estimated by specifying main effects for all covariates $U_k$ and the treatment indicator, first-order interactions between all pairs of covariates $U_k$, and second-order interactions between all pairs of covariates $U_k$ and treatment. Continuous predictor variables were centered by subtracting the mean value, before calculating and including the interaction term in the GLMM. Predictions were obtained by refitting a GLMM on the training data, using only the predictors with p-values $< .05$ in the original GLMM. Predictions for test observations were calculated using the refitted GLMMs.

\subsubsection{Assessment of performance} 
Predictive accuracy was assessed in terms of the correlation between the true and predicted treatment-effect differences in test datasets. To identify the most important predictors of predictive accuracy, we used an ANOVA with algorithm type and the parameters of the data-generating process as independent variables. First-order interactions between algorithm type and each of the data-generating parameters were also included. The most important predictors were further assessed with a graphical display.

<<linear-simulation-data>>=
load("pw_treespecs_long.dat")
treespecs.long$N <- factor(as.numeric(as.character(treespecs.long$N)), ordered=T)
treespecs.long$sigmab <- factor(as.numeric(as.character(treespecs.long$sigmab)), ordered=T)
treespecs.long$type <- factor(treespecs.long$type, ordered = TRUE, levels = 
c("linear", "both", "piecewise"))
levels(treespecs.long$type) <- gsub("linear","continuous", levels(treespecs.long$type))
@

\subsection{Study III: Results}

\begin{figure}[t!]
\centering
<<correlation-pw-xyplot, fig=TRUE, width=9, height=7>>=
aggdata.cor <- aggregate(formula=correlation.values ~ correlation.ind + np + type + N, FUN=mean, 
data=treespecs.long)
xyplot(correlation.values ~ type | N + np, data = aggdata.cor, groups=correlation.ind, type="b",
ylab="correlation", xlab="interaction type", par.settings=standard.theme("pdf",color=F), 
auto.key=list(space="top", columns=2, title=" ", cex.title=1,lines=T, points=T),
scales=list(y=list(limits = c(-.1,1.1))))
@
\caption{Average predictive accuracy of GLMMs and GLMM trees. Predictive accuracy of trees is defined as the correlation between the true and predicted differences between Treatment~1 and~2. Columns represent sample size, rows represent the number of covariates.}
\label{fig:xyplot_correlation_pw}
\end{figure}

GLMM tree showed somewhat higher average correlations between the true and predicted treatment-effect differences: average accuracy was .54 (SD = .40) for GLMM trees and .54 (SD = .43) for GLMMs. The most important predictors of predictive accuracy were algorithm type, sample size, the number of covariates, and type of interaction (continuous, piecewise or both). These effects were further assessed by means of the graphical display depicted in Figure~\ref{fig:xyplot_correlation_pw}.

As Figure~\ref{fig:xyplot_correlation_pw} indicates, GLMM trees show highest predictive accuracy in datasets with purely piecewise interactions, whereas GLMMs show highest predictive accuracy in datasets with purely continuous interactions. Performance of both methods improves with increasing sample size. Furthermore, performance of GLMM tree is not affected by the number of covariates, whereas the predictive accuracy of GLMMs deteriorates when the number of covariates increases, especially when the true interactions are not purely continuous. This indicates that GLMM tree may be especially useful for exploratory purposes, where there are many potential moderator variables. Also, GLMM trees may provide simpler models, as the average number of fixed-effects predictor variables (main effects and interactions) evaluated for prediction with GLMMs was 12.30, whereas the average number of inner nodes that need to be evaluated for prediction with GLMM tree was only 3.38.

\section{Application: Individual patient-level meta-analysis dataset on treatments for depression}

<<depression-data>>=
metadata <- read.dta("Database IPDMA CBT PHA Version 11.dta")
metadata[metadata == 999] <- NA
metadata[metadata == 888] <- NA
vars <- c("studyid", "Tx_group", "Age", "Gender", "education", "ComorbidAnxietyDisorder", "HRSDt0", "HRSDt1")
factors <- c("studyid", "Tx_group", "Gender", "education", "ComorbidAnxietyDisorder")
metadata$education <- factor(metadata$education, ordered = TRUE)
for (i in 1:length(factors)) {metadata[, factors[i]] <- factor(metadata[,factors[i]])}
metadata <- metadata[vars] # select only relevant variables
metadata <- metadata[complete.cases(metadata[, vars]), ] # select only complete data
metadata <- metadata[!metadata$Tx_group == "placebo", ] # remove placebo observations
metadata$Tx_group <- factor(metadata$Tx_group)
@

\subsection{Method}

To further illustrate the use of GLM and GLMM tree, we applied both algorithms to a dataset from an individual-patient data meta-analysis of \citeA{CuijyWeit14}. This meta-analysis was based on patient-level observations from 14 RCTs, comparing the effects of psychotherapy (cognitive behavioral therapy; CBT) and pharmacotherapy (PHA) in the treatment of depression. The study of \citeA{CuijyWeit14} was aimed at establishing whether gender is a predictor or moderator of the outcomes of psychological and pharmacological treatments for depression. Treatment outcomes were assessed by means of the 17-item Hamilton Rating Scale for Depression (HAM-D; \citeNP{Hami60}). \citeA{CuijyWeit14} found no indication that gender predicted or moderated treatment outcomes. Further details on the dataset can be found in \citeA{CuijyWeit14}.

In our analyses, posttreatment HAM-D score was the outcome variable, and potential partitioning variables were age, gender, level of education, presence of a comorbid anxiety disorder at baseline, and pretreatment HAM-D score. The predictor variable in the linear model was treatment type (0 = CBT and 1 = PHA). An indicator for study was used as the cluster indicator.

In RCTs, ANCOVAs are often employed, to linearly control posttreatment values on the outcome measure for pretreatment values. Therefore, we estimated the GLM and GLMM trees using posttreatment HAM-D scores, controlled for the linear effects of pretreatment HAM-D scores. The trees were grown using data of patients with complete observations; that is, observations with non-missing values for potential partitioning variables, and pre- and posttreatment HAM-D score. As a result, data from 694 patients from 7 studies were included in the analyses. Results of our analysis may therefore not be representative of the complete dataset of the meta-analysis by \citeA{CuijyWeit14}.

To provide a standardized estimate of the treatment effect differences in the final nodes of the trees, we calculated node-specific Cohen's $d$ values. Cohen's $d$ was calculated by dividing the node-specific predicted treatment outcome difference by the node-specific pooled standard deviation. Confidence intervals for Cohen's $d$ could not be calculated, as these can not take into account the exploratory nature of our analyses (i.e., variable and split point selection).

To compare the two tree-based approaches with a more classical approach of interaction detection, we also fitted a GLMM with pre-specified interactions on the dataset. In the GLMM, the posttreatment HAM-D scores were controlled for the linear effects of pretreatment HAM-D scores, and then regressed on a random intercept, main effects of treatment and the potential moderators (partitioning variables), and interactions between treatment and the potential moderators. As it is not known in advance how to interact the potential moderators, higher-order interactions were not included.

The predictive accuracy of the GLM and GLMM trees and the GLMM was assessed by calculating average correlations between observed and predicted HAM-D posttreatment scores, based on 50-fold cross validation.

The results of recursive partitioning techniques are known to be potentially unstable, in the sense that small changes in the dataset may substantially alter the variables or values selected for partitioning. Therefore, we used the R~package \verb|stablelearner| \cite{PhilyZeil16} to assess the stability of the selected splitting variables and values. Using the \verb|stabletree| function, we calculated the number of times a variable and value were selected for partitioning over 500 subsamples of size $.9 \times N$ of the dataset.

<<depression-trees>>=
## offset
metadata$HRSDfit <- fitted(lm(HRSDt1 ~ HRSDt0, data = metadata))

## GLM tree
lmtree_app <- lmtree(HRSDt1 ~ Tx_group | Age + Gender + education + ComorbidAnxietyDisorder + HRSDt0,
  data = metadata, offset = HRSDfit)

## GLMM tree
lmertree_app <- lmertree(HRSDt1 ~ Tx_group | (1 | studyid) + offset(HRSDfit) | Age + Gender + education + ComorbidAnxietyDisorder + HRSDt0,
  data = metadata, ranefstart = metadata$HRSDfit)

## GLM tree with effect sizes: 
GLMtree_app_eff <- party(
	partynode(1L,
		split = partysplit(1L, breaks = 2), kids = list(
			partynode(2L, info = c(
				expression(''),
				expression(widehat(mu)[CBT] == '13.25'),
				expression(''),
				expression(widehat(mu)[PHA] == '10.29'),
				expression(''),
				expression(widehat(sigma)[pooled] == '7.52'),
				expression(''),
				expression(widehat(d) == '0.39'),
				expression(''))),
			partynode(3L, split = partysplit(2L, breaks = 1), kids = list(
				partynode(4L, info = c(
				expression(''),
				expression(widehat(mu)[CBT] == '7.82'),
				expression(''),
				expression(widehat(mu)[PHA] == '7.55'),
				expression(''),
				expression(widehat(sigma)[pooled] == '5.51'),
				expression(''),
				expression(widehat(d) == '0.05'),
				expression(''))),
			partynode(5L, info = c(
				expression(''),
				expression(widehat(mu)[CBT] == '10.35'),
				expression(''),
				expression(widehat(mu)[PHA] == '7.69'),
				expression(''),
				expression(widehat(sigma)[pooled] == '6.64'),
				expression(''),
				expression(widehat(d) == '0.40'),
				expression(''))))))),
	data.frame(education = metadata$education, ComorbidAnxietyDisorder = metadata$ComorbidAnxietyDisorder)
)

## GLMM tree with effect sizes:
GLMMtree_app_eff <- party(
	partynode(1L,
		split = partysplit(2L, breaks = 1), kids = list(
			partynode(2L, info = c(
				expression(''),
				expression(widehat(mu)[CBT] == '8.27'),
				expression(''),
				expression(widehat(mu)[PHA] == '7.99'),
				expression(''),
				expression(widehat(sigma)[pooled] == '5.86'),
				expression(''),
				expression(widehat(d) == '0.05'),
				expression(''))),
			partynode(3L, info = c(
				expression(''),
				expression(widehat(mu)[CBT] == '10.84'),
				expression(''),
				expression(widehat(mu)[PHA] == '8.21'),
				expression(''),
				expression(widehat(sigma)[pooled] == '6.80'),
				expression(''),
				expression(widehat(d) == '0.39'),
				expression(''))))),
	data.frame(education = metadata$education, ComorbidAnxietyDisorder = metadata$ComorbidAnxietyDisorder)
)
@

\begin{figure}[t!]
\centering
\setkeys{Gin}{width=0.75\textwidth}
<<depression-glmtree, fig=TRUE, height=9, width=8>>=
combine_panels <- function(panel1, panel2, party2 = NULL) {
  function(node) {
    nid <- id_node(node)
    pushViewport(viewport(
      layout = grid.layout(nrow = 2, ncol = 1, heights = c(1.1, 0.7)),
      name = paste("node_mob_mypanel", nid, sep = "")))
    grid.rect(gp = gpar(fill = "white", col = 0))
    pushViewport(viewport(layout.pos.col = 1, layout.pos.row = 1))
    panel1(node)
    popViewport()
    pushViewport(viewport(layout.pos.col = 1, layout.pos.row = 2))
    node2 <- if(is.null(party2)) node else node_party(party2[nid])
    panel2(node2)
    popViewport(2)
  }
}

plot(lmtree_app, tnex = 3.5, terminal_panel = combine_panels(
  panel1 = node_bivplot(lmtree_app),
  panel2 = node_terminal(GLMtree_app_eff, FUN = identity, height = 10, width = 14, id = FALSE),
  party2 = GLMtree_app_eff
))
@
\caption{GLM tree for prediction of posttreatment total scores on the Hamilton Rating Scale for Depression (HAM-D). Upper panel: The y-axes of the boxplots represent posttreatment HAM-D scores, and the x-axes represent treatment levels (cognitive behavior therapy, CBT vs.\ pharmacotherapy, PHA). Lower panel: Subgroup-specific descriptive statistics.}
\label{fig:lmtree_C&W}
\end{figure}

\subsection{Results}

The tree and effects sizes resulting from application of GLM tree are presented in Figure~\ref{fig:lmtree_C&W}. Those resulting from application of GLMM tree are presented in Figure~\ref{fig:lmertree_C&W}.

The GLM tree (Figure~\ref{fig:lmtree_C&W}) selected level of education as the first partitioning variable, and presence of a comorbid anxiety disorder as a second partitioning variable, for observations with a higher level of education. By taking into account study-specific intercepts, the GLMM tree (Figure~\ref{fig:lmertree_C&W}) indicates that the first split made by GLM tree may be a spurious one. The GLMM tree selected presence of a comorbid anxiety disorder as the only partitioning variable. The terminal nodes of Figure~\ref{fig:lmertree_C&W} show only a single treatment-subgroup interaction: for patients without a comorbid anxiety disorder, CBT and PHA provide more or less the same reduction in HAM-D scores (Cohen's $d = 0.05$; Figure~\ref{fig:lmertree_C&W}). For patients with a comorbid anxiety disorder, PHA provides a greater reduction in HAM-D scores (Cohen's $d = 0.39$; Figure~\ref{fig:lmertree_C&W}). The estimated intraclass correlation coefficient for the GLMM tree was .05.

The GLMM with pre-specified treatment interactions yielded three significant predictors of treatment outcome: like in the GLMM tree, an effect of the presence of a comorbid anxiety disorder was found (main effect: b = 2.29, p = .002; interaction with treatment: b = -2.10, p = .028). Also, the GLMM indicated an interaction between treatment and age (b = .10, p = .018).

Assessment of predictive accuracy by means of 50-fold cross validation indicated better predictive accuracy for GLMM tree than for GLM tree and the GLMM with pre-specified interactions. The correlation between true and predicted posttreatment HAM-D total scores averaged over 50 folds was .272 ($\mathit{var}=.067$) for GLMM tree, .233 ($\mathit{var}=.064$) for the GLMM with prespecified interactions and .190 ($\mathit{var}=.084$) for GLM tree.

Table \ref{tab:stability} presents statistics on the variables selected for partitioning in subsamples of the dataset. Note that the selection frequencies do not add up to 1, as trees may involve multiple, or no splits. Table \ref{tab:stability} indicates that the presence of a comorbid anxiety disorder was selected for partitioning in the majority of GLMM trees grown on subsamples of the dataset, and all other variables were selected in less than 3\% of the subsamples. As the comorbid anxiety disorder variable involved only a single splitting value, further assessment of the stability of splitting values was not necessary.

\begin{figure}
\centering
\setkeys{Gin}{width=0.5\textwidth}
<<depression-glmmtree, fig=TRUE, height=7.5, width=5.5>>=
plot(lmertree_app$tree, tnex = 3.5, terminal_panel = combine_panels(
  panel1 = node_bivplot(lmertree_app$tree),
  panel2 = node_terminal(GLMMtree_app_eff, FUN = identity, height = 10, width = 14, id = FALSE),
  party2 = GLMMtree_app_eff
))
@
\caption{GLMM tree for prediction of posttreatment total scores on the Hamilton Rating Scale for Depression (HAM-D). Left panel: The y-axes of the boxplots represent posttreatment HAM-D scores, and the x-axes represent treatment levels (cognitive behavior therapy, CBT vs.\ pharmacotherapy, PHA). Right panel: Subgroup-specific descriptive statistics.}
\label{fig:lmertree_C&W}
\end{figure}

\begin{table}
	\small
	\caption{Variable selection statistics}
	\begin{tabular}{lcc}
		\thickline
		&	\multicolumn{2}{c}{Selection frequency} \\
		Variable				&	GLM tree&	GLMM tree	\\
		\hline
		Education               &	.956	&	.014		\\
		ComorbidAnxietyDisorder	&	.398	&	.528		\\
		HRSDt0                  &	.034	&	.002		\\
		Age                     &	.000	&	.022		\\
		Gender                  &	.002	&	.004		\\
		\hline
		\multicolumn{3}{l}{\textit{Note.} Frequencies are calculated over 500 random}\\
		\multicolumn{3}{l}{subsamples of the complete dataset. Frequencies}\\
		\multicolumn{3}{l}{do not add up to 1, as trees may involve multiple}\\
		\multicolumn{3}{l}{or no splits.}\\
	\end{tabular}
	\label{tab:stability}
\end{table}

\FloatBarrier

\section{Discussion}

In the current paper, we presented the GLMM tree algorithm, which allows for estimation of a GLM-based recursive partition, as well as estimation of random-effects parameters. As such, we hypothesized GLMM tree to be well suited for the detection of treatment-subgroup interactions in clustered datasets, which was confirmed by our simulation studies. We performed simulation studies to assess the accuracy of GLMM tree in detecting treatment-subgroup interactions and illustrated the application of GLMM tree on an existing dataset.

GLMM tree performed very well in recovering treatment-subgroup interactions, accurately recovering interactions in 90\% of datasets with treatment-subgroup interactions. In contrast, GLM tree accurately recovered interactions in only 61\% of these datasets. In the absence of treatment-subgroup interactions, GLMM tree erroneously detected subgroups in 4\% of the datasets, whereas GLM tree erroneously detected subgroups in 33\% of those datasets. Put differently, the Type-I error rate of GLMM tree very closely approximated the $\alpha$ level used for testing parameter stability, whereas the Type-I error rate of GLM tree clearly exceeded this value.

The better performance of GLMM tree was mostly observed when random effects in the datasets were sizable, and random intercepts were correlated with potential partitioning variables. In these instances, random effects gave rise to spurious subgroup detection (spurious splits) by GLM tree, both in datasets with and without treatment-subgroup interactions.

GLMM tree also provided better predictive accuracy than GLM tree, with an average correlation between true and predicted treatment differences of .94 for GLMM tree, and .88 for GLM tree. GLMM tree clearly outperformed GLM tree when random effects in the datasets were sizable, differences in treatment effects across terminal nodes were relatively small (i.e., Cohen's $d=.5$), and/or sample size was small (i.e., $ N < 1,000$). Such treatment-effect differences and sample sizes may be quite common, even in multi-center clinical trials, and GLMM tree may provide a helpful tool for subgroup detection in those instances.

As expected, when random effects were absent from the simulated datasets, GLM and GLMM tree yielded very similar predictive accuracy. This indicates that GLMM tree can be applied whenever cluster-specific random effects are expected: In the absence of random effects, GLM tree and GLMM tree are expected to perform about equally well and in the presence of random effects GLMM tree is expected to outperform GLM tree.

Compared to treatment-interaction detection by means of GLMMs with pre-specified interaction effects, GLMM trees provided somewhat better accuracy, on average. When interactions were purely continuous, GLMMs outperformed GLMM trees, and when interactions were partly or purely piecewise, GLMM trees outperformed GLMMs. In addition, GLMM tree may have a clear advantage when there are a large number of potential moderator variables (i.e., $> 5$). Therefore, GLMM trees may be better suited than GLMMs for exploratory analyses, in which moderator variables need to be selected from a larger number of covariates. Furthermore, the number of terms in a GLMM increases quadratically with the number of potential moderator variables, yielding complex models. The trees in our simulations were limited to a maximum number of 7 inner nodes, and may therefore be much easier to use for prediction in practical decision-making contexts.

These findings are encouraging for the use of GLMM tree in the detection of treatment-subgroup interactions in datasets with clustered structures. However, it should be noted that the simulations show that GLMM tree performs very well, if the model is correctly specified. That is, if there are subgroups with respect to the variables specified as potential partitioning variables and these subgroups have different values for the parameters of the GLM, then GLMM tree will accurately recover those subgroups. However, misspecification of the model will negatively affect performance. The most important source of misspecification would be the omission of relevant variables, either in the GLM or as partitioning variables. When relevant variables are omitted, GLMM tree can only approximate the true subgroups using the specified variables. Another source of misspecification would be the inclusion of irrelevant variables. Our simulations indicate that the performance of GLMM tree was not negatively affected by the number of potential moderator variables specified, but the power to detect subgroups may still be reduced with larger numbers of potential partitioning variables. Including irrelevant variables in the GLM may also negatively affect performance, although we have not assessed this in our simulations.

In the Application, we found GLMM tree to provide results that were more easily interpretable, and also more accurate than a GLM tree without random effects. In addition, to judge clinical relevance of the findings, we calculated node-specific effect sizes, as would often be done in RCTs or meta-analysis. Although we have limited ourselves to calculating Cohen's $d$ in the current paper, equivalent values of the success rate difference or the number needed to treat can be calculated, but this would involve additional distributional assumptions \cite{KraeyKupf06}. Node-specific effect sizes can also be used to prune trees, when a researcher prefers to have a final tree which is based on statistical as well as clinical significance. A topic for further research would be the development of splitting procedures based on effect sizes, as this would allow for taking into account clinical significant in the tree-growing process.

As discussed in the Introduction, several tree-based methods for treatment-subgroup interaction detection are available. These methods have different objectives, and there is not yet an agreed-upon single best method. In a simulation study, \citeA{SiesyVanM16} found the method of \citeA{ZhanyTsia12Stat} to perform best, followed by model-based recursive partitioning. However, the method of Zhang et al. performed worst under some conditions of the simulation study in terms of the Type I error rate. Further research comparing tree-based methods for treatment-subgroup interaction detection is needed, especially  for clustered datasets, as our simulations were limited to GLM and GLMM-based MOB.

Furthermore, it should be stressed that tree-based methods are exploratory tools. They can be used to detect predictors, interactions and non-linear effects in a data-driven way, but users should take the exploratory nature of such analyses into account. The resulting trees are potentially unstable, and stability of the results should be assessed, preferably in a dataset not used for training, or by multi-fold cross-validation or resampling techniques. In the Application, we have shown how the stability of splitting variable selection can be assessed using resampling techniques.

In conclusion, GLMM tree provided highly accurate recovery of treatment-subgroup interactions and predictions of treatment effect differences, both in the presence and absence of cluster-specific random effects. Therefore, GLMM tree is a promising algorithm for the detection of treatment-subgroup interactions in datasets with a clustered structure, like for example in multi-center trials or individual-level patient data meta-analyses.

\nolinenumbers
\bibliographystyle{apacite}
\bibliography{bib}

\end{document}