### 2017-05-18 pkg/partykit/inst/urp/urp.Rnw
mv vignette skeleton for the time being
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1345 255f4249-2263-4417-b1ea-8498779e27a7

\documentclass[nojss]{jss}

%\VignetteIndexEntry{urp: Internals of Unbiased Recursive Partitioning}
%\VignetteKeywords{conditional inference, model-based recursive partitioning}
%\VignettePackage{partykit}

%% packages
\usepackage{amstext}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{thumbpdf}
\usepackage{rotating}
%% need no \usepackage{Sweave}

%% commands
\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
\renewcommand{\Prob}{\mathbb{P} }
\renewcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Var}{\mathbb{V}}
\newcommand{\R}{\mathbb{R} }
\newcommand{\N}{\mathbb{N} }
\newcommand{\C}{\mathbb{C} }
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}
\newcommand{\LS}{\mathcal{L}_n}
\newcommand{\TS}{\mathcal{T}_n}
\newcommand{\LSc}{\mathcal{L}_{\text{comb},n}}
\newcommand{\LSbc}{\mathcal{L}^*_{\text{comb},n}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\yn}{y_{\text{new}}}
\newcommand{\z}{\mathbf{z}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\sX}{\mathcal{X}}
\newcommand{\sY}{\mathcal{Y}}
\newcommand{\T}{\mathbf{T}}
\newcommand{\x}{\mathbf{x}}
\renewcommand{\a}{\mathbf{a}}
\newcommand{\xn}{\mathbf{x}_{\text{new}}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\ws}{\mathbf{w}_\cdot}
\renewcommand{\t}{\mathbf{t}}
\newcommand{\M}{\mathbf{M}}
\renewcommand{\vec}{\text{vec}}
\newcommand{\B}{\mathbf{B}}
\newcommand{\K}{\mathbf{K}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\bS}{\mathbf{S}}
\newcommand{\cellx}{\pi_n[\x]}
\newcommand{\partn}{\pi_n(\mathcal{L}_n)}
\newcommand{\err}{\text{Err}}
\newcommand{\ea}{\widehat{\text{Err}}^{(a)}}
\newcommand{\ecv}{\widehat{\text{Err}}^{(cv1)}}
\newcommand{\ecvten}{\widehat{\text{Err}}^{(cv10)}}
\newcommand{\eone}{\widehat{\text{Err}}^{(1)}}
\newcommand{\eplus}{\widehat{\text{Err}}^{(.632+)}}
\newcommand{\eoob}{\widehat{\text{Err}}^{(oob)}}
\newcommand{\bft}{\mathbf{t}}

\hyphenation{Qua-dra-tic}

\title{URP: Internals of Unbiased Recursive Partitioning}
\Plaintitle{URP: urp: Internals of Unbiased Recursive Partitioning}

\author{Torsten Hothorn\\Universit\"at Z\"urich
   \And Achim Zeileis\\Universit\"at Innsbruck}
\Plainauthor{Torsten Hothorn, Achim Zeileis}

\Abstract{
tbd
}
\Keywords{conditional inference, model-based recursive partitioning}

\Address{
  Torsten Hothorn\\
  Institut f\"ur Sozial- und Pr\"aventivmedizin, Abteilung Biostatistik \\
  Universit\"at Z\"urich \\
  Hirschengraben 84\\
  CH-8001 Z\"urich, Switzerland \\
  E-mail: \email{Torsten.Hothorn@R-project.org}\\
  URL: \url{http://user.math.uzh.ch/hothorn/}\\

  Achim Zeileis \\
  Department of Statistics \\
  Faculty of Economics and Statistics \\
  Universit\"at Innsbruck \\
  Universit\"atsstr.~15 \\
  6020 Innsbruck, Austria \\
  E-mail: \email{Achim.Zeileis@R-project.org} \\
  URL: \url{http://eeecon.uibk.ac.at/~zeileis/}
}

\begin{document}

<<setup, echo = FALSE, results = hide>>=
options(width = 70, SweaveHooks = list(leftpar = 
    function() par(mai = par("mai") * c(1, 1.1, 1, 1))))
require("partykit")
set.seed(290875)
@

\setkeys{Gin}{width=\textwidth}

\section{Overview}

\section{Exported: URP tree}

Main workhorse

<<.urptree, eval = FALSE>>=
.urp_tree <- function
(
    call,               ### match.call of user-visible function
    frame,              ### parent.frame of user-visible function
    data = NULL, 
    data_asis = FALSE, 
    control,            ### .urp_control() or more
    trafofun,           ### function for transformations
    doFit = TRUE        ### grow tree or set-up only
) {}
@

\begin{description}
\item[\code{call}] a call with \code{formula}, 
                   \code{data}, \code{subset}, 
                   \code{weights}, \code{offset}, 
                   \code{cluster}, \code{strata} arguments.
\code{formula} is of the form \code{y ~ x | z} where \code{y} is the
response, \code{x} are covariates in the model to be partitioned and
\code{z} are partitioning variables. \code{y ~ z} is interpreted as
\code{y ~ 1 | z}.
\item[\code{frame}] a frame to evaluate the call in.
\item[\code{data}] optional \code{data.frame}.
\item[\code{data\_asis}] a logical indicating to use \code{data} and
  not reevaluating the call.
\item[\code{control}] control arguments
\item[\code{trafofun}] a function for the computation of scores / data
transformations, see below.
\item[\code{doFit}] a logical indicating to actually fit the tree or to
evaluate the call only.
\end{description}

The control arguments allow to change the behaviour of the tree growing
algorithm, for example to use permutation or $M$-fluctuation tests for
variable selection. The most important choice is are the scores to be used,
these are defined by \code{trafofun}.

\subsection{Argument trafofun}

This is a function with arguments

<<trafofun, eval = FALSE>>=
trafofun(formula, data, ctrl)
@

Here, \code{formula} is the model part \code{y ~ x} of the full formula
\code{y ~ x | z}. This function returns a function with arguments
<<trafo, eval = FALSE>>=
function(subset, estfun, object, info = NULL) {}
@

\begin{description}
\item[\code{subset}] an integer vector specifying the subset of
  observations in the current node (\proglang{R} indexing).
\item[\code{estfun}] a logical indicating to return the scores.
\item[\code{object}] a logical indicating to return a fitted model.
\item[\code{info}] additional information, typically a list with a
\code{coefficients} slot to be used as starting values for model fitting.
\end{description}

This function returns a list with slots
\begin{description}
\item[\code{estfun}] the matrix of score functions
\item[\code{coefficients}] optional model coefficients
\item[\code{objfun}] optional objective function to be maximised
\item[\code{object}] optional model object
\item[\code{converged}] a logical indicating model convergance
\end{description}

\section{Not Exported: URP fit}

<<urp_fit, eval = FALSE>>=
.urp_fit <- function
(
    data, 
    partyvars, 
    cluster, 
    ctrl
) {}
@

returns a function

<<urp_fit-ret, eval = FALSE>>=
function(trafo, subset, weights) {}
@

with returns a function

<<selectfun, eval = FALSE>>=
function(y = NULL, trafo, subset = integer(0), 
         weights = integer(0), whichvar, info = NULL)  {}
@

which returns a list with arguments

\begin{description}
\item[\code{criteria}] a $2 \times J$ matrix with rows \code{statistic} and
                       \code{p.value} on the LOG scale. 
\item[\code{splitfun}] a function \code{function(whichvar, minbucket)}
returning a \code{partysplit} object or \code{NULL}.
\end{description}

\section{No Exported: URP node}

<<urp-node, eval = FALSE>>=
.urp_node <- function
(
    id = 1L,                    ### id of this node
    data,                       ### full data, readonly
    selectfun,                  ### variable selection
                                ### and split function
    svselectfun = selectfun,    ### same for surrogate splits
    partyvars,                  ### partytioning variables
                                ### a subset of 1:ncol(data)
    weights = integer(0L),      ### optional case weights
    subset,                     ### subset of 1:nrow(data)
                                ### for identifying obs for this node
    ctrl,                       ### .urp_control()
    info = NULL,
    cenv = NULL                 ### environment for depth and maxid
) {}
@

%% \bibliography{party}

\end{document}