### 2017-05-25 pkg/model4you/inst/JORS/pmworkflow.tex
flow charts and model for one observation (does not work yet)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1381 255f4249-2263-4417-b1ea-8498779e27a7

\tikzstyle{block} = [rectangle, draw,
    text width=13em, text centered, rounded corners, minimum height=3em]
\tikzstyle{exp} = [rectangle,
    text width=10em, rounded corners, minimum height=4em]
\tikzstyle{line} = [draw, ->, line width=0.2em]

\begin{tikzpicture}[node distance = 7em, auto]
  \node [block] (clin) {Clinical trial};
  \node [block, below of=clin] (pdat) {Patient and treatment information, trial outcome};
  \node [block, below of=pdat] (frst) {Model-based forest};
  \node [block, below of=frst] (pmod) {Personalised models};
  \node [block, below of=pmod] (plot) {Dependence plot};

  \node [exp, right=1em of clin] {Real world};
  \node [exp, right=1em of pdat] {Data};
  \node [exp, right=1em of frst] {Black box model};
  \node [exp, right=1em of pmod] {Interpretable model};
  \node [exp, right=1em of plot] {Visual explanation};

  \path [line] (clin) -- (pdat) node [pos=0.5, left] {data collection};
  \path [line] (pdat) -- (frst) node [pos=0.5, left] {computation of forest};
  \path [line] (frst) -- (pmod) node [pos=0.5, left] {similarity weights};
  \path [line] (pmod) -- (plot) node [pos=0.5, left] {model coefficients};
\end{tikzpicture}