### 2017-06-22 pkg/disttree/inst/useR2017/slides.Rnw
added function 'gensimlist.R' and rda-files with data needed for plots; in 'slides.Rnw' it is first checked whether these rda-files already exist, otherwise they are generated; for 'simlist.rda' an external function 'gensimlist.R' is sourced and generates simlist
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1479 255f4249-2263-4417-b1ea-8498779e27a7

\documentclass[11pt,compress,t]{beamer}

\setbeamertemplate{caption}{\insertcaption} 
\usetheme{Z}
%% includes a replacement for \usepackage{Sweave}
\usepackage{amsmath,tikz}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap}
%\usepackage[style=authoryear, backend=bibtex]{biblatex}
\graphicspath{{plots/}}

%% colors
\definecolor{HighlightOrange}{rgb}{0.9490196,0.5725490,0.0000000}
\definecolor{HighlightBlue}{rgb}{0.4784314,0.7490196,0.9803922}

%% need no \usepackage{Sweave}
\definecolor{InputColor}{rgb}{0,0,0.3}
\definecolor{OutputColor}{rgb}{0.2,0.2,0.2}

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
library("ggplot2")
theme_set(theme_bw(base_size = 18))
library("disttree")
#library("partykitR1")
library("gamlss")
library("lattice")
library("crch")
library("latex2exp")
library("parallel")
library("gamlss.cens")
gen.cens(NO, type = "left")

pal <- hcl(c(10, 128, 260), 100, 50)
names(pal) <- c("forest", "tree", "gamlss")

## define distribution list:
# dist_list_normal
{

  dist_list_normal <- list()

  parnames <- c("mu", "sigma")
  etanames <- c("mu", "log(sigma)")

  ddist <-  function(y, eta, log = TRUE, weights = NULL, sum = FALSE) {

    val <- -1/2 * (log(2*pi) + 2*eta[2] + exp(log((y-eta[1])^2) - 2*eta[2]))
    if(!log) val <- exp(val)

    # par <- c(eta[1], exp(eta[2]))
    # val <- dnorm(y, mean = par[1], sd = par[2], log = log)

    if(sum) {
      if(is.null(weights)) weights <- rep.int(1, length(y))
      val <- sum(weights * val, na.rm = TRUE)
    }
    return(val)
  }

  sdist <- function(y, eta, weights = NULL, sum = FALSE) {

    score <- cbind(exp(-2*eta[2]) * (y-eta[1]), 
                   -1 + exp(-2*eta[2] + log((y-eta[1])^2)))

    # par <- c(eta[1], exp(eta[2])) 
    # score <- cbind(1/par[2]^2 * (y-par[1]), 
    #                (-1/par[2] + ((y - par[1])^2)/(par[2]^3)) * exp(eta[2]))

    score <- as.matrix(score)
    colnames(score) <- etanames
    if(sum) {
      if(is.null(weights)) weights <- rep.int(1, length(y))
      # if score == Inf replace score with 1.7e308 because Inf*0 would lead to NaN -> gradient is NaN
      score[score==Inf] = 1.7e308
      score <- colSums(weights * score, na.rm = TRUE)
    }
    return(score)
  }

  hdist <- function(y, eta, weights = NULL) {    
    ny <- length(y)
    if(is.null(weights)) weights <- rep.int(1, ny)

    d2ld.etamu2 <- sum(weights * rep.int(-exp(-2*eta[2]), ny))
    d2ld.etamu.d.etasigma <- sum(weights * (-2)*(y-eta[1]) * exp(-2*eta[2]), na.rm = TRUE)          # should be 0 for exact parameters (here: observed hess)
    d2ld.etasigma2 <- sum(weights * (-2)*exp(log((y-eta[1])^2) - 2*eta[2]), na.rm = TRUE)

    # par <- c(eta[1], exp(eta[2]))                           
    # d2ld.etamu2 <- sum(weights * rep.int(-1/par[2]^2, ny))
    # d2ld.etamu.d.etasigma <- sum(weights * (-2)*(y-par[1])/par[2]^2), na.rm = TRUE)          # should be 0 for exact parameters (here: observed hess)
    # d2ld.etasigma2 <- sum(weights * (-2)*(y-par[1])^2/par[2]^2, na.rm = TRUE)

    hess <- matrix(c(d2ld.etamu2, d2ld.etamu.d.etasigma, d2ld.etamu.d.etasigma, d2ld.etasigma2), nrow = 2)
    colnames(hess) <- rownames(hess) <-  etanames

    return(hess)
  }

  ## additional functions pdist, qdist, rdist
  pdist <- pnorm
  qdist <- qnorm
  rdist <- rnorm

  link <- c("identity", "log")

  linkfun <- function(par) {
    eta <- c(par[1], log(par[2]))
    names(eta) <- etanames
    return(eta)
  }

  linkinv <- function(eta) {
    par <- c(eta[1], exp(eta[2]))
    names(par) <- parnames
    return(par)
  }

  linkinvdr <- function(eta) {
    dpardeta <- c(1, exp(eta[2]))
    names(dpardeta) <- parnames
    return(dpardeta)
  }

  startfun <- function(y, weights = NULL){
    if(is.null(weights)) {
      mu <- mean(y)
      sigma <- sqrt(1/length(y) * sum((y - mu)^2))
    } else {
      mu <- weighted.mean(y, weights)
      sigma <- sqrt(1/sum(weights) * sum(weights * (y - mu)^2))
    }
    starteta <- c(mu, log(sigma))
    names(starteta) <- etanames
    return(starteta)
  }

  mle <- TRUE

  dist_list_normal <- list(family.name = "Normal Distribution",
                           ddist = ddist, 
                           sdist = sdist, 
                           hdist = hdist,
                           pdist = pdist,
                           qdist = qdist,
                           rdist = rdist,
                           link = link, 
                           linkfun = linkfun, 
                           linkinv = linkinv, 
                           linkinvdr = linkinvdr,
                           startfun = startfun,
                           mle = mle
  )
}

# dist_list_cens_normal
{

  dist_list_cens_normal <- list()

  parnames <- c("mu", "sigma")
  etanames <- c("mu", "log(sigma)")

  ddist <-  function(y, eta, log = TRUE, weights = NULL, sum = FALSE, left = 0, right = Inf) {     
    par <- c(eta[1], exp(eta[2]))
    val <- crch::dcnorm(x = y, mean = par[1], sd = par[2], left = left, right = right, log = log)
    if(sum) {
      if(is.null(weights)) weights <- if(is.matrix(y)) rep.int(1, dim(y)[1]) else rep.int(1, length(y))
      val <- sum(weights * val, na.rm = TRUE)
    }
    return(val)
  }

  sdist <- function(y, eta, weights = NULL, sum = FALSE, left = 0, right = Inf) {   
    par <- c(eta[1], exp(eta[2]))
    # y[y==0] <- 1e-323

    score_m <- crch:::scnorm(x = y, mean = par[1], sd = par[2], which = "mu", left = left, right = right)
    score_s <- crch:::scnorm(x = y, mean = par[1], sd = par[2], which = "sigma", left = left, right = right) * exp(eta[2]) # inner derivation exp(eta[2])
    score <- cbind(score_m, score_s)
    score <- as.matrix(score)
    colnames(score) <- etanames
    if(sum) {
      if(is.null(weights)) weights <- rep.int(1, length(y)[1])
      # if score == Inf replace score with 1.7e308 because Inf*0 would lead to NaN (0 in weights)
      score[score==Inf] = 1.7e308
      score <- colSums(weights * score, na.rm = TRUE)
      #if(any(is.nan(score))) print(c(eta, "y", y))
    }
    return(score)
  }

  hdist <- function(y, eta, weights = NULL, left = 0, right = Inf) {    
    ny <- length(y)
    if(is.null(weights)) weights <- rep.int(1, ny)

    par <- c(eta[1], exp(eta[2]))                           
    # y[y==0] <- 1e-323

    d2mu <- crch:::hcnorm(x = y, mean = par[1], sd = par[2], which = "mu", left = left, right = right)
    d2sigma <- crch:::hcnorm(x = y, mean = par[1], sd = par[2], which = "sigma", left = left, right = right)
    dmudsigma <- crch:::hcnorm(x = y, mean = par[1], sd = par[2], which = "mu.sigma", left = left, right = right) # FIX: order?
    dsigmadmu <- crch:::hcnorm(x = y, mean = par[1], sd = par[2], which = "sigma.mu", left = left, right = right) # FIX: order?
    dsigma <- crch:::scnorm(x = y, mean = par[1], sd = par[2], which = "sigma", left = left, right = right)

    d2ld.etamu2 <- sum(weights * d2mu, na.rm = TRUE)
    d2ld.etamu.d.etasigma <- sum(weights * dmudsigma * par[2], na.rm = TRUE)
    d2ld.etasigma.d.etamu <- sum(weights * dsigmadmu * par[2], na.rm = TRUE)
    d2ld.etasigma2 <- sum(weights * (d2sigma * exp(2*eta[2]) + dsigma * par[2]), na.rm = TRUE)

    hess <- matrix(c(d2ld.etamu2, d2ld.etamu.d.etasigma, d2ld.etasigma.d.etamu, d2ld.etasigma2), nrow = 2)
    colnames(hess) <- rownames(hess) <-  etanames

    return(hess)
  }

  ## additional functions pdist, qdist, rdist
  pdist <- function(q, eta, lower.tail = TRUE, log.p = FALSE) crch:::pcnorm(q, mean = eta[1], sd = eta[2], 
                                                                            lower.tail = lower.tail, log.p = log.p, 
                                                                            left = left, right = right)
  qdist <- function(p, eta, lower.tail = TRUE, log.p = FALSE) crch:::qcnorm(p, mean = eta[1], sd = eta[2], 
                                                                            lower.tail = lower.tail, log.p = log.p, 
                                                                            left = left, right = right)
  rdist <- function(n, eta) crch:::rcnorm(n, mean = eta[1], sd = eta[2], left = left, right = right)

  link <- c("identity", "log")

  linkfun <- function(par) {
    eta <- c(par[1], log(par[2]))
    names(eta) <- etanames
    return(eta)
  }

  linkinv <- function(eta) {
    par <- c(eta[1], exp(eta[2]))
    names(par) <- parnames
    return(par)
  }

  linkinvdr <- function(eta) {
    dpardeta <- c(1, exp(eta[2]))
    names(dpardeta) <- parnames
    return(dpardeta)
  }

  startfun <- function(y, weights = NULL){
    yc <- pmax(0,y)  # optional ?
    if(is.null(weights)) {
      mu <- mean(yc)
      sigma <- sqrt(1/length(yc) * sum((yc - mu)^2))
    } else {
      mu <- weighted.mean(yc, weights)
      sigma <- sqrt(1/sum(weights) * sum(weights * (yc - mu)^2))
    }
    starteta <- c(mu, log(sigma))
    names(starteta) <- etanames
    return(starteta)
  }

  mle <- FALSE

  dist_list_cens_normal <- list(family.name = "censored Normal Distribution",
                                ddist = ddist, 
                                sdist = sdist, 
                                hdist = hdist,
                                pdist = pdist,
                                qdist = qdist,
                                rdist = rdist,
                                link = link, 
                                linkfun = linkfun, 
                                linkinv = linkinv, 
                                linkinvdr = linkinvdr,
                                startfun = startfun,
                                mle = mle
  )
}

######## data generating process based on a predefined tree or a parameter function
# if tree: 2 splits => 3 terminal nodes
# 10 split variables (x1, ..., x10) are availabe (5 numeric, 2 binary, 3 categorical)
#
# input: n            ..... nr of observations
#        family       ..... distribution of the generated observations
#        split.matrix ..... indices of the variables used for the splits together with the corresponding split points
#        parm.matrix  ..... set of distribution parameters for each subgroup
#        fun          ..... parameter function
#
# output: data.frame with generated observations y (generated seperatly in each subgroups with given distributions parameters),
#                         the given split variables x1, ..., x10 for each observation and
#                         the index of the subgroup for each observation or
#                         the distribution parameter for each observation

# FIX: until now only complete lists can be handed over (dist_list...)
dgp <- function(n, family = dist_list_normal, 
                fun = NULL,
                split.matrix = matrix(nrow = 2, ncol = 2), 
                par.matrix = matrix(nrow = 3, ncol = 2), 
                round.sp = 3)
{

  # generating the possible split variables
  x1 <- runif(n,-0.4,1)
  x2 <- runif(n,-10,10)
  x3 <- runif(n,0,100)
  x4 <- x1 + rnorm(n, sd = 0.1)
  x5 <- x1 + rnorm(n, sd = 0.3)
  x6 <- rbinom(n,1,0.5)
  x7 <- rbinom(n,1,0.5)
  x8 <- sample(1:4, n, replace = TRUE) 
  x9 <- sample(1:5, n, replace = TRUE) 
  x10 <- sample(1:7, n, replace = TRUE)
  #x11 <- runif(n,-0.5,1)
  x <- cbind(x1,x2,x3,x4,x5,x6,x7,x8,x9,x10)
  # reduce nr of possible split points by rounding values of split variables
  x <- round(x, digits = round.sp)

  y <- vector(mode = "numeric", length = n)

  # getting the random function
  if(is.function(family)) family <- family()
  if(inherits(family, "gamlss.family")) {
    rfun <- get(paste0("r",family$family[[1]]))
  } else {
    if(family$family.name == "Normal Distribution") rfun <- rNO
    if(family$family.name == "censored Normal Distribution") rfun <- rNO
    if(family$family.name == "Poisson Distribution") rfun <- rPO
    if(family$family.name == "Weibull Distribution") rfun <- survival:::rsurvreg 
  }

  if(is.null(fun)) {
    index <- vector(mode = "numeric", length = n)
    sv = split.matrix[,1]    # index of split variables
    sp = split.matrix[,2]    # split points
    par = par.matrix

    # splitting based on the given split variables and split points and
    # generating the observations seperatly for each subgroup (with the corresponding given distribution parameters)
    for(k in 1:n){
      if(x[k, sv[1]] <= sp[1]){
        y[k] <- do.call(rfun, as.list(c(1,par[1,])))
        index[k] <- 2L
      } else {
        if(x[k, sv[2]] <= sp[2]){
          y[k] <- do.call(rfun, as.list(c(1,par[2,]))) 
          index[k] <- 4L
        } else {
          y[k] <- do.call(rfun, as.list(c(1,par[3,])))
          index[k] <- 5L
        }
      }
    }
    d <- as.data.frame(cbind(y, x, index))
    colnames(d) <- c("y", paste0("x", c(1:10)), "index")

  } else {

    dpar <- fun(x)

    if(!(family$family.name == "Poisson Distribution")){
      y <- rfun(n, dpar[,1], dpar[,2])
      d <- as.data.frame(cbind(y, x, dpar))
      colnames(d) <- c("y", paste0("x", c(1:10)), "mu", "sigma")
    } else {
      y <- rfun(n, dpar)
      d <- as.data.frame(cbind(y, x, dpar))
      colnames(d) <- c("y", paste0("x", c(1:10)), "lambda")
    }
  }

  if(family$family.name == "censored Normal Distribution") {
    #d$ystar <- d$y
    d$y <- pmax(d$y, 0)
  }

  return(d)
}

############################################ 
# parameter functions

### plotting function for one data set
plot1 <- function(learndata, dt=NULL, df=NULL, g=NULL, all.par = FALSE, df.par = FALSE, g.par = FALSE, dt.par = FALSE) 
{
  plotdata <- cbind(learndata[,c("y","x1","mu","sigma")],
                    df$fitted.par,
                    dt$fitted.par,
                    #coef(dt)[paste(dt$fitted[,1]),],
                    g$mu.fv, g$sigma.fv)

  colnames(plotdata) <- c("y","x1","true.mu","true.sigma", 
                          "fitted.mu.df","fitted.sigma.df",
                          "fitted.mu.dt","fitted.sigma.dt",
                          "fitted.mu.g","fitted.sigma.g")
  sp <- plotdata[order(plotdata["x1"]),]

  if(!(any(dt.par, df.par, g.par, all.par))){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "True parameters", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", lty = 2, col = 'black')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", lty = 2, col = 'black')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$')), 
           col = c('black','black'), lty = c(1,2,2), bty = 'n')
  }

  # disttree
  if(dt.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "disttree", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.dt, type = "l", col = pal["tree"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.dt + sp$fitted.sigma.dt, type = "l", col = 'greenyellow')
    lines(x = sp$x, y = sp$fitted.mu.dt - sp$fitted.sigma.dt, type = "l", col = 'greenyellow')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
                        TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')), 
           col = c('black','grey',pal["tree"], 'greenyellow'), lty = 1, bty = 'n')
  }

  # distforest
  if(df.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "distforest", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.df, type = "l", col = pal["forest"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.df + sp$fitted.sigma.df, type = "l", col = 'orange')
    lines(x = sp$x, y = sp$fitted.mu.df - sp$fitted.sigma.df, type = "l", col = 'orange')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
                        TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')), 
           col = c('black','grey',pal["forest"], 'orange'), lty = 1, bty = 'n')
  }

  # gamlss
  if(g.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey",  main = "gamlss", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.g, type = "l", col = pal["gamlss"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.g + sp$fitted.sigma.g, type = "l", col = 'lightblue')
    lines(x = sp$x, y = sp$fitted.mu.g - sp$fitted.sigma.g, type = "l", col = 'lightblue')
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
                        TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')),  
           col = c('black','grey',pal["gamlss"], 'lightblue'), lty = 1, bty = 'n')
  }

  # disttree vs distforest vs gamlss
  if(all.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n', xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$fitted.mu.dt, type = "l", col = pal["tree"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.df, type = "l", col = pal["forest"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.g, type = "l", col = pal["gamlss"], lwd = 2)
    legend('topleft', c(TeX('$\\mu$'), TeX('$\\hat{\\mu}$ disttree'),
                        TeX('$\\hat{\\mu}$ distforest'), TeX('$\\hat{\\mu}$ gamlss')), 
           col = c('black',pal["tree"],pal["forest"],pal["gamlss"]), lty = 1, bty = 'n')
  }
}

@

%\bibliography{ref.bib}

\begin{document}

\title{Distributional Trees and Forests}
\author{Lisa Schlosser, Torsten Hothorn, Achim Zeileis}
\URL{https://R-Forge.R-project.org/projects/partykit/}

\section{Distributional Trees and Forests}

\subsection{Motivation}

\begin{frame}[fragile]
\frametitle{Motivation}
\vspace{-0.41cm}
\begin{figure}[!htb]
\minipage{0.29\textwidth}
\begin{center}
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@
\visible<1->{
<<plot_motivation_GLM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt='n', yaxt='n', ann=FALSE, col = "darkgrey")
box(lwd=5)
lines(x = x, y = ytrue, col = "red", lwd = 5, main = "")
@
}
\end{center}
\endminipage
\visible<2->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
\begin{center}
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@
\visible<2->{
<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt='n', yaxt='n', ann=FALSE, col = "darkgrey")
box(lwd=5)
lines(x = x, y = ytrue, col = "red", lwd = 5, main = "")
@
}
\end{center}
\endminipage
\visible<3->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
\begin{center}
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\visible<3->{
<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt='n', yaxt='n', ann=FALSE, col = "darkgrey")
box(lwd=5)
lines(x = x, y = ytrue, col = "red", lwd=5)
lines(x = x, y = ytrue+0.1+var, col = "red", lty = 2)
lines(x = x, y = ytrue-0.1-var, col = "red", lty = 2)
@
}
\end{center}
\endminipage

\vspace{0.5cm}
\minipage{0.25\textwidth}
\begin{center}
\visible<1->{
LM, GLM\\
\vspace{0.5cm}
\code{lm}\\
\code{glm}\\
\vspace{1.5cm}}
\end{center}
\endminipage
\hspace{1.1cm}
\minipage{0.25\textwidth}
\begin{center}
\visible<2->{
GAM\\
\vspace{0.5cm}
\code{mgcv}\\
\code{VGAM}\\
$\ldots$\\
\vspace{1cm}}
\end{center}
\endminipage
\hspace{1.1cm}
\minipage{0.25\textwidth}
\begin{center}
\visible<3->{
GAMLSS\\
\vspace{0.5cm}
\code{gamlss}\\
\code{mgcv}\\
\code{VGAM}\\
\code{gamboostLSS}\\
$\ldots$}
\end{center}
\endminipage
\end{figure}
\end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}
\vspace{-0.2cm}
\begin{figure}[!htb]
\minipage{0.29\textwidth}
\begin{center}
<<motivation_regtree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- ytree <- yforest <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))
y <- ytrue + rnorm(nobs,0,0.3)
for(i in 1:nobs) ytree[i] <- if(x[i]<1/3) 0.5 else {if(x[i]<2/3) 2 else 1}
plot(x = x, y = y, xaxt='n', yaxt='n', ann=FALSE, col = "darkgrey")
box(lwd=5)
#lines(x = x, y = ytrue, col = "grey", lwd = 1.5, main = "")
lines(x = x, y = ytree, col = "red", lwd=5)
@
\visible<1->{
<<plot_motivation_regtree, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt='n', yaxt='n', ann=FALSE, col = "darkgrey")
box(lwd=5)
#lines(x = x, y = ytrue, col = "grey", lwd=5, main = "")
lines(x = x, y = ytree, col = "red", lwd=5)
@
}
\end{center}

\endminipage
\visible<2->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
 \begin{center}
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@
\visible<2->{
<<plot_motivation_randforest, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt='n', yaxt='n', ann=FALSE, col = "darkgrey")
box(lwd=5)
#lines(x = x, y = ytrue, col = "grey", lwd=5, main = "")
lines(x = x, y = yforest, col = "red", lwd=5, main = "")
@
}
\end{center}

\endminipage
\visible<3->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
  \visible<3->{
  \begin{center}
  \begin{tikzpicture}
  \draw[line width=0.2mm, black] (0,0) rectangle (2.6,2.24);
  \node[rectangle, align=center] (sq) at (1.3, 1.12) {\Huge{?}};
  \end{tikzpicture}
  \end{center}}
\endminipage

\minipage{0.29\textwidth}
\begin{center}
\vspace{0.0cm}
\visible<1->{
Regression Tree\\
\vspace{0.5cm}
\resizebox{0.3\textwidth}{!}{
\begin{tikzpicture}
  \node[ellipse, fill=HighlightBlue!70, align=center] (n0) at (1, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n1) at (0.5, 1) {};
  \draw[-, line width=1pt] (n0) -- (n1);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n2) at (1.5, 1) {};
  \draw[-, line width=1pt] (n0) -- (n2);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n3) at (1, 0) {};
  \draw[-, line width=1pt] (n2) -- (n3);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n4) at (2, 0) {};
  \draw[-, line width=1pt] (n2) -- (n4);
\end{tikzpicture}}
\vspace{0.3cm}\\
\code{rpart}\\
\code{party(kit)}\\
\vspace{1cm}}
\end{center}
\endminipage
\hspace{0.65cm}
\minipage{0.29\textwidth}
\begin{center}
\vspace{0.0cm}
\visible<2->{
Random Forest\\
\vspace{0.5cm}
\resizebox{0.9\textwidth}{!}{
\begin{tikzpicture}
  \node[ellipse, fill=HighlightBlue!70, align=center] (n00) at (1, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n01) at (0.5, 1) {};
  \draw[-, line width=1pt] (n00) -- (n01);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n02) at (1.5, 1) {};
  \draw[-, line width=1pt] (n00) -- (n02);

  \node[ellipse, fill=HighlightBlue!70, align=center] (n10) at (3, 2) {};
  \node[ellipse, fill=HighlightBlue!70, align=center] (n11) at (2.5, 1) {};
  \draw[-, line width=1pt] (n10) -- (n11);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n12) at (3.5, 1) {};
  \draw[-, line width=1pt] (n10) -- (n12);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n13) at (2, 0) {};
  \draw[-, line width=1pt] (n11) -- (n13);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n14) at (2.8, 0) {};
  \draw[-, line width=1pt] (n11) -- (n14);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n15) at (3.2, 0) {};
  \draw[-, line width=1pt] (n12) -- (n15);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n16) at (4, 0) {};
  \draw[-, line width=1pt] (n12) -- (n16);

  \node[ellipse, fill=HighlightBlue!70, align=center] (n20) at (5, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n21) at (4.5, 1) {};
  \draw[-, line width=1pt] (n20) -- (n21);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n22) at (5.5, 1) {};
  \draw[-, line width=1pt] (n20) -- (n22);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n23) at (5, 0) {};
  \draw[-, line width=1pt] (n22) -- (n23);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n24) at (6, 0) {};
  \draw[-, line width=1pt] (n22) -- (n24);
\end{tikzpicture}
}
\vspace{0.5cm}\\
\code{randomForest}\\
\code{ranger}\\
\code{party(kit)}\\
$\ldots$}
\end{center}
\endminipage
\hspace{0.65cm}
\minipage{0.28\textwidth}
\begin{center}
\vspace{0.68cm}
\visible<3->{
Distributional trees and forests\\
\vspace{1.7cm}
\code{disttree}\\
based on \code{partykit}\\
\vspace{1cm}}
\end{center}
\endminipage
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Goals}

\textbf{Distributional:} \\
\hspace{0.3cm} Specify the complete probability distribution\\
\hspace{0.3cm} (including location, scale and shape).

\medskip

{\bf Tree:}
\begin{itemize}
  \item Automatic detection of steps and abrupt changes. %(data driven)
  \item Capture non-linear and non-additive effects and interactions.
\end{itemize}

\medskip

{\bf Forest:}
\begin{itemize}
  \item Smoother effects.
  \item Stabilization and regularization of the model.
\end{itemize}
\end{frame}

\subsection{Building Distributional Trees and Forests}
\begin{frame}
\frametitle{Building Distributional Trees and Forests}
{\bf Tree:}
\begin{enumerate}
\item Specify a distribution with log-likelihood function $\ell(\theta; y)$.
\item Estimate $\hat{\theta}$ via maximum likelihood.
\item Test for associations or instabilities of the scores $\frac{\partial \ell}{\partial \theta}(\hat{\theta};y_i)$ and each partitioning variable $x_i$.
%\item Assess whether the \emph{model scores} are associated with
%    (or change along) any of the available covariates -- e.g.,
%    using parameter instability tests (\emph{strucchange}) or
%    conditional inference (\emph{coin}).
\item Split the sample along the partitioning variable with the strongest association or instability.
    Choose breakpoint with highest improvement in log-likelihood.
\item Repeat steps 2--4 recursively in the subgroups
    until some stopping criterion is met.
    %-- e.g., for significance or sample size.
%\item Choose the variable with the strongest association
%\item Choose the split point which leads to the highest improvement of the model
%\item Split and repeat 2-6 in each node until a stopping criterion is met
\end{enumerate}

\vspace{0.4cm}
{\bf Forest:} Ensemble of trees.
\begin{itemize}
\item Bootstrap or subsamples.
\item Random input variable sampling.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Prediction}
%\vspace{1cm}
{\bf Tree:}\\
\begin{enumerate}
\item Select the subsample of the learning data which ends up in the same terminal node as the new observation.
\item Estimate $\hat{\theta}$ on this subsample.
\end{enumerate}
\vspace{0.5cm}

{\bf Forest:}\\
\begin{enumerate}
\item For each learning observation calculate a weight by counting the number of trees in which it ends up in the same terminal node as the new observation.
\item Estimate $\hat{\theta}$ on the wheighted learning data.
\end{enumerate}
% but weighted by similarities of the learning data and the new observation over all trees
%\item Evaluate final node for new observation in each of the trees and investigate similarities with each of the learning observations
%\item Fit a distributional model for the new observation with different influences of the learning observations defined by weights
\end{frame}

\begin{frame}
\frametitle{Fitting a Tree}
\vspace*{-0.5cm}
\begin{center} 
DGP:  $Y \sim \mathcal{N}(\mu(X), \sigma(X))$
\vspace*{-0.21cm}
<<dgp_tree, echo=FALSE, results=hide>>=
data <- data.frame(x = numeric(0), x = numeric(0), x = numeric(0))
names(data) <- c("x","x","x")
fig <- party(
  partynode(1L,
            split = partysplit(2L, breaks = 0.4),
            kids = list(
              partynode(2L, info = c(
                "n = 200",
                "   True parameters:   ",
                expression(mu == '4'),
                expression(sigma == '1')
              )),
              partynode(3L,
                        split = partysplit(3L, breaks = 0.8),
                        kids = list(
                          partynode(4L, info = c(
                            "n = 200",
                            "   True parameters:   ",
                            expression(mu == '12'),
                            expression(sigma == '3')
                          )),
                          partynode(5L, info = c(
                            "n = 100",
                            "   True parameters:   ",
                            expression(mu == '4'),
                            expression(sigma == '3')
                          )))))),
  data
)

node_inner_ext <- function (obj, id = TRUE, pval = TRUE, abbreviate = FALSE, fill = "white", 
    gp = gpar()) 
{
    meta <- obj$data
    nam <- names(obj)
    extract_label <- function(node) {
        if (is.terminal(node)) 
            return(rep.int("", 2L))
        varlab <- character_split(split_node(node), meta)$name
        if (abbreviate > 0L) 
            varlab <- abbreviate(varlab, as.integer(abbreviate))
        if (pval) {
            nullna <- function(x) is.null(x) || is.na(x)
            pval <- suppressWarnings(try(!nullna(info_node(node)$p.value), 
                silent = TRUE))
            pval <- if (inherits(pval, "try-error")) 
                FALSE
            else pval
        }
        if (pval) {
            pvalue <- node$info$p.value
            plab <- ifelse(pvalue < 10^(-3L), paste("p <", 10^(-3L)), 
                paste("p =", round(pvalue, digits = 3L)))
        }
        else {
            plab <- ""
        }
        return(c(varlab, plab))
    }
    maxstr <- function(node) {
        lab <- extract_label(node)
        klab <- if (is.terminal(node)) 
            ""
        else unlist(lapply(kids_node(node), maxstr))
        lab <- c(lab, klab)
        lab <- unlist(lapply(lab, function(x) strsplit(x, "\n")))
        lab <- lab[which.max(nchar(lab))]
        if (length(lab) < 1L) 
            lab <- ""
        return(lab)
    }
    nstr <- maxstr(node_party(obj))
    if (nchar(nstr) < 6) 
        nstr <- "aAAAAa"
    rval <- function(node) {
        node_vp <- viewport(x = unit(0.5, "npc"), y = unit(0.5, 
            "npc"), width = unit(1, "strwidth", nstr) * 1.3, 
            height = unit(3, "lines"), name = paste("node_inner", 
                id_node(node), sep = ""), gp = gp)
        pushViewport(node_vp)
        xell <- c(seq(0, 0.2, by = 0.01), seq(0.2, 0.8, by = 0.05), 
            seq(0.8, 1, by = 0.01))
        yell <- sqrt(xell * (1 - xell))
        xell <- xell*1.11 - 0.055             # to adapt size of the ellipse to the size with p-value
        lab <- extract_label(node)
        fill <- rep(fill, length.out = 2L)
        grid.polygon(x = unit(c(xell, rev(xell)), "npc"), y = unit(c(yell, 
            -yell) + 0.5, "npc"), gp = gpar(fill = fill[1]))
        grid.text(lab[1L], y = unit(1.5 + 0.5,                  # to adapt position of x to its position with p-value
            "lines"))
        #grid.text(lab[1L], y = unit(1.5 + 0.5 * (lab[2L] != ""), 
        #    "lines"))
        grid.text(lab[2L], y = unit(1, "lines"))
        if (id) {
            nodeIDvp <- viewport(x = unit(0.5, "npc"), y = unit(1, 
                "npc"), width = max(unit(1, "lines"), unit(1.3, 
                "strwidth", nam[id_node(node)])), height = max(unit(1, 
                "lines"), unit(1.3, "strheight", nam[id_node(node)])))
            pushViewport(nodeIDvp)
            grid.rect(gp = gpar(fill = fill[2]))
            grid.text(nam[id_node(node)])
            popViewport()
        }
        upViewport()
    }
    return(rval)
}

class(node_inner_ext) <- "grapcon_generator"

@

\setkeys{Gin}{width=0.74\linewidth}
<<plottree_dgp, fig=TRUE, echo=FALSE>>=
pal2 <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(fig, inner_panel = node_inner_ext,
     tp_args = list(FUN = identity, width = 18, fill = pal2[c(1, 3)]), 
     ip_args = list(fill = pal2[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1.4)
@
\end{center}
\end{frame}

<<tree, echo=FALSE, results=hide>>=
set.seed(7)
nobs <- 500
x <- runif(nobs, 0, 1)
mu <- sigma <- ytrue <- numeric(length = nobs)
for(i in 1:nobs) sigma[i] <- if(x[i]<=0.4) 1 else 3
for(i in 1:nobs) mu[i] <- if(x[i]<= 0.4|| x[i]>0.8) 4 else 12
y <- rnorm(nobs, mean = mu, sd = sigma)
#y <- rcnorm(nobs, mean = mu, sd = sigma, left = 0)
ytrue <- mu
data <- data.frame(cbind(y,x, ytrue))
tree <- disttree(y ~ x, data = data, family = NO(), type.tree = "mob")
#tree <- disttree(y ~ x, data = data, family = dist_list_cens_normal)
@

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-0.5cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.74\linewidth}
<<plottree_estpar, fig=TRUE, echo=FALSE>>=
# function for output in terminal panels
FUN <- function (x) 
{
  cf <- x$coefficients
  cf <- matrix(cf, ncol = 1, dimnames = list(names(cf), ""))
  c(sprintf("n = %s", x$nobs), "Estimated parameters:", parse(text = paste("mu == ", round(cf[1],3))), 
                                                        parse(text = paste("sigma == ", round(cf[2],3))))
}

# alternative to using FUN:
# change info slot in each terminal node and use FUN = identity
if(FALSE){
info2 <- tree$node$kids[[1]]$info 
info4 <- tree$node$kids[[2]]$kids[[1]]$info 
info5 <- tree$node$kids[[2]]$kids[[2]]$info

cf <- info2$coefficients
info2 <- c(sprintf("n = %s", info2$nobs), "Estimated parameters:", parse(text = paste("mu == ", round(cf[1],3))), 
                                                          parse(text = paste("sigma == ", round(cf[2],3))))

cf <- info4$coefficients
info4 <- c(sprintf("n = %s", info4$nobs), "Estimated parameters:", parse(text = paste("mu == ", round(cf[1],3))), 
                                                          parse(text = paste("sigma == ", round(cf[2],3))))

cf <- info5$coefficients
info5 <- c(sprintf("n = %s", info5$nobs), "Estimated parameters:", parse(text = paste("sigma == ", round(cf[2],3))), 
                                                          parse(text = paste("mu == ", round(cf[1],3))))

tree$node$kids[[1]]$info <- info2
tree$node$kids[[2]]$kids[[1]]$info <- info4 
tree$node$kids[[2]]$kids[[2]]$info <- info5

class(tree) <- "party"
}

pal2 <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))

## plot version using FUN and tree of class 'disttree'
plot(tree, drop = TRUE, tnex = 1.4, FUN = FUN,
     tp_args = list(fill = pal2[c(1, 3)], width = 18), 
     ip_args = list(fill = pal2[c(2, 3)]))

## plot version using replaced info slots and tree of class 'party'
if(FALSE){
  plot(tree, drop = TRUE, tnex = 1.4,
       tp_args = list(fill = pal2[c(1, 3)], FUN = identity, width = 18), 
       ip_args = list(fill = pal2[c(2, 3)]))
}
@
\end{center}
\end{frame}

%\begin{frame}
%\frametitle{Fitting a Tree}
%\center
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{tree_estpar}
%\end{frame}

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-0.5cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.74\linewidth}
<<plottree_box, fig=TRUE, echo=FALSE>>=
plot(as.constparty(tree), tnex = 1.4,
     tp_args = list(fill = pal2[c(1, 3)]), 
     ip_args = list(fill = pal2[c(2, 3)]))
@
\end{center}
\end{frame}

%\begin{frame}
%\frametitle{Fitting a Tree}
%\center
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{tree_box}
%\end{frame}

<<treedata, echo=FALSE, results=hide>>=
set.seed(54)
nobs <- 500
x <- runif(nobs, 0, 1)
mu <- sigma <- ytrue <- numeric(length = nobs)
for(i in 1:nobs) sigma[i] <- if(x[i]<=0.4) 1 else 3
for(i in 1:nobs) mu[i] <- if(x[i]<= 0.4|| x[i]>0.8) 4 else 12
y <- rnorm(nobs, mean = mu, sd = sigma)
ytrue <- mu
data <- data.frame(cbind(y,x, ytrue))

alldata <- cbind(data, mu, sigma)
odata <- alldata[order(alldata["x"]),]
@

\begin{frame}[fragile]
\frametitle{Fitting a Tree}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<plottree_xyplot, fig=TRUE, echo=FALSE>>=
par(mar=c(5.1,4.1,4.1,3.1))
plot(y=odata$y, x=odata$x, ylab = "y", xlab = "x", col = "darkgrey")
#plot(y=odata$y, x=odata$x, xaxt='n', yaxt='n', ann=FALSE, col = "darkgrey")
lines(x = odata$x, y = odata$mu, col = "red", lwd = 1.5, main = "")
lines(x = odata$x, y = odata$mu + odata$sigma, col = "red", lwd = 1.5, lty = 2, main = "")
lines(x = odata$x, y = odata$mu - odata$sigma, col = "red", lwd = 1.5, lty = 2, main = "")
@
\end{center}
\end{frame}

%\begin{frame}
%\frametitle{Fitting a Tree}
%\center
%\vspace*{-0.7cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{tree_xyplot}
%\end{frame}

%\SweaveOpts{eval=FALSE}
<<sim_1dataset, echo=FALSE, results=hide>>=
set.seed(723)
fun <- function(x){cbind(10*exp(-((4*x[,1]-2)^4)), 0.5 + 2*abs(x[,1]))}
learndata <- dgp(300, family = dist_list_normal, round.sp = 4, fun = fun)
control <-  ctree_control(teststat = "quad", testtype = "Bonferroni", mincriterion = 0.7, minbucket = 20L)
dt <- disttree(y~x1, data=learndata, family=dist_list_normal, type.tree = "ctree", control = control)

if(file.exists("sim_forest.rda")){
  load("sim_forest.rda")
} else {
  control <-  ctree_control(teststat = "quad", testtype = "Univ", mincriterion = 0, minbucket = 7L)
  df <- distforest(y~x1, data=learndata, family=dist_list_normal, ntree = 500L, type.tree = "ctree", control = control)
  df$info <- df$fitted <- df$weights <- df$nodes <- df$data <- df$terms <- df$loglik <- df$trafo <- NULL
  save(df, file = "sim_forest.rda")
}

g <- gamlss(y ~ pb(x1), sigma.formula = ~pb(x1), data = learndata, trace = FALSE)
@

\begin{frame}[fragile]
\frametitle{Simulation}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<oneforest_exp1, fig=TRUE, echo=FALSE>>=
plot1(learndata, dt=dt, df=df, g=g)
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Simulation}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<oneforest_exp1_dt, fig=TRUE, echo=FALSE>>=
plot1(learndata, dt=dt, df=df, g=g, dt.par = TRUE)
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Simulation}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<oneforest_exp1_g, fig=TRUE, echo=FALSE>>=
plot1(learndata, dt=dt, df=df, g=g, g.par = TRUE)
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Simulation}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<oneforest_exp1_df, fig=TRUE, echo=FALSE>>=
plot1(learndata, dt=dt, df=df, g=g, df.par = TRUE)
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Simulation}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<oneforest_exp1_all, fig=TRUE, echo=FALSE>>=
plot1(learndata, dt=dt, df=df, g=g, all.par = TRUE)
@
\end{center}
\end{frame}

%\begin{frame}[fragile]
%\frametitle{Simulation}
%\begin{center}
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%<<oneforest_exp1_sigma, fig=TRUE, echo=FALSE>>=
%plot1sigma(learndata,dt, df, g)
%@
%\end{center}
%\end{frame}

<<get_simlist, echo=FALSE, results=hide>>=
if(file.exists("simlist.rda")){
  load("simlist.rda")
} else {
  source("gensimlist.R")
  simlist <- gensimlist(seed = 7, nrep = 100, ntree = 100)
  save(simlist, file = "simlist.rda")
}

# plots for various data sets
# results given in simlist
# plot RMSE
plot_rmse <- function(simlist, type = c("exp", "par")){

  if(type == "exp"){
    #dev.off() 
    rmse <- cbind(simlist$av.rmse.exp.true.dt, simlist$av.rmse.exp.true.df, simlist$av.rmse.exp.true.g,
                  simlist$av.rmse.exp.obs.dt, simlist$av.rmse.exp.obs.df, simlist$av.rmse.exp.obs.g)
    colnames(rmse) <- c("dt.true", "df.true", "g.true", "dt.obs", "df.obs", "g.obs")
    ylim <- c(min(rmse), max(rmse))
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(x = simlist$x.axis, y = rmse[,"dt.true"], type = "l", col = pal["tree"], ylim = ylim,
         xlab = "", ylab = "", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n', lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"dt.obs"], type = "l", lty = 2, col = pal["tree"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"df.true"], type = "l", col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"df.obs"], type = "l", lty = 2, col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"g.true"], type = "l", col = pal["gamlss"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"g.obs"], type = "l", lty = 2, col = pal["gamlss"], lwd = 2)
    #legend('left', c("disttree", "distforest", "gamlss"), col = c(pal["tree"], pal["forest"], pal["gamlss"]), lty = 1)
    mtext(text= "smooth", side = 1, line = 1, adj = 0)
    mtext(text= "steep", side = 1, line = 1, adj = 1)
    mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
    mtext(text= "RMSE", side = 2, line = 1)
    mtext(text= "disttree", side = 2, col = pal["tree"], las = 1, line = 0.2, padj = 5)
    mtext(text= "distforest", side = 2, col = pal["forest"], las = 1, line = 0.2, padj = 13)
    mtext(text= "gamlss", side = 2, col = pal["gamlss"], las = 1, line = 0.3, padj = 16.5)

    # add arrows
    parsave <- par(new = TRUE,  mar = c(0,0,0,0))
    plot(0,0,xlim=c(0,1),ylim=c(0,1),type='n',xlab='', ylab='', col='white', axes = FALSE)

    #add arrow on the right side (downwards, for RMSE)
    arrows(0.99,0.47, 0.99,0.3,lwd=1) ## add arrow
    text(0.99,0.5, 'better')
    segments(0.99,0.53,0.99,0.7,lwd=1)

    #add arrows below
    arrows(0.45,0.085,0.3,0.085,lwd=1) ## add arrow
    arrows(0.6,0.085,0.75,0.085,lwd=1) ## add arrow

    # plot parameters  reset to prior values
    par(parsave)
  }

  if(type == "par"){
    rmse <- cbind(simlist$av.rmse.mu.dt, simlist$av.rmse.mu.df, simlist$av.rmse.mu.g, 
                  simlist$av.rmse.sigma.dt, simlist$av.rmse.sigma.df, simlist$av.rmse.sigma.g)
    colnames(rmse) <- c("mu.dt", "mu.df", "mu.g", "sigma.dt", "sigma.df", "sigma.g")
    ylim <- c(min(rmse), max(rmse))
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(x = simlist$x.axis, y = rmse[,"mu.dt"], type = "l", col = pal["tree"], ylim = ylim,
         xlab = "", ylab = "", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n', lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"sigma.dt"], type = "l", lty = 2, col = pal["tree"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"mu.df"], type = "l", col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"sigma.df"], type = "l", lty = 2, col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"mu.g"], type = "l", col = pal["gamlss"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"sigma.g"], type = "l", lty = 2, col = pal["gamlss"], lwd = 2)
    #legend('topleft', c("disttree", "distforest", "gamlss"), col = c(pal["tree"], pal["forest"],pal["gamlss"]), lty = 1)
    mtext(text= "smooth", side = 1, line = 1, adj = 0)
    mtext(text= "steep", side = 1, line = 1, adj = 1)
    mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
    mtext(text= "RMSE", side = 2, line = 1)
    mtext(text= "disttree", side = 2, col = pal["tree"], las = 1, line = 0.2, padj = -7)
    mtext(text= "distforest", side = 2, col = pal["forest"], las = 1, line = 0.2, padj = 8)
    mtext(text= "gamlss", side = 2, col = pal["gamlss"], las = 1, line = 0.3, padj = 14)

    # add arrows
    parsave <- par(new = TRUE,  mar = c(0,0,0,0))
    plot(0,0,xlim=c(0,1),ylim=c(0,1),type='n',xlab='',ylab='',col='white', axes = FALSE)

    #add arrow on the right side (downwards, for RMSE)
    arrows(0.99,0.47,0.99,0.3,lwd=1) ## add arrow
    text(0.99,0.5, 'better')
    segments(0.99,0.53,0.99,0.7,lwd=1)

    #add arrows below
    arrows(0.45,0.085,0.3,0.085,lwd=1) ## add arrow
    arrows(0.6,0.085,0.75,0.085,lwd=1) ## add arrow

    # plot parameters  reset to prior values
    par(parsave)
  }
}

# plot loglikelihood
plot_ll <- function(simlist){
  ll <- cbind(simlist$av.loglik.dt, simlist$av.loglik.df, simlist$av.loglik.g)
  colnames(ll) <- c("dt", "df","g")
  ylim <- c(min(ll), max(ll))
  par(mar=c(5.1,4.1,4.1,3.1))
  plot(x = simlist$x.axis, y = ll[,"dt"], type = "l", col = pal["tree"], ylim = ylim,
       xlab = "", ylab = "", main = "disttree vs distforest vs gamlss", xaxt='n', yaxt='n', lwd = 2)
  lines(x = simlist$x.axis, y = ll[,"df"], type = "l", col = pal["forest"], lwd = 2)
  lines(x = simlist$x.axis, y = ll[,"g"], type = "l", col = pal["gamlss"], lwd = 2)
  #legend('topright', c("disttree", "distforest", "gamlss"), col = c(pal["tree"], pal["forest"],pal["gamlss"]), lty = 1)
  mtext(text= "smooth", side = 1, line = 1, adj = 0)
  mtext(text= "steep", side = 1, line = 1, adj = 1)
  mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
  mtext(text= "Log-Likelihood", side = 2, line = 1)
  mtext(text= "disttree", side = 2, col = pal["tree"], las = 1, line = 0.2, padj = 16)
  mtext(text= "distforest", side = 2, col = pal["forest"], las = 1, line = 0.2, padj = -8)
  mtext(text= "gamlss", side = 2, col = pal["gamlss"], las = 1, line = 0.3, padj = -16)

  # add arrows
  parsave <- par(new = TRUE,  mar = c(0,0,0,0))
  plot(0,0,xlim=c(0,1),ylim=c(0,1),type='n',xlab='',ylab='',col='white', axes = FALSE)

  #add arrow on the right side (upwards, for LL)
  arrows(0.99,0.53,0.99,0.7,lwd=1) ## add arrow
  text(0.99,0.5, 'better')
  segments(0.99,0.3,0.99,0.47,lwd=1)

  #add arrows below
  arrows(0.45,0.085,0.3,0.085,lwd=1) ## add arrow
  arrows(0.6,0.085,0.75,0.085,lwd=1) ## add arrow

  # plot parameters  reset to prior values
  par(parsave)
}
@

\begin{frame}[fragile]
\frametitle{Simulation}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<plotsim_ll, fig=TRUE, echo=FALSE>>=
plot_ll(simlist)
@
\end{center}
\end{frame}

%\begin{frame}[fragile]
%\frametitle{Simulation}
%\begin{center}
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{sim_ll}
%\end{center}
%\end{frame}

%\begin{frame}[fragile]
%\frametitle{Simulation}
%\begin{center}
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{sim_RMSE_par}
%\end{center}
%\end{frame}

%\begin{frame}[fragile]
%\frametitle{Simulation}
%\begin{center}
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{sim_RMSE_exp}
%\end{center}
%\end{frame}

\begin{frame}
\frametitle{Software}
\vspace{0.4cm}
R-package \pkg{disttree} available on R-Forge:\\
\vspace{0.2cm}
\url{https://R-Forge.R-project.org/projects/partykit/}\\
\vspace{1cm}
Main functions:
\vspace{0.2cm}
\begin{itemize}
\item \code{distfit()}
\item \code{disttree()}
\item \code{distforest()}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{References}
\nocite{*}
%\nocite{hothorn2006unbiased, zeileis2008model, hothorn2015partykit, stasinopoulos2007generalized, seibold2016model}
{\footnotesize
\bibliography{ref.bib}}
\end{frame}

\begin{frame}[fragile]
\frametitle{Simulation}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<plotsim_RMSE_par, fig=TRUE, echo=FALSE>>=
plot_rmse(simlist, type = "par")
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Simulation}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<plotsim_RMSE_exp, fig=TRUE, echo=FALSE>>=
plot_rmse(simlist, type = "exp")
@
\end{center}
\end{frame}

\end{document}