### 2017-12-05 pkg/disttree/inst/CMStatistics2017/slides_rain.Rnw
edited application section and references
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1813 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 252f6545fa6410ba6db6141f20a5ec1c441237cd

\documentclass[11pt,t,usepdftitle=false,aspectratio=43]{beamer}
\usetheme[nototalframenumber, noslidenumber]{uibk}

\title{Distributional Trees and Forests}
\author{Lisa Schlosser, Torsten Hothorn,\\ Heidi Seibold, Reto Stauffer, Achim Zeileis}
\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
%\renewcommand{\mysize}{\fontsize{7.7pt}{9pt}\selectfont}
\URL{https://R-Forge.R-project.org/projects/partykit/}
\headerimage{3}

\setbeamertemplate{caption}{\insertcaption} 
%% includes a replacement for \usepackage{Sweave}
\usepackage{amsmath,tikz}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap}
\graphicspath{{plots/}}

%% colors
\definecolor{HighlightOrange}{rgb}{0.9490196,0.5725490,0.0000000}
\definecolor{HighlightBlue}{rgb}{0.4784314,0.7490196,0.9803922}

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
library("ggplot2")
theme_set(theme_bw(base_size = 18))
library("disttree")
library("Formula")
#library("partykitR1")
library("gamlss")
library("lattice")
library("crch")
library("latex2exp")
library("parallel")
library("gamlss.cens")
gen.cens(NO, type = "left")

## HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)
names(pal) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")

pallight <- hcl(c(10, 128, 260, 290, 70), 100, 50, alpha = 0.25)
names(pallight) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")

transpgrey <- rgb(0.190,0.190,0.190, alpha = 0.2)

## define distribution list:
# dist_list_normal
{

  dist_list_normal <- list()

  parnames <- c("mu", "sigma")
  etanames <- c("mu", "log(sigma)")

  ddist <-  function(y, eta, log = TRUE, weights = NULL, sum = FALSE) {

    val <- -1/2 * (log(2*pi) + 2*eta[2] + exp(log((y-eta[1])^2) - 2*eta[2]))
    if(!log) val <- exp(val)

    # par <- c(eta[1], exp(eta[2]))
    # val <- dnorm(y, mean = par[1], sd = par[2], log = log)

    if(sum) {
      if(is.null(weights)) weights <- rep.int(1, length(y))
      val <- sum(weights * val, na.rm = TRUE)
    }
    return(val)
  }

  sdist <- function(y, eta, weights = NULL, sum = FALSE) {

    score <- cbind(exp(-2*eta[2]) * (y-eta[1]), 
                   -1 + exp(-2*eta[2] + log((y-eta[1])^2)))

    # par <- c(eta[1], exp(eta[2])) 
    # score <- cbind(1/par[2]^2 * (y-par[1]), 
    #                (-1/par[2] + ((y - par[1])^2)/(par[2]^3)) * exp(eta[2]))

    score <- as.matrix(score)
    colnames(score) <- etanames
    if(sum) {
      if(is.null(weights)) weights <- rep.int(1, length(y))
      # if score == Inf replace score with 1.7e308 because Inf*0 would lead to NaN -> gradient is NaN
      score[score==Inf] = 1.7e308
      score <- colSums(weights * score, na.rm = TRUE)
    }
    return(score)
  }

  hdist <- function(y, eta, weights = NULL) {    
    ny <- length(y)
    if(is.null(weights)) weights <- rep.int(1, ny)

    d2ld.etamu2 <- sum(weights * rep.int(-exp(-2*eta[2]), ny))
    d2ld.etamu.d.etasigma <- sum(weights * (-2)*(y-eta[1]) * exp(-2*eta[2]), na.rm = TRUE)          # should be 0 for exact parameters (here: observed hess)
    d2ld.etasigma2 <- sum(weights * (-2)*exp(log((y-eta[1])^2) - 2*eta[2]), na.rm = TRUE)

    # par <- c(eta[1], exp(eta[2]))                           
    # d2ld.etamu2 <- sum(weights * rep.int(-1/par[2]^2, ny))
    # d2ld.etamu.d.etasigma <- sum(weights * (-2)*(y-par[1])/par[2]^2), na.rm = TRUE)          # should be 0 for exact parameters (here: observed hess)
    # d2ld.etasigma2 <- sum(weights * (-2)*(y-par[1])^2/par[2]^2, na.rm = TRUE)

    hess <- matrix(c(d2ld.etamu2, d2ld.etamu.d.etasigma, d2ld.etamu.d.etasigma, d2ld.etasigma2), nrow = 2)
    colnames(hess) <- rownames(hess) <-  etanames

    return(hess)
  }

  ## additional functions pdist, qdist, rdist
  pdist <- pnorm
  qdist <- qnorm
  rdist <- rnorm

  link <- c("identity", "log")

  linkfun <- function(par) {
    eta <- c(par[1], log(par[2]))
    names(eta) <- etanames
    return(eta)
  }

  linkinv <- function(eta) {
    par <- c(eta[1], exp(eta[2]))
    names(par) <- parnames
    return(par)
  }

  linkinvdr <- function(eta) {
    dpardeta <- c(1, exp(eta[2]))
    names(dpardeta) <- parnames
    return(dpardeta)
  }

  startfun <- function(y, weights = NULL){
    if(is.null(weights)) {
      mu <- mean(y)
      sigma <- sqrt(1/length(y) * sum((y - mu)^2))
    } else {
      mu <- weighted.mean(y, weights)
      sigma <- sqrt(1/sum(weights) * sum(weights * (y - mu)^2))
    }
    starteta <- c(mu, log(sigma))
    names(starteta) <- etanames
    return(starteta)
  }

  mle <- TRUE

  dist_list_normal <- list(family.name = "Normal Distribution",
                           ddist = ddist, 
                           sdist = sdist, 
                           hdist = hdist,
                           pdist = pdist,
                           qdist = qdist,
                           rdist = rdist,
                           link = link, 
                           linkfun = linkfun, 
                           linkinv = linkinv, 
                           linkinvdr = linkinvdr,
                           startfun = startfun,
                           mle = mle,
                           gamlssobj = FALSE,
                           censored = FALSE
  )
}

# dist_list_cens_normal
{

  dist_list_cens_normal <- list()

  parnames <- c("mu", "sigma")
  etanames <- c("mu", "log(sigma)")

  ddist <-  function(y, eta, log = TRUE, weights = NULL, sum = FALSE, left = 0, right = Inf) {     
    par <- c(eta[1], exp(eta[2]))
    val <- crch::dcnorm(x = y, mean = par[1], sd = par[2], left = left, right = right, log = log)
    if(sum) {
      if(is.null(weights)) weights <- if(is.matrix(y)) rep.int(1, dim(y)[1]) else rep.int(1, length(y))
      val <- sum(weights * val, na.rm = TRUE)
    }
    return(val)
  }

  sdist <- function(y, eta, weights = NULL, sum = FALSE, left = 0, right = Inf) {   
    par <- c(eta[1], exp(eta[2]))
    # y[y==0] <- 1e-323

    score_m <- crch:::scnorm(x = y, mean = par[1], sd = par[2], which = "mu", left = left, right = right)
    score_s <- crch:::scnorm(x = y, mean = par[1], sd = par[2], which = "sigma", left = left, right = right) * exp(eta[2]) # inner derivation exp(eta[2])
    score <- cbind(score_m, score_s)
    score <- as.matrix(score)
    colnames(score) <- etanames
    if(sum) {
      if(is.null(weights)) weights <- rep.int(1, length(y)[1])
      # if score == Inf replace score with 1.7e308 because Inf*0 would lead to NaN (0 in weights)
      score[score==Inf] = 1.7e308
      score <- colSums(weights * score, na.rm = TRUE)
      #if(any(is.nan(score))) print(c(eta, "y", y))
    }
    return(score)
  }

  hdist <- function(y, eta, weights = NULL, left = 0, right = Inf) {    
    ny <- length(y)
    if(is.null(weights)) weights <- rep.int(1, ny)

    par <- c(eta[1], exp(eta[2]))                           
    # y[y==0] <- 1e-323

    d2mu <- crch:::hcnorm(x = y, mean = par[1], sd = par[2], which = "mu", left = left, right = right)
    d2sigma <- crch:::hcnorm(x = y, mean = par[1], sd = par[2], which = "sigma", left = left, right = right)
    dmudsigma <- crch:::hcnorm(x = y, mean = par[1], sd = par[2], which = "mu.sigma", left = left, right = right) # FIX: order?
    dsigmadmu <- crch:::hcnorm(x = y, mean = par[1], sd = par[2], which = "sigma.mu", left = left, right = right) # FIX: order?
    dsigma <- crch:::scnorm(x = y, mean = par[1], sd = par[2], which = "sigma", left = left, right = right)

    d2ld.etamu2 <- sum(weights * d2mu, na.rm = TRUE)
    d2ld.etamu.d.etasigma <- sum(weights * dmudsigma * par[2], na.rm = TRUE)
    d2ld.etasigma.d.etamu <- sum(weights * dsigmadmu * par[2], na.rm = TRUE)
    d2ld.etasigma2 <- sum(weights * (d2sigma * exp(2*eta[2]) + dsigma * par[2]), na.rm = TRUE)

    hess <- matrix(c(d2ld.etamu2, d2ld.etamu.d.etasigma, d2ld.etasigma.d.etamu, d2ld.etasigma2), nrow = 2)
    colnames(hess) <- rownames(hess) <-  etanames

    return(hess)
  }

  ## additional functions pdist, qdist, rdist
  pdist <- function(q, eta, lower.tail = TRUE, log.p = FALSE) crch:::pcnorm(q, mean = eta[1], sd = eta[2], 
                                                                            lower.tail = lower.tail, log.p = log.p, 
                                                                            left = left, right = right)
  qdist <- function(p, eta, lower.tail = TRUE, log.p = FALSE) crch:::qcnorm(p, mean = eta[1], sd = eta[2], 
                                                                            lower.tail = lower.tail, log.p = log.p, 
                                                                            left = left, right = right)
  rdist <- function(n, eta) crch:::rcnorm(n, mean = eta[1], sd = eta[2], left = left, right = right)

  link <- c("identity", "log")

  linkfun <- function(par) {
    eta <- c(par[1], log(par[2]))
    names(eta) <- etanames
    return(eta)
  }

  linkinv <- function(eta) {
    par <- c(eta[1], exp(eta[2]))
    names(par) <- parnames
    return(par)
  }

  linkinvdr <- function(eta) {
    dpardeta <- c(1, exp(eta[2]))
    names(dpardeta) <- parnames
    return(dpardeta)
  }

  startfun <- function(y, weights = NULL){
    yc <- pmax(0,y)  # optional ?
    if(is.null(weights)) {
      mu <- mean(yc)
      sigma <- sqrt(1/length(yc) * sum((yc - mu)^2))
    } else {
      mu <- weighted.mean(yc, weights)
      sigma <- sqrt(1/sum(weights) * sum(weights * (yc - mu)^2))
    }
    starteta <- c(mu, log(sigma))
    names(starteta) <- etanames
    return(starteta)
  }

  mle <- FALSE

  dist_list_cens_normal <- list(family.name = "censored Normal Distribution",
                                ddist = ddist, 
                                sdist = sdist, 
                                hdist = hdist,
                                pdist = pdist,
                                qdist = qdist,
                                rdist = rdist,
                                link = link, 
                                linkfun = linkfun, 
                                linkinv = linkinv, 
                                linkinvdr = linkinvdr,
                                startfun = startfun,
                                mle = mle,
                                gamlssobj = FALSE,
                                censored = TRUE
  )
}

######## data generating process based on a predefined tree or a parameter function
# if tree: 2 splits => 3 terminal nodes
# 10 split variables (x1, ..., x10) are availabe (5 numeric, 2 binary, 3 categorical)
#
# input: n            ..... nr of observations
#        family       ..... distribution of the generated observations
#        split.matrix ..... indices of the variables used for the splits together with the corresponding split points
#        parm.matrix  ..... set of distribution parameters for each subgroup
#        fun          ..... parameter function
#
# output: data.frame with generated observations y (generated seperatly in each subgroups with given distributions parameters),
#                         the given split variables x1, ..., x10 for each observation and
#                         the index of the subgroup for each observation or
#                         the distribution parameter for each observation

# FIX: until now only complete lists can be handed over (dist_list...)
dgp <- function(n, family = dist_list_normal, 
                fun = NULL,
                split.matrix = matrix(nrow = 2, ncol = 2), 
                par.matrix = matrix(nrow = 3, ncol = 2), 
                round.sp = 3)
{

  # generating the possible split variables
  x1 <- runif(n,-0.4,1)
  x2 <- runif(n,-10,10)
  x3 <- runif(n,0,100)
  x4 <- x1 + rnorm(n, sd = 0.1)
  x5 <- x1 + rnorm(n, sd = 0.3)
  x6 <- rbinom(n,1,0.5)
  x7 <- rbinom(n,1,0.5)
  x8 <- sample(1:4, n, replace = TRUE) 
  x9 <- sample(1:5, n, replace = TRUE) 
  x10 <- sample(1:7, n, replace = TRUE)
  #x11 <- runif(n,-0.5,1)
  x <- cbind(x1,x2,x3,x4,x5,x6,x7,x8,x9,x10)
  # reduce nr of possible split points by rounding values of split variables
  x <- round(x, digits = round.sp)

  y <- vector(mode = "numeric", length = n)

  # getting the random function
  if(is.function(family)) family <- family()
  if(inherits(family, "gamlss.family")) {
    rfun <- get(paste0("r",family$family[[1]]))
  } else {
    if(family$family.name == "Normal Distribution") rfun <- rNO
    if(family$family.name == "censored Normal Distribution") rfun <- rNO
    if(family$family.name == "Poisson Distribution") rfun <- rPO
    if(family$family.name == "Weibull Distribution") rfun <- survival:::rsurvreg 
  }

  if(is.null(fun)) {
    index <- vector(mode = "numeric", length = n)
    sv = split.matrix[,1]    # index of split variables
    sp = split.matrix[,2]    # split points
    par = par.matrix

    # splitting based on the given split variables and split points and
    # generating the observations seperatly for each subgroup (with the corresponding given distribution parameters)
    for(k in 1:n){
      if(x[k, sv[1]] <= sp[1]){
        y[k] <- do.call(rfun, as.list(c(1,par[1,])))
        index[k] <- 2L
      } else {
        if(x[k, sv[2]] <= sp[2]){
          y[k] <- do.call(rfun, as.list(c(1,par[2,]))) 
          index[k] <- 4L
        } else {
          y[k] <- do.call(rfun, as.list(c(1,par[3,])))
          index[k] <- 5L
        }
      }
    }
    d <- as.data.frame(cbind(y, x, index))
    colnames(d) <- c("y", paste0("x", c(1:10)), "index")

  } else {

    dpar <- fun(x)

    if(!(family$family.name == "Poisson Distribution")){
      y <- rfun(n, dpar[,1], dpar[,2])
      d <- as.data.frame(cbind(y, x, dpar))
      colnames(d) <- c("y", paste0("x", c(1:10)), "mu", "sigma")
    } else {
      y <- rfun(n, dpar)
      d <- as.data.frame(cbind(y, x, dpar))
      colnames(d) <- c("y", paste0("x", c(1:10)), "lambda")
    }
  }

  if(family$family.name == "censored Normal Distribution") {
    #d$ystar <- d$y
    d$y <- pmax(d$y, 0)
  }

  return(d)
}

############################################ 
# parameter functions

### plotting function for one data set
plot1 <- function(learndata, dt=NULL, df=NULL, g=NULL, all.par = FALSE, df.par = FALSE, g.par = FALSE, dt.par = FALSE, all.var = FALSE) 
{
  plotdata <- cbind(learndata[,c("y","x1","mu","sigma")],
                    df$fitted.par,
                    dt$fitted.par,
                    #coef(dt)[paste(dt$fitted[,1]),],
                    g$mu.fv, g$sigma.fv)

  colnames(plotdata) <- c("y","x1","true.mu","true.sigma", 
                          "fitted.mu.df","fitted.sigma.df",
                          "fitted.mu.dt","fitted.sigma.dt",
                          "fitted.mu.g","fitted.sigma.g")
  sp <- plotdata[order(plotdata["x1"]),]

  if(!(any(dt.par, df.par, g.par, all.par, all.var))){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "True parameters", cex.main = 1.2, xaxt="n", yaxt="n", xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
  col = transpgrey, border = "transparent")
    #lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", lty = 1, col = 'grey')
    #lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", lty = 1, col = 'grey')
    #legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$')), 
    #       col = c('black', transpgrey), lty = 1, bty = "n", lwd = 2.5)
    legend(x = -0.5, y = 14, expression(mu %+-% sigma), bty = "n")
  }

  # disttree
  if(dt.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "disttree", col.main = pal["tree"], cex.main = 1.2, xaxt="n", yaxt="n", xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    #polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
    #col = transpgrey, border = "transparent")
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.dt, type = "l", col = pal["tree"], lwd = 2.5)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.dt + sp$fitted.sigma.dt, rev(sp$fitted.mu.dt - sp$fitted.sigma.dt)),
  col = pallight["tree"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.dt + sp$fitted.sigma.dt, type = "l", col = 'greenyellow')
    #lines(x = sp$x, y = sp$fitted.mu.dt - sp$fitted.sigma.dt, type = "l", col = 'greenyellow')
    #legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
    #                    TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')), 
    #       col = c('black','grey', pal["tree"], pallight["tree"]), lty = 1, bty = "n", lwd = 2.5)
    legend(x = -0.5, y = 14, expression(mu %+-% sigma), bty = "n")
  }

  # distforest
  if(df.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "distforest", col.main = pal["forest"], cex.main = 1.2, xaxt="n", yaxt="n", xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    #polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
    #col = transpgrey, border = "transparent")
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.df, type = "l", col = pal["forest"], lwd = 2.5)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.df + sp$fitted.sigma.df, rev(sp$fitted.mu.df - sp$fitted.sigma.df)),
  col = pallight["forest"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.df + sp$fitted.sigma.df, type = "l", col = 'orange')
    #lines(x = sp$x, y = sp$fitted.mu.df - sp$fitted.sigma.df, type = "l", col = 'orange')
    #legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
    #                    TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')), 
    #       col = c('black','grey',pal["forest"], pallight["forest"]), lty = 1, bty = "n", lwd = 2.5)
    legend(x = -0.5, y = 14, expression(mu %+-% sigma), bty = "n")
  }

  # gamlss
  if(g.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey",  main = "gamlss", col.main = pal["gamlss"], cex.main = 1.2, xaxt="n", yaxt="n", xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    #polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
    #col = transpgrey, border = "transparent")
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.g, type = "l", col = pal["gamlss"], lwd = 2.5)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.g + sp$fitted.sigma.g, rev(sp$fitted.mu.g - sp$fitted.sigma.g)),
  col = pallight["gamlss"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.g + sp$fitted.sigma.g, type = "l", col = 'lightblue')
    #lines(x = sp$x, y = sp$fitted.mu.g - sp$fitted.sigma.g, type = "l", col = 'lightblue')
    #legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
    #                    TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')),  
    #       col = c('black','grey',pal["gamlss"], pallight["gamlss"]), lty = 1, bty = "n", lwd = 2.5)
    legend(x = -0.5, y = 14, expression(mu %+-% sigma), bty = "n")
  }

  # disttree vs. distforest vs. gamlss
  if(all.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", xaxt="n", yaxt="n", xlab = "", ylab = "")
    title(main = "disttree", col.main = pal["tree"], cex.main = 1.2, font.main = 2, adj = 0.19)
    title(main = "distforest", col.main = pal["forest"], cex.main = 1.2, font.main = 2, adj = 0.5)
    title(main = "gamlss", col.main = pal["gamlss"], cex.main = 1.2, font.main = 2, adj = 0.8)
    title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.35)
    title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.65)

    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$fitted.mu.dt, type = "l", col = pal["tree"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.df, type = "l", col = pal["forest"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.g, type = "l", col = pal["gamlss"], lwd = 2)
    #legend('topleft', c(TeX('$\\mu$'), TeX('$\\hat{\\mu}$ disttree'),
    #                    TeX('$\\hat{\\mu}$ distforest'), TeX('$\\hat{\\mu}$ gamlss')), 
    #       col = c('black',pal["tree"],pal["forest"],pal["gamlss"]), lty = 1, bty = "n", lwd = 2.5)
    legend(x = -0.5, y = 13.95, expression(mu), bty = "n")
  }

  # variance: disttree vs. distforest vs. gamlss
  if(all.var){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", xaxt="n", yaxt="n", xlab = "", ylab = "")
    title(main = "disttree", col.main = pal["tree"], cex.main = 1.2, font.main = 2, adj = 0.19)
    title(main = "distforest", col.main = pal["forest"], cex.main = 1.2, font.main = 2, adj = 0.5)
    title(main = "gamlss", col.main = pal["gamlss"], cex.main = 1.2, font.main = 2, adj = 0.8)
    title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.35)
    title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.65)

    #polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
    #col = transpgrey, border = "transparent")
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'black')
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.dt + sp$fitted.sigma.dt, rev(sp$fitted.mu.dt - sp$fitted.sigma.dt)),
  col = pallight["tree"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.dt + sp$fitted.sigma.dt, type = "l", col = pal["tree"], lwd = 2)
    #lines(x = sp$x, y = sp$fitted.mu.dt - sp$fitted.sigma.dt, type = "l", col = pal["tree"], lwd = 2)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.df + sp$fitted.sigma.df, rev(sp$fitted.mu.df - sp$fitted.sigma.df)),
  col = pallight["forest"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.df + sp$fitted.sigma.df, type = "l", col = pal["forest"], lwd = 2)
    #lines(x = sp$x, y = sp$fitted.mu.df - sp$fitted.sigma.df, type = "l", col = pal["forest"], lwd = 2)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.g + sp$fitted.sigma.g, rev(sp$fitted.mu.g - sp$fitted.sigma.g)),
  col = pallight["gamlss"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.g + sp$fitted.sigma.g, type = "l", col = pal["gamlss"], lwd = 2)
    #lines(x = sp$x, y = sp$fitted.mu.g - sp$fitted.sigma.g, type = "l", col = pal["gamlss"], lwd = 2)
    #legend('topleft', c(TeX('$\\mu \\pm \\sigma$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$ disttree'),
    #                    TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$ distforest'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$ gamlss')), 
    #       col = c('black',pal["tree"],pal["forest"],pal["gamlss"]), lty = 1, bty = "n", lwd = 2.5)
  legend(x = -0.5, y = 14, expression(mu %+-% sigma), bty = "n")
  }

}

@

\begin{document}

\section{Distributional Trees and Forests}

\subsection{Motivation}

\begin{frame}[fragile]
\frametitle{Motivation}
\vspace{-0.41cm}
\begin{figure}[!htb]
\minipage{0.29\textwidth}
\begin{center}
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@
\visible<2->{
<<plot_motivation_GLM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
\end{center}
\endminipage
\visible<3->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
\begin{center}
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@
\visible<3->{
<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
\end{center}
\endminipage
\visible<4->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
\begin{center}
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\visible<4->{
<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@

}
\end{center}
\endminipage

\vspace{0.5cm}
\minipage{0.25\textwidth}
\begin{center}
\visible<2->{
LM, GLM\\
\vspace{0.5cm}
\code{lm}\\
\code{glm}\\
\vspace{1.5cm}}
\end{center}
\endminipage
\hspace{1.1cm}
\minipage{0.25\textwidth}
\begin{center}
\visible<3->{
GAM\\
\vspace{0.5cm}
\code{mgcv}\\
\code{VGAM}\\
$\ldots$\\
\vspace{1cm}}
\end{center}
\endminipage
\hspace{1.1cm}
\minipage{0.25\textwidth}
\begin{center}
\visible<4->{
GAMLSS\\
\vspace{0.5cm}
\code{gamlss}\\
\code{mgcv}\\
\code{VGAM}\\
\code{gamboostLSS}\\
$\ldots$}
\end{center}
\endminipage
\end{figure}
\end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}
\vspace{-0.2cm}
\begin{figure}[!htb]
\minipage{0.29\textwidth}
\begin{center}
<<motivation_regtree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- ytree <- yforest <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))
y <- ytrue + rnorm(nobs,0,0.3)
for(i in 1:nobs) ytree[i] <- if(x[i]<1/3) 0.5 else {if(x[i]<2/3) 2 else 1}
@
\visible<1->{
<<plot_motivation_regtree, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "grey", lwd=5, main = "")
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
@
}
\end{center}

\endminipage
\visible<2->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
 \begin{center}
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@
\visible<2->{
<<plot_motivation_randforest, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "grey", lwd=5, main = "")
lines(x = x, y = yforest, col = pal["forest"], lwd=7, main = "")
@
}
\end{center}

\endminipage
\visible<3->{{\LARGE$\rightarrow$}}
\minipage{0.29\textwidth}
  \visible<3->{
  \begin{center}
  \begin{tikzpicture}
  \draw[line width=0.2mm, black] (0,0) rectangle (2.6,2.24);
  \node[rectangle, align=center] (sq) at (1.3, 1.12) {\Huge{?}};
  \end{tikzpicture}
  \end{center}}
\endminipage

\minipage{0.29\textwidth}
\begin{center}
\vspace{0.0cm}
\visible<1->{
Regression Tree\\
\vspace{0.5cm}
\resizebox{0.3\textwidth}{!}{
\begin{tikzpicture}
  \node[ellipse, fill=HighlightBlue!70, align=center] (n0) at (1, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n1) at (0.5, 1) {};
  \draw[-, line width=1pt] (n0) -- (n1);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n2) at (1.5, 1) {};
  \draw[-, line width=1pt] (n0) -- (n2);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n3) at (1, 0) {};
  \draw[-, line width=1pt] (n2) -- (n3);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n4) at (2, 0) {};
  \draw[-, line width=1pt] (n2) -- (n4);
\end{tikzpicture}}
\vspace{0.3cm}\\
\code{rpart}\\
\code{party(kit)}\\
\vspace{1cm}}
\end{center}
\endminipage
\hspace{0.65cm}
\minipage{0.29\textwidth}
\begin{center}
\vspace{0.0cm}
\visible<2->{
Random Forest\\
\vspace{0.5cm}
\resizebox{0.9\textwidth}{!}{
\begin{tikzpicture}
  \node[ellipse, fill=HighlightBlue!70, align=center] (n00) at (1, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n01) at (0.5, 1) {};
  \draw[-, line width=1pt] (n00) -- (n01);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n02) at (1.5, 1) {};
  \draw[-, line width=1pt] (n00) -- (n02);

  \node[ellipse, fill=HighlightBlue!70, align=center] (n10) at (3, 2) {};
  \node[ellipse, fill=HighlightBlue!70, align=center] (n11) at (2.5, 1) {};
  \draw[-, line width=1pt] (n10) -- (n11);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n12) at (3.5, 1) {};
  \draw[-, line width=1pt] (n10) -- (n12);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n13) at (2, 0) {};
  \draw[-, line width=1pt] (n11) -- (n13);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n14) at (2.8, 0) {};
  \draw[-, line width=1pt] (n11) -- (n14);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n15) at (3.2, 0) {};
  \draw[-, line width=1pt] (n12) -- (n15);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n16) at (4, 0) {};
  \draw[-, line width=1pt] (n12) -- (n16);

  \node[ellipse, fill=HighlightBlue!70, align=center] (n20) at (5, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n21) at (4.5, 1) {};
  \draw[-, line width=1pt] (n20) -- (n21);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n22) at (5.5, 1) {};
  \draw[-, line width=1pt] (n20) -- (n22);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n23) at (5, 0) {};
  \draw[-, line width=1pt] (n22) -- (n23);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n24) at (6, 0) {};
  \draw[-, line width=1pt] (n22) -- (n24);
\end{tikzpicture}
}
\vspace{0.5cm}\\
\code{randomForest}\\
\code{ranger}\\
\code{party(kit)}\\
$\ldots$}
\end{center}
\endminipage
\hspace{0.65cm}
\minipage{0.28\textwidth}
\begin{center}
\vspace{0.68cm}
\visible<3->{
Distributional trees and forests\\
\vspace{1.7cm}
\code{disttree}\\
based on \code{partykit}\\
\vspace{1cm}}
\end{center}
\endminipage
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Goals}

\textbf{Distributional:} \\
\vspace{0.1 cm}
\hspace{0.5cm} Specify the complete probability distribution\\
\hspace{0.5cm} (including location, scale and shape).

\medskip

{\bf Tree:}
\begin{itemize}
  \item Automatic detection of steps and abrupt changes. %(data driven)
  \item Capture non-linear and non-additive effects and interactions.
\end{itemize}

\medskip

{\bf Forest:}
\begin{itemize}
  \item Smoother effects.
  \item Stabilization and regularization of the model.
\end{itemize}
\end{frame}

\subsection{Building Distributional Trees and Forests}
\begin{frame}
\frametitle{Building Distributional Trees and Forests}
{\bf Tree:}
\begin{enumerate}
\item Specify a distribution with log-likelihood function $\ell(\theta; y)$.
\item Estimate $\hat{\theta}$ via maximum likelihood.
\item Test for associations or instabilities of the scores $\frac{\partial \ell}{\partial \theta}(\hat{\theta};y_i)$ and each partitioning variable $z_i$.
%\item Assess whether the \emph{model scores} are associated with
%    (or change along) any of the available covariates -- e.g.,
%    using parameter instability tests (\emph{strucchange}) or
%    conditional inference (\emph{coin}).
\item Split the sample along the partitioning variable with the strongest association or instability.
    Choose breakpoint with highest improvement in log-likelihood.
\item Repeat steps 2--4 recursively in the subgroups
    until some stopping criterion is met.
    %-- e.g., for significance or sample size.
%\item Choose the variable with the strongest association
%\item Choose the split point which leads to the highest improvement of the model
%\item Split and repeat 2-6 in each node until a stopping criterion is met
\end{enumerate}

\vspace{0.4cm}
{\bf Forest:} Ensemble of trees.
\begin{itemize}
\item Bootstrap or subsamples.
\item Random input variable sampling.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Prediction}
%\vspace{1cm}
{\bf Tree:}\\
\begin{enumerate}
\item Select the subsample of the learning data which ends up in the same terminal node as the new observation.
\item Estimate $\hat{\theta}$ on this subsample.
\end{enumerate}
\vspace{0.5cm}

{\bf Forest:}\\
\begin{enumerate}
\item For each learning observation calculate a weight by counting the number of trees in which it ends up in the same terminal node as the new observation.
\item Estimate $\hat{\theta}$ on the weighted learning data.
\end{enumerate}
% but weighted by similarities of the learning data and the new observation over all trees
%\item Evaluate final node for new observation in each of the trees and investigate similarities with each of the learning observations
%\item Fit a distributional model for the new observation with different influences of the learning observations defined by weights
\end{frame}

%\SweaveOpts{eval=FALSE}
<<sim_1dataset, echo=FALSE, results=hide>>=
set.seed(723)
fun <- function(x){cbind(10*exp(-((4*x[,1]-2)^4)), 0.5 + 2*abs(x[,1]))}
learndata <- dgp(300, family = dist_list_normal, round.sp = 4, fun = fun)
control <-  ctree_control(teststat = "quad", testtype = "Bonferroni", mincriterion = 0.7, minbucket = 20L)
dt <- disttree(y~x1, data=learndata, family=dist_list_normal, type.tree = "ctree", control = control)

if(file.exists("sim_forest.rda")){
  load("sim_forest.rda")
} else {
  control <-  ctree_control(teststat = "quad", testtype = "Univ", mincriterion = 0, minbucket = 7L)
  df <- distforest(y~x1, data=learndata, family=dist_list_normal, ntree = 500L, type.tree = "ctree", control = control)
  df$info <- df$fitted <- df$weights <- df$nodes <- df$data <- df$terms <- df$loglik <- df$trafo <- NULL
  save(df, file = "sim_forest.rda")
}

g <- gamlss(y ~ pb(x1), sigma.formula = ~pb(x1), data = learndata, trace = FALSE)
@

\begin{frame}[fragile]
\frametitle{Simulation}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<oneforest_exp1, fig=TRUE, echo=FALSE>>=
plot1(learndata, dt=dt, df=df, g=g)
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Simulation}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<oneforest_exp1_dt, fig=TRUE, echo=FALSE>>=
plot1(learndata, dt=dt, df=df, g=g, dt.par = TRUE)
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Simulation}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<oneforest_exp1_g, fig=TRUE, echo=FALSE>>=
plot1(learndata, dt=dt, df=df, g=g, g.par = TRUE)
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Simulation}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<oneforest_exp1_df, fig=TRUE, echo=FALSE>>=
plot1(learndata, dt=dt, df=df, g=g, df.par = TRUE)
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Simulation}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<oneforest_exp1_all, fig=TRUE, echo=FALSE>>=
plot1(learndata, dt=dt, df=df, g=g, all.par = TRUE)
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Simulation}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<oneforest_exp1_allvar, fig=TRUE, echo=FALSE>>=
plot1(learndata, dt=dt, df=df, g=g, all.var = TRUE)
@
\end{center}
\end{frame}

%\begin{frame}[fragile]
%\frametitle{Simulation}
%\begin{center}
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%<<oneforest_exp1_sigma, fig=TRUE, echo=FALSE>>=
%plot1sigma(learndata,dt, df, g)
%@
%\end{center}
%\end{frame}

<<get_simlist, echo=FALSE, results=hide>>=
if(file.exists("simlist.rda")){
  load("simlist.rda")
} else {
  source("gensimlist.R")
  simlist <- gensimlist(seed = 9, nrep = 150, ntree = 100)
  save(simlist, file = "simlist.rda")
}

# plots for various data sets
# results given in simlist
# plot RMSE
plot_rmse <- function(simlist, type = c("exp", "par")){

  if(type == "exp"){
    #dev.off() 
    rmse <- cbind(simlist$av.rmse.exp.true.dt, simlist$av.rmse.exp.true.df, simlist$av.rmse.exp.true.g,
                  simlist$av.rmse.exp.obs.dt, simlist$av.rmse.exp.obs.df, simlist$av.rmse.exp.obs.g)
    colnames(rmse) <- c("dt.true", "df.true", "g.true", "dt.obs", "df.obs", "g.obs")
    ylim <- c(min(rmse), max(rmse))
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(x = simlist$x.axis, y = rmse[,"dt.true"], type = "l", col = pal["tree"], ylim = ylim,
         xlab = "", ylab = "", xaxt="n", yaxt="n", lwd = 2)
    title(main = "disttree", col.main = pal["tree"], cex.main = 1.2, font.main = 2, adj = 0.19)
    title(main = "distforest", col.main = pal["forest"], cex.main = 1.2, font.main = 2, adj = 0.5)
    title(main = "gamlss", col.main = pal["gamlss"], cex.main = 1.2, font.main = 2, adj = 0.8)
    title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.35)
    title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.65)
    lines(x = simlist$x.axis, y = rmse[,"dt.obs"], type = "l", lty = 2, col = pal["tree"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"df.true"], type = "l", col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"df.obs"], type = "l", lty = 2, col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"g.true"], type = "l", col = pal["gamlss"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"g.obs"], type = "l", lty = 2, col = pal["gamlss"], lwd = 2)
    #legend('left', c("disttree", "distforest", "gamlss"), col = c(pal["tree"], pal["forest"], pal["gamlss"]), lty = 1)
    mtext(text= "smooth", side = 1, line = 1, adj = 0)
    mtext(text= "steep", side = 1, line = 1, adj = 1)
    mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
    mtext(text= "RMSE", side = 2, line = 1)
    mtext(text= "disttree", side = 2, col = pal["tree"], las = 1, line = 0.2, padj = 5)
    mtext(text= "distforest", side = 2, col = pal["forest"], las = 1, line = 0.2, padj = 13)
    mtext(text= "gamlss", side = 2, col = pal["gamlss"], las = 1, line = 0.3, padj = 16.5)

    # add arrows
    parsave <- par(new = TRUE,  mar = c(0,0,0,0))
    plot(0,0,xlim=c(0,1),ylim=c(0,1),type="n",xlab='', ylab='', col='white', axes = FALSE)

    #add arrow on the right side (downwards, for RMSE)
    arrows(0.99,0.47, 0.99,0.3,lwd=1) ## add arrow
    text(0.99,0.5, 'better')
    segments(0.99,0.53,0.99,0.7,lwd=1)

    #add arrows below
    arrows(0.45,0.085,0.3,0.085,lwd=1) ## add arrow
    arrows(0.6,0.085,0.75,0.085,lwd=1) ## add arrow

    # plot parameters  reset to prior values
    par(parsave)
  }

  if(type == "par"){
    rmse <- cbind(simlist$av.rmse.mu.dt, simlist$av.rmse.mu.df, simlist$av.rmse.mu.g, 
                  simlist$av.rmse.sigma.dt, simlist$av.rmse.sigma.df, simlist$av.rmse.sigma.g)
    colnames(rmse) <- c("mu.dt", "mu.df", "mu.g", "sigma.dt", "sigma.df", "sigma.g")
    ylim <- c(min(rmse), max(rmse))
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(x = simlist$x.axis, y = rmse[,"mu.dt"], type = "l", col = pal["tree"], ylim = ylim,
         xlab = "", ylab = "", xaxt="n", yaxt="n", lwd = 2)
    title(main = "disttree", col.main = pal["tree"], cex.main = 1.2, font.main = 2, adj = 0.19)
    title(main = "distforest", col.main = pal["forest"], cex.main = 1.2, font.main = 2, adj = 0.5)
    title(main = "gamlss", col.main = pal["gamlss"], cex.main = 1.2, font.main = 2, adj = 0.8)
    title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.35)
    title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.65)
    lines(x = simlist$x.axis, y = rmse[,"sigma.dt"], type = "l", lty = 2, col = pal["tree"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"mu.df"], type = "l", col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"sigma.df"], type = "l", lty = 2, col = pal["forest"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"mu.g"], type = "l", col = pal["gamlss"], lwd = 2)
    lines(x = simlist$x.axis, y = rmse[,"sigma.g"], type = "l", lty = 2, col = pal["gamlss"], lwd = 2)
    #legend('topleft', c("disttree", "distforest", "gamlss"), col = c(pal["tree"], pal["forest"],pal["gamlss"]), lty = 1)
    mtext(text= "smooth", side = 1, line = 1, adj = 0)
    mtext(text= "steep", side = 1, line = 1, adj = 1)
    mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
    mtext(text= "RMSE", side = 2, line = 1)
    mtext(text= "disttree", side = 2, col = pal["tree"], las = 1, line = 0.2, padj = -7)
    mtext(text= "distforest", side = 2, col = pal["forest"], las = 1, line = 0.2, padj = 8)
    mtext(text= "gamlss", side = 2, col = pal["gamlss"], las = 1, line = 0.3, padj = 14)

    # add arrows
    parsave <- par(new = TRUE,  mar = c(0,0,0,0))
    plot(0,0,xlim=c(0,1),ylim=c(0,1),type="n",xlab='',ylab='',col='white', axes = FALSE)

    #add arrow on the right side (downwards, for RMSE)
    arrows(0.99,0.47,0.99,0.3,lwd=1) ## add arrow
    text(0.99,0.5, 'better')
    segments(0.99,0.53,0.99,0.7,lwd=1)

    #add arrows below
    arrows(0.45,0.085,0.3,0.085,lwd=1) ## add arrow
    arrows(0.6,0.085,0.75,0.085,lwd=1) ## add arrow

    # plot parameters  reset to prior values
    par(parsave)
  }
}

# plot loglikelihood
plot_ll <- function(simlist){
  ll <- cbind(simlist$av.loglik.dt, simlist$av.loglik.df, simlist$av.loglik.g)
  colnames(ll) <- c("dt", "df","g")
  ylim <- c(min(ll), max(ll))
  par(mar=c(5.1,4.1,4.1,3.1))
  plot(x = simlist$x.axis, y = ll[,"dt"], type = "l", col = pal["tree"], ylim = ylim,
       xlab = "", ylab = "", xaxt="n", yaxt="n", lwd = 2)
  title(main = "disttree", col.main = pal["tree"], cex.main = 1.2, font.main = 2, adj = 0.19)
  title(main = "distforest", col.main = pal["forest"], cex.main = 1.2, font.main = 2, adj = 0.5)
  title(main = "gamlss", col.main = pal["gamlss"], cex.main = 1.2, font.main = 2, adj = 0.8)
  title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.35)
  title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.65)

  lines(x = simlist$x.axis, y = ll[,"df"], type = "l", col = pal["forest"], lwd = 2)
  lines(x = simlist$x.axis, y = ll[,"g"], type = "l", col = pal["gamlss"], lwd = 2)
  #legend('topright', c("disttree", "distforest", "gamlss"), col = c(pal["tree"], pal["forest"],pal["gamlss"]), lty = 1)
  #mtext(text= "disttree", side = 3, col = pal["tree"], line = 1, adj = 0.3)
  #mtext(text= "distforest", side = 3, col = pal["forest"], line = 1, adj = 0.5)
  #mtext(text= "gamlss", side = 3, col = pal["gamlss"], line = 1, adj = 0.7)
  #mtext(text= "vs.", side = 3, col = "black", line = 1, adj = 0.4)
  #mtext(text= "vs.", side = 3, col = "black", line = 1, adj = 0.6)
  mtext(text= "smooth", side = 1, line = 1, adj = 0)
  mtext(text= "steep", side = 1, line = 1, adj = 1)
  mtext(text= TeX('$\\kappa$'), side = 1, line = 1)
  mtext(text= "Log-Likelihood", side = 2, line = 1)
  mtext(text= "disttree", side = 2, col = pal["tree"], las = 1, line = 0.2, padj = 16.7)
  mtext(text= "distforest", side = 2, col = pal["forest"], las = 1, line = 0.2, padj = -8)
  mtext(text= "gamlss", side = 2, col = pal["gamlss"], las = 1, line = 0.3, padj = -16)

  # add arrows
  parsave <- par(new = TRUE,  mar = c(0,0,0,0))
  plot(0,0,xlim=c(0,1),ylim=c(0,1),type="n",xlab='', ylab='', col='white', axes = FALSE)

  #add arrow on the right side (upwards, for LL)
  arrows(0.99,0.53,0.99,0.7,lwd=1) ## add arrow
  text(0.99,0.5, 'better')
  segments(0.99,0.3,0.99,0.47,lwd=1)

  #add arrows below
  arrows(0.45,0.085,0.3,0.085,lwd=1) ## add arrow
  arrows(0.6,0.085,0.75,0.085,lwd=1) ## add arrow

  # plot parameters  reset to prior values
  par(parsave)
}
@

\begin{frame}[fragile]
\frametitle{Simulation}
\begin{center}
\vspace*{-1cm}
\setkeys{Gin}{width=0.8\linewidth}
<<plotsim_ll, fig=TRUE, echo=FALSE>>=
plot_ll(simlist)
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.4cm}
\textbf{Goal:} Predict total precipitation amount in alpine regions.\\
\medskip
\textbf{Input:} Numerical Weather Prediction (NWP):\\ 
Global physical model numerically simulating atmospheric processes on $50\times 50 \text{km}^2$ grid.\\
\medskip
\textbf{Problem:} Uncertain initial conditions, unresolved processes.\\ %$\rightarrow$ requires post-processing.\\
\vspace{1.3cm}
%\bigskip
\textbf{Distribution Assumption:}
%\vspace{0.1cm}
%\hspace{1cm}
$(\text{Total precipitation})^\frac{1}{1.6} \sim \textit{c}\mathcal{N}(\mu,\sigma^2)$\\
\bigskip
\textbf{Idea:} Statistical model with numerical predictions as covariates.
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.4cm}
\textbf{Observation data} (National Hydrographical Service):\\
\begin{itemize}
\item 95 observation stations in Tyrol, Austria.\\
\item Daily 24h precipitation sums from July of 28 years (1985--2012).
\end{itemize}
\vspace{1cm}
%\bigskip
\textbf{NWP model output} (Global Ensemble Forecast System):\\
\begin{itemize}
\item Numerical ensemble predictions of various meteorological quantities.\\
\item 11 members each using slightly different perturbed initial conditions.
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
\textbf{Base Variables:}
\begin{itemize}
\item Total precipitation
\item Convective available potential energy
\item Downwards short wave radiation flux (``sunshine'')
\item Mean sea level pressure
\item Preciptable water
\item 2m maximum temperature
\item Total column-integrated condensate
\item Temperature 
\item Temperature differences in altitude
\end{itemize}
\medskip
\textbf{Variations:} Ensemble min/max/mean/standard deviation\\
\vspace{0.1cm}
$\rightarrow$ 82 covariates.

\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
Application for one station:
\begin{itemize}
\item Learn forest model on data from 24 years (1985--2008).
\item Evaluate on 4 years (2009--2012).
\end{itemize}
\end{frame}

<<pred_axams, eval=TRUE, echo=FALSE, results=hide>>=
#setwd("~/svn/partykit/pkg/disttree/inst/CMStatistics2017")
load("data_Axams.rda")

# tree and forest formula
{
  dt.formula <- df.formula <- 
      robs ~ tppow_mean + tppow_sprd + tppow_min + tppow_max + 
      tppow_mean0612 + tppow_mean1218 + tppow_mean1824 + tppow_mean2430 + 
      tppow_sprd0612 + tppow_sprd1218 + tppow_sprd1824 + tppow_sprd2430 + 
      capepow_mean + capepow_sprd + capepow_min + capepow_max + 
      capepow_mean0612 + capepow_mean1218 + capepow_mean1224 + capepow_mean1230 +
      capepow_sprd0612 + capepow_sprd1218 + capepow_sprd1224 + capepow_sprd1230 +
      dswrf_mean_mean + dswrf_mean_max +  
      dswrf_sprd_mean + dswrf_sprd_max + 
      msl_mean_mean + msl_mean_min + msl_mean_max + 
      msl_sprd_mean + msl_sprd_min + msl_sprd_max +
      pwat_mean_mean + pwat_mean_min + pwat_mean_max + 
      pwat_sprd_mean + pwat_sprd_min + pwat_sprd_max +
      tmax_mean_mean + tmax_mean_min + tmax_mean_max +
      tmax_sprd_mean + tmax_sprd_min + tmax_sprd_max +
      tcolc_mean_mean + tcolc_mean_min + tcolc_mean_max +
      tcolc_sprd_mean + tcolc_sprd_min + tcolc_sprd_max +
      t500_mean_mean + t500_mean_min + t500_mean_max +
      t700_mean_mean + t700_mean_min + t700_mean_max +
      t850_mean_mean + t850_mean_min + t850_mean_max +
      t500_sprd_mean + t500_sprd_min + t500_sprd_max +
      t700_sprd_mean + t700_sprd_min + t700_sprd_max +
      t850_sprd_mean + t850_sprd_min + t850_sprd_max +
      tdiff500850_mean + tdiff500850_min + tdiff500850_max +
      tdiff700850_mean + tdiff700850_min + tdiff700850_max +
      tdiff500700_mean + tdiff500700_min + tdiff500700_max +
      msl_diff

  }

learndata <- data_axams[data_axams$year < 109,]
testdata <- data_axams[data_axams$year %in% c(109, 110, 111, 112),]

##############################################################
# fitting the models
set.seed(7)

df <- distforest(df.formula, data = learndata, family = dist_list_cens_normal, 
                 type.tree = "ctree", ntree = 100, censtype = "left", censpoint = 0,
                 control = ctree_control(teststat = "quad", testtype = "Univ",
                                         intersplit = TRUE,
                                         mincriterion = 0, minsplit = 50,
                                         minbucket = 20), mtry = 27)

#### prepare data for plot of estimated density functions
# predictions for one day (in each of the four years) 
# (19th of July 2011 is missing)
pday <- 24  # 2 (hohe Beobachtung zu niedrig geschaetzt), 4, 15, evtl. auch 7, 8, 23 (eine 0-Beobachtung und 2 sehr aehnliche),

pdays <- if(pday<19) c(pday, pday + 31, pday + 62, pday + 92) else c(pday, pday + 31, pday + 61, pday + 92)
pdf <- predict(df, newdata = testdata[pdays,], type = "parameter")
df_mu <- pdf$mu
df_sigma <- pdf$sigma
df_exp <- pnorm(df_mu/df_sigma) * (df_mu + df_sigma * (dnorm(df_mu/df_sigma) / pnorm(df_mu/df_sigma)))
cbind(df_exp, testdata[pdays,"robs"])

# plot predicted distributions together with observations
#set.seed(res$call$seedconst)
set.seed(7)
x <- c(0.01, sort(runif(500,0.01,8)))
y1 <- crch::dcnorm(x, mean = df_mu[1], sd = df_sigma[1], left = 0)
y2 <- crch::dcnorm(x, mean = df_mu[2], sd = df_sigma[2], left = 0)
y3 <- crch::dcnorm(x, mean = df_mu[3], sd = df_sigma[3], left = 0)
y4 <- crch::dcnorm(x, mean = df_mu[4], sd = df_sigma[4], left = 0)
dayending <- if(pday > 3) "th" else switch(pday, "1" = {dayending <- "st"}, "2" = {dayending <- "nd"}, "3" = {dayending <- "rd"})

# point mass (slightly shifted)
pm1 <- c(0.05, crch::dcnorm(-1, mean = df_mu[1], sd = df_sigma[1], left = 0))
pm2 <- c(0.01, crch::dcnorm(-1, mean = df_mu[2], sd = df_sigma[2], left = 0))
pm3 <- c(-0.03, crch::dcnorm(-1, mean = df_mu[3], sd = df_sigma[3], left = 0))
pm4 <- c(-0.07, crch::dcnorm(-1, mean = df_mu[4], sd = df_sigma[4], left = 0))

# predictions
pred1 <- c(testdata[pdays,"robs"][1], crch::dcnorm(testdata[pdays,"robs"][1], mean = df_mu[1], sd = df_sigma[1], left = 0))
pred2 <- c(testdata[pdays,"robs"][2], crch::dcnorm(testdata[pdays,"robs"][2], mean = df_mu[2], sd = df_sigma[2], left = 0))
pred3 <- c(testdata[pdays,"robs"][3], crch::dcnorm(testdata[pdays,"robs"][3], mean = df_mu[3], sd = df_sigma[3], left = 0))
pred4 <- c(testdata[pdays,"robs"][4], crch::dcnorm(testdata[pdays,"robs"][4], mean = df_mu[4], sd = df_sigma[4], left = 0))

#legendheight
lh1 <- crch::dcnorm(0.01, mean = df_mu[1], sd = df_sigma[1], left = 0)
lh2 <- crch::dcnorm(0.01, mean = df_mu[2], sd = df_sigma[2], left = 0)
lh3 <- crch::dcnorm(0.01, mean = df_mu[3], sd = df_sigma[3], left = 0)
lh4 <- crch::dcnorm(0.01, mean = df_mu[4], sd = df_sigma[4], left = 0)
@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\begin{center}
\vspace*{-0.2cm}
<<plot_pred_axams_24July, fig=TRUE, echo=FALSE, eval=TRUE, height=5.2>>=
plot(x = x, y = y1, type = "l", col = "red", 
     main = paste0("Axams, July ", pday), ylab = "Density", 
     xlab = expression(Total~precipitation~"["~mm^(1/1.6)~"/"~"24h"~"]"),
     #xlab = expression(Total~precipitation~group("[", mm^(1/1.6)~per~day,"]")),
     #{\frac{1}{1.6}}$,
     ylim = c(0,max(y1, y2, y3, y4, pm1, pm2, pm3, pm4) + 0.01),
     xlim = c(-1.5,8))

lines(x = x, y = y2, type = "l", col = "blue")
lines(x = x, y = y3, type = "l", col = "darkgreen")
lines(x = x, y = y4, type = "l", col = "purple")
legend('topright', c("predicted distribution", "point mass at censoring point", "observation"),
       bty = "n", col = "black", lty = c(1, NA, NA), pch = c(NA, 19, 4), cex = 0.8)

# plot point mass
lines(x = c(pm1[1], pm1[1]), y = c(pm1[2], 0), col = "red", type = "l", lwd = 1)
lines(x = c(pm2[1], pm2[1]), y = c(pm2[2], 0), col = "blue", type = "l", lwd = 1)
lines(x = c(pm3[1], pm3[1]), y = c(pm3[2], 0), col = "darkgreen", type = "l", lwd = 1)
lines(x = c(pm4[1], pm4[1]), y = c(pm4[2], 0), col = "purple", type = "l", lwd = 1)

points(x = pm1[1], y = pm1[2], col = "red", pch = 19)
points(x = pm2[1], y = pm2[2], col = "blue", pch = 19)
points(x = pm3[1], y = pm3[2], col = "darkgreen", pch = 19)
points(x = pm4[1], y = pm4[2], col = "purple", pch = 19)

# plot predictions
points(x = pred1[1], y = pred1[2], col = "red", pch = 4)
points(x = pred2[1], y = pred2[2], col = "blue", pch = 4)
points(x = pred3[1], y = pred3[2], col = "darkgreen", pch = 4)
points(x = pred4[1], y = pred4[2], col = "purple", pch = 4)

lines(x = c(pred1[1], pred1[1]), y = c(pred1[2], 0), col = "darkgrey", type = "l", lty = 2)
lines(x = c(pred2[1], pred2[1]), y = c(pred2[2], 0), col = "darkgrey", type = "l", lty = 2)
lines(x = c(pred3[1], pred3[1]), y = c(pred3[2], 0), col = "darkgrey", type = "l", lty = 2)
lines(x = c(pred4[1], pred4[1]), y = c(pred4[2], 0), col = "darkgrey", type = "l", lty = 2)

# add labels
text(x = -0.8, y = lh1, labels = "2009", col = "red", cex = 0.8)
text(x = -0.8, y = lh2, labels = "2010", col = "blue", cex = 0.8)
text(x = -0.8, y = lh3, labels = "2011", col = "darkgreen", cex = 0.8)
text(x = -0.8, y = lh4, labels = "2012", col = "purple", cex = 0.8)
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.2cm}
Application for one station:
\begin{itemize}
\item Learn forest model on data from 24 years.
\item Evaluate on 4 years.
\item 10 times 7-fold cross validation.
\item Compare to commonly used heteroscedastic censored Gaussian models:
\end{itemize}
%\begin{itemize}
%\item Prespecified GAM: variable selection based on meteorological expert knowledge.
%\item Boosted GAM: automatic variable selection.
%\item Ensemble Model Output Statistics (EMOS): linear predictors using only total precipitation predictions as covariates.
%\end{itemize}
\vspace{0.1cm}
\begin{table}[t!]
\footnotesize
\nodindent
\hskip-0.4cm\begin{tabular}{ l l l }
\hline
Model & Type & Variable Selection \\
\hline
\vspace{-0.15cm}
\textbf{Forest} & recursive    & automatic \\ 
                & partitioning &           \\ 
\hline
\vspace{-0.15cm}
\textbf{prespecified GAM} & spline  & based on expert \\
                          & in each & knowledge      \\
\hline
\vspace{-0.15cm}
\textbf{boosted GAM} & spline  & automatic  \\
                     & in each &            \\
\hline
\vspace{-0.15cm}
\textbf{EMOS} & linear & total precipitation mean \\
              &        & and standard deviation   \\
\hline
\end{tabular}
\end{table}

\end{frame}

<<echo=FALSE, results=hide>>=
#### cross validation rain
if(file.exists("rain_Axams_7x10.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/rain_Axams_7x10.rda")
  load("rain_Axams_7x10.rda")
} else {
  #source("~/svn/partykit/pkg/disttree/inst/draft/rain_cross_7x10.R")
  source("rain_cross_7x10.R")
  library("gamlss.cens")
  gen.cens("NO", type = "left")
  rainres <- rain_cross(stationname = "Axams", 
                        seedconst = 7, ntree = 100,
                        forest_mtry = 27,
                        gamboost_cvr = TRUE,
                        frac = FALSE)

  save(rainres, file = "rain_Axams_7x10.rda")
}

rain_crps <- matrix(0, ncol = ncol(rainres[[1]]$crps), length(rainres)-1)

for(i in 1:(length(rainres)-1)){
  rain_crps[i,] <- colMeans(rainres[[i]]$crps, na.rm = TRUE)
}

colnames(rain_crps) <- colnames(rainres[[1]]$rmse)

@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\begin{center}
\vspace*{-0.2cm}
<<rain_cross_axams_crps_skills_score, fig=TRUE, echo=FALSE, height=5.2>>=
boxplot(1 - rain_crps[,c(2,3,4)] / rain_crps[,6], ylim = c(-0.005, 0.065),
        main = "Cross validation (with reference model EMOS)",
        names = c("Forest", "prespecified GAM", "boosted GAM"),
        ylab = "CRPS skills score", col = "lightgray") 
abline(h = 0, col = pal["EMOS"], lwd = 2)
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.2cm}
Application for all 95 stations:
\begin{itemize}
\item Learn forest model on data from 24 years (1985--2008).
\item Evaluate on 4 years (2009--2012).
\item Compare to commonly used heteroscedastic censored Gaussian models:\\
\end{itemize}
%\begin{itemize}
%\item Prespecified GAM: variable selection based on meteorological expert knowledge.
%\item Boosted GAM: automatic variable selection.
%\item Ensemble Model Output Statistics (EMOS): linear predictors using only total precipitation predictions as covariates.
%\end{itemize}
\vspace{0.1cm}
\begin{table}[t!]
\footnotesize
\nodindent
\hskip-0.4cm\begin{tabular}{ l l l }
\hline
Model & Type & Variable Selection \\
\hline
\vspace{-0.15cm}
\textbf{Forest} & recursive    & automatic \\ 
                & partitioning &           \\ 
\hline
\vspace{-0.15cm}
\textbf{prespecified GAM} & spline  & based on expert \\
                          & in each & knowledge      \\
\hline
\vspace{-0.15cm}
\textbf{boosted GAM} & spline  & automatic  \\
                     & in each &            \\
\hline
\vspace{-0.15cm}
\textbf{EMOS} & linear & total precipitation mean \\
              &        & and standard deviation   \\
\hline
\end{tabular}
\end{table}
\end{frame}

<<echo = FALSE>>=
#### prediction over all stations 24 - 4
if(file.exists("rain_pred_24to4.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/rain_pred_24to4.rda")
  load("rain_pred_24to4.rda")
} else {

  source("rain_pred_24to4.R")
  library("gamlss.cens")
  gen.cens("NO", type = "left")

  res <- rain_pred(seedconst = 7,
                   gamboost_cvr = TRUE,
                   frac = FALSE)

  save(res, file = "rain_pred_24to4.rda")
}

crps_all <- res[[1]]$results["crps",]
for(i in 2:(length(res)-1))  crps_all <- rbind(crps_all, res[[i]]$results["crps",])

colnames(crps_all) <- colnames(res[[1]]$results)

# skills score
s <- 1 - crps_all[, 2:4]/crps_all[,6]
colnames(s) <- c("Forest", "prespecified GAM", "boosted GAM")

## prepare data for map which shows where distforest performed better than gamlss or gamboostLSS based on the crps

crps <- crps_all[,c("distforest", "gamlss", "gamboostLSS", "EMOS log")]

## best method
bst <- apply(crps, 1, which.min)

## distance of forest to best other method
dst <- crps[,1] - crps[cbind(1:nrow(crps), apply(crps[, -1], 1, which.min) + 1)]

## breaks/groups
brk <- c(-0.1, -0.05, -0.005, 0.005, 0.05, 0.1)
#brk <- c(-0.1, -0.05, -0.01, 0.01, 0.05, 0.1)
grp <- cut(dst, breaks = brk)

## HCL colors (relatively flashy, essentially CARTO Tropic)
clr <- colorspace::diverge_hcl(5, h = c(195, 325), c = 80, l = c(50, 90), power = 1.3)

library("raster") # dem (digital elevation model)
library("sp")     # gadm www.gadm.org/country

load("../draft/plot_map_rain/demo/tirol.gadm.rda")
load("../draft/plot_map_rain/demo/tirol.dem.rda")
load("../draft/plot_map_rain/demo/ehyd.statlist.rda")

# Create SpatialPointsDataFrame from station list
sp <- SpatialPointsDataFrame(subset(ehyd.statlist[res$complete_stations,],
                                    select=c(lon,lat)),
                             data = subset(ehyd.statlist[res$complete_stations,],
                                           select = -c(lon,lat)),
                             proj4string = crs(tirol.dem))

@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\begin{center}
\vspace*{0cm}

<<map, fig=TRUE, echo=FALSE, width=10, height = 6.5>>=

layout(cbind(1, 2), width = c(9, 1))
par(mar = c(5,4,4,0.1))
raster::image(tirol.dem, col = rev(gray.colors(100)),
              main="Observation Stations", ylab = "Latitude", xlab = "Longitude", 
              xlim = c(9.8,13.2), 
              ylim = c(46.6, 47.8), 
              cex.lab=1.5, cex.axis=1.2, cex.main=2)
plot(tirol.gadm, add = TRUE)
points(sp, pch = c(21, 24, 25, 22)[bst], bg = clr[grp], col = "black", las = 1, cex = 1.9)
legend("topleft", pch=c(21,24,25,22), legend = c("Forest", "prespecified GAM", "boosted GAM", "EMOS"), cex = 1.2, bty = "n")

par(mar = c(0.5,0.2,0.5,2.3))
## legend
plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
     xlim = c(0, 1), ylim = c(-0.2, 0.2), xaxs = "i", yaxs = "i")
rect(0, brk[-6], 0.5, brk[-1], col = rev(clr))
axis(4, at = brk, las = 1, mgp=c(0,0,-1.2), cex = 1.4)

@
\end{center}
\end{frame}

%\begin{frame}[fragile]
%\frametitle{Simulation}
%\begin{center}
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{sim_ll}
%\end{center}
%\end{frame}

%\begin{frame}[fragile]
%\frametitle{Simulation}
%\begin{center}
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{sim_RMSE_par}
%\end{center}
%\end{frame}

%\begin{frame}[fragile]
%\frametitle{Simulation}
%\begin{center}
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{sim_RMSE_exp}
%\end{center}
%\end{frame}

\begin{frame}
\frametitle{Software}
\vspace{0.4cm}
\textbf{Package:} \emph{disttree} available on R-Forge at\\

\medskip

\url{https://R-Forge.R-project.org/projects/partykit/}\\

\bigskip
\bigskip

\textbf{Main functions:}

\medskip

\begin{tabular}{ll}
\code{distfit}    & Distributional fits (ML, \code{gamlss.family}/custom \code{list}).\\
                  & No covariates. \\
\code{disttree}   & Distributional trees (\code{ctree}/\code{mob} + \code{distfit}).\\
                  & Covariates as partitioning variables. \\
\code{distforest} & Distributional forests (ensemble of \code{disttree}s).\\
                  & Covariates as partitioning variables.
\end{tabular}

\end{frame}

\subsection{References}

\begin{frame}
\frametitle{References}

\vspace{-0.2cm}

\footnotesize
Hothorn T, Hornik K, Zeileis A (2006).
 \dquote{Unbiased Recursive Partitioning: A Conditional Inference Framework.}
 \textit{Journal of Computational and Graphical Statistics},
 \textbf{15}(3), 651--674.
 \doi{10.1198/106186006X133933}

\medskip

Zeileis A, Hothorn T, Hornik K (2008).
 \dquote{Model-Based Recursive Partitioning.}
  \textit{Journal of Computational and Graphical Statistics},
  \textbf{17}(2), 492--514.
  \doi{10.1198/106186008X319331}

\medskip

Hothorn T, Zeileis A (2015).
 \dquote{partykit: A Modular Toolkit for Recursive Partytioning in R.}
 \textit{Journal of Machine Learning Research},
 \textbf{16}, 3905--3909.
 \url{http://www.jmlr.org/papers/v16/hothorn15a.html}

\medskip

Stasinopoulos DM, Rigby RA (2007).
  \dquote{Generalized Additive Models for Location Scale and Shape (GAMLSS) in R.}
  \textit{Journal of Statistical Software}, 
  \textbf{23}(7), 1--46.
  \doi{10.18637/jss.v023.i07}

%\medskip

%Seibold H, Zeileis A, Hothorn T (2017).
%  \dquote{Individual Treatment Effect Prediction for Amyotrophic Lateral Sclerosis Patients.}
%  \textit{Statistical Methods in Medical Research}, 
%  \textbf{12}(1), 45--63.
%  \doi{10.1177/0962280217693034}

%\medskip

%Hothorn T, Zeileis A (2017).
%  \dquote{Transformation Forests.}
%  \emph{arXiv 1701.02110}, arXiv.org E-Print Archive.
%  \url{http://arxiv.org/abs/1701.02110}

\medskip

Gneiting T, Raftery A E, Westveld III A H, Goldman T (2005).
  \dquote{Calibrated probabilistic forecasting using ensemble model output statistics and minimum CRPS estimation}
  \textit{Monthly Weather Review}, 
  \textbf{133}(5), 1098--1118.
  \doi{10.1175/MWR2904.1}

\medskip

Hofner B, Mayr A, Fenske N, Schmid M (2017).
  \dquote{{gamboostLSS}: Boosting Methods for {GAMLSS} Models}
  \textit{{R} package version 2.0-0},
  \url{https://CRAN.R-project.org/package=gamboostLSS}

\end{frame}

\begin{frame}
\frametitle{References (2)}

\vspace{-0.2cm}

\footnotesize

% Reference for the observation data set
BMLFUW (2016).
  \dquote{{B}undesministerium f\"ur {L}and und {F}orstwirtschaft, {U}mwelt und {W}asserwirtschaft ({BMLFUW}), {A}bteilung {IV/4} - {W}asserhaushalt}
  \text{Available at \url{http://ehyd.gv.at}}, 
  \text{Accessed: {F}ebruary 29 2016}.

\medskip

% Reference for the GFS reforecast version 2 data set 
Hamill T M, Bates G T, Whitaker J S, Murray D R, Fiorino M, Galarneau T J Jr., Zhu Y, Lapenta W (2013).
  \dquote{{NOAA}'s Second-Generation Global Medium-Range Ensemble Reforecast Dataset}
  \textit{Bulletin of the American Meteorological Society}, 
  \textbf{94}(10), 1553--1565.
  \doi{10.1175/BAMS-D-12-00014.1}

\medskip

%Breiman L (2001).
%  \dquote{Random {Forests}}
%  \textit{Machine Learning}, 
%  \textbf{45}(1), 5--32.
%  \doi{10.1023/A:1010933404324}

Messner J W,  Mayr G J, Zeileis A (2016).
  \dquote{Heteroscedastic Censored and Truncated Regression with {crch}}
  \textit{The R Journal}, 
  \textbf{8}(1), 173--181.\\
  \scriptsize
  \url{https://journal.R-project.org/archive/2016-1/messner-mayr-zeileis.pdf}
  \footnotesize

\medskip

Stauffer R, Mayr G J, Messner J W, Umlauf N, Zeileis A (2017).
  \dquote{Spatio-Temporal Precipitation Climatology Over Complex Terrain Using a Censored Additive Regression Model}
  \textit{International Journal of Climatology}, 
  \textbf{37}(7), 3264--3275.
  \doi{10.1002/joc.4913}

\medskip

Gebetsberger M, Messner J W, Mayr G J, Zeileis A (2017).
  \dquote{Fine-Tuning Non-Homogeneous Regression for Probabilistic Precipitation Forecasts: Unanimous Predictions, Heavy Tails, and Link Functions}
  \textit{Monthly Weather Review}, 
  \textbf{Forthcoming}.
  \doi{10.1175/MWR-D-16-0388.1}

%\medskip

%{CGIAR-CSI} (2016).
%  \dquote{{SRTM} 90m Digital Elevation Database {v4.1}}
%  \text{Available at \url{http://srtm.csi.cgiar.org}}, 
%  \text{Accessed: {J}uly 10 2016}.

\end{frame}

\end{document}