### 2017-12-22 pkg/model4you/inst/reproducing_papers/2017_individual_treatment_effect_als_smmr.Rmd

---
title: model4you reproducing "Individual treatment effect prediction for amyotrophic
  lateral sclerosis patients"
author: "Heidi Seibold"
date: "December 22, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

```{r}
library("model4you")
library("survival")
library("ggplot2")
load("data/ALSFRSdata.rda")
load("data/ALSsurvdata.rda")
```

## Base models
```{r}
bm_alsfrs <- glm(ALSFRS.Start ~ Riluzole + offset(log(ALSFRS.halfYearAfter)), 
                 data = ALSFRSdata, family = gaussian(link = "log"))
summary(bm_alsfrs)

bm_alssurv <- survreg(Surv(survival.time, cens) ~ Riluzole, 
                      data = ALSsurvdata, dist = "weibull")
summary(bm_alssurv)
```
## Forests
```{r}
set.seed(123)

frst_alsfrs <- pmforest(bm_alsfrs, ntree = 100, 
                        control = ctree_control(teststat = "max", testtype = "Univ",
                                                mincriterion = 0.95, minsplit = 40, 
                                                minbucket = 30, lookahead = TRUE))

frst_alssurv <- pmforest(bm_alssurv, ntree = 100, 
                         control = ctree_control(teststat = "max", testtype = "Univ",
                                                 mincriterion = 0.95, minsplit = 40, 
                                                 minbucket = 30, lookahead = TRUE))
```

## Personalised models
```{r}
pm_alsfrs <- pmodel(frst_alsfrs)

coeffun <- function(model) {
  ## model coefficients
  coefs <- c(coef(model), scale = model$scale)

  ## difference in median survival 
  p = 0.5
  coefs["median_s0"] <- qweibull(p = p, shape = 1/coefs["scale"], 
                                 scale = exp(coefs["(Intercept)"]))
  coefs["median_s1"] <- qweibull(p = p, shape = 1/coefs["scale"], 
                                 scale = exp(coefs["(Intercept)"] + coefs["aA"]))
  coefs["median_sdiff"] <- coefs["median_s1"] - coefs["median_s0"]

  return(coefs)
}
pm_alssurv <- pmodel(frst_alssurv, fun = coeffun)
```

## Variable importance
```{r}
vi_alsfrs <- varimp(frst_alsfrs, 
                    risk = function(x, ...) -objfun(x, sum = TRUE, ...), 
                    nperm = 1)
var <- names(vi_alsfrs)
vi_alsfrs_d <- data.frame(variable = factor(var, levels = var[order(vi_alsfrs)]), 
                          VI = vi_alsfrs)

ggplot(vi_alsfrs_d, aes(x = variable, y = VI)) + 
  geom_bar(stat = "identity", width = .1) +
  coord_flip() +
  theme(panel.grid.major.y = element_blank())

vi_alssurv <- varimp(frst_alssurv, 
                     risk = function(x, ...) -objfun(x, sum = TRUE, ...), 
                     nperm = 1)
var <- names(vi_alssurv)
vi_alssurv_d <- data.frame(variable = factor(var, levels = var[order(vi_alssurv)]), 
                          VI = vi_alssurv)

ggplot(vi_alssurv_d, aes(x = variable, y = VI)) + 
  geom_bar(stat = "identity", width = .1) +
  coord_flip() +
  theme(panel.grid.major.y = element_blank())
```

## dependence plot
```{r}
dp_alsfrs <- cbind(pm_alsfrs, ALSFRSdata)

dp_plot_alsfrs <- function(var) {
  print(ggplot(dp_alsfrs, aes_string(y = "RiluzoleYes", x = var)) +
    geom_point() + geom_smooth(se = FALSE))
}
dp_plot_alsfrs("time_onset_treatment")
dp_plot_alsfrs("value_creatinine")
dp_plot_alsfrs("value_phosphorus")
dp_plot_alsfrs("subjectliters_fvc")
```