### 2018-02-26 pkg/partykit/inst/guideliketest/results.Rnw
comparing results of different test statistics
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2039 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 86884d7441449eb5bfc473319355ac3b9648023f

\documentclass{article}

<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
library("partykit")
library("Formula")
library("parallel")
library("lattice")
source("guidelike.R")
load("sim/simres20180224_1step.rda")
@

\title{Comparison of different testing functions}

\begin{document}

\section{All: 1 step}
\setkeys{Gin}{width=\linewidth}
<<all_TTF1, fig=TRUE, echo=TRUE, height=10, width=8>>=
simres$test <- factor(simres$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = simres, 
       subset = binary_beta == TRUE & binary_regressor == TRUE & only_intercept == FALSE, 
       type = "b", auto.key = FALSE, ylim = c(0,1))
@

\newpage  
<<all_TFF1, fig=TRUE, echo=TRUE, height=10, width=8>>=
simres$test <- factor(simres$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = simres, 
       subset = binary_beta == TRUE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = FALSE, ylim = c(0,1))
@

\newpage
\section{Influence of binarization of scores and/or categorization of split variables: 1 step}

\subsection{ctree}
<<ctree_cat_bin_TFF1, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("ctree", "ctree_cat", "ctree_bin", "ctree_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata, 
       subset = binary_beta == TRUE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage  
<<ctree_cat_bin_FFF1, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("ctree", "ctree_cat", "ctree_bin", "ctree_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata, 
       subset = binary_beta == FALSE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage
\subsection{mfluc}

<<mfluc_cat_bin_TFF1, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("mfluc", "mfluc_cat", "mfluc_bin", "mfluc_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == TRUE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage  
<<mfluc_cat_bin_FFF1, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("mfluc", "mfluc_cat", "mfluc_bin", "mfluc_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == FALSE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage
\section{GUIDE vs ctree and mfluc with binarized scores and categorized splitvariables: 1 step}
<<guide_vs_cat_bin_TFF1, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("mfluc_cat_bin", "ctree_cat_bin", "ctree_max_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == TRUE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage  
<<guide_vs_cat_bin_FFF1, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("mfluc_cat_bin", "ctree_cat_bin", "ctree_max_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == FALSE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage
\section{GUIDE versions: 1 step}

<<guide_versions_TTF1, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("ctree_cat_bin", "ctree_max_cat_bin", "guide_coin_12", "guide_sum_12", "guide_sum_1_cor"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == TRUE & binary_regressor == TRUE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage  
<<guide_versions_TFF1, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("ctree_cat_bin", "ctree_max_cat_bin", "guide_coin_12", "guide_sum_12", "guide_sum_1_cor"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == TRUE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage  
<<guide_versions_FTF1, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("ctree_cat_bin", "ctree_max_cat_bin", "guide_coin_12", "guide_sum_12", "guide_sum_1_cor"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == FALSE & binary_regressor == TRUE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage  
<<guide_versions_FFF1, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("ctree_cat_bin", "ctree_max_cat_bin", "guide_coin_12", "guide_sum_12", "guide_sum_1_cor"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == FALSE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage
\section{check alpha: 1 step}
\subsection{ctree}
<<check_alpha_ctree1, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subdata <- subset(simres, test %in% c("ctree", "ctree_cat", "ctree_bin", "ctree_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == TRUE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,0.1))
@

\newpage  
\subsection{mfluc}
<<check_alpha_mfluc1, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subdata <- subset(simres, test %in% c("mfluc", "mfluc_cat", "mfluc_bin", "mfluc_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == TRUE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,0.1))
@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%load data set for 2 step function
<<>>=
load("sim/simres20180224_2step.rda")
@

\newpage
\section{All: 2 steps}
<<all_TTF2, fig=TRUE, echo=TRUE, height=10, width=8>>=
simres$test <- factor(simres$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = simres, 
       subset = binary_beta == TRUE & binary_regressor == TRUE & only_intercept == FALSE, 
       type = "b", auto.key = FALSE, ylim = c(0,1))
@

\newpage  
<<all_TFF2, fig=TRUE, echo=TRUE, height=10, width=8>>=
simres$test <- factor(simres$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = simres, 
       subset = binary_beta == TRUE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = FALSE, ylim = c(0,1))
@

\newpage  
<<all_FTF2, fig=TRUE, echo=TRUE, height=10, width=8>>=
simres$test <- factor(simres$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = simres, 
       subset = binary_beta == FALSE & binary_regressor == TRUE & only_intercept == FALSE, 
       type = "b", auto.key = FALSE, ylim = c(0,1))
@

\newpage  
<<all_FFF2, fig=TRUE, echo=TRUE, height=10, width=8>>=
simres$test <- factor(simres$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = simres, 
       subset = binary_beta == FALSE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = FALSE, ylim = c(0,1))
@

\newpage
\section{Influence of binarization of scores and/or categorization of split variables: 2 steps}

\subsection{ctree}
<<ctree_cat_bin_TFF2, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("ctree", "ctree_cat", "ctree_bin", "ctree_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata, 
       subset = binary_beta == TRUE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage  
<<ctree_cat_bin_FFF2, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("ctree", "ctree_cat", "ctree_bin", "ctree_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata, 
       subset = binary_beta == FALSE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage
\subsection{mfluc}

<<mfluc_cat_bin_TTF2, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("mfluc", "mfluc_cat", "mfluc_bin", "mfluc_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == TRUE & binary_regressor == TRUE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage  
<<mfluc_cat_bin_TFF2, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("mfluc", "mfluc_cat", "mfluc_bin", "mfluc_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == TRUE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage  
<<mfluc_cat_bin_FTF2, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("mfluc", "mfluc_cat", "mfluc_bin", "mfluc_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == FALSE & binary_regressor == TRUE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage  
<<mfluc_cat_bin_FFF2, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("mfluc", "mfluc_cat", "mfluc_bin", "mfluc_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == FALSE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage
\section{GUIDE vs ctree and mfluc with binarized scores and categorized splitvariables: 2 steps}
<<guide_vs_cat_bin_TFF2, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("mfluc_cat_bin", "ctree_cat_bin", "ctree_max_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == TRUE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage  
<<guide_vs_cat_bin_FFF2, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("mfluc_cat_bin", "ctree_cat_bin", "ctree_max_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == FALSE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage
\section{GUIDE versions}
<<guide_versions_TTF2, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("ctree_cat_bin", "ctree_max_cat_bin", "guide_coin_12", "guide_sum_12", "guide_sum_1_cor"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == TRUE & binary_regressor == TRUE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage  
<<guide_versions_TFF2, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("ctree_cat_bin", "ctree_max_cat_bin", "guide_coin_12", "guide_sum_12", "guide_sum_1_cor"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == TRUE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage  
<<guide_versions_FTF2, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("ctree_cat_bin", "ctree_max_cat_bin", "guide_coin_12", "guide_sum_12", "guide_sum_1_cor"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == FALSE & binary_regressor == TRUE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage  
<<guide_versions_FFF2, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subset(simres, test %in% c("ctree_cat_bin", "ctree_max_cat_bin", "guide_coin_12", "guide_sum_12", "guide_sum_1_cor"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == FALSE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,1))
@

\newpage
\section{check alpha: 2 steps}
\subsection{ctree}
<<check_alpha_ctree2, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subdata <- subset(simres, test %in% c("ctree", "ctree_cat", "ctree_bin", "ctree_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == TRUE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,0.1))
@

\newpage  
\subsection{mfluc}
<<check_alpha_mfluc2, fig=TRUE, echo=TRUE, height=9, width=8>>=
subdata <- subdata <- subset(simres, test %in% c("mfluc", "mfluc_cat", "mfluc_bin", "mfluc_cat_bin", "guide_coin_12"))
subdata$test <- factor(subdata$test)
xyplot(prop_T ~ delta | xi + vary_beta, groups = ~ test, 
       data = subdata,
       subset = binary_beta == TRUE & binary_regressor == FALSE & only_intercept == FALSE, 
       type = "b", auto.key = TRUE, ylim = c(0,0.1))
@
\end{document}