### 2018-04-04 pkg/disttree/inst/MIP-Seminar/slides.Rnw
revised slides
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2125 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: fbe982d0fbe31324ad2cb9e32e15b077fefcc8b5

\documentclass[11pt,t,usepdftitle=false,aspectratio=43]{beamer}
\usetheme[nototalframenumber, noslidenumber]{uibk}

\title{Flexible Probabilistic Forecasting with Distributional Trees and Forests}
\author{Lisa Schlosser}
%\author{Lisa Schlosser, Torsten Hothorn,\\ Heidi Seibold, Reto Stauffer, Achim Zeileis}
\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
%\URL{https://R-Forge.R-project.org/projects/partykit/}
\headerimage{3}

\setbeamertemplate{caption}{\insertcaption} 
%% includes a replacement for \usepackage{Sweave}
\usepackage{changepage}
\usepackage{amsmath,tikz}
\usepackage{calc}
\usepackage{graphicx}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap,trees,tikzmark,decorations.pathreplacing}
\graphicspath{{plots/}}

%% colors
\definecolor{HighlightOrange}{rgb}{0.9490196,0.5725490,0.0000000}
\definecolor{HighlightBlue}{rgb}{0.4784314,0.7490196,0.9803922}

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)

set.seed(7)

library("ggplot2")
theme_set(theme_bw(base_size = 18))
library("disttree")
#library("partykitR1")
library("gamlss")
library("lattice")
library("crch")
library("latex2exp")
library("parallel")
library("RainTyrol")
library("gamlss.cens")
gen.cens(NO, type = "left")

## HCL palette
pal <- hcl(c(10, 128, 260), 100, 50)
names(pal) <- c("forest", "tree", "gamlss")

pallight <- hcl(c(10, 128, 260), 100, 50, alpha = 0.25)
names(pallight) <- c("forest", "tree", "gamlss")

transpgrey <- rgb(0.190,0.190,0.190, alpha = 0.2)

######## data generating process based on a predefined tree or a parameter function
# if tree: 2 splits => 3 terminal nodes
# 10 split variables (x1, ..., x10) are available (5 numeric, 2 binary, 3 categorical)
#
# input: n            ..... nr of observations
#        family       ..... distribution of the generated observations
#        split.matrix ..... indices of the variables used for the splits together with the corresponding split points
#        parm.matrix  ..... set of distribution parameters for each subgroup
#        fun          ..... parameter function
#
# output: data.frame with generated observations y (generated seperatly in each subgroups with given distributions parameters),
#                         the given split variables x1, ..., x10 for each observation and
#                         the index of the subgroup for each observation or
#                         the distribution parameter for each observation

# FIX: until now only complete lists can be handed over (dist_list...)
dgp <- function(n, family = dist_list_normal, 
                fun = NULL,
                split.matrix = matrix(nrow = 2, ncol = 2), 
                par.matrix = matrix(nrow = 3, ncol = 2), 
                round.sp = 3)
{

  # generating the possible split variables
  x1 <- runif(n,-0.4,1)
  x2 <- runif(n,-10,10)
  x3 <- runif(n,0,100)
  x4 <- x1 + rnorm(n, sd = 0.1)
  x5 <- x1 + rnorm(n, sd = 0.3)
  x6 <- rbinom(n,1,0.5)
  x7 <- rbinom(n,1,0.5)
  x8 <- sample(1:4, n, replace = TRUE) 
  x9 <- sample(1:5, n, replace = TRUE) 
  x10 <- sample(1:7, n, replace = TRUE)
  #x11 <- runif(n,-0.5,1)
  x <- cbind(x1,x2,x3,x4,x5,x6,x7,x8,x9,x10)
  # reduce nr of possible split points by rounding values of split variables
  x <- round(x, digits = round.sp)

  y <- vector(mode = "numeric", length = n)

  # getting the random function
  if(is.function(family)) family <- family()
  if(inherits(family, "gamlss.family")) {
    rfun <- get(paste0("r",family$family[[1]]))
  } else {
    if(family$family.name == "Normal Distribution") rfun <- rNO
    if(family$family.name == "censored Normal Distribution") rfun <- rNO
    if(family$family.name == "Poisson Distribution") rfun <- rPO
    if(family$family.name == "Weibull Distribution") rfun <- survival:::rsurvreg 
  }

  if(is.null(fun)) {
    index <- vector(mode = "numeric", length = n)
    sv = split.matrix[,1]    # index of split variables
    sp = split.matrix[,2]    # split points
    par = par.matrix

    # splitting based on the given split variables and split points and
    # generating the observations seperatly for each subgroup (with the corresponding given distribution parameters)
    for(k in 1:n){
      if(x[k, sv[1]] <= sp[1]){
        y[k] <- do.call(rfun, as.list(c(1,par[1,])))
        index[k] <- 2L
      } else {
        if(x[k, sv[2]] <= sp[2]){
          y[k] <- do.call(rfun, as.list(c(1,par[2,]))) 
          index[k] <- 4L
        } else {
          y[k] <- do.call(rfun, as.list(c(1,par[3,])))
          index[k] <- 5L
        }
      }
    }
    d <- as.data.frame(cbind(y, x, index))
    colnames(d) <- c("y", paste0("x", c(1:10)), "index")

  } else {

    dpar <- fun(x)

    if(!(family$family.name == "Poisson Distribution")){
      y <- rfun(n, dpar[,1], dpar[,2])
      d <- as.data.frame(cbind(y, x, dpar))
      colnames(d) <- c("y", paste0("x", c(1:10)), "mu", "sigma")
    } else {
      y <- rfun(n, dpar)
      d <- as.data.frame(cbind(y, x, dpar))
      colnames(d) <- c("y", paste0("x", c(1:10)), "lambda")
    }
  }

  if(family$family.name == "censored Normal Distribution") {
    #d$ystar <- d$y
    d$y <- pmax(d$y, 0)
  }

  return(d)
}

############################################ 
# parameter functions

### plotting function for one data set
plot1 <- function(learndata, dt=NULL, df=NULL, g=NULL, all.par = FALSE, df.par = FALSE, g.par = FALSE, dt.par = FALSE, all.var = FALSE) 
{
  plotdata <- cbind(learndata[,c("y","x1","mu","sigma")],
                    df$fitted.par,
                    dt$fitted.par,
                    #coef(dt)[paste(dt$fitted[,1]),],
                    g$mu.fv, g$sigma.fv)

  colnames(plotdata) <- c("y","x1","true.mu","true.sigma", 
                          "fitted.mu.df","fitted.sigma.df",
                          "fitted.mu.dt","fitted.sigma.dt",
                          "fitted.mu.g","fitted.sigma.g")
  sp <- plotdata[order(plotdata["x1"]),]

  if(!(any(dt.par, df.par, g.par, all.par, all.var))){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "True parameters", cex.main = 1.2, xaxt="n", yaxt="n", xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
  col = transpgrey, border = "transparent")
    #lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", lty = 1, col = 'grey')
    #lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", lty = 1, col = 'grey')
    #legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$')), 
    #       col = c('black', transpgrey), lty = 1, bty = "n", lwd = 2.5)
    legend(x = -0.5, y = 14, expression(mu %+-% sigma), bty = "n")
  }

  # disttree
  if(dt.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "disttree", col.main = pal["tree"], cex.main = 1.2, xaxt="n", yaxt="n", xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    #polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
    #col = transpgrey, border = "transparent")
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.dt, type = "l", col = pal["tree"], lwd = 2.5)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.dt + sp$fitted.sigma.dt, rev(sp$fitted.mu.dt - sp$fitted.sigma.dt)),
  col = pallight["tree"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.dt + sp$fitted.sigma.dt, type = "l", col = 'greenyellow')
    #lines(x = sp$x, y = sp$fitted.mu.dt - sp$fitted.sigma.dt, type = "l", col = 'greenyellow')
    #legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
    #                    TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')), 
    #       col = c('black','grey', pal["tree"], pallight["tree"]), lty = 1, bty = "n", lwd = 2.5)
    legend(x = -0.5, y = 14, expression(mu %+-% sigma), bty = "n")
  }

  # distforest
  if(df.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", main = "distforest", col.main = pal["forest"], cex.main = 1.2, xaxt="n", yaxt="n", xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    #polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
    #col = transpgrey, border = "transparent")
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.df, type = "l", col = pal["forest"], lwd = 2.5)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.df + sp$fitted.sigma.df, rev(sp$fitted.mu.df - sp$fitted.sigma.df)),
  col = pallight["forest"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.df + sp$fitted.sigma.df, type = "l", col = 'orange')
    #lines(x = sp$x, y = sp$fitted.mu.df - sp$fitted.sigma.df, type = "l", col = 'orange')
    #legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
    #                    TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')), 
    #       col = c('black','grey',pal["forest"], pallight["forest"]), lty = 1, bty = "n", lwd = 2.5)
    legend(x = -0.5, y = 14, expression(mu %+-% sigma), bty = "n")
  }

  # gamlss
  if(g.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey",  main = "gamlss", col.main = pal["gamlss"], cex.main = 1.2, xaxt="n", yaxt="n", xlab = "", ylab = "")
    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    #polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
    #col = transpgrey, border = "transparent")
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'grey')
    lines(x = sp$x, y = sp$fitted.mu.g, type = "l", col = pal["gamlss"], lwd = 2.5)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.g + sp$fitted.sigma.g, rev(sp$fitted.mu.g - sp$fitted.sigma.g)),
  col = pallight["gamlss"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.g + sp$fitted.sigma.g, type = "l", col = 'lightblue')
    #lines(x = sp$x, y = sp$fitted.mu.g - sp$fitted.sigma.g, type = "l", col = 'lightblue')
    #legend('topleft', c(TeX('$\\mu$'), TeX('$\\mu \\pm \\sigma$'), 
    #                    TeX('$\\hat{\\mu}$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$')),  
    #       col = c('black','grey',pal["gamlss"], pallight["gamlss"]), lty = 1, bty = "n", lwd = 2.5)
    legend(x = -0.5, y = 14, expression(mu %+-% sigma), bty = "n")
  }

  # disttree vs. distforest vs. gamlss
  if(all.par){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", xaxt="n", yaxt="n", xlab = "", ylab = "")
    title(main = "disttree", col.main = pal["tree"], cex.main = 1.2, font.main = 2, adj = 0.19)
    title(main = "distforest", col.main = pal["forest"], cex.main = 1.2, font.main = 2, adj = 0.5)
    title(main = "gamlss", col.main = pal["gamlss"], cex.main = 1.2, font.main = 2, adj = 0.8)
    title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.35)
    title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.65)

    lines(x = sp$x, y = sp$true.mu, type = "l", col = 'black')
    lines(x = sp$x, y = sp$fitted.mu.dt, type = "l", col = pal["tree"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.df, type = "l", col = pal["forest"], lwd = 2)
    lines(x = sp$x, y = sp$fitted.mu.g, type = "l", col = pal["gamlss"], lwd = 2)
    #legend('topleft', c(TeX('$\\mu$'), TeX('$\\hat{\\mu}$ disttree'),
    #                    TeX('$\\hat{\\mu}$ distforest'), TeX('$\\hat{\\mu}$ gamlss')), 
    #       col = c('black',pal["tree"],pal["forest"],pal["gamlss"]), lty = 1, bty = "n", lwd = 2.5)
    legend(x = -0.5, y = 13.95, expression(mu), bty = "n")
  }

  # variance: disttree vs. distforest vs. gamlss
  if(all.var){
    par(mar=c(5.1,4.1,4.1,3.1))
    plot(y = sp$y, x = sp$x, type = "p", col="grey", xaxt="n", yaxt="n", xlab = "", ylab = "")
    title(main = "disttree", col.main = pal["tree"], cex.main = 1.2, font.main = 2, adj = 0.19)
    title(main = "distforest", col.main = pal["forest"], cex.main = 1.2, font.main = 2, adj = 0.5)
    title(main = "gamlss", col.main = pal["gamlss"], cex.main = 1.2, font.main = 2, adj = 0.8)
    title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.35)
    title(main = "vs.", col.main = "black", cex.main = 1.2, font.main = 2, adj = 0.65)

    #polygon(c(sp$x, rev(sp$x)), c(sp$true.mu + sp$true.sigma, rev(sp$true.mu - sp$true.sigma)),
    #col = transpgrey, border = "transparent")
    lines(x = sp$x, y = sp$true.mu + sp$true.sigma, type = "l", col = 'black')
    lines(x = sp$x, y = sp$true.mu - sp$true.sigma, type = "l", col = 'black')
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.dt + sp$fitted.sigma.dt, rev(sp$fitted.mu.dt - sp$fitted.sigma.dt)),
  col = pallight["tree"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.dt + sp$fitted.sigma.dt, type = "l", col = pal["tree"], lwd = 2)
    #lines(x = sp$x, y = sp$fitted.mu.dt - sp$fitted.sigma.dt, type = "l", col = pal["tree"], lwd = 2)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.df + sp$fitted.sigma.df, rev(sp$fitted.mu.df - sp$fitted.sigma.df)),
  col = pallight["forest"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.df + sp$fitted.sigma.df, type = "l", col = pal["forest"], lwd = 2)
    #lines(x = sp$x, y = sp$fitted.mu.df - sp$fitted.sigma.df, type = "l", col = pal["forest"], lwd = 2)
    polygon(c(sp$x, rev(sp$x)), c(sp$fitted.mu.g + sp$fitted.sigma.g, rev(sp$fitted.mu.g - sp$fitted.sigma.g)),
  col = pallight["gamlss"], border = "transparent")
    #lines(x = sp$x, y = sp$fitted.mu.g + sp$fitted.sigma.g, type = "l", col = pal["gamlss"], lwd = 2)
    #lines(x = sp$x, y = sp$fitted.mu.g - sp$fitted.sigma.g, type = "l", col = pal["gamlss"], lwd = 2)
    #legend('topleft', c(TeX('$\\mu \\pm \\sigma$'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$ disttree'),
    #                    TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$ distforest'), TeX('$\\hat{\\mu} \\pm \\hat{\\sigma}$ gamlss')), 
    #       col = c('black',pal["tree"],pal["forest"],pal["gamlss"]), lty = 1, bty = "n", lwd = 2.5)
  legend(x = -0.5, y = 14, expression(mu %+-% sigma), bty = "n")
  }

}

@

\begin{document}

\section{Distributional Trees and Forests}

\begin{frame}[fragile]
\frametitle{Statistical Modeling}
\visible<2,3, 6-> {Assumption:}
\vspace{-0.7cm}
\begin{center}
\begin{tikzpicture}
\visible<2,3, 6-> {\node[] at (0,0.4) {X};}
\visible<4-5>{
\node[inner sep=0pt] (today) at (-1.8,-0.4)
    {\includegraphics[width=.365\textwidth]{today.png}};
\node[font=\scriptsize] at (-1.9,-2) {15.03.2018};
}
\visible<2->{
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {nature};
\draw[->] (2.6,0.4)--(3.1,0.4);
}
\visible<2,3, 6->{
\node[] at (3.5,0.4) {Y};
}
\visible<4-5>{
\node[inner sep=0pt] (future) at (5.3,-0.4)
    {\includegraphics[width=.365\textwidth]{future.png}};
\node[font=\scriptsize] at (5.4,-2) {16.03.2018};
}
\end{tikzpicture}
\end{center}
\visible<3, 6->{
\vspace{-2cm}
Approximation of reality based on a set of observations.\\
}
\visible<5>{
\vspace{2cm}
X = observed precipitation of today,\\
\qquad current temperature / humidity / \dots,\\
\qquad physical model output\\
\vspace{0.15cm}
Y = daily precipitation sums\\
}

\vspace{-3.5cm}
\visible<7->{
\textbf{Application/Motivation:}
\begin{itemize}
\item Prediction.
\item Caption of dependencies and interactions.
\end{itemize}
}
\vspace{0.5cm}
\visible<8->{
\textbf{Two main approaches / ``cultures'':}
\begin{itemize}
\item Data modeling.
\item Algorithmic modeling.
\end{itemize}
}
\end{frame}

\begin{frame}[fragile]
\frametitle{Data Modeling}
\vspace{-0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {f(X; par)};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\only<1-6>{
\begin{itemize}
\item Approximate data generating process up to a few unknown processes.
\item Mathematical equations, model assumptions.
\item Data: estimation of parameters.
\end{itemize}
}
\only<7-10>{
\vspace{1cm}
Summary:
\begin{itemize}
\item<8-10> Stochastic model to approximate data generating process.\\
\only<9-10>{$\Rightarrow$ fixed $f$.\\}
\item<10> Disadvantage: Risk of incorrect assumptions.
\end{itemize}
}
\only<1>{\vspace{1cm}}

\vspace{0.1cm}

\visible<2-6>{
\begin{minipage}{0.65\textwidth}
\textbf{Example:} 
\only<2-3>{Linear model}\only<4>{Generalized additive model}\only<5-6>{Generalized additive model\\ for location, scale, and shape}
\\
\only<2-3>{
$$\visible<3>{\mathbb{E}(Y|X) = } f(X; \alpha) = \alpha_0 + \alpha_1 \cdot X$$  
\vspace{0.15cm}
}
\only<4>{
$$\mathbb{E}(Y|X) = f(X; \alpha) = \alpha + s(X)$$  
\vspace{0.15cm}
}
\only<5-6>{
\vspace{-1.3cm}

\begin{align*}
\visible<6>{\mathbb{E}(Y|X) = }\mu(X) = \alpha_{\mu} + s_{\mu}(X)\\  
\visible<6>{\sqrt{Var(Y|X)} = }\sigma(X) = \alpha_{\sigma} + s_{\sigma}(X)
\end{align*}
}
\end{minipage}
%\vspace{-0.6cm}
%\hfill
\begin{minipage}{0.3\textwidth}
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<2-3>{
<<plot_motivation_GLM_single, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<4>{
<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\only<5-6>{
<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
}
\end{minipage}
}
\end{frame}

<<raintree, echo=FALSE, results=hide>>=
load("weather.rda")
#d <- data.frame(rain24 = numeric(0), rain = logical(0), wind = numeric(0), 
#                humidity = numeric(0), airpressure = numeric(0), temperature = numeric(0))

sp_rain <- partysplit(2L, index = 1:2)
sp_humidity <- partysplit(4L, breaks = 70)

pn <- partynode(1L, split = sp_rain, kids = list(
  partynode(2L, split = sp_humidity, kids = list(
      partynode(3L, info = "4 mm/24h"),
      partynode(4L, info = "9 mm/24h"))),
  partynode(5L, info = "15 mm/24h")))

p <- party(pn, weather)
@
<<motivation_regtree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
nobs <- 120
x <- c(1:nobs)/nobs
ytrue <- ytree <- ytree2 <- ytree3 <- ytree4 <- 
  ytree_up <- ytree_up2 <- ytree_up3 <- ytree_up4 <- 
  yforest <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<0.55) 4 else 4+5*plogis(kappa*(3*(x[i]-0.36)-1))
y <- ytrue + rnorm(nobs,0,2)
y_up <- rep.int(15, length(x)) + rnorm(nobs,0,1)
for(i in 1:nobs) ytree[i] <- if(x[i]<0.7) 4 else 9
for(i in 1:nobs) ytree2[i] <- if(x[i]<0.63) 4.3 else 9.4
for(i in 1:nobs) ytree3[i] <- if(x[i]<0.75) 3.7 else 9.2
for(i in 1:nobs) ytree4[i] <- if(x[i]<0.55) 4.4 else 8.7
ytree_up <- rep.int(15, length(x))
ytree_up2 <- rep.int(14.5, length(x))
ytree_up3 <- rep.int(15.3, length(x))
ytree_up4 <- rep.int(15.5, length(x))
for(i in 1:nobs) yforest[i] <- if(x[i]<0.55) 4 else {if(x[i]<0.63) 5 else {if(x[i]<0.7) 6.5 else {if(x[i]<0.75) 8 else 9}}}
@
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
#%for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@

\begin{frame}[fragile]
\frametitle{Algorithmic Modeling}
\vspace{-0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {nature};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\draw[<->] (0.1,0.1) .. controls (0.2, -0.4) and (0.4, -0.5) .. (0.8,-0.5);
\draw[<->] (2.7,-0.5) .. controls (3.1, -0.5) and (3.3, -0.4) .. (3.4,0.1);
\node[font=\scriptsize] at (1.75,-0.5) {algorithm};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\visible<2->{
\begin{itemize}
\item Flexible approximation with no/few assumptions. %\\
%$\Rightarrow$ no fixed $f$.
\item Data-driven model fitting.
\end{itemize}
}
\only<2>{\vspace{1cm}}
\visible<3->{
\begin{minipage}{0.6\textwidth}
%\vspace{-0.5cm}
\textbf{Example:} \only<3>{Tree model}\only<4-10>{Subsampling}\only<3>{\\}\only<11->{Forest model}\\
%\vspace{-0.8cm}
\begin{center}
\only<3>{
\vspace{-0.96cm}
\resizebox{0.6\textwidth}{!}{
<<plot_raintree, fig=TRUE, echo=FALSE, width=4, height=2.7>>=
par(mar=c(0,0,0,0), oma=c(0,0,0.4,0))
paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(p,
     tp_args = list(FUN = identity, width = 9, fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1)
@
}

\vspace{-0.7cm}
\begin{flushleft}
\noindent {\small X = rain now, humidity\\
Y = rain 24h}
\end{flushleft}}
\only<4->{
\vspace{-0.5cm}
\resizebox{0.55\textwidth}{!}{
\begin{tikzpicture}
%\draw[ellipse, draw=black, align=center, scale = 0.7, minimum width=170pt, minimum height = 35pt] (set) at (1, 3) {};
\draw[gray,rounded corners=10pt] (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\draw[gray] (1,3.5) ellipse (3 and 0.8);
\begin{scope}
\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\clip (1,3.5) ellipse (3 and 0.8);
\pgfmathsetseed{7}
\foreach \p in {1,...,200} {\fill[gray] (1+3*rand,3.5+0.8*rand) circle (0.04);}
\end{scope}
\visible<5->{
\node[ellipse, draw=black!50!green, align=center, minimum width=80pt, minimum height = 40pt, line width = 3pt] (subset1) at (0.1, 3.5) {};
}
\visible<7->{
\node[ellipse, draw = blue, align=center, minimum width=100pt, minimum height = 30pt, line width = 3pt] (subset2) at (1.2, 3.6) {};
}
\visible<9->{
\node[ellipse, draw = orange, align=center, minimum width=120pt, minimum height = 25pt, line width = 3pt] (subset3) at (1.6, 3.3) {};
}
\visible<6->{
  \node[ellipse, fill=black!50!green, align=center] (n00) at (-1, 2) {};
  \node[ellipse, fill=black!50!green, align=center] (n01) at (-1.25, 1.25) {};
  \draw[-, line width=1pt] (n00) -- (n01);
  \node[rectangle, fill=black!50!green, align=center] (n02) at (-1.5, 0.5) {};
  \draw[-, line width=1pt] (n01) -- (n02);
  \node[rectangle, fill=black!50!green, align=center] (n03) at (-1, 0.5) {};
  \draw[-, line width=1pt] (n01) -- (n03);
  \node[rectangle, fill=black!50!green, align=center] (n04) at (-0.5, 0.5) {};
  \draw[-, line width=1pt] (n00) -- (n04);
}
\visible<8->{
  \node[ellipse, fill=blue, align=center] (n10) at (1, 2) {};
  \node[ellipse, fill=blue, align=center] (n11) at (0.75, 1.25) {};
  \draw[-, line width=1pt] (n10) -- (n11);
  \node[rectangle, fill=blue, align=center] (n12) at (0.5, 0.5) {};
  \draw[-, line width=1pt] (n11) -- (n12);
  \node[rectangle, fill=blue, align=center] (n13) at (1, 0.5) {};
  \draw[-, line width=1pt] (n11) -- (n13);
  \node[rectangle, fill=blue, align=center] (n14) at (1.5, 0.5) {};
  \draw[-, line width=1pt] (n10) -- (n14);
}
\visible<10->{
  \node[ellipse, fill=orange, align=center] (n20) at (3, 2) {};
  \node[ellipse, fill=orange, align=center] (n21) at (2.75, 1.25) {};
  \draw[-, line width=1pt] (n20) -- (n21);
  \node[rectangle, fill=orange, align=center] (n22) at (2.5, 0.5) {};
  \draw[-, line width=1pt] (n21) -- (n22);
  \node[rectangle, fill=orange, align=center] (n23) at (3, 0.5) {};
  \draw[-, line width=1pt] (n21) -- (n23);
  \node[rectangle, fill=orange, align=center] (n24) at (3.5, 0.5) {};
  \draw[-, line width=1pt] (n20) -- (n24);
}
\visible<11->{
\draw[-, line width=1pt] (-1, -0.2) -- (0.2,-0.7);
\draw[-, line width=1pt] (1, -0.2) -- (1,-0.7);
\draw[-, line width=1pt] (3, -0.2) -- (1.8,-0.7);
\node[rectangle, align=center] (combine) at (1, -1) {\large average};
}
\end{tikzpicture}
}
}
\end{center}
\end{minipage}
\hspace{-1cm}
\begin{minipage}{0.4\textwidth}
\begin{center}
\vspace{0.1cm}

\only<3-5>{
%\vspace{0.01cm}
<<plot_motivation_regtree, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=7)
mtext("rain", side = 4, las = TRUE, at = 16, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("now", side = 4, las = TRUE, at = 14.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10.3, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 9, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("now", side = 4, las = TRUE, at = 7.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.5, cex = 3)
@
}
\only<6-7>{
%\vspace{0.01cm}
\hspace{-0.2cm}
<<plot_motivation_randforest1, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=7)
lines(x = x, y = ytree2, col = pal["tree"], lwd=7)
lines(x = x, y = ytree_up2, col = pal["tree"], lwd=7)
mtext("rain", side = 4, las = TRUE, at = 16, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("now", side = 4, las = TRUE, at = 14.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10.3, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 9, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("now", side = 4, las = TRUE, at = 7.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.5, cex = 3)
@
}
\only<8-9>{
%\vspace{0.01cm}
\hspace{-0.3cm}
<<plot_motivation_randforest2, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=7)
lines(x = x, y = ytree2, col = pal["tree"], lwd=7)
lines(x = x, y = ytree3, col = "blue", lwd=7)
lines(x = x, y = ytree_up2, col = pal["tree"], lwd=7)
lines(x = x, y = ytree_up3, col = "blue", lwd=7)
mtext("rain", side = 4, las = TRUE, at = 16, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("now", side = 4, las = TRUE, at = 14.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10.3, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 9, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("now", side = 4, las = TRUE, at = 7.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.5, cex = 3)
@
}
\only<10>{
%\vspace{0.01cm}
\hspace{-0.3cm}
<<plot_motivation_randforest3, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=7)
lines(x = x, y = ytree2, col = pal["tree"], lwd=7)
lines(x = x, y = ytree3, col = "blue", lwd=7)
lines(x = x, y = ytree_up2, col = pal["tree"], lwd=7)
lines(x = x, y = ytree_up3, col = "blue", lwd=7)
lines(x = x, y = ytree4, col = "orange", lwd=7)
lines(x = x, y = ytree_up4, col = "orange", lwd=7)
mtext("rain", side = 4, las = TRUE, at = 16, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("now", side = 4, las = TRUE, at = 14.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10.3, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 9, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("now", side = 4, las = TRUE, at = 7.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.5, cex = 3)
@
}
\only<11->{
%\vspace{0.01cm}
\hspace{-0.3cm}
<<plot_motivation_randforest4, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=7)
lines(x = x, y = yforest, col = pal["forest"], lwd=7)
mtext("rain", side = 4, las = TRUE, at = 16, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("now", side = 4, las = TRUE, at = 14.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("no", side = 4, las = TRUE, at = 10.3, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain", side = 4, las = TRUE, at = 9, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("now", side = 4, las = TRUE, at = 7.7, line = 0.5, cex = 2.5, col = pal["forest"])
mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
mtext("humidity", side = 1, line = 1.5, cex = 3)
@
}
\end{center}
\end{minipage}
}
\end{frame}

\begin{frame}
\frametitle{Goals}
\vspace{-0.4cm}
\begin{minipage}{0.7\textwidth}
\textbf{Available:} \\
\vspace{-0.5 cm}
\begin{itemize}
  \item Distributional regression:\\
  Probabilistic forecast.
  \item Regression Trees:\\
  Automatic detection of abrupt changes,
  and selection of variables/interactions.
  \item Random Forests:\\
  Smoothing and regularization.
\end{itemize}

\vspace{1cm}

\visible<2->{
\textbf{Missing:} Synthesis

\vspace{0.2cm}

$\Rightarrow$ PhD-Thesis

\vspace{0.1cm}

%{\small 
(what I've been working on the last 2 years)
}
\end{minipage}
<<motivation_GAMLSS_syn, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
nobs <- 200
x_g <- c(1:nobs)/nobs
x_g <- 2*(x_g-0.5)
ytrue_g <- x_g^3
var_g <- exp(-(2*x_g)^2)/2
y_g <- ytrue_g + rnorm(nobs, 0, 0.1 + var_g)
@
<<motivation_regtree_syn, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(74)
nobs <- 170
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- ytree <- ytree2 <- ytree3 <- ytree4 <- ytree_up <- yforest <- var <- var_up <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<0.2) 4 else 4+5*plogis(kappa*(3*(x[i]-0.33)-1))
ytrue_up <- rep.int(15, length(x))
for(i in 1:nobs) var_up[i] <- if(x[i]<0.15) 3 else (if(x[i]<0.35) 2 else (if(x[i]<0.5) 2 else (if(x[i]<0.7) 1 else (if(x[i]<0.8) 1.5 else (if(x[i]<0.9) 2 else (2.5))))))
for(i in 1:nobs) var[i] <- if(x[i]<0.15) 4 else (if(x[i]<0.2) 3 else (if(x[i]<0.25) 2 else (if(x[i]<0.4) 1 else (if(x[i]<0.8) 2 else (if(x[i]<0.9) 2.5 else (3))))))
y <- ytrue + rnorm(nobs,0, sqrt(var))
y_up <- ytrue_up + rnorm(nobs,0, sqrt(var_up))
for(i in 1:nobs) ytree[i] <- if(x[i]<0.7) 4 else 9
ytree_up <- rep.int(15, length(x))
for(i in 1:nobs) yforest[i] <- if(x[i]<0.55) 4 else {if(x[i]<0.6) 5 else {if(x[i]<0.7) 6.5 else {if(x[i]<0.75) 8 else 9}}}
@
\begin{minipage}{0.25\textwidth}
\vspace{0.4cm}
<<plot_motivation_GAMLSS_syn, fig=TRUE, echo=FALSE, width=8.5>>=
par(mar=c(1,0,1,0))
plot(x_g, y_g, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x_g, rev(x_g)), c(ytrue_g + 0.1 + var_g, rev(ytrue_g - 0.1 - var_g)),
  col = pallight["forest"], border = "transparent")
lines(x_g, ytrue_g, col = pal["forest"], lwd=10)
points(x_g, y_g, col = "slategray", pch = 19)
box(lwd = 5)
@

<<plot_motivation_regtree_syn, fig=TRUE, echo=FALSE, width=8.5>>=
par(mar=c(1,0,1,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree, col = pal["forest"], lwd=10)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=10)
@

<<plot_motivation_randforest4_syn, fig=TRUE, echo=FALSE, width=8.5>>=
par(mar=c(1,0,1,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=10)
lines(x = x, y = yforest, col = pal["forest"], lwd=10)
@

\vspace{0.2cm}

\visible<2->{
<<plot_motivation_distforest, fig=TRUE, echo=FALSE, width=6>>=
par(mar=c(0,0,3,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(c(y,y_up)), max(c(y,y_up))))
polygon(c(x, rev(x)), c(yforest + var, rev(yforest - var)),
  col = pallight["forest"], border = "transparent")
polygon(c(x, rev(x)), c(ytrue_up + var_up, rev(ytrue_up - var_up)),
  col = pallight["forest"], border = "transparent")
points(x=x, y = y_up, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytree_up, col = pal["forest"], lwd=10)
lines(x = x, y = yforest, col = pal["forest"], lwd=10)
@
}
\end{minipage}
\end{frame}

<<dgp_tree, echo=FALSE, results=hide>>=
data <- data.frame(x = numeric(0), x = numeric(0), x = numeric(0))
names(data) <- c("x","x","x")
fig <- party(
  partynode(1L,
            split = partysplit(2L, breaks = 0.4),
            kids = list(
              partynode(2L, info = c(
                "n = 200",
                "   True parameters:   ",
                expression(mu == '4'),
                expression(sigma == '1')
              )),
              partynode(3L,
                        split = partysplit(3L, breaks = 0.8),
                        kids = list(
                          partynode(4L, info = c(
                            "n = 200",
                            "   True parameters:   ",
                            expression(mu == '12'),
                            expression(sigma == '3')
                          )),
                          partynode(5L, info = c(
                            "n = 100",
                            "   True parameters:   ",
                            expression(mu == '4'),
                            expression(sigma == '3')
                          )))))),
  data
)

node_inner_ext <- function (obj, id = TRUE, pval = TRUE, abbreviate = FALSE, fill = "white", 
    gp = gpar()) 
{
    meta <- obj$data
    nam <- names(obj)
    extract_label <- function(node) {
        if (is.terminal(node)) 
            return(rep.int("", 2L))
        varlab <- character_split(split_node(node), meta)$name
        if (abbreviate > 0L) 
            varlab <- abbreviate(varlab, as.integer(abbreviate))
        if (pval) {
            nullna <- function(x) is.null(x) || is.na(x)
            pval <- suppressWarnings(try(!nullna(info_node(node)$p.value), 
                silent = TRUE))
            pval <- if (inherits(pval, "try-error")) 
                FALSE
            else pval
        }
        if (pval) {
            pvalue <- node$info$p.value
            plab <- ifelse(pvalue < 10^(-3L), paste("p <", 10^(-3L)), 
                paste("p =", round(pvalue, digits = 3L)))
        }
        else {
            plab <- ""
        }
        return(c(varlab, plab))
    }
    maxstr <- function(node) {
        lab <- extract_label(node)
        klab <- if (is.terminal(node)) 
            ""
        else unlist(lapply(kids_node(node), maxstr))
        lab <- c(lab, klab)
        lab <- unlist(lapply(lab, function(x) strsplit(x, "\n")))
        lab <- lab[which.max(nchar(lab))]
        if (length(lab) < 1L) 
            lab <- ""
        return(lab)
    }
    nstr <- maxstr(node_party(obj))
    if (nchar(nstr) < 6) 
        nstr <- "aAAAAa"
    rval <- function(node) {
        node_vp <- viewport(x = unit(0.5, "npc"), y = unit(0.5, 
            "npc"), width = unit(1, "strwidth", nstr) * 1.3, 
            height = unit(3, "lines"), name = paste("node_inner", 
                id_node(node), sep = ""), gp = gp)
        pushViewport(node_vp)
        xell <- c(seq(0, 0.2, by = 0.01), seq(0.2, 0.8, by = 0.05), 
            seq(0.8, 1, by = 0.01))
        yell <- sqrt(xell * (1 - xell))
        xell <- xell*1.11 - 0.055             # to adapt size of the ellipse to the size with p-value
        lab <- extract_label(node)
        fill <- rep(fill, length.out = 2L)
        grid.polygon(x = unit(c(xell, rev(xell)), "npc"), y = unit(c(yell, 
            -yell) + 0.5, "npc"), gp = gpar(fill = fill[1]))
        grid.text(lab[1L], y = unit(1.5 + 0.5,                  # to adapt position of x to its position with p-value
            "lines"))
        #grid.text(lab[1L], y = unit(1.5 + 0.5 * (lab[2L] != ""), 
        #    "lines"))
        grid.text(lab[2L], y = unit(1, "lines"))
        if (id) {
            nodeIDvp <- viewport(x = unit(0.5, "npc"), y = unit(1, 
                "npc"), width = max(unit(1, "lines"), unit(1.3, 
                "strwidth", nam[id_node(node)])), height = max(unit(1, 
                "lines"), unit(1.3, "strheight", nam[id_node(node)])))
            pushViewport(nodeIDvp)
            grid.rect(gp = gpar(fill = fill[2]))
            grid.text(nam[id_node(node)])
            popViewport()
        }
        upViewport()
    }
    return(rval)
}

class(node_inner_ext) <- "grapcon_generator"

@

<<tree, echo=FALSE, results=hide>>=
set.seed(7)
nobs <- 500
x <- runif(nobs, 0, 1)
mu <- sigma <- ytrue <- numeric(length = nobs)
for(i in 1:nobs) sigma[i] <- if(x[i]<=0.4) 1 else 3
for(i in 1:nobs) mu[i] <- if(x[i]<= 0.4|| x[i]>0.8) 4 else 12
y <- rnorm(nobs, mean = mu, sd = sigma)
#y <- rcnorm(nobs, mean = mu, sd = sigma, left = 0)
ytrue <- mu
data <- data.frame(cbind(y,x, ytrue))
tree <- disttree(y ~ x, data = data, family = NO(), type.tree = "mob")
#tree <- disttree(y ~ x, data = data, family = dist_list_cens_normal)
@

\def\checkmark{\tikz\fill[scale=0.4](0,.35) -- (.25,0) -- (1,.7) -- (.25,.15) -- cycle;}

\begin{frame}[fragile]
\frametitle{Distributional Tree}
\begin{minipage}{0.78\textwidth}
\begin{center}
\vspace{0.2cm}
\visible<2->{
Model: \code{disttree(y $\sim$ x)}\\
}
\vspace{-0.2cm}
\only<2-3>{
%\setkeys{Gin}{width=0.\linewidth, height=0.68\linewidth}
<<plottree_estpar, fig=TRUE, echo=FALSE, width = 7>>=
# function for output in terminal panels
FUN <- function (x) 
{
  cf <- x$coefficients
  cf <- matrix(cf, ncol = 1, dimnames = list(names(cf), ""))
  c(sprintf("n = %s", x$nobs), "Estimated parameters:", parse(text = paste0("mu == '", format(round(cf[1], 2), nsmall = 2), "'")), 
                                                        parse(text = paste0("sigma == '", format(round(cf[2], 2), nsmall = 2), "'")))
}

paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))

## plot version using FUN and tree of class 'disttree'
plot(tree, drop = TRUE, tnex = 1.7, FUN = FUN,
     tp_args = list(fill = paltrees[c(1, 3)], width = 18), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
}
\only<4->{
%\setkeys{Gin}{width=0.68\linewidth, height=0.68\linewidth}
<<plottree_dens, fig=TRUE, echo=FALSE, width = 7>>=
node_density <- function (tree, xscale = NULL, yscale = NULL, horizontal = FALSE,
                          main = "", xlab = "", ylab = "Density", id = TRUE, rug = TRUE,
                          fill = paltrees[c(1, 3)], col = "black", lwd = 0.5, ...) 
{
  yobs <- tree$data[,as.character(tree$info$formula[[2]])]
  ylines <- 1.5
  if (is.null(xscale)) xscale <- c(-5.1,22.5)
  if (is.null(yscale)) yscale <- c(-0.05,0.45)
  xr <- xscale
  yr <- yscale

  if (horizontal) {
    yyy <- xscale
    xscale <- yscale
    yscale <- yyy
  }

  rval <- function(node) {
    yrange <- seq(from = -20, to = 90)/4
    ydens <- node$info$object$ddist(yrange)

    top_vp <- viewport(layout = grid.layout(nrow = 2, ncol = 3, 
                                            widths = unit(c(ylines, 1, 1), c("lines", "null", "lines")), 
                                            heights = unit(c(1, 1), c("lines", "null"))), 
                       width = unit(1, "npc"), 
                       height = unit(1, "npc") - unit(2, "lines"), 
                       name = paste("node_density",node$id, sep = ""))
    pushViewport(top_vp)
    grid.rect(gp = gpar(fill = "white", col = 0))
    top <- viewport(layout.pos.col = 2, layout.pos.row = 1)
    pushViewport(top)
    mainlab <- paste(ifelse(id, paste("Node", node$id, "(n = "), "n = "), node$info$nobs, ifelse(id, ")", ""), sep = "")

    grid.text(mainlab)
    popViewport()
    plot <- viewport(layout.pos.col = 2, layout.pos.row = 2, 
                     xscale = xscale, yscale = yscale, 
                     name = paste("node_density",  node$id, "plot", sep = ""))
    pushViewport(plot)
    yd <- ydens
    xd <- yrange
    if (horizontal) {
      yyy <- xd
      xd <- yd
      yd <- yyy
      yyy <- xr
      xr <- yr
      yr <- yyy
      rxd <- rep(0, length(xd))
      ryd <- rev(yd)
    } else {
      rxd <- rev(xd)
      ryd <- rep(0, length(yd))
    }

    if (rug) {
      nodeobs <- node$info$object$y
      if (horizontal) {
        grid.rect(x = xscale[1], y = nodeobs , height = 0, width = xscale[1], 
                  default.units = "native", just = c("right", "bottom"),
		  gp = gpar(lwd = 2, col = gray(0, alpha = 0.18)))
      } else {
        grid.rect(x = nodeobs, y = yscale[1], 
                  width = 0, height = abs(yscale[1]), default.units = "native", 
                  just = c("center", "bottom"),
		  gp = gpar(lwd = 2, col = gray(0, alpha = 0.18)))
        #grid.lines(x = xr, y = yr, gp = gpar(col = "lightgray"), 
        #           default.units = "native")
        #grid.lines(x = xr, y = yr, gp = gpar(col = "lightgray"), 
        #           default.units = "native")
      }
    }

    grid.polygon(x = c(xd, rxd), y = c(yd, ryd), default.units = "native",
              gp = gpar(col = "black", fill = fill, lwd = lwd))
    #grid.lines(x = xd, y = yd, default.units = "native", 
    #           gp = gpar(col = col, lwd = lwd))
    grid.xaxis()
    grid.yaxis()
    grid.rect(gp = gpar(fill = "transparent"))
    upViewport(2)
  }
  return(rval)
}

class(node_density) <- "grapcon_generator"

plot(tree, tnex = 1.7, drop = TRUE,
     terminal_panel = node_density,
     tp_args = list(fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
}
\visible<3->{
R-package \code{disttree}
}
\end{center}
\end{minipage}
\begin{minipage}{0.21\textwidth}

\vspace{1.9cm}

\center
\visible<5->{
Algorithmic \\
Modeling\\
}

\vspace{1.3cm}

\visible<6->{
Data \\
Modeling
}

\vspace{1.4cm}

\visible<7->{
Combined \visible<8->{\checkmark}
}
\end{minipage}
\end{frame}

%<<densities, eval=TRUE, echo=FALSE, results=hide>>=
%x <- c(-300:300)/100
%set.seed(7)
%y1 <- dnorm(x, mean = 0, sd = 0.7)
%y2 <- dnorm(x, mean = 0.8, sd = 0.5)
%y3 <- dnorm(x, mean = -0.7, sd = 1.2)
%@

%<<density1, fig=TRUE, echo=FALSE>>=
%par(mar=c(0,0,0,0))
%plot(x, y1, type = "l", col = "red", ylim = c(0,0.8), lwd=2,
%     xaxt='n', yaxt='n')
%@

%<<density2, fig=TRUE, echo=FALSE>>=
%par(mar=c(0,0,0,0))
%plot(x, y2, type = "l", col = "red", ylim = c(0,0.8), lwd=2,
%     xaxt='n', yaxt='n')
%@

%<<density3, fig=TRUE, echo=FALSE>>=
%par(mar=c(0,0,0,0))
%plot(x, y3, type = "l", col = "red", ylim = c(0,0.8), lwd=2,
%     xaxt='n', yaxt='n')
%@

\begin{frame}
\frametitle{Building Distributional Trees and Forests}
\begin{minipage}{0.7\textwidth}
{\bf Tree:}
\begin{enumerate}
\item<3-> Fit global distributional model $\mathcal{D}(Y; \theta)$: \\ % to the whole data set:\\
Estimate model parameters $\hat{\theta}$.
\item<6-> Evaluate goodness of fit \\
(for each parameter and each observation).
\item<8-> Choose covariate $X$ with strongest influence on goodness of fit of $\mathcal{D}(Y; \hat{\theta})$ as split variable.
\end{enumerate}
\end{minipage}
\begin{minipage}{0.25\textwidth}
\vspace{-1cm}
\begin{tikzpicture}
\visible<2-3>{
\node[ellipse, fill=HighlightBlue!70, align=center, scale = 0.7, minimum width=60pt, minimum height = 30pt] (n0) at (0.8, 2) {$Y$};
}
\visible<4>{
\node[ellipse, fill=HighlightBlue!70, align=center, scale = 0.7, minimum width=60pt, minimum height = 30pt] (n0) at (0.8, 2) {$\mathcal{D}(Y;\hat{\theta}$)};
}
\visible<5->{
\node[inner sep=0pt] (density1) at (0.8,2)
    {\includegraphics[width=0.6\textwidth]{density1.jpeg}};
}
\visible<7-8>{
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n1) at (0, 0) {?};
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n2) at (1.6, 0) {?};
\draw[-, line width=0.8pt] (0.7, 1.4) -- (n1);
\draw[-, line width=0.8pt] (0.9, 1.4) -- (n2);
}
\visible<9-10>{
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n1) at (0, 0) {$Y_1$};
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n2) at (1.6, 0) {$Y_2$};
\draw[-, line width=0.8pt] (0.7, 1.4) -- (n1) node [midway, left] {\scriptsize $X \leq p$};
\draw[-, line width=0.8pt] (0.9, 1.4) -- (n2) node [midway, right] {\scriptsize $X > p$};
}
\visible<11>{
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n1) at (0, 0) {$\mathcal{D}(Y_1;\hat{\theta_1}$)};
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n2) at (1.6, 0) {$\mathcal{D}(Y_2;\hat{\theta_2}$)};
}
\visible<11->{
\draw[-, line width=0.8pt] (0.7, 1.4) -- (n1) node [midway, left] {\scriptsize $X \leq p$};
\draw[-, line width=0.8pt] (0.9, 1.4) -- (n2) node [midway, right] {\scriptsize $X > p$};
}
\visible<12->{
\node[inner sep=0pt] (density2) at (-0.1,-0.25)
    {\includegraphics[width=0.5\textwidth]{density2.jpeg}};
}
\visible<12->{
\node[inner sep=0pt] (density3) at (1.7,-0.25)
    {\includegraphics[width=0.5\textwidth]{density3.jpeg}};
}
\end{tikzpicture}
\end{minipage}
\vspace{0.1cm}
\begin{adjustwidth}{-0.0em}{-3em}
\begin{enumerate}
\setcounter{enumi}{3}
\item<8-> Find the split point $p$ which leads to the highest improvement.
\item<10-> Repeat steps 1--4 recursively in the subgroups
    until some stopping criterion is met.
\end{enumerate}
\end{adjustwidth}
\vspace{0.4cm}
\visible<13->{
{\bf Forest:} Ensemble of trees.
\begin{itemize}
\item Varying subsamples.
\item Combine / average over trees.
\end{itemize}
}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SweaveOpts{eval=TRUE}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.4cm}
\textbf{Goal:} Predict daily precipitation amount in alpine regions.\\
\medskip
\textbf{Physical knowledge:} Numerical Weather Prediction (NWP) system.\\
\medskip
\textbf{Statistical post-processing} necessary due to uncertain initial conditions and unresolved processes.

%Numerical Weather Prediction (NWP) based on a global physical model numerically simulating atmospheric processes on a $50\times 50 \text{km}^2$ grid.\\
\end{frame}

<<echo = FALSE, results=hide>>=
library("zoo")

Sys.setenv("TZ"="UTC")

# -------------------------------------------------------------------
# GFS ensemble forecast
# -------------------------------------------------------------------

GFS   <- read.table("ensemble_plot/GENS_00_innsbruck-flughafen.txt", header = TRUE)
init  <- min(as.POSIXct(GFS$timestamp-GFS$step*3600,origin="1970-01-01"))
title <- sprintf("%s\n%s",
                 "Global Forecast System (GFS) Ensemble Forecast for Innsbruck, Airport",
                 sprintf("Forecast initialized %s",strftime(init,"%Y-%m-%d %H:%M UTC")))

# observations
load("STAGEobs_tawes_11121.rda")
# "2018-03-14 00:00:00" in line 321 (first entry for temperature)
# "2018-03-13 06:00:00" in line 317 (first entry for rain -> aggregated to 24h sums)
# "2018-03-23 00:00:00" in line 357
obs <- obs[318:357,]

getZoo <- function( x, variable ) {
  x <- subset( x, varname == variable )
  if ( nrow(x) == 0 ) stop("Ups, variable seems not to exist at all!")
  # Else create zoo
  x <- zoo( subset(x,select=-c(varname,timestamp,step)),
            as.POSIXct(x$timestamp,origin="1970-01-01") )
  x
}
GFS_t2m  <- getZoo( GFS, "tmp2m" )
# temperature bias: model simulates temperature for height 1675 (~1070 m higher than true height)
# => add 10.7 degrees to account for bias (approx. 1 degree / 100 m)
GFS_t2m <- GFS_t2m + 10.7
GFS_t2m$obs <- obs[,"obs$tl"]
# first value of rain is 18 hours later than temperature 
#(both start at 6:00 but rain is cumulated over 24 hours)
# -> drop first 3 values of temperature
GFS_t2m <- GFS_t2m[-c(1:3),]

GFS_rain <- getZoo( GFS, "apcpsfc" )
GFS_rain$obs <- obs[,"rr6"]
# precipitaion sums over 24 hours (00:00 UTC - 00:00 UTC +1day)
GFS_rain24 <- GFS_rain[c((1:10)*4),]
for(i in 1:NROW(GFS_rain24)){
  GFS_rain24[i,] <- colSums(GFS_rain[c((4*i-3):(4*i)),])
}

# Two POSIXct vectors for the axis
main  <- seq(min(as.POSIXct(as.Date(index(GFS_t2m)))),max(index(GFS_t2m)),by=86400)
minor <- seq(min(index(GFS_t2m)),max(index(GFS_t2m)),by=3*3600)

# GFS_t2m_mean <- rowMeans(GFS_t2m)
GFS_t2m_mean <- GFS_t2m
GFS_t2m_mean[,1] <- rowMeans(GFS_t2m[,-NCOL(GFS_t2m)])
GFS_t2m_mean <- GFS_t2m_mean[,1]

# GFS_rain_mean <- rowMeans(GFS_rain)
GFS_rain24_mean <- GFS_rain24
GFS_rain24_mean[,1] <- rowMeans(GFS_rain24[,-NCOL(GFS_rain24)])
GFS_rain24_mean <- GFS_rain24_mean[,1]
@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{-0.3cm}
\setkeys{Gin}{width=1.1\linewidth}
<<ensemble_1_plot, fig=TRUE, echo=FALSE, width=9>>=
par(mar=c(0.4,0,0.2,0), oma=c(6,3.2,5,4))
layout( matrix(1:2,ncol=1) )
plot( GFS_t2m[,1], screen=1, xaxs="i", xaxt="n", ylim = range(GFS_t2m)*1.05, col = "slategray" )
#plot( GFS_t2m_mean, screen=1, xaxs="i", xaxt="n", ylim = range(GFS_t2m)*1.05)
abline( v = main, lty = 2 )
mtext( side = 3, line = 1, cex = 1.2, font = 2, title )
mtext( side = 2, line = 2.3, cex = 1, font = 1, text = "Temperature [°C]" )
plot( GFS_rain24[,1], screen=1, xaxs="i", xaxt="n", yaxs="i", ylim=range(GFS_rain24)*c(0,1.05), type = "o", col = "slategray" )
#plot( GFS_rain[,1], screen=1, xaxs="i", xaxt="n", yaxs="i", ylim=range(GFS_rain)*c(0,1.05) )
#plot( GFS_rain_mean, screen=1, xaxs="i", xaxt="n", yaxs="i", ylim=range(GFS_rain)*c(0,1.05) )
abline( v = main, lty = 2 )
mtext( side = 2, line = 2.3, cex = 1, font = 1, text = "Rain [mm/6h]" )

#axis( side = 1, at = minor, strftime(minor,"%H") )
axis( side = 1, at = main + 42300,  strftime(main,"%b %d"),
      line = 0, col.ticks=NA, col.axis="gray30", col = NA )
#     line = 2, col.ticks=NA, col.axis="gray30", col = NA )

@
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{-0.3cm}
\setkeys{Gin}{width=1.1\linewidth}
<<ensemble_1_plot+obs, fig=TRUE, echo=FALSE, width=9>>=
par(mar=c(0.4,0,0.2,0), oma=c(6,3.2,5,4))
layout( matrix(1:2,ncol=1) )
plot( GFS_t2m[,1], screen=1, xaxs="i", xaxt="n", 
      ylim = range(GFS_t2m)*1.05, col = "slategray" )
lines( GFS_t2m[,"obs"], col = "red", lwd = 2)
abline( v = main, lty = 2 )
mtext( side = 3, line = 1, cex = 1.2, font = 2, title )
mtext( side = 2, line = 2.3, cex = 1, font = 1, text = "Temperature [°C]" )
plot( GFS_rain24[,1], screen=1, xaxs="i", xaxt="n", yaxs="i", 
      ylim=range(GFS_rain24)*c(0,1.05), type = "o", col = "slategray" )
lines( GFS_rain24[,"obs"], col = "red", lwd = 2)
abline( v = main, lty = 2 )
mtext( side = 2, line = 2.3, cex = 1, font = 1, text = "Rain [mm/6h]" )

#axis( side = 1, at = minor, strftime(minor,"%H") )
axis( side = 1, at = main + 42300,  strftime(main,"%b %d"),
      line = 0, col.ticks=NA, col.axis="gray30", col = NA )
#     line = 2, col.ticks=NA, col.axis="gray30", col = NA )

@
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{-0.3cm}
\setkeys{Gin}{width=1.1\linewidth}
<<ensemble_2_plot+obs, fig=TRUE, echo=FALSE, width=9>>=
par(mar=c(0.4,0,0.2,0), oma=c(6,3.2,5,4))
layout( matrix(1:2,ncol=1) )
plot( GFS_t2m[,c(1,2)], screen=1, xaxs="i", xaxt="n", 
      ylim = range(GFS_t2m)*1.05, col = "slategray" )
lines( GFS_t2m[,"obs"], col = "red", lwd = 2)
abline( v = main, lty = 2 )
mtext( side = 3, line = 1, cex = 1.2, font = 2, title )
mtext( side = 2, line = 2.3, cex = 1, font = 1, text = "Temperature [°C]" )
plot( GFS_rain24[,c(1,2)], screen=1, xaxs="i", xaxt="n", yaxs="i", 
      ylim=range(GFS_rain24)*c(0,1.05), type = "o", col = "slategray" )
lines( GFS_rain24[,"obs"], col = "red", lwd = 2)
abline( v = main, lty = 2 )
mtext( side = 2, line = 2.3, cex = 1, font = 1, text = "Rain [mm/6h]" )

#axis( side = 1, at = minor, strftime(minor,"%H") )
axis( side = 1, at = main + 42300,  strftime(main,"%b %d"),
      line = 0, col.ticks=NA, col.axis="gray30", col = NA )
#     line = 2, col.ticks=NA, col.axis="gray30", col = NA )

@
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{-0.3cm}
\setkeys{Gin}{width=1.1\linewidth}
<<ensemble_3_plot+obs, fig=TRUE, echo=FALSE, width=9>>=
par(mar=c(0.4,0,0.2,0), oma=c(6,3.2,5,4))
layout( matrix(1:2,ncol=1) )
plot( GFS_t2m[,c(1,2,3)], screen=1, xaxs="i", xaxt="n", 
      ylim = range(GFS_t2m)*1.05, col = "slategray" )
lines( GFS_t2m[,"obs"], col = "red", lwd = 2)
abline( v = main, lty = 2 )
mtext( side = 3, line = 1, cex = 1.2, font = 2, title )
mtext( side = 2, line = 2.3, cex = 1, font = 1, text = "Temperature [°C]" )
plot( GFS_rain24[,c(1,2,3)], screen=1, xaxs="i", xaxt="n", yaxs="i", 
      ylim=range(GFS_rain24)*c(0,1.05), type = "o", col = "slategray" )
lines( GFS_rain24[,"obs"], col = "red", lwd = 2)
abline( v = main, lty = 2 )
mtext( side = 2, line = 2.3, cex = 1, font = 1, text = "Rain [mm/6h]" )

#axis( side = 1, at = minor, strftime(minor,"%H") )
axis( side = 1, at = main + 42300,  strftime(main,"%b %d"),
      line = 0, col.ticks=NA, col.axis="gray30", col = NA )
#     line = 2, col.ticks=NA, col.axis="gray30", col = NA )

@
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{-0.3cm}
\setkeys{Gin}{width=1.1\linewidth}
<<ensemble_20_plot+obs, fig=TRUE, echo=FALSE, width=9>>=
par(mar=c(0.4,0,0.2,0), oma=c(6,3.2,5,4))
layout( matrix(1:2,ncol=1) )
plot( GFS_t2m[,-NCOL(GFS_t2m)], screen=1, xaxs="i", xaxt="n", 
      ylim = range(GFS_t2m)*1.05, col = "slategray" )
lines( GFS_t2m[,"obs"], col = "red", lwd = 2)
abline( v = main, lty = 2 )
mtext( side = 3, line = 1, cex = 1.2, font = 2, title )
mtext( side = 2, line = 2.3, cex = 1, font = 1, text = "Temperature [°C]" )
plot( GFS_rain24[,-NCOL(GFS_rain24)], screen=1, xaxs="i", xaxt="n", yaxs="i", 
      ylim=range(GFS_rain24)*c(0,1.05), type = "o", col = "slategray" )
lines( GFS_rain24[,"obs"], col = "red", lwd = 2)
abline( v = main, lty = 2 )
mtext( side = 2, line = 2.3, cex = 1, font = 1, text = "Rain [mm/6h]" )

#axis( side = 1, at = minor, strftime(minor,"%H") )
axis( side = 1, at = main + 42300,  strftime(main,"%b %d"),
      line = 0, col.ticks=NA, col.axis="gray30", col = NA )
#     line = 2, col.ticks=NA, col.axis="gray30", col = NA )

@
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.1cm}
\textbf{Idea:} Statistical model with numerical predictions as covariates.\\
\vspace{0.4cm}
\textbf{Distribution assumption:}\\
\vspace{-0.25cm}
$$(\text{precipitation})^\frac{1}{1.6} \sim \textit{c}\mathcal{N}(\mu,\sigma^2) \quad \text{censored at 0.}$$
%\vspace{2cm}
\visible<2->{
\textbf{Data:}
%\vspace{-1cm}
\begin{itemize}
\item \textbf{Y:} Observation data (National Hydrographical Service):\\
\begin{itemize}
\item 95 observation stations in Tyrol, Austria.\\
\item Daily 24h precipitation sums (July, 1985--2012).
%\item Daily 24h precipitation sums from July over 28 years (1985--2012).
\end{itemize}
\vspace{0.1cm}
\item \textbf{X:} NWP model output (Global Ensemble Forecast System):\\
\vspace{0.14cm}
\quad \small{Ensemble predictions of various meteorological quantities\\
\quad (precipitation, temperature, air pressure, sunshine, \dots)\\
\quad $\rightarrow$ 80 covariates.}
\end{itemize}
}
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
\textbf{Application for one station:}
\begin{itemize}
\item Learn forest model on data from 24 years (1985--2008).
\item Evaluate on 4 years (2009--2012).
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\SweaveOpts{eval=TRUE}

<<pred_axams, echo=FALSE, results=hide>>=
load("data_Axams.rda")

# tree and forest formula
{
  dt.formula <- df.formula <- 
      robs ~ tppow_mean + tppow_sprd + tppow_min + tppow_max + 
      tppow_mean0612 + tppow_mean1218 + tppow_mean1824 + tppow_mean2430 + 
      tppow_sprd0612 + tppow_sprd1218 + tppow_sprd1824 + tppow_sprd2430 + 
      capepow_mean + capepow_sprd + capepow_min + capepow_max + 
      capepow_mean0612 + capepow_mean1218 + capepow_mean1224 + capepow_mean1230 +
      capepow_sprd0612 + capepow_sprd1218 + capepow_sprd1224 + capepow_sprd1230 +
      dswrf_mean_mean + dswrf_mean_max +  
      dswrf_sprd_mean + dswrf_sprd_max + 
      msl_mean_mean + msl_mean_min + msl_mean_max + 
      msl_sprd_mean + msl_sprd_min + msl_sprd_max +
      pwat_mean_mean + pwat_mean_min + pwat_mean_max + 
      pwat_sprd_mean + pwat_sprd_min + pwat_sprd_max +
      tmax_mean_mean + tmax_mean_min + tmax_mean_max +
      tmax_sprd_mean + tmax_sprd_min + tmax_sprd_max +
      tcolc_mean_mean + tcolc_mean_min + tcolc_mean_max +
      tcolc_sprd_mean + tcolc_sprd_min + tcolc_sprd_max +
      t500_mean_mean + t500_mean_min + t500_mean_max +
      t700_mean_mean + t700_mean_min + t700_mean_max +
      t850_mean_mean + t850_mean_min + t850_mean_max +
      t500_sprd_mean + t500_sprd_min + t500_sprd_max +
      t700_sprd_mean + t700_sprd_min + t700_sprd_max +
      t850_sprd_mean + t850_sprd_min + t850_sprd_max +
      tdiff500850_mean + tdiff500850_min + tdiff500850_max +
      tdiff700850_mean + tdiff700850_min + tdiff700850_max +
      tdiff500700_mean + tdiff500700_min + tdiff500700_max +
      msl_diff

  }

learndata <- data_axams[data_axams$year < 109,]
testdata <- data_axams[data_axams$year %in% c(109, 110, 111, 112),]

##############################################################
# fitting the models
set.seed(7)

df <- distforest(df.formula, data = learndata, family = dist_list_cens_normal, 
                 type.tree = "ctree", ntree = 100, censtype = "left", censpoint = 0,
                 control = ctree_control(teststat = "quad", testtype = "Univ",
                                         intersplit = TRUE,
                                         mincriterion = 0, minsplit = 50,
                                         minbucket = 20), mtry = 27)

#### prepare data for plot of estimated density functions
# predictions for one day (in each of the four years) 
# (19th of July 2011 is missing)
pday <- 24  # 2 (hohe Beobachtung zu niedrig geschaetzt), 4, 15, evtl. auch 7, 8, 23 (eine 0-Beobachtung und 2 sehr aehnliche),

pdays <- if(pday<19) c(pday, pday + 31, pday + 62, pday + 92) else c(pday, pday + 31, pday + 61, pday + 92)
pdf <- predict(df, newdata = testdata[pdays,], type = "parameter")
df_mu <- pdf$mu
df_sigma <- pdf$sigma
df_exp <- pnorm(df_mu/df_sigma) * (df_mu + df_sigma * (dnorm(df_mu/df_sigma) / pnorm(df_mu/df_sigma)))
cbind(df_exp, testdata[pdays,"robs"])

# plot predicted distributions together with observations
#set.seed(res$call$seedconst)
set.seed(7)
x <- c(0.01, sort(runif(500,0.01,8)))
y1 <- crch::dcnorm(x, mean = df_mu[1], sd = df_sigma[1], left = 0)
y2 <- crch::dcnorm(x, mean = df_mu[2], sd = df_sigma[2], left = 0)
y3 <- crch::dcnorm(x, mean = df_mu[3], sd = df_sigma[3], left = 0)
y4 <- crch::dcnorm(x, mean = df_mu[4], sd = df_sigma[4], left = 0)
dayending <- if(pday > 3) "th" else switch(pday, "1" = {dayending <- "st"}, "2" = {dayending <- "nd"}, "3" = {dayending <- "rd"})

# point mass (slightly shifted)
pm1 <- c(0.05, crch::dcnorm(-1, mean = df_mu[1], sd = df_sigma[1], left = 0))
pm2 <- c(0.01, crch::dcnorm(-1, mean = df_mu[2], sd = df_sigma[2], left = 0))
pm3 <- c(-0.03, crch::dcnorm(-1, mean = df_mu[3], sd = df_sigma[3], left = 0))
pm4 <- c(-0.07, crch::dcnorm(-1, mean = df_mu[4], sd = df_sigma[4], left = 0))

# predictions
pred1 <- c(testdata[pdays,"robs"][1], crch::dcnorm(testdata[pdays,"robs"][1], mean = df_mu[1], sd = df_sigma[1], left = 0))
pred2 <- c(testdata[pdays,"robs"][2], crch::dcnorm(testdata[pdays,"robs"][2], mean = df_mu[2], sd = df_sigma[2], left = 0))
pred3 <- c(testdata[pdays,"robs"][3], crch::dcnorm(testdata[pdays,"robs"][3], mean = df_mu[3], sd = df_sigma[3], left = 0))
pred4 <- c(testdata[pdays,"robs"][4], crch::dcnorm(testdata[pdays,"robs"][4], mean = df_mu[4], sd = df_sigma[4], left = 0))

#legendheight
lh1 <- crch::dcnorm(0.01, mean = df_mu[1], sd = df_sigma[1], left = 0)
lh2 <- crch::dcnorm(0.01, mean = df_mu[2], sd = df_sigma[2], left = 0)
lh3 <- crch::dcnorm(0.01, mean = df_mu[3], sd = df_sigma[3], left = 0)
lh4 <- crch::dcnorm(0.01, mean = df_mu[4], sd = df_sigma[4], left = 0)
@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\begin{center}
\vspace*{-0.2cm}
<<plot_pred_axams_24July, fig=TRUE, echo=FALSE, height=5.2>>=
plot(x = x, y = y1, type = "l", col = "red", 
     main = paste0("Axams, July ", pday), ylab = "Density", 
     xlab = expression(Total~precipitation~"["~mm^(1/1.6)~"/"~"24h"~"]"),
     #xlab = expression(Total~precipitation~group("[", mm^(1/1.6)~per~day,"]")),
     #{\frac{1}{1.6}}$,
     ylim = c(0,max(y1, y2, y3, y4, pm1, pm2, pm3, pm4) + 0.01),
     xlim = c(-1.5,8))

lines(x = x, y = y2, type = "l", col = "blue")
lines(x = x, y = y3, type = "l", col = "darkgreen")
lines(x = x, y = y4, type = "l", col = "purple")
legend('topright', c("predicted distribution", "point mass at censoring point", "observation"),
       bty = "n", col = "black", lty = c(1, NA, NA), pch = c(NA, 19, 4), cex = 0.8)

# plot point mass
lines(x = c(pm1[1], pm1[1]), y = c(pm1[2], 0), col = "red", type = "l", lwd = 1)
lines(x = c(pm2[1], pm2[1]), y = c(pm2[2], 0), col = "blue", type = "l", lwd = 1)
lines(x = c(pm3[1], pm3[1]), y = c(pm3[2], 0), col = "darkgreen", type = "l", lwd = 1)
lines(x = c(pm4[1], pm4[1]), y = c(pm4[2], 0), col = "purple", type = "l", lwd = 1)

points(x = pm1[1], y = pm1[2], col = "red", pch = 19)
points(x = pm2[1], y = pm2[2], col = "blue", pch = 19)
points(x = pm3[1], y = pm3[2], col = "darkgreen", pch = 19)
points(x = pm4[1], y = pm4[2], col = "purple", pch = 19)

# plot predictions
points(x = pred1[1], y = pred1[2], col = "red", pch = 4)
points(x = pred2[1], y = pred2[2], col = "blue", pch = 4)
points(x = pred3[1], y = pred3[2], col = "darkgreen", pch = 4)
points(x = pred4[1], y = pred4[2], col = "purple", pch = 4)

lines(x = c(pred1[1], pred1[1]), y = c(pred1[2], 0), col = "darkgrey", type = "l", lty = 2)
lines(x = c(pred2[1], pred2[1]), y = c(pred2[2], 0), col = "darkgrey", type = "l", lty = 2)
lines(x = c(pred3[1], pred3[1]), y = c(pred3[2], 0), col = "darkgrey", type = "l", lty = 2)
lines(x = c(pred4[1], pred4[1]), y = c(pred4[2], 0), col = "darkgrey", type = "l", lty = 2)

# add labels
text(x = -0.8, y = lh1, labels = "2009", col = "red", cex = 0.8)
text(x = -0.8, y = lh2, labels = "2010", col = "blue", cex = 0.8)
text(x = -0.8, y = lh3, labels = "2011", col = "darkgreen", cex = 0.8)
text(x = -0.8, y = lh4, labels = "2012", col = "purple", cex = 0.8)
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace{0.2cm}
\textbf{Application for all 95 stations:}
\begin{itemize}
\item Learn forest model on data from 24 years.
\item Evaluate on 4 years.
\item Compare to commonly used heteroscedastic censored Gaussian models.
\end{itemize}
\end{frame}

<<echo = FALSE>>=
#### prediction over all stations 24 - 4
#### prediction over all stations 24 - 4
if(file.exists("crps_24to4_all.rda")){
  load("crps_24to4_all.rda")
} else {

  data("StationsTyrol")
  stations <- StationsTyrol$name
  test <- 2009:2012
  train <- 1985:2008

  res_24to4_all <- mclapply(1:length(stations),
                            function(i){

                              set.seed(7)

                              res <- evalmodels(station = stations[i],
                                                train = train,
                                                test = test,
                                                gamboost_cvr = TRUE)

                              return(res)
                            },
                            mc.cores = detectCores() - 1
  )

  # extract crps
  crps_24to4_all <- matrix(nrow = length(stations), ncol = 7)
  # loop over all stations
  for(i in 1:length(stations)){
    crps_24to4_all[i,] <- res_24to4_all[[i]]$crps
  }

  colnames(crps_24to4_all) <- names(res_24to4_all[[1]]$crps)
  rownames(crps_24to4_all) <- stations

  save(crps_24to4_all, file = "crps_24to4_all.rda")
  save(res_24to4_all, file = "res_24to4_all.rda")

}

# skill score
s <- 1 - crps_24to4_all[, 2:4]/crps_24to4_all[,6] 
colnames(s) <- c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS")

## prepare data for map which shows where distforest performed better than gamlss or gamboostLSS based on the crps

crps_map <- crps_24to4_all[,c("distforest", "gamlss", "gamboostLSS", "emos_log")]

# best method
bst <- apply(crps_map, 1, which.min)

# distance of forest to best other method
dst <- crps_map[,1] - crps_map[cbind(1:nrow(crps_map), apply(crps_map[, -1], 1, which.min) + 1)]

# breaks/groups
brk <- c(-0.1, -0.05, -0.005, 0.005, 0.05, 0.1)
#brk <- c(-0.1, -0.05, -0.01, 0.01, 0.05, 0.1)
grp <- cut(dst, breaks = brk)

# HCL colors (relatively flashy, essentially CARTO Tropic)
clr <- colorspace::diverge_hcl(5, h = c(195, 325), c = 80, l = c(50, 90), power = 1.3)

library("raster") # dem (digital elevation model)
library("sp")     # gadm www.gadm.org/country

load("../draft/plot_map_rain/data/tirol.rda")
load("../draft/plot_map_rain/data/dem.rda")

data(StationsTyrol)
# Create SpatialPointsDataFrame from station list
sp <- SpatialPointsDataFrame(subset(StationsTyrol,
                                    select=c(lon,lat)),
                             data = subset(StationsTyrol,
                                           select = -c(lon,lat)),
                             proj4string = crs(dem))
@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\begin{center}
%\vspace*{-2cm}

\setkeys{Gin}{width=0.95\textwidth}
<<map, fig=TRUE, echo=FALSE, width=10, height=6.5>>=

  ## plot map of Tyrol with all 95 observations
  layout(cbind(1, 2), width = c(9, 1))
  par(mar = c(5,4,4,0.1))
  raster::image(dem, col = rev(gray.colors(100)),
                main="Stations in Tyrol", cex.main=1.7,
                ylab = "Latitude", xlab = "Longitude",
                cex.lab = 1.3,
                xlim = c(9.8,13.2), 
                ylim = c(46.6, 47.87))
  plot(tirol, add = TRUE)
  points(sp, pch = c(21, 24, 25, 22)[bst], bg = clr[grp], col = "black", las = 1, cex = 1.5)
  legend(x = 9.8, y = 47.815, pch = c(21, 24, 25, 22), legend = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS", "EMOS"), cex = 1, bty = "n")
  text(x = 10.25, y = 47.82, labels = "Models with lowest CRPS")
  mtext("CRPS\ndifference", side=4, las = TRUE, at = c(x = 13.5, y = 47.76), line = 0.3)
  par(mar = c(0.5,0.2,0.5,2.3))
  ## legend
  plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
       xlim = c(0, 1), ylim = c(-0.2, 0.2), xaxs = "i", yaxs = "i")
  rect(0, brk[-6], 0.5, brk[-1], col = rev(clr))
  axis(4, at = brk, las = 1, mgp=c(0,-0.4,-1.2))

@
\end{center}
\end{frame}

\begin{frame}
\frametitle{Software}
\vspace{0.4cm}
\textbf{Package:} \emph{disttree} available on R-Forge at\\

\medskip

\url{https://R-Forge.R-project.org/projects/partykit/}\\

\bigskip
\bigskip

\textbf{Main functions:}

\medskip

\begin{tabular}{ll}
\code{distfit}    & Distributional fits (ML, \code{gamlss.family}/custom \code{list}).\\
                  & No covariates. \\
\code{disttree}   & Distributional trees (\code{ctree}/\code{mob} + \code{distfit}).\\
                  & Covariates as partitioning variables. \\
\code{distforest} & Distributional forests (ensemble of \code{disttree}s).\\
                  & Covariates as partitioning variables.
\end{tabular}

\end{frame}

\subsection{References}

\begin{frame}
\frametitle{References}

\vspace{0.5cm}
%\vspace{-0.2cm}

\footnotesize

%Hothorn T, Hornik K, Zeileis A (2006).
% \dquote{Unbiased Recursive Partitioning: A Conditional Inference Framework.}
% \textit{Journal of Computational and Graphical Statistics},
% \textbf{15}(3), 651--674.
% \doi{10.1198/106186006X133933}

%\medskip

%Zeileis A, Hothorn T, Hornik K (2008).
% \dquote{Model-Based Recursive Partitioning.}
%  \textit{Journal of Computational and Graphical Statistics},
%  \textbf{17}(2), 492--514.
%  \doi{10.1198/106186008X319331}

%\medskip

Schlosser L, Hothorn T, Stauffer R, Zeileis A (2018).
  \dquote{Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain.}
  \emph{Manuscript under preparation}.

\bigskip

Hothorn T, Zeileis A (2015).
 \dquote{\textbf{partykit}: A Modular Toolkit for Recursive Partytioning in \textsf{R}.}
 \textit{Journal of Machine Learning Research},
 \textbf{16}, 3905--3909.
 \url{http://www.jmlr.org/papers/v16/hothorn15a.html}

\bigskip

Stasinopoulos DM, Rigby RA (2007).
  \dquote{Generalized Additive Models for Location Scale and Shape (GAMLSS) in \textsf{R}.}
  \textit{Journal of Statistical Software}, 
  \textbf{23}(7), 1--46.
  \doi{10.18637/jss.v023.i07}

%\medskip

%Seibold H, Zeileis A, Hothorn T (2017).
%  \dquote{Individual Treatment Effect Prediction for Amyotrophic Lateral Sclerosis Patients.}
%  \textit{Statistical Methods in Medical Research}, 
%  \textbf{12}(1), 45--63.
%  \doi{10.1177/0962280217693034}

%\medskip

%Hothorn T, Zeileis A (2017).
%  \dquote{Transformation Forests.}
%  \emph{arXiv 1701.02110}, arXiv.org E-Print Archive.
%  \url{http://arxiv.org/abs/1701.02110}

%\medskip

%Gneiting T, Raftery A E, Westveld III A H, Goldman T (2005).
%  \dquote{Calibrated Probabilistic Forecasting using Ensemble Model Output Statistics and Minimum CRPS Estimation}
%  \textit{Monthly Weather Review}, 
%  \textbf{133}(5), 1098--1118.
%  \doi{10.1175/MWR2904.1}

%\medskip

%Hofner B, Mayr A, Schmid M (2016).
%  \dquote{\textbf{gamboostLSS}: An \textsf{R} Package for Model Building and Variable Selection in the {GAMLSS} Framework}
%  \textit{Journal of Statistical Software},
%  \textbf{74}(1), 1--31.
%  \doi{10.18637/jss.v074.i01}

\end{frame}

%\begin{frame}
%\frametitle{References (2)}

%\vspace{-0.2cm}

%\footnotesize

%% Reference for the observation data set
%BMLFUW (2016).
%  \dquote{{B}undesministerium f\"ur {L}and und {F}orstwirtschaft, {U}mwelt und {W}asserwirtschaft ({BMLFUW}), {A}bteilung {IV/4} -- {W}asserhaushalt}
%  \text{Available at \url{http://ehyd.gv.at}}, 
%  \text{Accessed: 2016--02--29}.

%\medskip

% Reference for the GFS reforecast version 2 data set 
%Hamill T M, Bates G T, Whitaker J S, Murray D R, Fiorino M, Galarneau T J Jr., Zhu Y, Lapenta W (2013).
%  \dquote{{NOAA}'s Second-Generation Global Medium-Range Ensemble Reforecast Dataset}
%  \textit{Bulletin of the American Meteorological Society}, 
%  \textbf{94}(10), 1553--1565.
%  \doi{10.1175/BAMS-D-12-00014.1}

%\medskip

%Breiman L (2001).
%  \dquote{Random {Forests}}
%  \textit{Machine Learning}, 
%  \textbf{45}(1), 5--32.
%  \doi{10.1023/A:1010933404324}

%Messner J W,  Mayr G J, Zeileis A (2016).
%  \dquote{Heteroscedastic Censored and Truncated Regression with \textbf{crch}}
%  \textit{The \textsf{R} Journal}, 
%  \textbf{8}(1), 173--181.\\
%  \scriptsize
%  \url{https://journal.R-project.org/archive/2016-1/messner-mayr-zeileis.pdf}
%  \footnotesize

%\medskip

%Stauffer R, Mayr G J, Messner J W, Umlauf N, Zeileis A (2017).
%  \dquote{Spatio-Temporal Precipitation Climatology Over Complex Terrain Using a Censored Additive Regression Model}
%  \textit{International Journal of Climatology}, 
%  \textbf{37}(7), 3264--3275.
%  \doi{10.1002/joc.4913}

%\medskip

%Gebetsberger M, Messner J W, Mayr G J, Zeileis A (2017).
%  \dquote{Fine-Tuning Non-Homogeneous Regression for Probabilistic Precipitation Forecasts: Unanimous Predictions, Heavy Tails, and Link %Functions}
%  \textit{Monthly Weather Review}, 
%  \text{Forthcoming}.
%  \doi{10.1175/MWR-D-16-0388.1}

%\medskip

%{CGIAR-CSI} (2016).
%  \dquote{{SRTM} 90m Digital Elevation Database {v4.1}}
%  \text{Available at \url{http://srtm.csi.cgiar.org}}, 
%  \text{Accessed: 2016--07--10}.

%\end{frame}

\end{document}