### 2018-08-29 pkg/disttree/inst/draft/draft.Rnw
highlighted changes
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2268 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 394c73009df503993cb104295fbe472d06896c85

\documentclass[nojss]{jss}
\shortcites{Hamill+Bates+Whitaker:2013}

%\VignetteIndexEntry{disttree: Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain}
%\VignetteDepends{disttree, crch, countreg, gamlss, gamlss.cens}
%\VignetteKeywords{parametric models, regression trees, random forests, recursive partitioning, probabilistic forecasting, GAMLSS}
%\VignettePackage{disttree}

%% packages
\usepackage{amstext,amsfonts,amsmath,bm,thumbpdf,lmodern}

%% tikz
\usepackage{array,makecell,tikz,color,soul}
\usetikzlibrary{arrows.meta,positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap}

%% need no \usepackage{Sweave}
\SweaveOpts{engine = R, concordance = FALSE, eps = FALSE, keep.source = TRUE, eval = TRUE}

%% commands
\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
\renewcommand{\Prob}{\mathbb{P} }
\renewcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Var}{\mathbb{V}}
\newcommand{\R}{\mathbb{R} }
\newcommand{\N}{\mathbb{N} }
\newcommand{\C}{\mathbb{C} }
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}
\newcommand{\LS}{\mathcal{L}_n}
\newcommand{\TS}{\mathcal{T}_n}
\newcommand{\LSc}{\mathcal{L}_{\text{comb},n}}
\newcommand{\LSbc}{\mathcal{L}^*_{\text{comb},n}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\yn}{y_{\text{new}}}
\newcommand{\z}{\mathbf{z}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\sX}{\mathcal{X}}
\newcommand{\sY}{\mathcal{Y}}
\newcommand{\T}{\mathbf{T}}
\newcommand{\x}{\mathbf{x}}
\renewcommand{\a}{\mathbf{a}}
\newcommand{\xn}{\mathbf{x}_{\text{new}}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\ws}{\mathbf{w}_\cdot}
\renewcommand{\t}{\mathbf{t}}
\newcommand{\M}{\mathbf{M}}
\renewcommand{\vec}{\text{vec}}
\newcommand{\B}{\mathbf{B}}
\newcommand{\K}{\mathbf{K}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\bS}{\mathbf{S}}
\newcommand{\cellx}{\pi_n[\x]}
\newcommand{\partn}{\pi_n(\mathcal{L}_n)}
\newcommand{\err}{\text{Err}}
\newcommand{\ea}{\widehat{\text{Err}}^{(a)}}
\newcommand{\ecv}{\widehat{\text{Err}}^{(cv1)}}
\newcommand{\ecvten}{\widehat{\text{Err}}^{(cv10)}}
\newcommand{\eone}{\widehat{\text{Err}}^{(1)}}
\newcommand{\eplus}{\widehat{\text{Err}}^{(.632+)}}
\newcommand{\eoob}{\widehat{\text{Err}}^{(oob)}}
\newcommand{\bft}{\mathbf{t}}

\numberwithin{equation}{subsection}

<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
library("disttree")
library("crch")
library("countreg")
library("gamlss")
library("gamlss.cens")
gen.cens(NO, type = "left")
library("RainTyrol")

## HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)
pallight <- hcl(c(10, 128, 260, 290, 70), 100, 50, alpha = 0.25)
transpgray <- rgb(0.190,0.190,0.190, alpha = 0.2)
@

\title{Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain}
\Shorttitle{Distributional Regression Forests for Probabilistic Precipitation Forecasting}

\author{Lisa Schlosser\\Universit\"at Innsbruck
   \And Torsten Hothorn\\Universit\"at Z\"urich
   \And Reto Stauffer\\Universit\"at Innsbruck
   \And Achim Zeileis\\Universit\"at Innsbruck}
\Plainauthor{Lisa Schlosser, Torsten Hothorn, Reto Stauffer, Achim Zeileis}

\Abstract{
To obtain a probabilistic model for a dependent variable based on some set of
explanatory variables, a distributional approach is often adopted where the
parameters of the distribution are linked to regressors. In many classical
models this only captures the location of the distribution but
over the last decade there has been increasing interest in distributional
regression approaches modeling all parameters including location, scale, and
shape. Notably, so-called non-homogenous Gaussian regression (NGR) models both
mean and variance of a Gaussian response and is particularly
popular in weather forecasting. More generally, the GAMLSS framework
allows to establish generalized additive models for location, scale, and shape
with smooth linear or nonlinear effects.
%
However, when variable selection is required and/or there are
non-smooth dependencies or interactions (especially unknown or of high-order),
it is challenging to establish a good GAMLSS. A natural alternative in these
situations would be the application of regression trees or random forests but,
so far, no general distributional framework is available for these. Therefore,
a framework for distributional regression trees and forests is proposed that blends
regression trees and random forests with classical distributions from the GAMLSS
framework as well as their censored or truncated counterparts.
%
To illustrate these novel approaches in practice, they are employed to obtain
probabilistic precipitation forecasts at numerous sites in a mountainous
region (Tyrol, Austria) based on a large number of numerical weather
prediction quantities. It is shown that the novel distributional regression forests
automatically select variables and interactions, performing on par or often even
better than GAMLSS specified either through prior meteorological knowledge or a
computationally more demanding boosting approach.
}
\Keywords{parametric models, regression trees, random forests, recursive partitioning, probabilistic forecasting, GAMLSS}

\Address{
  Lisa Schlosser, Reto Stauffer, Achim Zeileis \\
  Universit\"at Innsbruck \\
  Department of Statistics \\
  Faculty of Economics and Statistics \\
  Universit\"atsstr.~15 \\
  6020 Innsbruck, Austria \\
  E-mail: \email{Lisa.Schlosser@uibk.ac.at}, \email{Reto.Stauffer@uibk.ac.at},\\
  \email{Achim.Zeileis@R-project.org} \\
  URL: \url{https://www.uibk.ac.at/statistics/personal/schlosser-lisa/},\\
  \phantom{URL: }\url{https://retostauffer.org/},\\
  \phantom{URL: }\url{https://eeecon.uibk.ac.at/~zeileis/}\\

  Torsten Hothorn\\
  Universit\"at Z\"urich \\
  Institut f\"ur Epidemiologie, Biostatistik und Pr\"avention \\
  Hirschengraben 84\\
  CH-8001 Z\"urich, Switzerland \\
  E-mail: \email{Torsten.Hothorn@R-project.org}\\
  URL: \url{http://user.math.uzh.ch/hothorn/}\\

}

\begin{document}

\section{Introduction}
\label{sec:introduction}

In regression analysis a wide range of models has been developed to describe the
relationship between a response variable and a set of covariates. The classical
model is the linear model (LM) where the conditional mean of the response
is modeled through a linear function of the covariates (see the left panel of
Figure~\ref{fig:gam} for a schematic illustration). Over the last decades
this has been extended in various directions including:

\pagebreak

\begin{itemize}
  \item \emph{Generalized linear models} (GLMs, \citealp{Nelder+Wedderburn:1972})
    encompassing an additional nonlinear link function for the conditional mean.
  \item \emph{Generalized additive models} (GAMs, \citealp{Hastie+Tibshirani:1986})
    allowing for smooth nonlinear effects in the covariates
    (Figure~\ref{fig:gam}, middle).
  \item \emph{Generalized additive models for location, scale, and shape}
    (GAMLSS, \citealt{Rigby+Stasinopoulos:2005}) adopting a probabilistic modeling
    approach. In GAMLSS, each parameter of a statistical distribution can depend
    on an additive predictor of the covariates comprising linear and/or
    smooth nonlinear terms (Figure~\ref{fig:gam}, right).
\end{itemize}
Thus, the above-mentioned models provide a broad toolbox for capturing different
aspects of the response (mean only vs.\ full distribution) and different types
of dependencies on the covariates (linear vs.\ nonlinear additive terms).

\begin{figure}[t!]
\begin{tikzpicture}
\node (a) at (0,0){
\minipage{0.29\textwidth}
\centering
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@

<<plot_motivation_GLM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal[1], lwd = 7, main = "")
@
\endminipage};
\node (b) at (5,0) 
{
\minipage{0.29\textwidth}
\centering
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@

<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal[1], lwd = 7, main = "")
@
\endminipage};
\node (c) at (10,0) 
{
\minipage{0.29\textwidth}
\centering
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@

<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight[1], border = "transparent")
lines(x, ytrue, col = pal[1], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
\endminipage};
\node (d) at (0,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
LM, GLM  
\endminipage};
\node (e) at (5,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
GAM 
\endminipage};
\node at (10,-2)
{
\minipage{0.28\textwidth}
\centering
\vspace{0.0cm}
GAMLSS 
\endminipage};
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](2.1,0)--(2.9,0);
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](7.1,0)--(7.9,0);
\end{tikzpicture}
\caption{\label{fig:gam}Parametric modeling developments. (Generalized) linear models (left),
generalized additive models (middle), generalized additive models for location, scale, and shape (right).}
\end{figure}

While in many applications conditional mean regression models have been receiving
the most attention, there has been a paradigm shift over the last decade towards
distributional regression models. An important reason for this is that in many
fields forecasts of the mean are not the only (or not even the main) concern but
instead there is an increasing interest in probabilistic forecasts. Quantities of
interest typically include exceedance probabilities for certain thresholds of the
response or quantiles of the response distribution. Specifically, consider
weather forecasting where there is less interest in the mean amount of
precipitation on the next day. Instead, the probability of rain vs.\ no rain
is typically more relevant or, in some situations, a prediction interval of
expected precipitation (say from the expected \textit{10\%} to \textit{90\%} quantile). Similar
considerations apply for other meteorological quantities and hence attention
in the weather forecasting literature has been shifting from classical linear 
deterministic models \citep{Glahn+Lowry:1972} towards probabilistic models such as the
non-homogeneous Gaussian regression (NGR) of \cite{Gneiting+Raftery+Westveld:2005}.
The NGR typically describes the mean of some meteorological response variable through
the average of the corresponding quantity from an ensemble of physically-based
numerical weather predictions (NWPs). Similarly, the variance of the response
is captured through the variance of the ensemble of NWPs. Thus, the NGR
considers both the mean as well as the uncertainty of the ensemble predictions
to obtain probabilistic forecasts calibrated to a particular site.

\begin{figure}[t!]
\begin{tikzpicture}
\node (a) at (0,0){
\minipage{0.29\textwidth}
\centering
<<motivation_regtree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- ytree <- yforest <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))
y <- ytrue + rnorm(nobs,0,0.3)
for(i in 1:nobs) ytree[i] <- if(x[i]<1/3) 0.5 else {if(x[i]<2/3) 2 else 1}
@

<<plot_motivation_regtree, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x, y = ytree, col = pal[1], lwd=7)
@
\endminipage};
\node (b) at (5,0) 
{
\minipage{0.29\textwidth}
\centering
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@

<<plot_motivation_randforest, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x, y = yforest, col = pal[1], lwd=7, main = "")
@
\endminipage};
\node (c) at (10,0) 
{
\minipage{0.29\textwidth}
\centering
<<motivation_distforest, echo=FALSE, results=hide>>=
## distforest
set.seed(4723)
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
xt <- 2*(x-0.5)
var <- exp(-(2*xt)^4)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
x <- c(1:nobs)/nobs
@

<<plot_motivation_distforest, fig=TRUE, echo=FALSE>>=
par(mar = c(2, 0, 2, 0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
usr <- par("usr")
rect(usr[1], usr[3], usr[2], usr[4], col = gray(0.94), border = "transparent")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight[1], border = "transparent")
lines(x = x, y = yforest, col = pal[1], lwd=7, main = "")
points(x, y, col = "slategray", pch = 19)
box(lwd=8)
@
\endminipage};
\node (d) at (0,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
Regression tree 
\endminipage};
\node (e) at (5,-2) 
{
\minipage{0.29\textwidth}
\centering
\vspace{0.0cm}
Random forest 
\endminipage};
\node at (10,-2)
{
\minipage{0.28\textwidth}
\centering
\vspace{0.0cm}
Distributional forest 
\endminipage};
\node (g) at (0,-5.4) 
{
\minipage{0.29\textwidth}
\vspace{0.2cm}
%\hspace{0.0cm}
\centering
<<motivation_disttree, echo=FALSE, results=hide>>=
## disttree
set.seed(74)
x <- c(1:nobs)/nobs
var <- ytree/3
y <- ytrue + rnorm(nobs,0,var)
@

<<plot_motivation_disttree, fig=TRUE, echo=FALSE>>=
par(mar = c(2, 0, 2, 0))
plot(x, y, xaxt="n", yaxt="n", ann=FALSE, type = "n")
usr <- par("usr")
rect(usr[1], usr[3], usr[2], usr[4], col = gray(0.94), border = "transparent")
lines(x = x, y = ytree, col = pal[1], lwd=7, main = "")
polygon(c(x, rev(x)), c(ytree + var, rev(ytree - var)),
  col = pallight[1], border = "transparent")
points(x, y, col = "slategray", pch = 19)
box(lwd=8)
@
\vspace{-0.2cm}%\\
Distributional tree
\vspace{0.2cm}
\endminipage};
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](2.1,0)--(2.9,0);
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](7.1,0)--(7.9,0);
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](0,-2.4)--(0,-3.2);
\draw[-{>[scale=4, length=2, width=3]},line width=0.4pt](2.2,-5.4)--(7.9,-2.5);
\end{tikzpicture}
\caption{\label{fig:tree}Tree and forest developments.
Regression tree (top left), distributional tree (bottom left), random forest
(top middle), and distributional forest (top right).}
\end{figure}

In summary, the models discussed so far provide a broad and powerful toolset for
parametric distributional fits depending on a specified set of additive linear or
smooth nonlinear terms. A rather different approach to capturing the dependence on
covariates are tree-based models.
\begin{itemize}
  \item \emph{Regression trees} (\citealt{Breiman+Friedman+Stone:1984}) recursively
    split the data into more homogeneous subgroups and can thus capture abrupt shifts
    (Figure~\ref{fig:tree}, top left) and approximate nonlinear functions.
    Furthermore, trees automatically carry out a forward selection of covariates and
    their interactions. 
  \item \emph{Random forests} (\citealt{Breiman:2001}) average the predictions of
    an ensemble of trees fitted to resampled versions of the learning data. This
    stabilizes the recursive partitions from individual trees and hence better
    approximates smooth functions (Figure~\ref{fig:tree}, top middle).
\end{itemize}
While classical regression trees and random forests only model the mean of the response
we propose to follow the ideas from GAMLSS modeling -- as outlined
in Figure~\ref{fig:gam} -- and combine tree-based methods with parametric distributional models,
yielding two novel techniques:
\begin{itemize}
  \item \emph{Distributional regression trees} (for short: \emph{distributional trees}) split the data into more homogeneous groups with
    respect to a parametric distribution, thus capturing changes in any distribution
    parameter like location, scale, or shape (Figure~\ref{fig:tree}, bottom left).
  \item \emph{Distributional regression forests} (for short: \emph{distributional forests}) utilize an ensemble of distributional trees
    for obtaining stabilized and smoothed parametric predictions (Figure~\ref{fig:tree},
    top right).
\end{itemize}
In the following, particular focus is given to distributional forests as a method for obtaining
probabilistic forecasts by leveraging the strengths of random forests: the ability to
capture both smooth and abruptly changing functions along with simultaneous selection of
variables and possibly complex interactions. Thus, these properties make the method particularly
appealing in case of many covariates with unknown effects and interactions where it would be 
challenging to specify a distributional regression model like GAMLSS.

<<echo=FALSE, results=hide>>=
#### Axams prediction 24 - 4
if(file.exists("Axams_24to4.rda")){
  load("Axams_24to4.rda")
} else {
  source("Axams_24to4.R")
  Axams_24to4 <- Axams_24to4()
  save(Axams_24to4, file = "Axams_24to4.rda")
}

#### prepare data for plot of estimated density functions
# predictions for one day (in each of the four years) 
# (19th of July 2011 is missing)
pday <- 24  # 2 (hohe Beobachtung zu niedrig geschaetzt), 4, 15, evtl. auch 7, 8, 23 (eine 0-Beobachtung und 2 sehr aehnliche),

pdays <- if(pday<19) c(pday, pday + 31, pday + 62, pday + 92) else c(pday, pday + 31, pday + 61, pday + 92)
pdf <- predict(Axams_24to4$df, newdata = Axams_24to4$testdata[pdays,], type = "parameter")
df_mu <- pdf$mu
df_sigma <- pdf$sigma
df_exp <- pnorm(df_mu/df_sigma) * (df_mu + df_sigma * (dnorm(df_mu/df_sigma) / pnorm(df_mu/df_sigma)))
cbind(df_exp, Axams_24to4$testdata[pdays,"robs"])

# plot predicted distributions together with observations
set.seed(7)
x <- c(0.01, sort(runif(500,0.01,8)))
y1 <- crch::dcnorm(x, mean = df_mu[1], sd = df_sigma[1], left = 0)
y2 <- crch::dcnorm(x, mean = df_mu[2], sd = df_sigma[2], left = 0)
y3 <- crch::dcnorm(x, mean = df_mu[3], sd = df_sigma[3], left = 0)
y4 <- crch::dcnorm(x, mean = df_mu[4], sd = df_sigma[4], left = 0)
dayending <- if(pday > 3) "th" else switch(pday, "1" = {dayending <- "st"}, "2" = {dayending <- "nd"}, "3" = {dayending <- "rd"})

# switch x-axis back to untransformed scale
x <- x^(1.6)

# point mass (slightly shifted)
pm1 <- c(0.15, crch::dcnorm(-1, mean = df_mu[1], sd = df_sigma[1], left = 0))     #  0.05
pm2 <- c(0.05, crch::dcnorm(-1, mean = df_mu[2], sd = df_sigma[2], left = 0))     #  0.01
pm3 <- c(-0.05, crch::dcnorm(-1, mean = df_mu[3], sd = df_sigma[3], left = 0))    # -0.03
pm4 <- c(-0.15, crch::dcnorm(-1, mean = df_mu[4], sd = df_sigma[4], left = 0))    # -0.07

# observations
obs1 <- c(Axams_24to4$testdata[pdays,"robs"][1]^(1.6), crch::dcnorm(Axams_24to4$testdata[pdays,"robs"][1], mean = df_mu[1], sd = df_sigma[1], left = 0))
obs2 <- c(Axams_24to4$testdata[pdays,"robs"][2]^(1.6), crch::dcnorm(Axams_24to4$testdata[pdays,"robs"][2], mean = df_mu[2], sd = df_sigma[2], left = 0))
obs3 <- c(Axams_24to4$testdata[pdays,"robs"][3]^(1.6), crch::dcnorm(Axams_24to4$testdata[pdays,"robs"][3], mean = df_mu[3], sd = df_sigma[3], left = 0))
obs4 <- c(Axams_24to4$testdata[pdays,"robs"][4]^(1.6), crch::dcnorm(Axams_24to4$testdata[pdays,"robs"][4], mean = df_mu[4], sd = df_sigma[4], left = 0))

#legendheight
lh1 <- crch::dcnorm(0.01, mean = df_mu[1], sd = df_sigma[1], left = 0)
lh2 <- crch::dcnorm(0.01, mean = df_mu[2], sd = df_sigma[2], left = 0)
lh3 <- crch::dcnorm(0.01, mean = df_mu[3], sd = df_sigma[3], left = 0)
lh4 <- crch::dcnorm(0.01, mean = df_mu[4], sd = df_sigma[4], left = 0)

############
# get predicted parameter of all models for testdata

# distributional tree
pdt <- predict(Axams_24to4$dt, newdata = Axams_24to4$testdata, type = "parameter")
dt_mu <- pdt$mu
dt_sigma <- pdt$sigma
dt_exp <- pnorm(dt_mu/dt_sigma) * (dt_mu + dt_sigma * (dnorm(dt_mu/dt_sigma) / pnorm(dt_mu/dt_sigma)))

# distributional forest
pdf <- predict(Axams_24to4$df, newdata = Axams_24to4$testdata, type = "parameter")
df_mu <- pdf$mu
df_sigma <- pdf$sigma
df_exp <- pnorm(df_mu/df_sigma) * (df_mu + df_sigma * (dnorm(df_mu/df_sigma) / pnorm(df_mu/df_sigma)))

# prespecified GAM
g_mu <- predict(Axams_24to4$g, newdata = Axams_24to4$testdata, what = "mu", type = "response", data = Axams_24to4$g_learndata)
g_sigma <- predict(Axams_24to4$g, newdata = Axams_24to4$testdata, what = "sigma", type = "response", data = Axams_24to4$g_learndata)
g_exp <- pnorm(g_mu/g_sigma) * (g_mu + g_sigma * (dnorm(g_mu/g_sigma) / pnorm(g_mu/g_sigma)))

# boosted GAM
pgb <- predict(Axams_24to4$gb, newdata = Axams_24to4$testdata, parameter = c("mu","sigma"), type = "response")
gb_mu <- pgb$mu
gb_sigma <- pgb$sigma
gb_exp <- pnorm(gb_mu/gb_sigma) * (gb_mu + gb_sigma * (dnorm(gb_mu/gb_sigma) / pnorm(gb_mu/gb_sigma)))

# EMOS
ml_mu <- predict(Axams_24to4$ml, type = "location", newdata = Axams_24to4$testdata)
ml_sigma <- predict(Axams_24to4$ml, type = "scale", newdata = Axams_24to4$testdata)
ml_exp <- pnorm(ml_mu/ml_sigma) * (ml_mu + ml_sigma * (dnorm(ml_mu/ml_sigma) / pnorm(ml_mu/ml_sigma)))

############
## PIT histograms and qqrplots

## get predicted parameter
# for learndata and testdata
{
  set.seed(7)

  # distributional tree
  pit_dt <- cbind(0, pnorm(Axams_24to4$testdata[,"robs"], mean = dt_mu, sd = dt_sigma))
  pit_dt[Axams_24to4$testdata[,"robs"]>0, 1] <- pit_dt[Axams_24to4$testdata[,"robs"]>0, 2]

  # distributional forest
  pit_df <- cbind(0, pnorm(Axams_24to4$testdata[,"robs"], mean = df_mu, sd = df_sigma))
  pit_df[Axams_24to4$testdata[,"robs"]>0, 1] <- pit_df[Axams_24to4$testdata[,"robs"]>0, 2]

  # prespecified GAM
  pit_g <- cbind(0, pnorm(Axams_24to4$testdata[,"robs"], mean = g_mu, sd = g_sigma))
  pit_g[Axams_24to4$testdata[,"robs"]>0, 1] <- pit_g[Axams_24to4$testdata[,"robs"]>0, 2]

  # boosted GAM
  pit_gb <- cbind(0, pnorm(Axams_24to4$testdata[,"robs"], mean = gb_mu, sd = gb_sigma))
  pit_gb[Axams_24to4$testdata[,"robs"]>0, 1] <- pit_gb[Axams_24to4$testdata[,"robs"]>0, 2]

  # EMOS
  pit_ml <- cbind(0, pnorm(Axams_24to4$testdata[,"robs"], mean = ml_mu, sd = ml_sigma))
  pit_ml[Axams_24to4$testdata[,"robs"]>0, 1] <- pit_ml[Axams_24to4$testdata[,"robs"]>0, 2]

}

## Variable importance
if(file.exists("vimp_crps.rda")){
  load("vimp_crps.rda")
} else {

  set.seed(7)
  # locally redefine logLik.distforest to use crps in varimp()
  logLik.distforest <- function(object, newdata = NULL, ...) {
    if(is.null(newdata)) {
      newdata <- object$data
    }

    # predict parameter
    pdf <- predict(object, newdata = newdata, type = "parameter")
    df_mu <- pdf$mu
    df_sigma <- pdf$sigma

    # calculate CRPS
    crps <- mean(crps_cnorm(newdata$robs, location = df_mu, scale = df_sigma, lower = 0, upper = Inf), na.rm = TRUE)

    return(structure(crps, df = NA, class = "logLik"))
  }

  vimp_crps <- varimp(Axams_24to4$df, nperm = 1L)

  save(vimp_crps, file = "vimp_crps.rda")

  rm(logLik.distforest) 
}

@

\begin{figure}[t!]
\centering
\setkeys{Gin}{width=0.75\textwidth}
<<rain_Axams_pred, fig=TRUE, echo=FALSE, width=6.5, height=3.8>>=
par(mar = c(3.8, 4, 2.3, 1.5))
plot(x = x, y = y1, type = "l", col = pal[1], lwd = 1.3, 
     main = paste0("July ", pday), ylab = "Density", 
     xlab = expression(Total~precipitation~"["~mm~"/"~"24h"~"]"),
     # xlab = expression(Total~precipitation~"["~mm^(1/1.6)~"/"~"24h"~"]"),
     ylim = c(0,max(y1, y2, y3, y4, pm1, pm2, pm3, pm4) + 0.01),
     xlim = c(-2,28))     #  xlim = c(-1.5,8))

lines(x = x, y = y2, type = "l", col = pal[3], lwd = 1.3)
lines(x = x, y = y3, type = "l", col = pal[2], lwd = 1.3)
lines(x = x, y = y4, type = "l", col = pal[4], lwd = 1.3)
legend("topright", c("Predicted distribution", "Point mass at censoring point", "Observation"),
       bty = "n", col = "black", lty = c(1, NA, NA), pch = c(NA, 19, 4), cex = 0.8, lwd = 1.3)

# plot point mass
lines(x = c(pm1[1], pm1[1]), y = c(pm1[2], 0), col = pal[1], type = "l", lwd = 1.3)
lines(x = c(pm2[1], pm2[1]), y = c(pm2[2], 0), col = pal[3], type = "l", lwd = 1.3)
lines(x = c(pm3[1], pm3[1]), y = c(pm3[2], 0), col = pal[2], type = "l", lwd = 1.3)
lines(x = c(pm4[1], pm4[1]), y = c(pm4[2], 0), col = pal[4], type = "l", lwd = 1.3)

points(x = pm1[1], y = pm1[2], col = pal[1], pch = 19)
points(x = pm2[1], y = pm2[2], col = pal[3], pch = 19)
points(x = pm3[1], y = pm3[2], col = pal[2], pch = 19)
points(x = pm4[1], y = pm4[2], col = pal[4], pch = 19)

# plot observations
points(x = obs1[1], y = obs1[2], col = pal[1], pch = 4, cex = 1.4)
points(x = obs2[1], y = obs2[2], col = pal[3], pch = 4, cex = 1.4)
points(x = obs3[1], y = obs3[2], col = pal[2], pch = 4, cex = 1.4)
points(x = obs4[1], y = obs4[2], col = pal[4], pch = 4, cex = 1.4)

lines(x = c(obs1[1], obs1[1]), y = c(obs1[2], 0), col = "darkgray", type = "l", lty = 2, lwd = 1.3)
lines(x = c(obs2[1], obs2[1]), y = c(obs2[2], 0), col = "darkgray", type = "l", lty = 2, lwd = 1.3)
lines(x = c(obs3[1], obs3[1]), y = c(obs3[2], 0), col = "darkgray", type = "l", lty = 2, lwd = 1.3)
lines(x = c(obs4[1], obs4[1]), y = c(obs4[2], 0), col = "darkgray", type = "l", lty = 2, lwd = 1.3)

# add labels
text(x = -1.7, y = lh1, labels = "2009", col = pal[1], cex = 0.8)     # -0.8
text(x = -1.7, y = lh2, labels = "2010", col = pal[3], cex = 0.8)     # -0.8
text(x = -1.7, y = lh3, labels = "2011", col = pal[2], cex = 0.8)     # -0.8
text(x = -1.7, y = lh4, labels = "2012", col = pal[4], cex = 0.8)     # -0.8
@
\caption{\label{fig:axams-dist}\hl{Total precipitation predictions by a distributional forest
at station Axams for July 24 in 2009, 2010, 2011 and 2012 learned on data from 1985--2008.
Observations are left-censored at 0. The corresponding predicted point mass is shown at the censoring point (0).}}
\end{figure}

In weather forecasting, these properties are especially appealing in 
mountainous regions and complex terrain where a wide range of local-scale
effects are not yet resolved by the NWP models. 
Thus, effects with abrupt changes and possibly nonlinear interactions 
might be required to account for site-specific unresolved features.
To illustrate this in practice,
precipitation forecasts are obtained with distributional forests at 95
meteorological stations in a mountainous region in the Alps, covering mainly
Tyrol, Austria, and adjacent areas (see the map in Figure~\ref{fig:all-map}).
More specifically, a Gaussian distribution, left-censored at zero, is employed to model 24-hour total precipitation so that the
zero-censored point mass describes the probability of observing no precipitation 
on a given day (see Figure~\ref{fig:axams-dist}). Forecasts for July are
established based on data from the same month over the years 1985--2012 including
80~covariates derived from a wide range of different NWP quantities. As
Figure~\ref{fig:axams-dist} shows, the station-wise forests yield a full
distributional forecast for each day -- here for one specific day (July~24) at one
station (Axams) over four years (2009--2012) -- based on the previous 24~years
as learning data. The corresponding observations conform reasonably well with
the predictions. In Section~\ref{sec:precipitation} we investigate the
performance of distributional forests in this forecasting task in more detail.
Compared to three alternative zero-censored Gaussian models distributional forests
perform at least on par and sometimes clearly better while requiring no expert
knowledge for the model specification. The three alternatives are: a standard ensemble model
output statistics approach \citep[EMOS,][]{Gneiting+Raftery+Westveld:2005} based
on an NGR, a GAMLSS with regressors prespecified based on meteorological expertise
\citep[following][]{Stauffer+Umlauf+Messner:2017}, and a boosted GAMLSS
\citep{Hofner+Mayr+Schmid:2016} using non-homogeneous boosting
\citep{Messner+Mayr+Zeileis:2017} as an alternative technique for variable selection among all
80~available regressors.

\section{Methodology}
\label{sec:methods}

To embed the distributional approach from GAMLSS into regression trees and random
forests, we proceed in three steps. (1)~To fix notation, we briefly review fitting
distributions using standard maximum likelihood in Section~\ref{sec:distfit}.
(2)~A recursive partitioning strategy based on the corresponding scores (or
gradients) is introduced in Section~\ref{sec:disttree}, leading to distributional
trees. (3)~Ensembles of distributional trees fitted to randomized subsamples are
employed to establish distributional forests in Section~\ref{sec:distforest}.

The general distributional notation is exemplified in all three steps for the
Gaussian distribution, left-censored at zero. The latter is employed in the empirical 
case study in Section~\ref{sec:precipitation} to model power-transformed daily 
precipitation amounts.

\subsection{Distributional fit}
\label{sec:distfit}

A distributional model $\mathcal{D}(Y, \bm{\theta})$ is considered for the response
variable $Y \in \mathcal{Y}$ using the distributional family $\mathcal{D}$ with
\textit{k}-dimensional parameter vector $\bm{\theta} \in \bold{\Theta}$ and
corresponding log-likelihood function $\ell(\bm{\theta}; Y)$. The GAMLSS
framework \citep{Rigby+Stasinopoulos:2005} provides a wide range of such
distributional families with parameterizations corresponding to location, scale,
and shape. Furthermore, censoring and/or truncation of these distributions
can be incorporated in the usual straightforward way
\citep[see e.g.,][Chapter~7.2]{Long:1997}.

To capture both location and scale of the probabilistic precipitation forecasts
while accounting for a point mass at zero (i.e., dry days without
rain), a Gaussian distribution left-censored at zero (for short: zero-censored Gaussian)
with location parameter $\mu$ and scale parameter $\sigma$ is employed. Therefore,
the corresponding log-likelihood function with parameter vector 
$\bm{\theta} = (\mu, \sigma)$ is
$$
\ell(\mu, \sigma; Y) = 
\begin{cases}
    \log\left\{\frac{1}{\sigma} \cdot \phi\left(\frac{Y - \mu}{\sigma}\right) \right\}, & \text{if } Y > 0\\[0.2cm]
    \log\left\{\Phi\left(\frac{-\mu}{\sigma}\right)\right\}, & \text{if } Y = 0
\end{cases}
$$

where $\phi$ and $\Phi$ are the probability density function and the 
distribution function of the standard normal distribution $\mathcal{N}(0,1)$. 
Other distributions $\mathcal{D}$ and corresponding log-likelihoods
$\ell(\mu, \sigma; Y)$ could be set up in the same way, e.g., for
censored shifted gamma distributions \citep{Scheuerer+Hamill:2015} or
zero-censored logistic distributions \citep{Gebetsberger+Messner+Mayr:2017}.

With the specification of the distribution family and its log-likelihood 
function the task of fitting a distributional model turns into the task 
of estimating the distribution parameter~$\bm{\theta}$. This is commonly done by
maximum likelihood (ML) based on the learning sample with observations
$\{y_i\}_{i = 1, \dots, n}$ of the response variable $Y$. The maximum
likelihood estimator (MLE) $\bm{\hat \theta}$ is given by
\begin{equation}
\label{eq:mle}
\bm{\hat \theta} = \argmax_{\bm{\theta} \in \bold{\Theta}} \sum_{i=1}^n \ell(\bm{\theta}; y_i).
\end{equation}
Equivalently, this can be defined based on the corresponding first-order
conditions
\begin{equation}
\label{eq:foc}
\sum_{i = 1}^n s(\bm{\hat{\theta}}, y_i) = 0,
\end{equation}
where $s(\bm{\theta}; y_i)$ is the associated score function
\begin{equation}
\label{eq:score}
s(\bm{\theta}; y_i) = \frac{\partial \ell}{\partial \bm{\theta}}(\bm{\theta}; y_i).
\end{equation}
The latter is subsequently employed as a general goodness-of-fit measure
to assess how well the distribution with parameters $\bm{\theta}$ fits
one individual observation $y_i$.

\subsection{Distributional tree}
\label{sec:disttree}

Typically, a single global model $\mathcal{D}(Y, \bm{\theta})$ is not sufficient
for reasonably representing the response distribution and covariates
$\bold{Z} = Z_1, \dots, Z_m \in \mathcal{Z}$ are
employed to capture differences in the distribution parameters $\bm{\theta}$.
In weather forecasting, these covariates typically include the output
from numerical weather prediction systems and/or lagged meteorological observations.

To incorporate the covariates into the distributional model, 
they are considered as regressors in additive predictors
$g_j(\theta_j) = f_{j,1}(\bold{Z}) + f_{j,2}(\bold{Z}) + \dots$
in GAMLSS.
Link functions $g_j(\cdot)$ are used for every parameter $\theta_j$
($j = 1, \dots, k$) based on smooth terms $f_{j,k}$ such as
nonlinear effects, spatial effects, random coefficients, or interaction surfaces
\citep{Klein+Kneib+Lang:2015}. However, this requires specifying the
additive terms and their functional forms in advance which can be challenging in
practice and potentially require domain knowledge, especially if the number of covariates $m$ is large.

Regression trees generally take a different approach for automatically
including covariates in a data-driven way and allowing for abrupt
changes, nonlinear and non-additive effects, and interactions. In the context of
distributional models the goal is to partition the covariate space
$\mathcal{Z}$ recursively into disjoint segments so that a homogenous
distributional model for the response $Y$ can be found with segment-specific
parameters. More specifically, the $B$ disjoint segments $\mathcal{B}_b$
($b = 1, \dots, B$) partition the covariate space
\begin{equation}
\label{eq:partition}
\mathcal{Z} = \dot{\bigcup\limits_{b = 1, \ldots, \textit{B}}} \mathcal{B}_b,
\end{equation}
and a local distributional model $\mathcal{D}(Y, \bm{\theta}^{(b)})$
(i.e., with segment-specific parameters $\bm{\theta}^{(b)}$) is fitted to the
response $Y$ in each segment.

To find the segments $\mathcal{B}_b$ that are (approximately) homogenous
with respect to the distributional model with given parameters, the idea
is to use a gradient-based recursive-partitioning approach. In a given
subsample of the learning data this fits the model by ML (see
Equation~\ref{eq:mle}) and then assesses the goodness of fit by assessing the
corresponding scores $s(\bm{\hat \theta}; y_i)$ (see Equation~\ref{eq:score}).

To sum up, distributional trees are fitted recursively via:
\begin{enumerate}
\item Estimate $\bm{\hat \theta}$ via maximum likelihood for the observations
  in the current subsample.
\item Test for associations or instabilities of the scores $s(\bm{\hat \theta}, y_i)$ 
  and $Z_{l,i}$ for each partitioning variable~$Z_l$ ($l = 1, \dots, m$).
\item Split the sample along the partitioning variable $Z_l^*$ with the
  strongest association or instability. Choose the breakpoint with the highest
  improvement in the log-likelihood or the highest discrepancy.
\item Repeat steps 1--3 recursively in the subsamples until these become too
  small or there is no significant association/instability (or some other
  stopping criterion is reached).
\end{enumerate}
%
Different inference techniques can be used for assessing the association between
scores and covariates in step~3. In the following we use the general class of
permutation tests introduced by \cite{Hothorn+Hornik+VanDeWiel:2006} which is
also the basis of conditional inference trees \citep[CTree,][]{Hothorn+Hornik+Zeileis:2006}.
Alternatively, one could use asymptotic M-fluctuation tests for parameter
instability \citep{Zeileis+Hornik:2007} as in model-based recursive partitioning
\citep[MOB,][]{Zeileis+Hothorn+Hornik:2008}. More details are provided in
Appendix~\ref{app:tree}.

For obtaining probabilistic predictions from the tree for a (possibly new) set
of covariates $\bold{z} = (z_1, \ldots, z_m)$, the observation simply has to
be ``sent down'' the tree and the corresponding segment-specific MLE has to be
obtained. This can also be understood as a weighted MLE where the weights
select those observations from the learning sample that fall into the same
segment:
$$
w^{\text{tree}}_i(\bold{z}) = \sum_{b=1}^B \mathbf{1}((\bold{z}_i \in \mathcal{B}_b) \land (\bold{z} \in \mathcal{B}_b)),
$$
where $\mathbf{1}(\cdot)$ is the indicator function. The predicted distribution
for a given $\bold{z}$ is then fully specified by the estimated parameter
$\bm{\hat \theta}(\bold{z})$ where
$$
\bm{\hat \theta}(\bold{z}) = \argmax_{\bm{\theta} \in \bm{\Theta}} \sum_{i=1}^n w^{\text{tree}}_i(\bold{z}) \cdot \ell(\bm{\theta}; y_i).
$$
\hl{Even though this parameter depends on the given $z$ it has to be noted that
it is the same for all other observations that are in the same terminal node of 
the tree as $z$ and therefore does not necessarily have to be recalculated for
each of them.}

\subsection{Distributional forest}
\label{sec:distforest}

While the simple recursive structure of a tree model is easy to visualize and
interpret, the abrupt changes are often too rough, instable, and impose steps
on the model even if the true underlying effect is smooth. Hence, ensemble
methods such as bagging or random forests \citep{Breiman:2001} are typically
applied to smooth the effects, stabilize the model, and improve predictive performance.

While Breiman and Cutler's random forests \citep{Breiman:2001} grow ensembles of trees
that pick up changes in the location of the response across the covariates,
\emph{distributional forests} employ an ensemble of $T$~\emph{distributional trees}. These
pick up changes in the ``direction'' of any distribution parameter by
considering the full score vector for choosing splitting variables and split
points. Each of the distributional trees is grown on a different data set
obtained through bootstrap sampling (or subsampling) and in each node only a
random subset of the covariates $\bold{Z}$ is considered. As usual in random
forests, this reduces the correlation among the trees and stabilizes the
variance of the model.

To obtain probabilistic predictions from a distributional forest, it still
needs to be specified how to compute the parameter estimates
$\bm{\hat \theta}(\bold{z})$ for a (potentially new) set of covariates $\bold{z}$.
Following \cite{Hothorn+Zeileis:2017} we interpret random forests as adaptive
local likelihood estimators using the averaged ``nearest neighbor weights''
\citep{Lin+Jeon:2006} from the $T$~trees in the forest
$$
w^{\text{forest}}_i(\bold{z}) = \frac{1}{T} \sum_{t=1}^T \sum_{b=1}^{B^t}
\frac{\mathbf{1}((\bold{z}_i \in \mathcal{B}^t_b) \land (\bold{z} \in \mathcal{B}^t_b))}{|\mathcal{B}^t_b|},
$$
where $|\mathcal{B}^t_b|$ denotes the number of observations in the $b$-th
segment of the $t$-th tree.
Thus, these $w^{\text{forest}}_i(\bold{z}) \in [0, 1]$ whereas
$w^{\text{tree}}_i(\bold{z}) \in \{0, 1\}$. Hence, weights cannot only be $0$
or $1$ but change more smoothly, giving high weight to those observations $i$
from the learning sample that co-occur in the same segment $\mathcal{B}_b^t$
as the new observation $\bold{z}$ for many of the trees $t = 1, \dots, T$.
Consequently, the parameter estimates may, in principle, change for every
observation and can be obtained by
$$
\bm{\hat \theta}(\bold{z}) =  \argmax_{\bm{\theta} \in \bm{\Theta}} \sum_{i=1}^n w^{\text{forest}}_i(\bold{z}) \cdot \ell(\bm{\theta}; y_i).
$$
In summary, this yields a parametric distributional regression model 
(through the score-based approach) that can capture both abrupt effects and
high-order interactions (through the trees) and smooth effects (through
the forest).

Distributional forests share some concepts and algorithmic aspects 
with other generalizations of Breiman and Cutler's random forests. Nearest 
neighbor weights are employed for aggregation in survival forests 
\citep{Hothorn+Lausen+Benner:2004},
quantile regression forests \citep{Meinshausen:2006},
transformation forests \citep{Hothorn+Zeileis:2017},
and generalized random forests for causal inferences 
\citep{Athey+Tibshirani+Wager:2017}.
These procedures aggregate over trees fitted to specific score functions 
(e.g., log rank scores in survival trees, model residuals in transformation 
or generalized forests). Distributional forests, in contrast to these 
nonparametric approaches, provide a compromise between model flexibility 
and interpretability: The parameters of a problem-specific distribution 
(zero-censored Gaussian for precipitation) have a clear meaning but 
may depend on external variables in a quite general way.

\section{Probabilistic precipitation forecasting in complex terrain}
\label{sec:precipitation}

Many statistical weather forecasting models leverage the strengths of modern
numerical weather prediction (NWP) systems
\citep[see][]{Bauer+Thorpe+Brunet:2015}.  
One frequently used method based on
distributional regression models is the ensemble model output statistics 
(EMOS) approach first proposed by
\cite{Gneiting+Raftery+Westveld:2005} to produce high-quality forecasts for
specific quantities and sites.  
In case of precipitation forecasting EMOS typically uses the ensemble mean of 
``total precipitation'' (\emph{tp}) forecasts as predictor for the location parameter and
the corresponding ensemble standard deviation for the scale part of the statistical model
to correct for possible errors of the ensemble in both, the expectation but also
the uncertainty of a specific forecast.

While this approach alone is already highly effective in the plains,
it typically does not perform as well in complex terrain due to unresolved
effects in the NWP system. For example, in the Tyrolean Alps -- considered
in the following case study -- the NWP grid cells of $50 \times 50$
km$^2$ are too coarse to capture single mountains, narrow valleys, etc. 
Therefore, it is often possible to substantially improve the predictive 
performance of the EMOS by including additional predictor variables, either 
from local meteorological observations or an NWP model. Unfortunately, it is typically unknown 
which variables are relevant for improving the predictions. Simply including
all available variables may be computationally burdensome and
can lead to overfitting but, on the other hand, excluding too many variables
may result in a loss of valuable information. Therefore, selecting
the relevant variables and interactions among all possible covariates is 
crucial for improving the statistical forecasting model.

In the following, it is illustrated how distributional forests can
solve this problem without requiring prior expert knowledge.  For fitting the
forest only the response distribution and the list of potential predictor
variables need to be specified (along with a few algorithmic details) and then
the relevant variables, interactions, and potentially nonlinear effects are
determined automatically in a data-driven way. Here, we employ a zero-censored
Gaussian distribution and 80~predictor variables computed from ensemble means
and spreads of various NWP outputs. The predictive performance of the forest is
compared to three other zero-censored Gaussian models: (a)~a standard basic
EMOS, (b)~a GAMLSS with prespecified effects and interactions based on
meteorological knowledge/experience, and (c)~a boosted GAMLSS with automatic
selection of smooth additive terms based on all 80~predictor variables.

\subsection{Data}
\label{sec:data}

\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
\begin{minipage}{\textwidth}
\caption[Table caption text]{Basic covariates together with the number ({\#}) and the type of variations.
Time periods indicate aggregation time periods in hours after NWP model
initialization (e.g., 6--30 corresponds to +6\,h to +30\,h ahead forecasts,
0600\,UTC to 0600\,UTC of the next day). $^*$Minimum values of \emph{dswrf}
over 24\,h are always zero and thus neglected.}
\label{tab:covariates}
\begin{tabular}{ l  c  l }
\hline
Basic covariates & {\#} & Variations\\
\hline
\emph{tp}: total precipitation,           & 12  & ensemble mean of sums over 24h, \\
\hspace*{0.5cm} power transformed (by $1.6^{-1}$) &    & ensemble std. deviation of sums over 24h, \\ 
\emph{cape}: convective available     &    & ensemble minimum of sums over 24h, \\
\hspace*{0.9cm} potential energy, &    & ensemble maximum of sums over 24h\\
\hspace*{0.9cm} power transformed (by $1.6^{-1}$)&    & \qquad all for 6--30 \\
\hspace*{\fill}                            &    & ensemble mean of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 \\
\hspace*{\fill}                            &    & ensemble std. deviation of sums over 6h\\
\hspace*{\fill}                            &    & \qquad for 6--12, 12--18, 18--24, 24--30 \\
\hline
\emph{dswrf}: downwards short wave      & 6 & ensemble mean of mean values, \\
\hspace*{1.13cm} radiation flux (``sunshine'') &   & ensemble mean of minimum values$^*$,\\
\emph{msl}: mean sea level pressure     &   &  ensemble mean of maximal values,\\
\emph{pwat}: precipitable water         &   & ensemble std. deviation of mean values,\\
\emph{tmax}: 2m maximum temperature     &   & ensemble std. deviation of minimum values$^*$,\\
\hspace*{1.15cm}                          &   & ensemble std. deviation of maximal values,\\
\emph{tcolc}: total column-integrated   &   & \qquad all over 6--30 \\
\hspace*{0.95cm} condensate               &   & \\
\emph{t500}: temperature on 500 hPa     &   & \\
\hspace*{1cm}                             &   & \\
\emph{t700}: temperature on 700 hPa     &   & \\
\hspace*{1cm}                             &   & \\
\emph{t850}: temperature on 850 hPa     &   & \\
\hspace{1cm}                              &   & \\
\hline
\emph{tdiff500850}: temperature         & 3 & ensemble mean of difference in mean,\\
\hspace*{\fill} difference 500 to 850 hPa &   & ensemble minimum of difference in mean,\\
\emph{tdiff500700}: temperature         &   & ensemble maximum of difference in mean\\
\hspace*{\fill} difference 500 to 700 hPa &   & \qquad all over 6--30 \\
\emph{tdiff700850}: temperature         &   & \\
\hspace*{\fill} difference 700 to 850 hPa &   & \\
\hline
\emph{msl{\_}diff}: mean sea level pressure & 1 & \emph{msl{\_}mean{\_}max} $-$ \emph{msl{\_}mean{\_}min}\\
\hspace{1.6cm} difference            &   & \qquad over 6--30 \\
\hline
\end{tabular}
\end{minipage} 
%}
\end{table}

Learning and validation data consist of observed 
daily precipitation sums provided by the National Hydrographical Service 
(\citealt{ehyd}) and numerical weather forecasts from the U.S.~National
Oceanic and Atmospheric Administration (NOAA).
Both, observations and forecasts are available for 1985--2012 and
the analysis is exemplified using July, the month with the most precipitation
in Tyrol.

Observations are obtained for 95~stations all over Tyrol and surroundings,
providing 24-hour precipitation sums measured at 0600\,UTC and rigorously
quality-checked by the National Hydrographical Service. NWP outputs
are obtained from the second generation reforecast data set of the
global ensemble forecast system \citep[GEFS,][]{Hamill+Bates+Whitaker:2013}. 
This data set consists of an 11-member ensemble based on a fixed version of the 
numerical model and a horizontal grid spacing of about {$50 \times 50$ km$^2$}
initialized daily at 0000~UTC from December 1984 to present 
providing forecasts on a 6-hourly temporal resolution. Each of the 
11 ensemble members uses slightly different perturbed initial 
conditions to predict the situation-specific uncertainty 
of the atmospheric state.

From the GEFS, 14 basic forecast variables are considered with up 
to 12 variations each such as mean/maximum/minimum
over different aggregation time periods. A detailed overview is
provided in Table~\ref{tab:covariates}, yielding 80~predictor variables in total.

To remove large parts of the skewness of precipitation data, 
a power transformation \citep{Box+Cox:1964} is often applied, e.g., using
cubic \citep{Stidd:1973} or square root \citep{Hutchinson:1998} transformations.
However, the power parameter may vary for different climatic zones or temporal 
aggregation periods and hence we follow \cite{Stauffer+Mayr+Messner:2017} 
in their choice of $1.6^{-1}$ as a suitable power parameter for the
region of Tyrol. The same power transformation is applied to both the
observed precipitation sums and the NWP outputs ``total precipitation'' (\emph{tp}) and
``convective available potential energy'' (\emph{cape}).

\subsection{Models and evaluation}
\label{sec:evaluation}

\begin{table}[t!]
\centering
\caption[Table caption text]{Overview of models with type of 
covariate dependency and included covariates for each distribution 
parameter. \emph{A} $\ast$ \emph{B} indicates an interaction between
covariate \emph{A} and \emph{B}.}
\label{tab:models}
\begin{tabular}{ l l l l }
\hline
Model & Type & Location ($\mu$) & Scale ($\log(\sigma)$)                                  \\ \hline
Distributional forest & recursive    & all                & all                           \\ 
                      & partitioning &                    &                               \\ \hline
EMOS                  & linear       & \emph{tp{\_}mean}  & \emph{tp{\_}sprd}             \\ \hline
Prespecified GAMLSS   & spline       & \emph{tp{\_}mean}, & \emph{tp{\_}sprd},            \\
                      & in each      & \emph{tp{\_}max},  & \emph{dswrf{\_}sprd{\_}mean}, \\
 & & \emph{tp{\_}mean1218} $\ast$ & \emph{tp{\_}sprd1218} $\ast$\\ 
 & & \quad \emph{cape{\_}mean1218}, & \quad \emph{cape{\_}mean1218},\\
 & & \emph{dswrf{\_}mean{\_}mean}, & \emph{tcolc{\_}sprd{\_}mean},\\
 & & \emph{tcolc{\_}mean{\_}mean}, & \emph{tdiff500850{\_}mean}\\
 & & \emph{pwat{\_}mean{\_}mean}, & \\
 & & \emph{tdiff500850{\_}mean}, & \\
 & & \emph{msl{\_}diff} & \\ \hline
Boosted GAMLSS        & spline  & all & all \\
                      & in each &     &     \\ \hline
\end{tabular}
\end{table}

The following zero-censored Gaussian regression models are employed
in the empirical case study, see Table~\ref{tab:models} for
further details:
%
\begin{itemize}

\item \emph{Distributional forest:} All 80~predictor variables are
  considered for learning a forest of 100 trees. Bootstrap
  sampling is employed for each tree using a third of the predictors
  in each split of the tree (\code{mtry}). Parameters are estimated by
  adaptive local likelihood based on the forest weights as described
  in Section~\ref{sec:methods}. 
  \hl{As stopping criteria the minimal number
  of observations to perform a split} (\code{minsplit}) \hl{is set to 50, 
  the minimal number of observations in a terminal node} (\code{minbucket})
  \hl{is set to 20 and the significance level for variable selection} (\code{alpha}) 
  \hl{is kept at the default value $1$.}
  % stopping criteria:
  % forest_minsplit = 50, forest_minbucket = 20, forest_mincrit = 0

  % default in ctree_control within distforest:
  % mincriterion = 0

  % default values in ctree_control:
  % alpha = 0.05, mincriterion = 1 - alpha,
  % minsplit = 20L, minbucket = 7L,
  % minprob = 0.01

\item \emph{EMOS:} Standard ensemble model output statistics
  models use the ensemble mean of total precipitation as regressor in the location
  submodel and the corresponding ensemble standard deviation in the scale submodel.
  The parameters are estimated by maximum likelihood, using an identity
  link for the location part and a log link for the scale part
  \citep[following the advice of][]{Gebetsberger+Messner+Mayr:2017}.

\item \emph{Prespecified GAMLSS:} Smooth additive splines are selected for
  the most relevant predictors based on meteorological expert knowledge
  following \cite{Stauffer+Umlauf+Messner:2017}. More specifically, based
  on the 80~available variables, eight terms are included in the location
  submodel and five in the scale submodel. Both involve an interaction of
  \emph{tp} and \emph{cape} in the afternoon (between 1200\,UTC and 1800\,UTC)
  to capture the potential for thunderstorms that frequently occur in
  summer afternoons in the Alps. The model is estimated by maximum
  penalized likelihood using a backfitting algorithm \citep{Stasinopoulos+Rigby:2007}.

\item \emph{Boosted GAMLSS:} Smooth additive splines are selected
  automatically from all 80~available variables, using non-cyclic boosting
  for parameter estimation \citep{Hofner+Mayr+Schmid:2016,Messner+Mayr+Zeileis:2017}.
  This updates the predictor terms for the location or scale submodels iteratively
  by maximizing the log-likelihood only for the variable yielding the
  biggest improvement. The iteration stops early -- before fully maximizing
  the in-sample likelihood -- based on a (computationally intensive)
  out-of-bag bootstrap estimate of the log-likelihood. The grid considered for
  the number of boosting iterations (``mstop'') is: $50, 75, \dots, 975, 1000$.

\end{itemize}

The predictive performance in terms of full probabilistic forecasts is
assessed using the continuous ranked probability score (CRPS, \citealt{Hersbach:2000}).
For each of the models this assesses the discrepancy of the
predicted distribution function $F$ from the observation $y$ by
$$
\text{CRPS}(y, F) = \int_{-\infty}^\infty (F(z) - \mathbf{1}(y \leq z))^2 dz
$$
where \(\mathbf{1}(\cdot)\) is the indicator function. In the subsequent
applications, the mean CRPS is always evaluated out of sample, 
either using cross-validation or a hold-out data set (2009--2012)
that was not used for learning (1985--2008). CRPS is a proper scoring rule
\citep{Gneiting+Raftery:2007} often used within the meteorological community.
Lower values indicate better performance.

To assess differences in the improvement of the forests and GAMLSS
models over the basic EMOS, a CRPS-based skill score with EMOS as
the reference method is computed:
$$
\text{CRPSS}_{\text{method}} = 1 - \frac{\text{CRPS}_{\text{method}}}{\text{CRPS}_{\text{EMOS}}}.
$$

\subsection{Application for one station}
\label{sec:axams}

\begin{figure}[t!]
\centering
\setkeys{Gin}{width=0.99\textwidth}
<<rain_axams_qqr_oosample, fig=TRUE, echo=FALSE, width = 6.5, pdf = FALSE, png = TRUE, resolution = 150>>=
# out of sample
set.seed(7)
par(mfrow = c(2, 2))
qqrplot(pit_df, nsim = 100, main = "Distributional forest", ylim = c(-5, 5), col = gray(0.04, alpha = 0.01), pch = 19)
qqrplot(pit_ml, nsim = 100, main = "EMOS",                  ylim = c(-5, 5), col = gray(0.04, alpha = 0.01), pch = 19)
qqrplot(pit_g,  nsim = 100, main = "Prespecified GAMLSS",   ylim = c(-5, 5), col = gray(0.04, alpha = 0.01), pch = 19)
qqrplot(pit_gb, nsim = 100, main = "Boosted GAMLSS",        ylim = c(-5, 5), col = gray(0.04, alpha = 0.01), pch = 19)
@
\caption{\label{fig:axams-qq} Out-of-sample residual QQ~plots (2009--2012) 
for station Axams based on models learned on data from 1985--2008. 
(Residuals of precipitation are on the powertransformed scale.)}
\end{figure}

In a first step, we show a detailed comparison of the competing models
for one observation site, Axams in Tyrol (in the center of the study area,
see Figure~\ref{fig:all-map}). As for all other stations, daily precipitation observations and
numerical weather predictions are available for the month of July from 1985
through 2012. In Figure~\ref{fig:axams-dist} in the introduction the probabilistic
forecasts from the distributional forest, trained on 1985--2008, for July 24
in 2009--2012 have already been shown as a motivational example. The figure
shows that the model properly
depicts the point masses at zero (i.e., the probability of a dry day) and
the forecasted probability density function for the total amount of precipitation.
The four sample forecasts differ considerably in location $\mu$, scale $\sigma$, and the amount of
censoring while conforming quite well with the actual observations
from these days. While this is a nice illustrative example we are  
interested in the overall predictive performance and calibration of the
distributional fits.

To assess calibration Figure~\ref{fig:axams-qq} shows residual QQ~plots for
out-of-sample predictions (2009--2012) from the different models trained on
1985--2008. Due to the point masses at zero 100 draws from the randomized
quantile residuals \citep{Dunn+Smyth:1996} are plotted in semi-transparent
gray. Overall, the randomized quantile residuals conform quite well with the
theoretical standard normal quantile (i.e., form a straight line close to the 
diagonal), indicating that all four models are sufficiently well calibrated. 
This is also supported by the corresponding probability integral transform
(PIT, \citealt{Gneiting+Balabdaoui+Raftery:2007}) histograms in
Appendix~\ref{app:pit}.

<<echo=FALSE, results=hide>>=
#### cross validation rain
if(file.exists("crps_cross.rda")){
  load("crps_cross.rda")
} else {

  nrep_cross <- 10
  seed <- 7

  res_cross <- mclapply(1:nrep_cross,
                        function(i){

                          set.seed(seed*i)

                          # randomly split data in 7 parts each including 4 years
                          years <- 1985:2012
                          testyears <- list()
                          for(j in 1:7){
                            testyears[[j]] <- sample(years, 4, replace = FALSE)
                            years <- years[!(years %in% testyears[[j]])]
                          }

                          #crps <- matrix(nrow = 7, ncol = 7)
                          reslist <- list()
                          for(k in 1:7){
                            test <- testyears[[k]]
                            train <- c(1985:2012)[!c(1985:2012) %in% test]

                            res <- evalmodels(station = "Axams",
                                              train = train,
                                              test = test,
                                              gamboost_cvr = TRUE)

                            #crps[k,] <- res$crps
                            reslist[[k]] <- res
                          }

                          #colnames(crps) <- names(res$crps)
                          return(reslist)
                        },
                        mc.cores = detectCores() - 1
  )

  # extract CRPS
  crps_cross <- matrix(nrow = nrep_cross, ncol = 7)
  # loop over all repetitions
  for(i in 1:length(res_cross)){
    #loop over all 7 folds (for 7 methods)
    crps_cross_int <- matrix(nrow = length(res_cross[[1]]), ncol = 7)
    for(j in 1:length(res_cross[[1]])){
      crps_cross_int[j,] <- res_cross[[i]][[j]]$crps
    }
    crps_cross[i,] <- colMeans(crps_cross_int, na.rm = TRUE)
  }
  colnames(crps_cross) <- names(res_cross[[1]][[1]]$crps)

  save(crps_cross, file = "crps_cross.rda")
  save(res_cross, file = "res_cross.rda")

}
@

\begin{figure}[t!]
\centering
<<rain_cross_axams_crps_skill_score, fig=TRUE, echo=FALSE, height=4.5, width = 7>>=
par(mar = c(2.5,2,1,2))
boxplot(1 - crps_cross[,c(2,3,4)] / crps_cross[,6], ylim = c(-0.005, 0.065),
        names = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS"),
        ylab = "CRPS skill score", col = "lightgray") 
abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:axams-crps}CRPS skill score from the 10 times 
7-fold cross validation at station Axams (1985--2012). The horizontal orange 
line pertains to the reference model EMOS.}
\end{figure}

To assess the predictive performance, a full cross-validation is carried out
rather than relying on just the one fixed test set for the years 2009--2012.
\hl{To do so, a 10 times 7-fold cross-validation is carried out where each
splits the available 28~years into 7~subsets of 4~randomly selected and 
therefore mostly not subsequent years.}
The models are learned on 6~folds (= 24~years) and evaluated on the 7-th fold
(= 4~years) using the average CRPS across all observations.
The resulting 10~CRPS skill scores are displayed by boxplots in
Figure~\ref{fig:axams-crps} using EMOS as the reference model (horizontal
line at a CRPSS of 0). Both GAMLSS models and the distributional forest 
perform distinctly better than the EMOS model. While the two GAMLSS lead to an improvement 
of around 4~percent, the distributional forest has a slightly higher improvement
of around 5.5~percent in median.

\begin{figure}[t!]
\centering
\setkeys{Gin}{width=0.7\textwidth}
<<rain_axams_varim, fig=TRUE, echo=FALSE, width=6.5, height=4.2>>=
par(mar = c(2.5,10,1,2))
barplot(sort(vimp_crps, decreasing = FALSE)[(length(vimp_crps)-9):length(vimp_crps)], 
        horiz = TRUE, las = 1, axes = FALSE,
        xlab = "Variable importance: mean decrease in CRPS",
        font.axis = 3, #list(family="HersheySerif", face=3),
        names.arg = gsub("pow", "", names(sort(vimp_crps, decreasing = FALSE)[(length(vimp_crps)-9):length(vimp_crps)])))
axis(1, at = seq(0,1.6,0.2), las = 1, mgp=c(0,1,0))
@
\caption{\label{fig:axams-varimp} CRPS-based variable importance for the top 10 covariates 
in the distributional forest. Based on data for station Axams, learning period 
1985--2008 and assessed in 2009--2012.}
\end{figure}

Finally, it is of interest how this improvement in predictive performance
by the distributional forest is accomplished, i.e., which of the 80~covariates
are selected in the trees of the forest. As the 100~trees of the forest
do not allow to simply assess the variables' role graphically, a common
solution for random forests in general is to consider variable importance
measures. Here, this is defined as the amount of change in CRPS when
the association between one covariate and the response variable is
artificially broken through permutation (and thus also breaking the
association to the remaining covariates).

Figure~\ref{fig:axams-varimp} shows the 10~covariates with the highest permutation
importance (i.e., change in CRPS) for station Axams. As expected the NWP outputs
for total precipitation (\emph{tp}) are particularly important along with total
column-integrated condensate (\emph{tcolc}). Also, both variables occur
in various transformations such as means (either of the full day or certain parts
of the afternoon), spreads, or minima/maxima. Thus, while the covariates
themselves are not surprising, selecting a GAMLSS with a particular combination
of all the transformations would be much more challenging.

%\pagebreak

\subsection{Application for all stations}
\label{sec:all}

<<echo = FALSE>>=
#### prediction over all stations 24 - 4
if(file.exists("crps_24to4_all.rda")){
  load("crps_24to4_all.rda")
} else {

  data("StationsTyrol")
  stations <- StationsTyrol$name
  test <- 2009:2012
  train <- 1985:2008

  res_24to4_all <- mclapply(1:length(stations),
                            function(i){

                              set.seed(7)

                              res <- evalmodels(station = stations[i],
                                                train = train,
                                                test = test,
                                                gamboost_cvr = TRUE)

                              return(res)
                            },
                            mc.cores = detectCores() - 1
  )

  # extract crps
  crps_24to4_all <- matrix(nrow = length(stations), ncol = 7)
  # loop over all stations
  for(i in 1:length(stations)){
    crps_24to4_all[i,] <- res_24to4_all[[i]]$crps
  }

  colnames(crps_24to4_all) <- names(res_24to4_all[[1]]$crps)
  rownames(crps_24to4_all) <- stations

  save(crps_24to4_all, file = "crps_24to4_all.rda")
  save(res_24to4_all, file = "res_24to4_all.rda")

}

# skill score
s <- 1 - crps_24to4_all[, 2:4]/crps_24to4_all[,6] 
colnames(s) <- c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS")

## prepare data for map which shows where distforest performed better than gamlss or gamboostLSS based on the crps

crps_map <- crps_24to4_all[,c("distforest", "gamlss", "gamboostLSS", "emos_log")]

# best method
bst <- apply(crps_map, 1, which.min)

# distance of forest to best other method
dst <- crps_map[,1] - crps_map[cbind(1:nrow(crps_map), apply(crps_map[, -1], 1, which.min) + 1)]

# breaks/groups
brk <- c(-0.1, -0.05, -0.005, 0.005, 0.05, 0.1)
#brk <- c(-0.1, -0.05, -0.01, 0.01, 0.05, 0.1)
grp <- cut(dst, breaks = brk)

# HCL colors (relatively flashy, essentially CARTO Tropic)
clr <- colorspace::diverge_hcl(5, h = c(195, 325), c = 80, l = c(50, 90), power = 1.3)

library("raster") # dem (digital elevation model)
library("sp")     # gadm www.gadm.org/country

#load("plot_map_rain/data/tirol.rda")
#load("plot_map_rain/data/dem.rda")

data(StationsTyrol, package = "RainTyrol")
data(MapTyrol, package = "RainTyrol")
# data(MapTyrol_border, package = "RainTyrol")
# Create SpatialPointsDataFrame from station list
sp <- SpatialPointsDataFrame(subset(StationsTyrol,
                                    select=c(lon,lat)),
                             data = subset(StationsTyrol,
                                           select = -c(lon,lat)),
                             proj4string = crs(MapTyrol$RasterLayer))
@

\begin{figure}[p!]
\centering
<<rain_all_crps_skill_score, fig=TRUE, echo=FALSE, height = 4.5, width = 7>>=
par(mar = c(3.1, 4.1, 1.1, 2.1))
matplot(t(s[,]), type = "l", lwd = 2, 
          col = gray(0.5, alpha = 0.2),
          lty = 1, axes = FALSE, 
          xlab = "", ylab = "CRPS skill score", xlim = c(0.5, 3.5))
lines(s[70,], col = pal[2], type = "o", pch = 19, lwd = 2)  
# Station Axams is the 77th station which is the 70th in the list of complete stations
  boxplot(s, add = TRUE, col = "transparent")
  abline(h = 0, col = pal[5], lwd = 2)

@
\caption{\label{fig:all-crps}CRPS skill score for each station (gray lines
with boxplots superimposed). Station Axams is highlighted in green 
and the horizontal orange line pertains to the reference model EMOS. The models are 
learned on 1985--2008 and validated for 2009--2012.}

\vspace*{0.4cm}

\setkeys{Gin}{width=0.99\textwidth}
<<map, fig=TRUE, echo=FALSE, height=6.1, width=9>>=

  ## plot map of Tyrol with all 95 observations
  layout(cbind(1, 2), width = c(9, 1))
  par(mar = c(5,4,4,0.1))
  raster::image(MapTyrol$RasterLayer, col = rev(gray.colors(100)),
                main="Stations in Tyrol", ylab = "Latitude", xlab = "Longitude", 
                xlim = c(9.8,13.2), 
                ylim = c(46.6, 47.87))
  plot(MapTyrol$SpatialPolygons, add = TRUE)
  points(sp[70,], pch = 19, col = "black", cex = 1.85)
  points(sp, pch = c(21, 24, 25, 22)[bst], bg = clr[grp], col = "black", las = 1, cex = 1.5)
  legend(x = 9.8, y = 47.815, pch = c(21, 24, 25, 22), legend = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS", "EMOS"), cex = 1, bty = "n")
  text(x = 10.3, y = 47.82, labels = "Models with lowest CRPS")
  mtext("CRPS\ndifference", side=4, las = TRUE, at = c(x = 13.5, y = 47.76), line = 0.3)
  par(mar = c(0.5,0.2,0.5,2.3))
  ## legend
  plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
       xlim = c(0, 1), ylim = c(-0.2, 0.2), xaxs = "i", yaxs = "i")
  rect(0, brk[-6], 0.5, brk[-1], col = rev(clr))
  axis(4, at = brk, las = 1, mgp=c(0,-0.5,-1))

@
\caption{\label{fig:all-map}Map of Tyrol coding the best-performing model
for each station (type of symbol). The color codes whether the distributional 
forest had higher (green) or lower (red) CRPS compared to the best of the 
other three models. The background shows the local topography \citep{Robinson:2014}.
Station Axams is highlighted in bold.}
\end{figure}

After considering only one observational site up to now, a second step
evaluates and compares the competing methods on all 95 available stations. As in the previous
section, all models are learned on the first 24~years and evaluated by the
average CRPS on the last 4~years. More specifically, the CRPS skill score
against the EMOS model is computed for the out-of-sample predictions at each
station and visualized by parallel coordinates plots with boxplots superimposed
in Figure~\ref{fig:all-crps}. Overall, distributional forests have a slightly
higher improvement in CRPSS compared to the two GAMLSS which is best seen
by looking at the boxplots and the green line representing the results
for station Axams. The underlying parallel coordinates additionally
bring out that the prespecified GAMLSS sometimes performs rather differently
(sometimes better, sometimes worse) compared to the two data-driven models.
Values below zero show that, for some stations, EMOS performs better
than the more complex statistical methods.
%NAs in gamlss in 13 stations!!!\\
%for example in 13th station in complete{\_}stations\\
%\text{Error in glim.fit(f = mu.object, X = mu.X, y = y, w = w, fv = mu, os = mu.offset,  :}\\ 
%\text{NA's in the working vector or weights for parameter mu}
%\\
% sum(is.na(crps[,3]))
%<<eval = TRUE, echo = FALSE>>=

To assess whether these differences in predictive performance are due to differences
in the topography, Figure~\ref{fig:all-map} shows a brief spatial summary of all
stations. Each station is illustrated by a symbol that conveys which
model performed best in terms of CRPS on the last 4~years of the data.
Additionally, the color of the symbol indicates the CRPS difference between 
distributional forest and the best-performing other model. Green signals that
the distributional forest performs better than the other models whereas red
signals that another model performs better. Overall the distributional forest
performs on par (gray) or better (green) for the majority of stations. Only
for a few stations in the north-east EMOS performs best, and in East Tyrol
the prespecified GAMLSS performs particularly well.

\section{Discussion}
\label{sec:discussion}

Distributional regression modeling is combined with tree-based modeling
to obtain a novel and flexible method for probabilistic forecasting.
The resulting distributional trees and forests can capture abrupt and
nonlinear effects and interactions in a data-driven way. By basing the
split point and split variable selection on a full likelihood and
corresponding score function, the trees and forests can not only pick
up changes in the location but also the scale or shape of any distributional
family.

Distributional forests are an attractive alternative when prespecifying or
boosting all possible effects and interactions in a GAMLSS model is
challenging.  Distributional forests are rather straightforward to specify
requiring only little prior subject matter knowledge and also work well in the
presence of many potential covariates. The application to precipitation
forecasting in complex terrain illustrates that distributional forests often
perform on par or even better than their GAMLSS counterparts. Hence, they form
a useful addition to the already available toolbox of probabilistic forecasts
for disciplines such as meteorology.

\section*{Computational details}

The proposed methods are in the \textsf{R} package \textbf{disttree}
(version~\Sexpr{packageVersion("disttree")}) based on the \textbf{partykit}
package (version~\Sexpr{packageVersion("partykit")}), both available 
on \textsf{R}-Forge at
(\url{https://R-Forge.R-project.org/projects/partykit/}). The function
\code{distfit} fits distributional models by maximum likelihood, which is
used as the basis for the tree-building function \code{disttree}, upon which
the \code{distforest} is built. All functions can either be used with
GAMLSS family objects from the \textsf{R} package \textbf{gamlss.dist} 
(\citealt{Stasinopoulos+Rigby:2007}, version~\Sexpr{packageVersion("gamlss.dist")})
or with custom lists containing all required information about the distribution family.

In addition to \textbf{disttree}, Section~\ref{sec:precipitation} employs
\textsf{R} package \textbf{crch} 
\citep[version~\Sexpr{packageVersion("crch")}]{Messner+Mayr+Zeileis:2016} for the EMOS models,
\textbf{gamlss}
\citep[version~\Sexpr{packageVersion("gamlss")}]{Stasinopoulos+Rigby:2007} for the prespecified GAMLSS, and
\textbf{gamboostLSS}
\citep[version~\Sexpr{packageVersion("gamboostLSS")}]{Hofner+Mayr+Schmid:2016} for the boosted GAMLSS.

The fitted distributional forest for July 24 and
observation station Axams (including Figure~\ref{fig:axams-dist})
is reproducible using 
\code{demo("RainAxams", package = "disttree")}.
This also includes fitting the other zero-censored
Gaussian models considered in this paper and generating the corresponding 
QQ~plots (Figure~\ref{fig:axams-qq}) and
PIT histograms (Figure~\ref{fig:axams-pit}).
Full replication of all results can be obtained with
\code{demo("RainTyrol", package = "disttree")}
requiring the companion \textsf{R} package \textbf{RainTyrol} (version~\Sexpr{packageVersion("RainTyrol")}),
also available within the \textsf{R}-Forge~project.

\bibliography{ref.bib}

\newpage
\begin{appendix}

\section{Tree algorithm}
\label{app:tree}

In the following, the tree algorithm applied in the empirical case study 
discussed in this paper is explained. For notational simplicity, 
the testing and splitting procedure is described for the root node, i.e.,
the entire learning sample with observations $\{y_i\}_{i = 1,\ldots,n}$, $n \in \mathbb{N}$.
In each child node the corresponding subsample depends on 
the foregoing split(s).

After fitting a distributional model $\mathcal{D}(Y, \bm{\theta})$ to 
the learning sample with observations $\{y_i\}_{i = 1,\ldots,n}$ as explained in 
Section~\ref{sec:distfit} the resulting estimated parameter 
$\bm{\hat{\theta}} = 
(\hat{\theta}_1, \ldots, \hat{\theta}_k)$, $k \in \mathbb{N}$ 
can be plugged in the score function $s(\bm{\theta}, Y)$.
In that way a goodness-of-fit measurement is obtained for each
parameter $\theta_j$ and each observation $y_i$.
To use this information, statistical tests are employed to detect
dependencies between the score values
$$
s(\bm{\hat{\theta}}, y) = 
%\begin{pmatrix} 
%s(\bm{\hat{\theta}}, y_1) \\
%s(\bm{\hat{\theta}}, y_2) \\
%\vdots \\
%s(\bm{\hat{\theta}}, y_n)
%\end{pmatrix} =
\begin{pmatrix} 
s(\bm{\hat{\theta}}, y_1)_1 & s(\bm{\hat{\theta}}, y_1)_2 & \ldots & s(\bm{\hat{\theta}}, y_1)_k\\
\vdots & \vdots & \ddots & \vdots \\
s(\bm{\hat{\theta}}, y_n)_1 & s(\bm{\hat{\theta}}, y_n)_2 & \ldots & s(\bm{\hat{\theta}}, y_n)_k
\end{pmatrix}
$$
and each variable $Z_l \in \{Z_1, \ldots, Z_m\}$.  
\begin{align*}
H_0^l:  s(\bm{\hat{\theta}}, Y) \qquad \bot \qquad Z_l 
\end{align*}

These hypotheses are assessed with permutation tests based on
the multivariate linear statistic
$$
T_l = vec\left(\sum_{i=1}^n v_l(Z_{li}) \cdot s(\bm{\hat{\theta}}, Y_i)\right)
$$
\hl{where $s(\bm{\hat{\theta}}, Y_i) \in \mathbb{R}^{1\times k}$ and the 
type of the transformation function $v_l$ depends on the type of the split 
variable $Z_l$. If $Z_l$ is numeric then $v_l$ is simply the identity 
function $v_l(Z_{li}) = Z_{li}$ and therefore $T_l \in \mathbb{R}^k$ as 
the ``vec'' operator converts the $1 \times k$ matrix into a $k$ column vector. 
If $Z_l$ is a categorical variable with $H$ categories then 
$v_l(Z_{li}) = e_H(Z_{li}) = (\I(Z_{li} = 1), \ldots, \I(Z_{li} = H))$ 
such that $v_l$ is a $H$-dimensional unit vector where the element corresponding to 
the value of $Z_{li}$ is $1$. In this case the statistic $T_l \in \mathbb{R}^{H \cdot k}$
as the ``vec'' operator converts the $H \times k$ matrix into a $k$ column vector 
by column-wise combination.} Observations with missing values are excluded from the sums.

With the conditional expectation $\mu_l$ and the covariance 
$\Sigma_l$ of $T_l$ as derived by 
\cite{Strasser+Weber:1999} the test statistic can be standardized. 
\hl{The observed multivariate linear statistic $t_l$ which is either an
$k$~-~or~$k \cdot H$~-~dimensional vector, depending on the scale of Y,
is mapped onto the real line by a univariate test statistic $c$.
In the application of this paper a quadratic form is chosen, such that}
$$
c_{\text{quad}}(t_l,\mu_l,\Sigma_l) = (t_l-\mu_l)\Sigma_l^+(t_l-\mu_l)^{\top}
$$
\hl{where $\Sigma_l^+$ is the Moore-Penrose inverse of $\Sigma_l$.} Alternatively,
the maximum of the absolute values of the standardized linear statistic
can be considered.

The smaller the $p$-value corresponding to the standardized test 
statistic $c(t_{l},\mu_{l},\Sigma_{l})$ is the stronger the discrepancy 
from the assumption of independence between the scores and the split 
variable $Z_l$.
%
After Bonferroni-adjusting the $p$-values it has to be assessed whether
any of the resulting $p$-values is beneath the selected significance level. 
If so, the partitioning variable $Z_{l^\ast}$ with the lowest $p$-value 
is chosen as splitting variable. \hl{Otherwise no further split is made is in this node.}

The breakpoint that leads to the highest discrepancy between score functions 
in the two resulting subgroups is selected as split point. 
This is measured by the linear statistic
$$
T_{l^{\ast}}^{qr} = \sum_{i \in \mathcal{B}_{qr}} s(\bm{\hat{\theta}}, Y_i)
$$
\hl{for $q \in \{1,2\}$ where $\mathcal{B}_{1r}$ and $\mathcal{B}_{2r}$ are the 
two new subgroups, without any particular ordering, that are defined by 
splitting in split point $r$ of variable $Z_{l^{\ast}}$.
The split point is then chosen as follows:}
$$
r^{\ast} = \argmin_{r} (\min_{q=1,2}(c(t_{l^{\ast}}^{qr},\mu_{l^{\ast}}^{qr},\Sigma_{l^{\ast}}^{qr})).
$$
One repeats the testing and splitting procedure in each of the resulting 
subgroups until some stopping criterion is reached. This criterion 
can for example be a minimal number of observations in a node or a 
minimal $p$-value for the statistical tests. In that way pre-pruning
is applied in order to find right-sized trees and hence 
avoid overfitting.

This permutation test based tree algorithm is presented in 
\cite{Hothorn+Hornik+Zeileis:2006} as the CTree algorithm. 
A different framework to build a tree is provided by the MOB algorithm 
which is based on M-fluctuation tests (\citealt{Zeileis+Hothorn+Hornik:2008}).

\section{PIT histograms}
\label{app:pit}

\begin{figure}[b!]
\centering
%\begin{subfigure}{0.8\textwidth}
<<rain_axams_pit_oosample, fig=TRUE, echo=FALSE, width = 7>>=

# out of sample
set.seed(4)

par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))
pithist(pit_df, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "Distributional forest", ylim = c(0,1.5))
pithist(pit_ml, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "EMOS", ylim = c(0,1.5))
pithist(pit_g, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "Prespecified GAMLSS", ylim = c(0,1.5))
pithist(pit_gb, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "Boosted GAMLSS", ylim = c(0,1.5))

@
%\end{subfigure}
\caption{\label{fig:axams-pit}Out-of-sample PIT histograms (2009--2012) for station Axams and models learned on data from 1985--2008.}
\end{figure}

As an alternative to QQ~plots based on randomized quantile residuals,
probability integral transform (PIT, \citealt{Gneiting+Balabdaoui+Raftery:2007}) 
histograms are a commonly used tool to assess how well-calibrated distributional 
fits are. Figure~\ref{fig:axams-pit} shows out-of-sample PIT histograms for the same
probabilistic forecasts as Figure~\ref{fig:axams-qq}, confirming that despite
some deviations the overall calibration of all models is acceptable.

\end{appendix}

\end{document}