### 2018-09-09 pkg/disttree/inst/draft/supplement/supplement_distributions.Rnw
first draft of supplements
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2277 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: ca029d6448db189b0c0d78097537cf73806205c0

\documentclass[nojss]{jss}

\usepackage{graphicx, natbib}
\pagestyle{empty}
\usepackage{Sweave}
\usepackage{amstext,amsfonts,amsmath,bm,thumbpdf,lmodern}

\title{Supplement 1: Testing different distributions}

\author{Lisa Schlosser\\Universit\"at Innsbruck
   \And Torsten Hothorn\\Universit\"at Z\"urich
   \And Reto Stauffer\\Universit\"at Innsbruck
   \And Achim Zeileis\\Universit\"at Innsbruck}
\Plainauthor{Lisa Schlosser, Torsten Hothorn, Reto Stauffer, Achim Zeileis}

\begin{document}
\SweaveOpts{concordance=TRUE}

\section{Introduction}
This application study is supplementary material to the paper ''Distributional 
Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain''. 
In Section~3 the presented distributional forest model is employed to obtain 
probabilistic precipitation forecasts in a mountainous region assuming that the 
response variable (the power-transformed daily precipitation sum) follows a 
left-censored Gaussian distribution with censoring point~$0$.

To assess how appropriate this distributional assumption is we test the robustness
of the results with respect to two questions: (1)~Are the tails of the distribution
captured appropriately by a Gaussian distribution or are \emph{heavier tails} required?
(2)~Is the point mass at zero (= no rain) driven by the same determinants as the
positive observations (= rain) or should this be captured by a \emph{separate parameter}?
For (1) we compare the censored Gaussian with a \emph{censored logistic} distribution and
for (2) a Gaussian \emph{hurdle model} (or two-part model) is employed instead of the
censored Gaussian.

The distribution can not only be exchanged for the distributional forest 
but for all the models evaluated in this application:

\begin{itemize}

\item \emph{Distributional forest:} All predictor variables are
  considered for learning a forest where bootstrap
  sampling is employed for each tree. Parameters are estimated by
  adaptive local likelihood based on forest weights.

\item \emph{EMOS:} Standard ensemble model output statistics
  models use the ensemble mean of total precipitation as regressor in the location
  submodel and the corresponding ensemble standard deviation in the scale submodel.
  The parameters are estimated by maximum likelihood, using an identity
  link for the location part and a log link for the scale part (following the advice of
  \cite{Gebetsberger+Messner+Mayr:2017}).

\item \emph{Prespecified GAMLSS:} Smooth additive splines are selected for
  the most relevant predictors based on meteorological expert knowledge
  following \cite{Stauffer+Umlauf+Messner:2017}. The model is estimated by maximum
  penalized likelihood using a backfitting algorithm (\cite{Stasinopoulos+Rigby:2007}).

\item \emph{Boosted GAMLSS:} Smooth additive splines are selected
  automatically from all available variables, using non-cyclic boosting
  for parameter estimation (\cite{Hofner+Mayr+Schmid:2016, Messner+Mayr+Zeileis:2017}).
  This updates the predictor terms for the location or scale submodels iteratively
  by maximizing the log-likelihood only for the variable yielding the
  biggest improvement. The iteration stops early -- before fully maximizing
  the in-sample likelihood -- based on a (computationally intensive)
  out-of-bag bootstrap estimate of the log-likelihood.

\end{itemize}

\section{Distributions applied / Distributional approaches}

The distribution applied in the paper is based on a Gaussian distribution 
for which the probability density function $f$ and the distribution 
function $F$ are defined by
\begin{align}
f_{\textit{Gaussian}}(Y; \mu, \sigma) &= 
     \frac{1}{\sigma} \cdot \phi\left(\frac{Y-\mu}{\sigma}\right) \\
F_{\textit{Gaussian}}(Y; \mu, \sigma) &=
    \Phi\left(\frac{Y-\mu}{\sigma}\right)
\end{align}
where $\phi$ and $\Phi$ are the probability density function and the 
distribution function of the standard normal distribution $\mathcal{N}(0,1)$.

Alternatively one could also choose a distribution with heavier tails 
such as a logistic distribution with corresponding probability density 
function and distribution function

\begin{align}
f_{\textit{logistic}}(Y; \mu, \sigma) &= 
     \frac{\exp{\left(-\frac{Y - \mu}{\sigma}\right)}}
     {\sigma \cdot \left(1 + \exp{\left(-\frac{Y - \mu}{\sigma}\right)}\right)^2} \\
F_{\textit{logistic}}(Y; \mu, \sigma) &=
     \frac{1}{1+ \exp{\left(\frac{\mu}{\sigma}\right)}}.
\end{align}

Regardless of the chosen distribution it has to be considered 
that precipitation data usually includes a point mass at zero.
In order to deal with this point mass one of two common strategies 
can be used.
\begin{itemize}
\item Censored distribution (\cite{Winkelmann:2006}): 
In general, censoring data usually leads 
to a point mass at the censoring point which has to be considered in 
the modeling process. Therefore, censored distributions are applied as 
they are able to capture this point mass. 
In precipitation data, observations can not be negative, thus, left-censoring
it at zero does not change the data. However, there is a naturally given
point mass at zero (days without rain) making a censored distribution an 
appropriate choice to model precipitation data.
For a chosen distribution with probability density function $f$ and
distribution function $F$ the log-likelihood function of the corresponding 
censored distribution is
\begin{equation}
\ell_{\textit{cens}}(\mu, \sigma; Y) = 
\begin{cases}
    \log\left\{f(Y; \mu, \sigma) \right\}, & \text{if } Y > 0 \\[0.2cm]
    \log\left\{F(0; \mu, \sigma) \right\}, & \text{if } Y = 0
\end{cases}
\end{equation}
as seen in Equation~2.1.1. of the paper for a Gaussian distribution.

\item Hurdle model / Two-part model (\cite{Cragg:1971}, \cite{Mullahy:1986}): 
An additional parameter $\nu$ describing the probability that there is 
any precipitation at all, 
i.e. $\nu = \mathbb{P}(Y>0)$, is estimated by a binary model. 
Then, a second model is fit to the subset of the data containing 
only positive observations applying a left-truncated at zero 
distribution to estimate the location parameter $\mu$ and scale 
parameter $\sigma$ of the underlying latent variable.
The three-parametric log-likelihood function for the full hurdle model 
with parameter vector $\bm{\theta} = (\mu, \sigma, \nu)$ is
\begin{equation} \label{eq:hurdle}
\ell_{\textit{hurdle}}(\mu, \sigma, \nu; Y) = 
\begin{cases}
    \log\left\{\nu \cdot \frac{f(Y; \mu, \sigma)}{1 - F(0; \mu, \sigma)} \right\}, & \text{if } Y > 0 \\[0.2cm]
    \log(1 - \nu), & \text{if } Y = 0.
\end{cases}
\end{equation}

\end{itemize}

In the paper a left-censored at zero Gaussian distribution is chosen.
Here, we additionally apply a left-censored at zero logistic distribution
and a hurdle model based on a Gaussian distribution.

In the log-likelihood function in \ref{eq:hurdle} properties of the
logarithmic function can be applied to split the term for $Y>0$ 
into two additive terms separating the parameters $\mu$ and $\sigma$
from the additional parameter $\nu$. Therefore, $\nu$ can be modeled 
by a separate model in the GAMLSS framework. This approach is also
chosen for the distributional forest model in this application study.

The models are compared based on the continuous ranked probability 
score (CRPS). For the censored models the closed form following 
\cite{Jordan:2018} is used while a numerical approximation is 
calculated for the hurdle models.

\section{Stations}
Out of the 95 observation stations 18 are considered for this 
benchmark study (see Figure~\ref{fig:map}). 
They have been selected based on the region
they are in, their altitude and the performance of the 
different models in the application in the paper such that
all of these features are represented equally.

\begin{figure}[h]
\vspace*{-3cm}
<<map, fig=TRUE, echo=FALSE>>=
##  map
data("StationsTyrol", package = "RainTyrol")
data("MapTyrol", package = "RainTyrol")
library("sp")
sp <- SpatialPointsDataFrame(subset(StationsTyrol, select = c(lon, lat)),
                             data = subset(StationsTyrol, select = -c(lon, lat)),
                             proj4string = raster::crs(MapTyrol$RasterLayer))

plot(MapTyrol$SpatialPolygons)
points(sp, pch = 21, col = "black", las = 1, cex = 1)
points(sp[c(5, 6, 16, 20, 23, 25, 32, 33, 46, 47, 48, 56, 57, 70, 71, 79, 82, 83),], pch = 21, bg = hcl(325, 100, 70), cex = 1.5)
@
\vspace*{-3cm}
\caption{\label{fig:map}Map of Tyrol with all 95 observation stations and the 18 considered stations highlighted in pink.}
\end{figure}

\section{Results}

\begin{figure}[h]
\setkeys{Gin}{width=0.99\textwidth}
<<results_lattice, fig=TRUE, echo=FALSE, results=hide, height=10, width=8>>=
library("lattice")
##### 
# HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)

load("Rain_distributions.rda")
means <- results$means[results$means$method != "disttree",]
means$method <- factor(means$method, 
                       levels = c("distforest",
                                  "gamlss",
                                  "gamboostLSS",
                                  "EMOS"),
                       labels = c("Distributional forest",
                                  "Prespecified GAMLSS",
                                  "Boosted GAMLSS",
                                  "EMOS"))

strip.background.settings <- trellis.par.get("strip.background")
strip.background.settings$col <- "gray"
trellis.par.set("strip.background", strip.background.settings)

strip.border.settings <- trellis.par.get("strip.border")
strip.border.settings$col <- "black"
trellis.par.set("strip.border", strip.border.settings)

superpose.line.settings <- trellis.par.get("superpose.line")
superpose.line.settings$col <- hcl(c(128, 260, 290, 50), 100, 50)
trellis.par.set("superpose.line", superpose.line.settings)

superpose.symbol.settings <- trellis.par.get("superpose.symbol")
superpose.symbol.settings$pch <- c(16,24,25,15)
superpose.symbol.settings$col <- hcl(c(128, 260, 290, 50), 100, 50)
superpose.symbol.settings$fill <- hcl(c(128, 260, 290, 50), 100, 50)
trellis.par.set("superpose.symbol", superpose.symbol.settings)

xyplot(crps ~ distribution | station, groups = ~ method, 
       data = means, auto.key = TRUE, 
       type = "o", lwd = 2, lty = 1, layout = c(3,6))
@
\caption{\label{fig:lattice}CRPS values of all 4 methods, once applied with each of the three distributions for each of the 18 observation stations.}
\end{figure}

\begin{figure}[h]
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crps, fig=TRUE, echo=FALSE, height=3, width=7>>=

  # crps
  par(mfrow = c(1,3))
  boxplot(crps~method, data = means, subset = distribution == "gaussian", main = "gaussian",
          ylim = c(0.65,1.15), axes = FALSE, ylab = "CRPS")
  axis(2, seq(0.6, 1.2, 0.1), seq(0.6, 1.2, 0.1), las = 2)
  axis(1, 0:5, c("",
                 "Distr.
                 forest",
                 "Presp.
                 GAMLSS",
                 "Boosted
                 GAMLSS",
                 "EMOS", ""), las=2)
  axis(3, 0:5, lwd.ticks = 0, labels = FALSE)
  axis(4, seq(0.6, 1.2, 0.1), lwd.ticks = 0, labels = FALSE)

  boxplot(crps~method, data = means, subset = distribution == "hgaussian", main = "hgaussian",
          ylim = c(0.65,1.15), las = 2, axes = FALSE)
  axis(2, seq(0.6, 1.2, 0.1), seq(0.6, 1.2, 0.1), las = 2)
  axis(1, 0:5, c("",
                 "Distr.
                 forest",
                 "Presp.
                 GAMLSS",
                 "Boosted
                 GAMLSS",
                 "EMOS", ""), las=2)
  axis(3, 0:5, lwd.ticks = 0, labels = FALSE)
  axis(4, seq(0.6, 1.2, 0.1), lwd.ticks = 0, labels = FALSE)
  boxplot(crps~method, data = means, subset = distribution == "logistic", main = "logistic",
          ylim = c(0.65,1.15), las = 2, axes = FALSE)
  axis(2, seq(0.6, 1.2, 0.1), seq(0.6, 1.2, 0.1), las = 2)
  axis(1, 0:5, c("",
                 "Distr.
                 forest",
                 "Presp.
                 GAMLSS",
                 "Boosted
                 GAMLSS",
                 "EMOS", ""), las=2)
  axis(3, 0:5, lwd.ticks = 0, labels = FALSE)
  axis(4, seq(0.6, 1.2, 0.1), lwd.ticks = 0, labels = FALSE)

@
\caption{\label{fig:box_crps}(Mean) CRPS values of each of the 18 observation stations, separated by method and distribution.}
\end{figure}

\begin{figure}[h]
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crpss_EMOS, fig=TRUE, echo=FALSE, height=3, width=7>>=
  # crps skill score by method (reference: EMOS)
  means$crps_ss_method <- means$crps
  means$crps_ss_method[means$method == "Distributional forest"] <- 
    1 - means$crps[means$method == "Distributional forest"] / means$crps[means$method == "EMOS"]
  means$crps_ss_method[means$method == "Prespecified GAMLSS"] <- 
    1 - means$crps[means$method == "Prespecified GAMLSS"] / means$crps[means$method == "EMOS"]
  means$crps_ss_method[means$method == "Boosted GAMLSS"] <- 
    1 - means$crps[means$method == "Boosted GAMLSS"] / means$crps[means$method == "EMOS"]
  means$crps_ss_method[means$method == "EMOS"] <- 
    1 - means$crps[means$method == "EMOS"] / means$crps[means$method == "EMOS"]

  means_sel <- means[means$method %in% c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS"),]
  means_sel$method <- factor(means_sel$method, levels(means_sel$method)[c(1:3)])

  par(mfrow = c(1,3))
  boxplot(crps_ss_method~method, data = means_sel, 
          subset = (distribution == "gaussian"), main = "gaussian",
          ylim = c(-0.12,0.23), axes = FALSE, ylab = "CRPS skill score")
  axis(2, seq(-0.15, 0.25, 0.05), c(seq(-0.15, 0, 0.05), seq(0.05, 0.25, 0.05)), las = 2)
  axis(1, 0:4, c("",
                 "Distr.
                 forest",
                 "Presp.
                 GAMLSS",
                 "Boosted
                 GAMLSS", ""), las=2)
  axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
  axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_method~method, data = means_sel, 
          subset = (distribution == "logistic"), main = "logistic",
          ylim = c(-0.12,0.23), axes = FALSE)
  axis(2, seq(-0.15, 0.25, 0.05), c(seq(-0.15, 0, 0.05), seq(0.05, 0.25, 0.05)), las = 2)
  axis(1, 0:4, c("",
                 "Distr.
                 forest",
                 "Presp.
                 GAMLSS",
                 "Boosted
                 GAMLSS", ""), las=2)
  axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
  axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_method~method, data = means_sel, 
          subset = (distribution == "hgaussian"), main = "hgaussian",
          ylim = c(-0.12,0.23), axes = FALSE)
  axis(2, seq(-0.15, 0.25, 0.05), c(seq(-0.15, 0, 0.05), seq(0.05, 0.25, 0.05)), las = 2)
  axis(1, 0:4, c("",
                 "Distr.
                 forest",
                 "Presp.
                 GAMLSS",
                 "Boosted
                 GAMLSS", ""), las=2)
  axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
  axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)
  abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:box_crpss_EMOS}(Mean) CRPS skill score values of each of the 18 observation stations, separated by method and distribution. EMOS is chosen as reference model.}
\end{figure}

\begin{figure}[h]
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crpss_gaussian, fig=TRUE, echo=FALSE, height=3, width=7>>=
  # crps skill score by distribution (reference: gaussian)
  means$crps_ss_dist <- means$crps
  means$crps_ss_dist[means$distribution == "gaussian"] <- 
    1 - means$crps[means$distribution == "gaussian"] / means$crps[means$distribution == "gaussian"]
  means$crps_ss_dist[means$distribution == "logistic"] <- 
    1 - means$crps[means$distribution == "logistic"] / means$crps[means$distribution == "gaussian"]
  means$crps_ss_dist[means$distribution == "hgaussian"] <- 
    1 - means$crps[means$distribution == "hgaussian"] / means$crps[means$distribution == "gaussian"]

  means_sel <- means[means$distribution %in% c("logistic", "hgaussian"),]
  means_sel$distribution <- factor(means_sel$distribution, levels(means_sel$distribution)[c(2:3)])

  par(mfrow = c(1,4))
  boxplot(crps_ss_dist~distribution, data = means_sel, 
          subset = (method == "Distributional forest"), main = "Distributional forest",
          ylim = c(-0.12,0.07), las = 2, ylab = "CRPS skill score")
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_dist~distribution, data = means_sel, 
          subset = (method == "Prespecified GAMLSS"), main = "Prespecified GAMLSS",
          ylim = c(-0.12,0.07), las = 2)
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_dist~distribution, data = means_sel, 
          subset = (method == "Boosted GAMLSS"), main = "Boosted GAMLSS",
          ylim = c(-0.12,0.07), las = 2)
  abline(h = 0, col = pal[5], lwd = 2)
  boxplot(crps_ss_dist~distribution, data = means_sel, 
          subset = (method == "EMOS"), main = "EMOS",
          ylim = c(-0.12,0.07), las = 2)
  abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:box_crpss_gaussian}(Mean) CRPS skill score values of each of the 18 observation stations, separated by method and distribution. Gaussian is chosen as reference distribution.}
\end{figure}

\section{Discussion}
As explained above the the parameter $\nu$ is modeled separately from $\mu$ and $\sigma$ 
in the GAMLSS framework which is why two separate models can be built. For the forest model
an alternative approach would be to build just one model employing the three-parametric
distribution family. However, testing both versions of the forest model in this application study shows that they both perform almost equally. Over the 18 considered stations the 
absolute difference in terms of CRPS between the two-forest model and the three-parametric forest is on average 0.0078 and lies between 0.0014 and 0.0267 for each station. Comparing
these values to the mean CRPS values in Figure~\ref{fig:box_crps} lying between 0.6 and 1.2
confirms the relatively small difference between the two versions of the forest model.

<<compare_forests, results=hide, echo=FALSE>>=
  load("crps_forests.rda")
  sum(crps_forests[,"forest_2p"] > crps_forests[,"forest_h"])
  summary(crps_forests[,"forest_h"] - crps_forests[,"forest_2p"])
  summary(abs(crps_forests[,"forest_h"] - crps_forests[,"forest_2p"]))
  colMeans(crps_forests)  
@

\bibliography{ref_supplement_distributions.bib}

\end{document}