### 2018-09-09 pkg/disttree/inst/draft/supplement/supplement_stationwise.Rnw

\documentclass[a4paper,11pt]{article}

\usepackage{Sweave}
\usepackage{amstext,amsfonts,amsmath,bm,thumbpdf,lmodern}

\begin{document}
\SweaveOpts{concordance=TRUE}

\title{Supplement 2: Stationwise Evaluation}
\maketitle

\section{Introduction}
This application study is supplementary material to the paper ''Distributional 
Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain''. 
In Section~3 the presented distributional forest model is employed to obtain 
probabilistic precipitation forecasts in a mountainous region.
Next to an evaluation over all 95 observation stations further investigations
including a cross-validation have been performed for one selected station, Axams.
Here, this more extensive evaluation is carried out for 18 observation stations.

\section{Models and evaluation}
As in the paper a distributional forest applying a left-censored at zero 
Gaussian distribution is evaluated together with three other zero-censored 
Gaussian models: (a)~a standard basic EMOS, (b)~a GAMLSS with prespecified 
effects and interactions based on meteorological knowledge/experience, and 
(c)~a boosted GAMLSS with automatic selection of smooth additive terms based 
on all 80~predictor variables.

First, they are learned on data from 24 years (1985--2008) and evaluated 
on the following 4 years (2009--2012). 
To compare the performance of the
models the continuous ranked probability score (CRPS) is calculated
for each observation in the testing data set and than averaged.
Moreover, for each model it is timed, how long the fitting process takes 
as well as how long it takes them to make one single prediction 
(on average over all testing observations).

For the forest model variable importance is analyzed and the top 10
variables are illustrated. All four models are also compared
visually by residual Q-Q-plots and PIT-histograms.

Second, all four models are evaluated in a 10 times 7-fold cross-validation.
Again, their performances are compared based on the average CRPS values 
and the fitting time is meassured for all models as well as the prediction
time. however, contrary to the first model, the prediction time now refers 
to the time needed on average to make predictions for the whole testing data, 
not just for one observation.
For the cross-validation the CRPS skill score is illustrated in a boxplot 
where the EMOS model is chosen as reference model, represented by the horizontal
line at 0.

\section{Stations}
Out of the 95 observation stations 18 are considered for this 
benchmark study. They have been selected based on the region
they are in, their altitude and the performance of the 
different models in the application in the paper such that
all of these features are represented equally.

\vspace*{-3cm}

\begin{figure}[h]
<<map_stationwise, fig=TRUE, echo=FALSE>>=
##  map
data("StationsTyrol", package = "RainTyrol")
data("MapTyrol", package = "RainTyrol")
library("sp")
sp <- SpatialPointsDataFrame(subset(StationsTyrol, select = c(lon, lat)),
                             data = subset(StationsTyrol, select = -c(lon, lat)),
                             proj4string = raster::crs(MapTyrol$RasterLayer))

plot(MapTyrol$SpatialPolygons)
points(sp, pch = 21, col = "black", las = 1, cex = 1)
points(sp[c(5, 6, 16, 20, 23, 25, 32, 33, 46, 47, 48, 56, 57, 70, 71, 79, 82, 83),], pch = 21, bg = hcl(325, 100, 70), cex = 1.5)
@
\vspace*{-3cm}
\caption{\label{fig:map}Map of Tyrol with all 95 observation stations and the 18 considered stations highlighted in pink.}
\end{figure}

\section{Discussion}
Looking at prediction times for one single observation, the distributional forest is on average
the slowest model, however, the difference is not too big. This changes regarding predictions
for the a set of observation due to the fact that the forest makes predictions based
on local neighbour weights. Therefor, for each observation its own weight matrix has to be included in the prediction process making it a more time consuming procedure.

%\bibliography{ref_supplement_stationwise.bib}

\end{document}