### 2018-09-21 pkg/disttree/inst/draft/supplement/supplement_distributions.Rnw
revised versions of the two supplements
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2288 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 4fd840560f1fff3def8ab5426de9972e78214d72

\documentclass[nojss, shortnames]{jss}

\usepackage{graphicx, natbib}
\pagestyle{empty}
\usepackage{Sweave}
\usepackage{amstext,amsfonts,amsmath,bm,thumbpdf,lmodern}

\title{Supplement 1: Testing different distributions}

\author{Lisa Schlosser\\Universit\"at Innsbruck
\And Torsten Hothorn\\Universit\"at Z\"urich
\And Reto Stauffer\\Universit\"at Innsbruck
\And Achim Zeileis\\Universit\"at Innsbruck}
\Plainauthor{Lisa Schlosser, Torsten Hothorn, Reto Stauffer, Achim Zeileis}

\Abstract{In the application discussed in the paper ''Distributional 
Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain''
the underlying distribution is assumed to be a left-censored at zero Gaussian distribution.
In this supplement the exact same evaluations as in the paper are performed employing
other distribution assumptions. In particular, a left-censored at zero logistic distribution 
and a hurdle-model are applied and their results are compared to those of the left-censored 
at zero Gaussian distribution presented in the paper.}

\Keywords{random forests, GAMLSS, hurdle model, logistic, Gaussian}

\Address{
Lisa Schlosser, Reto Stauffer, Achim Zeileis \\
Universit\"at Innsbruck \\
Department of Statistics \\
Faculty of Economics and Statistics \\
Universit\"atsstr.~15 \\
6020 Innsbruck, Austria \\
E-mail: \email{Lisa.Schlosser@uibk.ac.at}, \email{Reto.Stauffer@uibk.ac.at},\\
\email{Achim.Zeileis@R-project.org} \\
URL: \url{https://www.uibk.ac.at/statistics/personal/schlosser-lisa/},\\
\phantom{URL: }\url{https://retostauffer.org/},\\
\phantom{URL: }\url{https://eeecon.uibk.ac.at/~zeileis/}\\

Torsten Hothorn\\
Universit\"at Z\"urich \\
Institut f\"ur Epidemiologie, Biostatistik und Pr\"avention \\
Hirschengraben 84\\
CH-8001 Z\"urich, Switzerland \\
E-mail: \email{Torsten.Hothorn@R-project.org}\\
URL: \url{http://user.math.uzh.ch/hothorn/}\\
}

\begin{document}
\SweaveOpts{concordance=TRUE}

\section{Introduction}
This application study is supplementary material to the application presented
in Section~3 of the paper.
There, the presented distributional forest model is employed to obtain 
probabilistic precipitation forecasts in a mountainous region assuming that the 
response variable (the power-transformed daily precipitation sum) follows a 
left-censored Gaussian distribution with censoring point~$0$.

To assess how appropriate this distributional assumption is we test the robustness
of the results with respect to two questions: (1)~Are the tails of the distribution
captured appropriately by a Gaussian distribution or are \emph{heavier tails} required?
(2)~Is the point mass at zero (= no rain) driven by the same determinants as the
positive observations (= rain) or should this be captured by a \emph{separate parameter}?
For (1) we compare the censored Gaussian with a \emph{censored logistic} distribution and
for (2) a Gaussian \emph{hurdle model} (or two-part model) is employed instead of the
censored Gaussian.

To compare the performance of the distributional forest to other commonly used
methods the exact same models as in the paper are evaluated on the same data set 
and applying the same settings while only the distribution is exchanged. Further 
details on the models can be found in Section~3.2. of the paper. In particular, 
Table~2 gives an overview over the models and their specifications. As a brief 
reminder the approaches are summarized below.

\begin{itemize}

\item \emph{Distributional forest:} All predictor variables are considered for 
learning a forest where subsampling is employed for each tree. Parameters are 
estimated by adaptive local likelihood based on forest weights.

\item \emph{EMOS} \citep{Gneiting+Raftery+Westveld:2005}: 
Standard ensemble model output statistics models use the ensemble 
mean of total precipitation as regressor in the location submodel 
and the corresponding ensemble standard deviation in the scale submodel.
The parameters are estimated by maximum likelihood.

\item \emph{Prespecified GAMLSS:} Smooth additive splines are selected for
the most relevant predictors based on meteorological expert knowledge
following \cite{Stauffer+Umlauf+Messner:2017}. The model is estimated by maximum
penalized likelihood using a backfitting algorithm (\cite{Stasinopoulos+Rigby:2007}).

\item \emph{Boosted GAMLSS:} Smooth additive splines are selected
automatically from all available variables, using non-cyclic boosting
for parameter estimation (\cite{Hofner+Mayr+Schmid:2016, Messner+Mayr+Zeileis:2017}).
The optimal stopping point for the iteration is evaluated based on 
a (computationally intensive) out-of-bag bootstrap estimate of the log-likelihood.

\end{itemize}

\section{Distributions applied / Distributional approaches}

The distribution applied in the paper is based on a Gaussian distribution 
for which the probability density function $f$ and the distribution 
function $F$ are defined by
%
\begin{align}
f_{\mathrm{Gaussian}}(Y; \mu, \sigma) &= 
\frac{1}{\sigma} \cdot \phi\left(\frac{Y-\mu}{\sigma}\right) \\
F_{\mathrm{Gaussian}}(Y; \mu, \sigma) &=
\Phi\left(\frac{Y-\mu}{\sigma}\right)
\end{align}
where $\phi$ and $\Phi$ are the probability density function and the 
distribution function of the standard normal distribution $\mathcal{N}(0,1)$.

Alternatively one could also choose a distribution with heavier tails 
such as a logistic distribution with corresponding probability density 
function and distribution function
%
\begin{align}
f_{\mathrm{logistic}}(Y; \mu, \sigma) &= 
\frac{\exp{\left(-\frac{Y - \mu}{\sigma}\right)}}
{\sigma \cdot \left(1 + \exp{\left(-\frac{Y - \mu}{\sigma}\right)}\right)^2} \\
F_{\mathrm{logistic}}(Y; \mu, \sigma) &=
\frac{1}{1+ \exp{\left(-\frac{Y - \mu}{\sigma}\right)}}.
\end{align}
%
Both distributions have to be adapted in our application to capture that
precipitation is non-negative with a point mass at zero (= no rain).
Two common strategies for this are to employ either \emph{left-censoring at zero}
or a \emph{hurdle} (or two-part) approach, combining a binary (zero vs.\ greater) and
a truncated (positive) part \citep[see][for an overview]{Winkelmann:2006}.
\begin{itemize}
\item Censored distribution \citep{Winkelmann:2006}: 
%In general, censoring data usually leads 
%to a point mass at the censoring point which has to be considered in 
%the modeling process. Therefore, censored distributions are applied as 
%they are able to capture this point mass. 
%In precipitation data, observations can not be negative, thus, left-censoring
%it at zero does not change the data. However, there is a naturally given
%point mass at zero (days without rain) making a censored distribution an 
%appropriate choice to model precipitation data.
For a chosen distribution with probability density function $f$ and
distribution function $F$ the log-likelihood function of the corresponding 
censored distribution is
%
\begin{equation}
\ell_{\mathrm{cens}}(\mu, \sigma; Y) = 
\begin{cases}
\log\left\{f(Y; \mu, \sigma) \right\}, & \text{if } Y > 0 \\[0.2cm]
\log\left\{F(0; \mu, \sigma) \right\}, & \text{if } Y = 0
\end{cases}
\end{equation}
%
as seen in Equation~2.1.1. of the paper for a Gaussian distribution.
For positive values of $Y$ the censored version of the density function
equals the original non-censored version. For $Y=0$ the censored density
function takes the values of the probability mass lying left of the 
censoring point. Therefore, the parameters $\mu$ and $\sigma$ control
both parts, the point mass and the positive part of the distribution.

\item Hurdle model / Two-part model (\cite{Cragg:1971}, \cite{Mullahy:1986}): 
An additional parameter $\nu$ describing the probability that there is 
any precipitation at all, i.e. $\nu = \mathbb{P}(Y>0)$, is included in the model.
The three-parametric log-likelihood function for the full hurdle model 
with parameter vector $\bm{\theta} = (\mu, \sigma, \nu)$ is
\begin{equation} \label{eq:hurdle}
\ell_{\mathrm{hurdle}}(\mu, \sigma, \nu; Y) = 
\begin{cases}
\log\left\{\nu \cdot \frac{f(Y; \mu, \sigma)}{1 - F(0; \mu, \sigma)} \right\}, & \text{if } Y > 0 \\[0.2cm]
\log(1 - \nu), & \text{if } Y = 0.
\end{cases}
\end{equation}
For $Y=0$ the probability that it won't rain is expressed by $1-\nu$. 
For $Y>0$ the probability that it will rain $\nu$ is multiplied with the 
left-truncated at zero density function. Contrary to the censored model, 
here it is only the parameter $\nu$ controlling the point mass while $\mu$ 
and $\sigma$ only influence the distribution of the positive observations.

\end{itemize}

In the paper a left-censored at zero Gaussian distribution is chosen.
Here, we additionally apply a left-censored at zero logistic distribution
and a hurdle model based on a Gaussian distribution.

In the log-likelihood function in Equation~\ref{eq:hurdle} properties of the
logarithmic function can be applied to split the term for $Y>0$ 
into two additive terms separating the parameters $\mu$ and $\sigma$
from the additional parameter $\nu$. Therefore, $\nu$ can be modeled 
by a separate model in the GAMLSS framework. The same approach is
chosen for the distributional forest model in this application study.

The models are compared based on the continuous ranked probability 
score (CRPS). For the censored models the closed form following 
\cite{Jordan:2018} is used while a numerical approximation is 
calculated for the hurdle models.

\section{Stations}
Out of the 95 observation stations 15 are considered for this benchmark 
study (see Figure~\ref{fig:map}). 
They have been selected based on the region they are in, their altitude 
and the performance of the different models in the application in the 
paper such that all of these features are represented equally.

Selected stations: Axams, Lech, Zuers, See im Paznaun, Jungholz, Ladis-Neuegg, 
Oetz, Ochsengarten-Obergut, Ginzling, Rotholz, Walchsee, Koessen, 
Innervillgraten, Matrei in Osttirol, St.Johann im Walde.

\begin{figure}[h]
\vspace*{-3cm}
\setkeys{Gin}{width=0.99\textwidth}
<<map_stationwise, fig=TRUE, echo=FALSE>>=
##  map
data("StationsTyrol", package = "RainTyrol")
data("MapTyrol", package = "RainTyrol")
library("sp")
sp <- SpatialPointsDataFrame(subset(StationsTyrol, select = c(lon, lat)),
                             data = subset(StationsTyrol, select = -c(lon, lat)),
                             proj4string = raster::crs(MapTyrol$RasterLayer))

plot(MapTyrol$SpatialPolygons)
points(sp, pch = 21, col = "darkgrey", las = 1, cex = 0.6)
points(sp[c(5, 6, 20, 23, 25, 32, 33, 46, 47, 56, 57, 70, 79, 82, 83),], 
       pch = 21, bg = hcl(325, 100, 70), cex = 1.5)
# stationname beneath
text(sp[c(25, 47, 57, 70),], 
     labels = StationsTyrol$name[c(25, 47, 57, 70)],
     pos=1, cex=0.75)
# stationname left
text(sp[c(20, 32, 46, 79, 83),], 
     labels = StationsTyrol$name[c(20, 32, 46, 79, 83)],
     pos=2, cex=0.75)
# stationname above
text(sp[c(6, 33, 82),], 
     labels = StationsTyrol$name[c(6, 33, 82)],
     pos=3, cex=0.75)
# stationname right
text(sp[c(5, 23, 56),], 
     labels = StationsTyrol$name[c(5, 23, 56)],
     pos=4, cex=0.75)
@
\vspace*{-3.5cm}
\caption{\label{fig:map}Map of Tyrol with all 95 observation stations and 
the 15 considered stations highlighted in pink.}
\end{figure}

\section{Results}
In Figure~\ref{fig:lattice} each panel belongs to one of the 15 observation stations. 
The three distributions \textit{gaussian} (left-censored at zero Gaussian distribution), 
\textit{hgaussian} (hurdle model based on the Gaussian distribution) and 
\textit{logistic} (left-censored at zero logistic distribution) are compared based 
on the resulting CRPS values for each of the four applied methods that are represented 
by different colors and symbols. 
Overall, the forest model performs almost equally well for all three distributions. 
Therefore it seems to be very robust against missspecifications of the distribution.
For the two GAMLSS models the censored Gaussian and the hurdle model perform similarly 
at most of the stations while applying the censored logistic distribution leads to an 
increase in CRPS, i.e. to a decline in performance, especially for the prespecified 
GAMLSS (for example at stations Zuers and Jungholz).
For the EMOS model exchanging the distribution does not have a strong effect on its 
performance, similar to the forest model.
All in all, exchanging the underlying distribution does not lead to clear 
improvements of any of the methods, especially the censored Gaussian and the 
hurdle model achieve almost the same results. However, the robustness against 
missspecification of the distribution varies slightly over stations. For example 
at station Matrei in Osttirol all methods seem to be very stable while stronger 
influences can be seen at station Jungholz.

\textit{Note:} When applying a hurdle model for the stations Oetz and 
Innvervillgraten the prespecified GAMLSS computationally failed to build/fit 
a model. Therefore this method is not represented in the corresponding panels 
of Figure~\ref{fig:lattice} for these two cases.

In Figure~\ref{fig:box_crpss_gaussian} the tested distributions are compared 
for each method based on CRPS skill score with the censored Gaussian distribution 
as reference represented by the horizontal line at zero. Again, the forest model 
and the EMOS show only slight variations (on average not even 1\%) over different 
distributions while the two GAMLSS models perform even or slightly better applying 
a hurdle model and loose strength when employing a logistic distribution compared 
to a Gaussian distribution.

\begin{figure}
\setkeys{Gin}{width=0.99\textwidth}
<<results_lattice, fig=TRUE, echo=FALSE, results=hide, height=10, width=8>>=
library("lattice")
##### 
# HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)

load("Rain_distributions.rda")
stations <- c("Axams", "Lech", "Zuers", "See im Paznaun", "Jungholz", 
              "Ladis-Neuegg", "Oetz", "Ochsengarten-Obergut",
              "Ginzling", "Rotholz", "Walchsee", "Koessen", 
              "Innervillgraten", "Matrei in Osttirol", 
              "St.Johann im Walde")

means <- results$means[(results$means$method != "disttree") & 
                         results$means$station %in% stations, ]
rownames(means) <- c(1:NROW(means))
means$station <- factor(means$station, levels = stations)
means$method <- factor(means$method, 
                       levels = c("distforest",
                                  "gamlss",
                                  "gamboostLSS",
                                  "EMOS"),
                       labels = c("Distributional forest",
                                  "Prespecified GAMLSS",
                                  "Boosted GAMLSS",
                                  "EMOS"))

strip.background.settings <- trellis.par.get("strip.background")
strip.background.settings$col <- "gray"
trellis.par.set("strip.background", strip.background.settings)

strip.border.settings <- trellis.par.get("strip.border")
strip.border.settings$col <- "black"
trellis.par.set("strip.border", strip.border.settings)

superpose.line.settings <- trellis.par.get("superpose.line")
superpose.line.settings$col <- hcl(c(128, 260, 290, 50), 100, 50)
trellis.par.set("superpose.line", superpose.line.settings)

superpose.symbol.settings <- trellis.par.get("superpose.symbol")
superpose.symbol.settings$pch <- c(16,24,25,15)
superpose.symbol.settings$col <- hcl(c(128, 260, 290, 50), 100, 50)
superpose.symbol.settings$fill <- hcl(c(128, 260, 290, 50), 100, 50)
trellis.par.set("superpose.symbol", superpose.symbol.settings)

xyplot(crps ~ distribution | station, groups = ~ method, 
       data = means, auto.key = TRUE, 
       type = "o", lwd = 2, lty = 1,  
       drop.unused.levels = TRUE, layout = c(3,5),
       as.table = TRUE, ylab = "CRPS", xlab = "Distribution")
@
\caption{\label{fig:lattice}CRPS values of all four methods, once applied with 
each of the three distributions for each of the 15 observation stations.}
\end{figure}

\begin{figure}
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crpss_gaussian, fig=TRUE, echo=FALSE, height=3, width=7>>=
# crps skill score by distribution (reference: gaussian)
means$crps_ss_dist <- means$crps
means$crps_ss_dist[means$distribution == "gaussian"] <- 
  1 - means$crps[means$distribution == "gaussian"] / means$crps[means$distribution == "gaussian"]
means$crps_ss_dist[means$distribution == "logistic"] <- 
  1 - means$crps[means$distribution == "logistic"] / means$crps[means$distribution == "gaussian"]
means$crps_ss_dist[means$distribution == "hgaussian"] <- 
  1 - means$crps[means$distribution == "hgaussian"] / means$crps[means$distribution == "gaussian"]

means_sel <- means[means$distribution %in% c("logistic", "hgaussian"),]
means_sel$distribution <- factor(means_sel$distribution, 
                                 levels(means_sel$distribution)[c(2:3)])

# CRPS skill score for distributional forest
s_df <- matrix(ncol = 2, nrow = 15)
s_df[,1] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "hgaussian", "crps_ss_dist"]
s_df[,2] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "logistic", "crps_ss_dist"]

# CRPS skill score for prespecified GAMLSS
s_g <- matrix(ncol = 2, nrow = 15)
s_g[,1] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "hgaussian", "crps_ss_dist"]
s_g[,2] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "logistic", "crps_ss_dist"]

# CRPS skill score for boosted GAMLSS
s_gb <- matrix(ncol = 2, nrow = 15)
s_gb[,1] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "hgaussian", "crps_ss_dist"]
s_gb[,2] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "logistic", "crps_ss_dist"]

# CRPS skill score for EMOS
s_emos <- matrix(ncol = 2, nrow = 15)
s_emos[,1] <- means_sel[means_sel$method == "EMOS" & means_sel$distribution == "hgaussian", "crps_ss_dist"]
s_emos[,2] <- means_sel[means_sel$method == "EMOS" & means_sel$distribution == "logistic", "crps_ss_dist"]

colnames(s_df) <- colnames(s_g) <- colnames(s_gb) <- colnames(s_emos) <- c("hgaussian", "logistic")

# boxplots with matplot
par(mfrow = c(1,4), mar = c(4,4,3,0.7))
matplot(t(s_df[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "Distributional forest",
        xlab = "", ylab = "CRPS skill score", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_df, add = TRUE, col = "transparent")
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Distributional forest"), main = "Distributional forest",
#        ylim = c(-0.12,0.07), las = 2, ylab = "CRPS skill score")
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,2.9,3,1.8))
matplot(t(s_g[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE, main = "Prespecified GAMLSS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_g, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Prespecified GAMLSS"), main = "Prespecified GAMLSS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,1.8,3,2.9))
matplot(t(s_gb[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE, main = "Boosted GAMLSS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_gb, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "Boosted GAMLSS"), main = "Boosted GAMLSS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
par(mar = c(4,0.7,3,4))
matplot(t(s_emos[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "EMOS",
        xlab = "", ylab = "", xlim = c(0.5, 2.5),
        ylim = c(-0.10, 0.05))
boxplot(s_emos, add = TRUE, col = "transparent", yaxt = 'n')
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_dist~distribution, data = means_sel, 
#        subset = (method == "EMOS"), main = "EMOS",
#        ylim = c(-0.12,0.07), las = 2)
#abline(h = 0, col = pal[5], lwd = 2)
@
\caption{\label{fig:box_crpss_gaussian} CRPS skill score values of each of 
the 15 observation stations, separated by method and distribution. Left-censored 
at zero Gaussian is chosen as reference distribution.}
\end{figure}

Figures~\ref{fig:box_crpss_EMOS} and \ref{fig:box_crps} compare the methods separately 
for each distribution, once based on CRPS skill score values with reference model EMOS 
and once based on CRPS values. Both confirm the conclusion that the censored Gaussian 
model and the hurdle model lead to almost the same results for all methods while only 
the logistic model shows small variations.

Altogether it can be stated that the left-censored at zero Gaussian distribution seems 
to be an appropriate choice while the other two distributions, especially the hurdle model, 
are also reasonable alternatives. However, they do not lead to an overall improvement in 
performance for this application.

\begin{figure}
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crpss_EMOS, fig=TRUE, echo=FALSE, height=3, width=7>>=
# crps skill score by method (reference: EMOS)
means$crps_ss_method <- means$crps
means$crps_ss_method[means$method == "Distributional forest"] <- 
  1 - means$crps[means$method == "Distributional forest"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "Prespecified GAMLSS"] <- 
  1 - means$crps[means$method == "Prespecified GAMLSS"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "Boosted GAMLSS"] <- 
  1 - means$crps[means$method == "Boosted GAMLSS"] / means$crps[means$method == "EMOS"]
means$crps_ss_method[means$method == "EMOS"] <- 
  1 - means$crps[means$method == "EMOS"] / means$crps[means$method == "EMOS"]

means_sel <- means[means$method %in% c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS"),]
means_sel$method <- factor(means_sel$method, levels(means_sel$method)[c(1:3)])

# CRPS skill score for gaussian
s_gaussian <- matrix(ncol = 3, nrow = 15)
s_gaussian[,1] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "gaussian", "crps_ss_method"]
s_gaussian[,2] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "gaussian", "crps_ss_method"]
s_gaussian[,3] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "gaussian", "crps_ss_method"]

# CRPS skill score for boosted hgaussian
s_hgaussian <- matrix(ncol = 3, nrow = 15)
s_hgaussian[,1] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "hgaussian", "crps_ss_method"]
s_hgaussian[,2] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "hgaussian", "crps_ss_method"]
s_hgaussian[,3] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "hgaussian", "crps_ss_method"]

# CRPS skill score for prespecified logistic
s_logistic <- matrix(ncol = 3, nrow = 15)
s_logistic[,1] <- means_sel[means_sel$method == "Distributional forest" & means_sel$distribution == "logistic", "crps_ss_method"]
s_logistic[,2] <- means_sel[means_sel$method == "Prespecified GAMLSS" & means_sel$distribution == "logistic", "crps_ss_method"]
s_logistic[,3] <- means_sel[means_sel$method == "Boosted GAMLSS" & means_sel$distribution == "logistic", "crps_ss_method"]

colnames(s_gaussian) <- colnames(s_hgaussian) <- colnames(s_logistic) <- c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS")

par(mfrow = c(1,3), mar = c(5,4,3,0))
matplot(t(s_gaussian[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "gaussian", ylim = c(-0.12,0.23), 
        xlab = "", ylab = "CRPS skill score", xlim = c(0.5, 3.5))
boxplot(s_gaussian, add = TRUE, col = "transparent", axes = FALSE)
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_method~method, data = means_sel, 
#        subset = (distribution == "gaussian"), main = "gaussian",
#        ylim = c(-0.12,0.23), axes = FALSE, ylab = "CRPS skill score")
axis(1, 0:4, c("",
               "Distr.
                 forest",
               "Presp.
                 GAMLSS",
               "Boosted
                 GAMLSS", ""), las=2)
axis(2, seq(-0.15, 0.25, 0.05), c(seq(-0.15, 0, 0.05), seq(0.05, 0.25, 0.05)), las = 0)
axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)

par(mar = c(5,2,3,2))
matplot(t(s_hgaussian[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "hgaussian", ylim = c(-0.12,0.23), 
        xlab = "", ylab = "", xlim = c(0.5, 3.5))
boxplot(s_hgaussian, add = TRUE, col = "transparent", axes = FALSE)
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_method~method, data = means_sel, 
#        subset = (distribution == "hgaussian"), main = "hgaussian",
#        ylim = c(-0.12,0.23), axes = FALSE)
axis(2, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)
axis(1, 0:4, c("",
               "Distr.
                 forest",
               "Presp.
                 GAMLSS",
               "Boosted
                 GAMLSS", ""), las=2)
axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)

par(mar = c(5,0,3,4))
matplot(t(s_logistic[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "logistic", ylim = c(-0.12,0.23), 
        xlab = "", ylab = "", xlim = c(0.5, 3.5))
boxplot(s_logistic, add = TRUE, col = "transparent", axes = FALSE)
abline(h = 0, col = pal[5], lwd = 2)
#boxplot(crps_ss_method~method, data = means_sel, 
#        subset = (distribution == "logistic"), main = "logistic",
#        ylim = c(-0.12,0.23), axes = FALSE)
axis(2, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)
axis(1, 0:4, c("",
               "Distr.
                 forest",
               "Presp.
                 GAMLSS",
               "Boosted
                 GAMLSS", ""), las=2)
axis(3, 0:4, lwd.ticks = 0, labels = FALSE)
axis(4, seq(-0.15, 0.25, 0.05), lwd.ticks = 0, labels = FALSE)

@
\caption{\label{fig:box_crpss_EMOS}CRPS skill score values of each of 
the 15 observation stations, separated by method and distribution. 
EMOS is chosen as reference model.}
\end{figure}

\begin{figure}
\setkeys{Gin}{width=0.99\textwidth}
<<results_box_crps, fig=TRUE, echo=FALSE, height=3, width=7>>=

# CRPS for gaussian
c_gaussian <- matrix(ncol = 4, nrow = 15)
c_gaussian[,1] <- means[means$method == "Distributional forest" & means$distribution == "gaussian", "crps"]
c_gaussian[,2] <- means[means$method == "Prespecified GAMLSS" & means$distribution == "gaussian", "crps"]
c_gaussian[,3] <- means[means$method == "Boosted GAMLSS" & means$distribution == "gaussian", "crps"]
c_gaussian[,4] <- means[means$method == "EMOS" & means$distribution == "gaussian", "crps"]

# CRPS skill score for boosted hgaussian
c_hgaussian <- matrix(ncol = 4, nrow = 15)
c_hgaussian[,1] <- means[means$method == "Distributional forest" & means$distribution == "hgaussian", "crps"]
c_hgaussian[,2] <- means[means$method == "Prespecified GAMLSS" & means$distribution == "hgaussian", "crps"]
c_hgaussian[,3] <- means[means$method == "Boosted GAMLSS" & means$distribution == "hgaussian", "crps"]
c_hgaussian[,4] <- means[means$method == "EMOS" & means$distribution == "hgaussian", "crps"]

# CRPS skill score for prespecified logistic
c_logistic <- matrix(ncol = 4, nrow = 15)
c_logistic[,1] <- means[means$method == "Distributional forest" & means$distribution == "logistic", "crps"]
c_logistic[,2] <- means[means$method == "Prespecified GAMLSS" & means$distribution == "logistic", "crps"]
c_logistic[,3] <- means[means$method == "Boosted GAMLSS" & means$distribution == "logistic", "crps"]
c_logistic[,4] <- means[means$method == "EMOS" & means$distribution == "logistic", "crps"]

colnames(c_gaussian) <- colnames(c_hgaussian) <- colnames(c_logistic) <- c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS", "EMOS")

par(mfrow = c(1,3), mar = c(5,4,3,0))
matplot(t(c_gaussian[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "gaussian", 
        xlab = "", ylab = "CRPS", xlim = c(0.5, 4.5),
        ylim = c(0.65,1.15))
boxplot(c_gaussian, add = TRUE, col = "transparent", axes = FALSE)
#boxplot(crps~method, data = means, subset = distribution == "gaussian", main = "gaussian",
#        ylim = c(0.65,1.15), axes = FALSE, ylab = "CRPS")
axis(2, seq(0.6, 1.2, 0.1), seq(0.6, 1.2, 0.1), las = 0)
axis(1, 0:5, c("",
               "Distr.
                 forest",
               "Presp.
                 GAMLSS",
               "Boosted
                 GAMLSS",
               "EMOS", ""), las=2)
axis(3, 0:5, lwd.ticks = 0, labels = FALSE)
axis(4, seq(0.6, 1.2, 0.1), lwd.ticks = 0, labels = FALSE)

par(mar = c(5,2,3,2))
matplot(t(c_hgaussian[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "hgaussian", 
        xlab = "", ylab = "", xlim = c(0.5, 4.5),
        ylim = c(0.65,1.15))
boxplot(c_hgaussian, add = TRUE, col = "transparent", axes = FALSE)
#boxplot(crps~method, data = means, subset = distribution == "hgaussian", main = "hgaussian",
#        ylim = c(0.65,1.15), las = 2, axes = FALSE)
axis(2, seq(0.6, 1.2, 0.1), lwd.ticks = 0, labels = FALSE)
axis(1, 0:5, c("",
               "Distr.
                 forest",
               "Presp.
                 GAMLSS",
               "Boosted
                 GAMLSS",
               "EMOS", ""), las=2)
axis(3, 0:5, lwd.ticks = 0, labels = FALSE)
axis(4, seq(0.6, 1.2, 0.1), lwd.ticks = 0, labels = FALSE)

par(mar = c(5,0,3,4))
matplot(t(c_logistic[,]), type = "l", lwd = 2, 
        col = gray(0.5, alpha = 0.2),
        lty = 1, axes = FALSE,  main = "logistic", 
        xlab = "", ylab = "CRPS", xlim = c(0.5, 4.5),
        ylim = c(0.65,1.15))
boxplot(c_logistic, add = TRUE, col = "transparent", axes = FALSE)
#boxplot(crps~method, data = means, subset = distribution == "logistic", main = "logistic",
#        ylim = c(0.65,1.15), las = 2, axes = FALSE)
axis(2, seq(0.6, 1.2, 0.1), lwd.ticks = 0, labels = FALSE)
axis(1, 0:5, c("",
               "Distr.
                 forest",
               "Presp.
                 GAMLSS",
               "Boosted
                 GAMLSS",
               "EMOS", ""), las=2)
axis(3, 0:5, lwd.ticks = 0, labels = FALSE)
axis(4, seq(0.6, 1.2, 0.1), lwd.ticks = 0, labels = FALSE)

@
\caption{\label{fig:box_crps}CRPS values of each of the 15 observation stations, 
separated by method and distribution.}
\end{figure}

As explained above, the parameter $\nu$ is modeled separately from $\mu$ and $\sigma$ 
in the GAMLSS framework which is why two separate models can be built. For the forest model
an alternative approach would be to build just one model employing the three-parametric
distribution family. However, testing both versions of the forest model in this application 
study shows that they both perform almost equally. Figure~\ref{fig:box_crpss_forests} shows
the CRPS skill score of the three-parametric forest model over the considered 15 observation
stations with the two-part forest model as reference. Even though this alternative version
of the forest model seems to perform better on average, according to the median this 
improvement is not even one percent, hence neglectably small. Therefore, the two forest 
versions can be stated to perform almost equally well.

\begin{figure}
\center
\setkeys{Gin}{width=0.5\textwidth}
<<compare_forests, fig=TRUE, results=hide, echo=FALSE, width=4, height=4>>=
load("crps_forests.rda")
crps_ss_forest_2p <- data.frame(1 - crps_forests[,"forest_2p"] / crps_forests[,"forest_h"])
names(crps_ss_forest_2p) <- "Three-parametric forest"
boxplot(crps_ss_forest_2p, ylab = "CRPS skill score", names = c("Three-parametric forest"))
abline(h = 0, col = pal[5], lwd = 2)
axis(1, 0:2, c("","Three-parametric forest", ""), las=1)
@
\caption{\label{fig:box_crpss_forests}CRPS skill score values of each of 
the 15 observation stations for the three-parametric forest model with 
the two-part forest model as reference model.}
\end{figure}

\newpage
%bibliographystyle{jss}
\bibliography{ref_supplement.bib}

\end{document}