### 2018-09-30 pkg/disttree/inst/draft/rejoinder.tex
first draft of rejoinder
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2302 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 3ccf8a68000484f182ed31433f38867e2d70f113

\documentclass[american,foldmarks=false,noconfig]{uibklttr}
%\documentclass{article}
\usepackage{sfmath,hyperref,color,enumitem}

\let\code=\texttt
\let\pkg=\emph
\let\proglang=\textsf
\newenvironment{review}{\fontshape{\itdefault}\fontseries{\bfdefault} \selectfont \smallskip}{\par}

\setkomavar{signature}{Lisa Schlosser}

\makeatletter
\@setplength{refvpos}{75mm}
\makeatother
\setlength{\textheight}{25cm}
\definecolor{quotecolor}{rgb}{0.4,0.4,0.5}

\newenvironment{reto}{
    \begin{color}{uibkorange}
    \textbf{Reto:~}
        \itshape
}{
    \end{color}
}
\usepackage{amsmath}

\begin{document}
\begin{letter}{
Nicoleta Serban\\
Area Editor\\
The Annals of Applied Statistics}

\setkomavar{subject}{Revision of AOAS1804-015: Distributional 
regression forests for probabilistic precipitation forecasting 
in complex terrain}

\opening{Dear Professor Serban,}

please find attached the revised version of our manuscript 
``Distributional regression forests for probabilistic precipitation 
forecasting in complex terrain'' (AOAS1804-015).

Thank you very much for the constructive and helpful feedback from you, 
the associate editor and the two referees. The manuscript has been 
improved and additional results haven been obtained from further evaluations.

The most important changes are:
\begin{itemize}

\item The more extensive evaluations for station Axams have been 
carried out for 14 other observation stations. Results are 
provided in an online supplement which comes with our revised manuscript.

\item The evaluations performed for all 95 observation stations 
have been carried out under different distribution assumptions 
for the response variable in order to test for robustness against
distributional misspecifications. Results are provided in an 
online supplement.

\item The mathematical notation in the Appendix is now more precise.

\end{itemize}

All changes and additons are explained in much more detail in the
point-to-point reply on the next pages.

\closing{Best regards,
\vspace*{1cm}
}

\vspace*{-2.8cm}
\hspace*{-0.3cm}signature %\includegraphics{zeileis-signature}

\end{letter}

\newpage

\textbf{\LARGE Associate Editor}

\begin{review}
This manuscript proposes a framework for distributional regression 
forests which combine the idea of random forests with distributional 
regression in order to obtain a flexible class of forecast models 
where the full forecast distribution depends on the predictors. 
This new approach is demonstrated in the context of probabilistic 
precipitation forecasting where it is used to construct full 
predictive distributions based on the output of an ensemble 
weather prediction model. The paper is very well written, the 
proposed methodology is useful, and the data example used to 
illustrate this method is practically relevant and discussed 
thoroughly.
\end{review}

Thank you very much for your advice and the positive feedback, 
it is highly appreciated!

\begin{review}
A typical limitation of very flexible modeling approaches like 
the one proposed here is that a sufficiently large sample is 
required to fit the model. Can the authors provide some 
guidelines as to how large the training sample needs to be for 
this method to be effective? Does the stopping criterion 
guarantee that the model is not overfitted? 
Is it obvious how to define the specific stopping criterion and 
if not, can general guidelines for doing so be provided?
\end{review}

It is difficult to set a minimal size of the learning data for
the method to be effective, however, the distributional forest
performed quite well in different applications with at least
50 observations. 
The stopping criteria that are based on the number of 
observations in a node need to be adapted to the size of the
learning data, especially if the data set is small.
For example, in the application presented in this paper the 
learning data consists of 744 observations and each tree is built
on a subsample of around 470 observations (63.2 percent, as 
this is the mean fraction for N-out-of-N bootstrapping with 
replacement and therefore commonly used for subsampling
in forest models). 
The stopping criteria are set such that no split will be performed 
if there are either less than 50 observations 
in the considered node or if there were less than 20 observations in 
one of the resulting subgroups of the considered possible split.
(This information is now also added in Section~3.2., 'Distributional forest').
Generally, these values (and also the significance level for
the statistical tests in the tree algorithm) should be kept 
rather low for a forest model as it is desirable to grow very 
large trees. Therefore, each of the trees in the forest will 
most likely overfit the subsample it is learned on, however, this 
is evened out by combining them to a forest model (with an adequate 
number of trees). For that reason, setting the stopping criteria too 
low does not lead to a decrease in goodness of fit, however, the 
computational costs might increase, especially for large data sets. 
On the other hand, if they are too high, splits might not be detected 
such that the corresponding effects are not captured by the model.

\textit{FIX ME: too long?}

\begin{review}
The equations are not numbered but the authors refer to equations 
by number (e.g. bottom of p7)
\end{review}

We apologize for the confusion the missing equation numbers 
might have caused. They are now added.

\begin{review}
In the last section where computational details are discussed, 
can you also provide some information on the computational cost? 
In a weather forecasting contest like the one discussed here, 
forecasts are often provided on a high-resolution grid, 
so it would be important to understand if the approach 
proposed here is computationally affordable. This is not clear 
to me because the nearest neighbor weights are specific 
to each new observation, so in a forecasting context these 
computations have to be done in realtime.
\end{review}

We have measured fitting time and prediction time in our 
application. For single predictions the distributional forest 
can keep up with the other models. However, as you already 
stated correctly, each prediction made by the forest is based 
on an observation-specific weight matrix and therefore has to 
be calculated separately. For that reason prediction time for a 
set of observations is clearly higher for the forest model than 
the GAMLSS models or the EMOS. The exact results and times can 
be found in an online supplement.

\begin{review}
abstract: the acronym 'GAMLSS' should be spelled out the first 
time it is used
\end{review}

\begin{review}
p3: I'm not sure if people outside the weather forecasting 
community is familiar with the notion of an 'ensemble', so 
this should be explained briefly
\end{review}

\begin{reto}
Wo baut man das am besten ein?
\end{reto}

\begin{review}
p7: What is meant by 'domain knowledge'?
\end{review}

With `domain knowledge' we are referring to expert knowledge 
in the field of the considered application. 
If variable selection and further model specifications have 
to be made in advance this additional knowledge is required 
in order to make reasonable decisions.

\begin{reto}
Im paper entsprechend adressieren (tbd).
\end{reto}

\begin{review}
p14: Was there any particular rationale for choosing Axams 
as the station to be studied in detail
\end{review}

Axams was chosen as the observation station being closest 
to Innsbruck, the capital of Tyrol and the work place of 
three of the authors, but not based on any other particular 
features. Furthermore, it has been checked, whether it 
is an adequate representative of all stations regarding 
the results of our evaluations. This is now also 
confirmed by further analysis in the online supplement which
comes along with our manuscript.

\begin{review}
Fig. 4: Are these the plots for residuals of precipitation 
at the original or at the transformed scale?
\end{review}

These are the residuals of precipitation on the transformed scale.

\begin{review}
p17: Is there any explanation why the prespecified GAMLSS 
method is the best-performing method in East Tyrol? Any 
explanation why the distributional regression forests 
performed less well than the simple EMOS method in this area?
\end{review}

East Tyrol is located south of the main Alpine Ridge and exhibits different
climatological characteristics than the part north of the Alpine Ridge.  It
seems that the choice of covariates and their interactions we made for the
prespecified GAMLSS method is particularly adequate for the region of East
Tyrol for this specific analysis. However, the advantages of the prespecified
GAMLSS models in this particular setting (training: 1985--2008, validation:
2009--2012) even out in a cross-validation framework for most of the stations
in East Tyrol. These results are presented in the online supplement which comes
along with our manuscript showing that the forest performs equally well as the
prespecified GAMLSS and better than the EMOS at most of the stations.

\newpage

\textbf{\LARGE Reviewer 1}

\bigskip

\textbf{General comments:}

\begin{review}
The article presents a new parametric methodology, called 
distributional forests, to estimate the probability 
distribution of some random variables, based on information 
from covariates. The methodology is new in that it mixes the
automatic variable selection properties of random forests 
with a parametric assumption on the forecast distribution. 
This allows to keep the interpretability of parametric 
distributions with the flexibility of random forests in 
dealing with local fitting, non-linearity, interactions 
and variable selection.

This approach may be used in applied statistics, as 
illustrated with its application to forecasting daily 
precipitation amount in the mountainous region of Tyrol, 
Austria. The mountains make it very difficult for numerical 
weather prediction models to forecast local precipitation. 
Statistical post-processing of weather forecasts allows 
downscaling the prediction.

The paper is well written and is a pleasant reading. 
The methodology is introduced clearly and well grounded 
in statistical practices, but some points need to be 
made clearer (see later).

The article shows on the Tyrol rain forecasts that the 
proposed method performs better than three other methods, 
among which is the current state-of-the-art in statistical 
post-processing in meteorology (EMOS). Thus distributional
forests may be added to the several statistical post-processing 
techniques, as an efficient and easy-to-use parametric 
probabilistic forecast.

I recommend the paper to be published after the following 
minor revisions have been taken into account.
\end{review}

Thank you very much for the positive and constructive 
comments regarding the manuscript. These helped to 
substantially improve the manuscript.

\bigskip

\textbf{Specific comments:}

\begin{review}
1. {\color{quotecolor} page 2: exceedence}\\
This spelling is not wrong, but less common than ``exceedance''.
Please change into \textit{exceedance}
\end{review}

\begin{review}
2. {\color{quotecolor}\textit{page 2:} 10$\%$ to 90\% quantile}\\
Please change into \textit{10\% to 90\% quantiles}
\end{review}

\begin{review}
3. {\color{quotecolor}\textit{page 5:} Fig 3, caption: Observations 
are left-censored at 0}\\
Since observations are precipitation amount, they are truncated, not 
left-censored (which may let think that they could be negative). 
Please change into \textit{Observations are truncated at 0}
\end{review}

You are right, the observations can not be negative and are 
therefore left-truncated at 0 and not left-censored (we have adapted 
this in the caption of Figure~3). However, due to days without 
rain there is a naturally given point mass at zero in precipitation 
data, similar to the point mass that results from censoring 
possibly negative data.
Thus, we use the concept of \textit{censoring} which allows us to directly
account for the large fraction of observations at zero and the non-negative
part of the observations at once.
The online supplement which comes along with our manuscript includes the
results for a hurdle model where the occurrence of precipitation is addressed
using a binary model while the non-negative part is
modeled by a \textit{truncated} Gaussian distribution.

\begin{review}
4. {\color{quotecolor}\textit{page 5:} Fig 3, caption: censoring point}\\
Please change into \textit{truncation point}
\end{review}

Since we are applying a censored distribution as explained above we 
prefer using the term censoring point, even though the data itself 
is not censored.

\begin{review}
5. {\color{quotecolor}\textit{page 5:} Fig 3, legend: Point mass at 
censoring point}\\
Please change into \textit{Point mass at truncation point}
\end{review}

See points 3 and 4.

\begin{review}
6. {\color{quotecolor}\textit{page 5:} Fig 3, x label: $mm^{(1/1.6)}$/24\textit{h}}\\
This is not a very common unit, which make the figure not easily 
interpretable. Furthermore, since you have not yet said that 
observations have been power transformed, this may be surprising 
here at first reading. Please redraw this figure with 
observations in $mm$/24$h$.
\end{review}

Thank you for the advice, the figure is now redrawn on the scale $mm$/24$h$.

\begin{review}
7. {\color{quotecolor}\textit{page 7:} see Equation 2.1}\\
Your equations have no numbering. Please add one.
\end{review}

We apologize for the confusion the missing equation numbers 
might have caused. They are now added.

\begin{review}
8. {\color{quotecolor}\textit{page 8:} $\hat{\Theta}(z) = \cdots$}\\
This expression may lead to think that the parameter is estimated 
for each new prediction. If I am right, this is not the case, 
since the weights $w_i^{tree}$ are either $0$ or $1$. Thus the 
estimated parameter vector is the same as the one for the 
observations in the same segments as the new observation.
Please make this clear in the text.
\end{review}

For each new observation $z$ the corresponding 
parameter vector $\hat{\Theta}(z)$ can be calculated following 
the given formula. But you are right, one can also just evaluate 
the predicted terminal node for $z$ if the predicted parameter
vector for this node is already known as there is one unique 
parameter vector for all observations in the same terminal node.
This is now also explained at the end of Section~2.2..

\begin{review}
9. {\color{quotecolor}\textit{page 9:} Each of the distributional trees}\\
Please change into \textit{Each of the distributional tree}
\end{review}

\textit{FIX ME: ?}

\begin{review}
10. {\color{quotecolor}\textit{page 9:} $w_i^{forest}$}\\
The normalization with $|B_b^t|$ makes the weight $w_i^{forest}$ 
take their value in $[0; 1]$. If I am not wrong, this means the 
parameter vector has to be estimated for each new prediction. 
How long does it take? For operational applications, the time 
to produce post-processed forecast may be a constraint. It may 
be interesting to compare the prediction time with distributional 
forest and with GAMLSS, since you claim in the abstract that
GAMLSS are computationally demanding. Please add some information
about the computation time to train your models and produce predictions.
\end{review}

Yes, in the forest model the parameter vector has to be estimated 
separately for each new prediction as it is based on an 
observation-specific weight matrix. For that reason, making 
predictions for a set of new observations is clearly more 
time consuming for the forest model than the GAMLSS models 
or the EMOS model. However, the time needed to make one single 
prediction is similar for all methods. The exact results 
concerning fitting and prediction times can be found in an online 
supplement.

\begin{review}
11. {\color{quotecolor}\textit{page 10:} While this approach 
alone is already highly effective in the plains, it typically 
does not perform as well in complex terrain due to unresolved 
effects in the NWP system. For example, in the Tyrolean 
Alps -- considered in the following case study -- the NWP grid 
cells of $50 \times 50$ km$^2$ are too coarse to capture single 
mountains, narrow valleys, etc.}\\
Do you have references about this lack of performance of NWP 
in complex terrains? If so, please add them.
\end{review}

\textit{TO DO}

\begin{reto}
Es gibt da ein klassisches MOS-Paper in welchem Postprocessing
mit einem Globalmodell und einem Lokalmodell gemacht werden.
Das ist nicht die Referenz die der Reviewer will.
Das Problem ist, dass es explizit dazu nix gibt. Das ist einfach
jedem von uns klar. M\"uller zeigt aber in seinen Figures (f\"ur
Wind und Temperatur), dass ein Modell mit h\"oherer Aufl\"osung
einen geringeren Fehler aufweist und sagt folgendes:

``\textit{The largest differences are found in the raw
model forecast for stations in complex terrain. For wind
speed, the increase in resolution is more beneficial than
for temperature but is again negligible for MOS forecasts.
[\dots]
A higher resolution will
further reduce the model bias, especially in complex
terrain, but this can also be achieved with statistical
postprocessing if observations are available. Furthermore,
a postprocessed temperature forecast at 12-km
resolution outperforms a raw forecast at 3-km, regardless
of the complexity of the terrain}''.

\begin{verbatim}
@article{mueller2011,
    author = {M{\"u}ller, M. D.},
    title = {Effects of Model Resolution and Statistical Postprocessing on Shelter Temperature and Wind Forecasts},
    journal = {Journal of Applied Meteorology and Climatology},
    volume = {50},
    number = {8},
    pages = {1627--1636},
    year = {2011},
    doi = {10.1175/2011JAMC2615.1},
}
\end{verbatim}

Oder Meyers and Steenburgh (2013) der aufzeigt, dass auch der Mensch sich mit
der Wettervorhersage im Gebirge schwer tut bzw. dort die groessten Challenges
zu finden sind mit all den lokalen und kleinraeumigen Eigenschaften und Einfluessen.
Habe das Buch grad nicht rumliegen, ist wohl aber auch nicht zwingend notwendig.

\begin{verbatim}
@Inbook{meyers2013,
    author = {Meyers, Michael P. and Steenburgh, W. James},
    editor = {Chow, Fotini K. and De Wekker, Stephan F.J. and Snyder, Bradley J.},
    title = {Mountain Weather Prediction: Phenomenological Challenges and Forecast Methodology},
    bookTitle = {Mountain Weather Research and Forecasting: Recent Progress and Current Challenges},
    year = {2013},
    publisher = {Springer Netherlands},
    pages = {1--34},
    isbn = {978-94-007-4098-3},
    doi = {10.1007/978-94-007-4098-3_1},
}
\end{verbatim}
\end{reto}

\begin{review}
12. {\color{quotecolor}\textit{page 11:} Table 1, caption}\\
Please move the caption above the table.
\end{review}

\begin{review}
13. {\color{quotecolor}\textit{page 12:} Parameters are estimated 
by adaptive local likelihood based on the forest weights as 
described in Section 2.}\\
Please explain what is your stopping criterion.
\end{review}

As stopping criteria the minimal number of observations to perform 
a split (\texttt{minsplit}) is set to 50,  the minimal number of 
observations in a terminal node (\texttt{minbucket}) is set to 20 
and the significance level for variable selection (\texttt{alpha}) 
is kept at the default value 1. We have added this information in 
the manuscript in Section~3.2., 'Distributional forest'.

\begin{review}
14. {\color{quotecolor}\textit{page 13:} Table 2, caption}\\
Please move the caption above the table.
\end{review}

\begin{review}
15. {\color{quotecolor}\textit{page 17:} Thus, while the covariates 
themselves are not surprising, selecting a GAMLSS with a particular 
combination of all the transformations would be much more challenging.}\\
Did you compare the most important covariates in distributional forests
with the ones in preselected GAMLSS? If so, is there some insightful
information you could add?
\end{review}

Figure 6 shows the variable importance of the distributional forest
for one specific case (similar for other sites and cross-folds) 
using a CRPS based variable importance measure. The subset of 
most important covariates is similar to the set of covariates 
used for the prespecified GAMLSS but not equivalent.
The use of different variations of the same covariate (cf., Table 2)
is not surprising as the covariates are often highly correlated.
Compared to the prespecified GAMLSS the forest seems to depict some
features via differences among the different variations of a specific
covariate (e.g., use \code{tp\_max}, \code{tp\_mean}, and \code{tp\_min}).
Unfortunately the analysis does not allow to get more insights
of (the importance of) interactions between different covariates.

\begin{reto}
Oder sowas \dots Oder g\"abe es daf\"ur tests in W\"aldern?
\end{reto}

\begin{review}
16. {\color{quotecolor}\textit{page 20:} If $Z_l$ is numeric 
then $v_l$ is simply the identity function $v_l(z_l^i) = z_l^i$. 
If $Z_l$ is a categorical variable with $H$ categories then 
$v_l(z_l^i) = e_H(z_l^i) = (I(z_j^i = 1), \dots , I(z_l^i = H))$ 
such that $v_l$ is a unit vector}\\
Based on this sentence, $v_l$ can be a vector or a scalar, 
which makes not very clear what $T_l$ is: is it a vector 
or a scalar? Please clarify.
\end{review}

Apologies for the unclear notation in Appendix A. This has been 
revised in the manuscript, also regarding your comments 17--20,
and we hope that it now clarifies the content.

Depending on whether the split variable $Z_l$ is a numeric 
or a categorical variable, $t_l$ is either a
$k$~-~or~$k \cdot H$~-~dimensional vector.

\begin{review}
17. {\color{quotecolor}\textit{page 20:} the observed 
multivariate linear statistic $t$}\\
What is $t$? Is it the vector with $m$ components $T_l$? 
Please clarify.
\end{review}

Apologies again for the confusing notation. Here we are 
actually referring to the linear statistic $t_l$ (instead of $t$) 
which is the observed value of $T_l$, the test statistic for 
the $l$-th split variable $Z_l$.

\begin{review}
18. {\color{quotecolor}\textit{page 20:} $c_{quad}(t, \mu, \Sigma)$}\\
$\mu$ and $\Sigma$ have not been defined. What is their relationship 
with $\mu_l$ and $\Sigma_l$? Please clarify.
\end{review}

As in the foregoing point this was just a general notation, but 
actually the variables corresponding to the $l$-th split variable 
are considered, i.e. $t_l$, $\mu_l$ and $\sigma_l$. Thank you for
pointing this out, now the missing indices are set.

\begin{review}
19. {\color{quotecolor}\textit{page 21:} it has to be assessed 
whether any of the resulting p-values is beneath the selected 
significance level. If so, the partitioning variable $Z_l$
with the lowest p-value is chosen as splitting variable.}\\
What if no p-value is beneath the selected significance level. 
Is it a stopping criterion? Please clarify.
\end{review}

Yes, one stopping criterion is that no p-value is beneath the 
selected significance level. Therefore, no further 
split is performed in the considered node in this case.

\begin{review}
20. {\color{quotecolor}\textit{page 21:} $B_{1r}$ is the first 
of the two new subgroups that are defined by splitting in split 
point $r$ of variable $Z_{l^{\ast}}$}\\
What do you mean by ``first of the two new subgroups''? Is it 
the subgroup with values $Z_{l^{\ast}} > r$? I don't see how 
such an ordering can be achieved if $Z_{l^{\ast}}$ is an 
unordered categorical variable. Please clarify how the two
subgroups are ordered.
\end{review}

Again, we are sorry for a confusing explanation. The two subgroups 
are not ordered which is also not necessary as only the sum of 
scores over the two subgroups is considered for the linear 
statistic $T_{l^{\ast}}^{qr}$.

\newpage

\textbf{\LARGE Reviewer 2}

\begin{review}
The authors propose an extension of non-homogeneous Gaussian models 
for the use in weather forecasting based on parametric distributional 
models (GAMLSS). While the latter have proved to be applicable to a 
wide range of empirical questions, the authors focus on precipitation 
forecasts in the Alps. The approach combines a distributional model 
with a random forest, which they call distributional random forests, 
in order to include variable selection and interactions or 
non-linearities of predictors. I have a number of major concerns 
given in the following. My impression is that the method is rather
an extension of existing methodology on random forests to censored 
normal problems without important with a lack in discussing 
advantages and disadvantages of distributional regression models 
and only a rather minor improvement over certain competitors.
\end{review}

Thank you very much for your feedback and helpful suggestions. We hope 
the improvements we made on the paper can clear your doubts and concerns.

\bigskip

\textbf{Main comments:}

\begin{review}
1. It is obvious in distributional regression that not only the 
variables influence the predictive performance but rather that 
the exact distributional assumption has a big impact on both the 
fit and the prediction. The authors do not consider any other
than the censored normal distribution, which I found odd. 
In earlier references, they used the same distribution but I 
did not find any justification why not to use a truncated normal, 
a zero-adjusted model or a continuous part for $y > 0$ other
than the normal. In any distributional model, checking the 
assumption on the error density is a crucial step which in 
my opinion should not be ignored or assumed not to be a problem.
\end{review}

Thank you for pointing this out. We have now rerun the exact 
same evaluations as presented in the paper with different 
distributional assumptions, including a logistic model and 
a hurdle model with a truncated part. The exact results can be 
found in an online supplement.

\begin{review}
2. The response and some predictors are power transformations 
of originally measured variables like total precipitation as 
it is known that skewness is usually present. Why? As far as I 
understand, one major advantage of distributional models is
that the response does not have to be normal but to model scale, 
skewness and the shape in general as a function of the covariates. 
By applying the transformation in advance and hence marginally, 
the dependence of covariates on further moments of the response 
is assumed to be not existing and simply ignored.
\end{review}

The skewness results from the characteristics of precipitation sums. 
Most of the time no or only moderate precipitation amounts are forecasted 
and observed. In some situations (e.g., days with thunderstorms, cold and 
warm fronts, \dots) large amounts of precipitation can be observed at 
specific sites. This is often also depict by the ensemble prediction 
system where a subset of all members predicts rather large amounts 
(compared to the overall ensemble mean or median). This yields strongly 
skewed ensemble forecasts which has to be addressed in some sort.\\
As you mention one way would be to directly apply a
distributional family with an additional parameter to capture
skewness. Another frequently used approach is to transform the 
data to remove large parts of the skewness. The latter has the 
advantage that the model assumptions get simpler as no estimate 
for the skewness parameter is required. 
Moreover, applying a transformation together with a censored 
at zero Gaussian distribution allows for a better comparison 
to the other models as they can all be built under the same 
settings.
In this study we are thus following Stauffer et al. (2017) 
using the power transformation which turned out to be optimal 
for a similar application for the very same study area.

\begin{review}
3. Is a random forest really the right thing to do? In the 
end, the analyst or meteorologist does not obtain an 
interpretable model and no explanation which variable 
affects the distribution in which direction is possible. 
I think, the reason that focussing on predictions should 
be made clearer and well justified.
\end{review}

This is the trade-off between interpretability and predictive 
skill. Regression-type models might be easier to interpret, 
however, if one is mainly interested in predictive skill and
a good performing fully automated weather forecasting tool the 
interpretability is not implicitly needed in some cases 
(machine learning performance vs. interpretability).
We adapted our manuscript as we definitely want to emphasize that 
in our application the focus is mainly on prediction but not on 
interpretation. We hope we could point this out clearly.

\begin{review}
4. The method itself is hard to follow in particular since 
authors reference to equations without using equation numbers. 
The part on random forests for instance is quite a black box 
without looking into a number of earlier works of some of 
the authors. From an applied perspective, there should be a 
better motivation and connection of the methodology and the 
data application.
\end{review}

We apologize for the confusion the missing equation numbers 
might have caused. They are now added.

\textit{FIX ME: comment on 'there should be a 
better motivation and connection of the methodology and the 
data application'?}

\begin{review}
5. In the introduction it is said that ``while requiring no 
expert knowledge for the model specification''. I do not think 
it is true given that the authors do not consider a possible 
distributional misspecification at all. In addition, If I am 
not wrong, the functional forms in the distributional 
regression models do not have to be prespecified. Applying a 
mixed model type representation it should be possible to
have for instance interaction surfaces separated into the pure 
interaction effects and the main effects. For instance, 
Wood et al. (2013) discuss a generic approach for defining 
well-identified and interpretable effect decompositions in 
tensor product interactions again based on the mixed model 
representation of penalised model components, while Goicoa 
et al. (2017); Ugarte et al. (2017) determine the identifiability
constraints from the null space of the penalties associated 
with the spatio-temporal interaction effect.
\end{review}

Thank you for your suggestions and the references. 
The prespecification of the distribution is expert knowledge in 
this context as in some of our previous publications we have been 
working on different response distributions and the censored normal 
distribution turned out to be one of the best performing. In this article 
we have not proven whether or not the chosen distribution is the overall
best for this specific application, however, the main part of the article 
is and should be the distributional forest and not the application itself. 
However, we are sure that, given our previous work, the selected methods 
and model assumptions are well performing and well suited for the practical 
example presented.
Moreover, as mentioned before we have tested robustness against 
distributional misspecification. The results show that applying a 
left-censored at zero logistic distribution or a Gaussian hurdle model 
instead of the left-censored at zero Gaussian distribution, as presented 
in the paper, does not lead to any significant improvements in goodness of 
fit of the applied models. The exact results are provided in an online 
supplement which comes with our manuscript.

\textit{FIX ME: refer to Wood, Goicoa, Ugarte}

\begin{review}
6. A simulation study designed according to the data set should 
be used in order to give an empirical evidence that the random 
forests are generally better able to perform predictions compared 
to the competing methods. The figures indicate that advantages 
for the chosen station are rather minor and one should keep in 
mind that GAMLSS and boosting yield interpretable results which 
the random forests do not do. also, what happens when choosing 
another station than Axams? The figure 8 indicates that the 
distributional forest does not perform better in all stations 
and I am wondering whether a meteorologist would really do a 
better job on a relative scale of precipitation.
\end{review}

You are right, there are clearly some stations where other models perform
better than the forest. However, at the majority of stations used in our study
it can at least keep up with them and therefore seems to provide a good and
reasonable alternative that is very easy to apply as no expert knowledge or
additional work in advance is required and the computational effort is kept
relatively small compared to a boosting method. In an online supplement the
same evaluations as for station Axams have been carried out for 14 other
stations, including some of those where the forest does not perform better than
the other models in the setting presented in Figure~8.  It shows that the
differences in performance even out in the cross-validation framework. There,
the forest performs at least equally well compared to the best of the other
models at most stations.

\textit{FIX ME: comments on simulation study? comments on use of
method by meteorologists? (last sentence)}

\begin{reto}
``[\dots] I am wondering whether a meteorologist would really do a
better job on a relative scale of precipitation.'' ist wohl eher eine
rethorische Frage und tut nix zur Sache wenn du mich fragst.
Unkommentiert lassen? Oder Z hat vielleicht eine Idee?
\end{reto}

\begin{review}
7. How are the folds chosen for performing the predictions? 
is it 4 subsequent years or randomly chosen 4 years out of 
the 28? I expect this to have a strong impact on the predictions 
and sensitivity with respect to this needs to be considered both 
in simulations and the real study.
\end{review}

The folds of the cross-validation are chosen randomly. The data 
set of 28 years is split randomly into 7 blocks of 4 years which 
are not necessarily successive. This is now also explained in the 
description of the cross-validation setup third paragraph of 
Section~3.3..

\begin{review}
8. I found the competing models to be chosen pretty unfair. 
First, for the prespecified GAMLSS, the authors use 
”relevant effects based on expert knowledge”. Why not all? 
Why not making a decision on some information criteria such as AIC/BIC? 
The latter would be done if the analyst decided to use a prespecified 
GAMLSS. I guess, the GAMLSS could improve a lot by this . 
Second, what does a basic EMOS exactly mean? Third, why not to include 
interactions in the boosting algorithm as well? this should be 
conceptually straight forward? Fourth, why are 100 trees chosen 
in the forest? Is this a a default value? May this number not also 
be dependent on the complexity of the model? Why not to choose 
the number of trees similar to what is done in boosting? this 
would also increase the computation time of the forest a lot.
\end{review}

We decided to evaluate two different versions of a GAMLSS: 
the prespecified GAMLSS where the covariates are selected 
in advance based on expert knowledge and the boosted GAMLSS
where all covariates are included. For the latter covariates 
are selected within the boosting algorithm. As stated by 
Hofner et al. (2016) the AIC-based variable selection 
methods implemented in the \textsf{R} package \textbf{gamlss} 
``can be unstable, especially when it comes to selecting 
possibly different sets of variables for multiple 
distribution parameters.''

By a `basic EMOS' we are referring to the linear distributional 
regression model based on the idea presented by Gneiting et al. (2005) 
where only the ensemble mean is included for the location parameter 
and the ensemble standard deviation for the scale parameter.

In the boosted GAMLSS we did not include interactions as this would
lead to an extremely high number of covariates that would be hardly
manageable. Even if we included only the pairwise interactions of all 
80 covariates we would end up with 3160 additional covariates.

% factorial(80)/factorial(80-2)/factorial(2)

\begin{reto}
K\"onnte man eventuell auch im Paper so erw\"ahnen?

\begin{equation}
    \frac{n!}{k! \cdot (n-k)!}~~\mbox{bei}~n=80~\mbox{und}~k=2~\mbox{die}~3160~\mbox{Kombinationen.}
\end{equation}
\end{reto}

By default the number of trees in a distributional forest
is set to 500. However, to keep computation time at a 
reasonable level we decided  to use only 100 trees. 
But of course, the more trees, the better the complexity of 
the model can be captured, especially smooth transitions can 
be modeled better by a higher number of trees.

\textit{FIX ME: comment on 'Why not to choose the number of trees 
similar to what is done in boosting? this would also increase the 
computation time of the forest a lot.'}

\bigskip

\textbf{Minor comments:}

\begin{review}
1. Given the length of the paper, I suggest to move the 
appendix to an online supplement.
\end{review}

\begin{review}
2. Does the EMOS method employ ensemble post-processing 
techniques? There are recent references using this in weather 
forecasting (Chen et al.; 2014)?
\end{review}

\textit{FIX ME: add formula for EMOS?}

\begin{reto}
Chen ist auf jeden Fall non-sense. Es gibt viele Papers die
sowas machen, die Frage ist, ob es 10 Referenzen braucht (unter
anderem macht das Jakob, Manu, ich, die Gneiting-Gruppe bzw.
Hamill-Gruppe hat auch was in die Richtung aber mit Gamma \dots).

Effizienter waere es wohl kurz das hier EMOS genannte Modell (was
eigentlich einfach ein non-homogeneous regression Modell ist)
im Appendix kurz hinzuschreiben?

    \begin{equation}
        \begin{split}
            y^{1/p} = \mathcal{N}_0(\mu, \sigma) \\
            \mu = \beta_0 + \beta_1 \cdot \bar{\mbox{tp}^{1/p}} + \mbox{split?} \\
            \log(\sigma) = \gamma_0 + \gamma_1 \cdot \mbox{sd}(\mbox{tp}^{1/p}) + \mbox{split?}
        \end{split}
    \end{equation}

\dots damit k\"onnte man auch auf Kommentar 8 von Reviewer 2 reagieren.
\end{reto}

\begin{review}
3. Looking at the PIT in the appendix, they seem to provide a 
worse fit as compared to the quantile residuals. Also, both 
just give a graphical device. If the focus is on prediction 
then my impression is that one should rather plot PIT and 
residuals for the predictions.
\end{review}

\textit{TO DO}

\begin{reto}
Beide? Verstehe ich den Kollegen da richtig?
\end{reto}

\begin{review}
4. Figure 5: Which most important variables are included in the 
boosting algorithm? Why are the 10 most important covariates chosen?
\end{review}

Figure 6 shows the 10 most important covariates of the forest model
in one particular setting (station Axams, learned on 1985--2008). 
Variable importance has been calculated for all covariates, however,
we decided to present only the top 10 as all following covariates 
are on a rather low and similar level not providing further useful
information. But this top 10 list is not considered for any 
specification of the boosted GAMLSS for which all available 
covariates are included. Out of them, in the exact same setting 
(Axams, 1985--2008), the boosting algorithm selects 28 covariates 
for the location parameter, mainly related to precipitation (7) 
and temperature (11), and 46 covariates for the scale parameter, 
mainly related to precipitation (10), convective available potential 
energy (8) and temperature (15).
% mu: 7 tppow, 1 cape, 2 dswrf, 2 msl, 2 pwat, 3 tcolc, 11 temperature
% sigma: 10 tppow, 8 cape, 1 dswrf, 5 msl, 5 pwat, 2 tcolc, 15 temperature

% R> names(coef(gb, parameter = "mu"))
% [1] "bbs(tppow_min)"        "bbs(tppow_max)"        "bbs(tppow_mean1218)"  
% [4] "bbs(tppow_mean1824)"   "bbs(tppow_mean2430)"   "bbs(tppow_sprd0612)"  
% [7] "bbs(tppow_sprd2430)"   "bbs(capepow_sprd0612)" "bbs(dswrf_mean_mean)" 
%[10] "bbs(dswrf_sprd_mean)"  "bbs(msl_sprd_min)"     "bbs(msl_sprd_max)"    
%[13] "bbs(pwat_sprd_min)"    "bbs(pwat_sprd_max)"    "bbs(tcolc_mean_mean)" 
%[16] "bbs(tcolc_mean_max)"   "bbs(tcolc_sprd_mean)"  "bbs(t500_mean_min)"   
%[19] "bbs(t850_mean_min)"    "bbs(t500_sprd_mean)"   "bbs(t500_sprd_max)"   
%[22] "bbs(t700_sprd_mean)"   "bbs(t700_sprd_min)"    "bbs(t700_sprd_max)"   
%[25] "bbs(t850_sprd_min)"    "bbs(t850_sprd_max)"    "bbs(tdiff700850_min)" 
%[28] "bbs(tdiff500700_max)"

% R> names(coef(gb, parameter = "sigma"))
% [1] "bbs(tppow_mean)"       "bbs(tppow_sprd)"       "bbs(tppow_min)"       
% [4] "bbs(tppow_max)"        "bbs(tppow_mean1824)"   "bbs(tppow_mean2430)"  
% [7] "bbs(tppow_sprd0612)"   "bbs(tppow_sprd1218)"   "bbs(tppow_sprd1824)"  
%[10] "bbs(tppow_sprd2430)"   "bbs(capepow_sprd)"     "bbs(capepow_min)"     
%[13] "bbs(capepow_max)"      "bbs(capepow_mean1218)" "bbs(capepow_mean1224)"
%[16] "bbs(capepow_mean1230)" "bbs(capepow_sprd0612)" "bbs(capepow_sprd1218)"
%[19] "bbs(dswrf_sprd_mean)"  "bbs(msl_mean_mean)"    "bbs(msl_mean_max)"    
%[22] "bbs(msl_sprd_mean)"    "bbs(msl_sprd_min)"     "bbs(pwat_mean_mean)"  
%[25] "bbs(pwat_mean_min)"    "bbs(pwat_mean_max)"    "bbs(pwat_sprd_min)"   
%[28] "bbs(pwat_sprd_max)"    "bbs(tmax_mean_mean)"   "bbs(tmax_mean_min)"   
%[31] "bbs(tcolc_mean_min)"   "bbs(tcolc_sprd_max)"   "bbs(t500_mean_mean)"  
%[34] "bbs(t700_mean_mean)"   "bbs(t700_mean_max)"    "bbs(t850_mean_min)"   
%[37] "bbs(t850_mean_max)"    "bbs(t500_sprd_max)"    "bbs(t850_sprd_min)"   
%[40] "bbs(t850_sprd_max)"    "bbs(tdiff700850_min)"  "bbs(tdiff700850_max)" 
%[43] "bbs(tdiff500700_mean)" "bbs(tdiff500700_min)"  "bbs(tdiff500700_max)" 
%[46] "bbs(msl_diff)"

Applying the stability selection procedure by Meinshausen et al. (2010) and
Hofner et al. (2015) on the same model shows that the 10 covariates with 
maximal selection probability are six covariates based on precipitation and 
one each based on mean sea level pressure, downwards short wave radiation 
flux, temperature and precipitable water respectively.

% load("~/svn/partykit/pkg/disttree/inst/draft/Axams_24to4.rda")
% sgb <- stabsel(gb, cutoff = 0.6, q = 30)

% R> sgb$selected
% tppow_min.mu         tppow_max.mu    tppow_mean1824.mu   dswrf_mean_mean.mu 
%      3                    4                    7                   25 
% tppow_min.sigma      tppow_max.sigma tppow_sprd2430.sigma  pwat_mean_max.sigma 
%      83                   84                   92                  117 
% t700_mean_max.sigma   msl_diff.sigma 
%      138                  160

% R> head(sort(sgb$max, decreasing = TRUE), 10)
% tppow_max.mu    tppow_mean1824.mu      tppow_max.sigma  t700_mean_max.sigma 
%    1.00                 0.94                 0.91                 0.87 
% dswrf_mean_mean.mu  pwat_mean_max.sigma  tppow_min.sigma tppow_sprd2430.sigma 
%    0.85                 0.84                 0.81                 0.74 
% msl_diff.sigma      tppow_min.mu 
%    0.70                 0.63

% R> head(sort(rowMeans(sgb$phat), decreasing = TRUE), 10)
% tppow_max.mu    tppow_mean1824.mu  t700_mean_max.sigma      tppow_max.sigma 
%    0.99516              0.86018              0.84936              0.84857 
% tppow_min.sigma  pwat_mean_max.sigma   dswrf_mean_mean.mu tppow_sprd2430.sigma 
%    0.79779              0.75813              0.74074              0.65091 
% msl_diff.sigma         tppow_min.mu 
%    0.63928              0.55443

Overall it can be stated that the most important covariates of the boosted 
GAMLSS are similar to the ones of the distributional forest even though they
are not exactly the same ones.

\bigskip

\noindent{\large\textbf{References}}

\begin{itemize}[leftmargin=*]

\item[] Chen, J., Brissette, F. P. and Lhi, Z. (2014). Postprocessing 
of ensemble weather forecasts using a stochastic weather generator, 
\textit{Monthly Weather Review}.

\item[] Goicoa, T., Adin, A., Ugarte, M. D. and Hodges, J. S. (2017). 
In spatio-temporal disease mapping models, identifiability constraints 
affet PQL and INLA results, \textit{Stochastic Environmental Research 
and Risk Assessment} p. to appear.

\item[] Ugarte, M. D., Adin, A. and Goicoa, T. (2017). One-dimensional, 
two-dimensional, and three-dimensional B-splines to specify space-time 
interations in bayesian disease mapping: Model fitting and model 
identifiability, \textit{Spatial Statistics} p. to appear.

\item[] Wood, S. N., Scheipl, F. and Faraway, J. J. (2013). 
Straightforward intermediate rank tensor product smoothing in mixed models, 
\textit{Statistics and Computing} \textbf{23}: 341--360.

%%%%%%%%%
% references used in answers
\item[] Gneiting, T., Raftery, A. E., Westveld III, A. H. and Goldman, T. (2005). 
Calibrated Probabilistic Forecasting Using Ensemble Model Output Statistics and 
Minimum CRPS Estimation,
\textit{Monthly Weather Review} \textbf{133, 5}: 1098--1118.

\item[] Stauffer, R., Mayr, G. J., Messner, J. W., Umlauf, N. and Zeileis, A. (2017). 
Spatio-Temporal Precipitation Climatology over Complex Terrain Using a Censored Additive Regression Model,
\textit{International Journal of Climatology} \textbf{37, 7}: 3264--3275.

\item[] Hofner, B., Mayr, A. and Schmid, M. (2016). 
\textbf{gamboostLSS}: An \textsf{R} Package for Model Building and 
Variable Selection in the {GAMLSS} Framework,
\textit{Journal of Statistical Software} \textbf{74, 1}: 1--31.

\item[] Hofner, B., Boccuto, L. and Goeker, M. (2015). 
Controlling false discoveries in high-dimensional situations: Boosting 
with stability selection,
\textit{BMC Bioinformatics} \textbf{16:144}.

\item[] Meinshausen, N. and Buehlmann, P. (2010). 
Stability selection,
\textit{Journal of the Royal Statistical Society, Series B} 
\textbf{72}: 417--473.

\end{itemize}

\end{document}