### 2018-11-04 pkg/disttree/inst/draft/supplement/Rain_stationwise_results.Rnw
final formatting
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2378 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 4503d2d73b067eabaa07a9184943d75e94a802fc

\documentclass[a4paper,11pt]{article}

%% layout
\usepackage[a4paper, total={15cm, 26.5cm}]{geometry}
\usepackage{graphicx}
\pagestyle{empty}
\setlength{\parskip}{0.5ex plus0.1ex minus0.1ex}
\setlength{\parindent}{0em}
\usepackage{Sweave}

%% fonts
\renewcommand{\rmdefault}{phv}
\renewcommand{\sfdefault}{phv}

<<echo = FALSE, results=hide>>=
## station
station <- "_STATION_"

if(station == "SeeimPaznaun") station <- "See im Paznaun"
if(station == "MatreiinOsttirol") station <- "Matrei in Osttirol"
if(station == "LadisNeuegg") station <- "Ladis-Neuegg"
if(station == "StJohannimWalde") station <- "St.Johann im Walde"
if(station == "OchsengartenObergut") station <- "Ochsengarten-Obergut"

## results data
load("stationwise.rda")
results <- stationwise[[station]]

#station_rda <- function(name, root = getwd()) {
#  for(i in c("~", "-", ".", " ")) name <- gsub(i, "", name, fixed = TRUE)  
#  file.path(root, paste0("res_", name, ".rda"))
#}
#load(station_rda(station))

## (i) map
data("StationsTyrol", package = "RainTyrol")
data("MapTyrol", package = "RainTyrol")
library("sp")
sp <- SpatialPointsDataFrame(subset(StationsTyrol, select = c(lon, lat)),
  data = subset(StationsTyrol, select = -c(lon, lat)),
  proj4string = raster::crs(MapTyrol$RasterLayer))

## convenience functions to unify results
cross <- Reduce("c", results$res_cross)

get_cross <- function(what = "crps", which = NULL) {
  ## methods
  if(is.null(which)) which <- c("distforest", "gamlss", "gamboostLSS", "gamboostLSS_cvr", "emos_log")
  if(what != "evaltime") which <- which[which != "gamboostLSS_cvr"]

  ## extract desired quantity (crps, evaltime, predtime)
  rval <- lapply(cross, "[[", what)
  rval <- if(what == "crps") {
    sapply(rval, function(x) x[which])
  } else {
    sapply(rval, function(x) x[which, "elapsed"])
  }

  ## reshape
  rval <- t(rval)
  rval <- aggregate(rval, list(rep(1:10, each = 7)), mean, na.rm = TRUE)[, -1]
  rval <- as.matrix(rval)
  which[which == "emos_log"] <- "EMOS"
  colnames(rval) <- which

  return(rval)
}

fmt <- function(x) {
  ## select and aggregate (if necessary)
  if(!is.null(dim(x))) {
    x <- if("elapsed" %in% colnames(x)) x[, "elapsed"] else colMeans(x, na.rm = TRUE)
  }

  ## round and format to three digits
  x <- format(round(x, digits = 3), nsmall = 3)

  ## collect results across methods (if necessary)
  if(length(x) == 1L) return(x)
  nam <- c("distforest", "gamlss", "gamboostLSS", "gamboostLSS_cvr", "EMOS")
  x <- x[nam]
  x[is.na(x)] <- ""
  names(x) <- nam
  return(x)
}

## (ii) summary table
tab <- cbind(
  "crps_24to4" =     fmt(results$crps),
  "fittime_24to4" =  fmt(results$fit_time),
  "predtime_24to4" = fmt(results$pred_time),
  "crps_cv" =        fmt(get_cross("crps")),
  "fittime_cv" =     fmt(get_cross("evaltime")),
  "predtime_cv" =    fmt(get_cross("predtime"))
)

## (iii) crps skill score
crps_cross <- get_cross("crps")
crps_cross_mean <- colMeans(crps_cross, na.rm = TRUE)
crpss_cross <- 1 - crps_cross[, c("distforest", "gamlss", "gamboostLSS")] / crps_cross[, "EMOS"]
crpss_24to4 <- 1 - colMeans(results$crps[, c("distforest", "gamlss", "gamboostLSS")], na.rm = TRUE) / mean(results$crps[, "EMOS"], na.rm = TRUE)

## (iv) variable importance
varimp10 <- rev(head(sort(results$vimp_crps, decreasing = TRUE), 10))
names(varimp10) <- gsub("pow", "", names(varimp10), fixed = TRUE)
@

\begin{document}
\SweaveOpts{concordance=TRUE}

\section*{\huge \Sexpr{station}}

Longitude: \Sexpr{fmt(StationsTyrol[StationsTyrol$name == station, "lon"])}

Latitude: \Sexpr{fmt(StationsTyrol[StationsTyrol$name == station, "lat"])}

Altitude: \Sexpr{fmt(StationsTyrol[StationsTyrol$name == station, "alt"])}

\setkeys{Gin}{width=0.7\textwidth}
\vspace*{-4.1cm}\hspace*{4cm}%
<<map, fig=TRUE, echo = FALSE, height = 5, width=7=>>=
plot(MapTyrol$SpatialPolygons)
points(sp, pch = 21, col = "black", las = 1, cex = 1)
points(sp[which(sp$name == station), ], pch = 21, bg = hcl(325, 100, 70), cex = 1.5)
@
\vspace*{-1.6cm}

\subsection*{Overview}

\begin{tabular}{lrrrrrr}
\hline
      & \multicolumn{3}{l}{Single split} & \multicolumn{3}{l}{10 times 7-fold cross-validation} \\
Model & CRPS & Fit time & Predict time                 & CRPS & Fit time & Predict time                       \\ \hline
<<tab, echo=FALSE, results=tex>>=
tab <- apply(tab, 1, paste, collapse = " & ")
tab <- paste(c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS", "\\emph{(mstop selection)}", "EMOS"), "&", tab, "\\\\")
writeLines(tab)
@
\hline
\end{tabular}

\subsection*{CRPS skill score (cross-validation, reference: EMOS)}

\setkeys{Gin}{width=0.73\textwidth}
<<boxplot_crps_skillscore, fig=TRUE, echo=FALSE, height=4, width=6.4>>=
par(mar = c(2.5, 4, 1, 2))
boxplot(crpss_cross, 
        ylim = c(min(-0.01, min(crpss_cross, na.rm = TRUE) - 0.01), 
                 max(-0.01, max(crpss_cross, na.rm = TRUE) + 0.01)),
        main = "",
        names = c("Distributional\nforest", "Prespecified\nGAMLSS", "Boosted\nGAMLSS"),
        ylab = "CRPS skill score", col = "lightgray", las = 1) 
#axis(1, 0:4, c("", "Distributional\nforest", "Prespecified\nGAMLSS", "Boosted\nGAMLSS", ""))
abline(h = 0, col = hcl(50, 100, 50), lwd = 2)
# lines(crpss_24to4, col = hcl(128, 100, 50), lwd = 2, type = "o", pch = 19)
@

\subsection*{Variable importance}

<<varimp, fig=TRUE, echo=FALSE, height=4, width=6.4>>=
par(mar = c(5, 10, 1, 2))
barplot(varimp10, horiz = TRUE, las = 1, axes = FALSE,
        xlab = "Mean increase in CRPS",
        xlim = c(0,max(varimp10)))
axis(1, at = seq(0, ceiling(max(varimp10) * 5)/5, 0.02), las = 1, mgp = c(0, 1, 0))
@

\newpage

\subsection*{Residual QQ plots}

\setkeys{Gin}{width=0.93\textwidth}
<<qq, fig=TRUE, echo=FALSE, height=6, width = 7, png=TRUE, pdf=FALSE>>=
library("countreg")
set.seed(7)
par(mfrow = c(2, 2))
qqrplot(results$pit_df, nsim = 100, main = "Distributional forest", ylim = c(-5, 5), col = gray(0.04, alpha = 0.01), pch = 19)
qqrplot(results$pit_ml, nsim = 100, main = "EMOS",		    ylim = c(-5, 5), col = gray(0.04, alpha = 0.01), pch = 19)
if(!all(is.na(results$pit_g))) qqrplot(results$pit_g,  nsim = 100, main = "Prespecified GAMLSS",   ylim = c(-5, 5), col = gray(0.04, alpha = 0.01), pch = 19)
qqrplot(results$pit_gb, nsim = 100, main = "Boosted GAMLSS",	    ylim = c(-5, 5), col = gray(0.04, alpha = 0.01), pch = 19)
@

\subsection*{PIT histograms}

<<pit, fig=TRUE, echo=FALSE, height=6, width = 7>>=
set.seed(4)
par(mfrow = c(2, 2))
pithist(results$pit_df, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "Distributional forest", ylim = c(0, 1.9))
pithist(results$pit_ml, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "EMOS", ylim = c(0, 1.9))
if(!all(is.na(results$pit_g))) pithist(results$pit_g,  nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "Prespecified GAMLSS", ylim = c(0, 1.9))
pithist(results$pit_gb, nsim = 1000, breaks = seq(0, 1, length.out = 9), main = "Boosted GAMLSS", ylim = c(0, 1.9))
@

\end{document}