### 2019-03-05 pkg/partykit/inst/guideliketest/DAGStat2019/slides.Rnw
small corrections in notation and illustrations
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2482 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: cca523f738777eba594966e8394465dd42c01660

\documentclass[11pt,t,usepdftitle=false,aspectratio=169]{beamer}
\usetheme[license, url]{uibk}

\title{The Power of Unbiased Recursive Partitioning:\\
A Unifying View of CTree, MOB, and GUIDE}
\author{Lisa Schlosser, Torsten Hothorn, Achim Zeileis}
\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
%\renewcommand{\mysize}{\fontsize{7.7pt}{9pt}\selectfont}
\URL{http://www.partykit.org/} 
\headerimage{3}

\setbeamertemplate{caption}{\insertcaption} 
%% includes a replacement for \usepackage{Sweave}
\usepackage{changepage}
\usepackage{amsmath,tikz}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap,trees,tikzmark,decorations.pathreplacing}
%\graphicspath{{plots/}}

\newcommand{\argmax}{\operatorname{argmax}\displaylimits}

%% colors
\definecolor{HighlightOrange}{rgb}{0.9490196,0.5725490,0.0000000}
\definecolor{HighlightBlue}{rgb}{0.4784314,0.7490196,0.9803922}

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
set.seed(7)
# setwd("~/svn/partykit/pkg/partykit/inst/guideliketest/DAGStat2019/")
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
library("partykit")
library("Formula")
library("parallel")
library("lattice")
library("scales")
library("colorspace")
#source("../guidelike.R")
source("../tests_guideprune.R")
#source("../ccprune.R")

load("../draft/stump.rda")
sim_stump <- simres$res
rm(simres)

load("../draft/tree.rda")
sim_tree <- simres$res
rm(simres)

load("../draft/pruning.rda")
pr <- simres$res
## extract data for xyplot comparing ARI and ARI_p
pr <- pr[,c(1,2,7,8,9,12,15)]
## extended data set with postpruning results being represented as separated tests
prex <- pr[,c(1,2,3,4,6)]
prex <- rbind(prex,prex)

prex$test <- factor(prex$test,
                    levels = c("ctree","mfluc",
                               "guide_sum_12", "guide_sum_1_cor",
                               "ctree_p","mfluc_p",
                               "guide_sum_12_p", "guide_sum_1_cor_p"),
                    labels = c("CTree","MOB",
                               "GUIDE_scores", "GUIDE",
                               "CTree_p","MOB_p",
                               "GUIDE_scores_p", "GUIDE_p"))

for(i in c((length(pr$test)+1) : (2*length(pr$test)))){
  if(prex$test[i] == "CTree") prex$test[i] <- "CTree_p"
  if(prex$test[i] == "MOB") prex$test[i] <- "MOB_p"
  if(prex$test[i] == "GUIDE_scores") prex$test[i] <- "GUIDE_scores_p"
  if(prex$test[i] == "GUIDE") prex$test[i] <- "GUIDE_p"
}

prex$nrsubgr[(length(pr$nrsubgr)+1) : (2*length(pr$nrsubgr))] <- pr$nrsubgr_p
prex$ari[(length(pr$ari)+1) : (2*length(pr$ari))] <- pr$ari_p

load("../draft/3way_xi00_delta03.rda")
load("../draft/3way_xi08_delta1.rda")
d00$xi <- 0
d08$xi <- 0.8
d3way <- rbind(d00, d08)
d3way$xi <- factor(d3way$xi, levels = c(0,0.8), labels = c(0,0.8))

## HCL palette
pal_dgp <- qualitative_hcl(3, "Set 2")

## HCL palette (light)
pal_results <- qualitative_hcl(4, "Dark 3")

#produce plots
set.seed(7)
x <- runif(70,-1,1)
e1 <- rnorm(70,0,1)
e2 <- rnorm(70,0,1)
y1 <- x*6 + e1 
y2 <- x*1 + e2
y <- c(y1,y2)

jpeg("lm1.jpeg")
par(mar=c(3.2,3.2,0,0))
plot(x,y1, ylab = "", xlab = "", xaxt="n", yaxt="n", ylim = c(-7,7))
mtext("Y",side = 2, las = 2, line = 2, cex = 2.5, col = gray.colors(1,0.3))
mtext("X",side = 1, las = 1, line = 2, cex = 2.5, col = gray.colors(1,0.3))
points(x, y2)
abline(lm(y~c(x,x)), col = "red", lwd = 3)
dev.off()

jpeg("lm2.jpeg")
par(mar=c(3.2,3.2,0,0))
plot(x,y1, ylab = "", xlab = "", xaxt="n", yaxt="n", ylim = c(-7,7))
mtext("Y",side = 2, las = 2, line = 2, cex = 2.5, col = gray.colors(1,0.3))
mtext("X",side = 1, las = 1, line = 2, cex = 2.5, col = gray.colors(1,0.3))
abline(lm(y1~x), col = "red", lwd = 3)
dev.off()

jpeg("lm3.jpeg")
par(mar=c(3.2,3.2,0,0))
plot(x,y2, ylab = "", xlab = "", xaxt="n", yaxt="n", ylim = c(-7,7))
mtext("Y",side = 2, las = 2, line = 2, cex = 2.5, col = gray.colors(1,0.3))
mtext("X",side = 1, las = 1, line = 2, cex = 2.5, col = gray.colors(1,0.3))
abline(lm(y2~x), col = "red", lwd = 3)
dev.off()

@

\begin{document}

\section{Unbiased Recursive Partitioning}

\begin{frame}%[fragile]
\frametitle{Motivation}

\vspace{-0.5cm}

\center
\begin{tikzpicture}
\visible<1-3>{
\node[inner sep=0pt] (lm1) at (4, 6)
    {\includegraphics[width=0.2\textwidth]{lm1.jpeg}};
}
\visible<2>{
\node[scale=1] at (4.15, 3.7) {other covariates $Z_1, \ldots, Z_p$?};
}
\visible<3>{
\node[inner sep=0pt] (lm2) at (2,1.7)
    {\includegraphics[width=0.2\textwidth]{lm2.jpeg}};
\node[inner sep=0pt] (lm3) at (6,1.7)
    {\includegraphics[width=0.2\textwidth]{lm3.jpeg}};
\draw[-, line width=0.2pt] (3.5,4.85) -- (2.5,3.1) node [midway, left] {\scriptsize $Z_j \leq \xi$};
\draw[-, line width=0.2pt] (4.7,4.85) -- (5.7,3.1) node [midway, right] {\scriptsize $Z_j > \xi$};
}
\visible<4->{
\node[ellipse, fill=HighlightBlue!70, align=center, scale = 1, minimum width=90pt, minimum height = 30pt] (model1) at (4.15, 5.35) {$\mathcal{M}(Y,X;\hat{\beta}$)};
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 1, minimum width=80pt, minimum height = 30pt] (model2) at (2.15, 2.55) {$\mathcal{M}(Y_1,X_1;\hat{\beta}$)};
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 1, minimum width=80pt, minimum height = 30pt] (model3) at (6.15, 2.55) {$\mathcal{M}(Y_2,X_2;\hat{\beta}$)};
\draw[-, line width=0.2pt] (3.5,4.85) -- (2.5,3.1) node [midway, left] {\scriptsize $Z_j \leq \xi$};
\draw[-, line width=0.2pt] (4.7,4.85) -- (5.7,3.1) node [midway, right] {\scriptsize $Z_j > \xi$};
}
\visible<5>{
\node[scale=1] at (4.15, 1) {$\mathcal{M}$ can also be a more general model (possibly without $X$).};
}
\end{tikzpicture}
\end{frame}

\begin{frame}%[fragile]
\frametitle{Unbiased Recursive Partitioning}
{\bf Basic tree algorithm:}
\medskip
\begin{enumerate}
\item Fit a model $\mathcal{M}(Y,X;\hat{\beta})$ to the response $Y$ (and possible covariates $X$).
\item Assess association of $Y$ (or a model-based transformation)
and each possible split variable~$Z_j$ and select the split~variable~$Z_{j^\ast}$ showing the strongest association.
\item Choose the corresponding split point leading to the highest improvement of model fit and split the data.
\item Repeat steps 1--3 recursively in each of the resulting subgroups until some stopping criterion is met.
\end{enumerate}

\vspace{0.8cm}

{\bf Here:} Focus on split variable selection (step 2).

\end{frame}

\begin{frame}%[fragile]
\frametitle{Split Variable Selection}

{\bf General testing strategy:}
\begin{enumerate}
\item Evaluate a discrepancy measure giving information on the goodness of fit of the model for each observation and each model parameter (transformation~of~$Y$).
\item Apply a statistical test assessing dependency of the discrepancy measure to each possible split variable $Z_j$.
\item Select the split variable $Z_j^\ast$ showing the smallest $p$-value.
\end{enumerate}

\bigskip

{\bf Possible discrepancy measures:}
\begin{itemize}
\item (Ranks of) Y
\item Residuals
\item Model scores
\item Absolute deviations from an average value
\item \ldots
\end{itemize}
\end{frame}

\begin{frame}%[fragile]
\frametitle{Discrepancy Measures}
As an example linear models $\mathcal{M}(Y,X;\beta_0, \beta_1)$
fitted via ordinary least squares~(OLS)
are considered with the following discrepancy measures:

\bigskip

{\bf Residuals:}
\begin{flalign*}
\hspace{0.7cm}
r(Y,X,\hat{\beta_0},\hat{\beta_1}) = Y - \hat{\beta_0} - \hat{\beta_1} \cdot X &&
\end{flalign*}

\bigskip

{\bf Model scores} (based on least squares):
\begin{flalign*}
\hspace{0.7cm}
s(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
\left(\frac{\partial r^2(Y, X, \hat{\beta_0}, \hat{\beta_1})}{\partial \beta_0} \quad , \quad  
\frac{\partial r^2(Y, X, \hat{\beta_0}, \hat{\beta_1})}{\partial \beta_1}\right) 
&&
\end{flalign*}

\visible<1>{
\medskip
Alternatively: Based on log-likelihood.
}
\visible<2>{
\vspace{-1cm}
$$\hspace{-0.3cm}{\color{red}\Downarrow} \hspace{3.4cm} {\color{red}\Downarrow} $$
$$
{\color{red} -2 \cdot r(Y, X, \hat{\beta_0},\hat{\beta_1})}  \hspace{0.7cm} {\color{red} -2 \cdot X \cdot r(Y, X, \hat{\beta_0},\hat{\beta_1})}
$$
}
\end{frame}

\begin{frame}%[fragile]
\frametitle{General Framework}

{\bf{Building blocks:}}

\begin{itemize}
\item {\only<4->{\color{lightgray}} Residuals vs.\ full model scores}
\item {\only<1-3,6->{\color{lightgray}} Binarization of residuals/scores}
\item {\only<1-5>{\color{lightgray}} Categorization of possible split variables}
\end{itemize}

\bigskip

\only<2>{
\small
$$
s(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
-2 \cdot \begin{pmatrix} 
r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} X \cdot r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})}\\
r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} X \cdot r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})}\\
\vdots & {\color{lightgray} \vdots} \\
r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} X \cdot r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})}
\end{pmatrix}
%\begin{pmatrix}
%\partial_{\beta_0} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})}\\
%\partial_{\beta_0} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})}\\
%\vdots & {\color{lightgray} \vdots} \\
%\partial_{\beta_0} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})}
%\end{pmatrix}
$$
}
\only<3>{
\small
$$
s(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
-2 \cdot \begin{pmatrix} 
r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & X \cdot r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})\\
r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & X \cdot r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})\\
\vdots & \vdots \\
r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & X \cdot r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})
\end{pmatrix}
%\begin{pmatrix} 
%\partial_{\beta_0} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})\\
%\partial_{\beta_0} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})\\
%\vdots & \vdots \\
%\partial_{\beta_0} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})
%\end{pmatrix}
$$
}
\only<4-5>{
\small
$$
r(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
\begin{pmatrix}
r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})\\
r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})\\
\vdots \\
r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})
\end{pmatrix}
%
\visible<5>{
\quad \Rightarrow \quad
\small
\begin{pmatrix}
\text{pos}\\
\text{non-pos}\\
\vdots \\
\text{pos}
\end{pmatrix}
}
$$
}
\only<6-7>{
\small
$$
Z_j = 
\begin{pmatrix}
Z_{j1}\\
Z_{j2}\\
\vdots \\
Z_{jn}\\
\end{pmatrix}
%
\visible<7>{
\quad \Rightarrow \quad
\small
\begin{pmatrix}
\text{Q}3\\
\text{Q}1\\
\vdots \\
\text{Q}2
\end{pmatrix}
}
$$
}
\end{frame}

\begin{frame}
\frametitle{Combinations}

The testing strategies of CTree, MOB, and GUIDE can be obtained from the general 
framework by specific combinations of the building blocks together with the 
corresponding approximation type of the underlying null distribution.

\bigskip

\begin{table}[t]
\begin{tabular}{ l c c c c }
\hline 
      & Scores            & Binarization  & Categorization  & Null distribution \\ 
\hline
CTree & full model scores & --          & --            & conditional\\
MOB   & full model scores & --          & --            & unconditional\\
GUIDE & residuals         & \checkmark  & \checkmark    & unconditional\\
\hline
\end{tabular}
\end{table}
\end{frame}

\begin{frame}%[fragile]
\frametitle{Pruning}

{\bf Goal:} Avoid overfitting.

\bigskip

{\bf Two Strategies:}

\begin{itemize}
\item Pre-pruning: \\
Internal stopping criterion based on Bonferroni-corrected $p$-values of the applied tests. Stop splitting when there is no significant association. 
\item Post-pruning: \\
First grow a very large tree and afterwards prune splits that do not improve the model fit, 
either via cross-validation (e.g.,~cost-complexity pruning as in CART) or 
based on information criteria (e.g.,~AIC or BIC). 
\end{itemize}
\end{frame}

\begin{frame}%[fragile]
\frametitle{Simulation}

\begin{table}
\begin{center}
\begin{tabular}{l l l}
\hline
Name                    & Notation                   & Specification \\
\hline
\textit{Variables:}     &                            & \\ 
Response                & $Y$                        & $=\beta_0(Z_1) + \beta_1(Z_1) \cdot X + \epsilon$ \\
Regressor               & $X$                        & $\mathcal{U}([-1,1])$ \\
Error                   & $\epsilon$                 & $\mathcal{N}(0,1)$ \\
True split variable     & $Z_1$                      & $\mathcal{U}([-1,1])$  or $\mathcal{N}(0, 1)$ \\
Noise split variables   & $Z_2, Z_3, \dots, Z_{10}$  & $\mathcal{U}([-1,1])$  or $\mathcal{N}(0, 1)$ \\
\hline
\textit{Parameters/functions:} &            & \\ 
Intercept                      & $\beta_0$  & $0$ or $\pm \delta$\\
Slope                          & $\beta_1$  & $1$ or $\pm \delta$\\
True split point               & $\xi$      & $\in \{0, 0.2, 0.5, 0.8\}$ \\
Effect size                    & $\delta$   & $\in \{0, 0.1, 0.2, \ldots , 1\}$ \\
\hline
\end{tabular}
\end{center}
\end{table}

\end{frame}

\begin{frame}%[fragile]
\frametitle{Simulation 1: True tree structure}

<<dgp, echo=FALSE>>=
set.seed(7)

nobs <- 250
sigma <- 0.5
xi <- 0
delta <- 1
beta0 <- 0
beta1 <- 1

d <- dgp_stump(nobs = nobs, binary_beta = TRUE, binary_regressor = FALSE, only_intercept = FALSE,
               vary_beta = "beta0", beta1 = beta1, delta = delta, xi = xi, sigma = sigma)

sp_xi1 <- partysplit(3L, breaks = 0)

pn <- partynode(1L, split = sp_xi1, 
                kids = list(
                  partynode(2L, info = c("true parameters:", 
                                         expression(beta[0] ==~" "~0~" or "~-delta), 
                                         expression(beta[1] ==~" "~1~" or "~+delta))),
                  partynode(3L, info = c("true parameters:", 
                                         expression(beta[0] ==~" "~0~" or "~+delta), 
                                         expression(beta[1] ==~" "~1~" or "~-delta)))))

p <- party(pn, d)

paltrees <- rgb(c(1, 0.9, 1), c(1, 0.9, 1), c(1, 0.9, 1))

ep <- function (obj, digits = 3, abbreviate = FALSE, justmin = Inf, 
                just = c("alternate", "increasing", "decreasing", "equal"), 
                fill = "white") 
{
  meta <- obj$data
  justfun <- function(i, split) {
    myjust <- if (mean(nchar(split)) > justmin) {
      match.arg(just, c("alternate", "increasing", "decreasing", "equal"))
    } else {
      "equal"
    }
    k <- length(split)
    rval <- switch(myjust, equal = rep.int(0, k), 
                   alternate = rep(c(0.5, -0.5), length.out = k), 
                   increasing = seq(from = -k/2, to = k/2, by = 1), 
                   decreasing = seq(from = k/2, to = -k/2, by = -1))
    unit(0.5, "npc") + unit(rval[i], "lines")
  }
  function(node, i) {
    split <-  c("<= xi", "> xi")
    y <- justfun(i, split)
    split <- split[i]
    if (any(grep(">", split) > 0) | any(grep("<", split) > 0)) {
      tr <- suppressWarnings(try(parse(text = paste("phantom(0)", split)), silent = TRUE))
      if (!inherits(tr, "try-error")) split <- tr
    }
    grid.rect(y = y, gp = gpar(fill = fill, col = 0), width = unit(1, "strwidth", split))
    grid.text(split, y = y, just = "center")
  }
}
class(ep) <- "grapcon_generator"

tp <- function (obj, digits = 3, abbreviate = FALSE, 
                fill = c("lightgray", "white"), id = TRUE, 
                just = c("center", "top"), top = 0.85, 
                align = c("center", "left", "right"), gp = NULL, FUN = NULL, 
                height = NULL, width = NULL) 
{
  nam <- names(obj)
  extract_label <- function(node) formatinfo_node(node, FUN = FUN, 
                                                  default = c("terminal", "node"))
  maxstr <- function(node) {
    lab <- extract_label(node)
    klab <- if (is.terminal(node)) 
      ""
    else unlist(lapply(kids_node(node), maxstr))
    lab <- c(lab, klab)
    lab <- try(unlist(lapply(lab, function(x) strsplit(x, 
                                                       "\n"))), silent = TRUE)
    if (inherits(lab, "try-error")) {
      paste(rep("a", 9L), collapse = "")
    }
    else {
      return(lab[which.max(nchar(lab))])
    }
  }
  nstr <- if (is.null(width)) 
    maxstr(node_party(obj))
  else paste(rep("a", width), collapse = "")
  just <- match.arg(just[1L], c("center", "centre", "top"))
  if (just == "centre") 
    just <- "center"
  align <- match.arg(align[1L], c("center", "centre", "left", 
                                  "right"))
  if (align == "centre") 
    align <- "center"
  rval <- function(node) {
    if(node$id == 2) fill <- c("white", pal_dgp[1])
    if(node$id == 3 | node$id == 4) fill <- c("white", pal_dgp[2])
    if(node$id == 5) fill <- c("white", pal_dgp[3])
    fill <- rep(fill, length.out = 2)
    lab <- extract_label(node)
    if (!is.null(gp)) {
      outer_vp <- viewport(gp = gp)
      pushViewport(outer_vp)
    }
    if (is.null(height)) 
      height <- length(lab) + 1L
    node_vp <- viewport(x = unit(0.5, "npc"), y = unit(if (just == 
                                                           "top") 
      top
      else 0.5, "npc"), just = c("center", just), width = unit(1, 
                                                               "strwidth", nstr) * 1.1, height = unit(height, "lines"), 
      name = paste("node_terminal", id_node(node), sep = ""), 
      gp = if (is.null(gp)) 
        gpar()
      else gp)
    pushViewport(node_vp)
    grid.rect(gp = gpar(fill = fill[1]))
    for (i in seq_along(lab)) grid.text(x = switch(align, 
                                                   center = unit(0.5, "npc"), left = unit(1, "strwidth", 
                                                                                          "a"), right = unit(1, "npc") - unit(1, "strwidth", 
                                                                                                                              "a")), y = unit(length(lab) - i + 1, "lines"), 
                                        lab[i], just = align)
    if (id) {
      nodeIDvp <- viewport(x = unit(0.5, "npc"), y = unit(1, 
                                                          "npc"), width = max(unit(1, "lines"), unit(1.3, 
                                                                                                     "strwidth", nam[id_node(node)])), height = max(unit(1, 
                                                                                                                                                         "lines"), unit(1.3, "strheight", nam[id_node(node)])))
      pushViewport(nodeIDvp)
      grid.rect(gp = gpar(fill = fill[2], lty = "solid"))
      grid.text(nam[id_node(node)])
      popViewport()
    }
    if (is.null(gp)) 
      upViewport()
    else upViewport(2)
  }
  return(rval)
}

class(tp) <- "grapcon_generator"
@

\vspace{-0.2cm}

\begin{figure}%[t]
\setkeys{Gin}{width=1\linewidth}
\minipage{0.36\textwidth}
<<dgp_stump1, fig=TRUE, echo=FALSE, width=4, height=3>>=
#par(mar=c(3,3,0,0), oma = c(3,3,0,0))
plot(p, 
     tp_args = list(FUN = identity, width = 15), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1, edge_panel = ep, terminal_panel = tp)
@
\endminipage
\minipage{0.36\textwidth}
<<dgp_stump2, fig=TRUE, echo=FALSE, width=4, height=3>>=
par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "", xlab = "", 
     main = expression("varying"~beta[0]),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
mtext("Y", side = 2, line = 2.5)
mtext("X", side = 1, line = 2.5)
abline(coef = c(-1,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(1,1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>xi], ylim = c(-5,5), pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = 3.2)
mtext(expression(beta[1]== 1), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = 2.2)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 0.2)
mtext(expression(beta[1]== 1), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = -0.8)
@
\endminipage

\minipage{0.36\textwidth}
<<dgp_stump3, fig=TRUE, echo=FALSE, width=4, height=3>>=
d <- dgp_stump(nobs = nobs, binary_beta = TRUE, binary_regressor = FALSE, only_intercept = FALSE,
               vary_beta = "beta1", beta0 = beta0, delta = delta, xi = xi, sigma = sigma)

par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "", xlab = "", main = expression("varying"~beta[1]),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
mtext("Y", side = 2, line = 2.5)
mtext("X", side = 1, line = 2.5)
abline(coef = c(0,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(0,-1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>xi], ylim = c(-5,5), pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -1)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 2)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1)
@
\endminipage
\minipage{0.36\textwidth}
<<dgp_stump4, fig=TRUE, echo=FALSE, width=4, height=3>>=
d <- dgp_stump(nobs = nobs, binary_beta = TRUE, binary_regressor = FALSE, only_intercept = FALSE,
               vary_beta = "all", delta = delta, xi = xi, sigma = sigma)
par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "", xlab = "",
     main = expression("varying"~beta[0]~"and"~beta[1]),
     xlim = c(-1,1.1),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
mtext("Y", side = 2, line = 2.5)
mtext("X", side = 1, line = 2.5)
abline(coef = c(-1,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(1,-1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>=xi], pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1.7)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 0.7)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -0.7)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -1.7)
@
\endminipage
\end{figure}
\end{frame}

\begin{frame}%[fragile]
\frametitle{Simulation 1: Residuals vs.\ full model scores}

\begin{figure}
\setkeys{Gin}{width=0.8\linewidth}
<<scores_stump, fig=TRUE, echo=FALSE, height=5, width=8>>=
subdata <- subset(sim_stump, test %in% c("ctree","mfluc",
                                         "guide_sum_12", "guide_sum_1_cor"))
subdata$test <- factor(subdata$test, 
                       levels = c("ctree","mfluc",
                                  "guide_sum_12", "guide_sum_1_cor"),
                       labels = c("CTree","MOB",
                                  "GUIDE_scores", "GUIDE"))

subdataxi <- subset(subdata, xi %in% c(0,0.8))

par.settings <- list(superpose.symbol = list(col = pal_results, 
                                             fill = pal_results), 
                     superpose.line = list(col = pal_results))

library("latticeExtra")
useOuterStrips(xyplot(prop_Tsplit ~ delta | vary_beta + xi, groups = ~ test, 
                      data = subdataxi, 
                      subset = binary_beta == TRUE & binary_regressor == FALSE & 
                        only_intercept == FALSE,# & xi == 0, 
                      type = "l", lty = c(1,1,1,1),
                      scales=list(x=list(at=seq(1,11,2))),
                      key =  list(text = list(levels(subdataxi$test)),
                                  lines = list(lty = c(1,1,1,1), type = "l", pch = 1, cex = 0.8,
                                               col = pal_results, lwd = 1)),
                      layout= c(3,2),
                      index.cond = list(c(2,3,1),c(2,1)),
                      xlab = expression(delta),
                      ylab = expression("Proportion of selecting"~Z[1]),
                      par.settings = par.settings),

               strip =strip.custom(factor.levels = c(expression("vary"~beta[0]~"and"~beta[1]), 
                                                     expression("vary"~beta[0]), 
                                                     expression("vary"~beta[1])),
                                   bg = "lightgray"),
               strip.left=strip.custom(factor.levels = c(expression(xi == 0),                                                              expression(xi == 0.8)),
                                       bg = "lightgray"))

@
\end{figure}

\end{frame}

\begin{frame}%[fragile]
\frametitle{Simulation 1: Full factorial analysis (vary both)}

\begin{figure}%[t]
\setkeys{Gin}{width=0.8\linewidth}
<<3way, fig=TRUE, echo=FALSE, height=4, width=8>>=
xyplot(pval_z1 ~ cat | xi, 
       groups = interaction(bin, res_scores), 
       data = d3way, 
       ylim = c(0,0.5),
       type = c("a"), 
       col = c("red", "blue", "red", "blue"),
       lty = c(1,1,2,2),
       xlab = "Categorization",
       ylab = expression(p~"-value of"~Z[1]),
       #index.cond = list(c(1,2,3,4)),
       strip = strip.custom(factor.levels = c(expression(xi==0~"and"~delta==0.3), 
                                              expression(xi==0.8~"and"~delta==1))),
       key = list(text = list(levels(interaction(d3way$bin, d3way$res_scores))),
                  lines = list(lty = c(1,1,2,2), 
                               type = "l",
                               col = c("red", "blue", "red", "blue"))))
@
\end{figure}
\end{frame}

\begin{frame}%[fragile]
\frametitle{Simulation 2: True tree structure}

<<true_tree_dgp, eval = TRUE, echo=FALSE, results=hide>>=

# d <- dgp_tree(nobs = 250)
set.seed(7)
d <- dgp_tree(nobs = 250, xi = 0, binary_regressor = FALSE, vary_beta = "all", 
              binary_beta = TRUE, nrsplitvar = 2, sigma = 0.5)

sp_xi2 <- partysplit(4L, breaks = 0)
sp_xi1 <- partysplit(3L, breaks = 0)

pn <- partynode(1L, split = sp_xi2, 
                kids = list(
                  partynode(2L, info = c("true parameters:", 
                                         expression(beta[0] == 0), 
                                         expression(beta[1] == +delta))),
                  partynode(3L, split = sp_xi1, 
                            kids = list(
                              partynode(4L, info = c("true parameters:", 
                                                     expression(beta[0] == -delta), 
                                                     expression(beta[1] == -delta))),
                              partynode(5L, info = c("true parameters:", 
                                                     expression(beta[0] == +delta), 
                                                     expression(beta[1] == -delta)))))))

p <- party(pn, d)

#ctest <- evaltests(y~x|z1+z2+z3+z4+z5+z6+z7+z8+z9+z10, data = d, testfun = "ctree", stump = FALSE)
@

\begin{figure}%[t]
\begin{center}
\setkeys{Gin}{width=1\linewidth}
\minipage{0.45\textwidth}
<<dgp_tree1, fig=TRUE, echo=FALSE, width=5, height=4.5>>=
par(mar=c(0,0,0,0), oma=c(0,0,0.4,0))
#paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
paltrees <- rgb(c(1, 0.9, 1), c(1, 0.9, 1), c(1, 0.9, 1))
plot(p, edge_panel = ep, terminal_panel = tp,
     tp_args = list(FUN = identity, width = 15), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1)
@
\endminipage
\minipage{0.4\textwidth}
<<dgp_tree2, fig=TRUE, echo=FALSE, width=4, height=4>>=
par(mar = c(4,4,2,4))
plot(x = d$x[d$z2<=xi], y = d$y[d$z2<=xi], ylim = c(-5,5), ylab = "Y", xlab = "X",
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
abline(coef = c(0,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(-1,-1), col = pal_dgp[2], lwd = 2)
abline(coef = c(1,-1), col = pal_dgp[3], lwd = 2)
points(x = d$x[d$z2>xi & d$z1<=xi], y = d$y[d$z2>xi & d$z1<=xi], pch = 19, 
       col = alpha(pal_dgp[2], 0.4), cex = 0.7)
points(x = d$x[d$z2>xi & d$z1>xi], y = d$y[d$z2>xi & d$z1>xi], pch = 19, 
       col = alpha(pal_dgp[3], 0.4), cex = 0.7)
legend("top", c(expression(z[2]<=xi), 
                expression(z[2]>xi~" and "~z[1]<=xi),
                expression(z[2]>xi~" and "~z[1]>xi)), 
       col = c(pal_dgp[1], pal_dgp[2], pal_dgp[3]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 2)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1.3)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[3], las = 1, line = 0.5, at = 0.1)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[3], las = 1, line = 0.5, at = -0.6)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2.7)
@
\endminipage
\end{center}
\end{figure}

\end{frame}

\begin{frame}%[fragile]
\frametitle{Simulation 2: Residuals vs.\ full model scores}

\begin{figure}%[t]
\setkeys{Gin}{width=0.8\linewidth}
<<tree_ari, fig=TRUE, echo=FALSE, height=4, width=8>>=
subdata <- subset(sim_tree, test %in% c("ctree","mfluc",
                                        "guide_sum_12", "guide_sum_1_cor"))
subdata$test <- factor(subdata$test, 
                       levels = c("ctree", "mfluc",
                                  "guide_sum_12", "guide_sum_1_cor"),
                       labels = c("CTree","MOB",
                                  "GUIDE_scores", "GUIDE"))

par.settings <- list(superpose.symbol = list(col = pal_results, 
                                             fill = pal_results), 
                     superpose.line = list(col = pal_results))

xyplot(ari ~ delta | xi, groups = ~ test, 
       data = subdata, 
       type = "l",
       lty = c(1,1,1,1),
       key =  list(text = list(levels(subdata$test)),
                   lines = list(lty = c(1,1,1,1), type = "l", pch = 1, cex = 0.8,
                                col = pal_results, lwd = 1)),
       layout= c(4,1),
       scales=list(x=list(at=seq(1,11,2))),
       xlab = expression(delta),
       ylab = "Adjusted Rand Index",
       strip = strip.custom(factor.levels = c(expression(xi==0), 
                                              expression(xi==0.2), 
                                              expression(xi==0.5), 
                                              expression(xi==0.8))),
       par.settings = par.settings)

@

\end{figure}
\end{frame}

\begin{frame}%[fragile]
\frametitle{Simulation 2: Pre-pruning vs.\ post-pruning}

\begin{figure}%[t]
\setkeys{Gin}{width=0.8\linewidth}
<<prune_ari, fig=TRUE, echo=FALSE, height=4, width=8>>=
par.settings <- list(superpose.symbol = list(col = pal_results, 
                                             fill = pal_results), 
                     superpose.line = list(col = pal_results))

xyplot(ari ~ delta | xi, groups = ~ test, 
       data = prex, 
       type = "l", lty = c(1,1,1,1,2,2,2,2),
       key =  list(text = list(levels(prex$test)[1:4]),
                   lines = list(lty = c(1,1,1,1), type = "l", pch = 1, cex = 0.8,
                                col = pal_results, lwd = 1)),
       layout= c(4,1),
       scales=list(x=list(at=seq(1,11,2))),
       xlab = expression(delta),
       ylab = "Adjusted Rand Index",
       strip = strip.custom(factor.levels = c(expression(xi==0), 
                                              expression(xi==0.2), 
                                              expression(xi==0.5), 
                                              expression(xi==0.8))),
       par.settings = par.settings
)
@
\end{figure}

\end{frame}

\begin{frame}%[fragile]
\frametitle{Recommendations}

Based on the simulation results the following general conclusions can be drawn:
\begin{itemize}
\item Including full model scores is advantageous over using only residuals.
\item Binarizing scores/residuals does not lead to any improvements in performance.
\item Categorizing split variables can simplify calculations, however, important
effects might be missed unless a binning point is close to the true split point.
\item If the applied statistical test performs well as significance test pre-pruning
is preferable. Otherwise post-pruning can compensate this deficiency at least partly.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Software}
\vspace{0.4cm}
\textbf{Package:} Available on CRAN at\\

\url{https://CRAN.R-project.org/package=partykit}\\

\bigskip

\textbf{Replication code:} Available on R-Forge at\\

\url{https://R-Forge.R-project.org/projects/partykit/}\\

%\bigskip

%\textbf{Applied functions:}

%\begin{itemize}
%\item \texttt{ctree}
%\item \texttt{mob}
%\item Adaptions of both of them
%\item Reimplementation of GUIDE
%\end{itemize}
\end{frame}

\subsection{References}

\begin{frame}
\frametitle{References}

\vspace{-0.2cm}

\footnotesize
Schlosser L, Hothorn T, Zeileis A (2019).
 \dquote{The Power of Unbiased Recursive Partitioning: A Unifying View of CTree, MOB, and GUIDE.}
 \emph{arXiv ...}, arXiv.org E-Print Archive.
 \doi{...}

\medskip

Breiman L, Friedman J H, Olshen R A, Stone C J (1984).
 \textit{Classification and Regression Trees.}
 Wadsworth,
 California.

\medskip

Hothorn T, Hornik K, Zeileis A (2006).
 \dquote{Unbiased Recursive Partitioning: A Conditional Inference Framework.}
 \textit{Journal of Computational and Graphical Statistics},
 \textbf{15}(3), 651--674.
 \doi{10.1198/106186006X133933}

\medskip

Hothorn T, Zeileis A (2015).
 \dquote{\textbf{partykit}: A Modular Toolkit for Recursive Partytioning in \textsf{R}.}
 \textit{Journal of Machine Learning Research},
 \textbf{16}, 3905--3909.
 \url{http://www.jmlr.org/papers/v16/hothorn15a.html}

\medskip

Loh W-Y (2002).
 \dquote{Regression Trees with Unbiased Variable Selection and Interaction Detection.}
 \textit{Statistica Sinica},
 \textbf{12}(2), 361--386.
 \url{http://www.jstor.org/stable/24306967}

\medskip

Zeileis A, Hothorn T, Hornik K (2008).
 \dquote{Model-Based Recursive Partitioning.}
  \textit{Journal of Computational and Graphical Statistics},
  \textbf{17}(2), 492--514.
  \doi{10.1198/106186008X319331}

\end{frame}

%\begin{frame}
%\frametitle{References (2)}

%\footnotesize

%\end{frame}

\end{document}