### 2019-04-03 pkg/partykit/inst/pkgdown/glmtree-titanic.Rmd
started playing around with possible pkgdown materials -> in separate directory inst/pkgdown and .Rbuildignore for the moment
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2513 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 611bc8b35daac863e6f9ed1c29e2591b144e23a7

---
title: "Logistic Regression Tree: Titanic Survival Data"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
bibliography: party.bib
vignette: >
  %\VignetteIndexEntry{Logistic Regression Tree: Titanic Survival Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{partykit}
  %\VignetteKeywords{GLM, tree, recursive partitioning, logit, Titanic}
  %\VignettePackage{partykit}
---

```{r preliminaries, echo = FALSE, message = FALSE}
library("partykit")
knitr::opts_chunk$set(message = FALSE, warning = FALSE,
  dev = "png", dpi = 100, out.width = "100%",
  fig.height = 6, fig.width = 6, fig.align = "center", fig.show = "hold")
```

## Motivation: Women and children first?

To illustrate how differences in treatment effects can be captured by MOB [@Zeileis+Hothorn+Hornik:2008],
the Titanic survival data [see @Friendly+Symanzik+Onder:2019 for an overview]
is considered: The question is whether "women and children first"
is applied in the same way for all subgroups of the
passengers of the Titanic. Or, in other words, whether the effectiveness of
preferential treatment for women/children differed across subgroups.

The `Titanic` data is provided in base R as a contingency
table and transformed here to a `data.frame` for use with `glmtree()`:

```{r Titanic-data}
data("Titanic", package = "datasets")
ttnc <- as.data.frame(Titanic)
ttnc <- ttnc[rep(1:nrow(ttnc), ttnc$Freq), 1:4]
names(ttnc)[2] <- "Gender"
ttnc <- transform(ttnc, Treatment = factor(
  Gender == "Female" | Age == "Child", levels = c(FALSE, TRUE),
  labels = c("Male&Adult", "Female|Child")))
```

The resulting `ttnc` data provides factors `Survived` (yes/no), `Class` (1st,
2nd, 3rd, crew), `Gender` (male, female), and `Age` (child, adult).
Additionally, a factor `Treatment` is added that distinguishes women/children
from male adults.

## Treatment-subgroup interactions

To investigate how the preferential treatment effect (`Survived ~ Treatment`)
differs across the remaining explanatory variables, the following
logistic-regression-based tree is considered. The significance
level of `alpha = 0.01` is employed here to avoid overfitting and separation
problems in the logistic regression.

```{r Titanic-tree}
ttnc_tree <- glmtree(Survived ~ Treatment | Class + Gender + Age,
  data = ttnc, family = binomial, alpha = 0.01)
ttnc_tree
```

The corresponding tree can be visualized with `plot(ttnc_tree)` where
spinograms in the leaves show survival status versus preferential treatment
(women or children). This shows that the treatment differs strongly across passengers
classes. The plot with some small tweaks in the layout is given by:

```{r Titanic-plot, fig = TRUE, fig.height = 6, fig.width = 10}
plot(ttnc_tree, tp_args = list(ylines = 1, margins = c(1.5, 1.5, 1.5, 2.5)))
```

The treatment effect is much lower in the 3rd class where women/children still have
higher survival rates than adult men but the odds ratio is much lower compared to
all remaining classes. The split between the 2nd and the remaining two classes (1st, crew)
is due to a lower overall survival rate (intercepts of `r round(coef(ttnc_tree)[2, 1], digits = 2)` and 
`r round(coef(ttnc_tree)[3, 1], digits = 2)`, respectively) while the odds ratios
associated with the preferential treatment are rather similar
(`r round(coef(ttnc_tree)[2, 2], digits = 2)` and `r round(coef(ttnc_tree)[3, 2], digits = 2)`,
respectively).

Another option for assessing the class effect would be to immediately split into
all four classes rather than using recursive binary splits. This can be obtained
by setting `catsplit = "multiway"` in the `glmtree()` call above. This
yields a tree with just a single split but four kid nodes.

## References