### 2019-05-22 pkg/partykit/inst/guideliketest/draft/draft.Rnw
results of pre- and post-pruning now illustrated in one combined plot; further text revisions, mainly in sections 4 and 5
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2585 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 1c8a91c149b9ea0c7379246f2af8a36a08b12967

\documentclass[nojss]{jss}

%% packages
\usepackage{amstext,amsfonts,amsmath,amssymb,bm,thumbpdf,lmodern,hyperref,dsfont}
\usepackage[all]{hypcap}
\usepackage{enumitem}
\usepackage{color,soul}

%% need no \usepackage{Sweave} 
\SweaveOpts{engine = R, concordance = FALSE, eps = FALSE, keep.source = TRUE, eval = TRUE}

%% commands
\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
\renewcommand{\Prob}{\mathbb{P} }
\renewcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Var}{\mathbb{V}}
\newcommand{\R}{\mathbb{R} }
\newcommand{\N}{\mathbb{N} }
\newcommand{\C}{\mathbb{C} }
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}
\newcommand{\LS}{\mathcal{L}_n}
\newcommand{\TS}{\mathcal{T}_n}
\newcommand{\LSc}{\mathcal{L}_{\text{comb},n}}
\newcommand{\LSbc}{\mathcal{L}^*_{\text{comb},n}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\yn}{y_{\text{new}}}
\newcommand{\z}{\mathbf{z}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\sX}{\mathcal{X}}
\newcommand{\sY}{\mathcal{Y}}
\newcommand{\T}{\mathbf{T}}
\newcommand{\x}{\mathbf{x}}
\renewcommand{\a}{\mathbf{a}}
\newcommand{\xn}{\mathbf{x}_{\text{new}}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\ws}{\mathbf{w}_\cdot}
\renewcommand{\t}{\mathbf{t}}
\newcommand{\M}{\mathbf{M}}
\renewcommand{\vec}{\text{vec}}
\newcommand{\B}{\mathbf{B}}
\newcommand{\K}{\mathbf{K}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\bS}{\mathbf{S}}
\newcommand{\cellx}{\pi_n[\x]}
\newcommand{\partn}{\pi_n(\mathcal{L}_n)}
\newcommand{\err}{\text{Err}}
\newcommand{\ea}{\widehat{\text{Err}}^{(a)}}
\newcommand{\ecv}{\widehat{\text{Err}}^{(cv1)}}
\newcommand{\ecvten}{\widehat{\text{Err}}^{(cv10)}}
\newcommand{\eone}{\widehat{\text{Err}}^{(1)}}
\newcommand{\eplus}{\widehat{\text{Err}}^{(.632+)}}
\newcommand{\eoob}{\widehat{\text{Err}}^{(oob)}}
\newcommand{\bft}{\mathbf{t}}

\definecolor{changes}{rgb}{1,1,0.7}
\newcommand{\hlc}[1]{{\sethlcolor{changes}\hl{#1}}}

<<preliminaries, echo=FALSE, results=hide>>=
# setwd("~/svn/partykit/pkg/partykit/inst/guideliketest/draft/")
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)

library("partykit")
library("Formula")
library("parallel")
library("lattice")
library("scales")
library("colorspace")

source("../guidelike.R")
source("../tests_guideprune.R")
source("../ccprune.R")

if(file.exists("stump.rda")){
  load("stump.rda")
} else {
  simres <- simwrapper_p(nobs = 250, nrep = 100, seed = 7,
                         nrsplitvar = 1, 
                         nrsteps = 1, nrlevels = 2,
                         delta = seq(from = 0, to = 1, by = 0.1),
                         xi = c(0, 0.2, 0.5, 0.8), 
                         vary_beta = c("all", "beta0", "beta1"),
                         binary_regressor = FALSE,
                         binary_beta = c(TRUE, FALSE),
                         only_intercept = FALSE,
                         test = c("ctree", "mfluc", "ctree_max",
                                  "ctree_cat", "mfluc_cat", "ctree_max_cat",
                                  "ctree_bin", "mfluc_bin", "ctree_max_bin",
                                  "ctree_cat_bin", "mfluc_cat_bin", "ctree_max_cat_bin",
                                  "guide_sum_12", "guide_coin_12", "guide_sum_1_cor"),
                       beta0 = 0, beta1 = 1,
                       stump = TRUE, z1dist = "unif", sigma = 1, alpha = 0.05)

  save(simres, file = "stump.rda")
}
sim_stump <- simres$res
rm(simres)

if(file.exists("tree.rda")){
  load("tree.rda")
} else {
  simres <- simwrapper_p(nobs = 250, nrep = 100, seed = 7,
                         nrsplitvar = 2, nrsteps = 1, nrlevels = 2,
                         delta = seq(from = 0, to = 1, by = 0.1),
                         xi = c(0, 0.2, 0.5, 0.8), vary_beta = "all",
                         binary_regressor = FALSE, binary_beta = TRUE,
                         only_intercept = FALSE,
                         test = c("ctree", "mfluc", "ctree_max",
                                  "ctree_cat", "mfluc_cat", "ctree_max_cat",
                                  "ctree_bin", "mfluc_bin", "ctree_max_bin",
                                  "ctree_cat_bin", "mfluc_cat_bin", "ctree_max_cat_bin",
                                  "guide_sum_12", "guide_coin_12", "guide_sum_1_cor"),
                         beta0 = NULL, beta1 = NULL,
                         stump = FALSE, z1dist = "unif", sigma = 1, alpha = 0.05)

  save(simres, file = "tree.rda")
}
sim_tree <- simres$res
rm(simres)

## HCL palette
pal <- qualitative_hcl(5, "Dark 3")
names(pal) <- c("CTree", "MOB", "GUIDE", "GUIDE+scores", "CTree+max")

## HCL palette for illustrations of data generating process
pal_dgp_light <- qualitative_hcl(3, "Dark 3", c1 = 50, l1 = 75)
pal_dgp <- qualitative_hcl(3, "Dark 3")

## HCL palette for tree panels
paltrees <- rgb(c(0.9, 1), c(0.9, 1), c(0.9, 1))

@

\title{The Power of Unbiased Recursive Partitioning:\\
A Unifying View of CTree, MOB, and GUIDE}
\Shorttitle{The Power of Unbiased Recursive Partitioning}
%Comparing Testing Strategies of Unbiased Tree Algorithms

\author{Lisa Schlosser\\Universit\"at Innsbruck
\And Torsten Hothorn\\Universit\"at Z\"urich
\And Achim Zeileis\\Universit\"at Innsbruck}
\Plainauthor{Lisa Schlosser, Torsten Hothorn, Achim Zeileis}

\Abstract{
A core step of every algorithm for learning regression trees is the decision if, 
how, and where to split the underlying data -- or, in other words, the selection 
of the best splitting variable from the available covariates and the corresponding 
split point. Early tree algorithms (e.g., AID, CART) employed greedy search strategies 
that directly compare all possible split points in all available covariates. However, 
subsequent research showed that this is biased towards selection of covariates with 
more potential split points. Therefore, unbiased recursive partitioning algorithms 
have been suggested (e.g., QUEST, GUIDE, CTree, MOB) that first select the covariate 
based on statistical inference using $p$-values that are appropriately adjusted for 
the possible split points. In a second step a split point optimizing some objective 
function is selected in the chosen split variable. However, different unbiased tree 
algorithms employ different inference frameworks for computing these $p$-values and 
their relative advantages or disadvantages are not well understood, yet.

Therefore, three different popular approaches are considered here: classical categorical 
association tests (as in GUIDE), conditional inference (as in CTree), and parameter instability 
tests (as in MOB). First, these are embedded into a common inference framework encompassing
parametric model trees, in particular linear model trees. Second, it is assessed how
different building blocks from this common framework affect the power of the 
tree algorithms to select the appropriate covariates for splitting: observation-wise
goodness-of-fit measure (residuals vs.\ model scores), binarization of residuals/scores
at zero, binning of covariates. This shows that specifically the goodness-of-fit measure 
is crucial for the power of the procedures, with model scores without binarization performing
much better in many scenarios.
}
\Keywords{classification and regression trees, independence tests, recursive partitioning}

\Address{
Lisa Schlosser, Achim Zeileis \\
Universit\"at Innsbruck \\
Department of Statistics \\
Faculty of Economics and Statistics \\
Universit\"atsstr.~15 \\
6020 Innsbruck, Austria \\
E-mail: \email{Lisa.Schlosser@uibk.ac.at}, \email{Achim.Zeileis@R-project.org} \\
URL: \url{https://www.uibk.ac.at/statistics/personal/schlosser-lisa/},\\
\phantom{URL: }\url{https://eeecon.uibk.ac.at/~zeileis/}\\

Torsten Hothorn\\
Universit\"at Z\"urich \\
Institut f\"ur Epidemiologie, Biostatistik und Pr\"avention \\
Hirschengraben 84\\
CH-8001 Z\"urich, Switzerland \\
E-mail: \email{Torsten.Hothorn@R-project.org}\\
URL: \url{http://user.math.uzh.ch/hothorn/}\\

}

\begin{document}

\vspace*{0.4cm}

\section{Introduction}\label{sec:introduction}
In many situations fitting one global model to a given data set can be very challenging, 
especially if the data contains lots of different features with strong variation 
and complex interactions. Therefore, separating the data into more homogeneous subgroups 
based on a set of covariates first can simplify the task and fitting a local model to 
each of the resulting subgroups often leads to better results. This separation can be 
done by applying a tree algorithm. While almost all algorithms proposed in the literature
follow the general idea of splitting the data such that some objective function is optimized, 
they differ in their specific approaches to selecting a split variable and the corresponding 
split point. 
Some of the first tree algorithms (e.g., AID, \citealp{Morgan+Sonquist:1963}; 
CART, \citealp{Breiman+Friedman+Stone:1984}) rely on exhaustive search procedures to find both, 
the best split point and split variable in one step by directly comparing all possible split
points in all possible split variables. However, it has been shown that this is not only
computationally expensive but also biased towards split variables with many possible split
points \citep{Doyle:1973, Kim+Loh:2001}. 
Therefore, selecting a split variable in a first step and then searching for the 
best split point only within this variable in a separated second step is a more 
promising strategy as applied for example by the algorithms QUEST \citep{Loh+Shih:1997},
%% CHAID \citep{Kass:1980},
GUIDE \citep{Loh:2002}, CTree \citep{Hothorn+Hornik+Zeileis:2006} 
and MOB \citep{Zeileis+Hothorn+Hornik:2008}.
For the first step of selecting a split variable they all share the same basic concept of 
choosing the covariate which shows the highest association to the response variable 
based on $p$-values provided by a statistical test. While QUEST
%%, CHAID,
and GUIDE employ 
statistical significance tests for contingency tables, CTree applies permutation tests in 
a conditional inference framework and MOB uses fluctuation tests based on central limit theorems
for the parameter estimators. 
All these approaches have shown to work well for various situations, however, the relative
(dis)advantages of the testing strategies have not yet been investigated and compared in detail.

Therefore, in this paper the focus will be put on that first step of tree algorithms,
i.e., the task of selecting the best split variable in a given (sub)sample. In particular, the approach of the 
GUIDE algorithm is compared to the one of the CTree algorithm and the MOB algorithm
by investigating the building blocks of their testing strategies in which they differ:
(1)~Variation of the goodness-of-fit measure for the response: using residuals or full model scores.
(2)~Binarization of these residuals or scores.
(3)~Categorization of covariates.
Apart from these three main factors further aspects such as the approximation of the null
distribution (conditional vs.\ unconditional) or the type of test statistic (maximally 
selected vs.\ sum of squares) will be considered as well.
For this purpose, a unifying framework for testing strategies in unbiased model-based tree 
algorithms is presented such that each of the three strategies in GUIDE, CTree, and MOB can be 
obtained by a specific combination of the available building blocks. This allows to systematically
vary the building blocks and assess the power of the resulting inference procedure. Moreover,
it is investigated whether the performance of the inference impacts the performance of the
trees differently under pre- vs.\ post-pruning.

In many of the considered scenarios the choice of goodness-of-fit measure heavily influences
the performance of testing strategies with overall using model scores leading to clearly better
results than employing residuals only. Moreover, the original
values of the goodness-of-fit measure are preferred over binarized versions of them.
Also regarding the effects of categorizing possible split variables and the selection
of a pruning strategy clear recommendations can be given based on the presented results.

The remainder of this manuscript is structured as follows:
Section~\ref{sec:urp} reviews unbiased tree models, starting with the general algorithmic idea
(Section~\ref{sec:generic_alg}) followed by the specific algorithms CTree (Section~\ref{sec:ctree}), 
MOB (\ref{sec:mob}), and GUIDE (\ref{sec:guide}), before possible pruning strategies 
(\ref{sec:pruning}) are discussed.
In Section~\ref{sec:unify} the unifying framework for testing strategies in unbiased model-based
recursive partitioning algorithms is presented.
The setting for the simulation study is introduced in Section~\ref{sec:simulation}
and the results are illustrated and discussed in Section~\ref{sec:results}.

\section{Unbiased recursive partitioning}
\label{sec:urp}

\subsection{Generic algorithm}
\label{sec:generic_alg}
The basic idea of building a regression tree model is to partition the data into smaller 
and more homogeneous subgroups based on a set of covariates. Various tree 
algorithms have been developed, following essentially the same general structure.
Starting at the root of the tree, pertaining to the full available data sample, the
algorithms proceed in the following steps:
%
\begin{enumerate}
  \item For the current sample a (possibly simple) model is fitted by optimizing
        some objective function (or loss function) that reflects the goodness of fit.
  \item Among all available covariates one is selected as the split variable,
        choosing the split point such that goodness of fit is maximized in the
        resulting subgroups.
  \item Steps 1 and 2 are repeated within each subgroup until some stopping
        criterion is attained.
\end{enumerate}
%
The term ``model'' is used here in a very broad sense and encompasses not
only least-squares or maximum-likelihood models but also simple constant
fits such as means or average proportions.
Therefore, depending on the type of model, the employed objective function can 
for example be the sum of deviations from a typical/average value, but it can 
also be based on a model for a response along with potential covariates.
For instance, different types of residuals (or the signs thereof) can be employed
\citep[see e.g.,][]{Loh:2002} as well as rank sums or logrank scores
\citep[as in][]{Hothorn+Hornik+VanDeWiel:2006}.

As explained in Section~\ref{sec:introduction} the considered tree algorithms CTree, 
MOB, and GUIDE first select the split variable and then, in a separate step, the 
split point. For the first step of selecting a split variable they all apply statistical 
tests following the same basic strategy: 
%
\begin{enumerate}
  \item To capture how the objective function changes with the observations in
        the current subgroup, a disaggregated, observation-wise goodness-of-fit
        measure is obtained. More formally, this is an $n \times k$ matrix
        where $n \in \mathbb{N}$ is the number of observations and $k \in \mathbb{N}$ 
        the number of goodness-of-fit measures per observation.
  \item The dependency or association of this goodness-of-fit matrix with
        each possible split variable $Z_j$, $j \in \{1, \dots, m\}$, is assessed
        using some suitable test statistic. The corresponding $p$-values allow for
        a comparison of all $m$ covariates on a standardized or unified scale and in
        that way for an unbiased split variable selection.
  \item The covariate corresponding to the smallest $p$-value -- and thus the
        highest influence on the goodness of fit of the model -- is selected for 
        splitting the data into subgroups.
\end{enumerate}
%
As an example -- and explained in more detail below -- consider a linear regression tree.
Thus, a linear regression model is fitted in each subgroup, minimizing the
residual sum of squares as the aggregated goodness-of-fit measure. The corresponding
observation-wise goodness-of-fit measure can be given by the residuals or the scores
(gradient contributions). Analogously, the log-likelihood and corresponding score function could be used.

While this basic approach is the same for the GUIDE, CTree, and MOB algorithms, they differ 
in their strategies on how to calculate test statistics. In order to point out these specific 
characteristics in Section~\ref{sec:unify} the strategies of the three tree algorithms are first 
explained in more detail in Sections~\ref{sec:ctree}, \ref{sec:mob}, and \ref{sec:guide}.

\subsection{CTree}\label{sec:ctree}

The CTree algorithm \citep{Hothorn+Hornik+Zeileis:2006} is based on the idea of providing 
non-parametric regression tree models in a conditional inference framework by applying 
permutation tests. 
To select a split variable it is tested whether there is any association between the transformed 
response $h(Y)$ and each possible transformed split variable $g(Z_j)$ ($j = 1, \dots, m$).
The only requirement for the function $h$ is to depend on $Y$ in a permutation-symmetric way but
this encompasses ranks, scores, indicator functions, etc.\ and can also be multidimensional.
For a numeric response the identity function $h(Y) = Y$ is a common 
choice while a categorical response can be mapped to a unity vector by an indicator function 
$h(Y) = (0,\ldots,1,\ldots,0)^{\top}$.
Alternatively, the function $h$ can capture location and scale of $Y$ via
$h(Y) = (Y, (Y-\bar{Y})^2)^{\top}$. 
If a parametric model is fitted to the response $Y$ with regressor variable(s)~$X$,
then a model-based transformation $h(Y) = s(Y,X,\hat{\beta})$ can be used, e.g.,
the residuals in a linear model with regression coefficients $\beta$. Moreover,
$s$ can be the score function pertaining to the objective function (or loss function)
$\ell$:
$$
s(Y, X, \beta) = \frac{\partial \ell(Y, X, \beta)}{\partial \beta}
$$
The estimate of the model parameters $\hat{\beta}$ is obtained by optimizing
the sum of the objective function $\ell$, aggregated over all observations.
This framework includes many different M-type estimators as special cases, including maximum 
likelihood and ordinary least squares estimation. For a $k$-dimensional parameter
$\beta$ the score function evaluated for the $i$-th observation $s(y_i, x_i, \beta)$
is also a $k$-dimensional vector, i.e., the gradient contribution of the $i$-th observation.
Thus, the $n \times k$-matrix consisting of these scores or gradient contributions for all 
$i = 1, \dots, n$ observations is a natural candidate for the observation-wise goodness-of-fit
measure as described in the previous Section~\ref{sec:generic_alg}.

Similarly, different types of functions can be chosen for the influence function $g$ depending
on a possible split variable $Z$. A simple choice in case of $Z$ being a numeric variable is again
the identity function $g(Z) = Z$. For categorical variables $g$ can also map its values to the 
corresponding unity vectors by an indicator function $g(Z) = (0,\ldots,1,\ldots,0)$.

To test for independence of $h(Y)$ and $g(Z)$ CTree calculates a linear association test statistic,
following the framework of \cite{Strasser+Weber:1999}. The corresponding conditional expectation and
covariance given all permutations of the response variable can be calculated and used to 
standardize the test statistic. This standardized statistic has an asymptotic normal distribution
which is in fact multivariate if either of the transformation $h(Y)$ and/or $g(Z)$ is
multivariate. The actual test is carried out by mapping this standardized statistic to
the real line either by taking the absolute maximum or using a quadratic form -- with $p$-values
being computed from the analogous transformation of the normal distribution
\citep[see][for a hands-on introduction]{Hothorn+Hornik+VanDeWiel:2006}.

If both variables $Y$ and $Z$ are numeric the default independence test corresponds to a Pearson 
correlation test. For one numeric and one categorical variable essentially a 1-way ANOVA
(analysis of variance) is employed while for two categorical variables a $\chi^2$~test is performed.
Thus, in the general CTree framework many types of tests can be specified by selecting suitable
transformations $g$ and $h$. While originally
conceived for nonparametric models, it is easy to adapt CTree to model-based testing and 
recursive partitioning by choosing a model-based $h$ transformation as argued above
\citep[see the concrete examples in][]{Zeileis+Hothorn:2013,Seibold+Zeileis+Hothorn:2016}.

% In the CTree algorithm the number of possible breakpoints is fixed as splits can only be made at
% observed values or optionally at an average value lying between two observations. To select the
% best split point two-sample linear statistics of the same structure as in the independence test 
% are evaluated for all possible subsets of the sample space of the selected split variable.

\subsection{MOB}\label{sec:mob}

In contrast to CTree the MOB algorithm \citep{Zeileis+Hothorn+Hornik:2008} was explicitly
designed for a model-based goodness-of-fit measure in order to embed parametric models
into a regression tree framework. Thus, MOB is based on an objective/loss function $\ell$
and corresponding score function $s$. The original paper considered generalized linear models
(GLMs) and survival regression models but subsequently various other models have been applied as well,
including beta regression \citep{Gruen+Kosmidis+Zeileis:2012}, psychometric item response theory
models \citep{Strobl+Kopf+Zeileis:2015}, or mixed effects models \citep{Fokkema+Smits+Zeileis:2018}.
But just like CTree can be applied to parametric models, MOB conversely also encompasses
simple regression and classification trees, e.g., by choosing an intercept-only model.

For selecting a split variable MOB employs a score-based test that relies on the central limit
theorem for the parameter estimate $\hat{\beta}$. The test assesses whether the scores
-- when ordered by the potential split variable $Z$ -- fluctuate randomly around their
zero mean or differ systematically in certain subgroups. The latter would indicate a
parameter instability that could be captured by fitting separate models (optimizing $\ell$)
in the resulting subgroups. In case of a numeric split variable $Z$ both the score-based statistic
and the partitioned objective function $\ell$ are maximized over all possible splits in $Z$
(subject to certain minimal subgroup size constraints). Unlike the CTree framework, MOB
relies on classical unconditional inference. For more details see \cite{Zeileis+Hornik:2007}.

\subsection{GUIDE}\label{sec:guide}

Building on earlier work for the QUEST algorithm \citep{Loh+Shih:1997}, \cite{Loh:2002}
proposed the GUIDE algorithm blending trees with parametric regression models and
encompassing simpler classification and regression trees as special cases. Thus,
linear regression models could be fitted in the nodes of a tree as well as
constant fits such as simple mean response $\hat{\beta} = \bar{y} = \frac{1}{n}\sum_{i=1,\ldots,n}^n y_i$.
The tests for selecting the split variable are then based on the corresponding residuals,
e.g., for a simple linear regression \citep[as used in][]{Loh:2002}:
$$
r(Y,X,\hat{\beta}) = Y - \hat{\beta_0} - \hat{\beta_1} \cdot X.
$$
In subsequent work other models together with an appropriate choice of residuals
have been applied such as in
regression trees for longitudinal and multiresponse data \citep{Loh+Zheng:2013}, 
%(http://www.stat.rice.edu/~jrojo/4th-Lehmann/slides/Loh.pdf)
quantile regression models \citep{Chaudhuri+Loh:2002}, 
and proportional hazards modeling via Poisson regression \citep{Loh+He+Man:2015}, among others.

To construct a statistical test two additional transformations are carried out:
(1)~The residuals $r(Y,X,\hat{\beta})$ are binarized at zero, yielding an indicator for positive vs.\ negative residuals.
(2)~Each possible split variable $Z$ is categorized, i.e., unless $Z$ is already categorical
it is split at its quartiles into four bins.
Subsequently, a $\chi^2$~test of independence is performed for the binarized residuals and
each categorized/categorical split variable.
After choosing the split variable showing the highest dependency by yielding the lowest 
$p$-value, the split point minimizing the overall goodness-of-fit measure is selected. 
Note that the split point in numeric variables $Z$ is searched over all possible splits,
not just the four bins that were constructed for the $\chi^2$~test.

%\newpage
\subsection{Pruning}
\label{sec:pruning}

To avoid overfitting recursive partitioning algorithms need to assure that trees do not grow
too large. Apart from certain minimal subgroup size or maximal tree depth constraints this is
classically accomplished by so-called ``pruning'' approaches. As the three tree algorithms
considered here (CTree, MOB, and GUIDE) differ in their default choice of pruning approach,
we briefly discuss these here. However, as all three algorithms can in principle be combined
with any of the pruning approaches this is done only relatively briefly.

The classical CART algorithm \citep{Breiman+Friedman+Stone:1984} proposed to first grow a large
tree and then prune those splits in the tree that did not increase predictive performance in a
cross-validation. This is also know as \emph{post-pruning} (after growing the initial tree) and
more specifically cost-complexity pruning.

In the unbiased recursive partitioning literature this post-pruning approach is also used frequently
\citep[e.g., in][]{Loh+Vanichsetakul:1988,Loh+Shih:1997,Kim+Loh:2001,Loh:2002} and the $p$-values
from the association tests are only employed for selecting the split variable on a unified scale.
However, \cite{Hothorn+Hornik+Zeileis:2006} proposed to also use these $p$-values for a 
so-called \emph{pre-pruning} strategy which stops growing the tree as soon as no significant association
can be found in a given subgroup. This approach is the default in CTree and also in
MOB. However, \cite{Zeileis+Hothorn+Hornik:2008} also pointed out that a natural strategy for
post-pruning in model-based partitioning is to use information criteria such as AIC (Akaike information
criterion) or BIC (Bayes information criterion), following the ideas of \cite{Su+Wang+Fan:2004}.

Clearly, for inference-based pre-pruning it is crucial that the association tests employed for
split selection work well as statistical significance test, i.e., conform with their nominal size
and have high power. In contrast, when using post-pruning (either based on cross-validation or
information criteria) it might not be as crucial that the significance test works well and has
high power.

Due to these considerations we first evaluate the significance tests underlying CTree, MOB,
and GUIDE by themselves, i.e., without growing an actual tree in combination with a pruning strategy.
Subsequently we combine the tests with a cost-complexity post-pruning approach in order to assess
whether shortcomings of the tests are mitigated by pruning.

\section{Unifying framework}\label{sec:unify}

Each of the algorithms CTree, GUIDE, and MOB can be characterized by its combination
of the type of model fits, tests, and pruning strategy employed to
grow the tree. Table~\ref{tab:combinations} provides an overview of the default
combinations.
However, as discussed above, subsequent publications have emphasized that all three
algorithms can be combined with different model fitting approaches and to some degree
different pruning strategies have been explored as well. Thus, the class of tests
employed for the unbiased splitting variable selection forms the core of each of the
algorithms: conditional inference (CTree) vs.\ score-based fluctuation tests (MOB)
vs.\ residual-based $\chi^2$ tests (GUIDE). Therefore, we consider a standard class
of model fits (namely, linear regression trees) and investigate the relative
advantages and disadvantages of the tests themselves (which are essential to pre-pruning)
as well as the combination with post-pruning. Thus, subsequently
the names CTree, MOB, and GUIDE distinguish the test-based variable and split selection
rather than the entire algorithm with all default settings.

\begin{table}[t!]
\centering
\begin{tabular}{ l l l l }
\hline 
      & Fit             & Test                     & Pruning \\ 
\hline
CTree & Non-parametric  & Conditional inference    & Pre\\
MOB   & Parametric      & Score-based fluctuation  & Pre (or post with AIC/BIC)\\
GUIDE & Parametric      & Residual-based $\chi^2$  & Post (cost-complexity pruning)\\
\hline
\end{tabular}
\caption{Default combinations of fitted model type, test type and pruning strategy
for the algorithms CTree, MOB, and GUIDE.
\label{tab:combinations}}
\end{table}

\subsection{Building blocks of testing strategies}\label{sec:buildingblocks}

Even though CTree, MOB, and GUIDE differ in the specific tests they apply, their approaches for split
variable selection follow the same basic structure as explained in Section~\ref{sec:generic_alg}. 
In fact, the tests can be embedded in a unifying conceptual framework that yields the
different tests by combining various building blocks.
These mostly differ in the way the model for the dependent variable on the one hand and
the splitting variables on the other are prepared or transformed:
%
\begin{itemize}
  \item \emph{Goodness-of-fit measure:}\\
        Different variations of the disaggregated, observation-wise goodness-of-fit measure of the 
        model for the response $Y$ and possible covariates $X$ can be considered. Either residuals
	$r(Y,X,\hat{\beta})$ can be used as proposed for GUIDE or model scores $s(Y,X,\hat{\beta})$
        as proposed for MOB. All three algorithms can, in principle, use both goodness-of-fit
	measures though, which is probably brought out most clearly in the CTree algorithm that
        explicitly allowed for different transformations $h(Y)$ in its original description already.
  \item \emph{Binarization of residuals/scores:}\\
        Rather than testing independence between the split variables and the residuals/scores
	themselves, it is possible	
	to binarize residuals/scores at $0$ so that only their signs are assessed (as proposed for GUIDE).
  \item \emph{Categorization of split variables:}\\
        Similarly, the split variables can also be categorized (for testing only). This was proposed for
	GUIDE, employing binning at the quartiles yielding four categories of approximately
	equal size.
\end{itemize}
%
The three algorithms combine these building blocks in different ways as shown in Table~\ref{tab:buildingblocks}:
When applying CTree for unbiased model-based recursive partitioning it has been suggested to use
the model scores without binarization and assess their association with the untransformed split
variables using a conditional inference test. This is similar to a squared correlation test statistic.
MOB also employs the scores without
binarization and maximally selects a score statistic over all potential split points in the
split variable. GUIDE employs the binarized residuals and assesses their association with
the categorized split variable in a classical (unconditional) $\chi^2$ test.

But the building blocks could be easily re-combined to yield new types of tests. For example,
in the GUIDE approach, a 1-way ANOVA can be used for assessing the association of the residuals
(without binarization) with the categorized split variable. Or alternatively, a multivariate
1-way ANOVA can be used for the non-binarized scores as opposed to the residuals etc.

Note that there are further differences in the testing strategies between the three algorithms, e.g.,
using conditional vs.\ unconditional approximations of the null distributions. However, this difference
has relatively little influence compared to the other building blocks considered in detail here.
Moreover, both similarities and relative differences between these approaches have been previously
discussed, e.g., in \cite{Hothorn+Zeileis:2008} and \cite{Zeileis+Hothorn:2013}.

\begin{table}[t!]
\centering
\begin{tabular}{ l l l l l }
\hline 
      & Scores       & Binarization  & Categorization  & Statistic\\
\hline
CTree & Model scores & --	     & -- 	       & Sum of squares\\
MOB   & Model scores & --	     & -- 	       & Maximally selected\\
GUIDE & Residuals    & \checkmark    & \checkmark      & Sum of squares\\
\hline
\end{tabular}
\caption{Testing strategies of CTree, MOB, and GUIDE with the corresponding
setting of the building blocks in the unifying framework and the type of test statistic.
\label{tab:buildingblocks}}
\end{table}

%\newpage

\subsection{Linear model tree}\label{sec:lmtrees}

To focus on the unified testing framework, as described in the previous section, we employ
the same model fits for all three algorithms. To do so, we employ linear regression models
because it is such a basic and widely used model and linear model trees were the leading
illustrations in both the original MOB \citep{Zeileis+Hothorn+Hornik:2008} and GUIDE \citep{Loh:2002} papers.
However, the conclusions drawn from this example also hold for many other model types.

To fix notation, we consider the following models for the simulation 
study in Section~\ref{sec:simulation}:
$$
Y = \beta_0 + \beta_1 \cdot X + \epsilon
$$
with response variable $Y$, regressor variable $X$, and error term $\epsilon$. This model is
fitted, as usual, by ordinary least squares (OLS) to the observations in each notation,
yielding the parameter estimates $\boldsymbol{\hat{\beta}} = (\hat{\beta_0}, \hat{\beta_1})^\top$.
To keep notation simple we present the following equations for the root node with all observations
$\{(y_i,x_i)\}_{i=1,\ldots,n}$ and $n \in \mathbb{N}$. In subsequent nodes the same equations
are used but just for a smaller subgroup.

The aggregated goodness-of-fit measure (or objective or loss function) in OLS estimation
is the sum of squared residuals: $\sum_{i=1}^n \ell(y_i,x_i,\beta_0,\beta_1)$ where
$$
\ell(y_i,x_i,\beta_0,\beta_1) = r(y_i,x_i,\beta_0,\beta_1)^2
$$
is the squared residual which is defined as
$$
r(y_i,x_i,\beta_0,\beta_1) = y_i - \beta_0 - \beta_1 \cdot x_i.
$$
These residuals can also intuitively be used as the corresponding disaggregated
observation-wise goodness-of-fit measure. Another natural candidate for this is the
score function for the $i$-th observation:
$$
s(y_i, x_i, \beta_0, \beta_1) = 
\frac{\partial \ell}{\partial \boldsymbol{\beta}}(y_i, x_i, \beta_0, \beta_1) = 
-2 \cdot r(y_i,x_i, \beta_0, \beta_1) \cdot (1,x_i)^\top
$$
Thus, up to a constant scaling factor of $-2$ (that could also be omitted) the first component
of the scores is in fact the residual. However, it is complemented by a second component
that captures the slope effect of $x_i$. Therefore, when computing the score for all $n$ observations
this yields an $n \times 2$ score matrix whose first column corresponds to the $n$ residuals.
As this is the derivative with respect to the intercept parameter~$\beta_0$, it
captures changes in the intercept. Moreover, the second column of the score matrix contains
derivatives with respect to the slope parameter~$\beta_1$ and thus captures changes in this.

\section{Simulation setting and evaluation}
\label{sec:simulation}

In this simulation study two different scenarios are considered for the linear model trees
presented in Section~\ref{sec:lmtrees}. First, the underlying tree structure based on which the
data is generated is a stump, i.e., a tree with only one split (``stump'' scenario, see 
Figure~\ref{fig:dgp_stump}). 
By keeping the tree so simple we can focus on the testing strategy only giving focus to their
power in terms of selecting the correct split variable.

In the second scenario the true tree structure contains two splits in two different variables
yielding a tree with three terminal nodes (``tree'' scenario, see Figure~\ref{fig:dgp_tree}). 
It employs the same basic structure as the first scenario but simply adds another split. This 
allows to evaluate the power of the three testing strategies in a more complex setting, in 
combination with using a post-pruning strategy.

\subsection{Data generating process}

\subsubsection{``Stump'' scenario}
Each generated data set consists of the response and regressor variables, one true split 
variable and nine noise split variables as listed in Table~\ref{tab:variables} 
together with the corresponding distributions.

\begin{table}
\centering
\begin{tabular}{l l l}
\hline
Name                    & Notation                   & Specification \\
\hline
\textit{Variables:}     &                            & \\ 
Response                & $Y$                        & $=\beta_0(Z_1) + \beta_1(Z_1) \cdot X + \epsilon$ \\
Regressor               & $X$                        & $\mathcal{U}([-1,1])$ \\
Error                   & $\epsilon$                 & $\mathcal{N}(0,1)$ \\
True split variable     & $Z_1$                      & $\mathcal{U}([-1,1])$\\
Noise split variables   & $Z_2, Z_3, \dots, Z_{10}$  & $\mathcal{U}([-1,1])$  or $\mathcal{N}(0, 1)$ (alternating) \\
\hline
\textit{Parameters/functions:} &            & \\ 
Intercept                      & $\beta_0$  & $0$ or $\pm \delta$\\
Slope                          & $\beta_1$  & $1$ or $\pm \delta$\\
True split point               & $\xi$      & $\in \{0, 0.2, 0.5, 0.8\}$ \\
Effect size                    & $\delta$   & $\in \{0, 0.1, 0.2, \ldots , 1\}$ \\
\hline
\end{tabular}
\caption{\label{tab:variables}Variables included in the data generating process 
as used for the ``stump'' scenario. In the ``tree'' scenario $Z_2$ is also a true split variable, 
not only $Z_1$, and hence $\beta_0$ and $\beta_1$ are functions $\beta_i(Z_1,Z_2)$, $i=1,2$.}
\end{table}

The location parameter $\mu$ of the normally distributed response variable $Y$ depends
linearly on the regressor variable $X$.
% which is a uniformly distributed variable taking values between $-1$ and $1$.
% can either be a binary variable (taking either $0$ or $1$) or a uniformly distributed 
% variable (taking values between $-1$ and $1$).
Moreover, the intercept $\beta_0$ and/or the slope parameter $\beta_1$
can depend on the true split variable $Z_1$. More specifically, three different variations
are considered for the coefficients $\beta_0$ and $\beta_1$:
%
\begin{enumerate}
  \item The intercept $\beta_0$ varies depending on $Z_1$ while $\beta_1$ is fixed (at $1$).
  \item The slope coefficient $\beta_1$ varies depending on $Z_1$ while $\beta_0$ is fixed 
        (at $0$).
  \item Both coefficients $\beta_0$ and $\beta_1$ vary depending on $Z_1$.
\end{enumerate}
%
See Figure~\ref{fig:dgp_stump} for an illustration. More precisely, the coefficient that
changes ($\beta_j$ with $j\in\{0,1\}$) switches between two values at the split point $\xi$:
%
\begin{align*}
\beta_{i}(Z_1) = \begin{cases}
-\delta \cdot (-1)^{j}  \quad \text{if } Z_1 \leq \xi \\
+\delta \cdot (-1)^{j}  \quad \text{if } Z_1 > \xi
\end{cases}
\end{align*}
%
%while in the continuous case a simple linear function is considered in the form of
%$$\beta_{i}(Z_1) = Z_1 \cdot \delta \cdot (-1)^{i}.$$
In that way, the type of variation is the same for $\beta_0$ and $\beta_1$, however, in 
opposite directions.

<<dgp, echo=FALSE>>=
set.seed(7)

nobs <- 250
sigma <- 0.5
xi <- 0
delta <- 1
beta0 <- 0
beta1 <- 1

d <- dgp_stump(nobs = nobs, binary_beta = TRUE, binary_regressor = FALSE, only_intercept = FALSE,
               vary_beta = "beta0", beta1 = beta1, delta = delta, xi = xi, sigma = sigma)

sp_xi1 <- partysplit(3L, breaks = 0)

pn <- partynode(1L, split = sp_xi1, 
                kids = list(
                  partynode(2L, info = c("true parameters:", 
                                         expression(beta[0] ==~" "~0~" or "~-delta), 
                                         expression(beta[1] ==~" "~1~" or "~+delta))),
                  partynode(3L, info = c("true parameters:", 
                                         expression(beta[0] ==~" "~0~" or "~+delta), 
                                         expression(beta[1] ==~" "~1~" or "~-delta)))))

p <- party(pn, d)

ep <- function (obj, digits = 3, abbreviate = FALSE, justmin = Inf, 
                just = c("alternate", "increasing", "decreasing", "equal"), 
                fill = "white") 
{
  meta <- obj$data
  justfun <- function(i, split) {
    myjust <- if (mean(nchar(split)) > justmin) {
      match.arg(just, c("alternate", "increasing", "decreasing", "equal"))
    } else {
      "equal"
    }
    k <- length(split)
    rval <- switch(myjust, equal = rep.int(0, k), 
                   alternate = rep(c(0.5, -0.5), length.out = k), 
                   increasing = seq(from = -k/2, to = k/2, by = 1), 
                   decreasing = seq(from = k/2, to = -k/2, by = -1))
    unit(0.5, "npc") + unit(rval[i], "lines")
  }
  function(node, i) {
    split <-  c("<= xi", "> xi")
    y <- justfun(i, split)
    split <- split[i]
    if (any(grep(">", split) > 0) | any(grep("<", split) > 0)) {
      tr <- suppressWarnings(try(parse(text = paste("phantom(0)", split)), silent = TRUE))
      if (!inherits(tr, "try-error")) split <- tr
    }
    grid.rect(y = y, gp = gpar(fill = fill, col = 0), width = unit(1, "strwidth", split))
    grid.text(split, y = y, just = "center")
  }
}
class(ep) <- "grapcon_generator"

tp <- function (obj, digits = 3, abbreviate = FALSE, 
                fill = c("lightgray", "white"), id = TRUE, 
                just = c("center", "top"), top = 0.85, 
                align = c("center", "left", "right"), gp = NULL, FUN = NULL, 
                height = NULL, width = NULL) 
{
  nam <- names(obj)
  extract_label <- function(node) formatinfo_node(node, FUN = FUN, 
                                                  default = c("terminal", "node"))
  maxstr <- function(node) {
    lab <- extract_label(node)
    klab <- if (is.terminal(node)) 
      ""
    else unlist(lapply(kids_node(node), maxstr))
    lab <- c(lab, klab)
    lab <- try(unlist(lapply(lab, function(x) strsplit(x, 
                                                       "\n"))), silent = TRUE)
    if (inherits(lab, "try-error")) {
      paste(rep("a", 9L), collapse = "")
    }
    else {
      return(lab[which.max(nchar(lab))])
    }
  }
  nstr <- if (is.null(width)) 
    maxstr(node_party(obj))
  else paste(rep("a", width), collapse = "")
  just <- match.arg(just[1L], c("center", "centre", "top"))
  if (just == "centre") 
    just <- "center"
  align <- match.arg(align[1L], c("center", "centre", "left", 
                                  "right"))
  if (align == "centre") 
    align <- "center"
  rval <- function(node) {
    if(node$id == 2) fill <- c("white", pal_dgp_light[1])
    if(node$id == 3 | node$id == 4) fill <- c("white", pal_dgp_light[2])
    if(node$id == 5) fill <- c("white", pal_dgp_light[3])
    fill <- rep(fill, length.out = 2)
    lab <- extract_label(node)
    if (!is.null(gp)) {
      outer_vp <- viewport(gp = gp)
      pushViewport(outer_vp)
    }
    if (is.null(height)) 
      height <- length(lab) + 1L
    node_vp <- viewport(x = unit(0.5, "npc"), y = unit(if (just == 
                                                           "top") 
      top
      else 0.5, "npc"), just = c("center", just), width = unit(1, 
                                                               "strwidth", nstr) * 1.1, height = unit(height, "lines"), 
      name = paste("node_terminal", id_node(node), sep = ""), 
      gp = if (is.null(gp)) 
        gpar()
      else gp)
    pushViewport(node_vp)
    grid.rect(gp = gpar(fill = fill[1]))
    for (i in seq_along(lab)) grid.text(x = switch(align, 
                                                   center = unit(0.5, "npc"), left = unit(1, "strwidth", 
                                                                                          "a"), right = unit(1, "npc") - unit(1, "strwidth", 
                                                                                                                              "a")), y = unit(length(lab) - i + 1, "lines"), 
                                        lab[i], just = align)
    if (id) {
      nodeIDvp <- viewport(x = unit(0.5, "npc"), y = unit(1, 
                                                          "npc"), width = max(unit(1, "lines"), unit(1.3, 
                                                                                                     "strwidth", nam[id_node(node)])), height = max(unit(1, 
                                                                                                                                                         "lines"), unit(1.3, "strheight", nam[id_node(node)])))
      pushViewport(nodeIDvp)
      grid.rect(gp = gpar(fill = fill[2], lty = "solid"))
      grid.text(nam[id_node(node)])
      popViewport()
    }
    if (is.null(gp)) 
      upViewport()
    else upViewport(2)
  }
  return(rval)
}

class(tp) <- "grapcon_generator"

@

\begin{figure}%[t]
\setkeys{Gin}{width=1\linewidth}
\minipage{0.45\textwidth}
<<dgp_stump1, fig=TRUE, echo=FALSE, width=4, height=3>>=
#par(mar=c(3,3,0,0), oma = c(3,3,0,0))
plot(p, 
     tp_args = list(FUN = identity, width = 15), 
     ip_args = list(fill = paltrees[c(1, 2)]),
     drop_terminal = TRUE, tnex = 1, edge_panel = ep, terminal_panel = tp)
@
\endminipage
\minipage{0.45\textwidth}
<<dgp_stump2, fig=TRUE, echo=FALSE, width=4, height=3>>=
par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "y", xlab = "x", 
     main = expression("varying"~beta[0]),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
abline(coef = c(-1,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(1,1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>xi], ylim = c(-5,5), pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = 3.2)
mtext(expression(beta[1]== 1), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = 2.2)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 0.2)
mtext(expression(beta[1]== 1), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = -0.8)
@
\endminipage

\minipage{0.45\textwidth}
<<dgp_stump3, fig=TRUE, echo=FALSE, width=4, height=3>>=
d <- dgp_stump(nobs = nobs, binary_beta = TRUE, binary_regressor = FALSE, only_intercept = FALSE,
               vary_beta = "beta1", beta0 = beta0, delta = delta, xi = xi, sigma = sigma)

par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "y", xlab = "x", main = expression("varying"~beta[1]),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
abline(coef = c(0,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(0,-1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>xi], ylim = c(-5,5), pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -1)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 2)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1)
@
\endminipage
\minipage{0.45\textwidth}
<<dgp_stump4, fig=TRUE, echo=FALSE, width=4, height=3>>=
d <- dgp_stump(nobs = nobs, binary_beta = TRUE, binary_regressor = FALSE, only_intercept = FALSE,
               vary_beta = "all", delta = delta, xi = xi, sigma = sigma)
par(mar = c(4,4,2,4))
plot(x = d$x[d$z1<=xi], y = d$y[d$z1<=xi], ylim = c(-5,5), ylab = "y", xlab = "x",
     main = expression("varying"~beta[0]~"and"~beta[1]),
     xlim = c(-1,1.1),
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
abline(coef = c(-1,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(1,-1), col = pal_dgp[2], lwd = 2)
points(x = d$x[d$z1>xi], y = d$y[d$z1>=xi], pch = 19, col = alpha(pal_dgp[2], 0.4), cex = 0.7)
legend("top", c(expression(z[1]<=xi), expression(z[1]>xi)), col = c(pal_dgp[1], pal_dgp[2]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1.7)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 0.7)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -0.7)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -1.7)
@
\endminipage
\caption{\label{fig:dgp_stump}{Top left panel: True stump structure applied in the data generating process for the ``stump'' scenario with the true model parameters $\beta_0$
and $\beta_1$, either varying or being fixed at $0$ or $1$ respectively.
Top right and bottom panels: Bivariate plot of the response $Y$ on regressor $X$ 
with effect size $\delta = 1$ and for three variations: 
varying $\beta_0$ and fixed $\beta_1$ (top right),
fixed $\beta_0$ and varying $\beta_1$ (bottom left),
varying $\beta_0$ and varying $\beta_1$ (bottom right).}}
\end{figure}

\subsubsection{``Tree'' scenario}
For the second simulation scenario (``tree'' scenario) the same basic setting with the same 
variables as in the ``stump'' scenario is applied (see Table~\ref{tab:variables}).
However, not only $Z_1$ but also $Z_2$ is used as a true split variable. Therefore, an additional split
is preformed yielding a tree with three terminal nodes. The first split (in $Z_2$, at~$\xi$) induces a 
change in the slope parameter $\beta_1$ while the second split (in $Z_1$, also at~$\xi$) corresponds 
to a change in the intercept $\beta_0$.
Hence, contrary to the ``stump'' scenario where three variations are considered for the parameters 
$\beta_0$ and $\beta_1$, only this one variation with both parameters varying is investigated for 
the ``tree'' scenario. In particular, the parameters depend on the split variables in the following way:
%
\begin{align*}
\beta_{0}(Z_1, Z_2) = \begin{cases}
0       \qquad \text{if } Z_2 \leq \xi \\
-\delta  \quad \text{ if } Z_1 \leq \xi \land Z_2 >\xi \\
+\delta  \quad \text{ if } Z_1 > \xi \land Z_2 >\xi
\end{cases}
\end{align*}
\begin{align*}
\beta_{1}(Z_1, Z_2) = \beta_{1}(Z_2)= \begin{cases}
+\delta \quad \text{if } Z_2 \leq \xi \\
-\delta \quad \text{if } Z_2 > \xi
\end{cases}
\end{align*}
%
The underlying tree structure for the data generating process is illustrated in 
Figure~\ref{fig:dgp_tree}.

<<true_tree_dgp, eval = TRUE, echo=FALSE, results=hide>>=

# d <- dgp_tree(nobs = 250)
set.seed(7)
d <- dgp_tree(nobs = 250, xi = 0, binary_regressor = FALSE, vary_beta = "all", 
              binary_beta = TRUE, nrsplitvar = 2, sigma = 0.5)

sp_xi2 <- partysplit(4L, breaks = 0)
sp_xi1 <- partysplit(3L, breaks = 0)

pn <- partynode(1L, split = sp_xi2, 
                kids = list(
                  partynode(2L, info = c("true parameters:", 
                                         expression(beta[0] == 0), 
                                         expression(beta[1] == +delta))),
                  partynode(3L, split = sp_xi1, 
                            kids = list(
                              partynode(4L, info = c("true parameters:", 
                                                     expression(beta[0] == -delta), 
                                                     expression(beta[1] == -delta))),
                              partynode(5L, info = c("true parameters:", 
                                                     expression(beta[0] == +delta), 
                                                     expression(beta[1] == -delta)))))))

p <- party(pn, d)

#ctest <- evaltests(y~x|z1+z2+z3+z4+z5+z6+z7+z8+z9+z10, data = d, testfun = "ctree", stump = FALSE)
@

\begin{figure}%[t]
\begin{center}
\setkeys{Gin}{width=1\linewidth}
\minipage{0.5\textwidth}
<<dgp_tree1, fig=TRUE, echo=FALSE, width=5, height=4.5>>=
par(mar=c(0,0,0,0), oma=c(0,0,0.4,0))
plot(p, edge_panel = ep, terminal_panel = tp,
     tp_args = list(FUN = identity, width = 15), 
     ip_args = list(fill = paltrees[c(1, 2)]),
     drop_terminal = TRUE, tnex = 1)
@
\endminipage
\minipage{0.45\textwidth}
<<dgp_tree2, fig=TRUE, echo=FALSE, width=4, height=4>>=
par(mar = c(4,4,2,4))
plot(x = d$x[d$z2<=xi], y = d$y[d$z2<=xi], ylim = c(-5,5), ylab = "y", xlab = "x",
     pch = 19, col = alpha(pal_dgp[1], 0.4), cex = 0.7)
abline(coef = c(0,1), col = pal_dgp[1], lwd = 2)
abline(coef = c(-1,-1), col = pal_dgp[2], lwd = 2)
abline(coef = c(1,-1), col = pal_dgp[3], lwd = 2)
points(x = d$x[d$z2>xi & d$z1<=xi], y = d$y[d$z2>xi & d$z1<=xi], pch = 19, 
       col = alpha(pal_dgp[2], 0.4), cex = 0.7)
points(x = d$x[d$z2>xi & d$z1>xi], y = d$y[d$z2>xi & d$z1>xi], pch = 19, 
       col = alpha(pal_dgp[3], 0.4), cex = 0.7)
legend("top", c(expression(z[2]<=xi), 
                expression(z[2]>xi~" and "~z[1]<=xi),
                expression(z[2]>xi~" and "~z[1]>xi)), 
       col = c(pal_dgp[1], pal_dgp[2], pal_dgp[3]), pch = c(19), bty = "n", cex = 0.9)
mtext(expression(beta[0]==0), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 2)
mtext(expression(beta[1]==+delta), side = 4, col = pal_dgp[1], las = 1, line = 0.5, at = 1.3)
mtext(expression(beta[0]==+delta), side = 4, col = pal_dgp[3], las = 1, line = 0.5, at = 0.1)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[3], las = 1, line = 0.5, at = -0.6)
mtext(expression(beta[0]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2)
mtext(expression(beta[1]==-delta), side = 4, col = pal_dgp[2], las = 1, line = 0.5, at = -2.7)
@
\endminipage
\caption{\label{fig:dgp_tree}Left panel: True tree structure
applied in the data generating process for the ``tree'' scenario 
with the true model parameters $\beta_0$ and $\beta_1$ for each terminal panel.
Right panel: Bivariate plot of the response $Y$ on regressor $X$ 
with effect size $\delta = 1$.}
\end{center}
\end{figure}

\subsection{Evaluation}
\label{sec:evaluation}
The testing strategies are evaluated over a stepwise increasing effect size $\delta$, 
on 100 replications per step each consisting of 250 observations. 
To compare the performance of the evaluated testing strategies for each step 
in the ``stump'' scenario the following criteria are considered:
the $p$-values pertaining only to the true split variable $Z_1$;
and the proportion of replications for which the $p$-value of $Z_1$
is the lowest and significant at 5\% level (i.e., where $Z_1$ would be
selected for splitting in a pre-pruning approach), denoted by the 
``selection probability''. In that way, the power of the
considered statistical tests can be compared as they are all applied
as significance tests answering two questions at once:
(1)~Should a split be performed at all? 
(2)~If so, in which variable?
For the first question the $p$-values regarding all available covariates are compared 
to a predefined level of significance $\alpha = 0.05$. Only if the smallest $p$-value is smaller
than $\alpha$ a split is performed and the covariate corresponding to this $p$-value is selected.

For the ``tree'' scenario the adjusted Rand index~(ARI) is calculated 
as a measure of similarity between the true tree structure and the fitted model tree.

As explained before, the aim of this simulation study is to investigate the effects
of each particular building block of the unifying framework presented in Section~\ref{sec:unify} 
rather than the whole testing strategies. Therefore, other combinations as presented in 
Table~\ref{tab:buildingblocks}, hence adapted versions of GUIDE, CTREE, and MOB are evaluated as well
as their original versions.

\section{Results}
\label{sec:results}

In the following we first investigate the properties of the testing
strategies in the ``stump'' scenario, focusing on the tests' $p$-values
for the true split variable $Z_1$ and its corresponding selection
probability (i.e., $Z_1$ being significant at 5\% level and having the
lowest $p$-value among all split variables). Section~\ref{sec:scores} begins by
highlighting the importance of using full model scores vs.\ residuals only
before Section~\ref{sec:3way} considers all building blocks (scores vs.\ residuals,
binarization of these, and categorization of the split variable).
Subsequently, the ``tree'' scenario is employed to investigate how the
performance of the tests affects growing the trees overall. This is
evaluated using the adjusted Rand index for trees grown by pre-pruning and 
cost-complexity post-pruning (Section~\ref{sec:tree}).

% To illustrate the effects of the building blocks in the unifying framework presented in 
% Section~\ref{sec:unify} different combinations of them are investigated for the first
% simulation scenario (stump) in Sections~\ref{sec:scores} and~\ref{sec:3way}. 
% Then the testing strategies are evaluated on the more
% complex tree structure of the second scenario in Section~\ref{sec:tree}. 
% The corresponding results are shown in the following plots.

\subsection[``Stump'' scenario: Residuals vs. full model scores]{``Stump'' scenario: Residuals vs.\ full model scores}
\label{sec:scores}

A crucial difference between the testing strategies of CTree/MOB and GUIDE
is the difference between using only the residuals vs. the full model
scores. In the literature on structural change tests there is evidence
that residual-based tests can only capture parameter differences that
affect the conditional mean \citep[see e.g.,][ and further discussion in 
Section~\ref{sec:discussion}]{Ploberger+Kraemer:1992}.
Hence we compare the CTree, MOB, and GUIDE algorithms  -- all three using the 
default specification as shown in Table~\ref{tab:buildingblocks} -- and 
additionally consider a new GUIDE flavor, denoted GUIDE+scores, that uses
binarized scores rather than residuals.

In Figure~\ref{fig:scores_stump} the performances of the testing strategies are 
represented by the corresponding selection probability, hence, a significance level 
is incorporated. To show that in general the drawn conclusions do not change if no 
significance level is included, a different illustration of the exact same evaluation 
is provided in Appendix~\ref{app:significance} where the performances are compared based 
on the smallest $p$-values, regardless of whether they are beneath any significance level.

Moreover, as defined by the true tree structure used in the data generating process, the
parameter functions $\beta_i(Z_1)$, $i=1,2$, are step functions which is particularly 
advantageous for the maximum selection in the test statistic applied by MOB while the linear 
selection in CTree is preferable for detecting monotonous dependencies. To show that employing 
a continuous parameter function instead of a step function only has an impact on whether CTree or
MOB performs better the exact same simulation setting as used for Figure~\ref{fig:scores_stump} 
but with continuously changing parameters has been evaluated and is illustrated in 
Appendix~\ref{app:contbeta}.

\begin{figure}
\setkeys{Gin}{width=\linewidth}
<<scores_stump, fig=TRUE, echo=FALSE, height=5, width=8>>=
subdata <- subset(sim_stump, test %in% c("ctree","mfluc",
                                         "guide_sum_12", "guide_sum_1_cor"))
subdata$test <- factor(subdata$test, 
                       levels = c("ctree","mfluc",
                                  "guide_sum_12", "guide_sum_1_cor"),
                       labels = c("CTree","MOB",
                                  "GUIDE+scores", "GUIDE"))

subdataxi <- subset(subdata, xi %in% c(0,0.8))

par.settings <- list(superpose.symbol = list(col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")], 
                                             fill = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")]), 
                     superpose.line = list(col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")]))

library("latticeExtra")
useOuterStrips(xyplot(prop_Tsplit ~ delta | vary_beta + xi, groups = ~ test, 
                      data = subdataxi, 
                      subset = binary_beta == TRUE & binary_regressor == FALSE & 
                        only_intercept == FALSE,# & xi == 0, 
                      type = "l", lty = c(1,1,1,1),
                      lwd = 2,
                      scales=list(x=list(at=seq(1,11,2))),
                      key =  list(text = list(levels(subdataxi$test)),
                                  lines = list(lty = c(1,1,1,1), type = "l", pch = 1, cex = 0.8,
                                               col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")], lwd = 2)),
                      layout= c(3,2),
                      index.cond = list(c(2,3,1),c(2,1)),
                      xlab = expression(delta),
                      ylab = expression("Selection probability of"~Z[1]),
                      par.settings = par.settings),

               strip =strip.custom(factor.levels = c(expression("varying"~beta[0]~"and"~beta[1]), 
                                                     expression("varying"~beta[0]), 
                                                     expression("varying"~beta[1])),
                                   bg = "gray90"),
               strip.left=strip.custom(factor.levels = c(expression(xi == 0~" (50%)"),                                                              expression(xi == 0.8~" (90%)" )),
                                   bg = "gray90"))

@
\caption{\label{fig:scores_stump}
Selection probability of the true split variable $Z_1$ for
testing strategies CTree, MOB, GUIDE, and GUIDE+scores in the ``stump''
scenario. Probabilities are estimated over increasing effect size $\delta$
with 100 replications of 250 observations per step. The true split point
in $Z_1$ is either the median 0 (top) or the 90\% quantile 0.8 (bottom)
for either varying intercept $\beta_0$ (left), varying slope $\beta_1$
(middle), or both coefficients varying (right).}
\end{figure}

Figure~\ref{fig:scores_stump} shows that overall CTree and MOB (both using full 
model scores) perform clearly better than the original GUIDE version (using residuals 
only) as they accomplish to select the true split variable $Z_1$ in a higher proportion
of replications. Even though the advantage of CTree and MOB evens out in certain situations
such as for a high effect size (see, e.g., for the median $0$ and varying $\beta_0$), 
GUIDE is never ahead of them in any of the considered settings. Moreover, the positive 
effect of including all available scores can be observed particularly well in the top middle 
panel where GUIDE+scores detects a shift in the slope parameter $\beta_1$ as opposed to GUIDE. 
%The fact that residual-based methods fail to detect changes that do not effect the
%expected value has already been elaborated in the literature of structural change
%analysis \citep[see, e.g.,][]{Ploberger+Kraemer:1992} and is also discussed in 
%Section~\ref{sec:discussion}. 
Furthermore, this advantage of including full models scores 
clearly outweighs the cost involved which is displayed by the small gain of GUIDE compared 
to GUIDE+scores in case of a fixed slope parameter (left panels).

For a central split point ($\xi=0$ corresponding to the median, upper panels) all strategies 
perform similarly well as long as the intercept varies (left and right panel) with CTree and 
MOB being only slightly ahead. However, for the split point at the $90$\% quantile $0.8$ 
(lower panels) the performance of all of the applied strategies decreases. 
While both GUIDE versions struggle to detect the correct split variable even for a high effect 
size, MOB is clearly ahead with CTree leading to the second best results.
As discussed before, this advantage of MOB over CTree is mainly due to the abrupt shift in the 
model parameters $\beta_0$ and $\beta_1$ and turns into an advantage of CTree over MOB for 
continuously changing parameters (see Appendix~\ref{app:contbeta}).
From the fact that GUIDE and GUIDE+scores are both performing worse than MOB and CTree it 
can be deduced that categorizing split variables and binarizing scores has a negative effect 
on the ability to detect the right split variable. To elaborate these effects more in detail 
the results for all possible combinations of the building blocks are discussed in a full 
factorial 3-way analysis in the following section.

\subsection{``Stump'' scenario: Full factorial analysis of building blocks}\label{sec:3way}
To investigate the effects of each of the building blocks separately the most general 
case of the ``stump'' scenario where the intercept and 
slope parameter are both varying is considered. In the underlying evaluation all 
possible combinations of the building blocks have been included. The different levels of 
each of the three building blocks are listed in Table~\ref{tab:levels} where 
$h$ and $g$ refer to the transformation functions applied to the response $Y$ or
a split variable $Z$ respectively, both as described in Section~\ref{sec:ctree}. 
(See Table~\ref{tab:combinations_appendix} in Appendix~\ref{app:combinations} 
for a more detailed overview of all 12 combinations.)
If quartiles are used as breakpoints to categorize a split variable the employed
function $g$ maps to a four-dimensional vector of zeros except for one value $1$
indication the bin to which an observation is assigned to. For a maximum selection
of a split variable $g$ maps to a vector containing only zeros until a certain
changepoint from which on the values are replace by $1$. This breakpoint indicates
for example a crossed quantile. In linear selection $g$ is simply the identity function.

\begin{table}[h!]
\centering
\begin{tabular}{ l l l }
\hline
Building block               & Levels              & Transformation \\
\hline
Residuals vs.\ scores        & \textit{residuals}  & $r(Y,X,\hat{\beta})$\\
\quad ($h$ transformation)   & \textit{scores}     & $s(Y,X,\hat{\beta})$\\
\hline
Binarization                 & \textit{yes}        & $\mathds{1}_{[0,\infty)}(h)$\\
\quad (of $h$)               & \textit{no}         & $h$ without binarization\\
\hline
Categorization               & \textit{cat}        & $g(Z)=(0,1,0,0)$\\
\quad ($g$ transformation)   &                     & \quad indicating the assigned bin\\
                             & \textit{max}        & $g(Z)=(0,\ldots,0,1,\ldots,1)$\\
                             &                     & \quad indicating the crossed quantile\\
                             & \textit{lin}        & $g(Z)=Z$\\
\hline
\end{tabular}
\caption{Available levels of the building blocks.
\label{tab:levels}}
\end{table}

Based on the results displayed in Figure~\ref{fig:3way} it can be stated that binarizing 
residuals/scores decreases the performance as it leads to higher $p$-values corresponding 
to the true split variable $Z_1$ and thus to lower power in the presented settings. 
For a central true split point ($\xi=0$, left panel) this effect remains at an almost constant 
level regardless of whether possible split variables are categorized or not. However, in case 
of the true split point at the $90$\% quantile $0.8$ (right panel), categorizing split 
variables increases the $p$-value of $Z_1$ even more. 
Here, the maximum selection performs best while the linear selection shows its disadvantage in 
detecting abrupt changes. Overall, it depends on the situation whether a linear or a maximum
selection of a split variable is advantageous. But unless one of the breakpoints used for binning 
possible split variables is close to the true split point (as for the median $0$, left panel),
this categorization weakens the performance of a testing strategy. 
While the effect of both of these transformations of the data (binarization of 
residuals/scores and categorization of split variables) can be observed
separately in Figure~\ref{fig:3way}, combining them increases the negative impact 
remarkably.

As already shown in Section~\ref{sec:scores} the use of scores vs.\ residuals has a minor 
effect if there is a change in the intercept which is
supported by the small differences between the dashed lines (scores) and the solid
lines (residuals) in Figure~\ref{fig:3way}.

Note that the choice of the two specific settings 
(left panel: true split point at the median $0$,  effect size $0.3$; 
right panel: true split point at the $90$\% quantile $0.8$ requiring a higher effect 
size $1$)
has been made based on the results shown in Figure~\ref{fig:scores_stump} in order to make 
the differences particularly well observable. (Additional evaluations for increasing $\delta$ 
as in Section~\ref{sec:scores} can be found in Appendix~\ref{app:increasingdelta}.)

<<3way_data, results=hide, echo=FALSE>>=
if(file.exists("3way_xi00_delta03.rda")){
  load("3way_xi00_delta03.rda")
} else {
  simres <- sim(nobs = 250, nrep = 100, seed = 7, stump = TRUE, nrsteps = 1,
                delta = 0.3,
                xi = 0, 
                binary_regressor = FALSE,
                binary_beta = TRUE,
                only_intercept = FALSE,
                nrlevels = 2, nrsplitvar = 1,
                vary_beta = "all", beta0 = 0, beta1 = 1, alpha = 0.05,
                return_matrices = TRUE,
                test = c("guide_sum_1_cor",
                         "mfluc_resid_bin",
                         "ctree_resid_bin",
                         "mfluc_resid_cat",
                         "mfluc_resid",
                         "ctree_resid",
                         "guide_sum_12",
                         "mfluc_bin",
                         "ctree_bin",
                         "mfluc_cat",
                         "mfluc",
                         "ctree"))
  d00 <- prep_3way(simres)
  save(d00, file = "3way_xi00_delta03.rda")
}

if(file.exists("3way_xi08_delta1.rda")){
  load("3way_xi08_delta1.rda")
} else {
  simres <- sim(nobs = 250, nrep = 100, seed = 7, stump = TRUE, nrsteps = 1,
                delta = 1,
                xi = 0.8, 
                binary_regressor = FALSE,
                binary_beta = TRUE,
                only_intercept = FALSE,
                nrlevels = 2, nrsplitvar = 1,
                vary_beta = "all", beta0 = 0, beta1 = 1, alpha = 0.05,
                return_matrices = TRUE,
                test = c("guide_sum_1_cor",
                         "mfluc_resid_bin",
                         "ctree_resid_bin",
                         "mfluc_resid_cat",
                         "mfluc_resid",
                         "ctree_resid",
                         "guide_sum_12",
                         "mfluc_bin",
                         "ctree_bin",
                         "mfluc_cat",
                         "mfluc",
                         "ctree"))
  d08 <- prep_3way(simres)
  save(d08, file = "3way_xi08_delta1.rda")
}

d00$xi <- 0
d08$xi <- 0.8

d3way <- rbind(d00, d08)
d3way$xi <- factor(d3way$xi, levels = c(0,0.8), labels = c(0,0.8))
@

\begin{figure}%[t]
\setkeys{Gin}{width=1\linewidth}
<<3way, fig=TRUE, echo=FALSE, height=4, width=8>>=
xyplot(pval_z1 ~ cat | xi, 
       groups = interaction(bin, res_scores), 
       data = d3way, 
       ylim = c(0,0.5),
       type = c("a"), 
       col = c(pal[1], pal[3], pal[1], pal[3]),
       lty = c(1,1,2,2),
       lwd = 2,
       scales = list(alternating=FALSE),
       xlab = "Categorization",
       ylab = expression("p-value of"~Z[1]),
       #index.cond = list(c(1,2,3,4)),
       strip = strip.custom(factor.levels = c(expression(xi==0~"(50%)"~" and"~delta==0.3), 
                                              expression(xi==0.8~"(90%)"~" and"~delta==1)),
                                   bg = "gray90"),
       key = list(text = list(c("Goodness of fit:", "  residuals", "  scores", "Binarization:", "  yes", "  no")),
                  columns = 2,
                  lines = list(lty = c(0,1,2,0,1,1), 
                               type = "l", lwd = 2,
                               col = c("black","black","black","black", pal[1], pal[3]))))

#par(mar = c(3,5,1,5.7))
#with(d00, interaction.plot(res_scores, interaction(cat, bin), pval_z1, ylim = c(0,1),
#                           ylab = expression("Mean"~p~"-value for"~Z[1])))
#with(d00, interaction.plot(interaction(cat, bin), res_scores, pval_z1))
#with(d00, interaction.plot(cat, interaction(bin, res_scores), pval_z1))
#with(d08, interaction.plot(res_scores, interaction(cat, bin), pval_z1, ylim = c(0,1)))
#with(d08, interaction.plot(interaction(cat, bin), res_scores, pval_z1))
#with(d08, interaction.plot(cat, interaction(bin, res_scores), pval_z1))
@
\caption{\label{fig:3way} Effects of the three building blocks \textit{residuals vs.\ scores},
\textit{binarization}, and \textit{categorization} in the ``stump'' scenario with the intercept
and the slope parameter both varying. All possible combinations have been 
evaluated in two different settings and their performances are compared based on the 
mean $p$-values corresponding to the true split variable $Z_1$. 
Left panel: true split point at the median $0$ and with
effect size $0.3$.
Right panel: true split point at the $90$\% quantile $0.8$ and with effect 
size $1$.}
\end{figure}

%\newpage
\subsection[``Tree'' scenario: Pre-pruning vs. post-pruning]{``Tree'' scenario: Pre-pruning vs.\ post-pruning}
\label{sec:tree}
To compare the performance of testing strategies in a more complex scenario they have
been evaluated on data sets based on the tree structure with two different split variables 
($Z_1$ and $Z_2$) both with the same split point $\xi$ as illustrated in 
Figure~\ref{fig:dgp_tree}.
The results presented in the top row of Figure~\ref{fig:tree_all} have been obtained applying pre-pruning.
Therefore, based on the corresponding ARI values conclusion can again be drawn on the power of the testing strategies 
as significance tests. CTree and MOB perform almost equally well for the true split points $0$, $0.2$ and $0.5$ 
($50$\%, $60$\%, and $75$\% quantile). GUIDE+scores performs slightly worse but it still achieves an ARI above 
$0.5$ for the highest considered effect size of $\delta = 1$. 
However, the original version of GUIDE is clearly behind which is mainly due to the fact that it doesn't detect 
the first split in $Z_2$ corresponding to a changing slope parameter as it only considers residuals in its testing 
strategy. 
For a split point farther away from the center such as the $90$\% quantile $0.8$ MOB 
outperforms all other methods and both GUIDE versions fail to find the correct tree structure.
Hence, the insights provided by the top row of Figure~\ref{fig:tree_all} conform with the conclusions
drawn from the ``stump'' scenario: Including all available scores in the testing strategy leads to a better performance 
while binarizing scores and categorizing split variables both do not cause any improvement but in many situations even 
have a negative effect on the power of a testing strategy.

<<tree_all_data, echo=FALSE, results=hide>>=
if(file.exists("pruning.rda")){
  load("pruning.rda")
} else {
  simres <- simwrapper_p(nobs = 250, nrep = 100, seed = 7, 
                         nrsteps = 1, nrlevels = 2,
                         delta = seq(from = 0, to = 1, by = 0.1),
                         xi = c(0, 0.2, 0.5, 0.8), 
                         vary_beta = "all", binary_regressor = FALSE,
                         binary_beta = TRUE, only_intercept = FALSE,
                         compare_pruning = TRUE,
                         test = c("ctree", "mfluc",
                                  "guide_sum_12", 
                                  "guide_sum_1_cor"),
                         beta0 = NULL, beta1 = NULL,
                         stump = FALSE, z1dist = "unif", sigma = 1, alpha = 0.05,
                         nrsplitvar = 2)

  save(simres, file = "pruning.rda")  
}

pr <- simres$res
## extract data for xyplot comparing ARI and ARI_p
pr <- pr[,c(1,2,7,8,9,12,15)]
## extended data set with post-pruning results being represented as separated tests
prex <- pr[,c(1,2,3,4,6)]
prex <- rbind(prex,prex)

prex$test <- factor(prex$test,
                    levels = c("ctree","mfluc",
                               "guide_sum_12", "guide_sum_1_cor"),
                    labels = c("CTree","MOB",
                               "GUIDE+scores", "GUIDE"))

prex$nrsubgr[(length(pr$nrsubgr)+1) : (2*length(pr$nrsubgr))] <- pr$nrsubgr_p
prex$ari[(length(pr$ari)+1) : (2*length(pr$ari))] <- pr$ari_p
prex$pruning <- c(rep("pre", NROW(pr)), rep("post", NROW(pr)))

par.settings <- list(superpose.symbol = list(col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")], 
                                             fill = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")]), 
                     superpose.line = list(col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")]))

@

\begin{figure}%[t]
\setkeys{Gin}{width=\linewidth}
<<tree_all, fig=TRUE, echo=FALSE, height=5, width=8>>=
library("latticeExtra")
useOuterStrips(xyplot(ari ~ delta | xi + pruning, groups = ~ test, 
                      data = prex, 
                      type = "l", lty = c(1,1,1,1),
                      lwd = 2,
                      key =  list(text = list(levels(prex$test)[1:4]),
                                  lines = list(lty = c(1,1,1,1), type = "l", pch = 1, cex = 0.8,
                                               col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")], lwd = 2)),
                      layout= c(4,2),
                      scales=list(x=list(at=seq(1,11,2))),
                      xlab = expression(delta),
                      ylab = "Adjusted Rand Index",
                      #index.cond = list(c(2,3,1),c(2,1)),
                      par.settings = par.settings),
               strip = strip.custom(factor.levels = c(expression(xi==0), 
                                                      expression(xi==0.2), 
                                                      expression(xi==0.5), 
                                                      expression(xi==0.8)),
                                    bg = "gray90"),
               strip.left=strip.custom(factor.levels = c("Post-pruning", "Pre-pruning"),
                                   bg = "gray90"))
@
\caption{\label{fig:tree_all}
Adjusted Rand index (ARI) for the testing strategies CTree, MOB, GUIDE, and GUIDE+scores in the
``tree'' scenario, once applying pre-pruning (top row) and once post-pruning (bottom row). 
The ARI is evaluated over
increasing effect size $\delta$ with 100 replications of 250 observations per step.
Each panel represents a different value of the true split point $\xi$ ($0$, $0.2$, $0.5$ and 
$0.8$ from left to right, corresponding to the $50$\%, $60$\%, $75$\%, and $90$\% quantile).
}
\end{figure}

To elaborate the impact of the choice of a pruning strategy the presented tree models
are built again in the exact same simulation setting but employing post-pruning (i.e., building very 
large trees without any significance level and applying cost-complexity pruning as in the CART 
algorithm \citep{Breiman+Friedman+Stone:1984}). 
The corresponding ARI values are illustrated in the bottom row of Figure~\ref{fig:tree_all}
allowing for a comparison to those obtained with pre-pruning (top row).

For MOB and CTree it has been shown that the tests work very well as significance tests.
Therefore, it can be recommended to take advantage of this power and use the information
given by the tests for pre-pruning. As it can be seen in Figure~\ref{fig:tree_all}
in most cases the results of these two testing strategies do not vary significantly over 
a change of the applied pruning strategy.
On the contrary, overall the tests in GUIDE have less power. For that reason, the pruning
strategy should not be based on their results which argues for the application of post-pruning.
This is also supported by the results in Figure~\ref{fig:tree_all} showing that post-pruning improves
the performance of both GUIDE versions. Particularly if the true split point is close to a binning 
point (such that categorizing split variables has no negative effect) the post-pruning strategy 
compensates the lack of power of GUIDE+scores compared to CTree and MOB almost completely (bottom 
row, first and third panel).
In conclusion, the choice of a pruning strategy should be based on knowledge about the 
power of the applied testing strategy. With a well working significance test additional expensive
calculations such as needed for post-pruning may be avoided being one of the main advantages of 
pre-pruning. However, deficiencies of a testing strategy in terms of power can be at least partly 
evened out by employing a post-pruning strategy.

%\newpage

\section{Discussion}
\label{sec:discussion}

From comparing the effects of the building blocks three main conclusions can be drawn: 
(1)~In many cases including all available score values rather than just residuals leads to 
clearly better results and performs in general at least equally well.
(2)~Binarizing scores/residuals does not lead to any improvement in performance.
(3)~Categorizing possible split variables also doesn't improve the performance 
unless the breakpoints used for the categorization are very close to the true split points.
However, even then the results are very similar to those achieved without any categorization.
But it has to be noted that a categorization of possible split variables can simplify
evaluations. Moreover, depending on the focus of interest, the results can be more
straightforward, for example in case of clearly separated subgroups being defined by the values 
of a continuous split variable. Therefore, categorizing possible split variable can have
advantages in practice, however, it has to be considered that in general
important effects might be neglected.

Linear models as presented in this study have been selected as they allow for a simple illustration
of the investigated subjects. However, the obtained results also hold for a variety of other models
as the introduced building blocks can be applied in the same way for all model types.
Specific characteristics of linear models regarding structural changes have been investigated
in detail in previous literature.
\cite{Ploberger+Kraemer:1992} have shown that residual-based tests can detect a change in the parameters
only if it also causes a shift in the expected value $\mathbb{E}(Y)$. 
This is not the case if changes are orthogonal to the mean regressor $(1,0)^{\top}$, hence, if
the parameter vector $\bm{\beta}$ is shifted 
%to $\bm{\beta} + \bm{\Delta}$ 
by $\bm{\Delta} = (0, \Delta_1)^{\top}$.
For that reason GUIDE breaks down when the slope parameter $\beta_1$ changes while the intercept
$\beta_0$ is fixed as seen in the middle panel of Figure~\ref{fig:scores_stump}.
While this might be an extreme example similar scenarios can appear in many applications, see, e.g., 
Figure~2 in \cite{Loh+He+Man:2015} where an additional separation of the data in each node is
required in order to fit a reasonable model. This can be avoided by including model scores
allowing for the detection of changes not only in the location parameter but also in all other 
parameters, i.e., not only in the expected value but also in the variance.
In situations where shifts appear only in the intercept testing procedures purely based on residuals 
perform best. However, the cost in terms of power that comes with including all scores, in particular
with testing more than one parameter, is negligibly small.

Also in more general models residual-based and score-based procedures are expected to perform equally 
well if all parameters are highly correlated. On the other hand, if parameters are orthogonal changes 
that are not connected to the mean might be missed when only considering residuals. This can be prevented 
by including scores which are a standard measurement of goodness of fit that can be easily employed for 
all model types.
Even though it also depends on the type of the applied residuals 
(see, e.g., \citealp{Loh+Zheng:2013}, \citealp{Chaudhuri+Loh:2002} 
and \citealp{Loh+He+Man:2015})
whether parameter changes can be detected successfully, a one-dimensional measurement of goodness of 
fit such as residuals provides less power than full model scores. 
Whether this difference affects the performance of a testing strategy and if so, to what extent, is 
determined by the way how residuals are connected to parameter shifts as well as the type of parameter
shift and the type of model. Therefore, a general statement for all types of models and residuals can 
hardly be made. However, the results from previous literature such as \cite{Ploberger+Kraemer:1992} and
the here presented simulation show that considering only residuals provides small gains in certain 
situations but remarkable deficits in general. On the contrary, the benefits of including full model 
scores clearly outweigh the price of of testing all parameters.

%Furthermore, with only two parameters determining a linear model the importance of including scores 
%regarding each parameter could already be pointed out clearly while for an increasing number of 
%parameters in a model this aspect becomes even more important.

%\newpage

\section*{Computational details}
\label{sec:comp}
The applied implementation is based on the \textsf{R} package \textbf{partykit}
(version~\Sexpr{packageVersion("partykit")}) which is available 
on \textsf{R}-Forge at (\url{https://R-Forge.R-project.org/projects/partykit/})
where the code to reproduce the simulation study can be found as well.

The functions \code{ctree} and \code{mob} provide an implementation of the
two tree algorithms in their original form. For their adapted versions additional
modifications have been applied within these functions allowing for a categorization
of possible split variables and a binarization of scores.
To evaluate the GUIDE algorithm a reimplementation of this algorithm has been 
established/constructed using the basic framework of \code{ctree} and \code{mob}.

\section*{Acknowledgments}
\hlc{An extended research stay of Torsten Hothorn in Innsbruck (August
2017 to January 2018) was financially supported by the Swiss National
Science Foundation, grant number SNF IZSEZ0 177091.}

\bibliography{ref.bib}

\newpage

\begin{appendix}

\section{Test statistics}
This section provides further information on the test statistics applied in CTree, MOB, 
and GUIDE. The presented details have been extracted from the original papers 
\cite{Hothorn+Hornik+Zeileis:2006}, \cite{Zeileis+Hothorn+Hornik:2008}, 
and \cite{Loh:2002}, however, notation is adapted to the main manuscript
in order to allow for better comparison.

\subsection{CTree}
To measure the association of a response $Y$ and each possible split variable $Z$ the
CTree algorithm applies a linear test statistic $T$ which is excerpted from Section~3.1.
(``Variable selection and stopping criteria'') of the original paper \citep{Hothorn+Hornik+Zeileis:2006} and is of the following form:
$$
T(Y, Z, w) = \text{vec}\left( \sum_{i=1}^n w_i g(Z_{i})h(Y_i, (Y_1,\ldots,Y_n))^{\top}\right)
$$
where $w$ are optional weights, $g: \mathcal{Z} \rightarrow \mathbb{R}^{p}$ is a nonrandom transformation of the split variable $Z$ and the influence function $h: \mathcal{Y} \times \mathcal{Y}^n \rightarrow \mathbb{R}^q$ depends on the responses $(Y_1, \ldots, Y_n)$ in a permutation symmetric way. Moreover, by applying the ``vec'' operator the resulting $p~\times~q$ matrix is converted into a $p q$ column vector.
Following \cite{Strasser+Weber:1999}
the conditional expectation $\mu$ and covariance $\Sigma$ of a linear test statistic can 
be calculated and used to standardize an observed linear test statistic $t$ within a 
function $c$ mapping into the real line.
For example, the maximum of the absolute values of the standardized linear statistic
$$
c_{\max}(t,\mu,\Sigma) = \max_{k=1,\ldots,pq} \left| \frac{(t-\mu)}{\sqrt{(\Sigma)_{kk}}} \right|
$$
or a quadratic form
$$
c_{\text{quad}}(t, \mu, \Sigma) = (t - \mu)\Sigma^{+} (t - \mu)^{\top}
$$
can be considered where $\Sigma^{+}$ is the Moore-Penrose inverse of $\Sigma$.
Since the asymptotic conditional distribution of a linear test statistic $T$ is a 
multivariate normal with parameters $\mu$ and $\Sigma$ \citep{Strasser+Weber:1999}, 
the asymptotic distribution of $c_{\text{max}}$ is normal while the quadratic form 
$c_{\text{quad}}$ follows an asymptotic $\chi^2$~distribution. Based on this knowledge 
the corresponding $p$-values can be calculated easily.

\subsection{MOB}
The MOB algorithm employs an empirical fluctuation process $W$ to measure deviations of
the model scores $s$ from zero. As described in detail in Section~3.2. (``Testing for 
parameter instability'') of the original paper \citep{Zeileis+Hothorn+Hornik:2008}
this process is of the following form:
$$
W(t) = \hat{J}^{-1/2} n^{-1/2} \sum_{i=1}^{\lfloor nt \rfloor} s_{\sigma(Z_{i})} \qquad (0 \leq t \leq 1)
$$
where the model scores $s$ are sorted by the split variable Z by including the ordering permutation $\sigma(Z_{i})$. 
To scale this partial sum process an estimate $\hat{J}$ of the covariance matrix $\text{cov}(s(Y,X,\hat{\beta}))$ is included.
Following \cite{Zeileis+Hornik:2007} $W(t)$ converges to a Brownian bridge $W^0$ under the null
hypothesis of parameter stability. To obtain a test statistic a scalar functional $\lambda(\cdot)$ capturing the fluctuation in the empirical process can be applied and the corresponding asymptotic distribution of $\lambda(W(t))$ can be obtained by employing the same functional to the limit process, i.e., $\lambda(W^0)$. 
For more details, particularly on calculating the corresponding $p$-values see Section~3.2. of the original paper \citep{Zeileis+Hothorn+Hornik:2008}.

\subsection{GUIDE}
The test statistic of the $\chi^2$~test as applied in the GUIDE algorithm is
$$
X^2 = \frac{(O_{11} - E_{11})^2}{E_{11}} + \ldots + \frac{(O_{11} - E_{11})^2}{E_{11}}
$$
where $O_{ij}$ are the observed frequencies of observations in a certain combination of binarized residuals ($i=1,2$) and categorized split variables ($j = 1,\ldots,4$). $E_{ij}$ are the corresponding expected frequencies under independence $E_{ij} = \frac{O_{i \cdot} \cdot O_{\cdot j}}{n}$. Under the null hypothesis of independence the asymptotic distribution of $X^2$ is a $\chi^2$~distribution allowing for a straight-forward calculation of $p$-values.
If full model scores are used instead of residuals a sum of the test statistic over the number of distribution parameters is considered such that each 
summand corresponds to one column of the score matrix.

\section{Combinations of building blocks applied}\label{app:combinations}

For the evaluation of all 12 possible combinations of the presented building blocks
the original testing strategies CTree, MOB, and GUIDE have been applied together
with modified versions of them. The employed versions and the corresponding combination
of building blocks are listed in Table~\ref{tab:combinations_appendix}.
Each suffix indicates a change of the original form of the considered testing strategy.
In particular, it relates to the transformations of the data before applying a statistical
test as represented by the different building blocks. In that way,
``+res'' refers to using only residuals (instead of all scores),
``+scores'' to using all scores (instead of only residuals),
``+cat'' to applying a categorization of split variables,
``+bin'' to applying a binarization of scores/residuals.

\begin{table}[h!]
\centering
\begin{tabular}{ l l l l }
\hline 
Testing Strategy  & Residuals/Scores   & Binarization & Categorization \\
\hline
GUIDE             & residuals          & bin          & cat \\
MOB+bin+res     & residuals          & bin          & max \\
CTree+bin+res   & residuals          & bin          & lin \\
MOB+cat+res    & residuals          & lin          & cat \\
MOB+res          & residuals          & lin          & max \\
CTree+res        & residuals          & lin          & lin \\
GUIDE+scores     & scores             & bin          & cat \\
MOB+bin          & scores             & bin          & max \\
CTree+bin        & scores             & bin          & lin \\
MOB+cat          & scores             & lin          & cat \\
MOB               & scores             & lin          & max \\
CTree             & scores             & lin          & lin \\
\hline
\end{tabular}
\caption{Testing strategies applied (representing the 12 combinations
of building blocks).
\label{tab:combinations_appendix}}
\end{table}

\section{Significance level}\label{app:significance}

\begin{figure}[t]
\setkeys{Gin}{width=\linewidth}
<<scores_stump_wosign, fig=TRUE, echo=FALSE, height=5, width=8>>=
subdata <- subset(sim_stump, test %in% c("ctree", "ctree_max", "mfluc", "guide_sum_12", "guide_sum_1_cor"))
subdata$test <- factor(subdata$test, 
                       levels = c("ctree", "ctree_max", "mfluc",
                                  "guide_sum_12", "guide_sum_1_cor"),
                       labels = c("CTree", "CTree+max", "MOB",
                                  "GUIDE+scores", "GUIDE"))

subdataxi <- subset(subdata, xi %in% c(0,0.8))

par.settings <- list(superpose.symbol = list(col = pal[c("CTree", "CTree+max", "MOB", "GUIDE+scores", "GUIDE")], 
                                             fill = pal[c("CTree", "CTree+max", "MOB", "GUIDE+scores", "GUIDE")]), 
                     superpose.line = list(col = pal[c("CTree", "CTree+max", "MOB", "GUIDE+scores", "GUIDE")]))

library("latticeExtra")
useOuterStrips(xyplot(prop_z1 ~ delta | vary_beta + xi, groups = ~ test, 
                      data = subdataxi, 
                      subset = binary_beta == TRUE & binary_regressor == FALSE & 
                        only_intercept == FALSE,# & xi == 0, 
                      type = "l", lty = c(1,1,1,1,1),
                      lwd = 2,
                      scales=list(x=list(at=seq(1,11,2))),
                      key =  list(text = list(levels(subdataxi$test)),
                                  lines = list(lty = c(1,1,1,1,1), type = "l", pch = 1, cex = 0.8,
                                               col = pal[c("CTree", "CTree+max", "MOB", "GUIDE+scores", "GUIDE")], lwd = 2)),
                      layout= c(3,2),
                      index.cond = list(c(2,3,1),c(2,1)),
                      xlab = expression(delta),
                      ylab = expression("Prop. of"~Z[1]~"having lowest p-value"),
                      par.settings = par.settings),

               strip =strip.custom(factor.levels = c(expression("varying"~beta[0]~"and"~beta[1]), 
                                                     expression("varying"~beta[0]), 
                                                     expression("varying"~beta[1])),
                                   bg = "gray90"),
               strip.left=strip.custom(factor.levels = c(expression(xi == 0~" (50%)"),                                                              expression(xi == 0.8~" (90%)")),
                                   bg = "gray90"))

@
\caption{\label{fig:scores_stump_wosign}
Proportion of replications where the true split variable $Z_1$ shows the smallest $p$-value, 
hence, where $Z_1$ is selected without any significance level in the ``stump'' scenario.
The testing strategies CTree, CTree+max, MOB, GUIDE, and GUIDE+scores are evaluated 
over increasing effect size $\delta$ with 100 replications of 250 observations per step. 
The true split point in $Z_1$ is either the median 0 (top) or the 90\% quantile 0.8 (bottom)
for either varying intercept $\beta_0$ (left), varying slope $\beta_1$
(middle), or both coefficients varying (right).}
\end{figure}

In Section~\ref{sec:scores} the investigated testing strategies are compared based
on the number of replications in which the true split variable $Z_1$ is detected with a $p$-value
smaller than the predefined significance level $\alpha = 0.05$. This choice of measurement
is due to the aim of investigating how well the tests perform as significance test.
However, ignoring this significance level yields the same conclusions 
which can be seen when comparing Figures~\ref{fig:scores_stump_wosign}
and \ref{fig:scores_stump}.
In Figure~\ref{fig:scores_stump_wosign} the proportion of replications 
in which the true splitting variable $Z_1$ is detected, but not necessarily with a $p$-value 
smaller than the significance level, is illustrated. In Figure~\ref{fig:scores_stump} 
the exact same simulation setting is applied with a significance level.

\section{Simulation results for continuous model parameters}
\label{app:contbeta}

In the simulation study presented in Sections~\ref{sec:simulation} and \ref{sec:results} the intercept
and slope parameters $\beta_0$ and $\beta_1$ are both either fixed or binary variables taking
either the positive or the negative value of the effect size $\delta$. Therefore, these parameters
are step functions for which MOB has shown to perform better than CTree. However, in case of
monotonous functions CTree is advantageous as it is constructed to detect monotonous effects.
To point out that the difference in performance of CTree and MOB depends on the type of effect
the same setting of which the results are presented in Figure~\ref{fig:scores_stump} is
evaluated again but this time with the varying parameter(s) changing continuously.
In particular, the parameters $\beta_0$ and $\beta_1$ are linear functions of the true
split variable $Z_1$:
\begin{align*}
\beta_{i}(Z_1) = \begin{cases}
-\delta \cdot (-1)^{i} \cdot Z_1\\
+\delta \cdot (-1)^{i} \cdot Z_1.
\end{cases}
\end{align*}
Moreover, a modified version of CTree considering maximum values in the test statistic is evaluated
(denoted by CTree+max). 
Looking at the results presented in Figure~\ref{fig:stump_contbeta} it can be observed that in this setting the 
original version of CTree is ahead while CTree+max and MOB perform almost equally well. Hence, the comparison of 
Figures \ref{fig:scores_stump} and \ref{fig:stump_contbeta} points out that it clearly depends on the type of 
effect whether the maximal selection (MOB, CTree+max) or the linear selection (CTree) is advantageous.

\begin{figure}%%[t]
\setkeys{Gin}{width=\linewidth}
<<contbeta, fig=TRUE, echo=FALSE, height=4, width=8>>=
subdata <- subset(sim_stump, test %in% c("ctree", "ctree_max","mfluc", "guide_sum_1_cor","guide_sum_12"))
#subdata <- subset(subdata, delta %in% c(0, 0.1, 0.2, 0.3, 0.4, 0.5))
subdata$test <- factor(subdata$test, 
                       levels = c("ctree", "ctree_max","mfluc", "guide_sum_1_cor","guide_sum_12"),
                       labels = c("CTree","CTree+max","MOB", "GUIDE","GUIDE+scores"))
par.settings <- list(superpose.symbol = list(col = pal[c("CTree","CTree+max","MOB", "GUIDE","GUIDE+scores")], 
                                             fill = pal[c("CTree","CTree+max","MOB", "GUIDE","GUIDE+scores")]), 
                     superpose.line = list(col = pal[c("CTree","CTree+max","MOB", "GUIDE","GUIDE+scores")]))

xyplot(prop_Tsplit ~ delta | vary_beta, groups = ~ test, 
       data = subdata, 
       subset = binary_beta == FALSE & binary_regressor == FALSE & 
         only_intercept == FALSE, 
       scales=list(x=list(at=seq(1,11,2))),
       type = "l", 
       layout= c(3,1),
       lwd = 2,
       key =  list(text = list(levels(subdata$test)),
                   lines = list(lty = c(1,1,1,1,1), 
                                type = "l", 
                                col = pal[c("CTree","CTree+max","MOB", "GUIDE","GUIDE+scores")], 
                                lwd = 2)), 
       xlab = expression(delta),
       ylab = expression("Selection probability of"~Z[1]),
       index.cond = list(c(2,3,1)),
       strip = strip.custom(factor.levels = c(expression("varying"~beta[0]~" and "~beta[1]),
                                              expression("varying"~beta[0]), 
                                              expression("varying"~beta[1])),
                                    bg = "gray90"),
       par.settings = par.settings)
@
\caption{\label{fig:stump_contbeta}
Selection probability of the true split variable $Z_1$ for testing strategies CTree, CTree+max,
MOB, GUIDE, and GUIDE+scores for continuous parameter functions $\beta_0(Z_1)$ and $\beta_1(Z_1)$.
Probabilities are estimated over increasing effect size $\delta$
with 100 replications of 250 observations per step. 
Each panel represents one variation for the coefficients: 
varying intercept $\beta_0$ (left), varying slope parameter $\beta_1$ (middle) and 
both coefficients varying (right).}
\end{figure}

\section{Results for increasing effect size}
\label{app:increasingdelta}

\subsection{Binarization of residuals/scores}
\label{app:binarization}
To elaborate over an increasing effect size whether a binarization of the residuals/scores at $0$ 
leads to an improvement or a deterioration of performance CTree and MOB are applied, once in their 
original version without a binarization and once in an adapted version with a binarization of 
scores at $0$ (CTree+bin, MOB+bin). Moreover, they are compared to the adapted GUIDE version 
which includes all available scores (GUIDE+scores).
%Hence, the five strategies are obtained from the unifying framework by setting the switches 
%as listed in Table~\ref{tab:buildingblocks_appendix}.

Figure~\ref{fig:bin_stump} shows the effect of binarizing the score values over
different values for the true split point $\xi$, however, all four situations
lead to the same conclusions. Binarizing the score values decreases the proportion
of replications where the split variable was selected correctly.

\begin{figure}%[t]
\setkeys{Gin}{width=\linewidth}
<<bin_stump, fig=TRUE, echo=FALSE, height=4, width=8>>=
subdata <- subset(sim_stump, test %in% c("ctree", "ctree_bin", 
                                         #"ctree_max", "ctree_max_bin",
                                         "mfluc", "mfluc_bin",
                                         "guide_sum_12"))

subdata <- subset(subdata, delta %in% c(0, 0.1, 0.2, 0.3, 0.4, 0.5))
subdata$test <- factor(subdata$test, 
                       levels = c("ctree", "ctree_bin", 
                                  #"ctree_max", "ctree_max_bin",
                                  "mfluc", "mfluc_bin",
                                  "guide_sum_12"),
                       labels = c("CTree","CTree_bin",
                                  #"CTree+max","CTree+max_bin",
                                  "MOB", "MOB_bin",
                                  "GUIDE+scores"))

par.settings <- list(superpose.symbol = list(col = pal[c("CTree", "CTree", 
                                                         #"CTree+max", "CTree+max", 
                                                         "MOB", "MOB",
                                                         "GUIDE+scores")], 
                                             fill = pal[c("CTree", "CTree", 
                                                          #"CTree+max", "CTree+max", 
                                                          "MOB", "MOB",
                                                          "GUIDE+scores")]), 
                     superpose.line = list(col = pal[c("CTree", "CTree", 
                                                       #"CTree+max", "CTree+max", 
                                                       "MOB", "MOB",
                                                       "GUIDE+scores")]))

xyplot(prop_Tsplit ~ delta | xi, groups = ~ test, 
       data = subdata, 
       subset = binary_beta == TRUE & binary_regressor == FALSE & 
         only_intercept == FALSE & vary_beta == "all", 
       type = "l",
       lty = c(1,2,#1,2,
               1,2,1),
       lwd = 2,
       layout= c(4,1),
       key =  list(text = list(levels(subdata$test)),
                   lines = list(lty = c(1,2,#1,2,
                                        1,2,1), 
                                type = "l", pch = 1, cex = 0.8,
                                col = pal[c("CTree", "CTree", 
                                            #"CTree+max", "CTree+max", 
                                            "MOB", "MOB",
                                            "GUIDE+scores")], lwd = 2)), 
       #index.cond = list(c(1,2,3,4)),
       xlab = expression(delta),
       ylab = expression("Selection probability of"~Z[1]),
       strip = strip.custom(factor.levels = c(expression(xi==0), 
                                              expression(xi==0.2), 
                                              expression(xi==0.5), 
                                              expression(xi==0.8)),
                                   bg = "gray90"),
       par.settings = par.settings)

@
\caption{\label{fig:bin_stump}
Selection probability of the true split variable $Z_1$ for
testing strategies CTree, CTree+bin, MOB, MOB+bin, and GUIDE+scores in the ``stump''
scenario with both parameter varying. 
Probabilities are estimated over increasing effect size $\delta$
with 100 replications of 250 observations per step. 
Each panel represents a different value of the true split point $\xi$ (0, 0.2, 0.5 
and 0.8 from left to right corresponding to the 50\%, 60\%, 75\%, and 90\% quantile).
}
\end{figure}

\subsection{Categorization of splitting variables}
\label{app:categorization}
To investigate the effect of categorizing the possible splitting variables CTree and MOB are
applied, once in their original version without a categorization and once in an adapted version
with a categorization of the possible split variables (CTree+cat and MOB+cat). 
Moreover, they are also compared to GUIDE+scores which includes all available scores.

In Figure~\ref{fig:cat_stump} the impact of categorizing split variables on the performance 
is illustrated over increasing effect size for four different values of the true split 
point $\xi$. For both, CTree and MOB, it can be stated that overall they perform better in their
original form without any categorization. Only if the true split point $\xi$
is close to the quartiles used as breakpoints for the categorizations both versions
lead to a similar proportion of replications where the true splitting variable is detected
(e.g., for CTree and $\xi=0.5$)

Therefore, it can be summarized that a categorization of the values of the split variable
does not lead to any advantages unless the true split point corresponds to one of the 
quartiles used for the categorization. In most situations it even causes the performance to 
get worse.

\begin{figure}%%[t]
\setkeys{Gin}{width=\linewidth}
<<cat_stump, fig=TRUE, echo=FALSE, height=4, width=8>>=
subdata <- subset(sim_stump, test %in% c("ctree", "ctree_cat",
                                         #"ctree_max", "ctree_max_cat",
                                         "mfluc", "mfluc_cat",
                                         "guide_sum_12"))
subdata <- subset(subdata, delta %in% c(0, 0.1, 0.2, 0.3, 0.4, 0.5))
subdata$test <- factor(subdata$test, 
                       levels = c("ctree", "ctree_cat", 
                                  #"ctree_max","ctree_max_cat",
                                  "mfluc", "mfluc_cat",
                                  "guide_sum_12"),
                       labels = c("CTree","CTree_cat",
                                  #"CTree+max","CTree+max_cat",
                                  "MOB", "MOB_cat",
                                  "GUIDE+scores"))

par.settings <- list(superpose.symbol = list(col = pal[c("CTree", "CTree", 
                                                         #"CTree+max", "CTree+max", 
                                                         "MOB", "MOB",
                                                         "GUIDE+scores")], 
                                             fill = pal[c("CTree", "CTree", 
                                                          #"CTree+max", "CTree+max", 
                                                          "MOB", "MOB",
                                                          "GUIDE+scores")]), 
                     superpose.line = list(col = pal[c("CTree", "CTree", 
                                                       #"CTree+max", "CTree+max", 
                                                       "MOB", "MOB",
                                                       "GUIDE+scores")]))

xyplot(prop_Tsplit ~ delta | xi, groups = ~ test, 
       data = subdata, 
       subset = binary_beta == TRUE & binary_regressor == FALSE & 
         only_intercept == FALSE & vary_beta == "all", 
       #xlim = c(0,0.1,0.5),
       type = "l", 
       lty = c(1,2,#1,2,
               1,2,1),
       lwd = 2,
       layout= c(4,1),
       key =  list(text = list(levels(subdata$test)),
                   lines = list(lty = c(1,2,#1,2,
                                        1,2,1), 
                                type = "l", pch = 1, cex = 0.8,
                                col = pal[c("CTree", "CTree", 
                                            #"CTree+max", "CTree+max", 
                                            "MOB", "MOB",
                                            "GUIDE+scores")], lwd = 2)), 
       #index.cond = list(c(1,2,3,4)),
       xlab = expression(delta),
       ylab = expression("Selection probability of"~Z[1]),
       strip = strip.custom(factor.levels = c(expression(xi==0), 
                                              expression(xi==0.2), 
                                              expression(xi==0.5), 
                                              expression(xi==0.8)),
                                   bg = "gray90"),
       par.settings = par.settings)

@
\caption{\label{fig:cat_stump}
Selection probability of the true split variable $Z_1$ for
testing strategies CTree, CTree+cat, MOB, MOB+cat, and GUIDE+scores in the ``stump''
scenario with both parameter varying. 
Probabilities are estimated over increasing effect size $\delta$
with 100 replications of 250 observations per step. 
Each panel represents a different value of the true split point $\xi$ (0, 0.2, 0.5 
and 0.8 from left to right corresponding to the 50\%, 60\%, 75\%, and 90\% quantile).
}
\end{figure}

\end{appendix}

\end{document}

%\begin{figure}%[t]
%\setkeys{Gin}{width=\linewidth}
%<<tree_ari, fig=TRUE, echo=FALSE, height=4, width=8>>=
%subdata <- subset(sim_tree, test %in% c("ctree","mfluc",
%                                        "guide_sum_12", "guide_sum_1_cor"))
%subdata$test <- factor(subdata$test, 
%                       levels = c("ctree", "mfluc",
%                                  "guide_sum_12", "guide_sum_1_cor"),
%                       labels = c("CTree","MOB",
%                                  "GUIDE+scores", "GUIDE"))
%
%par.settings <- list(superpose.symbol = list(col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")], 
%                                             fill = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")]), 
%                     superpose.line = list(col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")]))
%
%
%xyplot(ari ~ delta | xi, groups = ~ test, 
%       data = subdata, 
%       type = "l",
%       lty = c(1,1,1,1),
%       lwd = 2,
%      key =  list(text = list(levels(subdata$test)),
%                   lines = list(lty = c(1,1,1,1), type = "l", pch = 1, cex = 0.8,
%                                col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")], lwd = 2)),
%       layout= c(4,1),
%       #index.cond = list(c(1,2,3,4)),
%       scales=list(x=list(at=seq(1,11,2))),
%       xlab = expression(delta),
%       ylab = "Adjusted Rand Index",
%       strip = strip.custom(factor.levels = c(expression(xi==0), 
%                                              expression(xi==0.2), 
%                                              expression(xi==0.5), 
%                                              expression(xi==0.8)),
%                                   bg = "gray90"),
%       par.settings = par.settings)
%
%@
%\caption{\label{fig:tree_ari}
%Adjusted Rand index (ARI) for the testing strategies CTree, MOB, GUIDE, and GUIDE+scores in the
%``tree'' scenario, all applying pre-pruning. The ARI is evaluated over
%increasing effect size $\delta$ with 100 replications of 250 observations per step.
%Each panel represents a different value of the true split point $\xi$ ($0$, $0.2$, $0.5$ and 
%$0.8$ from left to right).}
%\end{figure}

%<<pruning_data, echo=FALSE, results=hide>>=
%if(file.exists("pruning.rda")){
%  load("pruning.rda")
%} else {
%  simres <- simwrapper_p(nobs = 250, nrep = 100, seed = 7, 
%                         nrsteps = 1, nrlevels = 2,
%                         delta = seq(from = 0, to = 1, by = 0.1),
%                         xi = c(0, 0.2, 0.5, 0.8), 
%                         vary_beta = "all", binary_regressor = FALSE,
%                         binary_beta = TRUE, only_intercept = FALSE,
%                         compare_pruning = TRUE,
%                         test = c("ctree", "mfluc",
%                                  "guide_sum_12", 
%                                  "guide_sum_1_cor"),
%                         beta0 = NULL, beta1 = NULL,
%                         stump = FALSE, z1dist = "unif", sigma = 1, alpha = 0.05,
%                         nrsplitvar = 2)
%  
%  save(simres, file = "pruning.rda")  
%}
%
%pr <- simres$res
%## extract data for xyplot comparing ARI and ARI_p
%pr <- pr[,c(1,2,7,8,9,12,15)]
%## extended data set with post-pruning results being represented as separated tests
%prex <- pr[,c(1,2,3,4,6)]
%prex <- rbind(prex,prex)
%
%prex$test <- factor(prex$test,
%                    levels = c("ctree","mfluc",
%                               "guide_sum_12", "guide_sum_1_cor",
%                               "ctree_p","mfluc_p",
%                               "guide_sum_12_p", "guide_sum_1_cor_p"),
%                    labels = c("CTree","MOB",
%                               "GUIDE+scores", "GUIDE",
%                               "CTree_p","MOB_p",
%                               "GUIDE+scores_p", "GUIDE_p"))
%
%for(i in c((length(pr$test)+1) : (2*length(pr$test)))){
%  if(prex$test[i] == "CTree") prex$test[i] <- "CTree_p"
%  if(prex$test[i] == "MOB") prex$test[i] <- "MOB_p"
%  if(prex$test[i] == "GUIDE+scores") prex$test[i] <- "GUIDE+scores_p"
%  if(prex$test[i] == "GUIDE") prex$test[i] <- "GUIDE_p"
%}
%
%prex$nrsubgr[(length(pr$nrsubgr)+1) : (2*length(pr$nrsubgr))] <- pr$nrsubgr_p
%prex$ari[(length(pr$ari)+1) : (2*length(pr$ari))] <- pr$ari_p
%
%par.settings <- list(superpose.symbol = list(col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")], 
%                                             fill = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")]), 
%                     superpose.line = list(col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")]))
%
%@

%\begin{figure}%[t]
%\setkeys{Gin}{width=\linewidth}
%<<prune_ari, fig=TRUE, echo=FALSE, height=4, width=8>>=
%xyplot(ari ~ delta | xi, groups = ~ test, 
%       data = prex, 
%       type = "l", lty = c(2,2,2,2,1,1,1,1),
%       lwd = 2,
%       key =  list(text = list(levels(prex$test)[1:4]),
%                   lines = list(lty = c(1,1,1,1), type = "l", pch = 1, cex = 0.8,
%                                col = pal[c("CTree", "MOB", "GUIDE+scores", "GUIDE")], lwd = 2)),
%       layout= c(4,1),
%       scales=list(x=list(at=seq(1,11,2))),
%       xlab = expression(delta),
%       ylab = "Adjusted Rand Index",
%       strip = strip.custom(factor.levels = c(expression(xi==0), 
%                                              expression(xi==0.2), 
%                                              expression(xi==0.5), 
%                                              expression(xi==0.8)),
%                                   bg = "gray90"),
%       par.settings = par.settings
%)
%@
%\caption{\label{fig:pruning_ari}
%Adjusted Rand index (ARI) for the testing strategies CTree, MOB, GUIDE, and GUIDE+scores in the
%``tree'' scenario, once applying pre-pruning (dashed lines) and once post-pruning (solid lines). 
%The ARI is evaluated over
%increasing effect size $\delta$ with 100 replications of 250 observations per step.
%Each panel represents a different value of the true split point $\xi$ ($0$, $0.2$, $0.5$ and 
%$0.8$ from left to right).}
%\end{figure}