### 2019-06-28 pkg/disttree/inst/IWSM2019/slides.Rnw
first draft of slides for IWSM 2019
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2626 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 227c0bb756d4097710ef42ed50c0b6a74658c2bd

\documentclass[11pt,t,usepdftitle=false,aspectratio=43]{beamer}
\usetheme[nototalframenumber, noslidenumber]{uibk}

\title{Distributional Trees 
%and Forests \\
for Circular Data}
\author{Lisa Schlosser, Moritz N. Lang, Torsten Hothorn, \\
Georg~J.~Mayr, Reto Stauffer, Achim Zeileis}
\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
%\renewcommand{\mysize}{\fontsize{7.7pt}{9pt}\selectfont}
\URL{https://R-Forge.R-project.org/projects/partykit/ \\
https://partykit.org}
\headerimage{3}

\setbeamertemplate{caption}{\insertcaption} 
%% includes a replacement for \usepackage{Sweave}
\usepackage{Sweave}
\usepackage{amsmath,tikz}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap}
\graphicspath{{plots/}}

\newcommand{\argmax}{\operatorname{argmax}\displaylimits}

%% colors
\definecolor{HighlightOrange}{rgb}{0.9490196,0.5725490,0.0000000}
\definecolor{HighlightBlue}{rgb}{0.4784314,0.7490196,0.9803922}

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)

set.seed(7)

# setwd("~/svn/partykit/pkg/disttree/inst/IWSM2019")
library("disttree")
library("Formula")
library("latex2exp")

## HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)
names(pal) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")

pallight <- hcl(c(10, 128, 260, 290, 70), 100, 50, alpha = 0.25)
names(pallight) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")

transpgrey <- rgb(0.190,0.190,0.190, alpha = 0.2)

@

\begin{document}
\SweaveOpts{concordance=TRUE}

\section{Distributional Trees and Forests for Circular Data}

\subsection{Motivation}

\begin{frame}[fragile]
\frametitle{Distributional Tree}

\vspace{1cm}

{\bf Starting point:}
\begin{itemize}
\item Response $Y$
\item Covariate(s) $X$ $\rightarrow$ split variable(s) $X$
\item Distribution family $\mathcal{D}(\theta)$
\end{itemize}

\vspace{1cm}

{\bf Goal:} Probabilistic model

%Idea: maximum-likelihood method
%\begin{enumerate}
%\item Specify a distribution with log-likelihood function $\ell(\theta; y)$.
%\item Estimate $\hat{\theta}$ via maximum likelihood.
%\item Test for associations or instabilities of the scores $\frac{\partial \ell}{\partial \theta}(\hat{\theta};y_i)$ and each partitioning variable $x_i$.
%\item Assess whether the \emph{model scores} are associated with
%    (or change along) any of the available covariates -- e.g.,
%    using parameter instability tests (\emph{strucchange}) or
%    conditional inference (\emph{coin}).
%\item Split the sample along the partitioning variable with the strongest association or instability.
%    Choose breakpoint with highest improvement in log-likelihood.
%\item Repeat steps 2--4 recursively in the subgroups
%    until some stopping criterion is met.
%    %-- e.g., for significance or sample size.
%\item Choose the variable with the strongest association
%\item Choose the split point which leads to the highest improvement of the model
%\item Split and repeat 2-6 in each node until a stopping criterion is met
%\end{enumerate}

\end{frame}

\begin{frame}
\frametitle{Distributional Tree}
\vspace*{-0.12cm}
\begin{center} 
DGP:  $\; Y\ |\ X = x \; \sim  \; \mathcal{N}(\mu(x), \sigma^2(x))$
\vspace*{-0.21cm}
<<dgp_tree, echo=FALSE, results=hide>>=
data <- data.frame(x = numeric(0), x = numeric(0), x = numeric(0))
names(data) <- c("x","x","x")
fig <- party(
  partynode(1L,
            split = partysplit(2L, breaks = 0.4),
            kids = list(
              partynode(2L, info = c(
                "n = 200",
                "   True parameters:   ",
                expression(mu == '4'),
                expression(sigma == '1')
              )),
              partynode(3L,
                        split = partysplit(3L, breaks = 0.8),
                        kids = list(
                          partynode(4L, info = c(
                            "n = 200",
                            "   True parameters:   ",
                            expression(mu == '12'),
                            expression(sigma == '3')
                          )),
                          partynode(5L, info = c(
                            "n = 100",
                            "   True parameters:   ",
                            expression(mu == '4'),
                            expression(sigma == '3')
                          )))))),
  data
)

node_inner_ext <- function (obj, id = TRUE, pval = TRUE, abbreviate = FALSE, fill = "white", 
    gp = gpar()) 
{
    meta <- obj$data
    nam <- names(obj)
    extract_label <- function(node) {
        if (is.terminal(node)) 
            return(rep.int("", 2L))
        varlab <- character_split(split_node(node), meta)$name
        if (abbreviate > 0L) 
            varlab <- abbreviate(varlab, as.integer(abbreviate))
        if (pval) {
            nullna <- function(x) is.null(x) || is.na(x)
            pval <- suppressWarnings(try(!nullna(info_node(node)$p.value), 
                silent = TRUE))
            pval <- if (inherits(pval, "try-error")) 
                FALSE
            else pval
        }
        if (pval) {
            pvalue <- node$info$p.value
            plab <- ifelse(pvalue < 10^(-3L), paste("p <", 10^(-3L)), 
                paste("p =", round(pvalue, digits = 3L)))
        }
        else {
            plab <- ""
        }
        return(c(varlab, plab))
    }
    maxstr <- function(node) {
        lab <- extract_label(node)
        klab <- if (is.terminal(node)) 
            ""
        else unlist(lapply(kids_node(node), maxstr))
        lab <- c(lab, klab)
        lab <- unlist(lapply(lab, function(x) strsplit(x, "\n")))
        lab <- lab[which.max(nchar(lab))]
        if (length(lab) < 1L) 
            lab <- ""
        return(lab)
    }
    nstr <- maxstr(node_party(obj))
    if (nchar(nstr) < 6) 
        nstr <- "aAAAAa"
    rval <- function(node) {
        node_vp <- viewport(x = unit(0.5, "npc"), y = unit(0.5, 
            "npc"), width = unit(1, "strwidth", nstr) * 1.3, 
            height = unit(3, "lines"), name = paste("node_inner", 
                id_node(node), sep = ""), gp = gp)
        pushViewport(node_vp)
        xell <- c(seq(0, 0.2, by = 0.01), seq(0.2, 0.8, by = 0.05), 
            seq(0.8, 1, by = 0.01))
        yell <- sqrt(xell * (1 - xell))
        xell <- xell*1.11 - 0.055             # to adapt size of the ellipse to the size with p-value
        lab <- extract_label(node)
        fill <- rep(fill, length.out = 2L)
        grid.polygon(x = unit(c(xell, rev(xell)), "npc"), y = unit(c(yell, 
            -yell) + 0.5, "npc"), gp = gpar(fill = fill[1]))
        grid.text(lab[1L], y = unit(1.5 + 0.5,                  # to adapt position of x to its position with p-value
            "lines"))
        #grid.text(lab[1L], y = unit(1.5 + 0.5 * (lab[2L] != ""), 
        #    "lines"))
        grid.text(lab[2L], y = unit(1, "lines"))
        if (id) {
            nodeIDvp <- viewport(x = unit(0.5, "npc"), y = unit(1, 
                "npc"), width = max(unit(1, "lines"), unit(1.3, 
                "strwidth", nam[id_node(node)])), height = max(unit(1, 
                "lines"), unit(1.3, "strheight", nam[id_node(node)])))
            pushViewport(nodeIDvp)
            grid.rect(gp = gpar(fill = fill[2]))
            grid.text(nam[id_node(node)])
            popViewport()
        }
        upViewport()
    }
    return(rval)
}

class(node_inner_ext) <- "grapcon_generator"

@

\setkeys{Gin}{width=0.68\linewidth, height=0.68\linewidth}
<<plottree_dgp, fig=TRUE, echo=FALSE>>=
paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(fig, inner_panel = node_inner_ext,
     tp_args = list(FUN = identity, width = 18, fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1.7)
@
\end{center}
\end{frame}

<<tree, echo=FALSE, results=hide>>=
set.seed(7)
nobs <- 500
x <- runif(nobs, 0, 1)
mu <- sigma <- ytrue <- numeric(length = nobs)
for(i in 1:nobs) sigma[i] <- if(x[i]<=0.4) 1 else 3
for(i in 1:nobs) mu[i] <- if(x[i]<= 0.4|| x[i]>0.8) 4 else 12
y <- rnorm(nobs, mean = mu, sd = sigma)
#y <- rcnorm(nobs, mean = mu, sd = sigma, left = 0)
ytrue <- mu
data <- data.frame(cbind(y,x, ytrue))
tree <- disttree(y ~ x, data = data, family = NO(), type.tree = "mob")
#tree <- disttree(y ~ x, data = data, family = dist_list_cens_normal)
@

\begin{frame}[fragile]
\frametitle{Distributional Tree}
\begin{center}
\vspace*{-0.12cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.68\linewidth, height=0.68\linewidth}
<<plottree_estpar, fig=TRUE, echo=FALSE>>=
# function for output in terminal panels
FUN <- function (x) 
{
  cf <- x$coefficients
  cf <- matrix(cf, ncol = 1, dimnames = list(names(cf), ""))
  c(sprintf("n = %s", x$nobs), "Estimated parameters:", parse(text = paste0("mu == '", format(round(cf[1], 2), nsmall = 2), "'")), 
                                                        parse(text = paste0("sigma == '", format(round(cf[2], 2), nsmall = 2), "'")))
}

paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))

## plot version using FUN and tree of class 'disttree'
plot(tree, drop = TRUE, tnex = 1.7, FUN = FUN,
     tp_args = list(fill = paltrees[c(1, 3)], width = 18), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Distributional Tree}
\begin{center}
\vspace*{-0.12cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.68\linewidth, height=0.68\linewidth}
<<plottree_box, fig=TRUE, echo=FALSE>>=
plot(as.constparty(tree), tnex = 1.7, drop = TRUE,
     tp_args = list(fill = paltrees[c(1, 3)], ylines = 1.5), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Distributional Tree}
\begin{center}
\vspace*{-0.12cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.68\linewidth, height=0.68\linewidth}
<<plottree_dens, fig=TRUE, echo=FALSE>>=
node_density <- function (tree, xscale = NULL, yscale = NULL, horizontal = FALSE,
                          main = "", xlab = "", ylab = "Density", id = TRUE, rug = TRUE,
                          fill = paltrees[c(1, 3)], col = "black", lwd = 0.5, ...) 
{
  yobs <- tree$data[,as.character(tree$info$formula[[2]])]
  ylines <- 1.5
  if (is.null(xscale)) xscale <- c(-5.1,22.5)
  if (is.null(yscale)) yscale <- c(-0.05,0.45)
  xr <- xscale
  yr <- yscale

  if (horizontal) {
    yyy <- xscale
    xscale <- yscale
    yscale <- yyy
  }

  rval <- function(node) {
    yrange <- seq(from = -20, to = 90)/4
    ydens <- node$info$object$ddist(yrange)

    top_vp <- viewport(layout = grid.layout(nrow = 2, ncol = 3, 
                                            widths = unit(c(ylines, 1, 1), c("lines", "null", "lines")), 
                                            heights = unit(c(1, 1), c("lines", "null"))), 
                       width = unit(1, "npc"), 
                       height = unit(1, "npc") - unit(2, "lines"), 
                       name = paste("node_density",node$id, sep = ""))
    pushViewport(top_vp)
    grid.rect(gp = gpar(fill = "white", col = 0))
    top <- viewport(layout.pos.col = 2, layout.pos.row = 1)
    pushViewport(top)
    mainlab <- paste(ifelse(id, paste("Node", node$id, "(n = "), "n = "), node$info$nobs, ifelse(id, ")", ""), sep = "")

    grid.text(mainlab)
    popViewport()
    plot <- viewport(layout.pos.col = 2, layout.pos.row = 2, 
                     xscale = xscale, yscale = yscale, 
                     name = paste("node_density",  node$id, "plot", sep = ""))
    pushViewport(plot)
    yd <- ydens
    xd <- yrange
    if (horizontal) {
      yyy <- xd
      xd <- yd
      yd <- yyy
      yyy <- xr
      xr <- yr
      yr <- yyy
      rxd <- rep(0, length(xd))
      ryd <- rev(yd)
    } else {
      rxd <- rev(xd)
      ryd <- rep(0, length(yd))
    }

    if (rug) {
      nodeobs <- node$info$object$y
      if (horizontal) {
        grid.rect(x = xscale[1], y = nodeobs , height = 0, width = xscale[1], 
                  default.units = "native", just = c("right", "bottom"),
		  gp = gpar(lwd = 2, col = gray(0, alpha = 0.18)))
      } else {
        grid.rect(x = nodeobs, y = yscale[1], 
                  width = 0, height = abs(yscale[1]), default.units = "native", 
                  just = c("center", "bottom"),
		  gp = gpar(lwd = 2, col = gray(0, alpha = 0.18)))
        #grid.lines(x = xr, y = yr, gp = gpar(col = "lightgray"), 
        #           default.units = "native")
        #grid.lines(x = xr, y = yr, gp = gpar(col = "lightgray"), 
        #           default.units = "native")
      }
    }

    grid.polygon(x = c(xd, rxd), y = c(yd, ryd), default.units = "native",
              gp = gpar(col = "black", fill = fill, lwd = lwd))
    #grid.lines(x = xd, y = yd, default.units = "native", 
    #           gp = gpar(col = col, lwd = lwd))
    grid.xaxis()
    grid.yaxis()
    grid.rect(gp = gpar(fill = "transparent"))
    upViewport(2)
  }
  return(rval)
}

class(node_density) <- "grapcon_generator"

plot(tree, tnex = 1.7, drop = TRUE,
     terminal_panel = node_density,
     tp_args = list(fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}

<<treedata, echo=FALSE, results=hide>>=
set.seed(54)
nobs <- 500
x <- runif(nobs, 0, 1)
mu <- sigma <- ytrue <- numeric(length = nobs)
for(i in 1:nobs) sigma[i] <- if(x[i]<=0.4) 1 else 3
for(i in 1:nobs) mu[i] <- if(x[i]<= 0.4|| x[i]>0.8) 4 else 12
y <- rnorm(nobs, mean = mu, sd = sigma)
ytrue <- mu
data <- data.frame(cbind(y,x, ytrue))

alldata <- cbind(data, mu, sigma)
odata <- alldata[order(alldata["x"]),]
@

\begin{frame}[fragile]
\frametitle{Distributional Tree}
\begin{center}
\vspace*{-1.03cm}
\setkeys{Gin}{width=0.8\linewidth}
<<plottree_xyplot, fig=TRUE, echo=FALSE>>=
par(mar=c(5.1,4.1,4.1,3.1))
plot(y=odata$y, x=odata$x, ylab = "y", xlab = "x", col = "grey")
#plot(y=odata$y, x=odata$x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray")
lines(x = odata$x, y = odata$mu, col = pal["forest"], lwd = 2.5, main = "")
polygon(c(odata$x, rev(odata$x)), c(odata$mu + odata$sigma, rev(odata$mu - odata$sigma)),
  col = pallight["forest"], border = "transparent")
legend(x = -0.0715, y = 21.62, expression(mu  %+-%  sigma), bty = "n")
#lines(x = odata$x, y = odata$mu + odata$sigma, col = hcl(10, 80, 70), lwd = 1.5, lty = 1)
#lines(x = odata$x, y = odata$mu - odata$sigma, col = hcl(10, 80, 70), lwd = 1.5, lty = 1)
@
\end{center}
\end{frame}

\begin{frame}
\frametitle{Building Distributional Trees and Forests}
\begin{minipage}{0.7\textwidth}
{\bf Tree:}
\begin{enumerate}
\item<3-> Fit global distributional model $\mathcal{D}(Y; \theta)$: \\ % to the whole data set:\\
Estimate $\hat{\theta}$ via maximum likelihood\\
$\hat{\theta} = \argmax_{\theta \in \Theta} \sum_{i=1}^n \ell(\theta; y_i)$
\item<7-> Test for associations/instabilities of the scores $\frac{\partial \ell}{\partial \theta}(\hat{\theta};y_i)$ and each covariate $X_i$.
\end{enumerate}
\end{minipage}
\begin{minipage}{0.23\textwidth}
%\vspace{-0.1cm}
\begin{tikzpicture}
\visible<2-3>{
\node[ellipse, fill=HighlightBlue!70, align=center, scale = 0.7, minimum width=60pt, minimum height = 30pt] (n0) at (0.8, 1.7) {$Y$};
}
\visible<4>{
\node[ellipse, fill=HighlightBlue!70, align=center, scale = 0.7, minimum width=60pt, minimum height = 30pt] (n0) at (0.8, 1.7) {$\mathcal{D}(Y;\hat{\theta}$)};
}
\visible<5->{
\node[inner sep=0pt] (density1) at (0.8, 1.7)
    {\includegraphics[width=0.6\textwidth]{density1.jpeg}};
}
\visible<6-8>{
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n1) at (0, 0) {?};
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n2) at (1.6, 0) {?};
\draw[-, gray, line width=0.5pt] (0.7, 1.18) -- (n1);
\draw[-, gray, line width=0.5pt] (0.9, 1.18) -- (n2);
}
\visible<9-10>{
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n1) at (0, 0) {$Y_1$};
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n2) at (1.6, 0) {$Y_2$};
\draw[-, gray, line width=0.5pt] (0.7, 1.18) -- (n1) node [midway, left] {\scriptsize $X \leq p$};
\draw[-, gray, line width=0.5pt] (0.9, 1.18) -- (n2) node [midway, right] {\scriptsize $X > p$};
}
\visible<11>{
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n1) at (0, 0) {$\mathcal{D}(Y_1;\hat{\theta_1}$)};
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n2) at (1.6, 0) {$\mathcal{D}(Y_2;\hat{\theta_2}$)};
}
\visible<11->{
\draw[-, gray, line width=0.5pt] (0.7, 1.18) -- (n1) node [midway, left] {\scriptsize $X \leq p$};
\draw[-, gray, line width=0.5pt] (0.9, 1.18) -- (n2) node [midway, right] {\scriptsize $X > p$};
}
\visible<12->{
\node[inner sep=0pt] (density2) at (-0.1,-0.1)
    {\includegraphics[width=0.5\textwidth]{density2.jpeg}};
}
\visible<12->{
\node[inner sep=0pt] (density3) at (1.7,-0.1)
    {\includegraphics[width=0.5\textwidth]{density3.jpeg}};
}
\end{tikzpicture}
\end{minipage}
\vspace{0.1cm}
%\begin{adjustwidth}{-0.0em}{-1em}
\begin{enumerate}
\setcounter{enumi}{2}
\item<8-> Split along the covariate $X$ with strongest association or instability and at breakpoint $p$ with highest improvement \\
in log-likelihood.
\item<10-> Repeat steps 1--3 recursively until some stopping criterion \\
is met, yielding $B$ subgroups $\mathcal{B}_b$ with $b = 1, \dots, B$.
\end{enumerate}
%\end{adjustwidth}
\vspace{0.4cm}
\visible<13->{
{\bf Forest:} Ensemble of $T$ trees.
\begin{itemize}
\item Bootstrap or subsamples.
\item Random input variable sampling.
\end{itemize}
}
\end{frame}

\begin{frame}
\frametitle{Adaptive Local Likelihood Estimation}

\textbf{Parameter estimator} for \only<1-3>{\textbf{a global}}\only<4->{\textbf{an adaptive local}}\\ model with learning data \only<1>{$\{y_i\}_{i=1,\ldots,n}$}\only<2->{$\{(y_i,\bold{x}_i)\}_{i=1,\ldots,n}$} :

\[
\hat{\theta}\visible<2->{(\bold{x})} =  \argmax_{\theta \in \Theta} \sum_{i=1}^n \visible<2->{w_i(\bold{x}) \cdot} \ell(\theta; y_i)
\]

\bigskip

\visible<3->{
\textbf{Weights:}
\begin{eqnarray*}
w^{\text{base}}_i(\bold{x})   & = & 1 \\[0.2cm]
\visible<4->{
w^{\text{tree}}_i(\bold{x})   & = & \sum_{b=1}^B I((\bold{x}_i \in \mathcal{B}_b) \land (\bold{x} \in \mathcal{B}_b)) \\[0.1cm]
\visible<5->{
w^{\text{forest}}_i(\bold{x}) & = & \frac{1}{T} \sum_{t=1}^T \sum_{b=1}^{B^t} I((\bold{x}_i \in \mathcal{B}^t_b) \land (\bold{x} \in \mathcal{B}^t_b))
\end{eqnarray*}
}}}
\end{frame}

\begin{frame}[fragile]
\frametitle{Circular Data}

\bigskip

\textbf{Examples/Applications:} 
\begin{itemize}
\item Wind directions: $[0^{\circ}, 360^{\circ})$ with $0^{\circ} = 360^{\circ}$
\item Hourly crime data: $[0\text{h}, 24\text{h})$ with $0\text{h} = 24\text{h}$
\item Directional behavior of animals: $[0^{\circ}, 360^{\circ})$ with $0^{\circ} = 360^{\circ}$
\item \ldots
\end{itemize}

\bigskip

\textbf{Task:} Model periodic response variable measured on the unit circle.

\medskip

\quad $\rightarrow$ Circular distribution

%\begin{center}
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{}
%\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Von Mises distribution}
Also known as ``the circular normal distribution'' with
\begin{itemize}
\item location parameter $\mu \in [0, 2\pi)$ 
\item concentration parameter $\kappa > 0$ 
\end{itemize}

\medskip

The density for $y \in [0, 2\pi)$ is given by:
$$
f_\mathrm{vM}(y; \mu, \kappa) = \frac{1}{2 \pi I_0(\kappa)}~e^{ \kappa \cos(y - \mu)}\label{schlosser:equ_vm}
$$
where $I_0(\kappa)$ is the modified Bessel function of the first kind and order $0$.

\bigskip

\textit{FIXME: include a plot of the density function on the unity circle}
%\begin{center}
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{}
%\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Wind Directions}

\textbf{Goal:} Predict wind directions at Innsbruck Airport (Austria).

\medskip

\textbf{Input:} Observation data of various meteorological quantities of the previous hour(s) from Innsbruck and 5 nearby weather stations.

\medskip

\textbf{Base variables:}
\begin{itemize}
\item Wind direction
\item Temperature
\item Air pressure
\item \dots 
\item \textit{FIXME: list of all included covariates}
\end{itemize}
%\begin{center}
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{}
%\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Wind Directions}

\textbf{Distributional tree model:}

\begin{center}
%\vspace*{1cm}
\setkeys{Gin}{width=\linewidth}
\includegraphics{schlosser-circtree_plot.pdf}
\end{center}

\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Wind Directions}

\textbf{Cross validation:}

\bigskip

\textit{FIXME: explain applied setting}

%\begin{center}
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{}
%\end{center}
\end{frame}

\begin{frame}
\frametitle{Software}
%\vspace{0.4cm}
\textbf{Packages:} \emph{disttree} and \emph{circmax} available on R-Forge at\\

\medskip

\small{\url{https://R-Forge.R-project.org/projects/partykit/}\\ 
and\\
\url{https://r-forge.r-project.org/projects/uibk-rprog-2018/}}\\

\bigskip
\bigskip

\textbf{Main functions:}

\medskip

\begin{tabular}{ll}

\code{disttree}   & Distributional trees (\code{ctree}/\code{mob} + \code{distfit}).\\
                  & Covariates as partitioning variables. \\
\code{distforest} & Distributional forests (ensemble of \code{disttree}s).\\
                  & Covariates as partitioning variables.
\end{tabular}

\bigskip
\bigskip

\textbf{Distribution family:} \code{dist\_vonmises}

\end{frame}

\subsection{References}

\begin{frame}
\frametitle{References (1)}

%\vspace{-0.4cm}

\small
Bai J, Perron P (2003).
 \dquote{Computation and Analysis of Multiple Structural Change Models.}
 \textit{Journal of Applied Econometrics}, 
 \textbf{18}, 1--22.
 \doi{}

\medskip

Breiman L (2001).
 \dquote{Random Forests.}
 \textit{Machine Learning}, 
 \textbf{45}(1), 5--32.
 \doi{10.1023/A:1010933404324}

\medskip

Fisher N I, Lee A J (1992).
 \dquote{Regression Models for an Angular Response.}
 \textit{Biometrics}, 
 \textbf{48}(3), 665--677. 
 \doi{}

\medskip

Hothorn T, Hornik K, Zeileis A (2006).
 \dquote{Unbiased Recursive Partitioning: A Conditional Inference Framework.}
 \textit{Journal of Computational and Graphical Statistics},
 \textbf{15}(3), 651--674.
 \doi{10.1198/106186006X133933}

\medskip

Hothorn T, Zeileis A (2008).
 \dquote{Generalized Maximally Selected Statistics.}
 \textit{Biometrics}, 
 \textbf{64}(4), 1263--1269. 
 \doi{}

\medskip

Jammalamadaka S R, Sengupta A (2001).
 \textit{Topics in Circular Statistics}.
 World Scientific.

\end{frame}

\begin{frame}
\frametitle{References (2)}

%\vspace{-0.4cm}

\small
Mulder K, Klugkist I (2017).
 \dquote{Bayesian Estimation and Hypothesis Tests for a Circular Generalized Linear Model.}
 \textit{Journal of Mathematical Psychology}, 
 \textbf{80}, 4--14. 
 \doi{}

\medskip

Schlosser L, Hothorn T, Stauffer R, Zeileis A (2019).
  \dquote{Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain.}
  \emph{arXiv 1804.02921}, arXiv.org E-Print Archive.
  \url{http://arxiv.org/abs/1804.02921}

\medskip

Zeileis A, Hothorn T, Hornik K (2008).
 \dquote{Model-Based Recursive Partitioning.}
  \textit{Journal of Computational and Graphical Statistics},
  \textbf{17}(2), 492--514.
  \doi{10.1198/106186008X319331}

%\medskip

%Hothorn T, Zeileis A (2015).
% \dquote{\textbf{partykit}: A Modular Toolkit for Recursive Partytioning in \textsf{R}.}
% \textit{Journal of Machine Learning Research},
% \textbf{16}, 3905--3909.
% \url{http://www.jmlr.org/papers/v16/hothorn15a.html}

\end{frame}

\end{document}