### 2019-11-05 pkg/extree_partykit/index.Rmd
create extree improvement branch
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2813 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 6ea65118b722de54b5770b6466add9909cf65a11

## A Toolkit for Recursive Partytioning

```{r ttnc, echo = FALSE, message = FALSE, results = "hide"}
library("partykit")
data("Titanic", package = "datasets")
ttnc <- as.data.frame(Titanic)
ttnc <- ttnc[rep(1:nrow(ttnc), ttnc$Freq), 1:4]
names(ttnc)[2] <- "Gender"
ttnc <- transform(ttnc, Treatment = factor(
  Gender == "Female" | Age == "Child", levels = c(FALSE, TRUE),
  labels = c("Normal\nMale&Adult", "Preferential\nFemale|Child")
))

```{r ttnc-mob, echo = FALSE, message = FALSE, out.width = "100%", fig.height = 6, fig.width = 10, fig.align = "center", dev = "png", dpi = 48}
ttnc_mob <- glmtree(Survived ~ Treatment | Class + Gender + Age,
  data = ttnc, family = binomial, alpha = 0.01)
plot(ttnc_mob, tp_args = list(ylines = 1.3, margins = c(1.5, 1, 1.5, 2.2)))
```

**Tree algorithms:** Conditional inference trees (`ctree()`) and model-based
recursive partitioning (`mob()`) along with interfaces `lmtree()` and `glmtree()`
for (generalized) linear model trees.

**Infrastructure:** Unified classes and methods for representing, summarizing, and
visualizing tree-structured regression and classification models. Common print/plot/predict methods,
enabling reading/coercing tree models from different sources (`rpart`, `RWeka`, `PMML`).

* `partysplit`: Split with integer ID for the splitting variable,
  breakpoint(s), indexes for the kids.
* `partynode`: Node specification with integer ID, a `partysplit`,
  and a list of kids (if any) that are `partynode` objects again.
* `party`: Tree with a recursive `partynode` and a `data.frame`
  (optionally empty), potentially plus information about fitted values
  and `terms` allowing to preprocess new data for predictions.