### 2019-12-14 pkg/circtree/inst/wind_paper/draft/draft.Rnw
some touch-ups in Sections 4 and 5
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2933 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: e596d25751143dde1f4ed92025bbd31262fd447e

\documentclass[nojss]{jss}

%% packages
\usepackage{amstext,amsfonts,amsmath,bm,thumbpdf,lmodern,hyperref}
\usepackage[all]{hypcap}

%% tikz
\usepackage{array,makecell,tikz,color,soul}
\usetikzlibrary{arrows.meta,positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap}

%% need no \usepackage{Sweave}
\SweaveOpts{engine = R, concordance = FALSE, eps = FALSE, keep.source = TRUE, eval = TRUE}

%% commands
\newcommand{\fixme}[1]{\emph{\marginpar{FIXME} (#1)}}
\renewcommand{\Prob}{\mathbb{P} }
\renewcommand{\E}{\mathbb{E}}
\newcommand{\V}{\mathbb{V}}
\newcommand{\Var}{\mathbb{V}}
\newcommand{\R}{\mathbb{R} }
\newcommand{\N}{\mathbb{N} }
\newcommand{\C}{\mathbb{C} }
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}
\newcommand{\LS}{\mathcal{L}_n}
\newcommand{\TS}{\mathcal{T}_n}
\newcommand{\LSc}{\mathcal{L}_{\text{comb},n}}
\newcommand{\LSbc}{\mathcal{L}^*_{\text{comb},n}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\yn}{y_{\text{new}}}
\newcommand{\z}{\mathbf{z}}
\newcommand{\X}{\mathbf{X}}
\newcommand{\Y}{\mathbf{Y}}
\newcommand{\sX}{\mathcal{X}}
\newcommand{\sY}{\mathcal{Y}}
\newcommand{\T}{\mathbf{T}}
\newcommand{\x}{\mathbf{x}}
\renewcommand{\a}{\mathbf{a}}
\newcommand{\xn}{\mathbf{x}_{\text{new}}}
\newcommand{\y}{\mathbf{y}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\ws}{\mathbf{w}_\cdot}
\renewcommand{\t}{\mathbf{t}}
\newcommand{\M}{\mathbf{M}}
\renewcommand{\vec}{\text{vec}}
\newcommand{\B}{\mathbf{B}}
\newcommand{\K}{\mathbf{K}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\I}{\mathbf{I}}
\newcommand{\bS}{\mathbf{S}}
\newcommand{\cellx}{\pi_n[\x]}
\newcommand{\partn}{\pi_n(\mathcal{L}_n)}
\newcommand{\err}{\text{Err}}
\newcommand{\ea}{\widehat{\text{Err}}^{(a)}}
\newcommand{\ecv}{\widehat{\text{Err}}^{(cv1)}}
\newcommand{\ecvten}{\widehat{\text{Err}}^{(cv10)}}
\newcommand{\eone}{\widehat{\text{Err}}^{(1)}}
\newcommand{\eplus}{\widehat{\text{Err}}^{(.632+)}}
\newcommand{\eoob}{\widehat{\text{Err}}^{(oob)}}
\newcommand{\bft}{\mathbf{t}}

\definecolor{changes}{rgb}{1,1,0.7}
\newcommand{\hlc}[1]{{\sethlcolor{changes}\hl{#1}}}

\numberwithin{equation}{section}

<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)
library("disttree")
library("circtree")

## HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)
pallight <- hcl(c(10, 128, 260, 290, 70), 100, 50, alpha = 0.25)
transpgray <- rgb(0.190,0.190,0.190, alpha = 0.2)

# set function for parallelization
applyfun <- function(X, FUN, ...) parallel::mclapply(X, FUN, ..., mc.cores = pmax(1, parallel::detectCores() - 1))

@

\title{Circular Regression Trees and Forests with an Application to Probabilistic Wind Direction Forecasting}
\Shorttitle{Circular Regression Trees and Forests}

\author{Moritz N. Lang\\Universit\"at Innsbruck
   \And Lisa Schlosser\\Universit\"at Innsbruck
   \And Torsten Hothorn\\Universit\"at Z\"urich
   \AND Georg~J.~Mayr\\Universit\"at Innsbruck
   \And Reto Stauffer\\Universit\"at Innsbruck
   \And Achim Zeileis\\Universit\"at Innsbruck}
\Plainauthor{Lang, Schlosser, Hothorn, Mayr, Stauffer, Zeileis}

\Abstract{
While circular data occur in a wide range of scientific fields,
the methodology for distributional modeling and probabilistic forecasting of circular
response variables is rather limited. Most of the existing methods are built on the framework of
generalized linear and additive models, which are often challenging to optimize
and to interpret. Therefore, we suggest circular regression trees and random forests
as an inuitive alternative approach that is relatively easy to fit. Building on
previous ideas for trees modeling circular means, we suggest a distributional
approach for both trees and forests yielding probabilistic forecasts based on the
von Mises distribution. The resulting tree-based models simplify
the estimation process by using the available covariates for partitioning the
data into sufficiently homogeneous subgroups so that a simple von Mises distribution
without further covariates can be fitted to the circular response in each subgroup.
These circular regression trees are straightforward to interpret, can capture
nonlinear effects and interactions, and automatically select the relevant
covariates that are associated with either location and/or scale changes in the
von Mises distribution. Combining an ensemble of circular regression trees
to a circular regression forest yields a local adaptive likelihood estimator
for the von Mises distribution that can regularize and smooth the covariate effects.
The new methods are evaluated in a case study on probabilistic wind direction
forecasting at two Austrian airports, considering other common approaches as
a benchmark.
}

\Keywords{circular data, regression trees, random forests, distributional regression, von Mises distribution}

\Address{
  Moritz N. Lang, Lisa Schlosser, Reto Stauffer, Achim Zeileis \\
  Universit\"at Innsbruck \\
  Department of Statistics \\
  Faculty of Economics and Statistics \\
  Universit\"atsstr.~15 \\
  6020 Innsbruck, Austria \\
  E-mail: \email{Moritz.Lang@uibk.ac.at}, \email{Lisa.Schlosser@uibk.ac.at},\\
  \email{Reto.Stauffer@uibk.ac.at},\email{Achim.Zeileis@R-project.org} \\
  URL: \url{https://www.uibk.ac.at/statistics/personal/moritz-lang/},\\
  \phantom{URL: }\url{https://www.uibk.ac.at/statistics/personal/schlosser-lisa/},\\
  \phantom{URL: }\url{https://retostauffer.org/},\\
  \phantom{URL: }\url{https://eeecon.uibk.ac.at/~zeileis/}\\

  Georg~J.~Mayr\\
  Universit\"at Innsbruck \\
  Department of Atmospheric and Cryospheric Science \\
  Faculty of Geo- and Atmospheric Sciences \\
  Innrain~52f \\
  6020 Innsbruck, Austria \\
  E-mail: \email{Georg.Mayr@uibk.ac.at} \\
  URL: \url{http://acinn.uibk.ac.at/persons/georg_mayr}\\

  Torsten Hothorn\\
  Universit\"at Z\"urich \\
  Institut f\"ur Epidemiologie, Biostatistik und Pr\"avention \\
  Hirschengraben 84\\
  CH-8001 Z\"urich, Switzerland \\
  E-mail: \email{Torsten.Hothorn@R-project.org}\\
  URL: \url{http://user.math.uzh.ch/hothorn/}\\

}

\begin{document}

% -------------------------------------------------------------------
% Introduction
% -------------------------------------------------------------------
\section{Introduction}
\label{sec:introduction}

Circular data can be found in a variety of applications and subject areas,
e.g., hourly crime rate in socio-economics, animal movement direction or
gene-structure in biology, and wind direction as one of the most important
weather variables in meteorology. Fitting a statistical model to this type of
data requires the incorporation of its specific feature of periodicity. For example,
angular data are restricted to an interval such
as $[0,2\pi)$ with $0$ being equivalent~to~$2\pi$.

\subsection[Circular regression: Conditional mean vs. distributional models]{Circular regression: Conditional mean vs.\ distributional models}

Many approaches to model circular data assume that the circular variable of
interest follows a circular distribution, in particular the von Mises
distribution which is also known as the ``circular normal distribution''. One
of the first regression models with a circular response variable and linear
covariates was presented by \cite{Gould:1969} where the circular mean
%% , i.e., the location parameter of the von Mises distribution, 
is predicted by a linear combination of covariates. \cite{Johnson+Wehrly:1978} refined this idea by
plugging in a link function transforming the linear predictor to a restricted
interval of length $2\pi$. This generalized linear model (GLM) type approach
was further extended by \cite{Fisher+Lee:1992} and subsequently by
\cite{Fisher:1993} introducing independent GLMs for either the location or
scale parameter with appropriate link functions while keeping the other
parameter constant. Additionally, they developed a combined heteroscedastic
or distributional version with alternating reestimation of the location and scale parameters
conditional on the respective sets of covariates until convergence. While all of
these models are built on well-elaborated theory, their application in practice
remains very challenging, mainly due to the complexity
encountered in optimizing the corresponding log-likelihood function which is
not globally concave. Therefore, highly-informative starting values are
crucial for such circular GLMs to converge
\citep{Pewsey+Neuhaeuser+Ruxton:2013, Gill+Hangartner:2010}. In order to avoid
this strong dependence on appropriate initial values,
\cite{Mulder+Klugkist:2017} present a Bayesian alternative of a
homoscedastic GLM for circular data. However, apart from potential
difficulties in the optimization procedure of circular GLMs, the interpretation
of the underlying additive effects is often challenging as well because the link function
is highly nonlinear and the representation of smooth transitions on the unit circle
is not straightforward. For example, the same rotation can be
obtained in either positive or negative direction on the circle leading to an
ambiguous interpretation.

As a very intuitive and data-driven alternative, we propose a flexible
tree-based regression approach for modeling circular data by applying the von
Mises distribution within the methodology of distributional trees and forests
\citep{Schlosser+Hothorn+Stauffer:2019}. The resulting circular regression
trees and forests avoid the discussed difficulties of circular GLMs by
using the available covariates for partitioning the
data into sufficiently homogeneous subgroups so that a simple von Mises distribution
without further covariates can be fitted to the circular response in each subgroup.
This obviates the need for a link function or for iterating between models for the
separate distribution parameters. By leveraging the distributional modeling approach, the
trees can automatically detect and capture differences in both
distribution parameters, providing a fully
specified circular response distribution in each terminal node offering a wide
range of statistical inference. In addition, the employed tree structure
allows to capture non-additive effects while forests enable the modeling of
smooth changes. Furthermore, covariates and their possible interactions do not
need to be specified in advance as they are selected automatically in the
recursive partitioning algorithm.

This novel approach to circular regression trees and forests complements the
literature on tree-based circular modeling. \cite{Lund:2002} already
introduced a circular regression tree algorithms where binary splits are
made based on an angular distance measure capturing node
homogeneity. However, this only models changes in the conditional mean
but not the conditional variance or full probabilistic distribution.
In contrast, here we introduce a distributional tree approach that considers
splits in all distribution parameters and yields a full probabilistic predictive model.
In addition, as a natural extension, ensembles or forests of circular
regression trees are presented in this study, showing that these can further
improve predictive power.

\begin{figure}[p!]
\centering
\includegraphics[width=0.9\textheight,,angle=90,origin=c]{_plot_circforest_finalexampletree_ibk_lag1_v14.pdf}
\caption{\label{fig:tree_ibk} Fitted tree based on the von Mises distribution for 1-hourly wind
direction forecasts at the airport of Innsbruck. In each terminal node the
empirical histogram (gray) and fitted density (red line) are depicted along
with the estimated location parameter (red hand). The covariates selected for
splitting are wind direction (degree), wind speed ($\text{ms}^{-1}$), and
pressure gradients ($\text{dpressure; hPa}$) west, east and south of the
airport, all lagged by one hour. Note that in the meteorological context wind
direction is defined on the scale $[0,360]$ degree and increases clockwise from
North ($0$ degree).}
\end{figure}

\subsection{Motivating example}

To provide a first impression of the presented methodology, a circular
regression tree is employed for probabilistic wind direction forecasting. Wind
direction is a classical circular quantity and accurate forecasts are of great
importance for decision-making processes, e.g., in air traffic management as
considered in this study. Figure~\ref{fig:tree_ibk} shows an estimated tree
for 1-hourly forecasts at Innsbruck Airport, located at the bottom of
a narrow valley within the European Alps. Topography channels the wind along
the west-east valley axis or along a tributary valley intersecting from the
south. Hence, pressure gradients to which valley wind regimes react are
considered as covariates along with other meteorological measurements (lagged
by one hour) and their derivatives, such as wind direction and wind speed at
the airport as well as spatial and temporal differences.

Figure~\ref{fig:tree_ibk} illustrates the resulting tree along with
the empirical (gray) and fitted von Mises (red) wind direction distribution
in each terminal node. Based on the fitted location parameters $\hat \mu$, the
subgroups can be distinguished into the following wind regimes: (1)~Up-valley
winds blowing from the valley mouth towards the upper valley (from east to
west, nodes~4 and 5); (2) Downslope winds blowing across the Alpine crest along
the intersecting valley towards Innsbruck (from south-east to north-west,
node~8); (3) Down-valley winds blowing in the direction of the valley mouth
(from west to east, nodes 10, 12 and 13). Node~7 captures observations with
rather low wind speeds that cannot be clearly distinguished into specific wind
regimes and are consequently associated with a very low estimated concentration
parameter~$\hat{\kappa}$, i.e., a high estimated variance. In terms of
covariates, the lagged wind direction (``persistence'') is mostly responsible
for distinguishing the broad range of wind regimes listed above while the
pressure gradients and wind speed separate the data into subgroups with high
vs.\ low precision. A more extensive case study of circular regression trees
and forests applied to probabilistic wind direction forecasting at
Innsbruck Airport and Vienna International Airport is presented in Section~\ref{sec:wind},
along with a benchmark against commonly-used alternative approaches.

The remainder of the paper is structured as follows: The theory on
probabilistic circular modeling introducing the von Mises distribution, and
circular regression models are discussed in Section~\ref{sec:prob_circ}.
The methodology of circular regression trees and forests and their features are
introduced in Section~\ref{sec:tree_forest}. After the case study presented in
Section~\ref{sec:wind}, a comprehensive summary and conclusions are given in
Section~\ref{sec:summary}.

% -------------------------------------------------------------------
% Probabilistic circular modeling
% -------------------------------------------------------------------
\section{Probabilistic circular modeling}
\label{sec:prob_circ}
Probabilistic modeling of circular data requires the selection of a probability
distribution which accounts for the periodicity of circular data. Generally,
this feature can be obtained by ``wrapping'' the probability density function
of any continuous distribution around the unit circle \citep{Mardia+Jupp:1999}.
In that way, the wrapped Cauchy distribution or the wrapped normal distribution
can be employed to model symmetric unimodal circular data. A close
approximation to the wrapped normal distribution that is mathematically simpler
and hence easier to use is provided by the von Mises
distribution \citep{Fisher:1993}, a purely circular distribution which is also known as ``the
circular normal distribution'' and is a common choice for probabilistic
modeling of circular data. Based on a location parameter $\mu \in [0, 2\pi)$
and a concentration parameter $\kappa > 0$ the density of the von Mises
distribution for an observation $y \in [0, 2\pi)$ is given by
%%
\begin{equation}
  f_\mathrm{vM}(y; \mu, \kappa) = \frac{1}{2 \pi I_0(\kappa)}~e^{ \kappa \cos(y - \mu)},\label{schlosser:equ_vm}
\end{equation}
%%
where $I_0(\kappa)$ is the modified Bessel function of the first kind and order
$0$ (see, e.g., \citealt{Jammalamadaka+Sengupta:2001} or
\citealt{Ley+Verdebout:2017} for a more detailed overview).

\begin{figure}[t]
  \begin{minipage}{.05\textwidth}
    \hfill
  \end{minipage}%
  \begin{minipage}{.45\textwidth}
    \vspace{-1em}
    \includegraphics[width=1.05\linewidth]{density_linear.pdf}
  \end{minipage}%
  \begin{minipage}{.45\textwidth}
    \includegraphics[width=1.2\linewidth]{density_circular.pdf}
  \end{minipage}
  \begin{minipage}{.05\textwidth}
    \hfill
  \end{minipage}%
  \caption{\label{fig:densities} Illustration of a von Mises model for circular data in
    the interval $[0, 2\pi)$ fitted by maximum likelihood. Left: Linearized
    scale. Right: Circular scale. In both panels the empirical histogram (gray bars)
    and fitted density (red line) are depicted along with the estimated location
    parameter (red hand).}
\end{figure}

The corresponding log-likelihood function is defined as
%%
\begin{equation}
    \ell(\mu, \kappa; y) = \log(f_\mathrm{vM}(y;\mu, \kappa))
    = -\log(2 \pi I_0(\kappa)) + \kappa \cos(y - \mu).\label{equ:vM:loglik}
\end{equation}
%%
To fit a probabilistic model $\text{vM}(y; \mu, \kappa)$ to a circular response $y$, the
distribution parameters $\mu$ and $\kappa$ need to be estimated.
This can be done by
maximizing the log-likelihood function $\ell(\mu, \kappa; y)$
%%
\begin{equation}\label{eq:optim}
  (\hat{\mu},\hat{\kappa}) = \argmax_{\mu,\kappa} \sum_{i=1}^n \ell(\mu, \kappa; y_i) 
\end{equation}
%%
yielding maximum likelihood estimators $\hat{\mu}$ and $\hat{\kappa}$ such that
a fully specified distributional model is fitted to the learning data $\{y_i\}_{i=1,\ldots,n}$.

The score function
%%
\begin{equation}
  \label{eq:scores}
  \begin{aligned}
    s(\mu, \kappa, y) &= \bigg(\frac{\partial \ell}{\partial \mu}(\mu,\kappa; y)
    \smallskip &,
    \quad & \frac{\partial \ell}{\partial \kappa}(\mu,\kappa; y) \bigg) 
    \\
    &= \bigg(\kappa \sin(y-\mu)
    \smallskip &,
    \quad &-\frac{I_1(\kappa)}{2\pi I_0(\kappa)}+\cos(y-\mu) \bigg)
  \end{aligned}
\end{equation}
%%
provides a way to obtain a measure of goodness of fit of the model for
each observation and fitted parameter. Then, the optimization problem in
Equation~\ref{eq:optim} can alternatively be specified as 
%%
\begin{equation}
  \sum_{i = 1}^n s(\hat{\mu}, \hat{\kappa}, y_i) = 0.
\end{equation}
%%
Figure~\ref{fig:densities} depicts von Mises model for circular data in
$[0, 2 \pi)$ fitted by maximum likelihood, either using a linearized
(left) or circular (right) scale. In both cases,
the empirical histogram (gray bars) is shown along with the fitted density (red line) and
estimated location parameter (red hand). However, this distributional model
only considers the circular response variable but no covariate. Of course,
including covariates is of interest in a regression setup for forecasting.

In most generalized linear model (GLM) or generalized additive model (GAM)
approaches to circular regression the location parameter $\mu$ depends on covariates $\mathbf{z}$
through a link function $g(\cdot)$, circular intercept $\mu_0$ and
coefficient vector $\mathbf{\beta}$:
\begin{equation}
\mu = \mu_0 + g\left(\mathbf{\beta}^{\top} \mathbf{z}\right).
\end{equation}
The link function transforms the additive predictor to an interval of length
$2\pi$.  Typically $g(x) = 2\cdot \text{arctan}(x)$ is employed, as suggested by
\citet{Fisher+Lee:1992}.  They also developed a heteroscedastic version by
combining two individual GLMs, each for one of the parameters $\mu$ and
$\kappa$. This provides a first approach to a fully probabilistic regression
model for circular data, albeit the parameters are not regressed
simultaneously on covariates as in the more general framework provided by
generalized additive models for location, scale, and shape \citep[GAMLSS,
][]{Rigby+Stasinopoulos:2005}. Nevertheless, other circular additive models
share the previously discussed difficulties induced by the
characteristics of the log-likelihood function and the strongly nonlinear link
function. Referring to additive models in general, it has to be considered that
a proper model specification can be very challenging, particularly for a high
number of covariates and no information on possible interactions. Moreover, the
additive structure might impose a smooth effect even if the true underlying
effect is an abrupt shift, which occurs, e.g., in atmospheric wind fields.

In contrast, the tree-based circular regression models proposed in the next sections
largely avoid the problems above by employing recursive partitioning in combination
with local adaptive likelihood estimation.

% -------------------------------------------------------------------
% Circular regression trees and forests
% -------------------------------------------------------------------
\section{Circular regression trees and forests}
\label{sec:tree_forest}

Starting out from the ideas of \citet{Lund:2002}, we introduce circular regression trees and
forests considering splits in all distribution parameters of the von Mises
distribution and providing a full probabilistic model.  Moreover, the resulting
tree-based models provide a very intuitive and data-driven alternative to
commonly used GLMs for circular data.

\subsection{Circular regression trees}
\label{sec:circtree}
Fitting a global model to a full data set can be very challenging, particularly
for complex data with substantial variations. Therefore, separating the data set
into more homogeneous subgroups based on covariates before fitting a local
model in each of these subgroups allows to capture (potential) group-specific
effects more precisely and hence can result in an overall better-fitting model.
This is the general idea of regression trees which are combined with
distributional modeling in \cite{Schlosser+Hothorn+Stauffer:2019}. Specifying a
full distributional model in each node of the tree yields a distributional
regression tree, where selecting the von Mises distribution enables an
application to circular data. The crucial step of how and where to split the
data can be accomplished with the unbiased recursive partitioning algorithms
MOB \citep{Zeileis+Hothorn+Hornik:2008} or CTree
\citep{Hothorn+Hornik+Zeileis:2006}. For this purpose, model scores are
obtained by evaluating the score function~$s(\cdot)$ for each individual observation
at the parameter estimates (Equation~\ref{eq:scores}). For the von Mises
distribution with its two distribution parameters ($\mu$ and $\kappa$) and a
data set of $n$ observations, this yields an $n \times 2$ matrix that can be
employed as a discrepancy measure, capturing how well each given observation
conforms with the estimated location $\hat{\mu}$ and precision $\hat{\kappa}$,
respectively.  To capture dependence on covariates, the association between the
model's scores and each available covariate is assessed using either a
parameter instability test~(MOB) or a permutation test~(CTree). By doing so in
each partitioning step, the covariate with the highest significant association
(i.e., lowest significant $p$-value, if any) is selected for splitting the
data. The corresponding split point is chosen either by optimizing the
log-likelihood~(MOB) or a two-sample test statistic~(CTree) over all possible
partitions. This procedure is repeated recursively until there are no
significant parameter instabilities or until another stopping criterion is met
(e.g., subgroup size or tree depth). A more detailed description of the applied
tree-building algorithm can be found in Appendix~\ref{app:algorithm}.

Once a distributional tree model is fitted it can be employed to obtain
probabilistic predictions for a possibly new set of observed covariates $\bm{z}
= (z_1, \ldots, z_m)$. Starting at the root node, the tree structure leads the
observation to a terminal node where the parameter pair $(\hat{\mu},
\hat{\kappa})$ is estimated for the corresponding subset of learning
observations. This can also be expressed by employing weights which indicate
whether the $i$-th learning observation and the observation $\bm{z}$ belong to
the same terminal node: 
%%
\begin{equation}
  w^{\text{tree}}_i(\bm{z}) = \sum_{b=1}^B \mathbf{1}((\bm{z}_i \in
  \mathcal{B}_b) \land (\bm{z} \in \mathcal{B}_b)).
\end{equation} 
%%
Here, $\mathbf{1}(\cdot)$ is the indicator function and $\mathcal{B}_b$ is the
$b$-th out of $B$ segments partitioning the covariate space in disjoint
subsets. Then the estimated parameter pair
$(\hat{\mu},\hat{\kappa})(\bm{z})$ specifying the predicted von Mises
distribution for a given $\bm{z}$ is obtained by a weighted maximum likelihood estimator:
%%
\begin{equation} 
  (\hat{\mu},\hat{\kappa})(\bm{z}) =
  \argmax_{\mu,\kappa} \sum_{i=1}^n w^{\text{tree}}_i(\bm{z}) \cdot
  \ell(\mu,\kappa; y_i).
\end{equation}
%%
Therefore, the same parameter pair is estimated for all observations belonging
to the same terminal node, which speeds up computation since the parameter
estimates do not need to be recalculated for each (new) observation via maximum
likelihood but can be extracted directly from the learning sample and the
fitted model.

While tree models can capture non-additive effects, their structure and the
consequential strict separation of data into subgroups hinders an adequate
depiction of smooth effects. They can be included by combining an ensemble of
trees in order to obtain a regression forest, which also stabilizes the model.

\subsection{Circular regression forests}
\label{sec:circforest} 
A natural extension of (circular) regression trees are ensembles or forests
that can improve forecasts by regularizing and stabilizing the model. Random
forests introduced by \cite{Breiman:2001} average the predictions of an
ensemble of trees, each built on a subsample or bootstrap of the original data.
A generalization of this strategy is to obtain weighted predictions by adaptive
local likelihood estimation of the distributional parameters
\citep[Section~2.3. of][]{Schlosser+Hothorn+Stauffer:2019,
Hothorn+Zeileis:2017}. More specifically, for each (possibly new)
observation~$\bm{z}$ a set of averaged ``nearest neighbor'' weights
$w_i^\text{forest}(\bm{z})$ is obtained that is based on the number of trees in
which $\bm{z}$ is assigned to the same terminal node as each learning
observation $y_i, i \in \{1,\ldots,n\}$. Hence, for a forest of $T$ trees, the
weights are calculated via
%%
\begin{equation}
  w^{\text{forest}}_i(\bm{z}) = \frac{1}{T} \sum_{t=1}^T \sum_{b=1}^{B^t}
  \frac{\mathbf{1}((\bm{z}_i \in \mathcal{B}^t_b) \land (\bm{z} \in \mathcal{B}^t_b))}{|\mathcal{B}^t_b|},
\end{equation}
%%
where $|\mathcal{B}^t_b|$ denotes the number of observations in the $b$-th
segment of the $t$-th tree. Therefore, similar observations ending up more
often in the same terminal node have higher weights and thus a stronger
influence in the weighted maximum likelihood estimation.

In that way a specific set of weights can be calculated for each observation
yielding its specific parameter estimates for the von Mises distribution
%%
\begin{equation}
  (\hat{\mu},\hat{\kappa})(\bm{z}) = \operatorname{argmax}\displaylimits_{\mu,
  \kappa} \sum_{i=1}^n w_i^\text{forest}(z) \cdot \ell(\mu, \kappa; y_i). 
\end{equation}
%%

Therefore, the resulting parameter estimates can smoothly adapt to the given
covariates $\bm{z}$ whereas $w_i(\bm{z}) = 1$ would correspond to the
unweighted full-sample estimates and $w_i(\bm{z}) \in \{0, 1\}$
corresponds to the subgroup selection from a tree. Thus, circular regression forests
can capture both smooth and abrupt changes, while covariates and possible
interactions are selected automatically and do not explicitly need to be
specified beforehand.

% -------------------------------------------------------------------
% Case study
% -------------------------------------------------------------------
\section{Case study: Probabilistic wind direction forecasting}
\label{sec:wind}
As motivated in Section~\ref{sec:introduction}, accurate forecasts of wind
directions are of great importance for risk management in various fields such
as agriculture, energy production, or aviation. For example, in order to direct
airplanes to a safe landing, precise knowledge of wind direction for the next
hour(s) at the respective airport is highly desirable and adequate prediction
methods are required. This section exemplifies the use of circular regression
trees and forests with wind direction forecasts for two Austrian airports --
one in flat terrain, the other one in mountainous terrain. The results are
benchmarked against alternative probabilistic forecasting methods. The study
is based on $1$\,h and $3$\,h forecasts employing lagged observations in the
vicinity of the airports as possible predictor variables.

\subsection{Data}
%%
\begin{figure}[t]
  \setkeys{Gin}{width=1.0\textwidth}
  \centering
  \includegraphics{_plot_circforest_stationmap_combined.pdf}
  \caption{Overview of the study area for Innsbruck Airports (left) and Vienna International Airport (right). 
    For Innsbruck, $4$~stations at the airport and $6$~stations along the
    intersecting valleys are used, whereas, for Vienna, $9$~stations at the airport
    and $13$~stations within the vicinity of approximately 30\,km are used.
    Elevation data are obtained from the TandDEM-X digital elevation model with a 
    horizontal resolution of $90$~m \citep{Wessel:2018}.}
  \label{fig:overview} 
\end{figure}
%%
\label{sec:wind:data}
The circular response variable considered in this case study is a $10$\,min-average
of wind direction measurements at Innsbruck Airport (INN) and Vienna International Airport (VIE) on
an hourly temporal resolution. Temporal information and $1$-hourly resolved $10$\,min mean observations of various
meteorological quantities are used as predictor variables, including wind direction, wind speed, temperature, air
pressure and humidity, all lagged by one or three hours according to the
respective forecasting step. The meteorological variables are measured either
directly at the airports or within their vicinities. For Innsbruck,
measurements at the airport and along the intersecting valleys are used,
whereas, for Vienna, measurements at the airport and within its vicinity of
approximately 30\,km are used. Figure~\ref{fig:overview} provides a topographical
overview of the airports and their surrounding areas with the station sites employed
in this study. In addition, we use derived quantities
such as 3-hourly means, minima and maxima, as well as 1-and 3-hourly temporal
changes and spatial differences towards the airport of the respective
quantities. An overview of the employed data sets can be found in
Table~\ref{tab:data}.

The data used in this study consists of four subsequent years from January 2015
to December 2018. After first eliminating predictor variables with more than
5\% missing values and then time points with any missing observations, the data set
consists of 41\,979 time points and 260 covariates for Innsbruck, and of
38\,985 time points and 494 covariates for Vienna, respectively.

%%
\begin{table}[t!]
 \caption[Table caption text]{Overview of the data sets employed in the case
    study: For Innsbruck and Vienna, various meteorological variables and derived
    quantities of these are considered at the respective stations, located either
    directly at the airports or in their vicinities.}
  \label{tab:data}
  \begin{center}
    \begin{tabular}{l  l}
      \hline
      Data components               & Description \\
      \hline
      Temporal information:         & Time of the day, day of the year \\
      \noalign{\vskip 1mm}   
      Meteorological variables:     & Wind direction, wind (gust) speed, \\
                                    & (reduced) air pressure, relative humidity, \\
                                    & temperature\\ 
      \noalign{\vskip 1mm}                           
      Derived quantities:           & 3-hourly means/minima/maxima, \\
                                    & 1-hourly and 3-hourly temporal changes, \\
                                    & spatial differences towards the airport\\
      \noalign{\vskip 1mm}                              
      Weather stations (Innsbruck): & 4 stations at the airport, \\
                                    & Igls, Kematen, Kufstein, Landeck, Patscherkofel, \\
                                    & Steinach\\
      \noalign{\vskip 1mm}
      Weather stations (Vienna):    & 9 stations at the airport, \\
                                    & Arsenal, Donaufeld, Exelberg, G\"anserndorf, \\
                                    & Gro{\ss}-Enzersdorf, Gumpoldskirchen, Hohe Warte,\\
                                    & Innere Stadt, Jubil\"aumswarte, Mariabrunn, \\
                                    & Seibersdorf, Unterlaa, Wolkersdorf\\
      \hline
    \end{tabular}
  \end{center}
\end{table}
%%

\subsection{Models and evaluation}
\label{sec:wind:models}
For a fair evaluation of circular regression trees and forests, and to
investigate whether they can be applied as a reasonable alternative to already
existing approaches, three additional statistical models are employed in this
study for probabilistic forecasting of wind directions. Two of them are based
on existing approaches used in the meteorological field, while the third is a
state-of-the-art GLM-type model to forecast circular response variables.

\begin{itemize}
  \item \emph{Climatological model:} 
    Accurate knowledge of weather quantities' climatologies can be important for a
    wide range of applications. While forecasts based on climatologies, by
    construction, do not adapt to the current weather situation they are still a
    useful baseline for the validation of newly developed forecasting systems
    \citep{Simon+Umlauf+Zeileis:2017, Stauffer+Mayr+Messner:2017}. 
    \newline
    Specifically, the climatology employed in the following uses all observations
    at the same time (to adapt to daily cycles) in a window of $31$~days centered
    around the day of interest (to adapt to seasonal cycles)
    in all available years in the sample. Based on these observations a probabilistic
    model is obtained by maximum likelihood estimation as described in
    Section~\ref{sec:prob_circ}. This approach follows \citet{Vogel+Knippertz+Fink:2018}
    and is discussed in a comprehensive summary on different time-adaptive training schemes in
    \citet{Lang+Lerch+Mayr:2019}.
    %%
  \item \emph{Persistency model:}
    The persistence describes the previous value of a single weather quantity
    in a time series. Like the climatology it is a very basic prediction model that
    is often applied as a baseline reference in weather forecasts
    \citep{noaasnationalweatherservice:2019}. Especially in nowcasting tasks with very short
    forecasting steps, the persistence can provide very good estimates.
    \newline
    To gain a full probabilistic persistency model, we proceed similarly as for
    the climatological model by using maximum likelihood estimation and fitting the
    distribution parameters of the von Mises distribution conditional on lagged
    response values according to the description in Section~\ref{sec:prob_circ}.
    We fit one model for every hour throughout the validation data set employing
    the previous six lagged response values as training data. In order to allow for
    a stronger influence of observations closer to the time of interest,
    exponential smoothing is employed with a smoothing factor of $0.5$;
    accordingly, for every prediction an equal influence rate of 50 percent is
    assigned both to the current observation and to the previous five observations
    together. Observations with longer time lags have exponential weights below
    $0.01$ and are therefore omitted from the training data.
    %%
  \item \emph{Generalized linear model:}
    Traditional approaches to forecast circular response variables are often
    based on circular GLM-type models \citep{Fisher:1993}. As discussed in
    Section~\ref{sec:introduction}, circular regression models often experience the
    problem that the likelihood function can be strongly irregular which makes
    optimization rather difficult. Hence, they often do not converge if no
    appropriate initial values are provided \citep{Pewsey+Neuhaeuser+Ruxton:2013,
    Gill+Hangartner:2010}. In this study, to be able to employ a GLM out of the box
    as a reference, we use the Bayesian implementation of
    \cite{Mulder+Klugkist:2017} which depends less on initial values due to an MCMC
    sampling algorithm using weakly informative priors.  \newline Following
    \cite{Mulder+Klugkist:2017} the model uses a link function~$g$ to keep the
    response values within an interval of length $2\pi$. As the implementation
    cannot handle circular covariates, we use the components of the lagged
    2-dimensional wind vector~$(u, v)^{\top}$ and the lagged
    wind speed $spd$ as predictor variables. The model formula for the location
    parameter $\mu$ of the von Mises distribution can be written as:
    %%
    \begin{align}
      \mu = \beta_0 + g(\beta_1 \cdot u + \beta_2 \cdot v + \beta_3 \cdot spd)
    \end{align}
    %%
    with $\beta_0$ being a circular intercept, $\beta_\bullet$ the regression
    coefficients and $g$ the link function $g(x) = 2 \cdot \text{arctan}(x)$. In
    addition, a constant concentration parameter $\kappa$ is fitted to the full
    learning sample. To allow for seasonally varying error characteristics in both
    bias and slope coefficients, and to allow for seasonal heteroscedasticity
    captured by the concentration parameter, we use the same time-adaptive
    training approach as for the climatological model; hence, separate models are
    estimated over all observation dates, using the same time of $31$~days
    centered around the day of interest over all available years in the training
    data \citep{Lang+Lerch+Mayr:2019}.
    %%
  \item \emph{Circular regression tree:}
    For the circular regression tree introduced in Section~\ref{sec:circtree},
    all covariates provided in the learning data can be considered due to an
    intrinsic automatic variable selection performed in the tree estimation. The
    tree is built with the newly developed \textsf{R} package \pkg{circtree} employing the
    CTree algorithm \citep{Hothorn+Hornik+Zeileis:2006} using a minimal number of
    2000 observations in each terminal node (argument \code{minbucket}).
    %%
  \item \emph{Circular regression forest:} Following the description in
    Section~\ref{sec:circforest}, the circular forest used in this study is
    constructed based on $100$~individual trees employing the \textsf{R} package
    \pkg{circtree}. Each of these trees is again built by the CTree algorithm on a
    subsample containing $30$~percent of the original learning data.  All
    covariates are included for building each tree which ensures that the lagged
    response variable is always considered for splitting. This bagging approach can
    be applied in \pkg{circtree} by setting the argument \code{mtry} to the total
    number of covariates. Since a high number of possible split points leads to
    high computational costs, the covariates are binned in a maximum of $50$
    classes (argument \code{nmax = c(yx = Inf, z = 50)}). Contrary to a
    single-tree model, forests usually consist of very large trees as they are not
    prone to overfitting the data due to the stabilization obtained by combining
    the individual trees. Therefore, we use the following control arguments to
    build rather large trees: The minimal number of observations to perform a split
    is set to 20 (argument \texttt{minsplit}), the minimal number of observations
    in each terminal node is set to 7 (argument \texttt{minbucket}), and the
    significance level for variable selection is kept at its maximum value of $1$
    (argument \texttt{alpha}).
\end{itemize}

To compare the predictive performance of all proposed models, a circular
analogue of the continuous ranked probability score (CRPS) as introduced by
\cite{Grimit+Gneiting+Berrocal:2006} is computed. Just as the linear version of
the CRPS \citep[for more details see][]{Hersbach:2000} it is a proper scoring
rule \citep{Gneiting+Raftery:2007} and measures the difference between an
observation and the corresponding predicted distribution function in order to
assess the probabilistic goodness of fit for the estimated model. Hence, the
lower the CRPS value the better the predictive performance. Contrary to the
linear version, the circular CRPS reduces not to the absolute error but to the
angular distance when the forecast is deterministic.

In addition to the raw CRPS, corresponding skill scores are computed to assess
differences in the improvement of the various statistical models over the
climatological model used as a reference:
%%
\begin{equation} 
  \text{CRPSS}_{\text{model}} = 1 -
  \frac{\text{CRPS}_{\text{model}}}{\text{CRPS}_{\text{climatology}}}.
\end{equation}
%%

All scores presented in the next section are computed out-of-sample based on
five years of data. For the persistency model only dates prior to the
time of interest are used and the validation is performed rolling over all
observations. For all other models a five-fold cross-validation is employed
using up to four calendar years for model training and the remaining single
calendar year for validation. Due to the large sample size of 24 hourly values
per day over five years, some kind of temporal aggregation is needed to ensure
a correct visual comparison of the individual methods. The analyses performed
have shown that for the employed models the variability of the predictive
performance over the five years is lower than over a single day or over a
single year. Hence, CRPS and CRPS skill scores are aggregated over the
respective five validation years which yields 24 hourly scores per month
averaged over the five validation years.

\subsection{Results} 
This section provides a detailed analysis on the predictive performance of the
different proposed statistical models applied to probabilistic wind direction
forecasting. To ensure a comprehensive comparison of the models, wind direction
forecasts are evaluated for two different lead times at two airports with
different climatological site characteristics.
Figure~\ref{fig:boxplot_crpsraw} shows the CRPS values of the employed models
at forecast steps 1\,h and 3\,h for Innsbruck (Panels~a,\,c) and
Vienna (Panels~b,\,d). The scores are aggregated over the five validation
years, yielding yearly mean values for every hour per calendar month, with a
lower score indicating better performance. The circular forest overall provides
the best predictive performance, followed by the circular regression tree and
the persistency model for both stations at both forecast steps; except for the
$3$\,h~forecasts at Innsbruck where the persistency model is outperformed by
all others. In comparison to the circular regression tree and forest, for both
stations and forecast steps, the climatological model and the linear model show
clearly higher CRPS values and hence a lower predictive performance.
%%
\begin{figure}[t]
  \centering
  \setkeys{Gin}{width=1.0\textwidth}
  \includegraphics{_plot_circforest_validation_crpsraw_agg_comparison_with_lowff_v14b.pdf}
  \caption{CRPS skill scores of wind direction forecasts based on the full
    predictive von~Mises distribution for $+1$\,h and $+3$\,h forecasts at the
    Innsbruck Airport and Vienna International Airport. Each box-and-whisker contains $24$~hourly scores
    for each of the $12$~months averaged over the $5$~validation years which
    yields a total of $288$~yearly mean values. The scores are shown for the
    climatological, persistency and the linear model as well as for the circular
    regression tree and forest.}
  \label{fig:boxplot_crpsraw} 
\end{figure}
%%
The different site characteristics of the airports Innsbruck
(Figure~\ref{fig:boxplot_crpsraw}\,a,\,c) and Vienna
(Figure~\ref{fig:boxplot_crpsraw}\,b,\,d) seem to have an effect on the
absolute level of the model performances and on their respective predictive
performance variances. At Innsbruck, due to the surrounding mountains only a
limited number of possible wind directions exists, namely the three wind
regimes discussed for Figure~\ref{fig:tree_ibk} in
Section~\ref{sec:introduction}. Therefore, for Innsbruck the wind direction
remains rather constant in one of these possible states, but once a change
takes place it is mostly a major wind direction shift, e.g., from up-valley to
down-valley. Due to the few wind regimes the rather inflexible climatological
and linear models score relatively well with similar CRPS values as the other
models (Figure~\ref{fig:boxplot_crpsraw}\,a,\,c). In addition, at Innsbruck the
potential high prediction errors in case of a change of the wind regime seem to
lead to a higher variation in the predictive performance for all models in
comparison to Vienna; this variation is especially high for the persistency
model due to its strong vulnerability to abrupt wind shifts. On the contrary,
at Vienna the geographic surroundings do not constrain any wind direction which
leads to smaller and less abrupt changes in the wind direction as well as less
pronounced wind regimes. This seems to weaken the predictive performance of the
climatological and linear models, and to reduce the performance variability for
all models (Figure~\ref{fig:boxplot_crpsraw}\,b,\,d).

The different forecast steps have apparently only a minor effect on the
predictive performance of the climatological model and the linear model at both
stations. As expected, for the persistency model, at both stations, higher
scores for the $3$\,h forecast (Figure~\ref{fig:boxplot_crpsraw}\,c,\,d) reveal
a lower performance for longer lead times; this is due to the lower information
content of 3-hourly instead of 1-hourly lagged response values employed
as covariates in the persistency model. The circular regression tree and
forest seem to partially compensate for the lower skill of the lagged response
values by other covariates, hence their predictive performance only slightly
decreases for the longer lead time. This compensation is especially evident for
Innsbruck, where the performance difference between the persistency model and
the tree-based methods significantly increases from the 1-hourly to the
3-hourly forecast.

In addition to the raw CRPS (Figure~\ref{fig:boxplot_crpsraw}), CRPS skill
scores with the climatological model as a reference are provided in
Figure~\ref{fig:boxplot_crpsskill}. Skill scores are in percent, where positive
values indicate an improvement in the predictive performance over the
reference. For all setups, the circular forest has the highest skill scores
with a mean performance gain of 13--25\% and 58--71\% for Innsbruck and Vienna,
respectively. As discussed for Figure~\ref{fig:boxplot_crpsraw}, this
improvement over the climatological model is lower for Innsbruck due to the low
number of predominant wind regimes and hence a relatively good performance of
the climatological model. Additionally, Figure~\ref{fig:boxplot_crpsskill}
shows that while the persistency model's performance is lower than the
reference (Panel~c) the tree-based models can compensate for the low skill of
the lagged response values employed as covariates and, hence, are still
significantly superior to the reference.

%%
\begin{figure}[t]
  \centering
  \setkeys{Gin}{width=1.0\textwidth}
  \includegraphics{_plot_circforest_validation_crpsskill_agg_comparison_with_lowff_v14b.pdf}
  \caption{As Figure~\ref{fig:boxplot_crpsraw}, but showing CRPS skill scores
    with the climatology model as reference. Skill scores are in percent; positive
    values indicate improvements over the reference.} 
  \label{fig:boxplot_crpsskill} 
\end{figure}
%%

% -------------------------------------------------------------------
% Summary
% -------------------------------------------------------------------
\section{Summary and conclusion}
\label{sec:summary}
Extending the toolbox for modeling circular data, circular regression trees and
forests are established by coupling model-based recursive partitioning with the
von Mises distribution. By separating the data into more homogeneous
subgroups, possible difficulties in circular regression are bypassed as
covariates are solely considered for splitting and group-specific models are
fitted without further covariates. In addition, by specifying the von Mises
distribution for each node and allowing for splits in both distribution
parameters $\mu$ and $\kappa$, fully probabilistic forecasts are provided.

The performance of the novel circular trees and forests is assessed in a 
case study for shortterm probabilistic wind direction forecasting at two
airports with different site characteristics. As benchline models,
probabilistic climatology and persistency models, as well as a state-of-the-art
circular GLM-type model are evaluated based on proper scoring rules. In summary, the
circular trees and forests have the highest predictive performance in this
setting. For cases without changes in the wind regime, lagged response values
provide already highly skillful estimates leading to a good performance of the
persistency model as observed for shortterm wind direction forecasts in this
study. While in these cases the trees and forests also benefit from the highly
informative lagged response, they can compensate for a lower information of
this covariate by incorporating other quantities and possible interactions of
these, contrary to the persistency model (see
Figure~\ref{fig:boxplot_crpsskill}). Hence, the tree-based models provide
reliable forecasts in all tested meteorological settings. For operational use,
a possible extension could be the incorporation of numerical weather
predictions as (additional) covariates. While this probably only slightly
improves the predictive skill for short leadtimes, it possibly extends the
potential forecast range of the different methods. Another possible
improvement for obtaining more parsimonious circular trees is to consider
splitting circular covariates into two circle segments by searching two split
points simultaneously rather than sequentially at different depths. While this
might slightly improve the predictive performance of circular trees, this
should not affect the performance of circular forests, as these consist of
very large trees with many different splits.

To conclude, in general the tree structure can capture nonlinear changes,
shifts, and potential interactions in covariates without prespecification of
such effects. As supported by the presented case study, this can be
particularly useful for modeling a highly fluctuating response, such as
typically observed for wind direction, or/and in case of a large number of
possible covariates. Moreover, the case study shows that building ensembles of
circular trees can even improve the forecasting performance, as the resulting
circular forests allow for modeling smooth effects and stabelize the model.

\section*{Acknowledgments}
This project was partially funded by the Austrian Research Promotion
Agency~(FFG) grant number~$858537$. Torsten Hothorn received funding from the
Swiss National Science Foundation, grant number~$200021\_184603$. Lisa
Schlosser received a PhD scholarship granted from the University of Innsbruck.

\section*{Computational details}
The corresponding implementation of the proposed methodology for circular
trees and forests is provided in the \textsf{R} package \textbf{circtree} 
(version~\Sexpr{packageVersion("circtree")}).
% The proposed circular tree and forest models are implemented in the \textsf{R}
% package \textbf{circtree} (version~\Sexpr{packageVersion("circtree")}). 
The package is based on the \textbf{disttree} package
(version~\Sexpr{packageVersion("disttree")}) which applies the main tree
building functions from the \textbf{partykit} package
(version~\Sexpr{packageVersion("partykit")}). All three packages are available
on \textsf{R}-Forge at \url{https://R-Forge.R-project.org/projects/partykit/}.

For the circular GLM considered as reference model the corresponding
implementation is provided in the \textsf{R} package \textbf{circglmbayes} by
\cite{Mulder+Klugkist:2017}. In particular the function \code{circGLM} is
applied to estimates the intercept and regression coefficient along with the
concentration parameter.

%\newpage
\bibliography{ref.bib}

\newpage
\begin{appendix}

\section{Tree algorithm}
\label{app:algorithm}
This section provides a more detailed overview on the permutation-test-based CTree algorithm 
\citep{Hothorn+Hornik+Zeileis:2006}, 
specifically for circular data as applied for building circular regression trees and forests 
presented in this case study. 
An alternative tree-building framework is provided by the MOB algorithm, which is based on 
M-fluctuation tests \citep[see][for more details]{Zeileis+Hothorn+Hornik:2008}.

%\subsection*{CTree}
In the following, the testing and splitting strategy is described for the root node of the tree
which contains the entire learning sample.
%consisting of circular observations denoted by $\{y_i\}_{i = 1,\ldots,n}$, $n \in \mathbb{N}$. 
For a complete tree model, the same procedure is applied iteratively to all resulting child nodes 
with the corresponding subsamples.

First, employing the von Mises distribution, a distributional model $\text{vM}(y; \mu,\kappa)$ 
is fitted to the learning sample of circular observations $\{y_i\}_{i = 1,\ldots,n}$ as explained 
in Section~\ref{sec:prob_circ}.
In a next step, a goodness-of-fit measurement %of $\text{vM}(Y; \mu,\kappa)$ 
is obtained for each parameter and each observation by evaluating the score 
function $s(\mu, \kappa, y)$ at the estimated location and concentration 
parameter $\hat{\mu}$ and $\hat{\kappa}$.
To detect dependencies between the resulting scoring matrix
\begin{equation}
\begin{pmatrix} 
s(\hat{\mu}, \hat{\kappa}, y_1)_1 \quad, & s(\hat{\mu}, \hat{\kappa}, y_1)_2\\
\vdots & \vdots\\
s(\hat{\mu}, \hat{\kappa}, y_n)_1 \quad,& s(\hat{\mu}, \hat{\kappa}, y_n)_2
\end{pmatrix}
\end{equation}
and each possible split variable $z_l \in \{z_1, \ldots, z_m\}$ a permutation test is applied. 
In particular, the null hypotheses
of independence of each split variable and the scores
%\begin{align}
%H_0^l:  s(\hat{\mu}, \hat{\kappa}, Y) \qquad \bot \qquad Z_l
%\end{align}
is assessed by employing the multivariate linear statistic
\begin{equation}
t_l = vec\left(\sum_{i=1}^n v_l(z_{li}) \cdot s(\hat{\mu}, \hat{\kappa}, y_i)\right)
\end{equation}
with $s(\hat{\mu}, \hat{\kappa}, y_i) \in \mathbb{R}^{1\times 2}$. 
For a numeric split variable $z_l$ the transformation function $v_l$
is simply the identity function $v_l(z_{li}) = z_{li}$ such that $t_l \in \mathbb{R}^2$ 
as the ``vec'' operator converts the matrix of dimension $1 \times 2$ into a $2$ column vector.
If $z_l$ is a categorical variable with $h$ categories then 
$v_l(z_{li}) = (\I(z_{li} = 1), \ldots, \I(z_{li} = h))$, hence, $v_l$ returns a unit vector
of dimension $h$ where the entry $1$ indicates the category of $z_{li}$. 
In this case the ``vec'' operator converts the $h \times 2$ matrix into a $h \cdot 2$ column vector 
by column-wise combination such that $t_l \in \mathbb{R}^{h \cdot 2}$. 
If there are any observations with missing values these are not included in the calculation of $t_l$.

To map the %observed 
multivariate linear statistic $t_l$ onto the real line a univariate test statistic~$c$ is employed, for example in a quadratic form
\begin{equation}
c_{\text{quad}}(t_l,\mu_l,\Sigma_l) = (t_l-\mu_l)\Sigma_l^+(t_l-\mu_l)^{\top}
\end{equation}
where $\mu_l$ and $\Sigma_l$ are the conditional expectation and the covariance of $t_l$,
as derived by \cite{Strasser+Weber:1999} and used for standardization, and $\Sigma_l^+$ 
is the Moore-Penrose inverse of $\Sigma_l$.
As an alternative, also a maximum form ($c_{\text{max}}$) can be considered such that
the maximum of the absolute values of the standardized linear statistic is returned.

The asymptotic conditional distribution of $c(t_{l},\mu_{l},\Sigma_{l})$ 
is either normal (for $c_{\text{max}}$) or $\chi^2$ (for $c_{\text{quad}}$) owing to 
the asymptotic conditional distribution of the linear statistic $t_l$ being a multivariate 
normal with parameters $\mu_l$ and $\Sigma_l$ \citep{Strasser+Weber:1999}.
With this knowledge at hand, the corresponding $p$-values can be calculated and used
to select the best splitting variable. A small $p$-value corresponding to 
$c(t_{l},\mu_{l},\Sigma_{l})$ indicates a strong discrepancy from the assumption of independence 
between the scores and the split variable $z_l$.
Therefore, if any of the Bonferroni-adjusted $p$-values is beneath the selected significance level,
the partitioning variable $z_{l^\ast}$ with the lowest $p$-value is selected as split variable, otherwise 
no split is performed. This early stopping induced by the significance level is referred to as 
``pre-pruning'' which is often avoided for forest models by setting the significance level to $1$.

To select the best split point within the already chosen split variable, again, a linear test statistic
is employed. In particular, for a breakpoint $r$ of the variable $z_{l^{\ast}}$ leading to two subgroups 
$\mathcal{B}_{1r}$ and $\mathcal{B}_{2r}$ the discrepancy between score functions in the subgroups
is measured by evaluating
\begin{equation}
t_{l^{\ast}}^{qr} = \sum_{i \in \mathcal{B}_{qr}} s(\hat{\mu}, \hat{\kappa}, y_i)
\end{equation}
for $q \in \{1,2\}$.
The breakpoint that leads to the highest discrepancy is then selected as split point as defined by
\begin{equation}
r^{\ast} = \argmin_{r} (\min_{q=1,2}(c(t_{l^{\ast}}^{qr},\mu_{l^{\ast}}^{qr},\Sigma_{l^{\ast}}^{qr})).
\end{equation}

Subsequently, the same testing and splitting procedure is repeated in each of the resulting 
subgroups until some stopping criterion is reached. Next to the already mentioned 
significance-level-based stopping criterion, i.e., a minimal $p$-value for the statistical tests,
also a maximal tree depth or a minimal number of observations in a node can be employed as stopping
criteria.

%This permutation-test-based tree algorithm is presented in 
%\cite{Hothorn+Hornik+Zeileis:2006} as the CTree algorithm. 
%A different framework to build a likelihood-based tree is provided by the MOB algorithm 
%which is based on M-fluctuation tests (\citealp{Zeileis+Hothorn+Hornik:2008}).

%\subsection{Teststatistic: CTree}\label{app:teststat_ctree}
%To measure the association of a response $Y$ and each possible split variable $Z_j$, $j=1,\ldots,J$,
%the CTree algorithm applies a linear test statistic $T_j$ which is excerpted from Section~3.1.
%(``Variable selection and stopping criteria'') of the original paper 
%\citep{Hothorn+Hornik+Zeileis:2006} and is of the following form:
%$$
%T_j(Y, Z_j, w) = \text{vec}\left( \sum_{i=1}^N w_i g_j(Z_{ji})h(Y_i)^{\top}\right)
%$$
%where $w$ are optional weights, $g_j: \mathcal{Z}_j \rightarrow \mathbb{R}^{P}$ is a nonrandom 
%transformation of the split variable $Z_j$ and the influence function 
%$h: \mathcal{Y} \times \mathcal{Y}^N \rightarrow \mathbb{R}^Q$ depends on the response 
%$Y = (Y_1, \ldots, Y_N)$ in a permutation symmetric way and is set to 
%$h(Y_i) = s(Y_i, X_i, \hat{\beta})$ for a score-based approach. 
%Moreover, by applying the ``vec'' operator the resulting
%$P~\times~Q$ matrix is converted into a $PQ$ column vector.
%Following \cite{Strasser+Weber:1999}
%the conditional expectation $\mu_j$ and covariance $\Sigma_j$ of a linear test statistic
%$T_j$ can be calculated and used to standardize an observed linear test statistic $t_j$ within a 
%function $c$ mapping into the real line.
%For example, the maximum of the absolute values of the standardized linear statistic
%$$
%c_{\max}(t_j,\mu_j,\Sigma_j) = \max \left| \frac{(t_j-\mu_j)}{\text{diag}(\Sigma_j)} \right|
%$$
%or a quadratic form
%$$
%c_{\text{quad}}(t_j, \mu_j, \Sigma_j) = (t_j - \mu_j)\Sigma_j^{+} (t_j - \mu_j)^{\top}
%$$
%can be considered where $\Sigma_j^{+}$ is the Moore-Penrose inverse of $\Sigma_j$.
%Since the asymptotic conditional distribution of a linear test statistic $T_j$ is a 
%multivariate normal with parameters $\mu_j$ and $\Sigma_j$ \citep{Strasser+Weber:1999}, 
%the asymptotic distribution of $c_{\text{max}}$ is normal while the quadratic form 
%$c_{\text{quad}}$ follows an asymptotic $\chi^2$~distribution. Based on this knowledge 
%the corresponding $p$-values can be calculated easily.

\end{appendix}

\end{document}