### 2020-03-22 pkg/circtree/inst/wind_paper/talks/ACINN_graduateseminar2020/slides/slides.Rnw
Add RNW for talk at ACINN
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@3000 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 51bc558f5ee3a861c5857f44883c50b46073638d

\documentclass[11pt,t,usepdftitle=false,aspectratio=169]{beamer}
\usetheme[url,license]{uibk}

\title[Circular Regression Trees and Forests with an Application to
Probabilistic Wind Direction Forecasting]{Circular Regression Trees and Forests 
with an\\\vspace{0.2em} Application to Probabilistic Wind Direction Forecasting}
\author{Moritz N. Lang, Lisa Schlosser, \\
Torsten~Hothorn, Georg~J.~Mayr, Reto Stauffer, Achim Zeileis}

%-------------------------------------------------------------------
% Preamble
%-------------------------------------------------------------------
\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
\headerimage{1}

\graphicspath{{images/}{./}}
\usepackage[utf8]{inputenc}

\setbeamertemplate{caption}{\insertcaption} 
%% includes a replacement for \usepackage{Sweave}
% \usepackage{Sweave}
\usepackage{changepage}
\usepackage{amsmath,tikz}
\usepackage{calc}
\usepackage{graphicx}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap,trees,tikzmark,decorations.pathreplacing}
\usepackage{xcolor}
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}

\newcommand{\backupbegin}{
   \newcounter{finalframe}
   \setcounter{finalframe}{\value{framenumber}}
}
\newcommand{\backupend}{
   \setcounter{framenumber}{\value{finalframe}}
}

%% colors
\definecolor{HighlightOrange}{rgb}{0.9490196,0.5725490,0.0000000}
\definecolor{HighlightBlue}{rgb}{0.4784314,0.7490196,0.9803922}
\definecolor{forestred}{RGB}{206,73,81}
\definecolor{treegreen}{RGB}{0,143,0}
\definecolor{lightblue}{RGB}{34,151,230}
\definecolor{lightorange}{RGB}{255,165,0}

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)

set.seed(7)

invisible(.Call(grDevices:::C_palette, grDevices::hcl(
  h = c(0,   5, 125, 245, 195, 315,  65,   0),
  c = c(0, 100,  90,  85,  63, 105,  90,   0),
  l = c(0,  55,  75,  60,  82,  48,  80,  65)
)))

# setwd("~/svn/partykit/pkg/disttree/inst/dissertation_overview/slides/")
library("disttree")
library("Formula")
library("latex2exp")
library("gamlss")
library("lattice")
library("crch")
library("RainTyrol")
library("parallel")
library("gamlss.cens")
gen.cens(NO, type = "left")
library("ggplot2")
theme_set(theme_bw(base_size = 18))

## HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)
names(pal) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")

pallight <- hcl(c(10, 128, 260, 290, 70), 100, 50, alpha = 0.25)
names(pallight) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")

transpgray <- rgb(0.190,0.190,0.190, alpha = 0.2)
@

%-------------------------------------------------------------------
% Main
%-------------------------------------------------------------------
\begin{document}
\section{Distributional Trees and Forests for Circular Data}

%-------------------------------------------------------------------
\subsection{Motivation}
%-------------------------------------------------------------------

%% SLIDE
\begin{frame}[fragile]
\frametitle{Motivation}

\pause

\textbf{Goal:} 
\begin{itemize}
  \item Statistical modeling of circular response variables.
  \item Probabilistic forecasting.
\end{itemize}

\medskip

\textbf{Examples:} Circular data.
\begin{itemize}
\item Hourly crime data: $[0\text{h}, 24\text{h})$ with $0\text{h} = 24\text{h}$.
\item Directional behavior of animals: $[0^{\circ}, 360^{\circ})$ with $0^{\circ} = 360^{\circ}$.
\item Wind directions: $[0^{\circ}, 360^{\circ})$ with $0^{\circ} = 360^{\circ}$.
\end{itemize}

\end{frame}

%% SLIDE
\begin{frame}[fragile]
\frametitle{Motivation}

\textbf{Application:} 
\begin{itemize}
  \item Probabilistic forecasting of wind directions for the next hours
  at\\Innsbruck and Vienna Airport, Austria.
\end{itemize}

\medskip

\textbf{Challenges:}
\begin{itemize}
  %\item Circular response in $[0^{\circ}, 360^{\circ})$ with $0^{\circ} = 360^{\circ}$.
  \item Possibly abrupt and short-term changes due to geographical position.
  \item Numerical weather predictions less useful due to short lead time.
  \item Large number of unknown interactions between covariates.
\end{itemize}

\bigskip
\pause
\begin{center}
$\boldsymbol{\Rightarrow}$ \textbf{What kind of statistical model should be used?}
\end{center}

\end{frame}

%% SLIDE 
\begin{frame}%[fragile]
\frametitle{Motivation}
\vspace{-0.41cm}
\begin{figure}[!htb]
\minipage{0.285\textwidth}
\begin{center}
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@
\visible<1->{
<<plot_motivation_GLM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
\end{center}
\endminipage
\visible<2->{{\LARGE$\rightarrow$}}
\minipage{0.285\textwidth}
\begin{center}
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3 
y <- ytrue + rnorm(nobs,0,0.3)
@
\visible<2->{
<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
\end{center}
\endminipage
\visible<3->{{\LARGE$\rightarrow$}}
\minipage{0.285\textwidth}
\begin{center}
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\visible<3->{
<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
}
\end{center}
\endminipage

\vspace{0.0cm}
\minipage{0.25\textwidth}
\vspace{-0.3cm}
\begin{center}
\visible<1->{
(Generalized) Linear model\\
\vspace{0.8cm}
{\small$\mathbb{E}(Y|X) =  \alpha_0 + \alpha_1 \cdot X$}
\vspace{0.3cm}
}
\end{center}
\endminipage
\hspace{1.1cm}
\minipage{0.25\textwidth}
\begin{center}
\vspace{-0.3cm}
\visible<2->{
Generalized additive model\\
\vspace{0.8cm}
{\small$\mathbb{E}(Y|X) = \alpha + f(X)$}\\
\vspace{0.3cm}
}
\end{center}
\endminipage
\hspace{1.1cm}
\minipage{0.27\textwidth}
\begin{center}
\visible<3->{
Generalized additive model for location, scale, and shape\\
\vspace{0.3cm}
{\small $\mathbb{E}(Y|X) = \alpha + f_1(X)$}\\\vspace{0.2em}
{\small $\sqrt{Var(Y|X)} = \beta + f_2(X)$}
}
\end{center}
\endminipage
\end{figure}
\end{frame}

%% SLIDE
\begin{frame}[fragile]
\frametitle{Motivation}
\vspace{-0.41cm}
\begin{figure}[!htb]
\minipage{0.285\textwidth}
\begin{center}
<<motivation_regtree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- ytree <- yforest <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))
y <- ytrue + rnorm(nobs,0,0.3)
for(i in 1:nobs) ytree[i] <- if(x[i]<1/3) 0.5 else {if(x[i]<2/3) 2 else 1}
@
\visible<1->{
<<plot_motivation_regtree, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
@
}
\end{center}

\endminipage
\visible<2->{{\LARGE$\rightarrow$}}
\minipage{0.285\textwidth}
 \begin{center}
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@
\visible<2->{
<<plot_motivation_randforest, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x, y = yforest, col = pal["forest"], lwd=7, main = "")
@
}
\end{center}

\endminipage
\visible<3->{{\LARGE$\rightarrow$}}
\minipage{0.285\textwidth}
\begin{center}
\visible<3->{
<<plot_motivation_question, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, type = "n")
box(lwd=5)
text(x = mean(range(x)), y = mean(range(y)), "?", cex = 12)
@
}
\end{center}

\endminipage

\minipage{0.285\textwidth}
\begin{center}
\vspace{0.0cm}
\visible<1->{
\vspace{-0.05cm}
Regression tree\\
\vspace{0.7cm}
\resizebox{0.2\textwidth}{!}{
\begin{tikzpicture}
  \node[ellipse, fill=HighlightBlue!70, align=center] (n0) at (1, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n1) at (0.5, 1) {};
  \draw[-, line width=1pt] (n0) -- (n1);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n2) at (1.5, 1) {};
  \draw[-, line width=1pt] (n0) -- (n2);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n3) at (1, 0) {};
  \draw[-, line width=1pt] (n2) -- (n3);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n4) at (2, 0) {};
  \draw[-, line width=1pt] (n2) -- (n4);
\end{tikzpicture}}
\vspace{0.2cm}\\
\vspace{1cm}}
\end{center}
\endminipage
\hspace{0.65cm}
\minipage{0.285\textwidth}
\begin{center}
\vspace{-1.0cm}
\visible<2->{
Random forest\\
\vspace{0.75cm}
\resizebox{0.6\textwidth}{!}{
\begin{tikzpicture}
  \node[ellipse, fill=HighlightBlue!70, align=center] (n00) at (1, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n01) at (0.5, 1) {};
  \draw[-, line width=1pt] (n00) -- (n01);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n02) at (1.5, 1) {};
  \draw[-, line width=1pt] (n00) -- (n02);

  \node[ellipse, fill=HighlightBlue!70, align=center] (n10) at (3, 2) {};
  \node[ellipse, fill=HighlightBlue!70, align=center] (n11) at (2.5, 1) {};
  \draw[-, line width=1pt] (n10) -- (n11);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n12) at (3.5, 1) {};
  \draw[-, line width=1pt] (n10) -- (n12);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n13) at (2, 0) {};
  \draw[-, line width=1pt] (n11) -- (n13);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n14) at (2.8, 0) {};
  \draw[-, line width=1pt] (n11) -- (n14);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n15) at (3.2, 0) {};
  \draw[-, line width=1pt] (n12) -- (n15);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n16) at (4, 0) {};
  \draw[-, line width=1pt] (n12) -- (n16);

  \node[ellipse, fill=HighlightBlue!70, align=center] (n20) at (5, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n21) at (4.5, 1) {};
  \draw[-, line width=1pt] (n20) -- (n21);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n22) at (5.5, 1) {};
  \draw[-, line width=1pt] (n20) -- (n22);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n23) at (5, 0) {};
  \draw[-, line width=1pt] (n22) -- (n23);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n24) at (6, 0) {};
  \draw[-, line width=1pt] (n22) -- (n24);
\end{tikzpicture}
}
\vspace{0.3cm}\\
}
\end{center}
\endminipage
\hspace{0.65cm}
\minipage{0.285\textwidth}
\begin{center}
\vspace{-0.4cm}
\visible<3->{
Distributional trees and forests\\
\vspace{1.15cm}
\vspace{1cm}}
\end{center}
\endminipage
\end{figure}
\end{frame}

%% SLIDE
\begin{frame}
\frametitle{Distributional trees}

\textbf{Distributional:}
\begin{itemize}
  \item Specify the complete probability distribution (location, scale, shape, \dots).
  \item Probabilistic / distributional forecasts.
\end{itemize}

\medskip

\textbf{Tree:}
\begin{itemize}
  \item Automatic detection of steps and abrupt changes. %(data driven)
  \item Capture non-linear and non-additive effects, and interactions.
\end{itemize}

\medskip

\textbf{Forest:}
\begin{itemize}
  \item Smoother effects.
  \item Stabilization and regularization of the model.
\end{itemize}

\medskip
\pause
\begin{center}
$\boldsymbol{\Rightarrow}$ \textbf{Ideal for capturing short-term and abrupt changes in wind direction.}
\end{center}

\end{frame}

%-------------------------------------------------------------------
\subsection{Distributional trees}
%-------------------------------------------------------------------

<<treedata, echo=FALSE, results=hide>>=
set.seed(54)
nobs <- 500
x <- runif(nobs, 0, 1)
mu <- sigma <- ytrue <- numeric(length = nobs)
for(i in 1:nobs) sigma[i] <- if(x[i]<=0.4) 1 else 3
for(i in 1:nobs) mu[i] <- if(x[i]<= 0.4|| x[i]>0.8) 4 else 12
y <- rnorm(nobs, mean = mu, sd = sigma)
ytrue <- mu
data <- data.frame(cbind(y,x, ytrue))

alldata <- cbind(data, mu, sigma)
odata <- alldata[order(alldata["x"]),]
@

<<dgp_tree, echo=FALSE, results=hide>>=
data <- data.frame(x = numeric(0), x = numeric(0), x = numeric(0))
names(data) <- c("x","x","x")
fig <- party(
  partynode(1L,
            split = partysplit(2L, breaks = 0.4),
            kids = list(
              partynode(2L, info = c(
                "n = 200",
                "   True parameters:   ",
                expression(mu == '4'),
                expression(sigma == '1')
              )),
              partynode(3L,
                        split = partysplit(3L, breaks = 0.8),
                        kids = list(
                          partynode(4L, info = c(
                            "n = 200",
                            "   True parameters:   ",
                            expression(mu == '12'),
                            expression(sigma == '3')
                          )),
                          partynode(5L, info = c(
                            "n = 100",
                            "   True parameters:   ",
                            expression(mu == '4'),
                            expression(sigma == '3')
                          )))))),
  data
)

node_inner_ext <- function (obj, id = TRUE, pval = TRUE, abbreviate = FALSE, fill = "white", 
    gp = gpar()) 
{
    meta <- obj$data
    nam <- names(obj)
    extract_label <- function(node) {
        if (is.terminal(node)) 
            return(rep.int("", 2L))
        varlab <- character_split(split_node(node), meta)$name
        if (abbreviate > 0L) 
            varlab <- abbreviate(varlab, as.integer(abbreviate))
        if (pval) {
            nullna <- function(x) is.null(x) || is.na(x)
            pval <- suppressWarnings(try(!nullna(info_node(node)$p.value), 
                silent = TRUE))
            pval <- if (inherits(pval, "try-error")) 
                FALSE
            else pval
        }
        if (pval) {
            pvalue <- node$info$p.value
            plab <- ifelse(pvalue < 10^(-3L), paste("p <", 10^(-3L)), 
                paste("p =", round(pvalue, digits = 3L)))
        }
        else {
            plab <- ""
        }
        return(c(varlab, plab))
    }
    maxstr <- function(node) {
        lab <- extract_label(node)
        klab <- if (is.terminal(node)) 
            ""
        else unlist(lapply(kids_node(node), maxstr))
        lab <- c(lab, klab)
        lab <- unlist(lapply(lab, function(x) strsplit(x, "\n")))
        lab <- lab[which.max(nchar(lab))]
        if (length(lab) < 1L) 
            lab <- ""
        return(lab)
    }
    nstr <- maxstr(node_party(obj))
    if (nchar(nstr) < 6) 
        nstr <- "aAAAAa"
    rval <- function(node) {
        node_vp <- viewport(x = unit(0.5, "npc"), y = unit(0.5, 
            "npc"), width = unit(1, "strwidth", nstr) * 1.3, 
            height = unit(3, "lines"), name = paste("node_inner", 
                id_node(node), sep = ""), gp = gp)
        pushViewport(node_vp)
        xell <- c(seq(0, 0.2, by = 0.01), seq(0.2, 0.8, by = 0.05), 
            seq(0.8, 1, by = 0.01))
        yell <- sqrt(xell * (1 - xell))
        xell <- xell*1.11 - 0.055             # to adapt size of the ellipse to the size with p-value
        lab <- extract_label(node)
        fill <- rep(fill, length.out = 2L)
        grid.polygon(x = unit(c(xell, rev(xell)), "npc"), y = unit(c(yell, 
            -yell) + 0.5, "npc"), gp = gpar(fill = fill[1]))
        grid.text(lab[1L], y = unit(1.5 + 0.5,                  # to adapt position of x to its position with p-value
            "lines"))
        #grid.text(lab[1L], y = unit(1.5 + 0.5 * (lab[2L] != ""), 
        #    "lines"))
        grid.text(lab[2L], y = unit(1, "lines"))
        if (id) {
            nodeIDvp <- viewport(x = unit(0.5, "npc"), y = unit(1, 
                "npc"), width = max(unit(1, "lines"), unit(1.3, 
                "strwidth", nam[id_node(node)])), height = max(unit(1, 
                "lines"), unit(1.3, "strheight", nam[id_node(node)])))
            pushViewport(nodeIDvp)
            grid.rect(gp = gpar(fill = fill[2]))
            grid.text(nam[id_node(node)])
            popViewport()
        }
        upViewport()
    }
    return(rval)
}

class(node_inner_ext) <- "grapcon_generator"

@

%% SLIDE
\begin{frame}[fragile]
\frametitle{Distributional trees}
\vspace*{-0.12cm}
\begin{center} 
DGP:  $\; Y\ |\ X = x \; \sim  \; \mathcal{N}(\mu(x), \sigma^2(x))$

\vspace*{-0.21cm}
\setkeys{Gin}{width=0.58\linewidth}
<<plottree_xyplot, fig=TRUE, echo=FALSE,width=7>>=
par(mar=c(5.1,4.1,2.4,1.1))
plot(y=odata$y, x=odata$x, ylab = "y", xlab = "x", col = "gray")
lines(x = odata$x, y = odata$mu, col = pal["forest"], lwd = 2.5, main = "")
polygon(c(odata$x, rev(odata$x)), c(odata$mu + odata$sigma, rev(odata$mu - odata$sigma)),
  col = pallight["forest"], border = "transparent")
legend("topleft", expression(mu  %+-%  sigma), bty = "n")
@
\end{center}
\end{frame}

%% SLIDE
\begin{frame}
\frametitle{Distributional trees}
\vspace*{-0.12cm}
\begin{center} 
DGP:  $\; Y\ |\ X = x \; \sim  \; \mathcal{N}(\mu(x), \sigma^2(x))$

\vspace*{-0.21cm}
\setkeys{Gin}{width=0.5\linewidth}
<<plottree_dgp, fig=TRUE, echo=FALSE>>=
paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
plot(fig, inner_panel = node_inner_ext,
     tp_args = list(FUN = identity, width = 18, fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]),
     drop_terminal = TRUE, tnex = 1.7)
@
\end{center}
\end{frame}

%% SLIDE
<<tree, echo=FALSE, results=hide>>=
set.seed(7)
nobs <- 500
x <- runif(nobs, 0, 1)
mu <- sigma <- ytrue <- numeric(length = nobs)
for(i in 1:nobs) sigma[i] <- if(x[i]<=0.4) 1 else 3
for(i in 1:nobs) mu[i] <- if(x[i]<= 0.4|| x[i]>0.8) 4 else 12
y <- rnorm(nobs, mean = mu, sd = sigma)
#y <- rcnorm(nobs, mean = mu, sd = sigma, left = 0)
ytrue <- mu
data <- data.frame(cbind(y,x, ytrue))
tree <- disttree(y ~ x, data = data, family = NO(), type.tree = "mob")
#tree <- disttree(y ~ x, data = data, family = dist_list_cens_normal)
@

\begin{frame}[fragile]
\frametitle{Distributional trees}
\begin{center}
\vspace*{-0.12cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.5\linewidth}
<<plottree_estpar, fig=TRUE, echo=FALSE>>=
# function for output in terminal panels
FUN <- function (x) 
{
  cf <- x$coefficients
  cf <- matrix(cf, ncol = 1, dimnames = list(names(cf), ""))
  c(sprintf("n = %s", x$nobs), "Estimated parameters:", parse(text = paste0("mu == '", format(round(cf[1], 2), nsmall = 2), "'")), 
                                                        parse(text = paste0("sigma == '", format(round(cf[2], 2), nsmall = 2), "'")))
}

paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))

## plot version using FUN and tree of class 'disttree'
plot(tree, drop = TRUE, tnex = 1.7, FUN = FUN,
     tp_args = list(fill = paltrees[c(1, 3)], width = 18), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}

%% SLIDE
\begin{frame}[fragile]
\frametitle{Distributional trees}
\begin{center}
\vspace*{-0.12cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.5\linewidth}
<<plottree_box, fig=TRUE, echo=FALSE>>=
plot(as.constparty(tree), tnex = 1.7, drop = TRUE,
     tp_args = list(fill = paltrees[c(1, 3)], ylines = 1.5), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}

%% SLIDE
\begin{frame}[fragile]
\frametitle{Distributional trees}
\begin{center}
\vspace*{-0.12cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.5\linewidth}
<<plottree_dens, fig=TRUE, echo=FALSE>>=
node_density <- function (tree, xscale = NULL, yscale = NULL, horizontal = FALSE,
                          main = "", xlab = "", ylab = "Density", id = TRUE, rug = TRUE,
                          fill = paltrees[c(1, 3)], col = "black", lwd = 0.5, ...) 
{
  yobs <- tree$data[,as.character(tree$info$formula[[2]])]
  ylines <- 1.5
  if (is.null(xscale)) xscale <- c(-5.1,22.5)
  if (is.null(yscale)) yscale <- c(-0.05,0.45)
  xr <- xscale
  yr <- yscale

  if (horizontal) {
    yyy <- xscale
    xscale <- yscale
    yscale <- yyy
  }

  rval <- function(node) {
    yrange <- seq(from = -20, to = 90)/4
    ydens <- node$info$object$ddist(yrange)

    top_vp <- viewport(layout = grid.layout(nrow = 2, ncol = 3, 
                                            widths = unit(c(ylines, 1, 1), c("lines", "null", "lines")), 
                                            heights = unit(c(1, 1), c("lines", "null"))), 
                       width = unit(1, "npc"), 
                       height = unit(1, "npc") - unit(2, "lines"), 
                       name = paste("node_density",node$id, sep = ""))
    pushViewport(top_vp)
    grid.rect(gp = gpar(fill = "white", col = 0))
    top <- viewport(layout.pos.col = 2, layout.pos.row = 1)
    pushViewport(top)
    mainlab <- paste(ifelse(id, paste("Node", node$id, "(n = "), "n = "), node$info$nobs, ifelse(id, ")", ""), sep = "")

    grid.text(mainlab)
    popViewport()
    plot <- viewport(layout.pos.col = 2, layout.pos.row = 2, 
                     xscale = xscale, yscale = yscale, 
                     name = paste("node_density",  node$id, "plot", sep = ""))
    pushViewport(plot)
    yd <- ydens
    xd <- yrange
    if (horizontal) {
      yyy <- xd
      xd <- yd
      yd <- yyy
      yyy <- xr
      xr <- yr
      yr <- yyy
      rxd <- rep(0, length(xd))
      ryd <- rev(yd)
    } else {
      rxd <- rev(xd)
      ryd <- rep(0, length(yd))
    }

    if (rug) {
      nodeobs <- node$info$object$y
      if (horizontal) {
        grid.rect(x = xscale[1], y = nodeobs , height = 0, width = xscale[1], 
                  default.units = "native", just = c("right", "bottom"),
		  gp = gpar(lwd = 2, col = gray(0, alpha = 0.18)))
      } else {
        grid.rect(x = nodeobs, y = yscale[1], 
                  width = 0, height = abs(yscale[1]), default.units = "native", 
                  just = c("center", "bottom"),
		  gp = gpar(lwd = 2, col = gray(0, alpha = 0.18)))
        #grid.lines(x = xr, y = yr, gp = gpar(col = "lightgray"), 
        #           default.units = "native")
        #grid.lines(x = xr, y = yr, gp = gpar(col = "lightgray"), 
        #           default.units = "native")
      }
    }

    grid.polygon(x = c(xd, rxd), y = c(yd, ryd), default.units = "native",
              gp = gpar(col = "black", fill = fill, lwd = lwd))
    #grid.lines(x = xd, y = yd, default.units = "native", 
    #           gp = gpar(col = col, lwd = lwd))
    grid.xaxis()
    grid.yaxis()
    grid.rect(gp = gpar(fill = "transparent"))
    upViewport(2)
  }
  return(rval)
}

class(node_density) <- "grapcon_generator"

plot(tree, tnex = 1.7, drop = TRUE,
     terminal_panel = node_density,
     tp_args = list(fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}

%% SLIDE
\begin{frame}[fragile]
\frametitle{Circular response distribution}
\pause

\vspace*{-0.3cm}

\begin{minipage}{0.5\textwidth}
\textbf{Distribution assumption:} Von Mises.
\begin{itemize}
\item Circular normal distribution.
\item Location parameter $\mu \in [0, 2 \pi)$.
\item Concentration parameter $\kappa > 0$.
\end{itemize}
\end{minipage}
\begin{minipage}{0.49\textwidth}
\only<2>{\includegraphics[width=\linewidth]{density_linear.pdf}}%
\only<3>{\includegraphics[width=\linewidth]{density_circular.pdf}}%
\end{minipage}

\bigskip
\medskip

\textbf{Log-likelihood:} $y \in [0, 2 \pi)$ and parameter vector $\theta = (\mu, \kappa)$.
$$
\ell(\theta; y) = \log \left\{ \frac{1}{2 \pi I_0(\kappa)}~e^{ \kappa \cos(y - \mu)} \right\}
$$
where $I_0(\kappa)$ is the modified Bessel function of the first kind and order $0$.

\end{frame}

%-------------------------------------------------------------------
\subsection{Learning}
%-------------------------------------------------------------------

%% SLIDE
\begin{frame}
\frametitle{Learning circular trees and forests}
      \begin{minipage}{0.7\textwidth}
      {\bf Tree:}
      \begin{enumerate}
      \item<3-> Fit global distributional model $\text{vM}(Y; \theta)$: \\ % to the whole data set:\\
      Estimate model parameters $\hat{\theta}$.
      \item<7-> Evaluate goodness of fit \\
      (for each parameter and each observation).
      \item<8-> Choose covariate $X$ with strongest influence on goodness of fit
        of $\text{vM}(Y; \hat{\theta})$ as split variable.
      \item<8-> Find the split point $p$ which leads to the highest improvement.
      \item<10-> Repeat steps 1--4 recursively in the subgroups
        until some stopping criterion is met.
      \end{enumerate}
      \end{minipage}
      \begin{minipage}{0.23\textwidth}
      %\vspace{-0.1cm}
      \begin{tikzpicture}
      \visible<2-3>{
      \node[ellipse, fill=HighlightBlue!70, align=center, scale = 0.7, minimum width=71pt, minimum height = 30pt] (n0) at (0.8, 1.7) {$Y$};
      }
      \visible<4>{
      \node[ellipse, fill=HighlightBlue!70, align=center, scale = 0.7, minimum width=71pt, minimum height = 30pt] (n0) at (0.8, 1.7) {$\text{vM}(Y;\hat{\theta}$)};
      }
      \visible<5->{
      \node[inner sep=0pt] (density_root) at (0.8, 1.7)
          {\includegraphics[width=0.6\textwidth]{density_root.pdf}};
      }
      \visible<6-8>{
      \node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=60pt, minimum height = 20pt] (n1) at (0, 0) {?};
      \node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=60pt, minimum height = 20pt] (n2) at (1.7, 0) {?};
      \draw[-, gray, line width=0.5pt] (0.7, 1.1) -- (n1);
      \draw[-, gray, line width=0.5pt] (1.0, 1.1) -- (n2);
      }
      \visible<9-10>{
      \node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=60pt, minimum height = 20pt] (n1) at (0, 0) {$Y_1$};
      \node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=60pt, minimum height = 20pt] (n2) at (1.7, 0) {$Y_2$};
      \draw[-, gray, line width=0.5pt] (0.7, 1.1) -- (n1) node [midway, left] {\scriptsize $X \leq p$};
      \draw[-, gray, line width=0.5pt] (1, 1.1) -- (n2) node [midway, right] {\scriptsize $X > p$};
      }
      \visible<11>{
      \node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=60pt, minimum height = 20pt] (n1) at (0, 0) {$\text{vM}(Y_1;\hat{\theta}_1$)};
      \node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=60pt, minimum height = 20pt] (n2) at (1.7, 0) {$\text{vM}(Y_2;\hat{\theta}_2$)};
      }
      \visible<11->{
      \draw[-, gray, line width=0.5pt] (0.7, 1.1) -- (n1) node [midway, left] {\scriptsize $X \leq p$};
      \draw[-, gray, line width=0.5pt] (1.0, 1.1) -- (n2) node [midway, right] {\scriptsize $X > p$};
      }
      \visible<12->{
      \node[inner sep=0pt] (density2) at (-0.1,-0.3)
          {\includegraphics[width=0.5\textwidth]{density_left.pdf}};
      }
      \visible<12->{
      \node[inner sep=0pt] (density3) at (1.7,-0.3)
          {\includegraphics[width=0.5\textwidth]{density_right.pdf}};
      }
      \end{tikzpicture}
      \end{minipage}
      \vspace{0.4cm}
\end{frame}

%% SLIDE
\begin{frame}
\frametitle{Learning circular trees and forests}
      \begin{minipage}{0.7\textwidth}
      {\bf Forest:} 
      \begin{enumerate}
      \item Learn an ensemble of $T$ trees:\\
       Each tree is built on a bootstrap or subsample of the data, with
      random input variable sampling.
      \item<2-> Combine / average over estimated $T$ trees:\\
      To stabilize and regularize the model.
      \end{enumerate}
      \end{minipage}

\end{frame}

%-------------------------------------------------------------------
\subsection{Wind forecasting}
%-------------------------------------------------------------------

%% SLIDE
\begin{frame}[fragile]
\frametitle{Wind forecasting}

\textbf{Here:} Nowcasting (1--3 hours ahead) of wind direction at Innsbruck Airport.

\bigskip
\pause

\textbf{Inputs:} Observation data only (41\,979 data points).
\begin{itemize}
  \item 4 stations at Innsbruck Airport, 6 nearby weather stations. 
  \item Base variables: Wind direction, wind (gust) speed, temperature, (reduced) air pressure, relative humidity.
  \item 260 covariates based on means/minima/maxima, temporal changes, spatial differences towards the airport.
\end{itemize}

\end{frame}

%% SLIDE
\begin{frame}
\frametitle{Wind forecasting}

\vspace*{-0.4cm}

\setkeys{Gin}{width=0.75\linewidth}
\only<1>{
  \vspace*{0.2cm}
  \begin{center}
  \includegraphics{_plot_circforest_stationmap_innsbruck.pdf}
  \end{center}
}
\only<2>{
  \vspace*{0.2cm}
  \begin{center}
  \includegraphics{_plot_circforest_stationmap_innsbruck_box.pdf}
  \end{center}
}
\only<3>{
  \framesubtitle{~}
  \vspace*{-0.55cm}
  \begin{center}
  \includegraphics{ibk_airport.png}
  \end{center}
}
\only<4>{
  \framesubtitle{Up-valley wind}
  \vspace*{-0.55cm}
  \begin{center}
  \includegraphics{ibk_airport_east.png}
  \end{center}
}
\only<5>{
  \framesubtitle{Down-valley wind}
  \vspace*{-0.55cm}
  \begin{center}
  \includegraphics{ibk_airport_west.png}
  \end{center}
}
\only<6>{
  \framesubtitle{Southerly wind}
  \vspace*{-0.55cm}
  \begin{center}
  \includegraphics{ibk_airport_south.png}
  \end{center}
}

\end{frame}

%% SLIDE
\begin{frame}[fragile]
\frametitle{Wind forecasting}

\vspace{-0.7cm}

\setkeys{Gin}{width=1\linewidth}
\begin{center}
\includegraphics{circtree_ibk.pdf}
\end{center}

\end{frame}

%% SLIDE
\begin{frame}[fragile]
\frametitle{Wind forecasting}

\textbf{Benchmark:} Against other naive and circular models.
\begin{itemize}
\item Climatology: Without covariates.
\item Persistency: Based on current wind direction.
\item Circular linear model: Based on current wind speed and wind vector $(u, v)^{\top}$.
\end{itemize}

\end{frame}

%% SLIDE
\begin{frame}[fragile]
\frametitle{Wind forecasting}

\textbf{Evaluation:} CRPS skill score for \only<1>{1}\only<2>{3}-hourly predictions (5-fold cross validation).
\begin{center}
\setkeys{Gin}{width=0.88\linewidth}
\only<1>{\includegraphics{circforest_crpsskill_ibk_lag1.pdf}}%
\only<2>{\includegraphics{circforest_crpsskill_ibk_lag3.pdf}}
\end{center}

\end{frame}

\begin{frame}[fragile]
\frametitle{Wind forecasting}

   \only<1-2>{
      \textbf{Summary}:
      \begin{itemize}
         \item Trees allow to capture abrupt changes, perform variable selection,
               and account for possible interactions.
         %\item Trees capture abrupt changes and possibly non-linear interactions.
         \item Forests additionally allow for modeling smooth changes.
         \item Circular trees are easy to interpret, in comparison to additive models.
          \pause
         \item Circular trees and forests outperform all other models.
         \item Even when persistence has no skill, trees and forests outperform climatology.
      \end{itemize}
   }

   \only<3->{
      \textbf{Outlook:} Wind shifts.
      \begin{itemize}
      \item Adapt already developed circular regression trees and forests to probabilistic nowcasting 
        of wind shifts.
      \item Build two step-model to predict (1) probability of occurrence of a wind shift
        and (2) possibly direction and magnitude.
      \end{itemize}
   }

\end{frame}

%-------------------------------------------------------------------
\subsection{Software}
%-------------------------------------------------------------------

%% SLIDE
\begin{frame}
\frametitle{Software}
%\vspace{0.4cm}
\textbf{Software:} \emph{circtree}
available on R-Forge at\\

\medskip

\small{\url{https://R-Forge.R-project.org/projects/partykit/}}\\

\bigskip
\medskip

\textbf{Main functions:}

\medskip

\begin{tabular}{ll}
\code{circfit}    & Circular distributional fits (ML).\\
& No covariates. \\
\code{circtree}   & Circular distributional trees (\code{ctree}/\code{mob} + \code{circfit}).\\
& Covariates as partitioning variables. \\
\code{circforest} & Circular distributional forests (ensemble of \code{circtree}s).\\
& Covariates as partitioning variables.
\end{tabular}

\bigskip
\medskip

\textbf{Correspondingly:} \emph{disttree} for any other response distribution (e.g., \code{gamlss.family}).

\end{frame}

%-------------------------------------------------------------------
\subsection{References}
%-------------------------------------------------------------------

%% SLIDE
\begin{frame}
\frametitle{References}

\vspace{-0.2cm}
\footnotesize

Lang MN, Schlosser L, Hothorn T, Mayr GJ, Stauffer R, Zeileis A (2020).
\dquote{Circular Regression Trees and Forests with an
Application to Probabilistic Wind Direction Forecasting.}
arXiv:12001.0041, {\it arXiv.org E-Print Archive}, in review, 2020.

\medskip

Schlosser L, Lang MN, Hothorn T, Mayr GJ, Stauffer R, Zeileis A (2019).
\dquote{Distributional Trees for Circular Data.}
\emph{Proceedings of the 34th International Workshop on Statistical Modelling}, 
\textbf{1}, 226--231.
\url{https://eeecon.uibk.ac.at/~zeileis/papers/Schlosser+Lang+Hothorn-2019.pdf}

\medskip

Schlosser L, Hothorn T, Stauffer R, Zeileis A (2019).
\dquote{Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain.}
\emph{The Annals of Applied Statistics}, \textbf{13}(3), 1564--1589.
\doi{10.1214/19-AOAS1247}

\bigskip
\medskip

Hothorn T, Hornik K, Zeileis A (2006).
 \dquote{Unbiased Recursive Partitioning: A Conditional Inference Framework.}
 \emph{Journal of Computational and Graphical Statistics},
 \textbf{15}(3), 651--674.
 \doi{10.1198/106186006X133933}

\medskip

Zeileis A, Hothorn T, Hornik K (2008).
 \dquote{Model-Based Recursive Partitioning.}
  \emph{Journal of Computational and Graphical Statistics},
  \textbf{17}(2), 492--514.
  \doi{10.1198/106186008X319331}

\end{frame}

%-------------------------------------------------------------------
\subsection{Appendix}
%-------------------------------------------------------------------
\backupbegin

%% SLIDE
\begin{frame}
~~~
\end{frame}

%% SLIDE
\begin{frame}
\frametitle{Wind forecasting: Vienna airport}

\vspace*{-0.2cm}

\setkeys{Gin}{width=0.75\linewidth}
\begin{center}
\includegraphics{_plot_circforest_stationmap_wien.pdf}
\end{center}
\end{frame}

%% SLIDE
\begin{frame}[fragile]
\frametitle{Wind forecasting: Vienna airport}

\vspace{-0.7cm}

\setkeys{Gin}{width=1\linewidth}
\begin{center}
\includegraphics{circtree_vie3h.pdf}
\end{center}

\end{frame}

%% SLIDE
\begin{frame}[fragile]
\frametitle{Wind forecasting: Vienna airport}

\textbf{Evaluation:} CRPS skill score for \only<1>{1}\only<2>{3}-hourly predictions (5-fold cross validation).
\begin{center}
\setkeys{Gin}{width=0.88\linewidth}
\only<1>{\includegraphics{circforest_crpsskill_vie_lag1.pdf}}%
\only<2>{\includegraphics{circforest_crpsskill_vie_lag3.pdf}}
\end{center}

\end{frame}
\backupend
\end{document}