### 2020-04-26 pkg/disttree/inst/PhD_thesis_defense/slides_169.Rnw
slides for PhD thesis defense
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@3011 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 2ac42af760b377bd2d56a4a05978e1eca5f1f0e9

\documentclass[11pt,t,usepdftitle=false,aspectratio=169]{beamer}
\usetheme[nototalframenumber, url,license]{uibk}

\title{Distributional Trees and Forests}
\author{\small{Lisa Schlosser\\
\vspace{0.4cm}
Supervisors: Achim Zeileis, Torsten~Hothorn, Michael Oberguggenberger\\}
\vspace{0.4cm}
\footnotesize{PhD thesis defense, 2020-04-24}
%Committee: Achim Zeileis, Michael Oberguggenberger, Georg Moser
}
%\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
%\URL{http://www.partykit.org/partykit/}
\renewcommand{\headerimage}[1]{%
   \IfStrEqCase{#1}{%
      {1}{%
         \gdef\myheaderimageid{#1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{forest.jpg}%
      }}[%
         \gdef\myheaderimageid{1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{forest.jpg}%
      ]%
}
\headerimage{1}

\graphicspath{{Figures/}}
\usepackage[utf8]{inputenc}
\usepackage[export]{adjustbox}

\setbeamertemplate{caption}{\insertcaption} 
%% includes a replacement for \usepackage{Sweave}
% \usepackage{Sweave}
\usepackage{changepage}
\usepackage{amsmath,tikz}
\usepackage{calc}
\usepackage{graphicx}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap,trees,tikzmark,decorations.pathreplacing}
\usepackage{xcolor}
%\usepackage[cal=boondoxo]{mathalfa}
%\graphicspath{{plots/}}
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}

%% colors
\definecolor{HighlightOrange}{rgb}{0.9490196,0.5725490,0.0000000}
\definecolor{HighlightBlue}{rgb}{0.4784314,0.7490196,0.9803922}
\definecolor{forestred}{rgb}{0.8078431, 0.2862745, 0.3176471}
\definecolor{treegreen}{rgb}{0.0000000, 0.5607843, 0.0000000}
\definecolor{lightblue}{RGB}{34,151,230}
\definecolor{lightorange}{RGB}{255,165,0}

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)

set.seed(7)

invisible(.Call(grDevices:::C_palette, grDevices::hcl(
  h = c(0,   5, 125, 245, 195, 315,  65,   0),
  c = c(0, 100,  90,  85,  63, 105,  90,   0),
  l = c(0,  55,  75,  60,  82,  48,  80,  65)
)))

# setwd("~/svn/partykit/pkg/disttree/inst/defensio/")
library("disttree")
library("Formula")
library("latex2exp")
library("gamlss")
library("lattice")
library("crch")
library("RainTyrol")
library("parallel")
library("gamlss.cens")
gen.cens(NO, type = "left")
library("ggplot2")
theme_set(theme_bw(base_size = 18))
library(colorspace)
## HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)
names(pal) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")

pallight <- hcl(c(10, 128, 260, 290, 70), 100, 50, alpha = 0.25)
names(pallight) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")

transpgray <- rgb(0.190,0.190,0.190, alpha = 0.2)

lightblue <- "#2297E6"
lightorange <- "#FFA500"
@

\begin{document}

\section{Distributional Trees and Forests}

\subsection{Statistical modeling}

\begin{frame}[fragile]
\frametitle{Statistical modeling}
\textbf{Assumption:}
\vspace{-0.543cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {nature};
\draw[->] (2.6,0.4)--(3.1,0.4);

\node[] at (3.5,0.4) {Y};

\end{tikzpicture}
\end{center}
\vspace{0.3cm}
\textbf{Goal:} Approximation of reality based on a set of observations.\\
\pause

\vspace{0.5cm}
\textbf{Motivation:}
\begin{itemize}
\item Prediction.
\item Capturing dependencies and interactions.
\end{itemize}
\pause

\vspace{0.5cm}
\textbf{Two cultures:} Breiman (2001, \emph{Statistical Science})
\begin{itemize}
\item Data modeling.
\item Algorithmic modeling.
\end{itemize}
\end{frame}

\subsubsection{Data modeling}

\begin{frame}[fragile]
\frametitle{Data modeling}
\vspace{-0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {f(X; par)};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\only<1->{
\begin{itemize}
\item Approximate data generating process up to a few unknown processes.
\item Mathematical equations, model assumptions.
\item Data: estimation of parameters.
\end{itemize}
}
\only<1>{\vspace{1cm}}

\vspace{0.1cm}

\visible<2->{
\begin{minipage}{0.65\textwidth}
\textbf{Example:} 
\only<2-3>{Linear model}\only<4>{Generalized additive model}\only<5>{Generalized additive model\\ for location, scale, and shape}
\\
\only<2-3>{
$$\visible<3>{\mathbb{E}(Y|X) = } f(X; \alpha) = \alpha_0 + \alpha_1 \cdot X$$  
\vspace{0.15cm}
}
\only<4>{
$$\mathbb{E}(Y|X) = f(X; \alpha) = \alpha + s(X)$$  
\vspace{0.15cm}
}
\only<5>{
\vspace{-1.29cm}

\begin{align*}
\visible<5>{\mathbb{E}(Y|X) = }\mu(X) = \alpha_{\mu} + s_{\mu}(X)\\  
\visible<5>{\sqrt{Var(Y|X)} = }\sigma(X) = \alpha_{\sigma} + s_{\sigma}(X)
\end{align*}
}
\end{minipage}
%\vspace{-0.6cm}
%\hfill
\begin{minipage}{0.3\textwidth}
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@
\only<2-3>{
<<plot_motivation_GLM_single, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(1.74*x)^2)/2.5
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\only<4>{
<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(1.74*x)^2)/2.5
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\only<5>{
<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
        col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
}
\end{minipage}
}
\end{frame}

<<motivation_tree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- numeric(length = length(x))
for(i in 1:(nobs)) ytrue[i] <- if(x[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))
y <- ytrue + rnorm(nobs,0,0.3)
# more points for precise illustration of true function
x100 <- c(1:(100*nobs))/(100*nobs)
ytrue <- ytree <- ytree2 <- ytree3 <- ytree4 <- yforest <- numeric(length = length(x100))
for(i in 1:(nobs*100)) ytrue[i] <- if(x100[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x100[i]-0.2)-1)))
for(i in 1:(nobs*100)) ytree[i] <- if(x100[i]<1/3) 0.5 else {if(x100[i]<2/3) 2 else 1}
for(i in 1:(nobs*100)) ytree2[i] <- if(x100[i]<0.31) 0.47 else {if(x100[i]<0.69) 1.9 else 1.1}
for(i in 1:(nobs*100)) ytree3[i] <- if(x100[i]<0.34) 0.52 else {if(x100[i]<0.54) 2.2 else {if(x100[i]<0.74) 1.74 else 0.9}}
for(i in 1:(nobs*100)) ytree4[i] <- if(x100[i]<0.32) 0.51 else {if(x100[i]<0.63) 1.8 else 1.15}
for(i in 1:(nobs*100)) yforest[i] <- if(x100[i]<=0.33) 0.52 else {if(x100[i]<=0.34) 1.5 else {if(x100[i]<=0.54) 2 else {if(x100[i]<=0.6) 1.9 else {if(x100[i]<=0.64) 1.7 else {if(x100[i]<=0.7) 1.4 else {if(x100[i]<=0.74) 1.2 else 1}}}}}}
@

\subsubsection{Algorithmic modeling}

\begin{frame}[fragile]
\frametitle{Algorithmic modeling}
\vspace{-0.1cm}
\begin{center}
\begin{tikzpicture}
\node[] at (0,0.4) {X};
\draw[->] (0.4,0.4)--(0.9,0.4);
\draw[] (1,0) rectangle (2.5,0.8);
\node[font=\scriptsize] at (1.75,0.4) {nature};
\draw[->] (2.6,0.4)--(3.1,0.4);
\node[] at (3.5,0.4) {Y};
\draw[<->] (0.1,0.1) .. controls (0.2, -0.4) and (0.4, -0.5) .. (0.8,-0.5);
\draw[<->] (2.7,-0.5) .. controls (3.1, -0.5) and (3.3, -0.4) .. (3.4,0.1);
\node[font=\scriptsize] at (1.75,-0.5) {algorithm};
\end{tikzpicture}
\end{center}
\vspace{-0.4cm}
\visible<2->{
\begin{itemize}
\item Flexible approximation with no/few assumptions. %\\
%$\Rightarrow$ no fixed $f$.
\item Data-driven model fitting.
\end{itemize}
}
\vspace{0.25cm}
\visible<3->{
\begin{minipage}{0.5425\textwidth}
\textbf{Example:} \only<3>{Tree model}\only<4-10>{Subsampling}\only<11->{Forest model}
%\vspace{-0.5cm}
%\vspace{-0.8cm}
\begin{center}
\only<3>{
%\vspace{0.2cm}
\resizebox{0.45\textwidth}{!}{
\begin{tikzpicture}
  \node[ellipse, fill=HighlightBlue!70, align=center] (n0) at (2, 3) {};
  \node[] (n00) at (1, 1.5) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n1) at (0, 0) {$\hat{Y}_1$};
  \draw[-] (n0) -- (n00) node [midway, left] {\small $X \leq p_1$};
  \draw[-] (n0) -- (n1);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n2) at (3, 1.5) {};
  \draw[-] (n0) -- (n2) node [midway, right] {\small $X > p_1$};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n3) at (2, 0) {$\hat{Y}_2$};
  \draw[-] (n2) -- (n3) node [midway, left] {\small $X \leq p_2$};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n4) at (4, 0) {$\hat{Y}_3$};
  \draw[-] (n2) -- (n4) node [midway, right] {\small $X > p_2$};
\end{tikzpicture}}
%\vspace{0.445cm}
}}
\only<4->{
%\vspace{0.2cm}
\resizebox{0.347\textwidth}{!}{
\begin{tikzpicture}
%\draw[ellipse, draw=black, align=center, scale = 0.7, minimum width=170pt, minimum height = 35pt] (set) at (1, 3) {};
\draw[gray,rounded corners=10pt] (-1.05,4.35) -- (4.05,4.35) -- (4.05,2.7) -- (-2.05,2.7) -- (-2.05,4.35) -- (0.05,4.35);
%\draw[gray] (1,3.5) ellipse (3 and 0.8);
%\visible<4>{
\begin{scope}
\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\clip (1,3.5) ellipse (3 and 0.8);
\pgfmathsetseed{7}
\foreach \p in {1,...,200} {\fill[gray] (1+3*rand,3.5+0.8*rand) circle (0.04);}
\end{scope}
%}
%%% color points
%\visible<5-6>{
%\begin{scope}
%\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\pgfmathsetseed{7}
%\foreach \p in {1,...,100} {\fill[treegreen] (1+3*rand,3.5+0.8*rand) circle (0.07);}
%\end{scope}
%}
\visible<5->{
\node[ellipse, draw=treegreen, align=center, minimum width=80pt, minimum height = 40pt, line width = 3pt] (subset1) at (0.1, 3.5) {};
}
%\visible<7-8>{
%\begin{scope}
%\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\pgfmathsetseed{7}
%\foreach \p in {1,...,49} {\fill[gray] (1+3*rand,3.5+0.8*rand) circle (0.04);}
%\foreach \p in {50,...,150} {\fill[lightblue] (1+3*rand,3.5+0.8*rand) circle (0.07);}
%\end{scope}
%}
\visible<7->{
\node[ellipse, draw = lightblue, align=center, minimum width=100pt, minimum height = 30pt, line width = 3pt] (subset2) at (1.2, 3.6) {};
}
%\visible<9-10>{
%\begin{scope}
%\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\pgfmathsetseed{7}
%\foreach \p in {1,...,99} {\fill[gray] (1+3*rand,3.5+0.8*rand) circle (0.04);}
%\foreach \p in {100,...,200} {\fill[lightorange] (1+3*rand,3.5+0.8*rand) circle (0.07);}
%\end{scope}
%}
\visible<9->{
\node[ellipse, draw = lightorange, align=center, minimum width=120pt, minimum height = 25pt, line width = 3pt] (subset3) at (1.6, 3.3) {};
}
\visible<6->{
\node[ellipse, fill=treegreen, align=center] (n00) at (-1, 2) {};
\node[ellipse, fill=treegreen, align=center] (n01) at (-1.25, 1.25) {};
\draw[-, line width=1pt] (n00) -- (n01);
\node[rectangle, fill=treegreen, align=center] (n02) at (-1.5, 0.5) {};
\draw[-, line width=1pt] (n01) -- (n02);
\node[rectangle, fill=treegreen, align=center] (n03) at (-1, 0.5) {};
\draw[-, line width=1pt] (n01) -- (n03);
\node[rectangle, fill=treegreen, align=center] (n04) at (-0.5, 0.5) {};
\draw[-, line width=1pt] (n00) -- (n04);
}
\visible<8->{
\node[ellipse, fill=lightblue, align=center] (n10) at (1, 2) {};
\node[ellipse, fill=lightblue, align=center] (n11) at (0.5, 1.25) {};
\draw[-, line width=1pt] (n10) -- (n11);
\node[rectangle, fill=lightblue, align=center] (n12) at (0.25, 0.5) {};
\draw[-, line width=1pt] (n11) -- (n12);
\node[rectangle, fill=lightblue, align=center] (n13) at (0.75, 0.5) {};
\draw[-, line width=1pt] (n11) -- (n13);
\node[ellipse, fill=lightblue, align=center] (n14) at (1.5, 1.25) {};
\draw[-, line width=1pt] (n10) -- (n14);
\node[rectangle, fill=lightblue, align=center] (n15) at (1.75, 0.5) {};
\draw[-, line width=1pt] (n14) -- (n15);
\node[rectangle, fill=lightblue, align=center] (n16) at (1.25, 0.5) {};
\draw[-, line width=1pt] (n14) -- (n16);

}
\visible<10->{
\node[ellipse, fill=lightorange, align=center] (n20) at (3, 2) {};
\node[rectangle, fill=lightorange, align=center] (n21) at (2.5, 0.5) {};
\draw[-, line width=1pt] (n20) -- (n21);
\node[ellipse, fill=lightorange, align=center] (n22) at (3.25, 1.25) {};
\draw[-, line width=1pt] (n20) -- (n22);
\node[rectangle, fill=lightorange, align=center] (n23) at (3, 0.5) {};
\draw[-, line width=1pt] (n22) -- (n23);
\node[rectangle, fill=lightorange, align=center] (n24) at (3.5, 0.5) {};
\draw[-, line width=1pt] (n22) -- (n24);
}
\visible<11->{
\draw[-, line width=1pt] (-0.5, -0.1) -- (0.2,-0.5);
\draw[-, line width=1pt] (1, -0.1) -- (1,-0.5);
\draw[-, line width=1pt] (2.5, -0.1) -- (1.8,-0.5);
\node[rectangle, align=center] (combine) at (1, -0.9) {\Large average};
}
\end{tikzpicture}
}
}
\end{center}
\end{minipage}
\hspace{0.7cm}
\begin{minipage}{0.3\textwidth}
%\begin{center}
%\vspace{0.1cm}
\only<3-5>{
%\vspace{0.01cm}
<<plot_motivation_tree, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(2,0,2,0))
#par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x100, y = ytree, col = pal["forest"], lwd=7)
@
% mtext("rain", side = 4, las = TRUE, at = 16, line = 0.5, cex = 2.5, col = pal["forest"])
% mtext("now", side = 4, las = TRUE, at = 14.7, line = 0.5, cex = 2.5, col = pal["forest"])
% mtext("no", side = 4, las = TRUE, at = 10.3, line = 0.5, cex = 2.5, col = pal["forest"])
% mtext("rain", side = 4, las = TRUE, at = 9, line = 0.5, cex = 2.5, col = pal["forest"])
% mtext("now", side = 4, las = TRUE, at = 7.7, line = 0.5, cex = 2.5, col = pal["forest"])
% mtext("rain 24h", side = 2, las = FALSE, line = 1, cex = 3)
% mtext("humidity", side = 1, line = 1.5, cex = 3)
}
\only<6-7>{
%\vspace{0.01cm}
%\hspace{-0.1cm}
<<plot_motivation_randforest1, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(2,0,2,0))
#par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x100, y = ytree, col = pal["forest"], lwd=7)
lines(x = x100, y = ytree2, col = pal["tree"], lwd=7)
@
}
\only<8-9>{
%\vspace{0.01cm}
%\hspace{-0.1cm}
<<plot_motivation_randforest2, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(2,0,2,0))
#par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x100, y = ytree, col = pal["forest"], lwd=7)
lines(x = x100, y = ytree2, col = pal["tree"], lwd=7)
lines(x = x100, y = ytree3, col = lightblue, lwd=7)
@
}
\only<10>{
%\vspace{0.01cm}
%\hspace{-0.1cm}
<<plot_motivation_randforest3, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(2,0,2,0))
#par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x100, y = ytree, col = pal["forest"], lwd=7)
lines(x = x100, y = ytree2, col = pal["tree"], lwd=7)
lines(x = x100, y = ytree3, col = lightblue, lwd=7)
lines(x = x100, y = ytree4, col = lightorange, lwd=7)
@
}
\only<11->{
%\vspace{0.01cm}
%\hspace{-0.1cm}
<<plot_motivation_randforest4, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x100, y = yforest, col = pal["forest"], lwd=7)
@
}
%\end{center}
\end{minipage}
\end{frame}

% \subsection{Motivation}
% 
% \begin{frame}%[fragile]
% \frametitle{Motivation}
% \vspace{-0.41cm}
% \begin{figure}[!htb]
% \minipage{0.285\textwidth}
% \begin{center}
% <<motivation_GLM, echo=FALSE, results=hide>>=
% nobs <- 200
% ## GLM
% set.seed(7)
% x <- c(1:nobs)/nobs
% ytrue <- 1+x*1.5
% y <- ytrue + rnorm(nobs,0,0.3)
% @
% \visible<2->{
% <<plot_motivation_GLM, fig=TRUE, echo=FALSE>>=
% par(mar=c(2,0,2,0))
% plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
% box(lwd=5)
% lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
% @
% }
% \end{center}
% \endminipage
% \visible<3->{{\LARGE$\rightarrow$}}
% \minipage{0.285\textwidth}
% \begin{center}
% <<motivation_GAM, echo=FALSE, results=hide>>=
% ## GAM
% set.seed(7)
% x <- c(1:nobs)/nobs
% x <- 2*(x-0.5)
% ytrue <- x^3 
% y <- ytrue + rnorm(nobs,0,0.3)
% @
% \visible<3->{
% <<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
% par(mar=c(2,0,2,0))
% plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
% box(lwd=5)
% lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
% @
% }
% \end{center}
% \endminipage
% \visible<4->{{\LARGE$\rightarrow$}}
% \minipage{0.285\textwidth}
% \begin{center}
% <<motivation_GAMLSS, echo=FALSE, results=hide>>=
% ## GAMLSS
% set.seed(7)
% x <- c(1:nobs)/nobs
% x <- 2*(x-0.5)
% ytrue <- x^3
% var <- exp(-(2*x)^2)/2
% y <- ytrue + rnorm(nobs, 0, 0.1 + var)
% @
% \visible<4->{
% <<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
% par(mar=c(2,0,2,0))
% plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
% polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
%   col = pallight["forest"], border = "transparent")
% lines(x, ytrue, col = pal["forest"], lwd=7)
% points(x, y, col = "slategray", pch = 19)
% box(lwd = 5)
% @
% }
% \end{center}
% \endminipage
% 
% \vspace{0.5cm}
% \minipage{0.25\textwidth}
% \begin{center}
% \visible<2->{
% LM, GLM\\
% \vspace{0.5cm}
% \code{lm}\\
% \code{glm}\\
% \vspace{1.5cm}}
% \end{center}
% \endminipage
% \hspace{1.1cm}
% \minipage{0.25\textwidth}
% \begin{center}
% \visible<3->{
% GAM\\
% \vspace{0.5cm}
% \code{mgcv}\\
% \code{VGAM}\\
% \vspace{1.5cm}}
% \end{center}
% \endminipage
% \hspace{1.1cm}
% \minipage{0.25\textwidth}
% \begin{center}
% \visible<4->{
% GAMLSS\\
% \vspace{0.5cm}
% \code{gamlss}\\
% \code{mgcv}\\
% \code{VGAM}\\
% \code{gamboostLSS}\\
% \code{bamlss}}
% \end{center}
% \endminipage
% \end{figure}
% \end{frame}
% 
% 
% 
% 
% \begin{frame}[fragile]
% \frametitle{Motivation}
% \vspace{-0.41cm}
% \begin{figure}[!htb]
% \minipage{0.285\textwidth}
% \begin{center}
% <<motivation_regtree, echo=FALSE, results=hide>>=
% ## Reg. Tree
% set.seed(7)
% kappa <- 12
% x <- c(1:nobs)/nobs
% ytrue <- ytree <- yforest <- numeric(length = length(x))
% for(i in 1:nobs) ytrue[i] <- if(x[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))
% y <- ytrue + rnorm(nobs,0,0.3)
% for(i in 1:nobs) ytree[i] <- if(x[i]<1/3) 0.5 else {if(x[i]<2/3) 2 else 1}
% @
% \visible<1->{
% <<plot_motivation_regtree, fig=TRUE, echo=FALSE>>=
% par(mar=c(2,0,2,0))
% plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
% box(lwd=5)
% #lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
% lines(x = x, y = ytree, col = pal["forest"], lwd=7)
% @
% }
% \end{center}
% 
% \endminipage
% \visible<2->{{\LARGE$\rightarrow$}}
% \minipage{0.285\textwidth}
%  \begin{center}
% <<motivation_randforest, echo=FALSE, results=hide>>=
% ## Random Forest
% for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
% @
% \visible<2->{
% <<plot_motivation_randforest, fig=TRUE, echo=FALSE>>=
% par(mar=c(2,0,2,0))
% plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
% box(lwd=5)
% #lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
% lines(x = x, y = yforest, col = pal["forest"], lwd=7, main = "")
% @
% }
% \end{center}
% 
% \endminipage
% \visible<3->{{\LARGE$\rightarrow$}}
% \minipage{0.285\textwidth}
% \begin{center}
% \visible<3->{
% <<plot_motivation_question, fig=TRUE, echo=FALSE>>=
% par(mar=c(2,0,2,0))
% plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, type = "n")
% box(lwd=5)
% text(x = mean(range(x)), y = mean(range(y)), "?", cex = 12)
% @
% }
% \end{center}
% 
% \endminipage
% 
% \minipage{0.285\textwidth}
% \begin{center}
% \vspace{0.0cm}
% \visible<1->{
% Regression tree\\
% \vspace{0.3cm}
% \resizebox{0.2\textwidth}{!}{
% \begin{tikzpicture}
%   \node[ellipse, fill=HighlightBlue!70, align=center] (n0) at (1, 2) {};
%   \node[rectangle, fill=HighlightOrange!70, align=center] (n1) at (0.5, 1) {};
%   \draw[-, line width=1pt] (n0) -- (n1);
%   \node[ellipse, fill=HighlightBlue!70, align=center] (n2) at (1.5, 1) {};
%   \draw[-, line width=1pt] (n0) -- (n2);
%   \node[rectangle, fill=HighlightOrange!70, align=center] (n3) at (1, 0) {};
%   \draw[-, line width=1pt] (n2) -- (n3);
%   \node[rectangle, fill=HighlightOrange!70, align=center] (n4) at (2, 0) {};
%   \draw[-, line width=1pt] (n2) -- (n4);
% \end{tikzpicture}}
% \vspace{0.2cm}\\
% \code{rpart}\\
% \code{party(kit)}\\
% \vspace{1cm}}
% \end{center}
% \endminipage
% \hspace{0.65cm}
% \minipage{0.285\textwidth}
% \begin{center}
% \vspace{-0.4cm}
% \visible<2->{
% Random forest\\
% \vspace{0.4cm}
% \resizebox{0.6\textwidth}{!}{
% \begin{tikzpicture}
%   \node[ellipse, fill=HighlightBlue!70, align=center] (n00) at (1, 2) {};
%   \node[rectangle, fill=HighlightOrange!70, align=center] (n01) at (0.5, 1) {};
%   \draw[-, line width=1pt] (n00) -- (n01);
%   \node[rectangle, fill=HighlightOrange!70, align=center] (n02) at (1.5, 1) {};
%   \draw[-, line width=1pt] (n00) -- (n02);
%   
%   \node[ellipse, fill=HighlightBlue!70, align=center] (n10) at (3, 2) {};
%   \node[ellipse, fill=HighlightBlue!70, align=center] (n11) at (2.5, 1) {};
%   \draw[-, line width=1pt] (n10) -- (n11);
%   \node[ellipse, fill=HighlightBlue!70, align=center] (n12) at (3.5, 1) {};
%   \draw[-, line width=1pt] (n10) -- (n12);
%   \node[rectangle, fill=HighlightOrange!70, align=center] (n13) at (2, 0) {};
%   \draw[-, line width=1pt] (n11) -- (n13);
%   \node[rectangle, fill=HighlightOrange!70, align=center] (n14) at (2.8, 0) {};
%   \draw[-, line width=1pt] (n11) -- (n14);
%   \node[rectangle, fill=HighlightOrange!70, align=center] (n15) at (3.2, 0) {};
%   \draw[-, line width=1pt] (n12) -- (n15);
%   \node[rectangle, fill=HighlightOrange!70, align=center] (n16) at (4, 0) {};
%   \draw[-, line width=1pt] (n12) -- (n16);
%   
%   \node[ellipse, fill=HighlightBlue!70, align=center] (n20) at (5, 2) {};
%   \node[rectangle, fill=HighlightOrange!70, align=center] (n21) at (4.5, 1) {};
%   \draw[-, line width=1pt] (n20) -- (n21);
%   \node[ellipse, fill=HighlightBlue!70, align=center] (n22) at (5.5, 1) {};
%   \draw[-, line width=1pt] (n20) -- (n22);
%   \node[rectangle, fill=HighlightOrange!70, align=center] (n23) at (5, 0) {};
%   \draw[-, line width=1pt] (n22) -- (n23);
%   \node[rectangle, fill=HighlightOrange!70, align=center] (n24) at (6, 0) {};
%   \draw[-, line width=1pt] (n22) -- (n24);
% \end{tikzpicture}
% }
% \vspace{0.3cm}\\
% \code{randomForest}\\
% \code{ranger}\\
% \code{party(kit)}}
% \end{center}
% \endminipage
% \hspace{0.65cm}
% \minipage{0.285\textwidth}
% \begin{center}
% \vspace{0.15cm}
% \visible<3->{
% Distributional trees and forests\\
% \vspace{1.15cm}
% \code{disttree}\\
% based on \code{partykit}\\
% \vspace{1cm}}
% \end{center}
% \endminipage
% \end{figure}
% \end{frame}

\begin{frame}
\frametitle{Building blocks}
%\vspace{0.5cm}

\begin{minipage}{0.79\textwidth}
\vspace{-0.5cm}
\textbf{Distributional modeling:}
\begin{itemize}
\vspace{-0.05cm}
  \item Specify the complete probability distribution \\
  \vspace{-0.05cm}
  (location, scale, shape, \dots).
\end{itemize}

\smallskip

\textbf{Tree:}
\begin{itemize}
\vspace{-0.05cm}
  \item Capture non-linear and non-additive effects.
  \vspace{-0.1cm}
  \item Automatic selection of covariates and interactions. %detection of steps and abrupt changes. %(data driven)
\end{itemize}

\smallskip

\textbf{Forest:}
\begin{itemize}
\vspace{-0.05cm}
  \item Smoother effects.
  \vspace{-0.1cm}
  \item Stabilization and regularization of the model.
\end{itemize}

\bigskip

\bigskip

$\Rightarrow$ \textbf{Synthesis:} Distributional Trees and Forests
\end{minipage}
<<motivation_GAMLSS_syn, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
nobs <- 200
x_g <- c(1:nobs)/nobs
x_g <- 2*(x_g-0.5)
ytrue_g <- x_g^3
var_g <- exp(-(1.74*x_g)^2)/2.5
y_g <- ytrue_g + rnorm(nobs, 0, 0.1 + var_g)
@
<<motivation_regtree_syn, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- ytree <- yforest <- var <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))
for(i in 1:nobs) var[i] <-  if(x[i]<1/3) 0.1 else 0.6 *(4/3 - x[i])
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
for(i in 1:nobs) ytree[i] <- if(x[i]<1/3) 0.5 else {if(x[i]<2/3) 2 else 1}
@
<<motivation_randforest_syn, echo=FALSE, results=hide>>=
## Random Forest
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@
\begin{minipage}{0.2\textwidth}
\vspace{-0.05cm}
<<plot_motivation_GAMLSS_syn, fig=TRUE, echo=FALSE, width=8.5>>=
par(mar=c(1,0,1,0))
plot(x_g, y_g, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x_g, rev(x_g)), c(ytrue_g + 0.1 + var_g, rev(ytrue_g - 0.1 - var_g)),
  col = pallight["forest"], border = "transparent")
lines(x_g, ytrue_g, col = pal["forest"], lwd=10)
points(x_g, y_g, col = "slategray", pch = 19)
box(lwd = 5)
@

\vspace{0.15cm}

<<plot_motivation_regtree_syn, fig=TRUE, echo=FALSE, width=8.5>>=
par(mar=c(1,0,1,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(y), max(y)))
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x, y = ytree, col = pal["forest"], lwd=10)
box(lwd=5)
@

\vspace{0.15cm}

<<plot_motivation_randforest4_syn, fig=TRUE, echo=FALSE, width=8.5>>=
par(mar=c(1,0,1,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(y), max(y)))
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x, y = yforest, col = pal["forest"], lwd=10, main = "")
box(lwd=5)
@

\vspace{0.15cm}

<<plot_motivation_distforest, fig=TRUE, echo=FALSE, width=8.5>>=
par(mar=c(1,0,1,0))
#plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
#box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
#lines(x = x, y = yforest, col = pal["forest"], lwd=7, main = "")
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(y), max(y)))
polygon(c(x, rev(x)), c(yforest + 0.1 + var, rev(yforest - (0.1 + var))),
       # c(ytrue_g + 0.1 + var_g, rev(ytrue_g - 0.1 - var_g))
  col = pallight["forest"], border = "transparent")
box(lwd=5)
lines(x = x, y = yforest, col = pal["forest"], lwd=10)
@
\end{minipage}
\end{frame}

\subsection{Distributional trees}

<<treedata, echo=FALSE, results=hide>>=
set.seed(54)
nobs <- 500
x <- runif(nobs, 0, 1)
mu <- sigma <- ytrue <- numeric(length = nobs)
for(i in 1:nobs) sigma[i] <- if(x[i]<=0.4) 1 else 3
for(i in 1:nobs) mu[i] <- if(x[i]<= 0.4|| x[i]>0.8) 4 else 12
y <- rnorm(nobs, mean = mu, sd = sigma)
ytrue <- mu
data <- data.frame(cbind(y,x, ytrue))

alldata <- cbind(data, mu, sigma)
odata <- alldata[order(alldata["x"]),]
@

<<dgp_tree, echo=FALSE, results=hide>>=
data <- data.frame(x = numeric(0), x = numeric(0), x = numeric(0))
names(data) <- c("x","x","x")
fig <- party(
  partynode(1L,
            split = partysplit(2L, breaks = 0.4),
            kids = list(
              partynode(2L, info = c(
                "n = 200",
                "   True parameters:   ",
                expression(mu == '4'),
                expression(sigma == '1')
              )),
              partynode(3L,
                        split = partysplit(3L, breaks = 0.8),
                        kids = list(
                          partynode(4L, info = c(
                            "n = 200",
                            "   True parameters:   ",
                            expression(mu == '12'),
                            expression(sigma == '3')
                          )),
                          partynode(5L, info = c(
                            "n = 100",
                            "   True parameters:   ",
                            expression(mu == '4'),
                            expression(sigma == '3')
                          )))))),
  data
)

node_inner_ext <- function (obj, id = TRUE, pval = TRUE, abbreviate = FALSE, fill = "white", 
    gp = gpar()) 
{
    meta <- obj$data
    nam <- names(obj)
    extract_label <- function(node) {
        if (is.terminal(node)) 
            return(rep.int("", 2L))
        varlab <- character_split(split_node(node), meta)$name
        if (abbreviate > 0L) 
            varlab <- abbreviate(varlab, as.integer(abbreviate))
        if (pval) {
            nullna <- function(x) is.null(x) || is.na(x)
            pval <- suppressWarnings(try(!nullna(info_node(node)$p.value), 
                silent = TRUE))
            pval <- if (inherits(pval, "try-error")) 
                FALSE
            else pval
        }
        if (pval) {
            pvalue <- node$info$p.value
            plab <- ifelse(pvalue < 10^(-3L), paste("p <", 10^(-3L)), 
                paste("p =", round(pvalue, digits = 3L)))
        }
        else {
            plab <- ""
        }
        return(c(varlab, plab))
    }
    maxstr <- function(node) {
        lab <- extract_label(node)
        klab <- if (is.terminal(node)) 
            ""
        else unlist(lapply(kids_node(node), maxstr))
        lab <- c(lab, klab)
        lab <- unlist(lapply(lab, function(x) strsplit(x, "\n")))
        lab <- lab[which.max(nchar(lab))]
        if (length(lab) < 1L) 
            lab <- ""
        return(lab)
    }
    nstr <- maxstr(node_party(obj))
    if (nchar(nstr) < 6) 
        nstr <- "aAAAAa"
    rval <- function(node) {
        node_vp <- viewport(x = unit(0.5, "npc"), y = unit(0.5, 
            "npc"), width = unit(1, "strwidth", nstr) * 1.3, 
            height = unit(3, "lines"), name = paste("node_inner", 
                id_node(node), sep = ""), gp = gp)
        pushViewport(node_vp)
        xell <- c(seq(0, 0.2, by = 0.01), seq(0.2, 0.8, by = 0.05), 
            seq(0.8, 1, by = 0.01))
        yell <- sqrt(xell * (1 - xell))
        xell <- xell*1.11 - 0.055             # to adapt size of the ellipse to the size with p-value
        lab <- extract_label(node)
        fill <- rep(fill, length.out = 2L)
        grid.polygon(x = unit(c(xell, rev(xell)), "npc"), y = unit(c(yell, 
            -yell) + 0.5, "npc"), gp = gpar(fill = fill[1]))
        grid.text(lab[1L], y = unit(1.5 + 0.5,                  # to adapt position of x to its position with p-value
            "lines"))
        #grid.text(lab[1L], y = unit(1.5 + 0.5 * (lab[2L] != ""), 
        #    "lines"))
        grid.text(lab[2L], y = unit(1, "lines"))
        if (id) {
            nodeIDvp <- viewport(x = unit(0.5, "npc"), y = unit(1, 
                "npc"), width = max(unit(1, "lines"), unit(1.3, 
                "strwidth", nam[id_node(node)])), height = max(unit(1, 
                "lines"), unit(1.3, "strheight", nam[id_node(node)])))
            pushViewport(nodeIDvp)
            grid.rect(gp = gpar(fill = fill[2]))
            grid.text(nam[id_node(node)])
            popViewport()
        }
        upViewport()
    }
    return(rval)
}

class(node_inner_ext) <- "grapcon_generator"

@

\begin{frame}[fragile]
\frametitle{Distributional trees}
\vspace*{-0.12cm}
\begin{center} 
DGP:  $\; Y\ |\ X = x \; \sim  \; \mathcal{N}(\mu(x), \sigma^2(x))$

\vspace*{-0.21cm}
\setkeys{Gin}{width=0.58\linewidth}
<<plottree_xyplot, fig=TRUE, echo=FALSE,width=7>>=
par(mar=c(5.1,4.1,2.4,1.1))
plot(y=odata$y, x=odata$x, ylab = "y", xlab = "x", col = "gray")
lines(x = odata$x, y = odata$mu, col = pal["forest"], lwd = 2.5, main = "")
polygon(c(odata$x, rev(odata$x)), c(odata$mu + odata$sigma, rev(odata$mu - odata$sigma)),
  col = pallight["forest"], border = "transparent")
legend("topleft", expression(mu  %+-%  sigma), bty = "n")
@
\end{center}
\end{frame}

% \begin{frame}
% \frametitle{Distributional trees}
% \vspace*{-0.12cm}
% \begin{center} 
% DGP:  $\; Y\ |\ X = x \; \sim  \; \mathcal{N}(\mu(x), \sigma^2(x))$
% 
% \vspace*{-0.21cm}
% \setkeys{Gin}{width=0.5\linewidth}
% <<plottree_dgp, fig=TRUE, echo=FALSE>>=
% paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
% plot(fig, inner_panel = node_inner_ext,
%      tp_args = list(FUN = identity, width = 18, fill = paltrees[c(1, 3)]), 
%      ip_args = list(fill = paltrees[c(2, 3)]),
%      drop_terminal = TRUE, tnex = 1.7)
% @
% \end{center}
% \end{frame}

<<tree, echo=FALSE, results=hide>>=
set.seed(7)
nobs <- 500
x <- runif(nobs, 0, 1)
mu <- sigma <- ytrue <- numeric(length = nobs)
for(i in 1:nobs) sigma[i] <- if(x[i]<=0.4) 1 else 3
for(i in 1:nobs) mu[i] <- if(x[i]<= 0.4|| x[i]>0.8) 4 else 12
y <- rnorm(nobs, mean = mu, sd = sigma)
#y <- rcnorm(nobs, mean = mu, sd = sigma, left = 0)
ytrue <- mu
data <- data.frame(cbind(y,x, ytrue))
tree <- disttree(y ~ x, data = data, family = NO(), type.tree = "mob")
#tree <- disttree(y ~ x, data = data, family = dist_list_cens_normal)
@

\begin{frame}[fragile]
\frametitle{Distributional trees}
\begin{center}
\vspace*{-0.12cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.5\linewidth}
<<plottree_estpar, fig=TRUE, echo=FALSE>>=
# function for output in terminal panels
FUN <- function (x) 
{
  cf <- x$coefficients
  cf <- matrix(cf, ncol = 1, dimnames = list(names(cf), ""))
  c(sprintf("n = %s", x$nobs), "Estimated parameters:", parse(text = paste0("mu == '", format(round(cf[1], 2), nsmall = 2), "'")), 
                                                        parse(text = paste0("sigma == '", format(round(cf[2], 2), nsmall = 2), "'")))
}

paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))

## plot version using FUN and tree of class 'disttree'
plot(tree, drop = TRUE, tnex = 1.7, FUN = FUN,
     tp_args = list(fill = paltrees[c(1, 3)], width = 18), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Distributional trees}
\begin{center}
\vspace*{-0.12cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.5\linewidth}
<<plottree_box, fig=TRUE, echo=FALSE>>=
plot(as.constparty(tree), tnex = 1.7, drop = TRUE,
     tp_args = list(fill = paltrees[c(1, 3)], ylines = 1.5), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Distributional trees}
\begin{center}
\vspace*{-0.12cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.5\linewidth}
<<plottree_dens, fig=TRUE, echo=FALSE>>=
node_density <- function (tree, xscale = NULL, yscale = NULL, horizontal = FALSE,
                          main = "", xlab = "", ylab = "Density", id = TRUE, rug = TRUE,
                          fill = paltrees[c(1, 3)], col = "black", lwd = 0.5, ...) 
{
  yobs <- tree$data[,as.character(tree$info$formula[[2]])]
  ylines <- 1.5
  if (is.null(xscale)) xscale <- c(-5.1,22.5)
  if (is.null(yscale)) yscale <- c(-0.05,0.45)
  xr <- xscale
  yr <- yscale

  if (horizontal) {
    yyy <- xscale
    xscale <- yscale
    yscale <- yyy
  }

  rval <- function(node) {
    yrange <- seq(from = -20, to = 90)/4
    ydens <- node$info$object$ddist(yrange)

    top_vp <- viewport(layout = grid.layout(nrow = 2, ncol = 3, 
                                            widths = unit(c(ylines, 1, 1), c("lines", "null", "lines")), 
                                            heights = unit(c(1, 1), c("lines", "null"))), 
                       width = unit(1, "npc"), 
                       height = unit(1, "npc") - unit(2, "lines"), 
                       name = paste("node_density",node$id, sep = ""))
    pushViewport(top_vp)
    grid.rect(gp = gpar(fill = "white", col = 0))
    top <- viewport(layout.pos.col = 2, layout.pos.row = 1)
    pushViewport(top)
    mainlab <- paste(ifelse(id, paste("Node", node$id, "(n = "), "n = "), node$info$nobs, ifelse(id, ")", ""), sep = "")

    grid.text(mainlab)
    popViewport()
    plot <- viewport(layout.pos.col = 2, layout.pos.row = 2, 
                     xscale = xscale, yscale = yscale, 
                     name = paste("node_density",  node$id, "plot", sep = ""))
    pushViewport(plot)
    yd <- ydens
    xd <- yrange
    if (horizontal) {
      yyy <- xd
      xd <- yd
      yd <- yyy
      yyy <- xr
      xr <- yr
      yr <- yyy
      rxd <- rep(0, length(xd))
      ryd <- rev(yd)
    } else {
      rxd <- rev(xd)
      ryd <- rep(0, length(yd))
    }

    if (rug) {
      nodeobs <- node$info$object$y
      if (horizontal) {
        grid.rect(x = xscale[1], y = nodeobs , height = 0, width = xscale[1], 
                  default.units = "native", just = c("right", "bottom"),
		  gp = gpar(lwd = 2, col = gray(0, alpha = 0.18)))
      } else {
        grid.rect(x = nodeobs, y = yscale[1], 
                  width = 0, height = abs(yscale[1]), default.units = "native", 
                  just = c("center", "bottom"),
		  gp = gpar(lwd = 2, col = gray(0, alpha = 0.18)))
        #grid.lines(x = xr, y = yr, gp = gpar(col = "lightgray"), 
        #           default.units = "native")
        #grid.lines(x = xr, y = yr, gp = gpar(col = "lightgray"), 
        #           default.units = "native")
      }
    }

    grid.polygon(x = c(xd, rxd), y = c(yd, ryd), default.units = "native",
              gp = gpar(col = "black", fill = fill, lwd = lwd))
    #grid.lines(x = xd, y = yd, default.units = "native", 
    #           gp = gpar(col = col, lwd = lwd))
    grid.xaxis()
    grid.yaxis()
    grid.rect(gp = gpar(fill = "transparent"))
    upViewport(2)
  }
  return(rval)
}

class(node_density) <- "grapcon_generator"

plot(tree, tnex = 1.7, drop = TRUE,
     terminal_panel = node_density,
     tp_args = list(fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}
%\begin{frame}
%\frametitle{Distributional trees}
%\center
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{tree_box}
%\end{frame}

\subsection{Learning}

\begin{frame}
\frametitle{Learning distributional trees and forests}
\begin{minipage}{0.7\textwidth}
{\bf Tree:}
\begin{enumerate}
\item<3-> Fit global distributional model $\mathcal{D}(Y; \theta)$: \\ % to the whole data set:\\
Estimate $\hat{\theta}$ via maximum likelihood \\
$\hat{\theta} = \argmax_{\theta \in \Theta} \sum_{i=1}^n \ell(\theta; y_i)$
\item<8-> Test for associations/instabilities of the \\
scores $\frac{\partial \ell}{\partial \theta}(\hat{\theta};y_i)$ and each possible split variable $X_j$.
\end{enumerate}
\end{minipage}
\begin{minipage}{0.23\textwidth}
%\vspace{-0.1cm}
\begin{tikzpicture}
\visible<2-3>{
\node[ellipse, fill=HighlightBlue!70, align=center, scale = 0.7, minimum width=60pt, minimum height = 30pt] (n0) at (0.8, 1.7) {$Y$};
}
\visible<4>{
\node[ellipse, fill=HighlightBlue!70, align=center, scale = 0.7, minimum width=60pt, minimum height = 30pt] (n0) at (0.8, 1.7) {$\mathcal{D}(Y;\hat{\theta}$)};
}
\visible<5>{
\node[inner sep=0pt] (density_all) at (0.8, 1.7)
{\includegraphics[width=0.7\textwidth]{density_all}};
}
\visible<6->{
\node[inner sep=0pt] (density_all) at (0.8, 1.7)
{\includegraphics[width=0.7\textwidth]{density_all_hist}};
}
\visible<7-9>{
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n1) at (0, 0.2) {?};
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n2) at (1.6, 0.2) {?};
\draw[-, gray, line width=0.5pt] (0.7, 1.16) -- (n1);
\draw[-, gray, line width=0.5pt] (0.9, 1.16) -- (n2);
}
\visible<10-11>{
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n1) at (0, 0.2) {$Y_1$};
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n2) at (1.6, 0.2) {$Y_2$};
\draw[-, gray, line width=0.5pt] (0.7, 1.16) -- (n1) node [midway, left] {\scriptsize $X \leq p$};
\draw[-, gray, line width=0.5pt] (0.9, 1.16) -- (n2) node [midway, right] {\scriptsize $X > p$};
}
\visible<12>{
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n1) at (0, 0.2) {$\mathcal{D}(Y_1;\hat{\theta}_1$)};
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n2) at (1.6, 0.2) {$\mathcal{D}(Y_2;\hat{\theta}_2$)};
}
\visible<12->{
\draw[-, gray, line width=0.5pt] (0.7, 1.16) -- (n1) node [midway, left] {\scriptsize $X \leq p$};
\draw[-, gray, line width=0.5pt] (0.9, 1.16) -- (n2) node [midway, right] {\scriptsize $X > p$};
}
\visible<13->{
\node[inner sep=0pt] (density_1_hist) at (-0.2,-0.04)
{\includegraphics[width=0.6\textwidth]{density_2_hist}};
}
\visible<13->{
\node[inner sep=0pt] (density_2_hist) at (1.8,-0.04)
{\includegraphics[width=0.6\textwidth]{density_1_hist}};
}
\end{tikzpicture}
\end{minipage}
\vspace{0.1cm}
%\begin{adjustwidth}{-0.0em}{-1em}
\begin{enumerate}
\setcounter{enumi}{2}
\item<9-> Split along the covariate $X$ with strongest association or instability \\
and at breakpoint $p$ with highest improvement in log-likelihood.
\item<11-> Repeat steps 1--3 recursively until some stopping criterion \\
is met, yielding $B$ subgroups $\mathcal{B}_b$ with $b = 1, \dots, B$.
\end{enumerate}
%\end{adjustwidth}
\vspace{0.4cm}
\visible<14->{
{\bf Forest:} Ensemble of $T$ trees.
\begin{itemize}
\item Bootstrap or subsamples.
\item Random input variable sampling.
\end{itemize}
}
\end{frame}

\begin{frame}
\frametitle{Adaptive local likelihood estimation}

\textbf{Parameter estimator} for \only<1-3>{\textbf{a global}}\only<4->{\textbf{an adaptive local}} model\\
with learning data \only<1>{$\{y_i\}_{i=1,\ldots,n}$}\only<2->{$\{(y_i,\bold{x}_i)\}_{i=1,\ldots,n}$} :

\[
\hat{\theta}\visible<2->{(\bold{x})} =  \argmax_{\theta \in \Theta} \sum_{i=1}^n \visible<2->{w_i(\bold{x}) \cdot} \ell(\theta; y_i)
\]

\medskip

\visible<3->{
\textbf{Weights:}
\begin{eqnarray*}
w^{\text{base}}_i(\bold{x})   & = & 1 \\[0.2cm]
\visible<4->{
w^{\text{tree}}_i(\bold{x})   & = & \sum_{b=1}^B I((\bold{x}_i \in \mathcal{B}_b) \land (\bold{x} \in \mathcal{B}_b)) \\[0.1cm]
\visible<5->{
w^{\text{forest}}_i(\bold{x}) & = & \frac{1}{T} \sum_{t=1}^T \sum_{b=1}^{B^t} \frac{I((\bold{x}_i \in \mathcal{B}^t_b) \land (\bold{x} \in \mathcal{B}^t_b))}{\left|\mathcal{B}^t_b\right|}
\end{eqnarray*}
}}}
\end{frame}

\subsection{Unbiased recursive partitioning}

\begin{frame}%[fragile]
\frametitle{}
\vspace{1.5cm}
\center
{\bf \large The Power of Unbiased Recursive Partitioning: \\
\smallskip
A Unifying View of CTree, MOB, and GUIDE}

\bigskip
\bigskip

{\small
Lisa Schlosser, Torsten Hothorn, Achim Zeileis}

\bigskip
\bigskip

{\footnotesize
Submitted to\\
\emph{Computational Statistics \& Data Analysis}

\medskip

Preprint version at \url{https://arxiv.org/abs/1906.10179}}

\end{frame}

\begin{frame}
\frametitle{Unbiased recursive partitioning}

\medskip

\textbf{Idea:} 
\begin{itemize}
\smallskip
  \item Separation of split variable and split point selection.
\smallskip
  \item Application of a statistical test to assess associations between the response~$Y$ (or corresponding scores) and each possible split variable~$X_j$.
\smallskip
  \item Split variable selection based on resulting $p$-values.
\end{itemize}
\end{frame}
% \begin{frame}%[fragile]
% \frametitle{Unbiased recursive partitioning}
% 
% {\bf General testing strategy for split variable selection:}
% \begin{enumerate}
% \item Evaluate a discrepancy measure capturing the observation-wise goodness of fit of
% the considered model.
% %$\mathcal{M}(Y,X;\hat{\beta})$.
% \item Apply a statistical test assessing dependency of the discrepancy measure to each possible split variable $X_j$.
% \item Select the split variable $X$ %$X_j^\ast$ 
% showing the smallest $p$-value.
% \end{enumerate}
% 
% \bigskip
% 
% {\bf Discrepancy measures:} (Model-based) transformations of $Y$, possibly for each model parameter.
% \begin{itemize}
% \item (Ranks of) $Y$.
% \item (Absolute) deviations $Y - \bar Y$.
% \item Residuals of the fitted model. %$\mathcal{M}(Y,X;\hat{\beta})$.
% \item Score matrix of the fitted model. %$\mathcal{M}(Y,X;\hat{\beta})$.
% \item \ldots
% \end{itemize}
% \end{frame}

\begin{frame}
\frametitle{Unbiased recursive partitioning algorithms}

\textbf{GUIDE:} Loh (2002, \emph{Statistica Sinica}).
\begin{itemize}
  \item First \emph{unbiased} algorithm for recursive partitioning of linear models.
  \item Based on categorical association tests. 
\end{itemize}

\medskip

\textbf{CTree:} Hothorn, Hornik, Zeileis (2006, \emph{JCGS}).
\begin{itemize}
  \item Proposed as unbiased recursive partitioning for \emph{nonparametric} modeling.
  \item Can be model-based via model scores as the response transformation.
  \item Based on conditional permutation tests.
\end{itemize}

\medskip

\textbf{MOB:} Zeileis, Hothorn, Hornik (2008, \emph{JCGS}).
\begin{itemize}
  \item Model-based unbiased recursive partitioning. % using M-estimation (ML, OLS, CRPS, \ldots)
  \item Based on asymptotic parameter instability tests.
\end{itemize}

\end{frame}

\subsection{A unifying view}
\begin{frame}
\frametitle{A unifying view}

\textbf{Algorithms:} CTree, MOB, GUIDE are all `flavors' of a general framework.

\bigskip

\textbf{Building blocks:} For standard setup.

\bigskip

\begin{tabular}{llccl}
\hline 
      & Scores  & Binarization  & Categorization  & Test statistic \\ 
\hline
CTree & Model scores & --	           & -- 	           & Sum of squares\\
MOB   & Model scores & --	           & -- 	           & Maximally selected\\
GUIDE & Residuals    & \checkmark    & \checkmark      & Sum of squares\\
\hline
\end{tabular}

\bigskip

\textbf{Remarks:}
\begin{itemize}
  \item All three algorithms allow for certain modifications of standard setup.
  \item Further differences, e.g., null distribution, pruning strategy, etc.
\end{itemize}

\end{frame}

% \begin{frame}%[fragile]
% \frametitle{General Framework}
% 
% {\bf{Building blocks:}}
% 
% \begin{itemize}
% \item {\only<4->{\color{lightgray}} Residuals vs.\ full model scores.}
% \item {\only<1-3,6->{\color{lightgray}} Binarization of residuals/scores.}
% \item {\only<1-5>{\color{lightgray}} Categorization of possible split variables.}
% \end{itemize}
% 
% \bigskip
% 
% 
% \only<2>{
% \small
% $$
% s(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
% -2 \cdot \begin{pmatrix} 
% r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) \cdot X_1}\\
% r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) \cdot X_2}\\
% \vdots & {\color{lightgray} \vdots} \\
% r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) \cdot X_n}
% \end{pmatrix}
% %\begin{pmatrix}
% %\partial_{\beta_0} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})}\\
% %\partial_{\beta_0} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})}\\
% %\vdots & {\color{lightgray} \vdots} \\
% %\partial_{\beta_0} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & {\color{lightgray} \partial_{\beta_1} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})}
% %\end{pmatrix}
% $$
% }
% \only<3>{
% \small
% $$
% s(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
% -2 \cdot \begin{pmatrix} 
% r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) \cdot X_1\\
% r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) \cdot X_2\\
% \vdots & \vdots \\
% r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) \cdot X_n
% \end{pmatrix}
% %\begin{pmatrix} 
% %\partial_{\beta_0} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})\\
% %\partial_{\beta_0} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})\\
% %\vdots & \vdots \\
% %\partial_{\beta_0} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1}) & \partial_{\beta_1} r^2(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})
% %\end{pmatrix}
% $$
% }
% \only<4-5>{
% \small
% $$
% r(Y, X, \hat{\beta_0}, \hat{\beta_1}) = 
% \begin{pmatrix}
% r(Y_1, X_1, \hat{\beta_0}, \hat{\beta_1})\\
% r(Y_2, X_2, \hat{\beta_0}, \hat{\beta_1})\\
% \vdots \\
% r(Y_n, X_n, \hat{\beta_0}, \hat{\beta_1})
% \end{pmatrix}
% %
% \visible<5>{
% \quad \Rightarrow \quad
% \small
% \begin{pmatrix}
% > 0\\
% \le 0\\
% \vdots \\
% > 0
% \end{pmatrix}
% }
% $$
% }
% \only<6-7>{
% \small
% $$
% Z_j = 
% \begin{pmatrix}
% Z_{j1}\\
% Z_{j2}\\
% \vdots \\
% Z_{jn}\\
% \end{pmatrix}
% %
% \visible<7>{
% \quad \Rightarrow \quad
% \small
% \begin{pmatrix}
% \text{Q}3\\
% \text{Q}1\\
% \vdots \\
% \text{Q}2
% \end{pmatrix}
% }
% $$
% }
% \end{frame}

% \begin{frame}%[fragile]
% \frametitle{Pruning}
% 
% {\bf Goal:} Avoid overfitting.
% 
% \bigskip
% 
% {\bf Two strategies:}
% 
% \begin{itemize}
% \item \emph{Pre-pruning:}
% Internal stopping criterion based on Bonferroni-corrected $p$-values of the underlying tests. Stop splitting when there is no significant association.
% \item \emph{Post-pruning:}
% First grow a very large tree and afterwards prune splits that do not improve the model fit, 
% either via cross-validation (e.g.,~cost-complexity pruning as in CART) or 
% based on information criteria (e.g.,~AIC or BIC). 
% \end{itemize}
% \end{frame}
% 
% \subsection{Recommendations}
% \begin{frame}%[fragile]
% \frametitle{Recommendations}
% 
% \vspace{0.2cm} 
% 
% \textbf{Based on evaluations in our simulation study:}
% 
% \vspace{0.2cm} 
% 
% \begin{itemize}
% \setlength\itemsep{0.2cm}
% \item Full model scores better than residuals only. %: Using scores allows for capturing changes in all model parameters.
% \item Original values of scores/residuals better than binarized values.
% \item Categorization is simpler, but less powerful in margins.
% \item Maximally-selected statistics (as in MOB) more powerful for abrupt shifts.
% \item Linear statistics (default in CTree) more powerful for linear changes.
% \item If the significance test performs well pre-pruning works well,\\
% otherwise post-pruning might be needed.
% %\item Including full model scores is advantageous over using only residuals.
% %\item Binarizing scores/residuals does not improve the performance.
% %\item Categorizing split variables can simplify calculations, however,
% %effects might be missed unless a binning point is close to the true split point.
% %\item If the applied test performs well as significance test pre-pruning
% %is preferable, otherwise post-pruning.
% \end{itemize}
% \end{frame}

\subsection{Recommendations}
\begin{frame}%[fragile]
\frametitle{Recommendations}

\vspace{0.34cm}

\textbf{Based on evaluations in our simulation study:}

\vspace{0.24cm}

\begin{itemize}
\setlength\itemsep{0.23cm}
\item Best performance with full model scores, without binarization, and without categorization of possible split variables.
\item Adapting GUIDE to employing full model scores improves its performance.
\item Preferable test statistic depends on the type of effects (maximally-selected for abrupt shifts, linear statistics for linear changes).
%\item If the significance test performs well pre-pruning works well,\\
%otherwise post-pruning might be needed.
%\item Including full model scores is advantageous over using only residuals.
%\item Binarizing scores/residuals does not improve the performance.
%\item Categorizing split variables can simplify calculations, however,
%effects might be missed unless a binning point is close to the true split point.
%\item If the applied test performs well as significance test pre-pruning
%is preferable, otherwise post-pruning.
\end{itemize}
\end{frame}

\subsection{Weather forecasting}

\begin{frame}%[fragile]
\frametitle{}
\vspace{1.5cm}
\center
{\bf \large Distributional Regression Forests for Probabilistic\\
\smallskip
Precipitation Forecasting in Complex Terrain}

\bigskip
\bigskip

{\small
Lisa Schlosser, Torsten Hothorn, Reto Stauffer, Achim Zeileis}

\bigskip
\bigskip

{\footnotesize
Published in\\
\emph{The Annals of Applied Statistics}

\medskip

\doi{10.1214/19-AOAS1247}}

\end{frame}

\begin{frame}[fragile]
\frametitle{Weather forecasting}

\textbf{Goal:}

\medskip

\begin{center}
\begin{tikzpicture}
\draw[->] (0.4,0)--(0.9,0);
\draw[] (1,-0.4) rectangle (2.5,0.4);
\node[font=\scriptsize] at (1.75,0) {nature};
\draw[->] (2.6,0)--(3.1,0);
\visible<1>{%
\node[] at (0.0,0) {X};
\node[] at (3.5,0) {Y};}
\visible<2->{%
\node[inner sep=0pt] (today) at (-2.3,-0.4)
{\includegraphics[width=.365\textwidth]{airport_20200412.jpg}};
\node[font=\scriptsize] at (-2.3,-2.4) {2020-04-12};
\node[inner sep=0pt] (future) at (5.8,-0.4)
{\includegraphics[width=.365\textwidth]{airport_20200413.jpg}};
\node[font=\scriptsize] at (5.8,-2.4) {2020-04-13};}
\end{tikzpicture}
\end{center}

%% https://innsbruck-airport.panomax.com/

\only<1-2>{%
\textbf{Data:}
\begin{itemize}
\item X: State of the atmosphere now (temperature, precipitation, wind, \dots).
\item Y: State of the atmosphere in the future (hours, days, weeks, \dots).
\end{itemize}}%
\only<3->{%
\textbf{Two stages:}
\begin{itemize}
\item Physical model: Numerical weather prediction (NWP).
\item Statistical model: Model output statistics (MOS).
\end{itemize}}

\end{frame}

\begin{frame}[fragile]
\frametitle{Weather forecasting}

\textbf{Numerical weather prediction (NWP):}
\begin{itemize}
  \item Based on a physical model.
  \item Massive numerical simulation of atmospheric processes.
  \item Here: Global model on a $50\times 50 \text{ km}^2$ grid.
\end{itemize}

\medskip

\textbf{Problem:} Uncertain initial conditions, unresolved processes.

\medskip

\textbf{Solution:} Ensemble of simulation runs under perturbed conditions.

\end{frame}

<<echo = FALSE, results=hide>>=
library("zoo")

Sys.setenv("TZ"="UTC")

# -------------------------------------------------------------------
# GFS ensemble forecast
# -------------------------------------------------------------------

GFS   <- read.table("Data/GENS_00_innsbruck_20200412.dat", header = TRUE)
init  <- min(as.POSIXct(GFS$timestamp-GFS$step*3600,origin="1970-01-01"))
title <- sprintf("%s\n%s",
                 "Global Forecast System (GFS) Ensemble Forecast for Innsbruck, Airport",
                 sprintf("Forecast initialized %s", strftime(init, "%Y-%m-%d %H:%M UTC")))

# observations
load("Data/STAGEobs_tawes_11121_defense.rda")
#load("Data/STAGEobs_tawes_11121.rda")
# "2018-03-14 00:00:00" in line 321 (first entry for temperature)
# "2018-03-13 06:00:00" in line 317 (first entry for rain -> aggregated to 24h sums)
# "2018-03-23 00:00:00" in line 357

getZoo <- function( x, variable ) {
  x <- subset( x, varname == variable )
  if ( nrow(x) == 0 ) stop("Ups, variable seems not to exist at all!")
  # Else create zoo
  x <- zoo( subset(x,select=-c(varname,timestamp,step)),
            as.POSIXct(x$timestamp,origin="1970-01-01") )
  x
}
GFS_t2m  <- getZoo( GFS, "tmp2m" ) - 273.15
# temperature bias: model simulates temperature for height 1675 (~1070 m higher than true height)
# => add 10.7 degrees to account for bias (approx. 1 degree / 100 m)
# Custom bias
GFS_t2m <- GFS_t2m + 6.5

# Subsetting obs
obs <- subset(obs, index(obs) %in% index(GFS_t2m))

GFS_t2m <- merge(GFS_t2m, obs = obs[, grep("tl", names(obs))])
###GFS_t2m$obs <- obs[,"obs$tl"]
# first value of rain is 18 hours later than temperature 
#(both start at 6:00 but rain is cumulated over 24 hours)
# -> drop first 3 values of temperature
GFS_t2m <- GFS_t2m[-c(1:3),]

GFS_rain <- getZoo( GFS, "apcpsfc" )
GFS_rain$obs <- obs[,"rr6"]
# precipitaion sums over 24 hours (00:00 UTC - 00:00 UTC +1day)
GFS_rain24 <- GFS_rain[c((1:10)*4),]
for(i in 1:NROW(GFS_rain24)){
  GFS_rain24[i,] <- colSums(GFS_rain[c((4*i-3):(4*i)),])
}

# Two POSIXct vectors for the axis
main  <- seq(min(as.POSIXct(as.Date(index(GFS_t2m)))),max(index(GFS_t2m)),by=86400)
minor <- seq(min(index(GFS_t2m)),max(index(GFS_t2m)),by=3*3600)

# GFS_t2m_mean <- rowMeans(GFS_t2m)
GFS_t2m_mean <- GFS_t2m
GFS_t2m_mean[,1] <- rowMeans(GFS_t2m[,-NCOL(GFS_t2m)])
GFS_t2m_mean <- GFS_t2m_mean[,1]

# GFS_rain_mean <- rowMeans(GFS_rain)
GFS_rain24_mean <- GFS_rain24
GFS_rain24_mean[,1] <- rowMeans(GFS_rain24[,-NCOL(GFS_rain24)])
GFS_rain24_mean <- GFS_rain24_mean[,1]
@

<<t2m, echo=FALSE, eval=FALSE>>=
par(mar=c(0.4,0,0.2,0), oma=c(6,3.2,5,4))
layout( matrix(1:2,ncol=1) )
plot(GFS_t2m[,1], screen=1, xaxs="i", xaxt="n", ylim = range(GFS_t2m, na.rm = TRUE)*1.05, col = "slategray" )
abline(v = main, lty = 2 )
mtext(side = 3, line = 1, cex = 1.2, font = 2, title )
mtext(side = 2, line = 2.3, cex = 1, font = 1, text = "Temperature [°C]" )
@

<<rain24, echo=FALSE, eval=FALSE>>=
plot(GFS_rain24[,1], screen=1, xaxs="i", xaxt="n", yaxs="i", ylim=range(GFS_rain24, na.rm = TRUE)*c(0,1.05), col = "slategray" )
abline(v = main, lty = 2 )
mtext(side = 2, line = 2.3, cex = 1, font = 1, text = "Rain [mm/6h]" )
axis(side = 1, at = main + 42300,  strftime(main,"%b %d"),
line = 0, col.ticks=NA, col.axis="gray30", col = NA )
@

\begin{frame}[fragile]
\frametitle{Weather forecasting}

\vspace{-0.6cm}
\setkeys{Gin}{width=0.95\linewidth}
\begin{center}
\only<1>{%
<<gfs1, fig=TRUE, echo=FALSE, width=9>>=
<<t2m>>
<<rain24>>
@
}%
\only<2>{%
<<gfs2, fig=TRUE, echo=FALSE, width=9>>=
<<t2m>>
lines(GFS_t2m[,3], col = "slategray")
<<rain24>>
lines(GFS_rain24[,3], col = "slategray")
@
}%
\only<3>{%
<<gfs3, fig=TRUE, echo=FALSE, width=9>>=
<<t2m>>
lines(GFS_t2m[,2], col = "slategray")
lines(GFS_t2m[,3], col = "slategray")
<<rain24>>
lines(GFS_rain24[,2], col = "slategray")
lines(GFS_rain24[,3], col = "slategray")
@
}%
\only<4>{%
<<gfs4, fig=TRUE, echo=FALSE, width=9>>=
<<t2m>>
for(j in 2:(NCOL(GFS_t2m) - 1)) lines(GFS_t2m[,j], col = "slategray")
<<rain24>>
for(j in 2:(NCOL(GFS_rain24) - 1)) lines(GFS_rain24[,j], col = "slategray")
@
}%
\only<5>{%
<<gfs5, fig=TRUE, echo=FALSE, width=9>>=
<<t2m>>
for(j in 2:(NCOL(GFS_t2m) - 1)) lines(GFS_t2m[,j], col = "slategray")
lines(GFS_t2m[,"obs"], col = 2, lwd = 2.3)
<<rain24>>
for(j in 2:(NCOL(GFS_rain24) - 1)) lines(GFS_rain24[,j], col = "slategray")
lines(GFS_rain24[,"obs"], col = 2, lwd = 2.3)
@
}%
\end{center}

\end{frame}

\subsection{Weather forecasting: precipitation}

\begin{frame}[fragile]
\frametitle{Weather forecasting: precipitation}

\textbf{Goal:} Predict daily precipitation amount in Tyrol. %complex terrain.

\bigskip
\pause

\textbf{Observation data:} National Hydrographical Service.
\begin{itemize}
\item Daily 24h precipitation sums from July over 28 years (1985--2012).
\item 95 observation stations in Tyrol.
\end{itemize}

\bigskip
\pause

\textbf{NWP:} Global Ensemble Forecast System.
\begin{itemize}
\item Model outputs: Precipitation, temperature, air pressure, 
  convective available potential energy, downwards short wave radiation flux, \dots
\item 80 covariates based on ensemble min/max/mean/standard deviation.
\end{itemize}

\bigskip
\pause

\textbf{Distribution assumption:} Power-transformed Gaussian, censored at 0.
\[ (\text{precipitation})^\frac{1}{1.6} \sim \textit{c}\mathcal{N}(\mu,\sigma^2) \]

\end{frame}

%% \begin{frame}
%% \frametitle{Weather forecasting: precipitation}
%% 
%% \textbf{Base variables:}
%% \begin{itemize}
%% \item Total precipitation.
%% \item Convective available potential energy.
%% \item Downwards short wave radiation flux (``sunshine'').
%% \item Mean sea level pressure.
%% \item Preciptable water.
%% \item 2m maximum temperature.
%% \item Total column-integrated condensate.
%% \item Temperature.
%% \item Temperature differences in altitude.
%% \end{itemize}
%% 
%% \medskip
%% 
%% \textbf{Variations:} 80 covariates based on ensemble min/max/mean/standard deviation.
%% 
%% \end{frame}

<<pred_axams, eval=TRUE, echo=FALSE, results=hide>>=
if(file.exists("Data/Axams_pred.rda") & file.exists("Data/Axams_testdata.rda")){
  load("Data/Axams_pred.rda")
  load("Data/Axams_testdata.rda")
} else {
  #####
  # load observations and covariates 
  data("RainAxams")

  # tree and forest formula
  {
    dt.formula <- df.formula <- 
      robs ~ tppow_mean + tppow_sprd + tppow_min + tppow_max + 
      tppow_mean0612 + tppow_mean1218 + tppow_mean1824 + tppow_mean2430 + 
      tppow_sprd0612 + tppow_sprd1218 + tppow_sprd1824 + tppow_sprd2430 + 
      capepow_mean + capepow_sprd + capepow_min + capepow_max + 
      capepow_mean0612 + capepow_mean1218 + capepow_mean1224 + capepow_mean1230 +
      capepow_sprd0612 + capepow_sprd1218 + capepow_sprd1224 + capepow_sprd1230 +
      dswrf_mean_mean + dswrf_mean_max +  
      dswrf_sprd_mean + dswrf_sprd_max + 
      msl_mean_mean + msl_mean_min + msl_mean_max + 
      msl_sprd_mean + msl_sprd_min + msl_sprd_max +
      pwat_mean_mean + pwat_mean_min + pwat_mean_max + 
      pwat_sprd_mean + pwat_sprd_min + pwat_sprd_max +
      tmax_mean_mean + tmax_mean_min + tmax_mean_max +
      tmax_sprd_mean + tmax_sprd_min + tmax_sprd_max +
      tcolc_mean_mean + tcolc_mean_min + tcolc_mean_max +
      tcolc_sprd_mean + tcolc_sprd_min + tcolc_sprd_max +
      t500_mean_mean + t500_mean_min + t500_mean_max +
      t700_mean_mean + t700_mean_min + t700_mean_max +
      t850_mean_mean + t850_mean_min + t850_mean_max +
      t500_sprd_mean + t500_sprd_min + t500_sprd_max +
      t700_sprd_mean + t700_sprd_min + t700_sprd_max +
      t850_sprd_mean + t850_sprd_min + t850_sprd_max +
      tdiff500850_mean + tdiff500850_min + tdiff500850_max +
      tdiff700850_mean + tdiff700850_min + tdiff700850_max +
      tdiff500700_mean + tdiff500700_min + tdiff500700_max +
      msl_diff

  }

  # learning data: 24 years (1985 - 2008, both inlcuded)
  # testing data: 4 successive years (2009, 2010, 2011, 2012)
  learndata <- RainAxams[RainAxams$year < 2009,]
  testdata <- RainAxams[RainAxams$year %in% c(2009, 2010, 2011, 2012),]
  save(file = "Data/Axams_testdata.rda", testdata)

  ##############################################################
  # fitting the models
  set.seed(7)

  df <- distforest(df.formula, data = learndata, family = dist_list_cens_normal, 
                   ntree = 100, censtype = "left", censpoint = 0,
                   control = disttree_control(teststat = "quad", testtype = "Univ",
                                              type.tree = "ctree",
                                              intersplit = TRUE,
                                              mincriterion = 0, minsplit = 50,
                                              minbucket = 20), mtry = 27)

  #### prepare data for plot of estimated density functions
  # predictions for one day (in each of the four years) 
  # (19th of July 2011 is missing)
  pday <- 24  # 2 (hohe Beobachtung zu niedrig geschaetzt), 4, 15, evtl. auch 7, 8, 23 (eine 0-Beobachtung und 2 sehr aehnliche),

  pdays <- if(pday<19) c(pday, pday + 31, pday + 62, pday + 92) else c(pday, pday + 31, pday + 61, pday + 92)
  pdf <- predict(df, newdata = testdata[pdays,], type = "parameter")
  pdf$pdays <- pdays
  save(file = "Data/Axams_pred.rda", pdf)
}

pdays <- pdf$pdays
pday <- pdays[1]
df_mu <- pdf$mu
df_sigma <- pdf$sigma

# plot predicted distributions together with observations
set.seed(7)
x <- c(0.01, sort(runif(500,0.01,8)))
y1 <- crch::dcnorm(x, mean = df_mu[1], sd = df_sigma[1], left = 0)
y2 <- crch::dcnorm(x, mean = df_mu[2], sd = df_sigma[2], left = 0)
y3 <- crch::dcnorm(x, mean = df_mu[3], sd = df_sigma[3], left = 0)
y4 <- crch::dcnorm(x, mean = df_mu[4], sd = df_sigma[4], left = 0)

# switch x-axis back to untransformed scale
# x <- x^(1.6)

# point mass (slightly shifted)
pm1 <- c(0.05, crch::dcnorm(-1, mean = df_mu[1], sd = df_sigma[1], left = 0))     #  0.15
pm2 <- c(0.01, crch::dcnorm(-1, mean = df_mu[2], sd = df_sigma[2], left = 0))     #  0.05
pm3 <- c(-0.03, crch::dcnorm(-1, mean = df_mu[3], sd = df_sigma[3], left = 0))    # -0.05
pm4 <- c(-0.07, crch::dcnorm(-1, mean = df_mu[4], sd = df_sigma[4], left = 0))    # -0.15

# predictions
pred1 <- c(testdata[pdays,"robs"][1], # ^(1.6), 
           crch::dcnorm(testdata[pdays,"robs"][1], mean = df_mu[1], sd = df_sigma[1], left = 0))
pred2 <- c(testdata[pdays,"robs"][2], # ^(1.6), 
           crch::dcnorm(testdata[pdays,"robs"][2], mean = df_mu[2], sd = df_sigma[2], left = 0))
pred3 <- c(testdata[pdays,"robs"][3], # ^(1.6), 
           crch::dcnorm(testdata[pdays,"robs"][3], mean = df_mu[3], sd = df_sigma[3], left = 0))
pred4 <- c(testdata[pdays,"robs"][4], # ^(1.6), 
           crch::dcnorm(testdata[pdays,"robs"][4], mean = df_mu[4], sd = df_sigma[4], left = 0))

#legendheight
lh1 <- crch::dcnorm(0.01, mean = df_mu[1], sd = df_sigma[1], left = 0)
lh2 <- crch::dcnorm(0.01, mean = df_mu[2], sd = df_sigma[2], left = 0)
lh3 <- crch::dcnorm(0.01, mean = df_mu[3], sd = df_sigma[3], left = 0)
lh4 <- crch::dcnorm(0.01, mean = df_mu[4], sd = df_sigma[4], left = 0)
@

\begin{frame}[fragile]
\frametitle{Weather forecasting: precipitation}

\textbf{Application for one station:} Axams.
\begin{itemize}
%\vspace{-0.05cm}
\item Learn forest model on data from 24 years (1985--2008).
%\vspace{-0.05cm}
\item Evaluate on 4 years (2009--2012). Here: July \Sexpr{24}.
\end{itemize}

\vspace*{-0.4cm}
\begin{center}
\setkeys{Gin}{width=0.54\textwidth}
<<plot_pred_axams_24July, fig=TRUE, echo=FALSE, eval=TRUE, height=4.3, width=7>>=
par(mar = c(4, 4, 1, 1))
plot(x = x, y = y1, type = "l", col = 2, ylab = "Density", lwd = 2,
     # xlab = expression(Total~precipitation~"["~mm~"/"~"24h"~"]"),
     xlab = expression(Total~precipitation~"["~mm^(1/1.6)~"/"~"24h"~"]"),
     ylim = c(0,max(y1, y2, y3, y4, pm1, pm2, pm3, pm4) + 0.01),
     xlim = c(-1.5,8))     #  xlim = c(-2,28))

lines(x = x, y = y2, type = "l", lwd = 2, col = 4)
lines(x = x, y = y3, type = "l", lwd = 2, col = 3)
lines(x = x, y = y4, type = "l", lwd = 2, col = 6)
legend('topright', c("Predicted distribution", "Point mass at censoring point", "Observation"),
       bty = "n", col = "black", lty = c(1, NA, NA), pch = c(NA, 19, 4), cex = 1)

# plot point mass
lines(x = c(pm1[1], pm1[1]), y = c(pm1[2], 0), col = 2, type = "l", lwd = 2)
lines(x = c(pm2[1], pm2[1]), y = c(pm2[2], 0), col = 4, type = "l", lwd = 2)
lines(x = c(pm3[1], pm3[1]), y = c(pm3[2], 0), col = 3, type = "l", lwd = 2)
lines(x = c(pm4[1], pm4[1]), y = c(pm4[2], 0), col = 6, type = "l", lwd = 2)

points(x = pm1[1], y = pm1[2], col = 2, pch = 19, cex = 1.4)
points(x = pm2[1], y = pm2[2], col = 4, pch = 19, cex = 1.4)
points(x = pm3[1], y = pm3[2], col = 3, pch = 19, cex = 1.4)
points(x = pm4[1], y = pm4[2], col = 6, pch = 19, cex = 1.4)

# plot predictions
points(x = pred1[1], y = pred1[2], col = 2, pch = 4, cex = 1.4, lwd = 2)
points(x = pred2[1], y = pred2[2], col = 4, pch = 4, cex = 1.4, lwd = 2)
points(x = pred3[1], y = pred3[2], col = 3, pch = 4, cex = 1.4, lwd = 2)
points(x = pred4[1], y = pred4[2], col = 6, pch = 4, cex = 1.4, lwd = 2)

lines(x = c(pred1[1], pred1[1]), y = c(pred1[2], 0), col = "darkgray", type = "l", lty = 2, lwd = 2)
lines(x = c(pred2[1], pred2[1]), y = c(pred2[2], 0), col = "darkgray", type = "l", lty = 2, lwd = 2)
lines(x = c(pred3[1], pred3[1]), y = c(pred3[2], 0), col = "darkgray", type = "l", lty = 2, lwd = 2)
lines(x = c(pred4[1], pred4[1]), y = c(pred4[2], 0), col = "darkgray", type = "l", lty = 2, lwd = 2)

# add labels
text(x = -0.8, y = lh1, labels = "2009", col = 2, cex = 1)      # -1.7 
text(x = -0.8, y = lh2, labels = "2010", col = 4, cex = 1)      # -1.7
text(x = -0.8, y = lh3, labels = "2011", col = 3, cex = 1)      # -1.7
text(x = -0.8, y = lh4, labels = "2012", col = 6, cex = 1)      # -1.7
@
\end{center}
\end{frame}

% \begin{frame}[fragile]
% \frametitle{Weather forecasting: precipitation}
% 
% \textbf{Application for one station:} Axams.
% \begin{itemize}
% %\vspace{-0.05cm}
% \item Learn forest model on data from 24 years.
% %\vspace{-0.05cm}
% \item Evaluate on 4 years.
% %\vspace{-0.05cm}
% \item 10 times 7-fold cross validation.
% \end{itemize}
% 
% \bigskip
% 
% \textbf{Benchmark:} Against other heteroscedastic censored Gaussian models.
% \begin{itemize}
% \item \emph{Ensemble MOS:} Linear predictors using only total precipitation.
% \item \emph{Prespecified GAMLSS:} Variable selection based on expert knowledge.
% \item \emph{Boosted GAMLSS:} Automatic variable selection.
% \end{itemize}
% 
% \bigskip
% 
% \textbf{Evaluation:} Continuous ranked probability skill score.
% 
% %% \vspace{0.1cm}
% %% \begin{table}[t!]
% %% \footnotesize
% %% 
% %% \hskip-0.4cm\begin{tabular}{ l l l }
% %% \hline
% %% Model & Type & Variable Selection \\
% %% \hline
% %% \vspace{-0.15cm}
% %% \textbf{Distributional forest} & recursive    & automatic \\ 
% %%                                & partitioning &           \\ 
% %% \vspace{-0.15cm}
% %% \textbf{Prespecified GAMLSS}   & spline  & based on expert \\
% %%                                & in each & knowledge      \\
% %% \vspace{-0.15cm}
% %% \textbf{Boosted GAMLSS}        & spline  & automatic  \\
% %%                                & in each &            \\
% %% \vspace{-0.15cm}
% %% \textbf{EMOS}                  & linear & total precipitation mean \\
% %%                                &        & and standard deviation   \\
% %% \hline
% %% \end{tabular}
% %% \end{table}
% 
% \end{frame}
% 
% 
% 
% 
% <<echo=FALSE, results=hide>>=
% #### cross validation rain
% if(file.exists("Data/crps_cross.rda")){
%   load("Data/crps_cross.rda")
% } else {
%   
%   nrep_cross <- 10
%   seed <- 7
%   
%   res_cross <- mclapply(1:nrep_cross,
%                         function(i){
%                           
%                           set.seed(seed*i)
%                           
%                           # randomly split data in 7 parts each including 4 years
%                           years <- 1985:2012
%                           testyears <- list()
%                           for(j in 1:7){
%                             testyears[[j]] <- sample(years, 4, replace = FALSE)
%                             years <- years[!(years %in% testyears[[j]])]
%                           }
%                           
%                           #crps <- matrix(nrow = 7, ncol = 7)
%                           reslist <- list()
%                           for(k in 1:7){
%                             test <- testyears[[k]]
%                             train <- c(1985:2012)[!c(1985:2012) %in% test]
%                             
%                             res <- evalmodels(station = "Axams",
%                                               train = train,
%                                               test = test,
%                                               gamboost_cvr = TRUE)
%                             
%                             #crps[k,] <- res$crps
%                             reslist[[k]] <- res
%                           }
%                           
%                           #colnames(crps) <- names(res$crps)
%                           return(reslist)
%                         },
%                         mc.cores = detectCores() - 1
%   )
%   
%   # extract CRPS
%   crps_cross <- matrix(nrow = nrep_cross, ncol = 7)
%   # loop over all repetitions
%   for(i in 1:length(res_cross)){
%     #loop over all 7 folds (for 7 methods)
%     crps_cross_int <- matrix(nrow = length(res_cross[[1]]), ncol = 7)
%     for(j in 1:length(res_cross[[1]])){
%       crps_cross_int[j,] <- res_cross[[i]][[j]]$crps
%     }
%     crps_cross[i,] <- colMeans(crps_cross_int, na.rm = TRUE)
%   }
%   colnames(crps_cross) <- names(res_cross[[1]][[1]]$crps) 
%   
%   save(crps_cross, file = "Data/crps_cross.rda")
% }
% @
% 
% 
% 
% 
% \begin{frame}[fragile]
% \frametitle{Weather forecasting: precipitation}
% \begin{center}
% \vspace*{-0.2cm}
% \setkeys{Gin}{width=0.65\textwidth}
% <<rain_cross_axams_crps_skills_score, fig=TRUE, echo=FALSE, height=5.6, width=7>>=
% #par(mar = c(2.5,2,1,2))
% boxplot(1 - crps_cross[,c(2,3,4)] / crps_cross[,6], ylim = c(-0.005, 0.065),
%         names = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS"),
%         main = "Cross validation (with reference model EMOS)",
%         cex.main=1.4,
%         cex.lab=1.2,
%         ylab = "CRPS skill score", col = "lightgray") 
% abline(h = 0, col = pal["EMOS"], lwd = 2)
% @
% \end{center}
% \end{frame}

\begin{frame}[fragile]
\frametitle{Weather forecasting: precipitation}
%\vspace{0.2cm}
\textbf{Application for all 95 stations:}
\begin{itemize}
\item Learn forest model on data from 24 years (1985--2008).
\item Evaluate on 4 years (2009--2012).
\item Benchmark against other heteroscedastic censored Gaussian models.
\end{itemize}

\end{frame}

<<echo = FALSE>>=
#### prediction over all stations 24 - 4
if(file.exists("Data/crps_24to4_all.rda")){
  load("Data/crps_24to4_all.rda")
} else {

  data("StationsTyrol")
  stations <- StationsTyrol$name
  test <- 2009:2012
  train <- 1985:2008

  res_24to4_all <- mclapply(1:length(stations),
                            function(i){

                              set.seed(7)

                              res <- evalmodels(station = stations[i],
                                                train = train,
                                                test = test,
                                                gamboost_cvr = TRUE)

                              return(res)
                            },
                            mc.cores = detectCores() - 1
  )

  # extract crps
  crps_24to4_all <- matrix(nrow = length(stations), ncol = 7)
  # loop over all stations
  for(i in 1:length(stations)){
    crps_24to4_all[i,] <- res_24to4_all[[i]]$crps
  }

  colnames(crps_24to4_all) <- names(res_24to4_all[[1]]$crps)
  rownames(crps_24to4_all) <- stations

  save(crps_24to4_all, file = "Data/crps_24to4_all.rda")
}

# skill score
s <- 1 - crps_24to4_all[, 2:4]/crps_24to4_all[,6] 
colnames(s) <- c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS")

## prepare data for map which shows where distforest performed better than gamlss or gamboostLSS based on the crps

crps_map <- crps_24to4_all[,c("distforest", "gamlss", "gamboostLSS", "emos_log")]

# best method
bst <- apply(crps_map, 1, which.min)

# distance of forest to best other method
dst <- crps_map[,1] - crps_map[cbind(1:nrow(crps_map), apply(crps_map[, -1], 1, which.min) + 1)]

# breaks/groups
brk <- c(-0.1, -0.05, -0.005, 0.005, 0.05, 0.1)
#brk <- c(-0.1, -0.05, -0.01, 0.01, 0.05, 0.1)
grp <- cut(dst, breaks = brk)

# HCL colors (relatively flashy, essentially CARTO Tropic)
clr <- colorspace::diverging_hcl(5, h = c(130, 320), c = 70, l = c(50, 90), power = 1.3)

library("raster") # dem (digital elevation model)
library("sp")     # gadm www.gadm.org/country

data("StationsTyrol", package = "RainTyrol")
data("MapTyrol", package = "RainTyrol")
# data(MapTyrol_border, package = "RainTyrol")
# Create SpatialPointsDataFrame from station list
sp <- SpatialPointsDataFrame(subset(StationsTyrol,
                                    select=c(lon,lat)),
                             data = subset(StationsTyrol,
                                           select = -c(lon,lat)),
                             proj4string = crs(MapTyrol$RasterLayer))
@

\begin{frame}[fragile]
\frametitle{Weather forecasting: precipitation}
\begin{center}
\vspace*{-0.5cm}

\setkeys{Gin}{width=0.84\textwidth}
<<map, fig=TRUE, echo=FALSE, width=10, height=6.5>>=

  ## plot map of Tyrol with all 95 observations
  layout(cbind(1, 2), width = c(9, 1))
  par(mar = c(5,4,4,0.1))
  raster::image(MapTyrol$RasterLayer, col = rev(gray.colors(100)),
                main="Stations in Tyrol", ylab = "Latitude", xlab = "Longitude", 
                xlim = c(9.8,13.2), 
                ylim = c(46.6, 47.87))
  plot(MapTyrol$SpatialPolygons, add = TRUE)
  points(sp[70,], pch = 19, col = "black", cex = 1.85)
  points(sp, pch = c(21, 24, 25, 22)[bst], bg = clr[grp], col = "black", las = 1, cex = 1.5)
  legend(x = 9.8, y = 47.815, pch = c(21, 24, 25, 22), legend = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS", "EMOS"), cex = 1, bty = "n")
  text(x = 10.3, y = 47.82, labels = "Models with lowest CRPS")
  mtext("CRPS\ndifference", side=4, las = TRUE, at = c(x = 13.5, y = 47.76), line = 0.3)
  par(mar = c(0.5,0.2,0.5,2.3))
  ## legend
  plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
       xlim = c(0, 1), ylim = c(-0.2, 0.2), xaxs = "i", yaxs = "i")
  rect(0, brk[-6], 0.5, brk[-1], col = rev(clr))
  axis(4, at = brk, las = 1, mgp=c(0,-0.5,-1))

@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Weather forecasting: precipitation}
\begin{center}
\vspace*{-0.5cm}

\setkeys{Gin}{width=0.47\textwidth}
<<map, fig=TRUE, echo=FALSE, width=10, height=6.5>>=

  ## plot map of Tyrol with all 95 observations
  layout(cbind(1, 2), width = c(9, 1))
  par(mar = c(5,4,4,0.1))
  raster::image(MapTyrol$RasterLayer, col = rev(gray.colors(100)),
                main="Stations in Tyrol", ylab = "Latitude", xlab = "Longitude", 
                xlim = c(9.8,13.2), 
                ylim = c(46.6, 47.87))
  plot(MapTyrol$SpatialPolygons, add = TRUE)
  points(sp[70,], pch = 19, col = "black", cex = 1.85)
  points(sp, pch = c(21, 24, 25, 22)[bst], bg = clr[grp], col = "black", las = 1, cex = 1.5)
  legend(x = 9.8, y = 47.815, pch = c(21, 24, 25, 22), legend = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS", "EMOS"), cex = 1, bty = "n")
  text(x = 10.3, y = 47.82, labels = "Models with lowest CRPS")
  mtext("CRPS\ndifference", side=4, las = TRUE, at = c(x = 13.5, y = 47.76), line = 0.3)
  par(mar = c(0.5,0.2,0.5,2.3))
  ## legend
  plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
       xlim = c(0, 1), ylim = c(-0.2, 0.2), xaxs = "i", yaxs = "i")
  rect(0, brk[-6], 0.5, brk[-1], col = rev(clr))
  axis(4, at = brk, las = 1, mgp=c(0,-0.5,-1))

@
\end{center}
\vspace{-0.2cm}
\textbf{Conclusions:}
\begin{itemize}
\item Distributional forests provide an easy-to-apply alternative to already existing methods.
\item Do not require expert knowledge for pre-specifications.
\item Competitive performance: mainly en par or even slightly better.
\end{itemize}

\end{frame}

\subsection{Weather forecasting: wind direction}

\begin{frame}%[fragile]
\frametitle{}
\vspace{1.5cm}
\center
{\bf \large 
Circular Regression Trees and Forests with an \\
\smallskip
Application to Probabilistic Wind Direction Forecasting}

\bigskip
\bigskip

\small{
Moritz N. Lang, Lisa Schlosser, Torsten Hothorn, \\
Georg J. Mayr, Reto Stauffer, Achim Zeileis}

\bigskip
\bigskip

{\footnotesize
Minor revision in\\
\emph{Journal of the Royal Statistical Society C}

\medskip

Preprint version at \url{https://arxiv.org/abs/2001.00412}}

\end{frame}

\begin{frame}[fragile]
\frametitle{Weather forecasting: wind direction}

\textbf{Goal:} Nowcasting (1--3 hours ahead) of wind direction at Innsbruck Airport.

\bigskip
\pause

\only<2>{
\vspace{-0.4cm}
\setkeys{Gin}{width=0.7\linewidth}
\begin{center}
\includegraphics{ibk_airport.png}
\end{center}
}

\only<3>{
\vspace{-0.5cm}
\setkeys{Gin}{width=0.7\linewidth}
\begin{center}
\includegraphics{ibk_airport_east.png}
\end{center}

\vspace{-1.4cm}
Up-valley wind
}

\only<4>{
\vspace{-0.5cm}
\setkeys{Gin}{width=0.7\linewidth}
\begin{center}
\includegraphics{ibk_airport_west.png}
\end{center}

\vspace{-1.4cm}
Down-valley wind
}

\only<5>{
\vspace{-0.5cm}
\setkeys{Gin}{width=0.7\linewidth}
\begin{center}
\includegraphics{ibk_airport_south.png}
\end{center}

\vspace{-1.4cm}
Southerly wind
}

\only<6->{
\textbf{Challenges:}
\begin{itemize}
  \item Possibly abrupt changes due to geographical position.
  \item Circular response in $[0^{\circ}, 360^{\circ})$ with $0^{\circ} = 360^{\circ}$.
  \item NWP outputs are less useful due to short lead time.
\end{itemize}

\bigskip

\visible<7->{
\textbf{Inputs:} Observation data only (41,979 data points).
\begin{itemize}
  \item 4 stations at Innsbruck Airport, 6 nearby weather stations. 
  \item Base variables: Wind direction, wind (gust) speed, temperature, \\
  (reduced) air pressure, relative humidity.
  \item 260 covariates based on means/minima/maxima, temporal changes, spatial differences towards the airport.
\end{itemize}
}}
\end{frame}

\begin{frame}[fragile]
\frametitle{Weather forecasting: wind direction}

\vspace*{-0.3cm}

\begin{minipage}{0.5\textwidth}
\textbf{Distribution assumption:} Von Mises.
\begin{itemize}
\item Circular normal distribution.
\item Location parameter $\mu \in [0, 2 \pi)$.
\item Concentration parameter $\kappa > 0$.
\end{itemize}
\end{minipage}
\begin{minipage}{0.49\textwidth}
\only<1>{\includegraphics[width=\linewidth]{density_linear.pdf}}%
\only<2>{\includegraphics[width=\linewidth]{density_circular.pdf}}%
\end{minipage}

\bigskip
\medskip

\textbf{Log-likelihood:} $y \in [0, 2 \pi)$ and parameter vector $\theta = (\mu, \kappa)$.
$$
\ell(\theta; y) = \log \left\{ \frac{1}{2 \pi I_0(\kappa)}~e^{ \kappa \cos(y - \mu)} \right\}
$$
where $I_0(\kappa)$ is the modified Bessel function of the first kind and order $0$.

\end{frame}

\begin{frame}[fragile]
\frametitle{Weather forecasting: wind direction}

\vspace{-0.7cm}

\setkeys{Gin}{width=1\linewidth}
\begin{center}
\includegraphics{circtree_ibk.pdf}
\end{center}

\end{frame}

\begin{frame}[fragile]
\frametitle{Weather forecasting: wind direction}

\vspace{-0.54cm}

\setkeys{Gin}{width=0.47\linewidth}
\begin{figure}
\includegraphics[center]{circtree_ibk.pdf}
\end{figure}

%\vspace{-0.3cm}
\textbf{Conclusions:}
\begin{itemize}
\item Circular trees are an easy-to-apply method with high interpretability. 
\item A single circular tree provides reliable predictions which can be improved by employing a circular forest model.
\item Competitive performance: better than persistence and linear model.
\end{itemize}

\end{frame}

% \begin{frame}[fragile]
% \frametitle{Weather forecasting: wind direction}
% 
% \textbf{Benchmark:} Against other naive and circular models.
% \begin{itemize}
% \item Climatology: Without covariates.
% \item Persistency: Based on current wind direction.
% \item Circular GLM: Based on current wind speed and wind vectors $(u, v)$.
% \end{itemize}
% 
% \end{frame}
% 
% 
% \begin{frame}[fragile]
% \frametitle{Weather forecasting: wind direction}
% 
% %\includegraphics{circforest_crpsraw_ibk_lag1.pdf}
% %\includegraphics{circforest_crpsraw_ibk_lag3.pdf}
% 
% \textbf{Evaluation:} CRPS skill score for \only<1>{1}\only<2>{3}-hourly predictions (5-fold cross validation).
% \begin{center}
% \setkeys{Gin}{width=0.88\linewidth}
% \only<1>{\includegraphics{circforest_crpsskill_ibk_lag1.pdf}}%
% \only<2>{\includegraphics{circforest_crpsskill_ibk_lag3.pdf}}
% \end{center}
% 
% \end{frame}

%\subsection{Transformation models}
%
% \begin{frame}
% \frametitle{Transformation models}
% 
% \textbf{Alternative:} When no obvious classic distribution assumption is available.
% 
% \medskip
% 
% \textbf{Advantages:}
% \begin{itemize}
% \item Does not require specification of distribution family.
% \item More flexible framework.
% \end{itemize}
% 
% \medskip
% 
% \textbf{Distribution function:}
% \[
% F(y; \theta) = \Phi(\bold{a}_{Bs,d}(y)^\top \theta) 
% \]
% \begin{itemize} 
% \item $\bold{a}_{Bs,d}(y)^\top \theta$ is a smooth, monotone Bernstein polynomial of degree $d$.
% \item $d=1$ corresponds to $\mathcal{N}(\mu,\sigma^2)$.
% \item $d=5$ is surprisingly flexible.
% \end{itemize}
% 
% \medskip
% 
% \textbf{Example:} Body Mass Index explained by lifestyle factors (Switzerland).
% 
% \end{frame}
% 
% \begin{frame}
% \frametitle{Transformation models}
% 
% \vspace*{0.2cm}
% \includegraphics[width=0.84\textwidth]{cmpx-tree.pdf}
% 
% \end{frame}

\subsection{Software}

\begin{frame}
\frametitle{Software}
%\vspace{0.4cm}
\textbf{Software:} \emph{disttree} and \emph{circtree} 
available on R-Forge at\\

\medskip

\small{\url{https://R-Forge.R-project.org/projects/partykit/}}\\ 
%and\\
%\url{https://R-Forge.R-project.org/projects/uibk-rprog-2018/}}\\

\bigskip
\medskip

\textbf{Main functions:}

\medskip

\begin{tabular}{ll}
\code{distfit}    & Distributional fits (ML, \code{gamlss.family}/custom \code{list}).\\
& No covariates. \\
\code{disttree}   & Distributional trees (\code{ctree}/\code{mob} + \code{distfit}).\\
& Covariates as partitioning variables. \\
\code{distforest} & Distributional forests (ensemble of \code{disttree}s).\\
& Covariates as partitioning variables.
\end{tabular}

\bigskip
\medskip

\textbf{Correspondingly:} \code{circtree}, \code{circforest}

\end{frame}

\subsection{Papers}

\begin{frame}
\frametitle{Papers}

%\vspace{-0.3cm}

\footnotesize
Schlosser L, Hothorn T, Stauffer R, Zeileis A (2019).
\dquote{Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain.}
\emph{The Annals of Applied Statistics}, \textbf{13}(3), 1564--1589.
\doi{10.1214/19-AOAS1247}

\bigskip

Schlosser L, Hothorn T, Zeileis A (2019). 
\dquote{The Power of Unbiased Recursive Partitioning: A Unifying View of CTree, MOB, and GUIDE}
\emph{arXiv 1906.10179}, arXiv.org E-Print Archive.
\url{https://arxiv.org/abs/1906.10179}.
Submitted to \emph{Computational Statistics \& Data Analysis}.

\bigskip

Lang MN, Schlosser L, Hothorn T, Mayr GJ, Stauffer R, Zeileis A (2020).
\dquote{Circular Regression Trees and Forests with an Application to Probabilistic Wind Direction Forecasting.}
Minor revision in \emph{Journal of the Royal Statistical Society C}.
Preprint version at \url{https://arxiv.org/abs/2001.00412}

% % \medskip
% % 
% % Hothorn T, Zeileis A (2017).
% %   \dquote{Transformation Forests.}
% %   \emph{arXiv 1701.02110}, arXiv.org E-Print Archive.
% %   \url{http://arxiv.org/abs/1701.02110}
%   
% \bigskip
% \bigskip
% %\medskip
%   
% Hothorn T, Hornik K, Zeileis A (2006).
%  \dquote{Unbiased Recursive Partitioning: A Conditional Inference Framework.}
%  \emph{Journal of Computational and Graphical Statistics},
%  \textbf{15}(3), 651--674.
%  \doi{10.1198/106186006X133933}
%  
% \medskip
% 
% Zeileis A, Hothorn T, Hornik K (2008).
%  \dquote{Model-Based Recursive Partitioning.}
%   \emph{Journal of Computational and Graphical Statistics},
%   \textbf{17}(2), 492--514.
%   \doi{10.1198/106186008X319331}
% 
% \medskip
% 
% Hothorn T, Zeileis A (2015).
%  \dquote{{partykit}: A Modular Toolkit for Recursive Partytioning in \textsf{R}.}
%  \emph{Journal of Machine Learning Research},
%  \textbf{16}, 3905--3909.
%  \url{http://www.jmlr.org/papers/v16/hothorn15a}
% 
% 
% 
% %% \medskip
% %% 
% %% Stasinopoulos DM, Rigby RA (2007).
% %%   \dquote{Generalized Additive Models for Location Scale and Shape (GAMLSS) in \textsf{R}.}
% %%   \emph{Journal of Statistical Software}, 
% %%   \textbf{23}(7), 1--46.
% %%   \doi{10.18637/jss.v023.i07}
% %%   
% %% \medskip
% %% 
% %% Breiman L (2001).
% %%   \dquote{Random {Forests}}
% %%   \emph{Machine Learning}, 
% %%   \textbf{45}(1), 5--32.
% %%   \doi{10.1023/A:1010933404324}

\end{frame}

\end{document}