### 2020-06-29 pkg/partykit/inst/ULGcourse-2020/slides_forest.Rnw
material for introductory course on trees and forests
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@3018 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 388e124f82296b07a34c1e60b134613ce0d783e6

\documentclass[11pt,t,usepdftitle=false,aspectratio=169]{beamer}
\usetheme[nototalframenumber,url,license]{uibk}

\title{Random Forests}
\author{Achim Zeileis, Lisa Schlosser}
\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
%\URL{http://www.partykit.org/partykit/}
\renewcommand{\headerimage}[1]{%
   \IfStrEqCase{#1}{%
      {1}{%
         \gdef\myheaderimageid{#1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{forest.jpg}%
      }}[%
         \gdef\myheaderimageid{1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{forest.jpg}%
      ]%
}
\headerimage{1}

\usepackage[utf8]{inputenc}

\setbeamertemplate{caption}{\insertcaption} 
%% includes a replacement for \usepackage{Sweave}
%\usepackage{Sweave}
\usepackage{changepage}
\usepackage{amsmath,tikz}
\usepackage{calc}
\usepackage{graphicx}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap,trees,tikzmark,decorations.pathreplacing}
\usepackage{xcolor}
%\usepackage[cal=boondoxo]{mathalfa}
%\graphicspath{{plots/}}
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}

%% colors
\definecolor{HighlightOrange}{rgb}{0.9490196,0.5725490,0.0000000}
\definecolor{HighlightBlue}{rgb}{0.4784314,0.7490196,0.9803922}
\definecolor{forestred}{RGB}{206,73,81}
\definecolor{treegreen}{RGB}{0,143,0}
\definecolor{lightblue}{RGB}{34,151,230}
\definecolor{lightorange}{RGB}{255,165,0}

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)

set.seed(7)

invisible(.Call(grDevices:::C_palette, grDevices::hcl(
  h = c(0,   5, 125, 245, 195, 315,  65,   0),
  c = c(0, 100,  90,  85,  63, 105,  90,   0),
  l = c(0,  55,  75,  60,  82,  48,  80,  65)
)))

library("disttree")
library("Formula")
library("latex2exp")
library("lattice")
library("MASS")
library(colorspace)

## HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)
names(pal) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")

pallight <- hcl(c(10, 128, 260, 290, 70), 100, 50, alpha = 0.25)
names(pallight) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")

transpgray <- rgb(0.190,0.190,0.190, alpha = 0.2)

lightblue <- "#2297E6"
lightorange <- "#FFA500"
@

<<motivation_tree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
nobs <- 200
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- numeric(length = length(x))
for(i in 1:(nobs)) ytrue[i] <- if(x[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))
y <- ytrue + rnorm(nobs,0,0.3)
# more points for precise illustration of true function
x100 <- c(1:(100*nobs))/(100*nobs)
ytrue <- ytree <- ytree2 <- ytree3 <- ytree4 <- yforest <- numeric(length = length(x100))
for(i in 1:(nobs*100)) ytrue[i] <- if(x100[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x100[i]-0.2)-1)))
for(i in 1:(nobs*100)) ytree[i] <- if(x100[i]<1/3) 0.5 else {if(x100[i]<2/3) 2 else 1}
for(i in 1:(nobs*100)) ytree2[i] <- if(x100[i]<0.31) 0.47 else {if(x100[i]<0.69) 1.9 else 1.1}
for(i in 1:(nobs*100)) ytree3[i] <- if(x100[i]<0.34) 0.52 else {if(x100[i]<0.54) 2.2 else {if(x100[i]<0.74) 1.74 else 0.9}}
for(i in 1:(nobs*100)) ytree4[i] <- if(x100[i]<0.32) 0.51 else {if(x100[i]<0.63) 1.8 else 1.15}
for(i in 1:(nobs*100)) yforest[i] <- if(x100[i]<=0.33) 0.52 else {if(x100[i]<=0.34) 1.5 else {if(x100[i]<=0.54) 2 else {if(x100[i]<=0.63) 1.9 else {if(x100[i]<=0.69) 1.6 else {if(x100[i]<=0.74) 1.3 else 1}}}}}
@

\begin{document}

\section{Forests models}

\subsection{Motivation: Forest models}
\begin{frame}[fragile]
\frametitle{Motivation: Forest models}

{\bf Idea:} \\
\begin{itemize}
\item Combination of an ensemble of trees. 
\item Each tree can capture non-linear and non-additive effects and select covariates and possible interactions automatically.
\item Combining trees to a forest model allows for an approximation of smooth effects.
\item A forest can regularize and stabilize the model.
\end{itemize}

%\bigskip

\end{frame}

\begin{frame}
\frametitle{Motivation: Forest models}

\vspace{0.2cm}

\begin{minipage}{0.4\textwidth}
\only<1>{\textbf{Tree model:\\}}\only<2-8>{\textbf{Subsampling:\\}}\only<9->{\textbf{Forest model:\\}}
\vspace{-0.4cm}
%\vspace{-0.8cm}
\begin{center}
\only<1>{
%\vspace{0.2cm}
\resizebox{0.63\textwidth}{!}{
\begin{tikzpicture}
  \node[ellipse, fill=HighlightBlue!70, align=center] (n0) at (2, 3) {};
  \node[] (n00) at (1, 1.5) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n1) at (0, 0) {$\hat{Y}_1$};
  \draw[-] (n0) -- (n00) node [midway, left] {\small $X \leq p_1$};
  \draw[-] (n0) -- (n1);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n2) at (3, 1.5) {};
  \draw[-] (n0) -- (n2) node [midway, right] {\small $X > p_1$};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n3) at (2, 0) {$\hat{Y}_2$};
  \draw[-] (n2) -- (n3) node [midway, left] {\small $X \leq p_2$};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n4) at (4, 0) {$\hat{Y}_3$};
  \draw[-] (n2) -- (n4) node [midway, right] {\small $X > p_2$};
\end{tikzpicture}}

\vspace{0.475cm}

}
\only<2->{
%\vspace{0.2cm}
\resizebox{0.6\textwidth}{!}{
\begin{tikzpicture}
%\draw[ellipse, draw=black, align=center, scale = 0.7, minimum width=170pt, minimum height = 35pt] (set) at (1, 3) {};
\draw[gray,rounded corners=10pt] (-1.05,4.35) -- (4.05,4.35) -- (4.05,2.7) -- (-2.05,2.7) -- (-2.05,4.35) -- (0.05,4.35);
%\draw[gray] (1,3.5) ellipse (3 and 0.8);
%\visible<4>{
\begin{scope}
\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\clip (1,3.5) ellipse (3 and 0.8);
\pgfmathsetseed{7}
\foreach \p in {1,...,200} {\fill[gray] (1+3*rand,3.5+0.8*rand) circle (0.04);}
\end{scope}
%}
%%% color points
%\visible<5-6>{
%\begin{scope}
%\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\pgfmathsetseed{7}
%\foreach \p in {1,...,100} {\fill[treegreen] (1+3*rand,3.5+0.8*rand) circle (0.07);}
%\end{scope}
%}
\visible<3->{
\node[ellipse, draw=treegreen, align=center, minimum width=80pt, minimum height = 40pt, line width = 3pt] (subset1) at (0.1, 3.5) {};
}
%\visible<7-8>{
%\begin{scope}
%\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\pgfmathsetseed{7}
%\foreach \p in {1,...,49} {\fill[gray] (1+3*rand,3.5+0.8*rand) circle (0.04);}
%\foreach \p in {50,...,150} {\fill[lightblue] (1+3*rand,3.5+0.8*rand) circle (0.07);}
%\end{scope}
%}
\visible<5->{
\node[ellipse, draw = lightblue, align=center, minimum width=100pt, minimum height = 30pt, line width = 3pt] (subset2) at (1.2, 3.6) {};
}
%\visible<9-10>{
%\begin{scope}
%\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\pgfmathsetseed{7}
%\foreach \p in {1,...,99} {\fill[gray] (1+3*rand,3.5+0.8*rand) circle (0.04);}
%\foreach \p in {100,...,200} {\fill[lightorange] (1+3*rand,3.5+0.8*rand) circle (0.07);}
%\end{scope}
%}
\visible<7->{
\node[ellipse, draw = lightorange, align=center, minimum width=120pt, minimum height = 25pt, line width = 3pt] (subset3) at (1.6, 3.3) {};
}
\visible<4->{
\node[ellipse, fill=treegreen, align=center] (n00) at (-1, 2) {};
\node[ellipse, fill=treegreen, align=center] (n01) at (-1.25, 1.25) {};
\draw[-, line width=1pt] (n00) -- (n01);
\node[rectangle, fill=treegreen, align=center] (n02) at (-1.5, 0.5) {};
\draw[-, line width=1pt] (n01) -- (n02);
\node[rectangle, fill=treegreen, align=center] (n03) at (-1, 0.5) {};
\draw[-, line width=1pt] (n01) -- (n03);
\node[rectangle, fill=treegreen, align=center] (n04) at (-0.5, 0.5) {};
\draw[-, line width=1pt] (n00) -- (n04);
}
\visible<6->{
\node[ellipse, fill=lightblue, align=center] (n10) at (1, 2) {};
\node[ellipse, fill=lightblue, align=center] (n11) at (0.5, 1.25) {};
\draw[-, line width=1pt] (n10) -- (n11);
\node[rectangle, fill=lightblue, align=center] (n12) at (0.25, 0.5) {};
\draw[-, line width=1pt] (n11) -- (n12);
\node[rectangle, fill=lightblue, align=center] (n13) at (0.75, 0.5) {};
\draw[-, line width=1pt] (n11) -- (n13);
\node[ellipse, fill=lightblue, align=center] (n14) at (1.5, 1.25) {};
\draw[-, line width=1pt] (n10) -- (n14);
\node[rectangle, fill=lightblue, align=center] (n15) at (1.75, 0.5) {};
\draw[-, line width=1pt] (n14) -- (n15);
\node[rectangle, fill=lightblue, align=center] (n16) at (1.25, 0.5) {};
\draw[-, line width=1pt] (n14) -- (n16);
}
\visible<8->{
\node[ellipse, fill=lightorange, align=center] (n20) at (3, 2) {};
\node[rectangle, fill=lightorange, align=center] (n21) at (2.5, 0.5) {};
\draw[-, line width=1pt] (n20) -- (n21);
\node[ellipse, fill=lightorange, align=center] (n22) at (3.25, 1.25) {};
\draw[-, line width=1pt] (n20) -- (n22);
\node[rectangle, fill=lightorange, align=center] (n23) at (3, 0.5) {};
\draw[-, line width=1pt] (n22) -- (n23);
\node[rectangle, fill=lightorange, align=center] (n24) at (3.5, 0.5) {};
\draw[-, line width=1pt] (n22) -- (n24);
}
\visible<10->{
\node[ellipse, draw = black, align=center, minimum width=15pt, minimum height = 15pt, line width = 2pt] (B1) at (-1.5, 0.5) {};
\node[ellipse, align=center, line width = 3pt, text=treegreen] (Y1) at (-1.5, -0.5) {\LARGE $\hat{y}_1$};
}
\visible<11->{
\node[ellipse, draw = black, align=center, minimum width=15pt, minimum height = 15pt, line width = 2pt] (B2) at (1.25, 0.5) {};
\node[ellipse, align=center, line width = 3pt, text=lightblue] (Y2) at (1.25, -0.5) {\LARGE $\hat{y}_2$};
}
\visible<12->{
\node[ellipse, draw = black, align=center, minimum width=15pt, minimum height = 15pt, line width = 2pt] (B2) at (3.5, 0.5) {};
\node[ellipse, align=center, line width = 3pt, text=lightorange] (Y3) at (3.5, -0.5) {\LARGE $\hat{y}_3$};(3.5, 0.5)
}
\end{tikzpicture}
}
}
\end{center}
\end{minipage}
\hspace{0.7cm}
\begin{minipage}{0.36\textwidth}
%\begin{center}
%\vspace{0.1cm}
\only<1-2>{
%\vspace{0.01cm}
<<plot_motivation_tree, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,0))
#par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x100, y = ytree, col = pal["forest"], lwd=7)
mtext(text = "X", side = 1, cex = 2.5, line = 2)  
mtext(text = "Y", side = 2, cex = 2.5, line = 1)  
@
}
\only<3>{
%\vspace{0.01cm}
%\hspace{-0.1cm}
<<plot_motivation_randforest_data, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,0))
#par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
mtext(text = "X", side = 1, cex = 2.5, line = 2)  
mtext(text = "Y", side = 2, cex = 2.5, line = 1)  
@
}
\only<4-5>{
%\vspace{0.01cm}
%\hspace{-0.1cm}
<<plot_motivation_randforest1, fig=TRUE, echo=FALSE, width=7>>=
<<plot_motivation_randforest_data>>
lines(x = x100, y = ytree2, col = pal["tree"], lwd=7)
@
}
\only<6-7>{
%\vspace{0.01cm}
%\hspace{-0.1cm}
<<plot_motivation_randforest2, fig=TRUE, echo=FALSE, width=7>>=
<<plot_motivation_randforest1>>
lines(x = x100, y = ytree3, col = lightblue, lwd=7)
@
}
\only<8>{
%\vspace{0.01cm}
%\hspace{-0.1cm}
<<plot_motivation_randforest3, fig=TRUE, echo=FALSE, width=7>>=
<<plot_motivation_randforest2>>
lines(x = x100, y = ytree4, col = lightorange, lwd=7)
@
}
\only<9-12>{
<<plot_motivation_randforest_x, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,0))
#par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x100, y = ytree2, col = pal["tree"], lwd=7)
lines(x = x100, y = ytree3, col = lightblue, lwd=7)
lines(x = x100, y = ytree4, col = lightorange, lwd=7)
lines(x = c(0.6,0.6), y = c(-1, 1.8), type = "l", lty = 2, lwd = 4)
mtext(text = "Y", side = 2, cex = 2.5, line = 1)  
mtext(text = "x", side = 1, cex = 4, line = 2, at=0.6)  
@
}
\only<13->{
<<plot_motivation_randforest4, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x100, y = yforest, col = pal["forest"], lwd=7)
lines(x = c(0.6,0.6), y = c(-1, 1.8), type = "l", lty = 2, lwd = 4)
mtext(text = "x", side = 1, cex = 4, line = 2, at=0.6)  
mtext(text = "Y", side = 2, cex = 2.5, line = 1)  
@
}
\end{minipage}

\vspace{0.5cm}

\only<13->{
$\hat{y}= \frac{1}{3} (\hat{y}_1 + \hat{y}_2 + \hat{y}_3)$
}

\vspace{0.3cm}

\visible<13->{
for $\hat{y}_t$ being the average response value in the corresponding segment $\mathcal{B}^t_x$ of the $t$-th tree for learning data $\{y_i\}_{i=1,\ldots,n}$:

\vspace{0.3cm}

$\hat{y}_t= \frac{1}{|\mathcal{B}^t_x|} \sum_{i: y_i \in \mathcal{B}^t_x} y_i$
\visible<14->{
$\ =\ \sum_{i=1}^n \frac{I(x_i \in \mathcal{B}^t_x)}{|\mathcal{B}^t_x|} \cdot y_i$
}
\visible<15->{
$\ =\ \sum_{i=1}^n w_i(x) \cdot y_i$
}
}
\end{frame}

\subsection{Aggregation of trees}

\begin{frame}
\frametitle{Aggregation of trees}
{\bf Strategy:} Combine trees to a forest model by employing tree-based weights.

The predicted response $\hat{y}$ or, more generally, the predicted model parameter (vector) $\hat{\theta}$ is the argument of the parameter space $\Theta$ that minimizes the weighted sum of the objective function $f$ evaluated for learning data $\{y_i\}_{i=1,\ldots,n}$:

\vspace{-0.2cm}

\[
\hat{\theta}(x) =  \argmin_{\theta \in \Theta} \sum_{i=1}^n w_i(x) \cdot f(\theta; y_i)
\]

\medskip

\only<2-4>{

\vspace{0.13cm}

\textbf{Weights:}

\vspace{-1.14cm}

\begin{eqnarray*}
w^{\text{base}}_i(x)   & = & 1 \\[0.2cm]
\visible<3-4>{
w^{\text{tree}}_i(x)   & = & \sum_{b=1}^B I((x_i \in \mathcal{B}_b) \land (x \in \mathcal{B}_b)) \\[0.1cm]
\visible<4>{
w^{\text{forest}}_i(x) & = & \frac{1}{T} \sum_{t=1}^T \sum_{b=1}^{B^t} \frac{I((x_i \in \mathcal{B}^t_b) \land (x \in \mathcal{B}^t_b))}{\left|\mathcal{B}^t_b\right|}
\end{eqnarray*}
}}}

\only<5->{
%\vspace{-3cm}
{\bf Note:} This general notation holds for classic random forests averaging over mean values of tree nodes by setting $f(\theta, y_i) = (\theta - y_i)^2$\\
but also allows for combining quantiles (quantile regression forests) or full distributions as in distributional forests with
$f(\theta, y_i) = log(likelihood(\theta, y_i))$.\\
}

\end{frame}

\subsection{Subsampling}

\begin{frame}
\frametitle{Subsampling}
{\bf Idea:} Build each tree on a different subset of the full data set.

\medskip

{\bf Types of subsampling:}
\begin{itemize}
\item Subsets (recommended size: 63.2\% of the full data set)
\item Bootstrapping (resampling with replacement)
\end{itemize}

\bigskip

{\bf Random input variable sampling:} In order to reduce the correlation among trees only a subset of possible split variables is provided for each split (recommended size: $\sqrt{\text{\#Var}}$).

\bigskip

{\bf Note:} Each tree is built as large as possible which can lead to overfitting for the single tree. However, this is compensated by aggregating the trees to a forest model.

\end{frame}

\subsection{Software}

\begin{frame}[fragile]
\frametitle{R software for forest models}

{\bf randomForest:}
\begin{itemize}
  \item R package for classic random forests.
  \item Based on implementations of CART.
\end{itemize}

\medskip

<<call_randomForest, eval=FALSE, echo=TRUE>>=
randomForest(formula, data=NULL, ..., subset, na.action=na.fail)
randomForest(x, y=NULL,  ...)
@

\medskip

Optional control arguments:
\begin{itemize}
\item \code{ntree}: nr. of trees, 
\item \code{mtry}: nr. of randomly sampled split variable candidates, 
\item \code{replace}: type of subsetting, 
\item \code{sampsize}: size of drawn subsamples, 
\item \code{nodesize}: minimum size of terminal nodes, 
\item \code{maxnodes}: maximum number of terminal nodes, 
\item \ldots 
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{R software for forest models}

{\bf partykit:}
\begin{itemize}
  \item R package offering a toolkit for unbiased recursive partitioning. 
  \item Provides the forest-building function \fct{cforest}, based on conditional inference trees.
\end{itemize}

\medskip

<<call_cforest, eval = FALSE>>=
cforest(formula, data, subset, ...)
@

\medskip

\only<1>{
Optional forest-specific control arguments:
\begin{itemize}
\item \code{ytrafo}: transformation function applied to the response,
\item \code{ntree}: nr. of trees, 
\item \code{mtry}: nr. of randomly sampled split variable candidates, 
\item \code{perturb}: type and size of subsetting, 
\item \ldots 
\end{itemize}
\vspace{0.9cm}
}
\only<2>{
Optional tree-specific arguments handed over via \fct{ctree\_control}:
\begin{itemize}
\item \code{minsplit}: minimum nr. of observations to perform a split, 
\item \code{minbucket}: minimum nr. of observations in a node after splitting, 
\item \code{maxdepth}: maximum depth of the tree, 
\item \code{alpha}: significance level for split variable selection,
\item \ldots 
\end{itemize}
\vspace{0.9cm}
}
\end{frame}

\begin{frame}[fragile]
\frametitle{R software for forest models}

{\bf Predictions:} For a (possibly new) set of observations the method \fct{predict} can be employed to obtain the corresponding predicted values of the response variable.

\bigskip

<<call_predict, eval=FALSE, echo=TRUE>>=
predict(object, newdata, type = "response", ...)
@

\end{frame}

\begin{frame}
\frametitle{R software for forest models}

\textbf{Further R packages:}
%
\begin{itemize}
  \item ranger (classic random forests),
  \item grf (generalized random forests),
  \item disttree (distributional forests, based on partykit),
  \item \ldots
\end{itemize}

\end{frame}

\subsection{Variable importance}

\begin{frame}
\frametitle{Variable importance}
{\bf Idea:} Asses the influence of a covariate by evaluating the change in performance of the model (mean decrease in accuracy) after dropping or permutating the selected covariate.

\bigskip

Depending on the fitted model the corresponding objective function can be employed as \emph{risk function} to measure accuracy by calculating importance scores (e.g., log-likelihood, nr. of misclassifications, \ldots).

\bigskip

{\bf Implementations:}
\begin{itemize}
\item \fct{importance} in the R package randomForest.
\item \fct{importance.ranger} in the R package ranger.
\item \fct{varimp} in the R package partykit.
\end{itemize}

\end{frame}

\subsection{References}

\begin{frame}
\frametitle{References}

\vspace{-0.2cm}

\scriptsize

Breiman L (2001).
  \dquote{Random {Forests}.}
  \emph{Machine Learning}, 
  \textbf{45}(1), 5--32.
  \doi{10.1023/A:1010933404324}

\medskip

Breiman L, Cutler A (2004).
 \dquote{Random Forests.},
 \url{https://www.stat.berkeley.edu/~breiman/RandomForests}
 (Accessed: 2018-02-22)

\medskip

%Hothorn T, Zeileis A (2017).
%  \dquote{Transformation Forests.}
%  \emph{arXiv 1701.02110}, arXiv.org E-Print Archive.
%  \url{http://arxiv.org/abs/1701.02110}
%  
%\smallskip

Hothorn T, Hornik K, Zeileis A (2006).
 \dquote{Unbiased Recursive Partitioning: A Conditional Inference Framework.}
 \emph{Journal of Computational and Graphical Statistics},
 \textbf{15}(3), 651--674.
 \doi{10.1198/106186006X133933}

\medskip

Hothorn T, Zeileis A (2015).
 \dquote{{partykit}: A Modular Toolkit for Recursive Partytioning in \textsf{R}.}
 \emph{Journal of Machine Learning Research},
 \textbf{16}, 3905--3909.
 \url{http://www.jmlr.org/papers/v16/hothorn15a.html}

\medskip

Liaw A, Wiener M (2002).
 \dquote{Classification and Regression by randomForest.}
 \emph{R News},
 \textbf{2}(3), 18--22.
 \url{https://CRAN.R-project.org/doc/Rnews/}

\medskip

% Zeileis A, Hothorn T, Hornik K (2008).
%  \dquote{Model-Based Recursive Partitioning.}
%   \emph{Journal of Computational and Graphical Statistics},
%   \textbf{17}(2), 492--514.
%   \doi{10.1198/106186008X319331}
% 
% \medskip

Athey S, Tibshirani J, Wager S (2019). 
 \dquote{Generalized Random Forests.} 
 \emph{Annals of Statistics}, \textbf{47}{2}, 1148--1178.
 \doi{10.1214/18-AOS1709"}

\medskip

Schlosser L, Hothorn T, Stauffer R, Zeileis A (2019).
\dquote{Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain.}
\emph{The Annals of Applied Statistics}, \textbf{13}(3), 1564--1589.
\doi{10.1214/19-AOAS1247}

\medskip

Wright M N, Ziegler A (2017).
\dquote{{ranger}: A Fast Implementation of Random Forests for High
Dimensional Data in {C++} and {R}.}
\emph{Journal of Statistical Software}, \textbf{77}{1}, 1--17.
\doi{10.18637/jss.v077.i01}

\end{frame}

\end{document}