### 2020-07-01 pkg/partykit/inst/ULGcourse-2020/slides_tree.Rnw
finished revision of tree slides
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@3028 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 937f161b6b4c726f83e9e0db39c1fcf3beaf8185

\documentclass[11pt,t,usepdftitle=false,aspectratio=169]{beamer}
\usetheme[nototalframenumber,license]{uibk}

\title{Classification and Regression Trees and Beyond}
\subtitle{Supervised Learning: Algorithmic Modeling}
\author{Lisa Schlosser, Achim Zeileis}

%% forest header image
\renewcommand{\headerimage}[1]{%
   \IfStrEqCase{#1}{%
      {1}{%
         \gdef\myheaderimageid{#1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{forest.jpg}%
      }}[%
         \gdef\myheaderimageid{1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{forest.jpg}%
      ]%
}
\headerimage{1}

%% custom subsection slides
\setbeamercolor*{subsectionfade}{use={normal text},parent={normal text},fg=structure.fg!30!normal text.bg}
\AtBeginSubsection[]{%
  \begin{frame}[c]
    \begin{center}
      \usebeamercolor[fg]{subsectionfade}
      \Large \insertsection \\[2ex]
      \usebeamercolor[fg]{structure}
      \huge\bfseries\insertsubsection
    \end{center}
  \end{frame}
}

%% for \usepackage{Sweave}
\SweaveOpts{engine=R, eps=FALSE, echo=FALSE, results=hide, keep.source=TRUE}

<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)

library("rpart")
library("partykit")
library("coin")

set.seed(7)
@

<<data>>=
data("CPS1985", package = "AER")
data("Titanic", package = "datasets")
ttnc <- as.data.frame(Titanic)
ttnc <- ttnc[rep(1:nrow(ttnc), ttnc$Freq), 1:4]
names(ttnc)[2] <- "Gender"
@

\begin{document}

\section{Classification and regression trees and beyond}

\subsection{Motivation}
%\subsectionpage

\begin{frame}[fragile]
\frametitle{Motivation}

\smallskip

\textbf{Idea:} ``Divide and conquer.''
\begin{itemize}
  \item \emph{Goal:} Split the data into small(er) and (rather) homogenous subgroups.
  \item \emph{Inputs:} Explanatory variables (or covariates/regressors) used for splitting.
  \item \emph{Output:} Prediction for dependent (or target) variable(s).
\end{itemize}

\bigskip
\pause

\textbf{Formally:}
\begin{itemize}
  \item Dependent variable $Y$ (possibly multivariate).
  \item Based on explanatory variables $X_1, \dots, X_m$.
  \item ``Learn'' subgroups of data by combining splits in $X_1, \dots, X_m$.
  \item Predict $Y$ with (simple) model in the subgroups, often simply the mean.
\end{itemize}

\bigskip
\pause

\textbf{Key features:}
\begin{itemize}
  \item Predictive power in nonlinear regression relationships.
  \item Interpretability (enhanced by tree visualization), i.e., no ``black box''.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}

\textbf{Example:} Survival (yes/no) on the Titanic. Women and children first?

\begin{center}
\setkeys{Gin}{width=0.9\textwidth}
<<ctree_Titanic, echo=FALSE, fig=TRUE, width=11>>=
ct_ttnc <- ctree(Survived ~ Gender + Age + Class, data = ttnc, alpha = 0.01)
plot(ct_ttnc)
@
\end{center}

\end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}

\textbf{Model formula:} \code{Survived ~ Gender + Age + Class}.

\bigskip

\textbf{Data:} Information on the survival of 2201~passengers on board the
ill-fated maiden voyage of the RMS~Titanic in 1912. A data frame containing
\Sexpr{nrow(ttnc)} observations on \Sexpr{ncol(ttnc)} variables.

\bigskip

\begin{tabular}{ll}
\hline 
Variable        & Description\\
\hline 
\code{Class}    & Factor: 1st, 2nd, 3rd, or Crew.\\
\code{Gender}   & Factor: Male, Female.\\
\code{Age}      & Factor: Child, Adult.\\
\code{Survived} & Factor: No, Yes.\\
\hline 
\end{tabular}

\end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}

\textbf{Example:} Determinants of wages and returns to education.

\begin{center}
\setkeys{Gin}{width=0.99\textwidth}
<<ctree_Wages1985, echo=FALSE, fig=TRUE, width=13>>=
ct_cps <- ctree(log(wage) ~ education + experience + age + ethnicity + gender + union, data = CPS1985, alpha = 0.01)
plot(ct_cps)
@
\end{center}

\end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}

\textbf{Model formula:} \code{log(wage) ~ education + experience + age + ethnicity + gender + union}.

\bigskip

\textbf{Data:} Random sample from the May 1985 US Current Population Survey.
A data frame containing \Sexpr{nrow(CPS1985)} observations on \Sexpr{ncol(CPS1985)} variables.

\bigskip

\begin{tabular}{ll}
\hline 
Variable & Description\\
\hline 
\code{wage}       & Wage (in US dollars per hour). \\
\code{education}  & Education (in years).  \\
\code{experience} & Potential work experience (in years, \code{age - education - 6}). \\
\code{age}        & Age (in years). \\
\code{ethnicity}  & Factor: Caucasian, Hispanic, Other. \\
\code{gender}     & Factor: Male, Female. \\
\code{union}      & Factor. Does the individual work on a union job? \\
\hline 
\end{tabular}

\end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}

\textbf{Alternatively:} Linear regression tree.

\begin{center}
\setkeys{Gin}{width=0.9\textwidth}
<<lmtree_Wages1985, echo=FALSE, fig=TRUE, width=11>>=
mob_cps <- lmtree(log(wage) ~ education | experience + age + ethnicity + gender + union, data = CPS1985)
plot(mob_cps)
@
\end{center}

\end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}

\textbf{Model formula:} \code{log(wage) ~ education | experience + age + ethnicity + gender + union}.

\bigskip

\textbf{Model fit:}
\begin{itemize}
  \item Not just a group-specific mean.
  \item But group-specific intercept and slope (i.e., returns to education).
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}

\textbf{Example:} Nowcasting (1--3 hours ahead) of wind direction at Innsbruck Airport.

\setkeys{Gin}{width=0.87\linewidth}
\begin{center}
\includegraphics{circtree_ibk.pdf}
\end{center}

\end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}

\textbf{Data:} 41,979 data points for various weather observations.
\begin{itemize}
  \item Dependent variable: Wind direction 1--3 hours ahead.
  \item Explanatory variables: Current weather observations including
    wind direction, wind speed, temperature, (reduced) air pressure, relative humidity.
  \item Circular response in $[0^{\circ}, 360^{\circ})$ with $0^{\circ} = 360^{\circ}$.
\end{itemize}

\bigskip

\textbf{Model fit:} Circular distribution (von Mises), fitted by maximum likelihood.

\end{frame}

% \begin{frame}[fragile]
% \frametitle{Motivation}
% {\small Wage data:  \quad $log(wage) \; \sim  \; education + experience + ethnicity + region + parttime$}
% \begin{center}
% \setkeys{Gin}{width=0.99\textwidth}
% <<ctree_Wages1988, echo=FALSE, fig=TRUE, width=13>>=
% data("CPS1988", package = "AER")
% ## ctree
% plot(ctree(log(wage)~education+experience+ethnicity+region+parttime,data=CPS1988,
%            control = ctree_control(alpha = 0.01, minsplit = 7000)))
% @
% \end{center}
% \end{frame}

% \begin{frame}[fragile]
% \frametitle{Motivation}
% {\small Boston housing:  \quad $medv \; \sim  \; lstat + crim + rm + age + black$}
% \begin{center}
% \setkeys{Gin}{width=0.99\textwidth}
% <<ctree_Boston, echo=FALSE, fig=TRUE, width=13>>=
% data("Boston")
% plot(ctree(medv~lstat + crim + rm + age + black, 
%            data = Boston, control = ctree_control(alpha = 0.01, minsplit = 150)))
% @
% \end{center}
% \end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}

\textbf{Original idea:} Trees are purely algorithmic models without assumptions.
\begin{itemize}
  \item Data-driven ``learning'' of homogenous subgroups.
  \item Simple constant fit in each group, e.g., an average or proportion.
\end{itemize}

\bigskip
\pause

\textbf{Subsequently:} Trees are well-suited for combination with classical models.
\begin{itemize}
  \item Group-wise models, e.g., fitted by least squares or maximum likelihood.
  \item Model-based learning accounts for model differences across subgroups.
\end{itemize}

\bigskip
\pause

\textbf{Trade-off:} 
\begin{itemize}
  \item Assume simple model and learn larger tree \emph{vs.}
  \item More complex model and potentially smaller tree.
\end{itemize}

\end{frame}

\subsection{Tree algorithm}

\begin{frame}[fragile]
\frametitle{Tree algorithm}

\textbf{Base algorithm:}
\begin{enumerate}
\item Fit a model to the response $Y$.
\item Assess association of $Y$ (or a corresponding transformation/goodness-of-fit measure) and each possible split variable $X_j$.
\item Split sample along the $X_{j^{\ast}}$ with strongest association: Choose split point with highest improvement of the model fit.
\item Repeat steps 1--3 recursively in the subsamples until some stopping criterion is met.
\item \emph{Optionally:} Reduce size of the tree by pruning branches of splits that do not improve the model fit sufficiently.
\end{enumerate}

\end{frame}

\begin{frame}[fragile]
\frametitle{Tree algorithm}

\textbf{Specific algorithms:} Many (albeit not all) can be derived from the base algorithm
by combining suitable building blocks.

\bigskip
\pause

\textbf{Models for $Y$:} Simple constant fits vs.\ more complex statistical models.

\bigskip
\pause

\textbf{Goodness of fit:} Suitable measure depends on type of response variable(s) $Y$
and the corresponding model.

\end{frame}

\begin{frame}[fragile]
\frametitle{Tree algorithm}

\textbf{Goodness of fit:}
\begin{itemize}
  \item Numeric response: 
  \begin{itemize}
    \item $Y$ or ranks of $Y$.
    \item (Absolute) deviations $Y - \bar Y$.
    \item Residual sum of squares $\sum (Y - \hat{Y})^2$.
  \end{itemize}
  \pause
  \item Categorical response: 
  \begin{itemize}
    \item Dummy variables for categories.
    \item Number of misclassifications.
    \item Gini impurity.
  \end{itemize}
  \pause
  \item Survival response:
  \begin{itemize}
    \item Log-rank scores.
  \end{itemize}
  \pause
  \item Parametric model:
  \begin{itemize}
    \item Residuals.
    \item Model scores (gradient contributions).
  \end{itemize}
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Tree algorithm}

\textbf{Split variable selection:} Optimize some criterion over all $X_j$ ($j = 1, \dots, m$).
\begin{itemize}
  \item Objective function (residual sum of squares, log-likelihood, misclassification rate, impurity, \dots).
  \item Test statistic or corresponding $p$-value.
\end{itemize}

\bigskip
\pause

\textbf{Split point selection:} Optimize some criterion over all (binary) splits in $X_{j^*}$.
\begin{itemize}
  \item Objective function.
  \item Two-sample test statistic or corresponding $p$-value.
\end{itemize}

\bigskip
\pause

\textbf{Stopping criteria:}
\begin{itemize}
  \item Constraints: Number of observations per node, tree depth, \dots
  \item Lack of improvement: Significance, information criteria, \dots
\end{itemize}

\end{frame}

\subsection{Split variable selection}

\begin{frame}[fragile]
\frametitle{Split variable selection}

\textbf{Idea:}
\begin{itemize}
  \item Select variable $X_j$ ($j = 1, \dots, m$) most associated with heterogeneity in $Y$.
  \item Heterogeneity captured by goodness-of-fit measure.
  \item \emph{Often:} Maximum association over all possible binary splits.
  \item \emph{Alternatively:} Overall association.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Split variable selection}

\textbf{Potential bias:} Variables with many potential splits may yield greater
association ``by chance'', e.g., continuous $X_j$ or categorical with many levels.

\bigskip
\pause

\textbf{Unbiased recursive partitioning:} Accounts for potential random variation by
employing $p$-values from appropriate statistical tests.

\bigskip
\pause

\textbf{Possible tests:} Depend on scales of $Y$, $X_j$, and the adopted model.
\begin{itemize}
  \item $\chi^2$ test, Pearson correlation test, two-sample t-test, ANOVA.
  \item Maximally-selected two-sample tests.
  \item Parameter instability tests.
  \item \dots
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Split variable selection}

\textbf{Examples:}
\begin{itemize}
  \item Titanic survival and wage determinants data.
  \item Selection of first split variable in root node.
  \item Employ classical statistical tests.
  \item Does not exactly match a particular tree algorithm but similar to CTree.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Split variable selection}
\textbf{Assess pairwise associations:} \code{Survived ~ Gender} (Titanic).

\begin{minipage}[t]{0.43\linewidth}
<<survival-gender, echo=TRUE, eval=FALSE>>=
plot(Survived ~ Gender, data = ttnc)
@
\vspace{-0.7cm}
<<survival-gender-plot, echo=FALSE, fig=TRUE, width=4.5, height=3.4>>=
<<survival-gender>>
@
\end{minipage}
\hfill
\pause
\begin{minipage}[t]{0.5\linewidth}
<<echo=TRUE, results=verbatim>>=
xtabs(~ Survived + Gender, data = ttnc)
@
\end{minipage}

\pause

<<echo=TRUE, results=verbatim>>=
chisq.test(xtabs(~ Survived + Gender, data = ttnc))
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Split variable selection}
\textbf{Assess pairwise associations:} \code{Survived ~ Age} (Titanic).

\begin{minipage}[t]{0.43\linewidth}
<<survival-age, echo=TRUE, eval=FALSE>>=
plot(Survived ~ Age, data = ttnc)
@
\vspace{-0.7cm}
<<survival-age-plot, echo=FALSE, fig=TRUE, width=4.5, height=3.4>>=
<<survival-age>>
@
\end{minipage}
\hfill
\pause
\begin{minipage}[t]{0.5\linewidth}
<<echo=TRUE, results=verbatim>>=
xtabs(~ Survived + Age, data = ttnc)
@
\end{minipage}

<<echo=TRUE, results=verbatim>>=
chisq.test(xtabs(~ Survived + Age, data = ttnc))
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Split variable selection}
\textbf{Assess pairwise associations:} \code{Survived ~ Class} (Titanic).

\begin{minipage}[t]{0.43\linewidth}
<<survival-class, echo=TRUE, eval=FALSE>>=
plot(Survived ~ Class, data = ttnc)
@
\vspace{-0.7cm}
<<survival-class-plot, echo=FALSE, fig=TRUE, width=4.5, height=3.4>>=
<<survival-class>>
@
\end{minipage}
\hfill
\pause
\begin{minipage}[t]{0.5\linewidth}
<<echo=TRUE, results=verbatim>>=
xtabs(~ Survived + Class, data = ttnc)
@
\end{minipage}

<<echo=TRUE, results=verbatim>>=
chisq.test(xtabs(~ Survived + Class, data = ttnc))
@

\end{frame}

% 
% \begin{frame}[fragile]
% \frametitle{Split variable selection} 
% \textbf{Compare pairwise associations} (Titanic): 
% Based on independence tests, e.g., $\chi^2$-test for categorical variables.
% \vspace{0.4cm}
% <<echo=TRUE, results=verbatim>>=
% chisq_test(Survived ~ Gender, data = ttnc)
% @
% 
% \vspace{0.2cm}
% 
% <<echo=TRUE, results=verbatim>>=
% chisq_test(Survived ~ Age, data = ttnc)
% @
% 
% \vspace{0.7cm}
% 
% $\Rightarrow$ Select the split variable showing the lowest $p$-value.
% 
% \end{frame}

\begin{frame}[fragile]
\frametitle{Split variable selection}

\begin{center}
\setkeys{Gin}{width=0.95\textwidth}
<<echo=FALSE, fig=TRUE, width=11>>=
<<ctree_Titanic>>
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Split variable selection}
\textbf{Assess pairwise associations:} \code{log(wage) ~ education} (Wages).

\begin{minipage}[t]{0.39\linewidth}
<<logwage-education, echo=TRUE, eval=FALSE>>=
plot(log(wage) ~ education,
  data = CPS1985)
@
\vspace{-0.7cm}
\setkeys{Gin}{width=\textwidth}
<<logwage-education-plot, echo=FALSE, fig=TRUE, width=4.5, height=3.4>>=
<<logwage-education>>
@
\end{minipage}
\hfill
\pause
\begin{minipage}[t]{0.59\linewidth}
<<echo=TRUE, eval=FALSE>>=
cor.test(~ log(wage) + education, data = CPS1985)
@
<<echo=FALSE, results=verbatim>>=
out <- capture.output(cor.test(~ log(wage) + education, data = CPS1985))
out <- gsub("alternative hypothesis: ", "alternative hypothesis:\n", out, fixed = TRUE)
writeLines(out)
@
\end{minipage}

\end{frame}

\begin{frame}[fragile]
\frametitle{Split variable selection}
\textbf{Assess pairwise associations:} \code{log(wage) ~ gender} (Wages).

\begin{minipage}[t]{0.39\linewidth}
<<logwage-gender, echo=TRUE, eval=FALSE>>=
plot(log(wage) ~ gender,
  data = CPS1985)
@
\vspace{-0.7cm}
\setkeys{Gin}{width=\textwidth}
<<logwage-gender-plot, echo=FALSE, fig=TRUE, width=4.5, height=3.4>>=
<<logwage-gender>>
@
\end{minipage}
\hfill
\pause
\begin{minipage}[t]{0.59\linewidth}
<<echo=TRUE, eval=FALSE>>=
t.test(log(wage) ~ gender, data = CPS1985)
@
<<echo=FALSE, results=verbatim>>=
out <- capture.output(t.test(log(wage) ~ gender, data = CPS1985))
out <- gsub("alternative hypothesis: ", "alternative hypothesis:\n", out, fixed = TRUE)
writeLines(out)
@
\end{minipage}

\end{frame}

\begin{frame}[fragile]
\frametitle{Split variable selection}

\begin{center}
\setkeys{Gin}{width=0.99\textwidth}
<<echo=FALSE, fig=TRUE, width=13>>=
<<ctree_Wages1985>> 
@
\end{center}

\end{frame}

\subsection{Split point selection \& pruning}

\begin{frame}[fragile]
\frametitle{Split point selection}

\textbf{Idea:}
\begin{itemize}
  \item Split $Y$ into most homogenous subgroups with respect to selected $X_{j^*}$.
  \item Homogeneity captured by goodness-of-fit measure.
  \item \emph{Often:} Consider only binary splits.
  \item Trivial for binary $X_{j^*}$.
  \item Otherwise typically exhaustive search over all binary splits.
  \item Possibly already done in split variable selection.
\end{itemize}

\bigskip
\pause

\textbf{Goodness-of-fit measure:} Depends on scale of $Y$ and the adopted model.
\begin{itemize}
  \item Objective function (residual sum of squares, log-likelihood, misclassification rate, impurity, \dots).
  \item Two-sample test statistic or corresponding $p$-value.
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Split point selection}

\textbf{Example:}
\begin{itemize}
  \item Wage determinants data.
  \item Selection of best split point in education in root node.
  \item Residual sum of squares vs.\ normal log-likelihood.
  \item Two-sample ANOVA-type $\chi^2$ test and corresponding $p$-value.
\end{itemize}

\end{frame}

<<split_point_search>>=
evalsplit <- function(sp) {
  m <- lm(log(wage) ~ factor(education <= sp), data = CPS1985)
  s <- independence_test(log(wage) ~ factor(education <= sp), data = CPS1985, teststat = "quadratic") 
  c(
    rss = deviance(m),
    loglik = logLik(m),
    logpval = pchisq(statistic(s), df = 1, log = TRUE, lower.tail = FALSE),
    teststatistic = statistic(s)
  )
}  
sp <- head(sort(unique(CPS1985$education)), -1)
eval <- sapply(sp, evalsplit)
@

\begin{frame}[fragile]
\frametitle{Split point selection}

<<plot_split_point_search1, fig=TRUE, width = 7.5, height = 5>>=
par(mar = c(5, 3, 1, 3))
plot(sp, eval["loglik",], type = "n", xlab = "education", ylab = "", axes = FALSE)
abline(v = 13, lty = 2, lwd = 1.5)
lines(sp, eval["loglik", ], type = "o", col = 2, lwd = 1.5)
axis(2, col.ticks = 2, col.axis = 2)
par(new = TRUE)
plot(sp, eval["teststatistic",], type = "o",
  col = 4, lwd = 1.5,
  xlab = "", ylab = "", axes = FALSE)
axis(4, col.ticks = 4, col.axis = 4)
axis(1)
axis(1, at = 13)
box()
legend("topleft", c("Log-likelihood", "Test statistic"), col = c(2, 4), lwd = 1.5, pch = 1, bty = "n")
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Split point selection}

<<plot_split_point_search2, fig=TRUE, width = 7.5, height = 5>>=
par(mar = c(5, 3, 1, 3))
plot(sp, eval["rss",], type = "n", xlab = "education", ylab = "", axes = FALSE)
abline(v = 13, lty = 2, lwd = 1.5)
lines(sp, eval["rss", ], type = "o", col = 2, lwd = 1.5)
axis(2, col.ticks = 2, col.axis = 2)
par(new = TRUE)
plot(sp, eval["logpval",], type = "o",
  col = 4, lwd = 1.5,
  xlab = "", ylab = "", axes = FALSE)
axis(4, col.ticks = 4, col.axis = 4)
axis(1)
axis(1, at = 13)
box()
legend("bottomleft", c("Residual sum of squares", "Log-p-value"), col = c(2, 4), lwd = 1.5, pch = 1, bty = "n")
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Split point selection}

\begin{center}
\setkeys{Gin}{width=0.99\textwidth}
<<echo=FALSE, fig=TRUE, width=13>>=
<<ctree_Wages1985>> 
@
\end{center}

\end{frame}

\begin{frame}[fragile]
\frametitle{Pruning}

\textbf{Goal:} Avoid overfitting.

\medskip

\textbf{Pre-pruning:} Internal stopping criterium. Stop splitting when there is no significant association to any of the possible split variables $X_j$.

\medskip

\textbf{Post-pruning:} Grow large tree and prune splits that do not improve the model fit (e.g., via cross-validation or information criteria).

\end{frame}

\begin{frame}[fragile]
\frametitle{Pruning}

\textbf{Pre-pruning:}
\begin{itemize}
\item[+] Does not require additional calculations as only already provided information from statistical test is used.
\item[+] Computationally less expensive as trees are prevented from getting too large.
\item[--] Performance depends on the power of the selected statistical test.
\end{itemize}

\vspace{0.4cm}

\textbf{Post-pruning:}
\begin{itemize}
\item[+] Stable method to avoid overfitting, regardless of the power of the employed statistical test.
\item[+] Can employ information criteria such as AIC or BIC for model-based partitioning. 
\item[--] Computationally more expensive as very large trees are grown and additional (out-of-bag) evaluations are required.
\end{itemize}

\end{frame}

\subsection{Conditional inference trees}

\begin{frame}
\frametitle{Conditional inference trees}

\textbf{CTree:}
\begin{itemize} 
  \item Employs general conditional inference (or permutation test) framework for association
    of $h(Y)$ vs.\ $g(X_j)$.
  \item Broadly applicable by choosing suitable transformations $h(\cdot)$ and $g(\cdot)$.
  \item \emph{Examples:} Univariate $Y$ of arbitrary scale, multivariate, model-based, etc.
  \item \emph{Default:} Pre-pruning based on direct asymptotic tests for split variable selection.
    Maximally-selected two-sample statistics for split point selection.
  \item \emph{Software:} R package \emph{partykit} (and previously: \emph{party}).
  \item \emph{Reference:} Hothorn, Hornik, Zeileis (2006).
\end{itemize}

\bigskip

\textbf{Note:} CTree used in previous illustrations of classification and regression trees.

\end{frame}

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{Illustration:} Titanic survival data.

\bigskip

\textbf{Data:} Preprocessing from four-way contingency table \code{Titanic} in base R.

<<echo=TRUE>>=
data("Titanic", package = "datasets")
ttnc <- as.data.frame(Titanic)
ttnc <- ttnc[rep(1:nrow(ttnc), ttnc$Freq), 1:4]
names(ttnc)[2] <- "Gender"
@

\bigskip

\textbf{CTree:} With default arguments.

<<ctree, echo=TRUE, eval=FALSE>>=
library("partykit")
ct_ttnc <- ctree(Survived ~ Gender + Age + Class, data = ttnc)
plot(ct_ttnc)
print(ct_ttnc)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\setkeys{Gin}{width=0.9\linewidth}
<<echo=FALSE, eval=TRUE, fig=TRUE, height=6, width=11>>=
plot(ct_ttnc)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

<<echo=FALSE, eval=TRUE, results=verbatim>>=
print(ct_ttnc)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{Predictions:} Different types.

<<echo=TRUE, results=verbatim>>=
ndm <- data.frame(Gender = "Male", Age = "Adult", Class = c("1st", "2nd", "3rd"))
predict(ct_ttnc, newdata = ndm, type = "node")
predict(ct_ttnc, newdata = ndm, type = "response")
predict(ct_ttnc, newdata = ndm, type = "prob")
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{Predictions:} Women and children first?

<<echo=TRUE, results=verbatim>>=
ndf <- data.frame(Gender = "Female", Age = "Adult", Class = c("1st", "2nd", "3rd"))
ndc <- data.frame(Gender = "Male", Age = "Child", Class = c("1st", "2nd", "3rd"))
cbind(
  Male   = predict(ct_ttnc, newdata = ndm, type = "prob")[, 2],
  Female = predict(ct_ttnc, newdata = ndf, type = "prob")[, 2],
  Child  = predict(ct_ttnc, newdata = ndc, type = "prob")[, 2]
)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{Hyperparameters:} See \code{?ctree_control}.
\begin{itemize}
  \item \code{alpha}: Significance level for pre-pruning in split variable selection.
  \item \code{minsplit}: Minimum number of observations in a node for splitting.
  \item \code{minbucket}: Minimum number of observations in a terminal node.
  \item \code{maxdepth}: Maximum depth of the tree.
  \item \code{multiway}: Use multiway splits for unordered factors with $> 2$ levels?
  \item \dots
\end{itemize}

\bigskip
\pause

\textbf{Example:}

<<echo=TRUE, eval=TRUE>>=
ct_ttnc2 <- ctree(Survived ~ Gender + Age + Class, data = ttnc,
  alpha = 0.01, minbucket = 5, minsplit = 15, maxdepth = 4)
plot(ct_ttnc2)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\setkeys{Gin}{width=\linewidth}
<<echo=FALSE, eval=TRUE, fig=TRUE, width=13, height=6.9>>=
plot(ct_ttnc2)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{Predictions:} Male children in 1st and 2nd class are now in their own terminal nodes.

<<ctree_predict, echo=TRUE, eval=TRUE, results=verbatim>>=
predict(ct_ttnc2, newdata = ndc, type = "prob")
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{Evaluations:} ``As usual.''
\begin{itemize}
  \item In-sample vs. out-of-sample.
  \item Scoring rules (e.g., misclassification rates or other loss functions).
  \item Confusion matrix.
  \item Receiver operator characteristic (ROC) curve.
  \item \dots
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{In-sample:} Augment learning data.

<<echo=TRUE, results=verbatim>>=
ttnc$Fit <- predict(ct_ttnc2, type = "response")
ttnc$Group <- factor(predict(ct_ttnc2, type = "node"))
@

\bigskip

\textbf{Confusion matrix:}

<<echo=TRUE, results=verbatim>>=
xtabs(~ Fit + Survived, data = ttnc)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{Terminal nodes:} Recompute empirical survival rates.

<<echo=TRUE, results=verbatim>>=
tab <- xtabs(~ Group + Survived, data = ttnc)
prop.table(tab, 1)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Conditional inference trees}

\textbf{More graphics:} \emph{ggplot2} support via \code{autoplot()} method from \emph{ggparty}.

\setkeys{Gin}{width=0.9\linewidth}
<<echo=FALSE, eval=TRUE, fig=TRUE, width=13, height=6.9>>=
library("ggparty")
theme_set(theme_minimal())
autoplot(ct_ttnc2)
@

\end{frame}

\subsection{Recursive partitioning}
\begin{frame}
\frametitle{Recursive partitioning}

\textbf{RPart:}
\begin{itemize} 
  \item Implements classic CART algorithm (classification and regression trees).
  \item Split variable and split point selection via exhaustive search based on objective functions.
  \item Hence biased towards split variables with many possible splits.
  \item Cross-validation-based post-pruning and no pre-pruning.
  \item \emph{Default:} Gini impurity for classification, residual sum of squares for regression.
  \item \emph{Software:} R package \emph{rpart}. Similar implementations in other languages (e.g., \emph{scikit-learn}).
  \item \emph{Reference:} Algorithm by Breiman \emph{et al.} (1984), open-source implementation by Therneau \& Atkinson (1997).
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Recursive partitioning}

\textbf{R package:} Part of every standard R installation.

\bigskip

\textbf{Hyperparameters:} \code{?rpart.control}.

\bigskip

\textbf{RPart:} With default arguments.

<<rpart, echo=TRUE, eval=TRUE>>=
library("rpart")
rp_ttnc <- rpart(Survived ~ Gender + Age + Class, data = ttnc)
plot(rp_ttnc)
text(rp_ttnc)
print(rp_ttnc)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Recursive partitioning}

\setkeys{Gin}{width=0.9\linewidth}
<<echo=FALSE, eval=TRUE, fig=TRUE, width=10, height = 6>>=
plot(rp_ttnc)
text(rp_ttnc)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Recursive partitioning}

\setkeys{Gin}{width=0.9\linewidth}
<<echo=FALSE, eval=TRUE, results=verbatim>>=
print(rp_ttnc)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Recursive partitioning}

\textbf{Coercion:}
\begin{itemize}
  \item \code{as.party()} method to coerce \code{rpart} objects to \code{constparty} objects.
  \item Based on generic infrastructure for recursive partitioning in \emph{partykit}.
  \item Enables unified interface (print, plot, predict, \dots).
\end{itemize}

\medskip

<<echo=TRUE>>=
py_ttnc <- as.party(rp_ttnc)
plot(py_ttnc)
print(py_ttnc)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Recursive partitioning}

\setkeys{Gin}{width=0.9\linewidth}
<<echo=FALSE, eval=TRUE, fig=TRUE, width=10, height = 6>>=
plot(py_ttnc)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Recursive partitioning}

\setkeys{Gin}{width=0.9\linewidth}
<<echo=FALSE, eval=TRUE, results=verbatim>>=
print(py_ttnc)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Recursive partitioning}

\textbf{Post-pruning:}
\begin{itemize}
  \item Cost-complexity pruning, similar to using information criteria.
  \item Complexity parameter \code{cp} controls trade-off between reduction of objective function and
    tree size.  
  \item Carried out automatically during fitting.
  \item Additionally, ten-fold cross-validation carried out as well.
  \item Results in \code{rpart} object as \code{cptable}.
\end{itemize}

\bigskip
\pause

\textbf{Here:} Fitted object also has optimal cross-validation error (\code{xerror}).

\medskip

<<echo=TRUE, results=verbatim>>=
rp_ttnc$cptable
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Recursive partitioning}

\textbf{Example:} Employ complexity parameter \code{cp = 0.1} for illustration.

\medskip

<<echo=TRUE, results=verbatim>>=
prune(rp_ttnc, cp = 0.1)
@

\end{frame}

\subsection{Model-based recursive partitioning}
\begin{frame}
\frametitle{Model-based recursive partitioning}

\textbf{MOB:}
\begin{itemize}
  \item Parametric model in subgroups (maximum likelihood, least squares, \dots).
  \item Split variable selection based on asymptotic parameter instability tests.
  \item Split point selection via exhaustive search based on objective function.
  \item Significance-based pre-pruning (default) and optionally
    post-pruning based on information criteria (AIC, BIC, \dots).  
  \item \emph{Examples:} (Generalized) linear model tree, distributional trees, \dots
  \item \emph{Software:} R package \emph{partykit} (and previously: \emph{party}), various ``mobsters'' in extension packages.
  \item \emph{Hyperparameters:} \code{?mob\_control}.
  \item \emph{Reference:} Zeileis, Hothorn, Hornik (2008).
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Model-based recursive partitioning}

\textbf{Example:} Investigate treatment heterogeneity for ``women and children first.''

\medskip

<<ttnc_treatment, echo=TRUE, eval=TRUE>>=
ttnc <- transform(ttnc,
  Treatment = factor(Gender == "Female" | Age == "Child", 
    levels = c(FALSE, TRUE), labels = c("Male&Adult", "Female|Child")
  )
)
@

\bigskip
\pause

\textbf{Model:} Tree based on binary logit GLM for survival.

\medskip

<<glmtree, echo=TRUE>>=
mob_ttnc <- glmtree(Survived ~ Treatment | Class + Gender + Age,
  data = ttnc, family = binomial, alpha = 0.01)
plot(mob_ttnc)
print(mob_ttnc)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Model-based recursive partitioning}

\setkeys{Gin}{width=0.83\textwidth}
<<echo=FALSE, eval=TRUE, fig=TRUE, width=9, height=5>>=
plot(mob_ttnc)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Model-based recursive partitioning}

\vspace{-0.4cm}
\small
<<glmtree_print_eval, echo=FALSE, eval=TRUE, results=verbatim>>=
mob_ttnc
@

\end{frame}

\subsection{Evolutionary learning of globally optimal trees}
\begin{frame}
\frametitle{Evolutionary learning of globally optimal trees}

\textbf{EvTree:}
\begin{itemize}
  \item \emph{Goal:} Globally optimal partition rather than locally optimal split in each step.
  \item NP-hard optimization problem, attempt solution via evolutionary learning.
  \item Based on so-called ``fitness'' function.
  \item Essentially a penalized objective function, similar to information criteria or cost-complexity.
  \item Thus, no additional pruning step.
  \item \emph{Software:} R package \emph{evtree} (leverages \emph{partykit}).
  \item \emph{Hyperparameters:} \code{?evtree\_control}.
  \item \emph{Reference:} Grubinger, Zeileis, Pfeiffer (2014).
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{Evolutionary learning of globally optimal trees}

\textbf{EvTree:} With default parameters.

\medskip

<<echo=TRUE, eval=FALSE>>=
library("evtree")
set.seed(1)
ev_ttnc <- evtree(Survived ~ Gender + Age + class, data = ttnc)
plot(ev_ttnc)
print(ev_ttnc)
@

<<echo=FALSE, eval=TRUE>>=
library("evtree")
if(file.exists("ev_ttnc.rds")) {
  ev_ttnc <- readRDS("ev_ttnc.rds")
} else {
  set.seed(1)
  ev_ttnc <- evtree(Survived ~ Gender + Age + Class, data = ttnc)
  saveRDS(ev_ttnc, file = "ev_ttnc.rds")
}
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Evolutionary learning of globally optimal trees}

\setkeys{Gin}{width=0.72\textwidth}
<<echo=FALSE, eval=TRUE, fig=TRUE, width=8, height=5>>=
plot(ev_ttnc)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{Evolutionary learning of globally optimal trees}

<<echo=FALSE, eval=TRUE, results=verbatim>>=
ev_ttnc
@

\end{frame}

\subsection{Other algorithms}
\begin{frame}
\frametitle{Other algorithms}

\textbf{Furthermore:} Based on \emph{partykit}.
\begin{itemize}
  \item \emph{model4you:} Model-based trees/forests with personalised treatment effects.
  \item \emph{disttree:} Distributional modeling with regression trees and random forests.
  \item \emph{circtree:} Circular distributional trees and forests.
  \item \emph{glmertree:} Generalized linear mixed-effects model trees.
  \item \emph{psychotree:} Psychometric model trees (Rasch, Bradley-Terry, \dots).
  \item \emph{stablelearner:} Stability assessment of trees (and other learners).
  \item \ldots
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Other algorithms}

\textbf{Other R packages:}
\begin{itemize}
  \item \emph{RWeka:} Interface to \emph{Weka} machine learning library, provided open-source
    implementations of C4.5 (J.48) and M5 (M5'), and LMT.
  \item \emph{C50:} C5.0 (successor to C4.5) decision trees and rule-based models.
  \item Many more: \emph{mvpart}, \emph{quint}, \emph{stima}, \dots.
\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Other algorithms}

\textbf{Beyond R:}
\begin{itemize}
  \item Unbiased recursive partitioning software (standalone) by Loh and co-workers: \emph{QUEST}, \emph{GUIDE}, \dots
  \item In Python: \emph{scikit-learn} implements classical machine learning algorithms,
    very limited availability of statistical inference-based algorithms.
  \item \dots
\end{itemize}

\end{frame}

\subsection{References}

\begin{frame}
\frametitle{References}

\small

Breiman L, Friedman JH, Olshen RA, Stone CJ (1984).
  \dquote{Classification and Regression Trees.}
  \emph{Wadsworth, California.},

\medskip

% Breiman L (2001).
%   \dquote{Random {Forests}.}
%   \emph{Machine Learning}, 
%   \textbf{45}(1), 5--32.
%   \doi{10.1023/A:1010933404324}
% 
% \medskip

Hothorn T, Hornik K, Zeileis A (2006).
 \dquote{Unbiased Recursive Partitioning: A Conditional Inference Framework.}
 \emph{Journal of Computational and Graphical Statistics},
 \textbf{15}(3), 651--674.
 \doi{10.1198/106186006X133933}

\medskip

Zeileis A, Hothorn T, Hornik K (2008).
 \dquote{Model-Based Recursive Partitioning.}
  \emph{Journal of Computational and Graphical Statistics},
  \textbf{17}(2), 492--514.
  \doi{10.1198/106186008X319331}

\medskip

Hothorn T, Zeileis A (2015).
 \dquote{{partykit}: A Modular Toolkit for Recursive Partytioning in \textsf{R}.}
 \emph{Journal of Machine Learning Research},
 \textbf{16}, 3905--3909.
 \url{http://www.jmlr.org/papers/v16/hothorn15a.html}

\medskip

Therneau T, Atkinson B (2019). 
  \dquote{{rpart}: Recursive Partitioning and Regression Trees.}
  \url{https://CRAN.R-project.org/package=rpart}

% medskip
%  
% Schlosser L, Hothorn T, Stauffer R, Zeileis A (2019).
% \dquote{Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain.}
% \emph{The Annals of Applied Statistics}, \textbf{13}(3), 1564--1589.
% \doi{10.1214/19-AOAS1247}

% \medskip

% Breiman L (2001).
%   \dquote{Statistical Modeling: The Two Cultures.}
%   \emph{Statistical Science}, 
%   \textbf{16}(3), 199--231.
%   \doi{10.1214/ss/1009213726}

\end{frame}

\end{document}