### 2020-07-02 pkg/partykit/inst/ULGcourse-2020/slides_forest.Rnw
small corrections in section 'aggregation'; added section on distributional forests
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@3035 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: c8e5510b3c8d99b30497ac311a8f39e92a900e30

\documentclass[11pt,t,usepdftitle=false,aspectratio=169]{beamer}
\usetheme[nototalframenumber,license]{uibk}

\title{Random Forests}
\subtitle{Supervised Learning: Algorithmic Modeling}
\author{Lisa Schlosser, Achim Zeileis}

%% forest header image
\renewcommand{\headerimage}[1]{%
   \IfStrEqCase{#1}{%
      {1}{%
         \gdef\myheaderimageid{#1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{forest.jpg}%
      }}[%
         \gdef\myheaderimageid{1}%
         \gdef\myheaderimageposition{nw}%
         \gdef\myheaderimage{forest.jpg}%
      ]%
}
\headerimage{1}

%% custom subsection slides
\setbeamercolor*{subsectionfade}{use={normal text},parent={normal text},fg=structure.fg!30!normal text.bg}
\AtBeginSubsection[]{%
  \begin{frame}[c]
    \begin{center}
      \usebeamercolor[fg]{subsectionfade}
      \Large \insertsection \\[2ex]
      \usebeamercolor[fg]{structure}
      \huge\bfseries\insertsubsection
    \end{center}
  \end{frame}
}

\usepackage[utf8]{inputenc}

\setbeamertemplate{caption}{\insertcaption} 
%% includes a replacement for \usepackage{Sweave}
%\usepackage{Sweave}
\usepackage{changepage}
\usepackage{amsmath,tikz}
\usepackage{calc}
\usepackage{graphicx}
\usetikzlibrary{positioning,shapes,arrows,decorations.pathreplacing,calc,automata,mindmap,trees,tikzmark,decorations.pathreplacing}
\usepackage{xcolor}
\usepackage{changepage}
%\usepackage[cal=boondoxo]{mathalfa}
%\graphicspath{{plots/}}
\newcommand{\argmax}{\operatorname{argmax}\displaylimits}
\newcommand{\argmin}{\operatorname{argmin}\displaylimits}

%% colors
\definecolor{HighlightOrange}{rgb}{0.9490196,0.5725490,0.0000000}
\definecolor{HighlightBlue}{rgb}{0.4784314,0.7490196,0.9803922}
\definecolor{forestred}{RGB}{206,73,81}
\definecolor{treegreen}{RGB}{0,143,0}
\definecolor{lightblue}{RGB}{34,151,230}
\definecolor{lightorange}{RGB}{255,165,0}

\SweaveOpts{engine=R, eps=FALSE, keep.source=TRUE}
<<preliminaries, echo=FALSE, results=hide>>=
options(prompt = "R> ", continue = "+  ", useFancyQuotes = FALSE, width = 70)

set.seed(7)

invisible(.Call(grDevices:::C_palette, grDevices::hcl(
  h = c(0,   5, 125, 245, 195, 315,  65,   0),
  c = c(0, 100,  90,  85,  63, 105,  90,   0),
  l = c(0,  55,  75,  60,  82,  48,  80,  65)
)))

library("partykit")
library("Formula")
library("latex2exp")
library("lattice")
library("MASS")
library("colorspace")
library("disttree")
library("latex2exp")
library("gamlss")
library("crch")
library("RainTyrol")
library("ggplot2")
theme_set(theme_bw(base_size = 18))
library(colorspace)
## HCL palette
pal <- hcl(c(10, 128, 260, 290, 50), 100, 50)
names(pal) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")

pallight <- hcl(c(10, 128, 260, 290, 70), 100, 50, alpha = 0.25)
names(pallight) <- c("forest", "tree", "gamlss", "gamboostLSS", "EMOS")

transpgray <- rgb(0.190,0.190,0.190, alpha = 0.2)

lightblue <- "#2297E6"
lightorange <- "#FFA500"
@

<<motivation_tree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
nobs <- 200
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- numeric(length = length(x))
for(i in 1:(nobs)) ytrue[i] <- if(x[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))
y <- ytrue + rnorm(nobs,0,0.3)
# more points for precise illustration of true function
x100 <- c(1:(100*nobs))/(100*nobs)
ytrue <- ytree <- ytree2 <- ytree3 <- ytree4 <- yforest <- numeric(length = length(x100))
for(i in 1:(nobs*100)) ytrue[i] <- if(x100[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x100[i]-0.2)-1)))
for(i in 1:(nobs*100)) ytree[i] <- if(x100[i]<1/3) 0.5 else {if(x100[i]<2/3) 2 else 1}
for(i in 1:(nobs*100)) ytree2[i] <- if(x100[i]<0.31) 0.47 else {if(x100[i]<0.69) 1.9 else 1.1}
for(i in 1:(nobs*100)) ytree3[i] <- if(x100[i]<0.34) 0.52 else {if(x100[i]<0.54) 2.2 else {if(x100[i]<0.74) 1.74 else 0.9}}
for(i in 1:(nobs*100)) ytree4[i] <- if(x100[i]<0.32) 0.51 else {if(x100[i]<0.63) 1.8 else 1.15}
for(i in 1:(nobs*100)) yforest[i] <- if(x100[i]<=0.33) 0.52 else {if(x100[i]<=0.34) 1.5 else {if(x100[i]<=0.54) 2 else {if(x100[i]<=0.63) 1.9 else {if(x100[i]<=0.69) 1.6 else {if(x100[i]<=0.74) 1.3 else 1}}}}}
@

\begin{document}

\section{Random forests}

\subsection{Motivation}
\begin{frame}[fragile]
\frametitle{Motivation}

{\bf Idea:} \\
\begin{itemize}
\item Combine an ensemble of trees. 
\item Each tree can capture non-linear and non-additive effects and select covariates and possible interactions automatically.
\item Combining trees to a forest model allows for an approximation of smooth effects.
\item A forest can regularize and stabilize the model.
\end{itemize}

%\bigskip

\end{frame}

\begin{frame}
\frametitle{Motivation}

\vspace{0.2cm}

\begin{minipage}{0.4\textwidth}
\only<1>{\textbf{Tree model:\\}}\only<2-8>{\textbf{Subsampling:\\}}\only<9->{\textbf{Forest model:\\}}
\vspace{-0.4cm}
%\vspace{-0.8cm}
\begin{center}
\only<1>{
%\vspace{0.2cm}
\resizebox{0.63\textwidth}{!}{
\begin{tikzpicture}
  \node[ellipse, fill=HighlightBlue!70, align=center] (n0) at (2, 3) {};
  \node[] (n00) at (1, 1.5) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n1) at (0, 0) {$\hat{Y}_1$};
  \draw[-] (n0) -- (n00) node [midway, left] {\small $X \leq p_1$};
  \draw[-] (n0) -- (n1);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n2) at (3, 1.5) {};
  \draw[-] (n0) -- (n2) node [midway, right] {\small $X > p_1$};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n3) at (2, 0) {$\hat{Y}_2$};
  \draw[-] (n2) -- (n3) node [midway, left] {\small $X \leq p_2$};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n4) at (4, 0) {$\hat{Y}_3$};
  \draw[-] (n2) -- (n4) node [midway, right] {\small $X > p_2$};
\end{tikzpicture}}

\vspace{0.475cm}

}
\only<2->{
%\vspace{0.2cm}
\resizebox{0.6\textwidth}{!}{
\begin{tikzpicture}
%\draw[ellipse, draw=black, align=center, scale = 0.7, minimum width=170pt, minimum height = 35pt] (set) at (1, 3) {};
\draw[gray,rounded corners=10pt] (-1.05,4.35) -- (4.05,4.35) -- (4.05,2.7) -- (-2.05,2.7) -- (-2.05,4.35) -- (0.05,4.35);
%\draw[gray] (1,3.5) ellipse (3 and 0.8);
%\visible<4>{
\begin{scope}
\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\clip (1,3.5) ellipse (3 and 0.8);
\pgfmathsetseed{7}
\foreach \p in {1,...,200} {\fill[gray] (1+3*rand,3.5+0.8*rand) circle (0.04);}
\end{scope}
%}
%%% color points
%\visible<5-6>{
%\begin{scope}
%\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\pgfmathsetseed{7}
%\foreach \p in {1,...,100} {\fill[treegreen] (1+3*rand,3.5+0.8*rand) circle (0.07);}
%\end{scope}
%}
\visible<3->{
\node[ellipse, draw=treegreen, align=center, minimum width=80pt, minimum height = 40pt, line width = 3pt] (subset1) at (0.1, 3.5) {};
}
%\visible<7-8>{
%\begin{scope}
%\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\pgfmathsetseed{7}
%\foreach \p in {1,...,49} {\fill[gray] (1+3*rand,3.5+0.8*rand) circle (0.04);}
%\foreach \p in {50,...,150} {\fill[lightblue] (1+3*rand,3.5+0.8*rand) circle (0.07);}
%\end{scope}
%}
\visible<5->{
\node[ellipse, draw = lightblue, align=center, minimum width=100pt, minimum height = 30pt, line width = 3pt] (subset2) at (1.2, 3.6) {};
}
%\visible<9-10>{
%\begin{scope}
%\clip (-1,4.3) -- (4,4.3) -- (4,2.7) -- (-2,2.7) -- (-2,4.3) -- (0,4.3);
%\pgfmathsetseed{7}
%\foreach \p in {1,...,99} {\fill[gray] (1+3*rand,3.5+0.8*rand) circle (0.04);}
%\foreach \p in {100,...,200} {\fill[lightorange] (1+3*rand,3.5+0.8*rand) circle (0.07);}
%\end{scope}
%}
\visible<7->{
\node[ellipse, draw = lightorange, align=center, minimum width=120pt, minimum height = 25pt, line width = 3pt] (subset3) at (1.6, 3.3) {};
}
\visible<4->{
\node[ellipse, fill=treegreen, align=center] (n00) at (-1, 2) {};
\node[ellipse, fill=treegreen, align=center] (n01) at (-1.25, 1.25) {};
\draw[-, line width=1pt] (n00) -- (n01);
\node[rectangle, fill=treegreen, align=center] (n02) at (-1.5, 0.5) {};
\draw[-, line width=1pt] (n01) -- (n02);
\node[rectangle, fill=treegreen, align=center] (n03) at (-1, 0.5) {};
\draw[-, line width=1pt] (n01) -- (n03);
\node[rectangle, fill=treegreen, align=center] (n04) at (-0.5, 0.5) {};
\draw[-, line width=1pt] (n00) -- (n04);
}
\visible<6->{
\node[ellipse, fill=lightblue, align=center] (n10) at (1, 2) {};
\node[ellipse, fill=lightblue, align=center] (n11) at (0.5, 1.25) {};
\draw[-, line width=1pt] (n10) -- (n11);
\node[rectangle, fill=lightblue, align=center] (n12) at (0.25, 0.5) {};
\draw[-, line width=1pt] (n11) -- (n12);
\node[rectangle, fill=lightblue, align=center] (n13) at (0.75, 0.5) {};
\draw[-, line width=1pt] (n11) -- (n13);
\node[ellipse, fill=lightblue, align=center] (n14) at (1.5, 1.25) {};
\draw[-, line width=1pt] (n10) -- (n14);
\node[rectangle, fill=lightblue, align=center] (n15) at (1.75, 0.5) {};
\draw[-, line width=1pt] (n14) -- (n15);
\node[rectangle, fill=lightblue, align=center] (n16) at (1.25, 0.5) {};
\draw[-, line width=1pt] (n14) -- (n16);
}
\visible<8->{
\node[ellipse, fill=lightorange, align=center] (n20) at (3, 2) {};
\node[rectangle, fill=lightorange, align=center] (n21) at (2.5, 0.5) {};
\draw[-, line width=1pt] (n20) -- (n21);
\node[ellipse, fill=lightorange, align=center] (n22) at (3.25, 1.25) {};
\draw[-, line width=1pt] (n20) -- (n22);
\node[rectangle, fill=lightorange, align=center] (n23) at (3, 0.5) {};
\draw[-, line width=1pt] (n22) -- (n23);
\node[rectangle, fill=lightorange, align=center] (n24) at (3.5, 0.5) {};
\draw[-, line width=1pt] (n22) -- (n24);
}
\visible<10->{
\node[ellipse, draw = black, align=center, minimum width=15pt, minimum height = 15pt, line width = 2pt] (B1) at (-1.5, 0.5) {};
\node[ellipse, align=center, line width = 3pt, text=treegreen] (Y1) at (-1.5, -0.5) {\LARGE $\hat{y}_1$};
}
\visible<11->{
\node[ellipse, draw = black, align=center, minimum width=15pt, minimum height = 15pt, line width = 2pt] (B2) at (1.25, 0.5) {};
\node[ellipse, align=center, line width = 3pt, text=lightblue] (Y2) at (1.25, -0.5) {\LARGE $\hat{y}_2$};
}
\visible<12->{
\node[ellipse, draw = black, align=center, minimum width=15pt, minimum height = 15pt, line width = 2pt] (B2) at (3.5, 0.5) {};
\node[ellipse, align=center, line width = 3pt, text=lightorange] (Y3) at (3.5, -0.5) {\LARGE $\hat{y}_3$};(3.5, 0.5)
}
\end{tikzpicture}
}
}
\end{center}
\end{minipage}
\hspace{0.7cm}
\begin{minipage}{0.36\textwidth}
%\begin{center}
%\vspace{0.1cm}
\only<1-2>{
%\vspace{0.01cm}
<<plot_motivation_tree, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,0))
#par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x100, y = ytree, col = pal["forest"], lwd=7)
mtext(text = "X", side = 1, cex = 2.5, line = 2)  
mtext(text = "Y", side = 2, cex = 2.5, line = 1)  
@
}
\only<3>{
%\vspace{0.01cm}
%\hspace{-0.1cm}
<<plot_motivation_randforest_data, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,0))
#par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
mtext(text = "X", side = 1, cex = 2.5, line = 2)  
mtext(text = "Y", side = 2, cex = 2.5, line = 1)  
@
}
\only<4-5>{
%\vspace{0.01cm}
%\hspace{-0.1cm}
<<plot_motivation_randforest1, fig=TRUE, echo=FALSE, width=7>>=
<<plot_motivation_randforest_data>>
lines(x = x100, y = ytree2, col = pal["tree"], lwd=7)
@
}
\only<6-7>{
%\vspace{0.01cm}
%\hspace{-0.1cm}
<<plot_motivation_randforest2, fig=TRUE, echo=FALSE, width=7>>=
<<plot_motivation_randforest1>>
lines(x = x100, y = ytree3, col = lightblue, lwd=7)
@
}
\only<8>{
%\vspace{0.01cm}
%\hspace{-0.1cm}
<<plot_motivation_randforest3, fig=TRUE, echo=FALSE, width=7>>=
<<plot_motivation_randforest2>>
lines(x = x100, y = ytree4, col = lightorange, lwd=7)
@
}
\only<9-12>{
<<plot_motivation_randforest_x, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,0))
#par(mar=c(3,3,2,5.5))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x100, y = ytree2, col = pal["tree"], lwd=7)
lines(x = x100, y = ytree3, col = lightblue, lwd=7)
lines(x = x100, y = ytree4, col = lightorange, lwd=7)
lines(x = c(0.6,0.6), y = c(-1, 1.8), type = "l", lty = 2, lwd = 4)
mtext(text = "Y", side = 2, cex = 2.5, line = 1)  
mtext(text = "x", side = 1, cex = 4, line = 2, at=0.6)  
@
}
\only<13->{
<<plot_motivation_randforest4, fig=TRUE, echo=FALSE, width=7>>=
par(mar=c(3,3,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x100, y = yforest, col = pal["forest"], lwd=7)
lines(x = c(0.6,0.6), y = c(-1, 1.8), type = "l", lty = 2, lwd = 4)
mtext(text = "x", side = 1, cex = 4, line = 2, at=0.6)  
mtext(text = "Y", side = 2, cex = 2.5, line = 1)  
@
}
\end{minipage}

\vspace{0.5cm}

\only<13->{
$\hat{y}= \frac{1}{3} (\hat{y}_1 + \hat{y}_2 + \hat{y}_3)$
}

\vspace{0.3cm}

\visible<14->{
for $\hat{y}_t$ being the average response value in the segment $\mathcal{B}^t_x$ which the observed covariate $x$ is assigned to in the $t$-th tree with learning data $\{(y_i,x_i)\}_{i=1,\ldots,n}$:

\vspace{0.3cm}

$\hat{y}_t= \frac{1}{|\mathcal{B}^t_x|} \sum_{i: y_i \in \mathcal{B}^t_x} y_i$
\visible<15->{
$\ =\ \sum_{i=1}^n \frac{I(x_i \in \mathcal{B}^t_x)}{|\mathcal{B}^t_x|} \cdot y_i$
}
\visible<16->{
$\ =\ \sum_{i=1}^n w_i(x) \cdot y_i$
}
}
\end{frame}

\subsection{Aggregation of trees\\
\&\\
random sampling}

\begin{frame}
\frametitle{Aggregation of trees}
{\bf Simple tree model:} Obtain the predicted response $\hat{y}$ for an observed covariate~$x$ by averaging over observations in the same terminal node $\mathcal{B}_x$: 
$$
\hat{y}= \frac{1}{|\mathcal{B}_x|} \sum_{i: y_i \in \mathcal{B}_x} y_i
\visible<2->{
\ =\ \sum_{i=1}^n w_i(x) \cdot y_i$$
}
\visible<3->{
$$\ =\ \argmin_{\mu} \sum_{i=1}^n w_i(x) \cdot (y_i - \mu)^2$$
}

\visible<4->{
{\bf Model-based tree:} Obtain the predicted model parameter(s) $\hat{\theta}$ for an observed covariate~$x$ by optimizing a loss function for all observations in the same terminal node $\mathcal{B}_x$:
$$
\hat{\theta}= \argmin_{\theta} \sum_{i: y_i \in \mathcal{B}_x} loss(\theta; y_i)
\ =\ \argmin_{\theta} \sum_{i=1}^n w_i(x) \cdot loss(\theta; y_i)$$
}

% {\bf Note:} The model-based approach allows for fitting complex models but also includes the specific case of averaging over all observations which corresponds to minimizing the residual sum of squares $\sum (y_i - \theta)^2$.

\end{frame}

\begin{frame}
\frametitle{Aggregation of trees}
{\bf Strategy:} Combine trees to a forest model by employing tree-based weights.

The predicted model parameter (vector) $\hat{\theta}$ 
is the argument of the parameter space $\Theta$ 
that minimizes the weighted sum of a loss function evaluated for learning data $\{(y_i,x_i)\}_{i=1,\ldots,n}$:

\vspace{-0.2cm}

\[
\hat{\theta}(x) =  \argmin_{\theta} \sum_{i=1}^n w_i(x) \cdot loss(\theta; y_i)
\]

\medskip

\only<2-4>{

\vspace{0.13cm}

\textbf{Weights:}

\vspace{-1.14cm}

\begin{eqnarray*}
w^{\text{base}}_i(x)   & = & 1 \\[0.2cm]
\visible<3-4>{
w^{\text{tree}}_i(x)   & = & \sum_{b=1}^B I((x_i \in \mathcal{B}_b) \land (x \in \mathcal{B}_b)) \\[0.1cm]
\visible<4>{
w^{\text{forest}}_i(x) & = & \frac{1}{T} \sum_{t=1}^T \sum_{b=1}^{B^t} \frac{I((x_i \in \mathcal{B}^t_b) \land (x \in \mathcal{B}^t_b))}{\left|\mathcal{B}^t_b\right|}
\end{eqnarray*}
}}}

\end{frame}

\begin{frame}
\frametitle{Random sampling}
{\bf Idea:} Build each tree on a different subset of the full data set.

\medskip

{\bf Sampling observations:}
\begin{itemize}
\item Bootstrapping: resampling with replacement, usually ``n out of n''.
\item Subsampling: draw a subset without replacement (typically of the size of 63.2\% of the full data set; for large data sets a smaller fraction can be advantageous).
\end{itemize}

\bigskip

{\bf Sampling covariates:}\\
\smallskip
In order to reduce the correlation among trees random input variable sampling is employed, i.e., for each split only a subset of covariates is selected randomly and provided as possible split variables (typical sampling size: $\sqrt{\text{\#Var}}$).
\end{frame}

\begin{frame}
\frametitle{Hyperparameter selection}

Apart from the size and type of sampling two further control parameters have to be set for a forest model: 
\begin{itemize}
\item {\bf Number of trees:} A high number of trees allows for a more stable model and for a better approximation of smooth effects, however, at the cost of higher computational effort.
\item {\bf Size of the trees:} The size of each single tree can be controlled by setting stopping criteria such as a minimal number of observations in a node, maximum tree depth or a significance level if a statistical test is employed in the tree algorithm. However, in a forest model each tree is usually built as large as possible which can lead to overfitting for a single tree, but this is compensated by aggregating the trees to a forest model.
\end{itemize}
\end{frame}

<<data_wages, eval=TRUE, echo=FALSE, results=hide>>=
data("CPS1985", package = "AER")
@

\begin{frame}[fragile]
\frametitle{Example: wages}

\textbf{Data:} Random sample from the May 1985 US Current Population Survey.
A data frame containing \Sexpr{nrow(CPS1985)} observations on \Sexpr{ncol(CPS1985)} variables.

\bigskip

\begin{tabular}{ll}
\hline 
Variable & Description\\
\hline 
\code{wage}       & Wage (in US dollars per hour). \\
\code{education}  & Education (in years).  \\
\code{experience} & Potential work experience (in years, \code{age - education - 6}). \\
\code{age}        & Age (in years). \\
\code{ethnicity}  & Factor: Caucasian, Hispanic, Other. \\
\code{gender}     & Factor: Male, Female. \\
\code{union}      & Factor. Does the individual work on a union job? \\
\hline 
\end{tabular}

\bigskip

\textbf{Model formula:} \code{log(wage) ~ education + experience + age + ethnicity + gender + union}.

\end{frame}

<<example_wages, eval=TRUE, echo=FALSE, results=hide>>=
set.seed(4)
f <- log(wage) ~ education + experience + age + ethnicity + gender + union
ct_cps <- ctree(f, data = CPS1985, alpha = 0.01)
cf <- cforest(f, data = CPS1985, ntree = 5) #, control =ctree_control(maxdepth = 5))
vimp <- varimp(cf , risk = "loglik")
@

\begin{frame}[fragile]
\frametitle{Example: wages}
Single tree model:
\setkeys{Gin}{width=0.99\textwidth}
<<example_wages_ctree, eval=TRUE, echo=FALSE, fig=TRUE, width=13>>=
plot(ct_cps)
@
\end{frame}

\begin{frame}[fragile]
\frametitle{Example: wages}
Forest model - tree I:
\setkeys{Gin}{width=0.99\textwidth}
<<example_wages_cforest1, eval=TRUE, echo=FALSE, fig=TRUE, width=13>>=
#par(mfrow=c(2,2))
plot(party(cf[[1]][[1]], data = CPS1985[,c("wage", "education", "experience", "age", "ethnicity", "gender", "union")]), drop_terminal = TRUE)
@
\end{frame}

\begin{frame}[fragile]
\frametitle{Example: wages}
Forest model - tree II:
\setkeys{Gin}{width=0.99\textwidth}
<<example_wages_cforest2, eval=TRUE, echo=FALSE, fig=TRUE, width=13>>=
#par(mfrow=c(2,2))
plot(party(cf[[1]][[2]], data = CPS1985[,c("wage", "education", "experience", "age", "ethnicity", "gender", "union")]), drop_terminal = TRUE)
@
\end{frame}

\begin{frame}[fragile]
\frametitle{Example: wages}
Forest model - tree III:
\setkeys{Gin}{width=0.99\textwidth}
<<example_wages_cforest3, eval=TRUE, echo=FALSE, fig=TRUE, width=13>>=
#par(mfrow=c(2,2))
plot(party(cf[[1]][[3]], data = CPS1985[,c("wage", "education", "experience", "age", "ethnicity", "gender", "union")]), drop_terminal = TRUE)
@
\end{frame}

\begin{frame}[fragile]
\frametitle{Example: wages}
Forest model - tree IV:
\setkeys{Gin}{width=0.99\textwidth}
<<example_wages_cforest4, eval=TRUE, echo=FALSE, fig=TRUE, width=13>>=
#par(mfrow=c(2,2))
plot(party(cf[[1]][[4]], data = CPS1985[,c("wage", "education", "experience", "age", "ethnicity", "gender", "union")]), drop_terminal = TRUE)
@
\end{frame}

\begin{frame}[fragile]
\frametitle{Example: wages}
\vspace{-0.5cm}
\setkeys{Gin}{width=0.94\textwidth}
<<example_wages_forest_prediction, eval=TRUE, echo=FALSE, fig=TRUE, width = 10>>=
nd <- data.frame(education = 13.02, #c(2:18),
                 experience = 17.82,
                 age = c((18:64)/1), #36.83,
                 ethnicity = "cauc",
                 region = "other",
                 gender = "male",
                 occupation = "worker",
                 sector = "other",
                 union = "no",
                 married = "yes")

{
  par(mfrow = c(2,2))
  set.seed(74)
  cf <- cforest(f, data = CPS1985, ntree = 2)
  predwage <- cbind(c((18:64)/1), predict(cf, newdata = nd))
  colnames(predwage) <- c("age", "predicted log(wage)")
  plot(x = predwage[,"age"], y = exp(predwage[,"predicted log(wage)"]), type = 'l',
       ylab = "predicted wage", xlab = "age", main = "Forest with 2 trees")

  cf <- cforest(f, data = CPS1985, ntree = 4)
  predwage <- cbind(c((18:64)/1), predict(cf, newdata = nd))
  colnames(predwage) <- c("age", "predicted log(wage)")
  plot(x = predwage[,"age"], y = exp(predwage[,"predicted log(wage)"]), type = 'l',
       ylab = "predicted wage", xlab = "age", main = "Forest with 4 trees")
  lines(x = predwage[,"age"], y = exp(predwage[,"predicted log(wage)"]), type = 'l',
       col = "gray")

  cf <- cforest(f, data = CPS1985, ntree = 20)
  predwage <- cbind(c((18:64)/1), predict(cf, newdata = nd))
  colnames(predwage) <- c("age", "predicted log(wage)")
  plot(x = predwage[,"age"], y = exp(predwage[,"predicted log(wage)"]), type = 'l',
       ylab = "predicted wage", xlab = "age", main = "Forest with 20 trees")

  cf <- cforest(f, data = CPS1985, ntree = 200)
  predwage <- cbind(c((18:64)/1), predict(cf, newdata = nd))
  colnames(predwage) <- c("age", "predicted log(wage)")
  plot(x = predwage[,"age"], y = exp(predwage[,"predicted log(wage)"]), type = 'l',
       ylab = "predicted wage", xlab = "age", main = "Forest with 200 trees")
}
@
\end{frame}

\begin{frame}
\frametitle{Variable importance}
{\bf Problem:} In a single tree model the effect of each covariate can easily be investigated based on a simple illustration of the tree structure. However, in a forest model this would require to investigate each tree on its own including its specific setting (induced by random sampling).

\bigskip

{\bf Idea:} Asses the influence of a covariate on the forest model by 
\begin{itemize} 
\item permuting the selected covariate and evaluating the resulting change in performance of the model (mean decrease in accuracy, measured for example based on the employed loss function) or
\item calculating the total decrease in node impurities from splitting in the selected covariate (measured for example by the Gini index for classification trees or the RMSE for regression trees), 
\end{itemize}
averaged over all trees.

\bigskip

%\bigskip

% {\bf Implementations:}
% \begin{itemize}
% \item \fct{importance} in the R package randomForest.
% \item \fct{importance.ranger} in the R package ranger.
% \item \fct{varimp} in the R package partykit.
% \end{itemize}

% \only<2->{
% \setkeys{Gin}{width=0.5\textwidth}
% <<example_wages_vimp, eval=TRUE, echo=FALSE, fig=TRUE, width=6, height=3>>=
% par(mar = c(4,5.5,0,2))
% barplot(sort(vimp, decreasing = FALSE),
%         horiz = TRUE, las = 1, axes = FALSE,
%         xlab = "Variable importance: mean decrease in log-likelihood")
% axis(1, at = seq(0,1,0.02), las = 1, mgp=c(0,1,0))
% @

% Depending on the fitted model the corresponding objective function can be employed as \emph{risk function} to measure accuracy by calculating importance scores (e.g., log-likelihood, number of misclassifications, \ldots).\\
% To measure impurity, for example, the Gini index can be employed for classification trees or the RMSE for regression trees.
%}
\end{frame}

\begin{frame}
\frametitle{Example: wages}
Evaluation of variable importance in the fitted forest model:

\bigskip

\setkeys{Gin}{width=0.57\textwidth}
<<example_wages_vimp, eval=TRUE, echo=FALSE, fig=TRUE, width=6, height=3>>=
par(mar = c(4,5.4,1,2))
barplot(sort(vimp, decreasing = FALSE),
        horiz = TRUE, las = 1, axes = FALSE,
        xlab = "Variable importance: mean decrease in log-likelihood")
axis(1, at = seq(0,1,0.02), las = 1, mgp=c(0,1,0))
@
\end{frame}

\subsection{Implementations}

\begin{frame}[fragile]
\frametitle{R software for forest models}

{\bf randomForest:}
\begin{itemize}
  \item R package for classic random forests.
  \item Based on implementations of CART.
  \item First implementation of random forests in R.
\end{itemize}

% \medskip
% 
% <<call_randomForest, eval=FALSE, echo=TRUE>>=
% randomForest(formula, data=NULL, ..., subset, na.action=na.fail)
% randomForest(x, y=NULL,  ...)
% @
% 
% \medskip
% 
% Optional control arguments:
% \begin{itemize}
% \item \code{ntree}: nr. of trees, 
% \item \code{mtry}: nr. of randomly sampled split variable candidates, 
% \item \code{replace}: type of subsetting, 
% \item \code{sampsize}: size of drawn subsamples, 
% \item \code{nodesize}: minimum size of terminal nodes, 
% \item \code{maxnodes}: maximum number of terminal nodes, 
% \item \ldots 
% \end{itemize}

\medskip

{\bf ranger}:
\begin{itemize}
\item RANdom  forest  GEneRator.
\item R package providing a fast implementation of classic random forests.
\end{itemize}

\medskip

{\bf partykit:}
\begin{itemize}
  \item R package offering a toolkit for unbiased recursive partitioning. 
  \item Provides the forest-building function \fct{cforest}, based on conditional inference trees (CTree).
\end{itemize}

\end{frame}

\begin{frame}[fragile]
\frametitle{R software for forest models}

{\bf ranger}:

<<call_ranger, eval=TRUE, echo=TRUE>>=
library("ranger")
rf <- ranger(log(wage) ~ education + experience + age + ethnicity + gender + union,
             data = CPS1985)
@

Optional control arguments:
\begin{itemize}
\item \code{num.trees}: number of trees, 
\item \code{mtry}: number of randomly sampled split variable candidates, 
\item \code{replace}: sample with replacement, 
\item \code{sample.fraction}: fraction of observations to sample, 
\item \code{min.node.size}: minimal size of terminal nodes, 
\item \code{max.depth}: maximal tree depth, 
\item \code{splitrule}: splitting rule, 
\item \ldots 
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{R software for forest models}

{\bf ranger}:

<<call_ranger_predict, eval=TRUE, echo=TRUE, results=verbatim>>=
newdata <- data.frame(education = 13, experience = 17,
                      age = 36, ethnicity = "cauc",
                      region = "other", gender = "male",
                      occupation = "worker", sector = "other",
                      union = "no", married = "yes")
pred_rf <- predict(rf, data = newdata)
pred_rf$predictions     
exp(pred_rf$predictions) 
rf <- ranger(log(wage) ~ education + experience + age + ethnicity + gender + union,
             data = CPS1985, importance = "impurity")
importance(rf)
@

\end{frame}

\begin{frame}[fragile]
\frametitle{R software for forest models}

{\bf partykit:}

<<call_cforest, eval = FALSE, echo=TRUE>>=
library("partykit")
cf <- cforest(log(wage) ~ education + experience + age + ethnicity + gender + union,
              data = CPS1985)
@

\only<1>{
Optional forest-specific control arguments:
\begin{itemize}
\item \code{ytrafo}: transformation function applied to the response,
\item \code{ntree}: number of trees, 
\item \code{mtry}: number of randomly sampled split variable candidates, 
\item \code{perturb}: type and size of subsetting, 
\item \ldots 
\end{itemize}
%\vspace{0.9cm}
}
\only<2->{
Optional tree-specific arguments handed over via \fct{ctree\_control}:
\begin{itemize}
\item \code{minsplit}: minimum number of observations to perform a split, 
\item \code{minbucket}: minimum number of observations in a node after splitting, 
\item \code{maxdepth}: maximum depth of the tree, 
\item \code{alpha}: significance level for split variable selection,
\item \ldots 
\end{itemize}
}
% \only<3>{
% 
% <<call_cforest_control, eval = FALSE, echo=TRUE>>=
% cf <- cforest(log(wage) ~ education + experience + age + ethnicity + gender + union,
%             data = CPS1985, control = ctree_control(minsplit = 20))
% @
% }

\end{frame}

\begin{frame}[fragile]
\frametitle{R software for forest models}

{\bf partykit:}

<<call_cforest, eval = FALSE, echo=TRUE>>=
library("partykit")
cf <- cforest(log(wage) ~ education + experience + age + ethnicity + gender + union,
              data = CPS1985)
@

Optional tree-specific arguments handed over via \fct{ctree\_control}:
\begin{itemize}
\item \code{minsplit}: minimum number of observations to perform a split, 
\item \code{minbucket}: minimum number of observations in a node after splitting, 
\item \code{maxdepth}: maximum depth of the tree, 
\item \code{alpha}: significance level for split variable selection,
\item \ldots 
\end{itemize}

<<call_cforest_control, eval = FALSE, echo=TRUE>>=
cf <- cforest(log(wage) ~ education + experience + age + ethnicity + gender + union,
            data = CPS1985, control = ctree_control(minsplit = 20))
@

\end{frame}
\begin{frame}[fragile]
\frametitle{R software for forest models}

{\bf partykit:}

<<call_cforest_predict, eval = TRUE, results=verbatim>>=
newdata <- data.frame(education = 13, experience = 17,
                      age = 36, ethnicity = "cauc",
                      region = "other", gender = "male",
                      occupation = "worker", sector = "other",
                      union = "no", married = "yes")
pred_cf <- predict(cf, newdata = newdata)
pred_cf
exp(pred_cf)
varimp(cf)
@

\end{frame}

\begin{frame}
\frametitle{R software for forest models}

\textbf{Further R packages:}
%
\begin{itemize}
  \item grf: generalized random forests,
  \item disttree: distributional forests, based on partykit,
  \item circtree: circular distributional forests, based on partykit,
  \item model4you: model-based random forests for personalized treatment effects,
  \item randomSurvivalForest: random forests for the analysis of right-censored survival data, 
  \item \ldots
\end{itemize}

\end{frame}

\subsection{Distributional forests}
\begin{frame}%[fragile]
\frametitle{Motivation}
\vspace{-0.41cm}
\begin{figure}[!htb]
\minipage{0.285\textwidth}
\begin{center}
<<motivation_GLM, echo=FALSE, results=hide>>=
nobs <- 200
## GLM
set.seed(7)
x <- c(1:nobs)/nobs
ytrue <- 1+x*1.5
y <- ytrue + rnorm(nobs,0,0.3)
@
\visible<2->{
<<plot_motivation_GLM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
\end{center}
\endminipage
\visible<3->{{\LARGE$\rightarrow$}}
\minipage{0.285\textwidth}
\begin{center}
<<motivation_GAM, echo=FALSE, results=hide>>=
## GAM
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
y <- ytrue + rnorm(nobs,0,0.3)
@
\visible<3->{
<<plot_motivation_GAM, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(y=y , x=x, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
lines(x = x, y = ytrue, col = pal["forest"], lwd = 7, main = "")
@
}
\end{center}
\endminipage
\visible<4->{{\LARGE$\rightarrow$}}
\minipage{0.285\textwidth}
\begin{center}
<<motivation_GAMLSS, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
x <- c(1:nobs)/nobs
x <- 2*(x-0.5)
ytrue <- x^3
var <- exp(-(2*x)^2)/2
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
@
\visible<4->{
<<plot_motivation_GAMLSS, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x, y, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x, rev(x)), c(ytrue + 0.1 + var, rev(ytrue - 0.1 - var)),
  col = pallight["forest"], border = "transparent")
lines(x, ytrue, col = pal["forest"], lwd=7)
points(x, y, col = "slategray", pch = 19)
box(lwd = 5)
@
}
\end{center}
\endminipage

\vspace{0.5cm}
\minipage{0.25\textwidth}
\begin{center}
\visible<2->{
LM, GLM\\
\vspace{0.5cm}
\code{lm}\\
\code{glm}\\
\vspace{1.5cm}}
\end{center}
\endminipage
\hspace{1.1cm}
\minipage{0.25\textwidth}
\begin{center}
\visible<3->{
GAM\\
\vspace{0.5cm}
\code{mgcv}\\
\code{VGAM}\\
\vspace{1.5cm}}
\end{center}
\endminipage
\hspace{1.1cm}
\minipage{0.25\textwidth}
\begin{center}
\visible<4->{
GAMLSS\\
\vspace{0.5cm}
\code{gamlss}\\
\code{mgcv}\\
\code{VGAM}\\
\code{gamboostLSS}\\
\code{bamlss}}
\end{center}
\endminipage
\end{figure}
\end{frame}

\begin{frame}[fragile]
\frametitle{Motivation}
\vspace{-0.41cm}
\begin{figure}[!htb]
\minipage{0.285\textwidth}
\begin{center}
<<motivation_regtree, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- ytree <- yforest <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))
y <- ytrue + rnorm(nobs,0,0.3)
for(i in 1:nobs) ytree[i] <- if(x[i]<1/3) 0.5 else {if(x[i]<2/3) 2 else 1}
@
\visible<1->{
<<plot_motivation_regtree, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x, y = ytree, col = pal["forest"], lwd=7)
@
}
\end{center}

\endminipage
\visible<2->{{\LARGE$\rightarrow$}}
\minipage{0.285\textwidth}
 \begin{center}
<<motivation_randforest, echo=FALSE, results=hide>>=
## Random Forest
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@
\visible<2->{
<<plot_motivation_randforest, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x, y = yforest, col = pal["forest"], lwd=7, main = "")
@
}
\end{center}

\endminipage
\visible<3->{{\LARGE$\rightarrow$}}
\minipage{0.285\textwidth}
\begin{center}
\visible<3->{
<<plot_motivation_question, fig=TRUE, echo=FALSE>>=
par(mar=c(2,0,2,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, type = "n")
box(lwd=5)
text(x = mean(range(x)), y = mean(range(y)), "?", cex = 12)
@
}
\end{center}

\endminipage

\minipage{0.285\textwidth}
\begin{center}
\vspace{0.0cm}
\visible<1->{
Regression tree\\
\vspace{0.3cm}
\resizebox{0.2\textwidth}{!}{
\begin{tikzpicture}
  \node[ellipse, fill=HighlightBlue!70, align=center] (n0) at (1, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n1) at (0.5, 1) {};
  \draw[-, line width=1pt] (n0) -- (n1);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n2) at (1.5, 1) {};
  \draw[-, line width=1pt] (n0) -- (n2);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n3) at (1, 0) {};
  \draw[-, line width=1pt] (n2) -- (n3);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n4) at (2, 0) {};
  \draw[-, line width=1pt] (n2) -- (n4);
\end{tikzpicture}}
\vspace{0.2cm}\\
\code{rpart}\\
\code{party(kit)}\\
\vspace{1cm}}
\end{center}
\endminipage
\hspace{0.65cm}
\minipage{0.285\textwidth}
\begin{center}
\vspace{-0.4cm}
\visible<2->{
Random forest\\
\vspace{0.4cm}
\resizebox{0.6\textwidth}{!}{
\begin{tikzpicture}
  \node[ellipse, fill=HighlightBlue!70, align=center] (n00) at (1, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n01) at (0.5, 1) {};
  \draw[-, line width=1pt] (n00) -- (n01);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n02) at (1.5, 1) {};
  \draw[-, line width=1pt] (n00) -- (n02);

  \node[ellipse, fill=HighlightBlue!70, align=center] (n10) at (3, 2) {};
  \node[ellipse, fill=HighlightBlue!70, align=center] (n11) at (2.5, 1) {};
  \draw[-, line width=1pt] (n10) -- (n11);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n12) at (3.5, 1) {};
  \draw[-, line width=1pt] (n10) -- (n12);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n13) at (2, 0) {};
  \draw[-, line width=1pt] (n11) -- (n13);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n14) at (2.8, 0) {};
  \draw[-, line width=1pt] (n11) -- (n14);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n15) at (3.2, 0) {};
  \draw[-, line width=1pt] (n12) -- (n15);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n16) at (4, 0) {};
  \draw[-, line width=1pt] (n12) -- (n16);

  \node[ellipse, fill=HighlightBlue!70, align=center] (n20) at (5, 2) {};
  \node[rectangle, fill=HighlightOrange!70, align=center] (n21) at (4.5, 1) {};
  \draw[-, line width=1pt] (n20) -- (n21);
  \node[ellipse, fill=HighlightBlue!70, align=center] (n22) at (5.5, 1) {};
  \draw[-, line width=1pt] (n20) -- (n22);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n23) at (5, 0) {};
  \draw[-, line width=1pt] (n22) -- (n23);
  \node[rectangle, fill=HighlightOrange!70, align=center] (n24) at (6, 0) {};
  \draw[-, line width=1pt] (n22) -- (n24);
\end{tikzpicture}
}
\vspace{0.3cm}\\
\code{randomForest}\\
\code{ranger}\\
\code{party(kit)}}
\end{center}
\endminipage
\hspace{0.65cm}
\minipage{0.285\textwidth}
\begin{center}
\vspace{0.15cm}
\visible<3->{
Distributional trees and forests\\
\vspace{1.15cm}
\code{disttree}\\
based on \code{partykit}\\
\vspace{1cm}}
\end{center}
\endminipage
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Building blocks}
%\vspace{0.5cm}

\begin{minipage}{0.79\textwidth}
\vspace{-0.5cm}
\textbf{Distributional modeling:}
\begin{itemize}
\vspace{-0.05cm}
  \item Specify the complete probability distribution \\
  \vspace{-0.05cm}
  (location, scale, shape, \dots).
\end{itemize}

\smallskip

\textbf{Tree:}
\begin{itemize}
\vspace{-0.05cm}
  \item Capture non-linear and non-additive effects.
  \vspace{-0.1cm}
  \item Automatic selection of covariates and interactions. %detection of steps and abrupt changes. %(data driven)
\end{itemize}

\smallskip

\textbf{Forest:}
\begin{itemize}
\vspace{-0.05cm}
  \item Smoother effects.
  \vspace{-0.1cm}
  \item Stabilization and regularization of the model.
\end{itemize}

\bigskip

\bigskip

$\Rightarrow$ \textbf{Synthesis:} Distributional Trees and Forests
\end{minipage}
<<motivation_GAMLSS_syn, echo=FALSE, results=hide>>=
## GAMLSS
set.seed(7)
nobs <- 200
x_g <- c(1:nobs)/nobs
x_g <- 2*(x_g-0.5)
ytrue_g <- x_g^3
var_g <- exp(-(1.74*x_g)^2)/2.5
y_g <- ytrue_g + rnorm(nobs, 0, 0.1 + var_g)
@
<<motivation_regtree_syn, echo=FALSE, results=hide>>=
## Reg. Tree
set.seed(7)
kappa <- 12
x <- c(1:nobs)/nobs
ytrue <- ytree <- yforest <- var <- numeric(length = length(x))
for(i in 1:nobs) ytrue[i] <- if(x[i]<1/3) 0.5 else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))
for(i in 1:nobs) var[i] <-  if(x[i]<1/3) 0.1 else 0.6 *(4/3 - x[i])
y <- ytrue + rnorm(nobs, 0, 0.1 + var)
for(i in 1:nobs) ytree[i] <- if(x[i]<1/3) 0.5 else {if(x[i]<2/3) 2 else 1}
@
<<motivation_randforest_syn, echo=FALSE, results=hide>>=
## Random Forest
for(i in 1:nobs) yforest[i] <- if(x[i]<0.27) 0.5 else { if(x[i]<0.39) 0.5 + 1.5*(plogis((x[i]-0.33)/6*700)) else 1+(1-plogis(kappa*(2*(x[i]-0.2)-1)))}
@
\begin{minipage}{0.2\textwidth}
\vspace{-0.05cm}
<<plot_motivation_GAMLSS_syn, fig=TRUE, echo=FALSE, width=8.5>>=
par(mar=c(1,0,1,0))
plot(x_g, y_g, xaxt = "n", yaxt = "n", ann = FALSE, type = "n")
polygon(c(x_g, rev(x_g)), c(ytrue_g + 0.1 + var_g, rev(ytrue_g - 0.1 - var_g)),
  col = pallight["forest"], border = "transparent")
lines(x_g, ytrue_g, col = pal["forest"], lwd=10)
points(x_g, y_g, col = "slategray", pch = 19)
box(lwd = 5)
@

\vspace{0.15cm}

<<plot_motivation_regtree_syn, fig=TRUE, echo=FALSE, width=8.5>>=
par(mar=c(1,0,1,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(y), max(y)))
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x, y = ytree, col = pal["forest"], lwd=10)
box(lwd=5)
@

\vspace{0.15cm}

<<plot_motivation_randforest4_syn, fig=TRUE, echo=FALSE, width=8.5>>=
par(mar=c(1,0,1,0))
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(y), max(y)))
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
lines(x = x, y = yforest, col = pal["forest"], lwd=10, main = "")
box(lwd=5)
@

\vspace{0.15cm}

<<plot_motivation_distforest, fig=TRUE, echo=FALSE, width=8.5>>=
par(mar=c(1,0,1,0))
#plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19)
#box(lwd=5)
#lines(x = x, y = ytrue, col = "gray", lwd=5, main = "")
#lines(x = x, y = yforest, col = pal["forest"], lwd=7, main = "")
plot(x = x, y = y, xaxt="n", yaxt="n", ann=FALSE, col = "slategray", pch = 19,
     ylim = c(min(y), max(y)))
polygon(c(x, rev(x)), c(yforest + 0.1 + var, rev(yforest - (0.1 + var))),
       # c(ytrue_g + 0.1 + var_g, rev(ytrue_g - 0.1 - var_g))
  col = pallight["forest"], border = "transparent")
box(lwd=5)
lines(x = x, y = yforest, col = pal["forest"], lwd=10)
@
\end{minipage}
\end{frame}

<<treedata, echo=FALSE, results=hide>>=
set.seed(54)
nobs <- 500
x <- runif(nobs, 0, 1)
mu <- sigma <- ytrue <- numeric(length = nobs)
for(i in 1:nobs) sigma[i] <- if(x[i]<=0.4) 1 else 3
for(i in 1:nobs) mu[i] <- if(x[i]<= 0.4|| x[i]>0.8) 4 else 12
y <- rnorm(nobs, mean = mu, sd = sigma)
ytrue <- mu
data <- data.frame(cbind(y,x, ytrue))

alldata <- cbind(data, mu, sigma)
odata <- alldata[order(alldata["x"]),]
@

<<dgp_tree, echo=FALSE, results=hide>>=
data <- data.frame(x = numeric(0), x = numeric(0), x = numeric(0))
names(data) <- c("x","x","x")
fig <- party(
  partynode(1L,
            split = partysplit(2L, breaks = 0.4),
            kids = list(
              partynode(2L, info = c(
                "n = 200",
                "   True parameters:   ",
                expression(mu == '4'),
                expression(sigma == '1')
              )),
              partynode(3L,
                        split = partysplit(3L, breaks = 0.8),
                        kids = list(
                          partynode(4L, info = c(
                            "n = 200",
                            "   True parameters:   ",
                            expression(mu == '12'),
                            expression(sigma == '3')
                          )),
                          partynode(5L, info = c(
                            "n = 100",
                            "   True parameters:   ",
                            expression(mu == '4'),
                            expression(sigma == '3')
                          )))))),
  data
)

node_inner_ext <- function (obj, id = TRUE, pval = TRUE, abbreviate = FALSE, fill = "white", 
    gp = gpar()) 
{
    meta <- obj$data
    nam <- names(obj)
    extract_label <- function(node) {
        if (is.terminal(node)) 
            return(rep.int("", 2L))
        varlab <- character_split(split_node(node), meta)$name
        if (abbreviate > 0L) 
            varlab <- abbreviate(varlab, as.integer(abbreviate))
        if (pval) {
            nullna <- function(x) is.null(x) || is.na(x)
            pval <- suppressWarnings(try(!nullna(info_node(node)$p.value), 
                silent = TRUE))
            pval <- if (inherits(pval, "try-error")) 
                FALSE
            else pval
        }
        if (pval) {
            pvalue <- node$info$p.value
            plab <- ifelse(pvalue < 10^(-3L), paste("p <", 10^(-3L)), 
                paste("p =", round(pvalue, digits = 3L)))
        }
        else {
            plab <- ""
        }
        return(c(varlab, plab))
    }
    maxstr <- function(node) {
        lab <- extract_label(node)
        klab <- if (is.terminal(node)) 
            ""
        else unlist(lapply(kids_node(node), maxstr))
        lab <- c(lab, klab)
        lab <- unlist(lapply(lab, function(x) strsplit(x, "\n")))
        lab <- lab[which.max(nchar(lab))]
        if (length(lab) < 1L) 
            lab <- ""
        return(lab)
    }
    nstr <- maxstr(node_party(obj))
    if (nchar(nstr) < 6) 
        nstr <- "aAAAAa"
    rval <- function(node) {
        node_vp <- viewport(x = unit(0.5, "npc"), y = unit(0.5, 
            "npc"), width = unit(1, "strwidth", nstr) * 1.3, 
            height = unit(3, "lines"), name = paste("node_inner", 
                id_node(node), sep = ""), gp = gp)
        pushViewport(node_vp)
        xell <- c(seq(0, 0.2, by = 0.01), seq(0.2, 0.8, by = 0.05), 
            seq(0.8, 1, by = 0.01))
        yell <- sqrt(xell * (1 - xell))
        xell <- xell*1.11 - 0.055             # to adapt size of the ellipse to the size with p-value
        lab <- extract_label(node)
        fill <- rep(fill, length.out = 2L)
        grid.polygon(x = unit(c(xell, rev(xell)), "npc"), y = unit(c(yell, 
            -yell) + 0.5, "npc"), gp = gpar(fill = fill[1]))
        grid.text(lab[1L], y = unit(1.5 + 0.5,                  # to adapt position of x to its position with p-value
            "lines"))
        #grid.text(lab[1L], y = unit(1.5 + 0.5 * (lab[2L] != ""), 
        #    "lines"))
        grid.text(lab[2L], y = unit(1, "lines"))
        if (id) {
            nodeIDvp <- viewport(x = unit(0.5, "npc"), y = unit(1, 
                "npc"), width = max(unit(1, "lines"), unit(1.3, 
                "strwidth", nam[id_node(node)])), height = max(unit(1, 
                "lines"), unit(1.3, "strheight", nam[id_node(node)])))
            pushViewport(nodeIDvp)
            grid.rect(gp = gpar(fill = fill[2]))
            grid.text(nam[id_node(node)])
            popViewport()
        }
        upViewport()
    }
    return(rval)
}

class(node_inner_ext) <- "grapcon_generator"

@

\begin{frame}[fragile]
\frametitle{Distributional trees}
\vspace*{-0.12cm}
\begin{center} 
DGP:  $\; Y\ |\ X = x \; \sim  \; \mathcal{N}(\mu(x), \sigma^2(x))$

\vspace*{-0.21cm}
\setkeys{Gin}{width=0.58\linewidth}
<<plottree_xyplot, fig=TRUE, echo=FALSE,width=7>>=
par(mar=c(5.1,4.1,2.4,1.1))
plot(y=odata$y, x=odata$x, ylab = "y", xlab = "x", col = "gray")
lines(x = odata$x, y = odata$mu, col = pal["forest"], lwd = 2.5, main = "")
polygon(c(odata$x, rev(odata$x)), c(odata$mu + odata$sigma, rev(odata$mu - odata$sigma)),
  col = pallight["forest"], border = "transparent")
legend("topleft", expression(mu  %+-%  sigma), bty = "n")
@
\end{center}
\end{frame}

% \begin{frame}
% \frametitle{Distributional trees}
% \vspace*{-0.12cm}
% \begin{center} 
% DGP:  $\; Y\ |\ X = x \; \sim  \; \mathcal{N}(\mu(x), \sigma^2(x))$
% 
% \vspace*{-0.21cm}
% \setkeys{Gin}{width=0.5\linewidth}
% <<plottree_dgp, fig=TRUE, echo=FALSE>>=
% paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))
% plot(fig, inner_panel = node_inner_ext,
%      tp_args = list(FUN = identity, width = 18, fill = paltrees[c(1, 3)]), 
%      ip_args = list(fill = paltrees[c(2, 3)]),
%      drop_terminal = TRUE, tnex = 1.7)
% @
% \end{center}
% \end{frame}

<<tree, echo=FALSE, results=hide>>=
set.seed(7)
nobs <- 500
x <- runif(nobs, 0, 1)
mu <- sigma <- ytrue <- numeric(length = nobs)
for(i in 1:nobs) sigma[i] <- if(x[i]<=0.4) 1 else 3
for(i in 1:nobs) mu[i] <- if(x[i]<= 0.4|| x[i]>0.8) 4 else 12
y <- rnorm(nobs, mean = mu, sd = sigma)
#y <- rcnorm(nobs, mean = mu, sd = sigma, left = 0)
ytrue <- mu
data <- data.frame(cbind(y,x, ytrue))
tree <- disttree(y ~ x, data = data, family = NO(), type.tree = "mob")
#tree <- disttree(y ~ x, data = data, family = dist_list_cens_normal)
@

\begin{frame}[fragile]
\frametitle{Distributional trees}
\begin{center}
\vspace*{-0.12cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.5\linewidth}
<<plottree_estpar, fig=TRUE, echo=FALSE>>=
# function for output in terminal panels
FUN <- function (x) 
{
  cf <- x$coefficients
  cf <- matrix(cf, ncol = 1, dimnames = list(names(cf), ""))
  c(sprintf("n = %s", x$nobs), "Estimated parameters:", parse(text = paste0("mu == '", format(round(cf[1], 2), nsmall = 2), "'")), 
                                                        parse(text = paste0("sigma == '", format(round(cf[2], 2), nsmall = 2), "'")))
}

paltrees <- rgb(c(0.97, 0.64, 1), c(0.70, 0.83, 1), c(0.30, 0.99, 1))

## plot version using FUN and tree of class 'disttree'
plot(tree, drop = TRUE, tnex = 1.7, FUN = FUN,
     tp_args = list(fill = paltrees[c(1, 3)], width = 18), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Distributional trees}
\begin{center}
\vspace*{-0.12cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.5\linewidth}
<<plottree_box, fig=TRUE, echo=FALSE>>=
plot(as.constparty(tree), tnex = 1.7, drop = TRUE,
     tp_args = list(fill = paltrees[c(1, 3)], ylines = 1.5), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}

\begin{frame}[fragile]
\frametitle{Distributional trees}
\begin{center}
\vspace*{-0.12cm}
Model: \code{disttree(y ~ x)}\\
\vspace*{-0.2cm}
\setkeys{Gin}{width=0.5\linewidth}
<<plottree_dens, fig=TRUE, echo=FALSE>>=
node_density <- function (tree, xscale = NULL, yscale = NULL, horizontal = FALSE,
                          main = "", xlab = "", ylab = "Density", id = TRUE, rug = TRUE,
                          fill = paltrees[c(1, 3)], col = "black", lwd = 0.5, ...) 
{
  yobs <- tree$data[,as.character(tree$info$formula[[2]])]
  ylines <- 1.5
  if (is.null(xscale)) xscale <- c(-5.1,22.5)
  if (is.null(yscale)) yscale <- c(-0.05,0.45)
  xr <- xscale
  yr <- yscale

  if (horizontal) {
    yyy <- xscale
    xscale <- yscale
    yscale <- yyy
  }

  rval <- function(node) {
    yrange <- seq(from = -20, to = 90)/4
    ydens <- node$info$object$ddist(yrange)

    top_vp <- viewport(layout = grid.layout(nrow = 2, ncol = 3, 
                                            widths = unit(c(ylines, 1, 1), c("lines", "null", "lines")), 
                                            heights = unit(c(1, 1), c("lines", "null"))), 
                       width = unit(1, "npc"), 
                       height = unit(1, "npc") - unit(2, "lines"), 
                       name = paste("node_density",node$id, sep = ""))
    pushViewport(top_vp)
    grid.rect(gp = gpar(fill = "white", col = 0))
    top <- viewport(layout.pos.col = 2, layout.pos.row = 1)
    pushViewport(top)
    mainlab <- paste(ifelse(id, paste("Node", node$id, "(n = "), "n = "), node$info$nobs, ifelse(id, ")", ""), sep = "")

    grid.text(mainlab)
    popViewport()
    plot <- viewport(layout.pos.col = 2, layout.pos.row = 2, 
                     xscale = xscale, yscale = yscale, 
                     name = paste("node_density",  node$id, "plot", sep = ""))
    pushViewport(plot)
    yd <- ydens
    xd <- yrange
    if (horizontal) {
      yyy <- xd
      xd <- yd
      yd <- yyy
      yyy <- xr
      xr <- yr
      yr <- yyy
      rxd <- rep(0, length(xd))
      ryd <- rev(yd)
    } else {
      rxd <- rev(xd)
      ryd <- rep(0, length(yd))
    }

    if (rug) {
      nodeobs <- node$info$object$y
      if (horizontal) {
        grid.rect(x = xscale[1], y = nodeobs , height = 0, width = xscale[1], 
                  default.units = "native", just = c("right", "bottom"),
		  gp = gpar(lwd = 2, col = gray(0, alpha = 0.18)))
      } else {
        grid.rect(x = nodeobs, y = yscale[1], 
                  width = 0, height = abs(yscale[1]), default.units = "native", 
                  just = c("center", "bottom"),
		  gp = gpar(lwd = 2, col = gray(0, alpha = 0.18)))
        #grid.lines(x = xr, y = yr, gp = gpar(col = "lightgray"), 
        #           default.units = "native")
        #grid.lines(x = xr, y = yr, gp = gpar(col = "lightgray"), 
        #           default.units = "native")
      }
    }

    grid.polygon(x = c(xd, rxd), y = c(yd, ryd), default.units = "native",
              gp = gpar(col = "black", fill = fill, lwd = lwd))
    #grid.lines(x = xd, y = yd, default.units = "native", 
    #           gp = gpar(col = col, lwd = lwd))
    grid.xaxis()
    grid.yaxis()
    grid.rect(gp = gpar(fill = "transparent"))
    upViewport(2)
  }
  return(rval)
}

class(node_density) <- "grapcon_generator"

plot(tree, tnex = 1.7, drop = TRUE,
     terminal_panel = node_density,
     tp_args = list(fill = paltrees[c(1, 3)]), 
     ip_args = list(fill = paltrees[c(2, 3)]))
@
\end{center}
\end{frame}
%\begin{frame}
%\frametitle{Distributional trees}
%\center
%\vspace*{-1cm}
%\setkeys{Gin}{width=0.8\linewidth}
%\includegraphics{tree_box}
%\end{frame}

\begin{frame}
\frametitle{Learning distributional trees and forests}
\begin{minipage}{0.7\textwidth}
{\bf Tree:}
\begin{enumerate}
\item<3-> Fit global distributional model $\mathcal{D}(Y; \theta)$: \\ % to the whole data set:\\
Estimate $\hat{\theta}$ via maximum likelihood \\
$\hat{\theta} = \argmax_{\theta \in \Theta} \sum_{i=1}^n \ell(\theta; y_i)$
\item<8-> Test for associations/instabilities of the \\
scores $\frac{\partial \ell}{\partial \theta}(\hat{\theta};y_i)$ and each possible split variable $X_j$.
\end{enumerate}
\end{minipage}
\begin{minipage}{0.23\textwidth}
%\vspace{-0.1cm}
\begin{tikzpicture}
\visible<2-3>{
\node[ellipse, fill=HighlightBlue!70, align=center, scale = 0.7, minimum width=60pt, minimum height = 30pt] (n0) at (0.8, 1.7) {$Y$};
}
\visible<4>{
\node[ellipse, fill=HighlightBlue!70, align=center, scale = 0.7, minimum width=60pt, minimum height = 30pt] (n0) at (0.8, 1.7) {$\mathcal{D}(Y;\hat{\theta}$)};
}
\visible<5>{
\node[inner sep=0pt] (density_all) at (0.8, 1.7)
{\includegraphics[width=0.7\textwidth]{density_all}};
}
\visible<6->{
\node[inner sep=0pt] (density_all) at (0.8, 1.7)
{\includegraphics[width=0.7\textwidth]{density_all_hist}};
}
\visible<7-9>{
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n1) at (0, 0.2) {?};
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n2) at (1.6, 0.2) {?};
\draw[-, gray, line width=0.5pt] (0.7, 1.16) -- (n1);
\draw[-, gray, line width=0.5pt] (0.9, 1.16) -- (n2);
}
\visible<10-11>{
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n1) at (0, 0.2) {$Y_1$};
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n2) at (1.6, 0.2) {$Y_2$};
\draw[-, gray, line width=0.5pt] (0.7, 1.16) -- (n1) node [midway, left] {\scriptsize $X \leq p$};
\draw[-, gray, line width=0.5pt] (0.9, 1.16) -- (n2) node [midway, right] {\scriptsize $X > p$};
}
\visible<12>{
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n1) at (0, 0.2) {$\mathcal{D}(Y_1;\hat{\theta}_1$)};
\node[rectangle, fill=HighlightOrange!70, align=center, scale = 0.7, minimum width=50pt, minimum height = 20pt] (n2) at (1.6, 0.2) {$\mathcal{D}(Y_2;\hat{\theta}_2$)};
}
\visible<12->{
\draw[-, gray, line width=0.5pt] (0.7, 1.16) -- (n1) node [midway, left] {\scriptsize $X \leq p$};
\draw[-, gray, line width=0.5pt] (0.9, 1.16) -- (n2) node [midway, right] {\scriptsize $X > p$};
}
\visible<13->{
\node[inner sep=0pt] (density_1_hist) at (-0.2,-0.04)
{\includegraphics[width=0.6\textwidth]{density_2_hist}};
}
\visible<13->{
\node[inner sep=0pt] (density_2_hist) at (1.8,-0.04)
{\includegraphics[width=0.6\textwidth]{density_1_hist}};
}
\end{tikzpicture}
\end{minipage}
\vspace{0.1cm}
%\begin{adjustwidth}{-0.0em}{-1em}
\begin{enumerate}
\setcounter{enumi}{2}
\item<9-> Split along the covariate $X$ with strongest association or instability \\
and at breakpoint $p$ with highest improvement in log-likelihood.
\item<11-> Repeat steps 1--3 recursively until some stopping criterion \\
is met, yielding $B$ subgroups $\mathcal{B}_b$ with $b = 1, \dots, B$.
\end{enumerate}
%\end{adjustwidth}
\vspace{0.4cm}
\visible<14->{
{\bf Forest:} Ensemble of $T$ trees.
\begin{itemize}
\item Bootstrap or subsamples.
\item Random input variable sampling.
\end{itemize}
}
\end{frame}

\begin{frame}[fragile]
\frametitle{Weather forecasting}

\textbf{Goal:}

\medskip

\begin{center}
\begin{tikzpicture}
\draw[->] (0.4,0)--(0.9,0);
\draw[] (1,-0.4) rectangle (2.5,0.4);
\node[font=\scriptsize] at (1.75,0) {nature};
\draw[->] (2.6,0)--(3.1,0);
\visible<1>{%
\node[] at (0.0,0) {X};
\node[] at (3.5,0) {Y};}
\visible<2->{%
\node[inner sep=0pt] (today) at (-2.3,-0.4)
{\includegraphics[width=.365\textwidth]{airport_20200412.jpg}};
\node[font=\scriptsize] at (-2.3,-2.4) {2020-04-12};
\node[inner sep=0pt] (future) at (5.8,-0.4)
{\includegraphics[width=.365\textwidth]{airport_20200413.jpg}};
\node[font=\scriptsize] at (5.8,-2.4) {2020-04-13};}
\end{tikzpicture}
\end{center}

%% https://innsbruck-airport.panomax.com/

\only<1-2>{%
\textbf{Data:}
\begin{itemize}
\item X: State of the atmosphere now (temperature, precipitation, wind, \dots).
\item Y: State of the atmosphere in the future (hours, days, weeks, \dots).
\end{itemize}}%
\only<3->{%
\textbf{Two stages:}
\begin{itemize}
\item Physical model: Numerical weather prediction (NWP).
\item Statistical model: Model output statistics (MOS).
\end{itemize}}

\end{frame}

% \begin{frame}[fragile]
% \frametitle{Weather forecasting}
% 
% \textbf{Numerical weather prediction (NWP):}
% \begin{itemize}
%   \item Based on a physical model.
%   \item Massive numerical simulation of atmospheric processes.
%   \item Here: Global model on a $50\times 50 \text{ km}^2$ grid.
% \end{itemize}
% 
% \medskip
% 
% \textbf{Problem:} Uncertain initial conditions, unresolved processes.
% 
% \medskip
% 
% \textbf{Solution:} Ensemble of simulation runs under perturbed conditions.
% 
% \end{frame}
% 
% 
% 
% <<echo = FALSE, results=hide>>=
% library("zoo")
% 
% Sys.setenv("TZ"="UTC")
% 
% # -------------------------------------------------------------------
% # GFS ensemble forecast
% # -------------------------------------------------------------------
% 
% GFS   <- read.table("Data/GENS_00_innsbruck_20200412.dat", header = TRUE)
% init  <- min(as.POSIXct(GFS$timestamp-GFS$step*3600,origin="1970-01-01"))
% title <- sprintf("%s\n%s",
%                  "Global Forecast System (GFS) Ensemble Forecast for Innsbruck, Airport",
%                  sprintf("Forecast initialized %s", strftime(init, "%Y-%m-%d %H:%M UTC")))
% 
% # observations
% load("Data/STAGEobs_tawes_11121_defense.rda")
% #load("Data/STAGEobs_tawes_11121.rda")
% # "2018-03-14 00:00:00" in line 321 (first entry for temperature)
% # "2018-03-13 06:00:00" in line 317 (first entry for rain -> aggregated to 24h sums)
% # "2018-03-23 00:00:00" in line 357
% 
% getZoo <- function( x, variable ) {
%   x <- subset( x, varname == variable )
%   if ( nrow(x) == 0 ) stop("Ups, variable seems not to exist at all!")
%   # Else create zoo
%   x <- zoo( subset(x,select=-c(varname,timestamp,step)),
%             as.POSIXct(x$timestamp,origin="1970-01-01") )
%   x
% }
% GFS_t2m  <- getZoo( GFS, "tmp2m" ) - 273.15
% # temperature bias: model simulates temperature for height 1675 (~1070 m higher than true height)
% # => add 10.7 degrees to account for bias (approx. 1 degree / 100 m)
% # Custom bias
% GFS_t2m <- GFS_t2m + 6.5
% 
% # Subsetting obs
% obs <- subset(obs, index(obs) %in% index(GFS_t2m))
% 
% GFS_t2m <- merge(GFS_t2m, obs = obs[, grep("tl", names(obs))])
% ###GFS_t2m$obs <- obs[,"obs$tl"]
% # first value of rain is 18 hours later than temperature 
% #(both start at 6:00 but rain is cumulated over 24 hours)
% # -> drop first 3 values of temperature
% GFS_t2m <- GFS_t2m[-c(1:3),]
% 
% GFS_rain <- getZoo( GFS, "apcpsfc" )
% GFS_rain$obs <- obs[,"rr6"]
% # precipitaion sums over 24 hours (00:00 UTC - 00:00 UTC +1day)
% GFS_rain24 <- GFS_rain[c((1:10)*4),]
% for(i in 1:NROW(GFS_rain24)){
%   GFS_rain24[i,] <- colSums(GFS_rain[c((4*i-3):(4*i)),])
% }
% 
% 
% # Two POSIXct vectors for the axis
% main  <- seq(min(as.POSIXct(as.Date(index(GFS_t2m)))),max(index(GFS_t2m)),by=86400)
% minor <- seq(min(index(GFS_t2m)),max(index(GFS_t2m)),by=3*3600)
% 
% # GFS_t2m_mean <- rowMeans(GFS_t2m)
% GFS_t2m_mean <- GFS_t2m
% GFS_t2m_mean[,1] <- rowMeans(GFS_t2m[,-NCOL(GFS_t2m)])
% GFS_t2m_mean <- GFS_t2m_mean[,1]
% 
% # GFS_rain_mean <- rowMeans(GFS_rain)
% GFS_rain24_mean <- GFS_rain24
% GFS_rain24_mean[,1] <- rowMeans(GFS_rain24[,-NCOL(GFS_rain24)])
% GFS_rain24_mean <- GFS_rain24_mean[,1]
% @
% 
% <<t2m, echo=FALSE, eval=FALSE>>=
% par(mar=c(0.4,0,0.2,0), oma=c(6,3.2,5,4))
% layout( matrix(1:2,ncol=1) )
% plot(GFS_t2m[,1], screen=1, xaxs="i", xaxt="n", ylim = range(GFS_t2m, na.rm = TRUE)*1.05, col = "slategray" )
% abline(v = main, lty = 2 )
% mtext(side = 3, line = 1, cex = 1.2, font = 2, title )
% mtext(side = 2, line = 2.3, cex = 1, font = 1, text = "Temperature [°C]" )
% @
% 
% <<rain24, echo=FALSE, eval=FALSE>>=
% plot(GFS_rain24[,1], screen=1, xaxs="i", xaxt="n", yaxs="i", ylim=range(GFS_rain24, na.rm = TRUE)*c(0,1.05), col = "slategray" )
% abline(v = main, lty = 2 )
% mtext(side = 2, line = 2.3, cex = 1, font = 1, text = "Rain [mm/6h]" )
% axis(side = 1, at = main + 42300,  strftime(main,"%b %d"),
% line = 0, col.ticks=NA, col.axis="gray30", col = NA )
% @
% 
% \begin{frame}[fragile]
% \frametitle{Weather forecasting}
% 
% \vspace{-0.6cm}
% \setkeys{Gin}{width=0.95\linewidth}
% \begin{center}
% \only<1>{%
% <<gfs1, fig=TRUE, echo=FALSE, width=9>>=
% <<t2m>>
% <<rain24>>
% @
% }%
% \only<2>{%
% <<gfs2, fig=TRUE, echo=FALSE, width=9>>=
% <<t2m>>
% lines(GFS_t2m[,3], col = "slategray")
% <<rain24>>
% lines(GFS_rain24[,3], col = "slategray")
% @
% }%
% \only<3>{%
% <<gfs3, fig=TRUE, echo=FALSE, width=9>>=
% <<t2m>>
% lines(GFS_t2m[,2], col = "slategray")
% lines(GFS_t2m[,3], col = "slategray")
% <<rain24>>
% lines(GFS_rain24[,2], col = "slategray")
% lines(GFS_rain24[,3], col = "slategray")
% @
% }%
% \only<4>{%
% <<gfs4, fig=TRUE, echo=FALSE, width=9>>=
% <<t2m>>
% for(j in 2:(NCOL(GFS_t2m) - 1)) lines(GFS_t2m[,j], col = "slategray")
% <<rain24>>
% for(j in 2:(NCOL(GFS_rain24) - 1)) lines(GFS_rain24[,j], col = "slategray")
% @
% }%
% \only<5>{%
% <<gfs5, fig=TRUE, echo=FALSE, width=9>>=
% <<t2m>>
% for(j in 2:(NCOL(GFS_t2m) - 1)) lines(GFS_t2m[,j], col = "slategray")
% lines(GFS_t2m[,"obs"], col = 2, lwd = 2.3)
% <<rain24>>
% for(j in 2:(NCOL(GFS_rain24) - 1)) lines(GFS_rain24[,j], col = "slategray")
% lines(GFS_rain24[,"obs"], col = 2, lwd = 2.3)
% @
% }%
% \end{center}
% 
% \end{frame}

\begin{frame}[fragile]
\frametitle{Weather forecasting: precipitation}

\textbf{Goal:} Predict daily precipitation amount in Tyrol. %complex terrain.

\bigskip
\pause

\textbf{Observation data:} National Hydrographical Service.
\begin{itemize}
\item Daily 24h precipitation sums from July over 28 years (1985--2012).
\item 95 observation stations in Tyrol.
\end{itemize}

\bigskip
\pause

\textbf{Numerical weather prediction:} Global Ensemble Forecast System.
\begin{itemize}
\item Model outputs: Precipitation, temperature, air pressure, 
  convective available potential energy, downwards short wave radiation flux, \dots
\item 80 covariates based on ensemble min/max/mean/standard deviation.
\end{itemize}

\bigskip
\pause

\textbf{Distribution assumption:} Power-transformed Gaussian, censored at 0.
\[ (\text{precipitation})^\frac{1}{1.6} \sim \textit{c}\mathcal{N}(\mu,\sigma^2) \]

\end{frame}

%% \begin{frame}
%% \frametitle{Weather forecasting: precipitation}
%% 
%% \textbf{Base variables:}
%% \begin{itemize}
%% \item Total precipitation.
%% \item Convective available potential energy.
%% \item Downwards short wave radiation flux (``sunshine'').
%% \item Mean sea level pressure.
%% \item Preciptable water.
%% \item 2m maximum temperature.
%% \item Total column-integrated condensate.
%% \item Temperature.
%% \item Temperature differences in altitude.
%% \end{itemize}
%% 
%% \medskip
%% 
%% \textbf{Variations:} 80 covariates based on ensemble min/max/mean/standard deviation.
%% 
%% \end{frame}

<<pred_axams, eval=TRUE, echo=FALSE, results=hide>>=
if(file.exists("Data/Axams_pred.rda") & file.exists("Data/Axams_testdata.rda")){
  load("Data/Axams_pred.rda")
  load("Data/Axams_testdata.rda")
} else {
  #####
  # load observations and covariates 
  data("RainAxams")

  # tree and forest formula
  {
    dt.formula <- df.formula <- 
      robs ~ tppow_mean + tppow_sprd + tppow_min + tppow_max + 
      tppow_mean0612 + tppow_mean1218 + tppow_mean1824 + tppow_mean2430 + 
      tppow_sprd0612 + tppow_sprd1218 + tppow_sprd1824 + tppow_sprd2430 + 
      capepow_mean + capepow_sprd + capepow_min + capepow_max + 
      capepow_mean0612 + capepow_mean1218 + capepow_mean1224 + capepow_mean1230 +
      capepow_sprd0612 + capepow_sprd1218 + capepow_sprd1224 + capepow_sprd1230 +
      dswrf_mean_mean + dswrf_mean_max +  
      dswrf_sprd_mean + dswrf_sprd_max + 
      msl_mean_mean + msl_mean_min + msl_mean_max + 
      msl_sprd_mean + msl_sprd_min + msl_sprd_max +
      pwat_mean_mean + pwat_mean_min + pwat_mean_max + 
      pwat_sprd_mean + pwat_sprd_min + pwat_sprd_max +
      tmax_mean_mean + tmax_mean_min + tmax_mean_max +
      tmax_sprd_mean + tmax_sprd_min + tmax_sprd_max +
      tcolc_mean_mean + tcolc_mean_min + tcolc_mean_max +
      tcolc_sprd_mean + tcolc_sprd_min + tcolc_sprd_max +
      t500_mean_mean + t500_mean_min + t500_mean_max +
      t700_mean_mean + t700_mean_min + t700_mean_max +
      t850_mean_mean + t850_mean_min + t850_mean_max +
      t500_sprd_mean + t500_sprd_min + t500_sprd_max +
      t700_sprd_mean + t700_sprd_min + t700_sprd_max +
      t850_sprd_mean + t850_sprd_min + t850_sprd_max +
      tdiff500850_mean + tdiff500850_min + tdiff500850_max +
      tdiff700850_mean + tdiff700850_min + tdiff700850_max +
      tdiff500700_mean + tdiff500700_min + tdiff500700_max +
      msl_diff

  }

  # learning data: 24 years (1985 - 2008, both inlcuded)
  # testing data: 4 successive years (2009, 2010, 2011, 2012)
  learndata <- RainAxams[RainAxams$year < 2009,]
  testdata <- RainAxams[RainAxams$year %in% c(2009, 2010, 2011, 2012),]
  save(file = "Data/Axams_testdata.rda", testdata)

  ##############################################################
  # fitting the models
  set.seed(7)

  df <- distforest(df.formula, data = learndata, family = dist_list_cens_normal, 
                   ntree = 100, censtype = "left", censpoint = 0,
                   control = disttree_control(teststat = "quad", testtype = "Univ",
                                              type.tree = "ctree",
                                              intersplit = TRUE,
                                              mincriterion = 0, minsplit = 50,
                                              minbucket = 20), mtry = 27)

  #### prepare data for plot of estimated density functions
  # predictions for one day (in each of the four years) 
  # (19th of July 2011 is missing)
  pday <- 24  # 2 (hohe Beobachtung zu niedrig geschaetzt), 4, 15, evtl. auch 7, 8, 23 (eine 0-Beobachtung und 2 sehr aehnliche),

  pdays <- if(pday<19) c(pday, pday + 31, pday + 62, pday + 92) else c(pday, pday + 31, pday + 61, pday + 92)
  pdf <- predict(df, newdata = testdata[pdays,], type = "parameter")
  pdf$pdays <- pdays
  save(file = "Data/Axams_pred.rda", pdf)
}

pdays <- pdf$pdays
pday <- pdays[1]
df_mu <- pdf$mu
df_sigma <- pdf$sigma

# plot predicted distributions together with observations
set.seed(7)
x <- c(0.01, sort(runif(500,0.01,8)))
y1 <- crch::dcnorm(x, mean = df_mu[1], sd = df_sigma[1], left = 0)
y2 <- crch::dcnorm(x, mean = df_mu[2], sd = df_sigma[2], left = 0)
y3 <- crch::dcnorm(x, mean = df_mu[3], sd = df_sigma[3], left = 0)
y4 <- crch::dcnorm(x, mean = df_mu[4], sd = df_sigma[4], left = 0)

# switch x-axis back to untransformed scale
# x <- x^(1.6)

# point mass (slightly shifted)
pm1 <- c(0.05, crch::dcnorm(-1, mean = df_mu[1], sd = df_sigma[1], left = 0))     #  0.15
pm2 <- c(0.01, crch::dcnorm(-1, mean = df_mu[2], sd = df_sigma[2], left = 0))     #  0.05
pm3 <- c(-0.03, crch::dcnorm(-1, mean = df_mu[3], sd = df_sigma[3], left = 0))    # -0.05
pm4 <- c(-0.07, crch::dcnorm(-1, mean = df_mu[4], sd = df_sigma[4], left = 0))    # -0.15

# predictions
pred1 <- c(testdata[pdays,"robs"][1], # ^(1.6), 
           crch::dcnorm(testdata[pdays,"robs"][1], mean = df_mu[1], sd = df_sigma[1], left = 0))
pred2 <- c(testdata[pdays,"robs"][2], # ^(1.6), 
           crch::dcnorm(testdata[pdays,"robs"][2], mean = df_mu[2], sd = df_sigma[2], left = 0))
pred3 <- c(testdata[pdays,"robs"][3], # ^(1.6), 
           crch::dcnorm(testdata[pdays,"robs"][3], mean = df_mu[3], sd = df_sigma[3], left = 0))
pred4 <- c(testdata[pdays,"robs"][4], # ^(1.6), 
           crch::dcnorm(testdata[pdays,"robs"][4], mean = df_mu[4], sd = df_sigma[4], left = 0))

#legendheight
lh1 <- crch::dcnorm(0.01, mean = df_mu[1], sd = df_sigma[1], left = 0)
lh2 <- crch::dcnorm(0.01, mean = df_mu[2], sd = df_sigma[2], left = 0)
lh3 <- crch::dcnorm(0.01, mean = df_mu[3], sd = df_sigma[3], left = 0)
lh4 <- crch::dcnorm(0.01, mean = df_mu[4], sd = df_sigma[4], left = 0)
@

\begin{frame}[fragile]
\frametitle{Weather forecasting: precipitation}

\textbf{Application for one station:} Axams.
\begin{itemize}
%\vspace{-0.05cm}
\item Learn forest model on data from 24 years (1985--2008).
%\vspace{-0.05cm}
\item Evaluate on 4 years (2009--2012). Here: July \Sexpr{24}.
\end{itemize}

\vspace*{-0.4cm}
\begin{center}
\setkeys{Gin}{width=0.54\textwidth}
<<plot_pred_axams_24July, fig=TRUE, echo=FALSE, eval=TRUE, height=4.3, width=7>>=
par(mar = c(4, 4, 1, 1))
plot(x = x, y = y1, type = "l", col = 2, ylab = "Density", lwd = 2,
     # xlab = expression(Total~precipitation~"["~mm~"/"~"24h"~"]"),
     xlab = expression(Total~precipitation~"["~mm^(1/1.6)~"/"~"24h"~"]"),
     ylim = c(0,max(y1, y2, y3, y4, pm1, pm2, pm3, pm4) + 0.01),
     xlim = c(-1.5,8))     #  xlim = c(-2,28))

lines(x = x, y = y2, type = "l", lwd = 2, col = 4)
lines(x = x, y = y3, type = "l", lwd = 2, col = 3)
lines(x = x, y = y4, type = "l", lwd = 2, col = 6)
legend('topright', c("Predicted distribution", "Point mass at censoring point", "Observation"),
       bty = "n", col = "black", lty = c(1, NA, NA), pch = c(NA, 19, 4), cex = 1)

# plot point mass
lines(x = c(pm1[1], pm1[1]), y = c(pm1[2], 0), col = 2, type = "l", lwd = 2)
lines(x = c(pm2[1], pm2[1]), y = c(pm2[2], 0), col = 4, type = "l", lwd = 2)
lines(x = c(pm3[1], pm3[1]), y = c(pm3[2], 0), col = 3, type = "l", lwd = 2)
lines(x = c(pm4[1], pm4[1]), y = c(pm4[2], 0), col = 6, type = "l", lwd = 2)

points(x = pm1[1], y = pm1[2], col = 2, pch = 19, cex = 1.4)
points(x = pm2[1], y = pm2[2], col = 4, pch = 19, cex = 1.4)
points(x = pm3[1], y = pm3[2], col = 3, pch = 19, cex = 1.4)
points(x = pm4[1], y = pm4[2], col = 6, pch = 19, cex = 1.4)

# plot predictions
points(x = pred1[1], y = pred1[2], col = 2, pch = 4, cex = 1.4, lwd = 2)
points(x = pred2[1], y = pred2[2], col = 4, pch = 4, cex = 1.4, lwd = 2)
points(x = pred3[1], y = pred3[2], col = 3, pch = 4, cex = 1.4, lwd = 2)
points(x = pred4[1], y = pred4[2], col = 6, pch = 4, cex = 1.4, lwd = 2)

lines(x = c(pred1[1], pred1[1]), y = c(pred1[2], 0), col = "darkgray", type = "l", lty = 2, lwd = 2)
lines(x = c(pred2[1], pred2[1]), y = c(pred2[2], 0), col = "darkgray", type = "l", lty = 2, lwd = 2)
lines(x = c(pred3[1], pred3[1]), y = c(pred3[2], 0), col = "darkgray", type = "l", lty = 2, lwd = 2)
lines(x = c(pred4[1], pred4[1]), y = c(pred4[2], 0), col = "darkgray", type = "l", lty = 2, lwd = 2)

# add labels
text(x = -0.8, y = lh1, labels = "2009", col = 2, cex = 1)      # -1.7 
text(x = -0.8, y = lh2, labels = "2010", col = 4, cex = 1)      # -1.7
text(x = -0.8, y = lh3, labels = "2011", col = 3, cex = 1)      # -1.7
text(x = -0.8, y = lh4, labels = "2012", col = 6, cex = 1)      # -1.7
@
\end{center}

\only<2>{
\vspace{-0.5cm}
{\small Replication: \texttt{demo("RainAxams", package = "disttree")}}
}
\end{frame}

% \begin{frame}[fragile]
% \frametitle{Weather forecasting: precipitation}
% 
% \textbf{Application for one station:} Axams.
% \begin{itemize}
% %\vspace{-0.05cm}
% \item Learn forest model on data from 24 years.
% %\vspace{-0.05cm}
% \item Evaluate on 4 years.
% %\vspace{-0.05cm}
% \item 10 times 7-fold cross validation.
% \end{itemize}
% 
% \bigskip
% 
% \textbf{Benchmark:} Against other heteroscedastic censored Gaussian models.
% \begin{itemize}
% \item \emph{Ensemble MOS:} Linear predictors using only total precipitation.
% \item \emph{Prespecified GAMLSS:} Variable selection based on expert knowledge.
% \item \emph{Boosted GAMLSS:} Automatic variable selection.
% \end{itemize}
% 
% \bigskip
% 
% \textbf{Evaluation:} Continuous ranked probability skill score.
% 
% %% \vspace{0.1cm}
% %% \begin{table}[t!]
% %% \footnotesize
% %% 
% %% \hskip-0.4cm\begin{tabular}{ l l l }
% %% \hline
% %% Model & Type & Variable Selection \\
% %% \hline
% %% \vspace{-0.15cm}
% %% \textbf{Distributional forest} & recursive    & automatic \\ 
% %%                                & partitioning &           \\ 
% %% \vspace{-0.15cm}
% %% \textbf{Prespecified GAMLSS}   & spline  & based on expert \\
% %%                                & in each & knowledge      \\
% %% \vspace{-0.15cm}
% %% \textbf{Boosted GAMLSS}        & spline  & automatic  \\
% %%                                & in each &            \\
% %% \vspace{-0.15cm}
% %% \textbf{EMOS}                  & linear & total precipitation mean \\
% %%                                &        & and standard deviation   \\
% %% \hline
% %% \end{tabular}
% %% \end{table}
% 
% \end{frame}
% 
% 
% 
% 
% <<echo=FALSE, results=hide>>=
% #### cross validation rain
% if(file.exists("Data/crps_cross.rda")){
%   load("Data/crps_cross.rda")
% } else {
%   
%   nrep_cross <- 10
%   seed <- 7
%   
%   res_cross <- mclapply(1:nrep_cross,
%                         function(i){
%                           
%                           set.seed(seed*i)
%                           
%                           # randomly split data in 7 parts each including 4 years
%                           years <- 1985:2012
%                           testyears <- list()
%                           for(j in 1:7){
%                             testyears[[j]] <- sample(years, 4, replace = FALSE)
%                             years <- years[!(years %in% testyears[[j]])]
%                           }
%                           
%                           #crps <- matrix(nrow = 7, ncol = 7)
%                           reslist <- list()
%                           for(k in 1:7){
%                             test <- testyears[[k]]
%                             train <- c(1985:2012)[!c(1985:2012) %in% test]
%                             
%                             res <- evalmodels(station = "Axams",
%                                               train = train,
%                                               test = test,
%                                               gamboost_cvr = TRUE)
%                             
%                             #crps[k,] <- res$crps
%                             reslist[[k]] <- res
%                           }
%                           
%                           #colnames(crps) <- names(res$crps)
%                           return(reslist)
%                         },
%                         mc.cores = detectCores() - 1
%   )
%   
%   # extract CRPS
%   crps_cross <- matrix(nrow = nrep_cross, ncol = 7)
%   # loop over all repetitions
%   for(i in 1:length(res_cross)){
%     #loop over all 7 folds (for 7 methods)
%     crps_cross_int <- matrix(nrow = length(res_cross[[1]]), ncol = 7)
%     for(j in 1:length(res_cross[[1]])){
%       crps_cross_int[j,] <- res_cross[[i]][[j]]$crps
%     }
%     crps_cross[i,] <- colMeans(crps_cross_int, na.rm = TRUE)
%   }
%   colnames(crps_cross) <- names(res_cross[[1]][[1]]$crps) 
%   
%   save(crps_cross, file = "Data/crps_cross.rda")
% }
% @
% 
% 
% 
% 
% \begin{frame}[fragile]
% \frametitle{Weather forecasting: precipitation}
% \begin{center}
% \vspace*{-0.2cm}
% \setkeys{Gin}{width=0.65\textwidth}
% <<rain_cross_axams_crps_skills_score, fig=TRUE, echo=FALSE, height=5.6, width=7>>=
% #par(mar = c(2.5,2,1,2))
% boxplot(1 - crps_cross[,c(2,3,4)] / crps_cross[,6], ylim = c(-0.005, 0.065),
%         names = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS"),
%         main = "Cross validation (with reference model EMOS)",
%         cex.main=1.4,
%         cex.lab=1.2,
%         ylab = "CRPS skill score", col = "lightgray") 
% abline(h = 0, col = pal["EMOS"], lwd = 2)
% @
% \end{center}
% \end{frame}

% \begin{frame}[fragile]
% \frametitle{Weather forecasting: precipitation}
% %\vspace{0.2cm}
% \textbf{Application for all 95 stations:}
% \begin{itemize}
% \item Learn forest model on data from 24 years (1985--2008).
% \item Evaluate on 4 years (2009--2012).
% \item Benchmark against other heteroscedastic censored Gaussian models.
% \end{itemize}
% 
% \end{frame}
% 
% 
% <<echo = FALSE>>=
% #### prediction over all stations 24 - 4
% if(file.exists("Data/crps_24to4_all.rda")){
%   load("Data/crps_24to4_all.rda")
% } else {
%   
%   data("StationsTyrol")
%   stations <- StationsTyrol$name
%   test <- 2009:2012
%   train <- 1985:2008
%   
%   
%   res_24to4_all <- mclapply(1:length(stations),
%                             function(i){
%                               
%                               set.seed(7)
%                               
%                               res <- evalmodels(station = stations[i],
%                                                 train = train,
%                                                 test = test,
%                                                 gamboost_cvr = TRUE)
%                               
%                               return(res)
%                             },
%                             mc.cores = detectCores() - 1
%   )
%   
%   # extract crps
%   crps_24to4_all <- matrix(nrow = length(stations), ncol = 7)
%   # loop over all stations
%   for(i in 1:length(stations)){
%     crps_24to4_all[i,] <- res_24to4_all[[i]]$crps
%   }
%   
%   colnames(crps_24to4_all) <- names(res_24to4_all[[1]]$crps)
%   rownames(crps_24to4_all) <- stations
%   
%   save(crps_24to4_all, file = "Data/crps_24to4_all.rda")
% }
% 
% 
% # skill score
% s <- 1 - crps_24to4_all[, 2:4]/crps_24to4_all[,6] 
% colnames(s) <- c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS")
% 
% 
% ## prepare data for map which shows where distforest performed better than gamlss or gamboostLSS based on the crps
% 
% crps_map <- crps_24to4_all[,c("distforest", "gamlss", "gamboostLSS", "emos_log")]  
% 
% # best method
% bst <- apply(crps_map, 1, which.min)
% 
% # distance of forest to best other method
% dst <- crps_map[,1] - crps_map[cbind(1:nrow(crps_map), apply(crps_map[, -1], 1, which.min) + 1)]
% 
% # breaks/groups
% brk <- c(-0.1, -0.05, -0.005, 0.005, 0.05, 0.1)
% #brk <- c(-0.1, -0.05, -0.01, 0.01, 0.05, 0.1)
% grp <- cut(dst, breaks = brk)
% 
% # HCL colors (relatively flashy, essentially CARTO Tropic)
% clr <- colorspace::diverging_hcl(5, h = c(130, 320), c = 70, l = c(50, 90), power = 1.3)
% 
% 
% library("raster") # dem (digital elevation model)
% library("sp")     # gadm www.gadm.org/country
% 
% data("StationsTyrol", package = "RainTyrol")
% data("MapTyrol", package = "RainTyrol")
% # data(MapTyrol_border, package = "RainTyrol")
% # Create SpatialPointsDataFrame from station list
% sp <- SpatialPointsDataFrame(subset(StationsTyrol,
%                                     select=c(lon,lat)),
%                              data = subset(StationsTyrol,
%                                            select = -c(lon,lat)),
%                              proj4string = crs(MapTyrol$RasterLayer))
% @
% 
% 
% \begin{frame}[fragile]
% \frametitle{Weather forecasting: precipitation}
% \begin{center}
% \vspace*{-0.5cm}
% 
% \setkeys{Gin}{width=0.84\textwidth}
% <<map, fig=TRUE, echo=FALSE, width=10, height=6.5>>=
% 
%   ## plot map of Tyrol with all 95 observations
%   layout(cbind(1, 2), width = c(9, 1))
%   par(mar = c(5,4,4,0.1))
%   raster::image(MapTyrol$RasterLayer, col = rev(gray.colors(100)),
%                 main="Stations in Tyrol", ylab = "Latitude", xlab = "Longitude", 
%                 xlim = c(9.8,13.2), 
%                 ylim = c(46.6, 47.87))
%   plot(MapTyrol$SpatialPolygons, add = TRUE)
%   points(sp[70,], pch = 19, col = "black", cex = 1.85)
%   points(sp, pch = c(21, 24, 25, 22)[bst], bg = clr[grp], col = "black", las = 1, cex = 1.5)
%   legend(x = 9.8, y = 47.815, pch = c(21, 24, 25, 22), legend = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS", "EMOS"), cex = 1, bty = "n")
%   text(x = 10.3, y = 47.82, labels = "Models with lowest CRPS")
%   mtext("CRPS\ndifference", side=4, las = TRUE, at = c(x = 13.5, y = 47.76), line = 0.3)
%   par(mar = c(0.5,0.2,0.5,2.3))
%   ## legend
%   plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
%        xlim = c(0, 1), ylim = c(-0.2, 0.2), xaxs = "i", yaxs = "i")
%   rect(0, brk[-6], 0.5, brk[-1], col = rev(clr))
%   axis(4, at = brk, las = 1, mgp=c(0,-0.5,-1))
% 
% 
% @
% \end{center}
% \end{frame}
% 
% 
% 
% \begin{frame}[fragile]
% \frametitle{Weather forecasting: precipitation}
% \begin{center}
% \vspace*{-0.5cm}
% 
% \setkeys{Gin}{width=0.47\textwidth}
% <<map, fig=TRUE, echo=FALSE, width=10, height=6.5>>=
% 
%   ## plot map of Tyrol with all 95 observations
%   layout(cbind(1, 2), width = c(9, 1))
%   par(mar = c(5,4,4,0.1))
%   raster::image(MapTyrol$RasterLayer, col = rev(gray.colors(100)),
%                 main="Stations in Tyrol", ylab = "Latitude", xlab = "Longitude", 
%                 xlim = c(9.8,13.2), 
%                 ylim = c(46.6, 47.87))
%   plot(MapTyrol$SpatialPolygons, add = TRUE)
%   points(sp[70,], pch = 19, col = "black", cex = 1.85)
%   points(sp, pch = c(21, 24, 25, 22)[bst], bg = clr[grp], col = "black", las = 1, cex = 1.5)
%   legend(x = 9.8, y = 47.815, pch = c(21, 24, 25, 22), legend = c("Distributional forest", "Prespecified GAMLSS", "Boosted GAMLSS", "EMOS"), cex = 1, bty = "n")
%   text(x = 10.3, y = 47.82, labels = "Models with lowest CRPS")
%   mtext("CRPS\ndifference", side=4, las = TRUE, at = c(x = 13.5, y = 47.76), line = 0.3)
%   par(mar = c(0.5,0.2,0.5,2.3))
%   ## legend
%   plot(0, 0, type = "n", axes = FALSE, xlab = "", ylab = "",
%        xlim = c(0, 1), ylim = c(-0.2, 0.2), xaxs = "i", yaxs = "i")
%   rect(0, brk[-6], 0.5, brk[-1], col = rev(clr))
%   axis(4, at = brk, las = 1, mgp=c(0,-0.5,-1))
% 
% 
% @
% \end{center}
% \vspace{-0.2cm}
% \textbf{Conclusions:}
% \begin{itemize}
% \item Distributional forests provide an easy-to-apply alternative to already existing methods.
% \item Do not require expert knowledge for pre-specifications.
% \item Competitive performance: mainly en par or even slightly better.
% \end{itemize}
% 
% \end{frame}

\subsection{References}

\begin{frame}
\frametitle{References}

\vspace{-0.2cm}

\scriptsize

Breiman L (2001).
  \dquote{Random {Forests}.}
  \emph{Machine Learning}, 
  \textbf{45}(1), 5--32.
  \doi{10.1023/A:1010933404324}

\medskip

Breiman L, Cutler A (2004).
 \dquote{Random Forests.},
 \url{https://www.stat.berkeley.edu/~breiman/RandomForests}
 (Accessed: 2018-02-22)

\medskip

%Hothorn T, Zeileis A (2017).
%  \dquote{Transformation Forests.}
%  \emph{arXiv 1701.02110}, arXiv.org E-Print Archive.
%  \url{http://arxiv.org/abs/1701.02110}
%  
%\smallskip

Hothorn T, Hornik K, Zeileis A (2006).
 \dquote{Unbiased Recursive Partitioning: A Conditional Inference Framework.}
 \emph{Journal of Computational and Graphical Statistics},
 \textbf{15}(3), 651--674.
 \doi{10.1198/106186006X133933}

\medskip

Hothorn T, Zeileis A (2015).
 \dquote{{partykit}: A Modular Toolkit for Recursive Partytioning in \textsf{R}.}
 \emph{Journal of Machine Learning Research},
 \textbf{16}, 3905--3909.
 \url{http://www.jmlr.org/papers/v16/hothorn15a.html}

\medskip

Liaw A, Wiener M (2002).
 \dquote{Classification and Regression by randomForest.}
 \emph{R News},
 \textbf{2}(3), 18--22.
 \url{https://CRAN.R-project.org/doc/Rnews/}

\medskip

Wright M N, Ziegler A (2017).
\dquote{{ranger}: A Fast Implementation of Random Forests for High
Dimensional Data in {C++} and {R}.}
\emph{Journal of Statistical Software}, \textbf{77}{1}, 1--17.
\doi{10.18637/jss.v077.i01}

\medskip

% Zeileis A, Hothorn T, Hornik K (2008).
%  \dquote{Model-Based Recursive Partitioning.}
%   \emph{Journal of Computational and Graphical Statistics},
%   \textbf{17}(2), 492--514.
%   \doi{10.1198/106186008X319331}
% 
% \medskip

Athey S, Tibshirani J, Wager S (2019). 
 \dquote{Generalized Random Forests.} 
 \emph{Annals of Statistics}, \textbf{47}{2}, 1148--1178.
 \doi{10.1214/18-AOS1709"}

\medskip

Schlosser L, Hothorn T, Stauffer R, Zeileis A (2019).
\dquote{Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain.}
\emph{The Annals of Applied Statistics}, \textbf{13}(3), 1564--1589.
\doi{10.1214/19-AOAS1247}

\end{frame}

\end{document}